### 2015-12-19 Rmd/lit-review-analyses.Rmd
Commit of full datasets, some moving around of files, first package function added (kdensity.R).

---
title: "Supporting analysis of paper metadata"
author: "Lyndon Estes"
date: "February 24, 2015"
output: html_document
---

From 10.5194/bg-11-463-2014. Figuring out parameters from MODIS data used in study. 
```{r}
library(sp)
library(rgeos)
library(raster)
library(geosphere)
library(rworldmap)
data(countriesLow)

r <- raster(extent(-70, 30, -55, -25))
res(r) <- 0.00833
areas <- area(r)
plot(areas)
areas_m <- mask(areas, countriesLow, inverse = TRUE)
plot(areas_m)
plot(countriesLow, add = TRUE)
cellStats(areas_m, sum) * 100  # 23795817
cellStats(areas_m, mean)
1000^2 * 0.64005
freq(areas_m)
```

### 2015-12-19 Rmd/plots.Rmd
Commit of full datasets, some moving around of files, first package function added (kdensity.R).

---
title: "Exploratory lit review analysis"
author: "Lyndon Estes"
output: 
  html_document:
    toc: yes
---

## Initial analyis of metadata

### Data
```{r}
library(data.table)
library(raster)
library(sp)
require(splancs)
p_root <- lmisc::full_path(lmisc::proj_root("lit_review"), "lit_review")
p_dat <- lmisc::full_path(p_root, "data/")
dat <- read.csv("data/metadata_lde.csv", skip = 1, stringsAsFactors = FALSE)

```

### Test plots
```{r}
unique(dat$DOI.title)  # 20
unlist(strsplit(unique(dat$DOI_data_source), ";"))  # 12 more
with(dat, plot(log(plot_res), n_sites))
with(dat, plot(plot_res, n_sites, pch = 16))
with(dat, plot(plot_res, log(sampled_area), pch = 16))
with(dat, plot(log10(plot_res), log10(sampled_area), pch = 16))
with(dat, plot(log10(plot_res), log10(n_sites), pch = 16))
with(dat, plot(log10(plot_res), n_sites, pch = 16))
with(dat, plot(log(samp_duration), log(t_btwn_samp + 0.00001), pch = 16))
with(dat, plot(log(samp_duration), log(study_duration), pch = 16))
with(dat, plot(log10(t_btwn_samp + 0.00001), log10(study_duration), pch = 16))
with(dat, plot(log10(t_btwn_samp + 0.00001), log10(study_span), pch = 16))

hist(dat$t_btwn_samp + (1 / (60 * 60 * 24)))
hist(log(dat$study_duration))
hist(dat$study_span)
table(dat$study_type)
```

## Spatal and temporal data

### Scaling in log10 space
```{r}
i <- 0.000001
j <- rep(0, 15)
for(k in 1:length(j)) {
  i <- i * 10
  j[k] <- i
}
tdt <- data.table("scaleval" = j, "time" = j)
# tdt <- data.table("scale" = c("1 msec", "1 sec", "1 min", "1 hr", "1 dy", 
#                               "1 wk", "1 mth", "1 yr", "1 de", " 1 ce"),
#                   "time" = c(1 / (1000 * 60 * 60 * 24), 1 / (60 * 60 * 24), 
#                              1 / (60 * 24), 1 / 24, 1, 7, 30, 365, 365 * 10, 
#                              365 * 100))
tdt[, tlog := log10(time)]

# i <- 0.0001
# j <- rep(0, 10)
# for(k in 1:length(j)) {
#   i <- i * 10
#   j[k] <- i
# }
rdt <- data.table("scaleval" = j, "area" = j^2)
# rdt <- data.table("scale" = c("0.000001m", "0.00001m", "0.0001m", "0.001m", 
#                               "0.01 m", "0.1 m", "1 m", "10 m", "100 m", 
#                               "1000 m", "10000 m", "100000 m"),
#                   "scaleval" = j, "area" = j^2)
#c(1/10000^2, 1/1000^2, 1/100^2, 0.1^2, 1, 10^2, 
                             #100^2, 1000^2, 10000^2, 100000^2))
rdt[, slog := log10(scaleval)]
rdt[, rlog := log10(area)]

# i <- 0.000001
# j <- rep(0, 16)
# for(k in 1:length(j)) {
#   i <- i * 10
#   j[k] <- i
# }
adt <- data.table("area" = j)
# adt <- data.table("scale" = c("<0.1 ha", "0.1 ha", "1 ha", "10 ha", "100 ha",
#                               "1000 ha", "10000 ha", "100000 ha", "1000000 ha", 
#                               "10000000 ha", "100000000 ha", "1000000000 ha"),
#                   "area" = j)
#                     c(0.1, 1, 10, 100, 1000, 10000, 100000, 1000000, 
#                               10000000, 100000000, 1000000000))
adt[, alog := log10(area)]
```

### Rasterize and calculate kernel densities

1. For spatial versus temporal version
```{r,eval = FALSE}
# raster for resolution (x dimension) versus area (y dimension)
# arst <- raster(xmn = -6, xmx = 9, ymn = -5, ymx = 10)
arst <- raster(xmn = -5, xmx = 10, ymn = -5, ymx = 10)
res(arst) <- 0.25
arst@crs <- CRS("")
arst[] <- 1
as(extent(arst), "SpatialPolygons")
spatdat <- data.table("x" = log10(dat$plot_res), "y" = log10(dat$sampled_area))
spatdat[y < -5, y := -4.5]
spatdat$n <- rep(1, nrow(spatdat))
coordinates(spatdat) <- ~x + y
plot(spatdat$x, spatdat$y)

# trst <- raster(xmn = -6, xmx = 4, ymn = -5, ymx = 5)
trst <- raster(xmn = -5, xmx = 10, ymn = -5, ymx = 10)
res(trst) <- 0.25
trst@crs <- CRS("")
trst[] <- 1

# tempdat <- data.table("x" = log10(dat$t_btwn_samp + 0.001), 
#                       "y" = log10(dat$study_duration + 0.001)) 
tempdat <- data.table("x" = log10(dat$t_btwn_samp + 0.001), 
                      "y" = log10(dat$study_span + 0.001)) 
tempdat[x < -5, x := -4.5]
tempdat[y < -5, y := -4.5]
tempdat$n <- rep(1, nrow(tempdat))
coordinates(tempdat) <- ~x + y
plot(tempdat$x, tempdat$y)

# set-up grid topologies
# sgrd <- GridTopology(cellcentre.offset = c(-5.75, -4.75), 
#                      cellsize = c(0.5, 0.5), cells.dim = c(30, 30))
# tgrd <- GridTopology(cellcentre.offset = c(-6 + (0.333 / 2), -5 + (0.333 / 2)), 
#                      cellsize = c(0.333, 0.333), cells.dim = c(30, 30))
sgrd <- tgrd <- GridTopology(cellcentre.offset = c(-4.875, -4.875), 
                             cellsize = c(0.25, 0.25), cells.dim = c(60, 60))

# kernel densities
# spatial
ppoly <- as(extent(arst), "SpatialPolygons")
crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
kern1 <- spkernel2d(spatdat, crds, h0 = 1.25, sgrd)
kgrid1 <- as(SpatialGridDataFrame(sgrd, data = data.frame(kern1)),
             "SpatialPixelsDataFrame")

# temporal
ppoly <- as(extent(trst), "SpatialPolygons")
crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
# kern2 <- spkernel2d(tempdat, crds, h0 = 1.34, tgrd)
kern2 <- spkernel2d(tempdat, crds, h0 = 1.25, tgrd)
kgrid2 <- SpatialGridDataFrame(tgrd, data = data.frame(kern2))

# rasterize
# space
sr <- raster(kgrid1)
# sr <- shift(sr, 0.125, 0.125)  # shift to line up with point offset
sr[is.na(sr)] <- 0
# time
tr <- raster(kgrid2)
# tr <- shift(tr, 0.125, 0.125)  # shift to line up with point offset
# tr <- shift(tr, 0.333 / 2, 0.333 / 2)
tr[is.na(tr)] <- 0

# have a look
plot(sr)
points(spatdat$x, spatdat$y)
plot(tr)
points(tempdat$x, tempdat$y)

# shift rasters to common axis (0-15 in x, y)
# rr <- raster(nrow = 10, ncol = 10)
# srs <- raster(extent(c(0, 15, 0, 15)))
# res(srs) <- 0.5
# srs[] <- values(sr)
# trs <- raster(extent(c(0, 15, 0, 15)))
# res(trs) <- 0.5
# trs[] <- values(tr)
# trs <- resample(trs, srs)
srs <- sr
trs <- tr
```

#### Convert to rgb space and plot
```{r, eval = FALSE}
plot(srs, axes = FALSE, box = FALSE)
plot(trs, axes = FALSE, box = FALSE)

hstretch <- function(x, brange = 0.15) return(brange + (x * (1 - brange) / 1))
srsv <- values(srs)
srsvn <- (srsv - min(srsv)) / diff(range(srsv))
srsvn[srsvn > 0] <- hstretch(srsvn[srsvn > 0])  # compress range in non-0
trsv <- values(trs)
trsvn <- (trsv - min(trsv)) / (diff(range(trsv)))
trsvn[trsvn > 0] <- hstretch(trsvn[trsvn > 0])  # compress range in non-0
hist(trsvn)
grsv <- rep(0, length(srsv))
# grsv[which(srsv 0 & trsv == 0)] <- 1
# trsvn[which(srsv == 0 & trsv == 0)] <- 1
# srsvn[which(srsv == 0 & trsv == 0)] <- 1
srtrvals <- rgb(red = srsvn, blue = trsvn, green = grsv)
z <- matrix(srtrvals, nrow=nrow(srs), ncol=ncol(srs), byrow = TRUE)
bb <- as.vector(t(bbox(srs)))

cx <- 1.2
png("paper/figures/spacevtime.png", width = 700, height = 700)
par(mar = c(3, 3, 3, 3))
plot(-5:10, -5:10, axes = F, pch = "", xlab = "", ylab = "")
rasterImage(z, bb[1], bb[3], bb[2], bb[4])
#axis(1, labels = FALSE, line = -0.5)
axis(1, at = -5:10, labels = c(rdt[, slog], 10), line = -1.6, 
     col.axis = "red", col = "red", col.ticks = "red")
axis(2, at = -5:10, labels = c(adt[, alog], 10), line = -1.6, 
     col.axis = "red", col = "red", col.ticks = "red", las = 2)
axis(3, at = -5:10, labels = c(tdt[, tlog], 10), line = -1.6, 
     col.axis = "blue", col = "blue", col.ticks = "blue")
axis(4, at = -5:10, labels = c(tdt[, tlog], 10), line = -1.6, 
     col.axis = "blue", col = "blue", col.ticks = "blue", las = 2)
mtext("Plot resolution log10(m)", side = 1, line = 1, col = "red", cex = cx)
mtext("Sampled area log10(ha)", side = 2, line = 1, col = "red", cex = cx)
mtext("Sampling interval log10(days)", side = 3, line = 1, col = "blue", 
      cex = cx)
text(16, y = 7.5, "Sampling duration log10(days)", col = "blue", srt = 270, 
     xpd = TRUE, cex = cx)
dev.off()

```

2. For resolution versus extent version
```{r,eval = FALSE}
# raster for resolution (x dimension) versus area (y dimension)
# arst <- raster(xmn = -6, xmx = 9, ymn = -5, ymx = 10)
rst <- raster(xmn = -5, xmx = 10, ymn = -5, ymx = 10)
res(rst) <- 0.25
# rst@crs <- CRS("")
rst[] <- 1
rst2 <- raster(xmn = 0, xmx = 10, ymn = 0, ymx = 10)
res(rst2) <- c(0.25, 0.25)
# rst@crs <- CRS("")
rst2[] <- 1
plot(rst2)

resdat <- data.table("x" = log10(dat$t_btwn_samp + 0.00001), 
                     y = log10(dat$plot_res))
resdat[x < -5, x := -4.5]
resdat[y < -5, y := -4.5]
resdat$n <- rep(1, nrow(resdat))
coordinates(resdat) <- ~x + y
plot(resdat$x, resdat$y)

# extdat <- data.table("x" = log10(dat$study_duration + 0.00001), 
#                       "y" = log10(dat$sampled_area)) 
extdat <- data.table("x" = log10(dat$study_span), 
                     "y" = log10(dat$sampled_area)) 
extdat[x < 0, x := 0]
extdat[y < 0, y := 0]  # setting minima to 1 day and 1 ha
extdat$n <- rep(1, nrow(extdat))
coordinates(extdat) <- ~x + y
plot(extdat$x, extdat$y)

# set-up grid topologies
sgrd <- GridTopology(cellcentre.offset = c(-4.875, -4.875), 
                     cellsize = c(0.25, 0.25), cells.dim = c(60, 60))
tgrd <- GridTopology(cellcentre.offset = c(0.125, 0.125), 
                     cellsize = c(0.25, 0.25), cells.dim = c(40, 40))
# seq(0, 10, 10 / 59)

# kernel densities
# resolution
ppoly <- as(extent(rst), "SpatialPolygons")
crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
kern1 <- spkernel2d(resdat, crds, h0 = 1, sgrd)
kgrid1 <- as(SpatialGridDataFrame(sgrd, data = data.frame(kern1)),
             "SpatialPixelsDataFrame")

# extent
ppoly <- as(extent(rst2), "SpatialPolygons")
crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
kern2 <- spkernel2d(extdat, crds, h0 = 1, tgrd)
kgrid2 <- SpatialGridDataFrame(tgrd, data = data.frame(kern2))

# rasterize
# space
rr <- raster(kgrid1)
rr[is.na(rr)] <- 0
# time
er <- raster(kgrid2)
er[is.na(er)] <- 0

# have a look
par(mfrow = c(2, 1))
plot(er)
points(extdat$x, extdat$y)
plot(rr)
points(resdat$x, resdat$y)

ers <- er
rrs <- rr
```

### Convert to rgb space and plot
```{r, eval = FALSE}
# As separate plots
png("paper/figures/res&extent.png", width = 900, height = 450)
par(mfrow = c(1, 2), oma = c(0, 0, 0, 0), mar = c(4, 4, 4, 4))
plot(rrs, xlim = c(-5, 15), ylim = c(-5, 15), 
     xlab = "Sample interval - log10(days)", 
     ylab = "Plot resolution - log10(m^2)")
plot(ers, box = FALSE, xlim = c(-5, 15), ylim = c(-5, 15), 
     xlab = "Study duration - log10(days)", ylab = "Sampled area - log10(ha)")
dev.off()

rrs2 <- rrs
# rrs2[rrs2 == 0] <- NA
ers2 <- ers
# ers2[ers2 == 0] <- NA
tcols <- rev(terrain.colors(100))
# tcols <- colorRampPalette(c("grey95", "orange", "orange", "yellow", "green"))(100)
# tcols[1] <- "grey"
png("paper/figures/res&extent_pres.png", width = 900, height = 450)
par(mfrow = c(1, 2), oma = c(0, 0, 0, 0), mar = c(5, 5, 4, 4), bg = "transparent")
plot(rrs2, xlim = c(-5, 15), ylim = c(-5, 15), 
     cex.lab = 2, col = tcols, col.lab = "white", 
     col.ticks = "white", col.axis = "white", cex.axis = 2,
     xlab = "Sample interval - log10(days)", 
     ylab = "Plot resolution - log10(m^2)")
plot(ers2, xlim = c(-5, 15), ylim = c(-5, 15), col.axis = "white", 
     cex.axis = 2, cex.lab = 2, col = tcols, col.lab = "white",
     xlab = "Study period - log10(days)", box = FALSE, col.ticks = "white", 
     ylab = "Study extent - log10(ha)")
dev.off()

# Switched off because need to reconcile scales first

# hstretch <- function(x, brange = 0.25) return(brange + (x * (1 - brange) / 1))
# ersv <- values(ers)
# ersvn <- (ersv - min(ersv)) / diff(range(ersv))
# ersvn[ersvn > 0] <- hstretch(ersvn[ersvn > 0])  # compress range in non-0
# rrsv <- values(rrs)
# rrsvn <- (rrsv - min(rrsv)) / (diff(range(rrsv)))
# rrsvn[rrsvn > 0] <- hstretch(rrsvn[rrsvn > 0])  # compress range in non-0
# hist(rrsvn)
# grsv <- rep(0, length(rrsv))
# # grsv[which(srsv 0 & trsv == 0)] <- 1
# # trsvn[which(srsv == 0 & trsv == 0)] <- 1
# # srsvn[which(srsv == 0 & trsv == 0)] <- 1
# ertrvals <- rgb(red = ersvn, blue = rrsvn, green = grsv)
# z <- matrix(ertrvals, nrow=nrow(rrs), ncol=ncol(rrs), byrow = TRUE)
# bb <- as.vector(t(bbox(rrs)))

cx <- 1.2
png("paper/figures/resvextent.png", width = 700, height = 700)
par(mar = c(3, 3, 3, 3))
plot(-5:10, -5:10, axes = F, pch = "", xlab = "", ylab = "")
rasterImage(z, bb[1], bb[3], bb[2], bb[4])
#axis(1, labels = FALSE, line = -0.5)
axis(1, at = -5:10, labels = c(rdt[, slog], 10), line = -1.6)
axis(2, at = -5:10, labels = c(adt[, alog], 10), line = -1.6)
mtext("Sample interval - log10(days)", side = 1, line = 0.8, col = "blue", 
      cex = cx)
mtext("Study duration - log10(days)", side = 1, line = 2, col = "red", cex = cx)
mtext("Plot resolution - log10(m^2)", side = 2, line = 0.8, col = "blue", 
      cex = cx)
mtext("Sampled area - log10(ha)", side = 2, line = 2, col = "red", cex = cx)
dev.off()

```

### 2015-12-19 Rmd/select-titles.Rmd
Commit of full datasets, some moving around of files, first package function added (kdensity.R).

---
title: "Select titles"
author: "Lyndon Estes"
date: "December 19, 2015"
output: 
  html_document:
    toc: yes
---

# Select papers
Organization and random selection of titles from literature review.

```{r, eval=FALSE}
library(lmisc)
p_root <- full_path(proj_root(proj.dir = "lit_review"), "lit_review")
p_data <- full_path(p_root, "data")

# Read in downloaded citations and merge
batch1 <- read.csv(full_path(p_data, "top30_from1982_set1.csv"))  
batch2 <- read.csv(full_path(p_data, "top30_from1982_set2.csv"))
articles <- rbind(batch1, batch2)

# Shuffle full list 
set.seed(234)
draw <- sample(1:nrow(articles), size = nrow(articles), replace = FALSE)
article.shuffle <- articles[draw, ]

# 10-year list\set.seed(234)
articles.2004 <- articles[articles$Year >=2004, ]

set.seed(234)
draw <- sample(1:nrow(articles.2004), size = nrow(articles.2004), 
               replace = FALSE)
article.2004.shuffle <- articles.2004[draw, ]

# draw subsets for each reviewer - note: as long as drawsize and random seed are the same, new reviewers can 
# be appended to the group without affecting the draw for existing reviewers
reviewers <- c("Chang", "Treuer", "Socolar", "Daskin", "Elsen", "Estes", 
               "Choi", "Margulies", "JChang", "Ahmed")
nauth <- length(reviewers)
#drawsize <- (round(nrow(articles.2004) * 0.025 / nauth) * nauth) / nauth  # number divisible by 
drawsize <- 400
set.seed(432)
sset <- sample(1:nrow(article.2004.shuffle), size = drawsize * nauth, 
               replace = FALSE)
inds <- seq(1, length(sset) + 1, by = length(sset) / nauth)
ind.mat <- cbind(inds[-length(inds)], 
                 c(inds[-c(1, length(inds))] - 1, length(sset)))

# New calibration set randomly selected (note, run on earlier version of random draw taken before Elsen, 
# so to replicate use commented out drawsize and remove Elsen from reviewers before re-running)
# article.2004.shuffle.red <- article.2004.shuffle[-sset, ]
# set.seed(124)
# cal.set <- article.2004.shuffle.red[sample(1:nrow(article.2004.shuffle.red), size = 20, replace = FALSE), ]
# # which(rownames(cal.set) %in% rownames(article.2004.shuffle[sset, ]))  # 0
# write.csv(cal.set, file = "new_calibration_set.csv")

# Then draw sets for each reviewer
fnames <- full_path(p_data, paste(reviewers, "2004_draw3.csv", sep = "_"))
sel.set <- vector("list", length = length(fnames))
names(sel.set) <- reviewers
for(i in 1:nrow(ind.mat)) {
  sel.set[[i]] <- article.2004.shuffle[sset[ind.mat[i, 1]:ind.mat[i, 2]], ]
  write.csv(sel.set[[i]], file = fnames[i])
}

# check that rownames aren't duplicated
sel.set[[1]][1:10, ]
for(i in 1:(length(sel.set) - 1)) {
   print(any(sel.set[[i]]$Title == sel.set[[i + 1]]$Title))
}

fn_old <- dir(p_data, 
              pattern = paste0("^", 
                               paste(reviewers[1:6], "draw2.csv", 
                                     sep = ".*.", collapse = "|")))
fn_old <- fn_old[sapply(reviewers[1:6], function(x) grep(x, fn_old))]
fn_new <- dir(p_data, 
              pattern = paste0("^", 
                               paste(reviewers[1:6], "draw3.csv", 
                                     sep = ".*.", collapse = "|")))
fn_new <- fn_new[sapply(reviewers[1:6], function(x) grep(x, fn_new))]

# comparing draw2 and draw3
for(i in 1:length(fn_old)) {
  old <- read.csv(full_path(p_data, fn_old[i]))
  neww <- read.csv(full_path(p_data, fn_new[i]))
  print(which(!old$Title %in% neww$Title))
}  # only last record of old set is not in new set (b/c bad indexing)

# comparing draw2 titles in sequence
for(i in 1:(length(fn_old) - 1)) {
  old <- read.csv(full_path(p_data, fn_old[i]))
  oldp1 <- read.csv(full_path(p_data, fn_old[i + 1]))
  print(paste("common title", which(old$Title %in% oldp1$Title)))
  print(which(old$X %in% oldp1$X))
}  # overlap between last of preceding and 1st of next (and if no title)

fnames_newall <- dir(p_data, pattern = "draw3.csv")
for(i in 1:(length(fn_new) - 1)) {
  neww <- read.csv(full_path(p_data, fn_new[i]))
  newp1 <- read.csv(full_path(p_data, fn_new[i + 1]))
  print(which(neww$X %in% newp1$X))
  print(which(neww$Title %in% newp1$Title))
}  # no duplicate names in new file sets (unless title is untitled)

```

### 2015-12-19 Rmd/space-time.Rmd
Commit of full datasets, some moving around of files, first package function added (kdensity.R).

---
title: "Space-time plots"
author: "Lyndon Estes"
date: "December 3, 2015"
output: 
  html_document:
    toc: yes
---

# Space versus time analysis
```{r, eval = FALSE}
library(data.table)
library(raster)
library(sp)
require(splancs)
library(lmisc)
p_root <- set_base_path("ecoscales")
p_dat <- full_path(p_root, "external/data")
dat <- fread(full_path(p_dat, "merged-data.csv"), stringsAsFactors = FALSE)
setnames(dat, "DOI/title", "DOI")
dat <- dat[, names(dat)[1:20], with = FALSE]
unique(dat$DOI)  # 139
```

## Spatal and temporal data

### Scaling in log10 space
```{r}
i <- 0.000001
j <- rep(0, 15)
for(k in 1:length(j)) {
  i <- i * 10
  j[k] <- i
}
tdt <- data.table("scaleval" = j, "time" = j)
tdt[, tlog := log10(time)]

# label axis for time
taxis <- cbind.data.frame(
  c("Sec", "Min", "Hour", "Day", "Month", "Year", "Decade", "Century", 
    "Millenium", "10000"),  
  c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 30, 365, 365 * 10, 
    365 * 100, 365 * 1000, 365 * 10000))
taxis <- data.table(taxis)
setnames(taxis, names(taxis), c("label", "days"))
taxis[, logdays := log10(days)]

rdt <- data.table("scaleval" = j, "area" = j^2)
rdt[, slog := log10(scaleval)]
rdt[, rlog := log10(area)]

adt <- data.table("area" = j)
adt[, alog := log10(area)]

```

### Resolution versus extent, gridded
```{r,eval = FALSE}
rres <- 0.25
# # raster for resolution (x dimension) versus area (y dimension)
# # arst <- raster(xmn = -6, xmx = 9, ymn = -5, ymx = 10)
# rst <- raster(xmn = -5, xmx = 10, ymn = -5, ymx = 10)
# res(rst) <- rres
# # rst@crs <- CRS("")
# rst[] <- 1
# rst2 <- raster(xmn = 0, xmx = 10, ymn = 0, ymx = 10)
# # res(rst2) <- c(0.25, 0.25)
# res(rst2) <- rres
# # rst@crs <- CRS("")
# rst2[] <- 1
# plot(rst2)

resdat <- data.table(x = log10(dat$t_btwn_samp + 0.00001), 
                     y = log10(dat$plot_res))
# resdat[x < -5, x := -4.5]
# resdat[y < -5, y := -4.5]
resdat$n <- rep(1, nrow(resdat))
coordinates(resdat) <- ~x + y
# plot(resdat$x, resdat$y)
rx <- range(resdat$x)
# rx <- c(floor(rx[1]), c(ceiling(rx[2])))
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rxd <- length(seq(rx[1], rx[2], rres))  # 0.25))
ry <- range(resdat$y)
# ry <- c(floor(ry[1]), c(ceiling(ry[2])))
ry <- round(ry / 0.5) * 0.5  # round to nearest 0.5
ryd <- length(seq(ry[1], ry[2], rres))#0.25))

# length(seq(-5, 4, 0.25))

extdat <- data.table("x" = log10(dat$study_duration + 0.00001), 
                     "y" = log10(dat$sampled_area + 0.00001)) 
# extdat <- data.table("x" = log10(dat$study_span), 
#                      "y" = log10(dat$sampled_area)) 
# extdat[x < 0, x := 0]
# extdat[y < 0, y := 0]  # setting minima to 1 day and 1 ha
extdat$n <- rep(1, nrow(extdat))
coordinates(extdat) <- ~x + y
# plot(extdat$x, extdat$y)
ex <- range(extdat$x)
# ex <- c(floor(ex[1]), c(ceiling(ex[2])))
ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
ex[1] <- -5
exd <- length(seq(ex[1], ex[2], rres))#0.25))
ey <- range(extdat$y)
# ey <- c(floor(ey[1]), c(ceiling(ey[2])))
ey <- round(ey / 0.5) * 0.5  # round to nearest 0.5
eyd <- length(seq(ey[1], ey[2], rres))#0.25))

# set-up grid topologies
rgrd <- GridTopology(cellcentre.offset = c(rx[1], ry[1]) + 0.125, 
                     #cellsize = c(0.25, 0.25), cells.dim = c(rxd, ryd))
                     cellsize = c(rres, rres), cells.dim = c(rxd, ryd))
egrd <- GridTopology(cellcentre.offset = c(ex[1], ey[1]) + 0.125, 
                     #cellsize = c(0.25, 0.25), cells.dim = c(exd, eyd))
                     cellsize = c(rres, rres), cells.dim = c(exd, eyd))
# seq(0, 10, 10 / 59)

rst <- raster(xmn = rx[1], xmx = rx[2], ymn = ry[1], ymx = ry[2])
# rst <- raster(xmn = min(resdat$x), xmx = ceiling(max(resdat$x)), 
#               ymn = min(resdat$y), ymx = ceiling(max(resdat$y)))
res(rst) <- rres
# rst@crs <- CRS("")
rst[] <- 1
# plot(rst)
rst2 <- raster(xmn = ex[1], xmx = ex[2], ymn = ey[1], ymx = ey[2])
# rst2 <- raster(xmn = min(extdat$x), xmx = ceiling(max(extdat$x)), 
#               ymn = min(extdat$y), ymx = ceiling(max(extdat$y)))
res(rst2) <- rres
# rst@crs <- CRS("")
rst2[] <- 1
# plot(rst2)

# kernel densities
# resolution
ppoly <- as(extent(rst), "SpatialPolygons")
crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
kern1 <- spkernel2d(resdat, crds, h0 = 1, rgrd)
kgrid1 <- as(SpatialGridDataFrame(rgrd, data = data.frame(kern1)),
             "SpatialPixelsDataFrame")

# extent
ppoly <- as(extent(rst2), "SpatialPolygons")
crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
kern2 <- spkernel2d(extdat, crds, h0 = 1, egrd)
kgrid2 <- SpatialGridDataFrame(egrd, data = data.frame(kern2))

# rasterize
# space
rr <- raster(kgrid1)
rr[is.na(rr)] <- 0
# time
er <- raster(kgrid2)
er[is.na(er)] <- 0

# have a look
par(mfrow = c(2, 1))
plot(er, axes = FALSE, add = FALSE)
arg <- list(at = taxis$logdays, labels = taxis$label)
image(er, col = rev(terrain.colors(100)), axes = FALSE, xlab = "", ylab = "", 
      xlim = c(-5, 6))
axis(1, at = taxis$logdays, labels = taxis$label, las = 2)
axis(2, at = -5:10, las = 2)

# points(extdat$x, extdat$y)
plot(rr)
# points(resdat$x, resdat$y)

ers <- er
rrs <- rr
```

### Convert to rgb space and plot
```{r, eval = FALSE}
# As separate plots
png("paper/figures/res&extent2.png", width = 900, height = 450)
par(mfrow = c(1, 2), oma = c(0, 0, 0, 0), mar = c(4, 4, 4, 4))
plot(rrs, xlim = c(-5, 15), ylim = c(-5, 15), 
     xlab = "Sample interval - log10(days)", 
     ylab = "Plot resolution - log10(m^2)")
plot(ers, box = FALSE, xlim = c(-5, 15), ylim = c(-5, 15), 
     xlab = "Study duration - log10(days)", ylab = "Sampled area - log10(ha)")
dev.off()
```

### 2015-12-19 R/kdensity.R

@@ -1,0 +1,40 @@
+#' Kernel density raster
+#' @description Creates a raster from kernel density 
+#' @param xmn Minimum x coordinate for output grid
+#' @param xmx Maximum x coordinate for output grid
+#' @param ymn Minimum y coordinate for output grid
+#' @param ymx Maximum y coordinate for output grid
+#' @param dxdy Resolution for output grid
+#' @param pts SpatialPointsDataFrame showing distribution of observations 
+#' @return Raster of density values
+#' @export
+kdensity <- function(xmn, xmx, ymn, ymx, dxdy, pts) {
+  
+  if(length(dxdy) == 1) dxdy <- rep(dxdy, 2)
+  adj <- dxdy / 2
+  
+  d1 <- length(seq(xmn[1], xmx[2], dxdy[1]))
+  d2 <- length(seq(ymn[1], ymx[2], dxdy[2]))
+  
+  # grid topology
+  grd <- GridTopology(cellcentre.offset = c(xmn + adj[1], ymn + adj[2]), 
+                      cellsize = c(dxdy[1], dxdy[2]), 
+                      cells.dim = c(d1, d2))
+  
+  # container raster
+  rst <- raster(xmn = xmn, xmx = xmx, ymn = ymn, ymx = ymx)
+  res(rst) <- dxdy
+  
+  # kernel densities
+  ppoly <- as(extent(rst), "SpatialPolygons")
+  crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
+  kern <- spkernel2d(resdat, crds, h0 = 1, grd)
+  kgrid <- as(SpatialGridDataFrame(grd, data = data.frame(kern)),
+              "SpatialPixelsDataFrame")
+
+  krast <- raster(kgrid)
+  return(krast)
+}
+
+
+

### 2015-12-19 R/selectTitles.R

@@ -1,69 +1,0 @@
-# selectTitles.R
-# Organization and random selection of titles from literature review
-
-library(lmisc)
-p_root <- full_path(proj_root(proj.dir = "lit_review"), "lit_review")
-p_data <- full_path(p_root, "data")
-
-setwd(p_data)
-
-# Read in downloaded citations and merge
-batch1 <- read.csv("top30_from1982_set1.csv")  
-batch2 <- read.csv("top30_from1982_set2.csv")
-articles <- rbind(batch1, batch2) 
-
-# Full list
-set.seed(234)
-draw <- sample(1:nrow(articles), size = nrow(articles), replace = FALSE)
-article.shuffle <- articles[draw, ]
-
-# 1982 set (superseded now)
-# # draw subsets for each author
-# drawsize <- (round(nrow(articles) * 0.015 / 5) * 5) / 5  # number divisible by 5
-# set.seed(432)
-# sset <- sample(1:nrow(article.shuffle), size = drawsize * 5, replace = FALSE)
-# inds <- seq(1, length(sset) + 1, by = length(sset) / 5)
-# ind.mat <- cbind(inds[-length(inds)], c(inds[-c(1, length(inds))], length(sset)))
-# 
-# fnames <- paste(c("Chang", "Treuer", "Socolar", "Daskin", "Estes"), "draw.csv", sep = "_")
-# for(i in 1:nrow(ind.mat)) write.csv(article.shuffle[sset[ind.mat[i, 1]:ind.mat[i, 2]], ], file = fnames[i])
-
-# 10-year list\set.seed(234)
-articles.2004 <- articles[articles$Year >=2004, ]
-
-set.seed(234)
-draw <- sample(1:nrow(articles.2004), size = nrow(articles.2004), replace = FALSE)
-article.2004.shuffle <- articles.2004[draw, ]
-
-# draw subsets for each reviewer - note: as long as drawsize and random seed are the same, new reviewers can 
-# be appended to the group without affecting the draw for existing reviewers
-reviewers <- c("Chang", "Treuer", "Socolar", "Daskin", "Elsen", "Estes")
-nauth <- length(reviewers)
-#drawsize <- (round(nrow(articles.2004) * 0.025 / nauth) * nauth) / nauth  # number divisible by 
-drawsize <- 400
-set.seed(432)
[…]

### 2016-04-02 Rmd/calibration.Rmd
Bulk commit of various changes, including library functions added. Edits to space-time.Rmd include near-final version of figure.

---
title: "calibration"
author: "Lyndon Estes"
date: "January 8, 2016"
output: 
  html_document:
    toc: yes
    fig_caption: true
---

# Calibration data
```{r, message=FALSE, warning = FALSE}
library(ecoscales)
library(readxl)
library(Agreement)
library(irr)

p_root <- set_base_path("ecoscales")
p_dat <- fp(p_root, "external/data/calibration")

# Read in and clean-up calibration data
# cal <- data.table(read_excel(fp(p_dat, "merged_calibration.xlsx")))
cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
cal[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
cal[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
dropnm <-  c("V30", "V31", "journal", "exclusion reason", "sensitivity")
cal[, c(dropnm) := NULL]

# check data types
for(j in names(cal)[8:24]) set(cal, j = j, value = as.numeric(cal[[j]]))

# use NA in n_sites as indicator that study was excluded
cal[, inout := ifelse(is.na(n_sites), 0, 1)]  
# write.csv(cal, file = fp(p_dat, "merged_calibrationf.csv"))

# external fixes and recodes, bring back in
cal <- data.table(read_excel(fp(p_dat, "merged_calibrationf.xlsx")))
cal[, notes := NULL]
for(j in names(cal)[c(4:5, 8:24)]) set(cal, j = j, value = as.numeric(cal[[j]]))
# length(which(is.na(cal$feature)))

```

## Agreement on studies included/excluded
```{r}
# ndois <- cal[, unique(DOI)]  # check on number of journals

# reduce within-observer DOI duplicates 
dois <- cal[, unique(DOI), by = .(observer, inout)]

# recast dataset to obsevers in columns, inout as value, DOI as case
dois_c1 <- dcast(dois, V1 ~ observer, value.var = "inout")

# Fleiss's Kappa
kappam.fleiss(dois_c1[, 2:7, with = FALSE])
```

## Inter-rater reliability using number of records per paper
```{r}
doisum <- dcast(cal, DOI ~ observer, value.var = "inout", fun = sum)
# kappam.fleiss(doisum[, 2:7, with = FALSE])
icc(ratings = doisum[, 2:7, with = FALSE], model = "twoway")

```

```{r, warning = FALSE, fig.cap="Figure 1: Number of records extracted by each rater (observer) for each of the 20 papers in the calibration set."}
doimat <- t(doisum[, -1, with = FALSE])
cols <- RColorBrewer::brewer.pal(n = 6, "Paired")#[c(3:1, 4:6)]
cxs <- seq(3, 0.8, length.out = 6)
pchs <- c(rep(20, 5), 4)
par(mar = c(4, 4, 0.1, 0.1))
plot(c(1, 20), c(0, 10), pch = "", xlab = "Paper reviewed", 
     ylab = "N extractable records", mgp = c(1.45, 0.7, -0.2))
polygon(x = c(0, 21, 21, 0, 0), y = c(-1, -1, 11, 11, 0), col = "grey70")
# plot(doimat[1, ], pch = 20, xlab = "Paper reviewed", 
#      ylab = "N extractable records", cex = cxs[1], col = cols[1])
for(i in 1:6) points(doimat[i, ], col = cols[i], pch = pchs[i], cex = cxs[i])
legend(x = "topleft", pch = pchs, pt.cex = cxs, col = cols, 
       legend = paste0("Rater", 1:6), bty = "n")

```

## Variance in extracted statistics
```{r}
varnms <- c("plot_res", "n_sites", "sampled_area", "t_btwn_samp", 
            "study_duration")
mus <- cal[inout == 1, lapply(.SD, mean, na.rm = TRUE), by = .(DOI, observer),             .SDcols = varnms]
vars <- mus[, lapply(.SD, sd, na.rm = TRUE), by = DOI, .SDcols = varnms]
musmu <- mus[, lapply(.SD, mean, na.rm = TRUE), by = DOI, .SDcols = varnms]

par(mfrow = c(3, 2), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
for(i in varnms) {
  mu <- musmu[, get(i)]
  sdev <- vars[, get(i)]
#   print(paste(i, ":", sdev))
  print(paste(i, ":", round(mu, 2), ":", round(sdev, 2), ":", 
              round(sdev / mu, 2)))
  # plot(mu, sdev / mu, xlab = "mean", ylab = "CV", main = i)
  plot(mu, sdev, xlab = "mean", ylab = "sd", main = i)
}

vars[, lapply(.SD, mean, na.rm = TRUE), .SDcols = varnms]
colMeans(sapply(varnms, function(i) {
  mu <- musmu[, get(i)]
  sdev <- vars[, get(i)]
  sdev / mu
}), na.rm = TRUE)

```

```{r}
par(mfcol = c(10, 4), mar = c(2, 2, 2, 2), mgp = c(1.5, 0.5, 0))

mus[, plot(plot_res), by = DOI]

```

Tim needs to fix paleo studies in calibration (and main set) giving too much credit to effective area.

Others need to fix their study duration figures. Paul on paleo one in calibration, Choi, Chang, etc in main set.

For N-sites, count sites only once if there are repeated measures taken at them, e.g. Jonathan in 10.1111/j.1466-8238.2009.00454.x for bird counts

### 2016-04-02 Rmd/space-time.Rmd
Bulk commit of various changes, including library functions added. Edits to space-time.Rmd include near-final version of figure.

@@
     toc: yes
+    number_sections: true
+    toc_depth: 3
 ---
@@
 ```{r, eval = FALSE}
-library(data.table)
-library(raster)
-library(sp)
-require(splancs)
-library(lmisc)
+library(ecoscales)
+library(readxl)
+library(stringdist)
+library(maptools)
+
 p_root <- set_base_path("ecoscales")
-p_dat <- full_path(p_root, "external/data")
-dat <- fread(full_path(p_dat, "merged-data.csv"), stringsAsFactors = FALSE)
+p_dat <- full_path(p_root, "external/data/result")
+# p_dat <- full_path(p_root, "external/data/")
+# dat <- fread(full_path(p_dat, "merged-data.csv"), stringsAsFactors = FALSE)
+dat <- as.data.table(read_excel(full_path(p_dat, "merged.xlsx")))
 setnames(dat, "DOI/title", "DOI")
@@
 ## Spatal and temporal data 
+### Adjustments
+Prior to scaling the data in log10 space, we need to make some adjustments.
+
+1.
+Time between samples has many 0 values because many observations are simply one-offs.
+Set these to an arbitrarily large value which will represent clear separation on axis from high frequency studies
+
+```{r, eval=FALSE}
+# dat[t_btwn_samp > 0.000001 & t_btwn_samp < 0.001, ]
+# dat[t_btwn_samp <0]
+# dat[t_btwn_samp == 0, t_btwn_samp := 365 * 100000]  # interval adjust
+# dat[study_duration > 365 * 1000]
+
+### Inconsistency in study_duration - multiplying by N sites in many cases
+# Paul a little bit, Ahmed, Chang. Jonathan a few places also. 
+```
+
+### Bootstrap on uncertainties
+
+Just for the four main variables. Rules:
+
++ If `plot_res` and/or `n_sites` uncertain, then `sampled_area` must be. Recalculate using random draw within the uncertain variable 
++ If `t_btwn_samp` and/or `samp_duration` uncertain, do the same as above, but we have no variable for n repeats for calculating `study_duration`.
+
+```{r, eval = FALSE}
+svars <- c("plot_res", "n_sites", "sampled_area", "samp_duration",
+           "t_btwn_samp", "study_duration")
+
+sens <- gsub("\\\\", ";", dat$sensitivity)
+sens <- gsub(", ", ";", gsub("; ", ";", sens))
+sens <- tolower(sens)
+sensl <- strsplit(sens, ";")
+unique(unlist(sensl))
+
+# sensul <- unlist(sensl)
+# smatch <- adist(sensul, svars, partial = TRUE, ignore.case = TRUE)
+# min_name <- apply(smatch, 1, min)
+# matchs1s2 <- NULL  
+# for(i in 1:nrow(smatch)) {  # i <- 10
+#     s2i <- match(min_name[i], smatch[i, ])
+#     s1i <- i
+#     ms1s2 <- rbind(data.frame(s2i=s2i, s1i=s1i,
+#                               s2name = svars[s2i],
+#                               s1name = sensul[s1i],
+#                               adist = min_name[i]), ms1s2)
+# }
+# # and we then can have a look at the results
+# View(ms1s2)
+
+sens2 <- gsub("plo_", "plot_", sens)  # fix this one
+sens2 <- gsub("[[:blank:]]", "_", sens2)  # replace spaces with _
+
+corrtab <- cbind(c("^res", "plo_res", "sample_area", "samples_area", 
+                   "sampled area",
+                   "t_btwn_samples", "t_btwn_sample", "t_btwn_samp_samp", 
+                   "t_twn_samp", "t_btw_samp", "samp_duratiion", 
+                   "sampling_duration", "sample_duration", 
+                   "plot_resolution", "number_of_sites", 
+                   "time_between_sampling", "plot_size", 
+                   "time_between_samples", "samp_study", ":", " ", ";_"), 
+                 c(svars[c(1, 1, 3, 3, 3, 5, 5, 5, 5, 5, 4, 4, 4, 1, 2,
+                           5, 1, 5)], "samp;study", ";", ";", ";"))
+for(i in 1:nrow(corrtab)) {
+  sens2 <- gsub(corrtab[i, 1], corrtab[i, 2], sens2)
+}
+# gsub("[a-z]( )[a-z]", "_", sens[[228]])
+# sens[[228]]
+# sens2[is.na(sens2)]
+sens2[is.na(sens2)] <- "none"
+sensl2 <- lapply(1:length(sens2), function(x) strsplit(sens2[x], ";")[[1]])
+# unique(unlist(sensl))
+unique(unlist(sensl2))
+
+# create a sensitivity table
+senst <- do.call(cbind.data.frame, lapply(svars, function(x) {
+  v <- sapply(sensl2, function(y) ifelse(any(y == x), 1, 0))
+  # ifelse(is.na(v), 0, v)
+}))
+colnames(senst) <- c("res", "n", "sarea", "sampd", "tbtwn", "studyd")
+
+# add column for uncertainty about n repeats
+# senst$nrep <- rep(0, nrow(senst))
+
+# i <- 300:305
+# sensl2[i]
+# senst[i, ]
+
+# apply fixes (mostly to account for observer omissions)
+# 1. if samp_duration != study_duration and t_btwn_samp == 0, set 
+#    study_duration == samp_duration
+# change made in dat
+# dat[which(dat$t_btwn_samp == 0 & dat$study_duration == 1), ]
+names(dat)
+dat[t_btwn_samp == 0 & (study_duration != samp_duration), c(8:13), 
+    with = FALSE]
+# which(dat$t_btwn_samp == 0 & senst$studyd == 1)
+dat[t_btwn_samp == 0 & study_duration == 1, 
+    study_duration := samp_duration]
+# dat[t_btwn_samp == 0 & study_duration == 1, ]
+
+# 2. if samp_duration OR t_btwn_samp is uncertain, study_duration must also be. 
+id <- which((senst$sampd == 1 | senst$tbtwn == 1) & senst$studyd == 0)
+senst[id, "studyd"] <- 1  
+# senst[id, ]
+
+# but no real need to set it to 1 because it will get automatically perturbed
+# via n_repeats
+
+# which(dat$t_btwn_samp == 0 & dat$samp_duration == 1)
+# which(dat$t_btwn_samp == 0 & dat$samp_duration == 1)
+# which(dat$t_btwn_samp == 0 & senst$sampd == 1)
+
+# 3. if study_duration is uncertain but t_btwn and samp_duration are not, treat
+# both as uncertain
+id <- which(senst$tbtwn == 0 & senst$sampd == 0 & senst$studyd == 1)
+senst[id, c("sampd", "tbtwn")] <- 1
+# senst[id, ]
+
+# although it is possible that just n_repeats is uncertain, but more likely 
+# that observers didn't note this is sensitive.
+
+# id <- which(senst$tbtwn == 0 & senst$sampd == 0 & senst$studyd == 1)
+# which(senst$sampd == 0 & senst$studyd == 1)
+# dat[which(senst$sampd == 0 & senst$studyd == 1), ]
+
+# 4. Can t_btwn_samps be 0 and study_duration be uncertain if samp_duration is 
+# not?  No, it shouldn't be.
+
+# note situations in which neither plot_res or n_samples are certain and 
+# study_area is not
+any(which(senst$res == 0 & senst$sarea == 1) %in% 
+      which(senst$n == 0 & senst$sarea == 1))
+which(senst$res == 1 & senst$n == 1 & senst$sarea == 1)
+which(senst$res == 1 & senst$sarea == 1)
+which(senst$n == 1 & senst$sarea == 1)
+which((senst$res == 1 & senst$sarea == 1) | (senst$n == 1 & senst$sarea == 1))
+```
+
+### Run resampling with 
+  - With 10% variation
+  - With 20% variation
+  - With 30%
+  - Linked to calibration stdev
+
+```{r, eval=FALSE}
+# function to recalculate study_duration by finding hidden sampling frequency 
+# variable from study_duration and samp_duration. 
+recalc_duration <- function(study, samp, tbtwn) {
+  ifelse(tbtwn == 0, samp, (study / samp) * samp)
+}
+
+# function to set t_btwn_samp to 365 * 100000 to indicate once-off studies 
+# when t_btwn_samp == 0.  
+to_infin_byond <- function(tbtwn) ifelse(tbtwn == 0, 365 * 10000, tbtwn)
+
+# dat[, round(t_btwn_samp, 6) == 0]
+# dat[1, n_repeats(t_btwn_samp, samp_duration, study_duration) * samp_duration]
+# round(dat[, recalc_duration(study_duration, samp_duration)] - 
+#         dat$study_duration, 8)
+
+# set.seed(123)
+# spct <- rnorm(1000, mean = 1, sd = 0.3)
+# # hist(runif(1000, min = 0.8, max = 1.2))
+# set.seed(123)
+
+# hist(rgamma(100, 1, rate = 0.5))
+# hist(rlnorm(100, 0.1, sd = 1))
+# hist(rexp(100, rate = 2))
+
+p <- 0.2
+lapply(1:10, function(x, slen = 1, p = 0) {
+
+  # a <- rnorm(1, mean = 1, sd = 0.3)
+  # b <- rnorm(1, mean = 1, sd = 0.3)
+  a <- runif(slen, min = 1 - p, max = 1 + p)  # plot_res perturb
+  b <- runif(slen, min = 1 - p, max = 1 + p)  # n_sites perturb
+  d <- runif(slen, min = 1 - p, max = 1 + p)  # samp_duration perturb
+  e <- runif(slen, min = 1 - p, max = 1 + p)  # t_btwn_samp perturb
+
+  dnew <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
+                       t_btwn_samp, study_duration)])
+  dnew[which(senst$res == 1), plot_res := plot_res * a]
+  dnew[which(senst$n == 1), n_sites := n_sites * b]
+  dnew[, sampled_area := (plot_res * n_sites) / 10000]
+  # plot(dnew$sampled_area, dat$sampled_area)
+  # dat[sampled_area < 1.2e+8, hist(sampled_area)]
+
+  # check for incorrect calculations
+  # dnew[which(senst$n == 1), (plot_res * n_sites) / 10000]
+  # dnew[sampled_area < 1e+9, plot((plot_res * n_sites) / 10000, sampled_area)]
+  # dat[sampled_area < 1e+7, plot((plot_res * n_sites) / 10000, sampled_area)]
+  # ii <- which(dat[, round((plot_res * n_sites) / 10000 - sampled_area, 4)] 
+  #            != 0)
+  # dat[ii, c(1:3, 8:10), with = FALSE]
+  # dat[56, c(1:3, 8:10), with = FALSE]
+  # dat[ii, round((plot_res * n_sites) / 10000 - sampled_area, 4)]
+
+  # temporal variables
+  dnew[which(senst$sampd == 1), samp_duration := samp_duration * d]
+  dnew[which(senst$tbtwn == 1), t_btwn_samp := t_btwn_samp * e]
+  dnew[, study_duration := {
+    recalc_duration(study_duration, samp_duration, t_btwn_samp)
+  }]
+  dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]
+  # dnew[7, recalc_duration(study_duration, samp_duration, t_btwn_samp)]
+
+  round(dat$study_duration - dnew$study_duration, 4)
+  # dnew[1:20, ]
+#   dnew[which(senst$n == 1), n_sites]
+#   tst <- which(dat$plot_res > 1)
+#   cbind(dat[tst, log10(plot_res)], dnew[tst, log10(plot_res)])
+#   cbind(dat[which(senst$res == 1), plot_res], 
+#         dnew[which(senst$res == 1), plot_res])
+#   plot(dat$n_sites, dnew$n_sites)
+})
+
+# full function for resampling
+resamp_func <- function(dat, p, iter) {
+  bout <- lapply(1:iter, function(x) {  # x <- 1
+    if((x / 100) %in% 1:100) print(x)
+    # amount to perturb
+    a <- runif(1, min = 1 - p, max = 1 + p)  # plot_res perturb
+    b <- runif(1, min = 1 - p, max = 1 + p)  # n_sites perturb
+    d <- runif(1, min = 1 - p, max = 1 + p)  # samp_duration perturb
+    e <- runif(1, min = 1 - p, max = 1 + p)  # t_btwn_samp perturb
+
+    # spatial variables
+    dnew <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
+                         t_btwn_samp, study_duration)])
+    dnew[which(senst$res == 1), plot_res := plot_res * a]
+    dnew[which(senst$n == 1), n_sites := n_sites * b]
+    dnew[, sampled_area := (plot_res * n_sites) / 10000]
+
+    # temporal variables
+    dnew[which(senst$sampd == 1), samp_duration := samp_duration * d]
+    dnew[which(senst$tbtwn == 1), t_btwn_samp := t_btwn_samp * e]
+    dnew[, study_duration := {
+      recalc_duration(study_duration, samp_duration, t_btwn_samp)
+    }]
+    dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
+  })
+  return(bout)
+}
+
+# perturbed samples
+p10 <- resamp_func(dat, 0.1, 1000)  # 10% perturb
+p20 <- resamp_func(dat, 0.2, 1000)  # 20% perturb
+p30 <- resamp_func(dat, 0.3, 1000)  # 30% perturb
+
+# bootstrap
+bootsamp <- lapply(1:1000, function(x) {
+  if((x / 100) %in% 1:100) print(x)
+  dnew <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
+                       t_btwn_samp, study_duration)])
+  dnew <- dnew[sample(1:.N, size = .N, replace = TRUE), ]
+  dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
+  dnew
+})
+
+# plot(bootsamp[[3]]$plot_res, dat$plot_res)
+# plot(bootsamp[[3]]$plot_res, bootsamp[[10]]$plot_res)
+# plot(bootsamp[[1]]$plot_res, bootsamp[[10]]$plot_res)
+# sd(sapply(bootsamp, function(x) mean(x$plot_res)))
+# mean(bootsamp[[3]]$plot_res - bootsamp[[10]]$plot_res)
+
+# a <- 10
+# b <- 10
+# 
+# u <- 0.2
+# rnorm(100, mean = a, sd = 0.2 * a)
+# a * b
+
+# dat[t_btwn_samp == 0, t_btwn_samp := 365 * 100000]  # interval adjust
+
+```
 
 ### Scaling in log10 space
-```{r}
-i <- 0.000001
-j <- rep(0, 15)
+```{r, eval = FALSE}
+i <- 0.0000001
+j <- rep(0, 16)
 for(k in 1:length(j)) {
@@
 }
+
+# temporal scales
 tdt <- data.table("scaleval" = j, "time" = j)
@@
 
-# label axis for time
-taxis <- cbind.data.frame(
-  c("Sec", "Min", "Hour", "Day", "Month", "Year", "Decade", "Century", 
-    "Millenium", "10000"),  
-  c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 30, 365, 365 * 10, 
+# labels for temporal axes
+# return interval
+tlab1 <- c(expression("second"^-1), expression("minute"^-1),
+           expression("hour"^-1), expression("day"^-1), 
+           expression("week"^-1), expression("month"^-1),            
+           expression("year"^-1), expression("decade"^-1),
+           expression("century"^-1), 
+           expression("millenium"^-1), "once-off")# expression(infinity))
+taxis1 <- cbind.data.frame( 
+  c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 30, 365, 365 * 10, 
     365 * 100, 365 * 1000, 365 * 10000))
-taxis <- data.table(taxis)
-setnames(taxis, names(taxis), c("label", "days"))
-taxis[, logdays := log10(days)]
+taxis1 <- data.table(taxis1)
+setnames(taxis1, names(taxis1), "days")
+taxis1[, logdays := log10(days)]
 
-rdt <- data.table("scaleval" = j, "area" = j^2)
-rdt[, slog := log10(scaleval)]
-rdt[, rlog := log10(area)]
+taxis2 <- cbind.data.frame(
+  c("second", "minute", "hour", "day", "week", "month", "year", "decade",
+    "century", "millenium", "10 KA"),  
+  c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 30, 365, 365 * 10, 
+    365 * 100, 365 * 1000, 365 * 10000))
+taxis2 <- data.table(taxis2)
+setnames(taxis2, names(taxis2), c("label", "days"))
+taxis2[, logdays := log10(days)]
 
-adt <- data.table("area" = j)
-adt[, alog := log10(area)]
+# spatial
+# plot(1:10, xlab = expression(paste("mm"^2)), xaxt = "n")
+# axis(1, at = 1:8, labels = lab, las = 2)
+
+# plot resolution
+alab1 <- c(expression("0.01 cm"^2), expression("0.1 cm"^2), 
+           expression("1 cm"^2), expression("10 cm"^2),
+           expression("100 cm"^2), expression("1000 cm"^2), 
+           expression("1 m"^2), expression("10 m"^2),
+           expression("100 m"^2), expression("1000 m"^2), 
+           "1 ha", "10 ha", "100 ha", "1000 ha", "10000 ha")
+
+aaxis1 <- cbind.data.frame(c(0.000001, 0.00001, 0.0001, 0.001, 0.01, 0.1,
+                             1^2, 10, 10^2, 10^3, 10^4, 10^5, 10^6, 10^7, 10^8))
+aaxis1 <- data.table(aaxis1)
+setnames(aaxis1, names(aaxis1), "res")
+aaxis1[, logres := log10(res)]
+
+# sampled area
+alab2 <- c(expression("1 m"^2), expression("10 m"^2), expression("100 m"^2),
+           expression("1000 m"^2), expression("1 ha"), expression("10 ha"),
+           expression("100 ha"), expression("1000 ha"), 
+           expression(paste(10^4 , " ha")), expression(paste(10^5 , " ha")), 
+           expression(paste(10^6 , " ha")), expression(paste(10^7 , " ha")), 
+           expression(paste(10^8 , " ha")), expression(paste(10^9 , " ha")), 
+           expression(paste(10^10 , " ha")))
+
+aaxis2 <- cbind.data.frame(
+  c(0.1 / 10000, 10 / 10000, 1000 / 10000, 1, 10, 10^2, 10^3, 10^4, 10^5, 10^6,
+    10^7, 10^8, 10^9, 10^10))
+aaxis2 <- data.table(aaxis2)
+setnames(aaxis2, names(aaxis2), "area")
+aaxis2[, logarea := log10(area)]
+
+```
+
+### Boot-strapped resolution versus extent
+```{r, eval = FALSE}
+p30dt <- rbindlist(p30)
+resdat <- data.table(x = log10(p30dt$t_btwn_samp), y = log10(p30dt$plot_res))
+resdat[y < min(aaxis2), y := min(aaxis2)]
+resdat[y > max(aaxis2), y := max(aaxis2)]
+resdat$n <- rep(1, nrow(resdat))
+coordinates(resdat) <- ~x + y
+rx <- range(resdat$x)
+# rx <- c(floor(rx[1]), ceiling(rx[2]))
+rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
+rx[2] <- 7.6
+ry <- range(aaxis1$logres)
+
+rkd <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat, 0.8)
+
+par(mar = c(5, 5, 3, 3))
+# plot(rkd)
+image(rkd, col = rev(terrain.colors(20)), axes = FALSE, xlab = "", 
+      ylab = "", xlim = rx, ylim = ry)
+axis(1, at = taxis1$logdays[-5], labels = tlab1[-5], las = 2)
+axis(2, at = aaxis2$logres, labels = alab2, las = 2)
+
+# main plot
+# bootdt
+# resolution
+bootdt <- rbindlist(bootsamp)
+resdat2 <- data.table(x = log10(bootdt$t_btwn_samp), y = log10(bootdt$plot_res))
+resdat2$n <- rep(1, nrow(resdat2))
+coordinates(resdat2) <- ~x + y
+rx <- range(resdat2$x)
+rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
+rx[2] <- 6.56
+# rx[2] <- 7.6
+
+# extent
+extdat2 <- data.table(x = log10(bootdt$study_duration), 
+                      y = log10(bootdt$sampled_area))
+extdat2$n <- rep(1, nrow(extdat2))
+coordinates(extdat2) <- ~x + y
+ex <- range(taxis2$logdays)
+ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
+# ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
+# ex[2] <- 7.6
+ey <- range(aaxis2$logarea)
+
+# densities
+# resolution
+rres2 <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat2, 1)
+rres2 <- (rres2 / cellStats(rres2, sum)) * 100
+rng <- range(rres2[is.finite(rres2)])
+bwidth <- (rng[2] - rng[1]) / 30
+brks <- seq(rng[1], rng[2], bwidth)
+brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
+
+# extent
+rext2 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat2, 1)
+rext2 <- (rext2 / cellStats(rext2, sum)) * 100
+rng <- range(rext2[is.finite(rext2)])
+bwidth <- (rng[2] - rng[1]) / 30
+brks2 <- seq(rng[1], rng[2], bwidth)
+brklabs2 <- round(brks2[seq(1, length(brks2), 3)], 2)
+
+cxa = 0.7
+pdf("paper/figures/res_v_extent_bstrap.pdf", width = 7, 
+    height = 2.5)
+par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
+    mgp = c(2, 0.5, 0))
+image(rres2, col = rev(terrain.colors(30)), breaks = brks,
+     axes = FALSE, xlab = "", ylab = "")
+# plot(rres2, col = rev(terrain.colors(20)))
+axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
+     las = 2, tcl = -0.2, cex.axis = cxa)
+axis(2, at = aaxis1$logres, labels = alab1, las = 2, tcl = -0.2, 
+     cex.axis = cxa)
+# plot(rres2, axes = FALSE, legend.only = TRUE)
+aargs <- list(mgp = c(3, 1, 0), at = brklabs, labels = brklabs, 
+              cex.axis = cxa)
+plot(rres2, legend.only = TRUE, axis.args = aargs)
+
+image(rext2, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
+      ylab = "", breaks = brks2)
+axis(1, at = taxis2$logdays[-5], labels = taxis2$label[-5], las = 2,
+     cex.axis = cxa, tcl = -0.2)
+axis(2, at = aaxis2$logarea, labels = alab2[-2], las = 2, cex.axis = cxa,
+     tcl = -0.2)
+aargs <- list(mgp = c(3, 1, 0), at = brklabs2, labels = brklabs2, 
+              cex.axis = cxa)
+plot(rext2, legend.only = TRUE, axis.args = aargs)
+dev.off()
+
+# image(rkd2 > qtiles[40])
+
+v <- values(rkd2)
+v <- v[(v > 0) & !is.na(v)]
+qtiles <- quantile(v, probs = seq(0, 1, 0.01))
+round(qtiles / sum(qtiles) * 100, 2)
+sum(v[v > qtiles[19]]) / sum(v)
+bootdt
+sum(qtiles)
+
+length(which(dat$study_type %in% 
+               c("remote sensing", "other geographic data"))) / nrow(dat)
+dat[t_btwn_samp == 0, .N] / dat[, .N]  # 31% of samples are once-offs
+dat[study_duration == samp_duration, .N]
+
+p50 <- rkd2 >= qtiles[11]
+p75 <- rkd2 >= qtiles[16]
+
+par(mfrow = c(2, 2))
+bootdt[, {
+  hist(log10(plot_res), main = "Plot res")
+  hist(log10(t_btwn_samp), main = "Frequency")
+  hist(log10(sampled_area), main = "Area")
+  hist(log10(study_duration), main = "Duration")
+}]
+
+plot(rkd2)
+image(rkern, add=TRUE, col=rev(terrain.colors(20)))
+plot(ContourLines2SLDF(contourLines(rkern)))
+
+dat[, {
+  plot(log10(plot_res), log10(sampled_area))
+  plot(log10(t_btwn_samp), log10(study_duration))
+}]
+
+points(bodmin$x, bodmin$y)
+data(bodmin)
+plot(bodmin$poly, asp=1, type="n")
+image(kernel2d(as.points(bodmin), bodmin$poly, h0=2, nx=100, ny=100), 
+      add=TRUE, col=terrain.colors(20))
+plot(p50)
+plot(p75)
+p50poly <- rasterToPolygons(p50, dissolve = TRUE)
+p75poly <- rasterToPolygons(p75, dissolve = TRUE)
+
+plot(p50poly, col = "transparent", add = TRUE)
+plot(p75poly, col = "transparent", add = TRUE)
 
@@
 ```{r,eval = FALSE}
-rres <- 0.25
-# # raster for resolution (x dimension) versus area (y dimension)
-# # arst <- raster(xmn = -6, xmx = 9, ymn = -5, ymx = 10)
-# rst <- raster(xmn = -5, xmx = 10, ymn = -5, ymx = 10)
-# res(rst) <- rres
-# # rst@crs <- CRS("")
-# rst[] <- 1
-# rst2 <- raster(xmn = 0, xmx = 10, ymn = 0, ymx = 10)
-# # res(rst2) <- c(0.25, 0.25)
-# res(rst2) <- rres
-# # rst@crs <- CRS("")
-# rst2[] <- 1
-# plot(rst2)
-
-resdat <- data.table(x = log10(dat$t_btwn_samp + 0.00001), 
-                     y = log10(dat$plot_res))
-# resdat[x < -5, x := -4.5]
-# resdat[y < -5, y := -4.5]
+# resolution data
+resdat <- data.table(x = log10(dat$t_btwn_samp), y = log10(dat$plot_res))
 resdat$n <- rep(1, nrow(resdat))
 coordinates(resdat) <- ~x + y
-# plot(resdat$x, resdat$y)
 rx <- range(resdat$x)
-# rx <- c(floor(rx[1]), c(ceiling(rx[2])))
+# rx <- range(aaxis$logarea)
 rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-rxd <- length(seq(rx[1], rx[2], rres))  # 0.25))
-ry <- range(resdat$y)
-# ry <- c(floor(ry[1]), c(ceiling(ry[2])))
+# rx[1] <- -6.5  # set temporal interval to -6.5 to have null returns centered
+# ry <- range(resdat$y)
+ry <- range(aaxis$logarea)
 ry <- round(ry / 0.5) * 0.5  # round to nearest 0.5
-ryd <- length(seq(ry[1], ry[2], rres))#0.25))
 
-# length(seq(-5, 4, 0.25))
+# extent data
+# fix area rounded to zero in Jon's plots
+extdat <- data.table("x" = log10(dat$study_duration), 
+                     "y" = log10(dat$sampled_area)) 
+# extdat[is.infinite(x), ]; extdat[is.infinite(y), ]
 
-extdat <- data.table("x" = log10(dat$study_duration + 0.00001), 
-                     "y" = log10(dat$sampled_area + 0.00001)) 
-# extdat <- data.table("x" = log10(dat$study_span), 
-#                      "y" = log10(dat$sampled_area)) 
-# extdat[x < 0, x := 0]
-# extdat[y < 0, y := 0]  # setting minima to 1 day and 1 ha
+### Note: we are going to have revisit assumption of 1 day duration on 
+#   sampling duration. Some people might have just put 1 day
+# plot(extdat$x, extdat$y)
+
 extdat$n <- rep(1, nrow(extdat))
 coordinates(extdat) <- ~x + y
-# plot(extdat$x, extdat$y)
-ex <- range(extdat$x)
-# ex <- c(floor(ex[1]), c(ceiling(ex[2])))
+# ex <- range(extdat$x)
+ex <- range(taxis2$logdays)
 ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
-ex[1] <- -5
-exd <- length(seq(ex[1], ex[2], rres))#0.25))
+# ex[1] <- -5
 ey <- range(extdat$y)
-# ey <- c(floor(ey[1]), c(ceiling(ey[2])))
 ey <- round(ey / 0.5) * 0.5  # round to nearest 0.5
-eyd <- length(seq(ey[1], ey[2], rres))#0.25))
 
-# set-up grid topologies
-rgrd <- GridTopology(cellcentre.offset = c(rx[1], ry[1]) + 0.125, 
-                     #cellsize = c(0.25, 0.25), cells.dim = c(rxd, ryd))
-                     cellsize = c(rres, rres), cells.dim = c(rxd, ryd))
-egrd <- GridTopology(cellcentre.offset = c(ex[1], ey[1]) + 0.125, 
-                     #cellsize = c(0.25, 0.25), cells.dim = c(exd, eyd))
-                     cellsize = c(rres, rres), cells.dim = c(exd, eyd))
-# seq(0, 10, 10 / 59)
-
-rst <- raster(xmn = rx[1], xmx = rx[2], ymn = ry[1], ymx = ry[2])
-# rst <- raster(xmn = min(resdat$x), xmx = ceiling(max(resdat$x)), 
-#               ymn = min(resdat$y), ymx = ceiling(max(resdat$y)))
-res(rst) <- rres
-# rst@crs <- CRS("")
-rst[] <- 1
-# plot(rst)
-rst2 <- raster(xmn = ex[1], xmx = ex[2], ymn = ey[1], ymx = ey[2])
-# rst2 <- raster(xmn = min(extdat$x), xmx = ceiling(max(extdat$x)), 
-#               ymn = min(extdat$y), ymx = ceiling(max(extdat$y)))
-res(rst2) <- rres
-# rst@crs <- CRS("")
-rst2[] <- 1
-# plot(rst2)
-
-# kernel densities
-# resolution
-ppoly <- as(extent(rst), "SpatialPolygons")
-crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
-kern1 <- spkernel2d(resdat, crds, h0 = 1, rgrd)
-kgrid1 <- as(SpatialGridDataFrame(rgrd, data = data.frame(kern1)),
-             "SpatialPixelsDataFrame")
-
-# extent
-ppoly <- as(extent(rst2), "SpatialPolygons")
-crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
-kern2 <- spkernel2d(extdat, crds, h0 = 1, egrd)
-kgrid2 <- SpatialGridDataFrame(egrd, data = data.frame(kern2))
-
-# rasterize
-# space
-rr <- raster(kgrid1)
-rr[is.na(rr)] <- 0
-# time
-er <- raster(kgrid2)
-er[is.na(er)] <- 0
+# density rasters
+# vary the kernel size about them
+kernsl <- lapply(c(0.4, 0.6, 0.8, 1), function(x) {
+  rr <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat, x)
+  er <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat, x)
+  kerns <- list("res" = rr, "ext" = er)
+})
 
 # have a look
-par(mfrow = c(2, 1))
-plot(er, axes = FALSE, add = FALSE)
-arg <- list(at = taxis$logdays, labels = taxis$label)
-image(er, col = rev(terrain.colors(100)), axes = FALSE, xlab = "", ylab = "", 
-      xlim = c(-5, 6))
-axis(1, at = taxis$logdays, labels = taxis$label, las = 2)
-axis(2, at = -5:10, las = 2)
-
+# par(mfrow = c(2, 1))
+# plot(er)
+# par(mar = c(3, 3, 3, 5))
+# plot(er, axes = FALSE, add = FALSE, legend = FALSE, box = FALSE)
+# image(er, col = rev(terrain.colors(100)), axes = FALSE, xlab = "", ylab = "")
+# plot(er, legend.only = TRUE)
+# axis(1, at = taxis$logdays, labels = taxis$label, las = 2)
+# axis(2, at = -5:10, las = 2)
 # points(extdat$x, extdat$y)
-plot(rr)
+# plot(rr)
 # points(resdat$x, resdat$y)
 
-ers <- er
-rrs <- rr
 ```
 
-### Convert to rgb space and plot
+### Plot
+#### Varying kernel sizes
 ```{r, eval = FALSE}
 # As separate plots
-png("paper/figures/res&extent2.png", width = 900, height = 450)
-par(mfrow = c(1, 2), oma = c(0, 0, 0, 0), mar = c(4, 4, 4, 4))
-plot(rrs, xlim = c(-5, 15), ylim = c(-5, 15), 
-     xlab = "Sample interval - log10(days)", 
-     ylab = "Plot resolution - log10(m^2)")
-plot(ers, box = FALSE, xlim = c(-5, 15), ylim = c(-5, 15), 
-     xlab = "Study duration - log10(days)", ylab = "Sampled area - log10(ha)")
+pdf("paper/figures/res_v_extent_ksize.pdf", width = 7, height = 12, )
+par(mfrow = c(4, 2), oma = c(2, 2, 0, 0), mar = c(4, 4, 4, 4))
+lapply(kernsl, function(x) {
+  # x <- kernsl[[1]]
+  image(x$res, col = rev(terrain.colors(100)), axes = FALSE, xlab = "", 
+        ylab = "", ylim = range(aaxis$logarea))
+  axis(1, at = taxis1$logdays, labels = tlab1, las = 2)
+  # axis(1, at = taxis$logdays, labels = taxis$label, las = 2)
+  axis(2, at = aaxis$logarea, labels = alab1, las = 2)
+  plot(x$res, axes = FALSE, legend.only = TRUE)
+  image(x$ext, col = rev(terrain.colors(100)), axes = FALSE, xlab = "", 
+        ylab = "")
+  axis(1, at = taxis2$logdays, labels = taxis2$label, las = 2)
+  axis(2, at = aaxis$logarea, labels = alab1, las = 2)
+  # axis(2, at = ey[1]:ey[2], las = 2)
+  plot(x$ext, legend.only = TRUE)
+})
+par(xpd = NA)
+xx <- grconvertX(0.5, from = "ndc", to = "inches")
+ii <- c(0.4, 0.6, 0.8, 1)
+yy <- c(0.95, 0.75, 0.5, 0.25)
+# yy <- c(-1, -5, -10, -20)
+for(i in 1:length(ii)) {
+  y <- grconvertY(yy[i], from = "ndc", to = "inches")
+  # text(x = xx, y = y, labels = paste("kernel =", ii[i]))
+  # par(xpd = NA)
+  # mtext(text = paste("kernel =", ii[i]), side = 3, line = yy[i])
+}
 dev.off()
+
+# plot(rrs, xlim = c(-5, 15), ylim = c(-5, 15), 
+#      xlab = "Sample interval - log10(days)", 
+#      ylab = "Plot resolution - log10(m^2)")
+# plot(ers, box = FALSE, xlim = c(-5, 15), ylim = c(-5, 15), 
+#      xlab = "Study duration - log10(days)", ylab = "Sampled area - log10(ha)")
+# dev.off()
 ```

### 2016-04-02 R/kdensity.R

@@ -5,16 +5,17 @@
 #' @param ymn Minimum y coordinate for output grid
 #' @param ymx Maximum y coordinate for output grid
 #' @param dxdy Resolution for output grid
-#' @param pts SpatialPointsDataFrame showing distribution of observations 
+#' @param pts SpatialPointsDataFrame showing distribution of observations
+#' @param kwidth Kernel smoothing width 
 #' @return Raster of density values
 #' @export
-kdensity <- function(xmn, xmx, ymn, ymx, dxdy, pts) {
+kdensity <- function(xmn, xmx, ymn, ymx, dxdy, pts, kwidth) {
   
   if(length(dxdy) == 1) dxdy <- rep(dxdy, 2)
   adj <- dxdy / 2
   
-  d1 <- length(seq(xmn[1], xmx[2], dxdy[1]))
-  d2 <- length(seq(ymn[1], ymx[2], dxdy[2]))
+  d1 <- length(seq(xmn, xmx, dxdy[1]))
+  d2 <- length(seq(ymn, ymx, dxdy[2]))
   
   # grid topology
   grd <- GridTopology(cellcentre.offset = c(xmn + adj[1], ymn + adj[2]), 
@@ -28,7 +29,7 @@ kdensity <- function(xmn, xmx, ymn, ymx, dxdy, pts) {
   # kernel densities
   ppoly <- as(extent(rst), "SpatialPolygons")
   crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
-  kern <- spkernel2d(resdat, crds, h0 = 1, grd)
+  kern <- spkernel2d(pts, crds, h0 = kwidth, grd)
   kgrid <- as(SpatialGridDataFrame(grd, data = data.frame(kern)),
               "SpatialPixelsDataFrame")
 

### 2016-04-04 Rmd/calibration.Rmd
Updated calibration.Rmd to include analysis of SD/CV of observations by DOI and feature

@@
 p_dat <- fp(p_root, "external/data/calibration")
+```
 
+```{r, eval=FALSE}
 # Read in and clean-up calibration data
-# cal <- data.table(read_excel(fp(p_dat, "merged_calibration.xlsx")))
-cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
+cal <- data.table(read_excel(fp(p_dat, "merged_calibration.xlsx")))
 cal[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
 cal[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
-dropnm <-  c("V30", "V31", "journal", "exclusion reason", "sensitivity")
+names(cal)
+dropnm <-  c("V29")#, "V31", "journal", "exclusion reason", "sensitivity")
 cal[, c(dropnm) := NULL]
@@
 # write.csv(cal, file = fp(p_dat, "merged_calibrationf.csv"))
+```
 
+```{r}
 # external fixes and recodes, bring back in
-cal <- data.table(read_excel(fp(p_dat, "merged_calibrationf.xlsx")))
-cal[, notes := NULL]
-for(j in names(cal)[c(4:5, 8:24)]) set(cal, j = j, value = as.numeric(cal[[j]]))
+# cal <- data.table(read_excel(fp(p_dat, "merged_calibrationf.xlsx")))
+cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
+names(cal)
+cal[, c("notes", "V29") := NULL]
+for(j in names(cal)[c(4:5, 10:24)]) {
+  set(cal, j = j, value = as.numeric(cal[[j]]))
+}
 # length(which(is.na(cal$feature)))
@@
 plot(c(1, 20), c(0, 10), pch = "", xlab = "Paper reviewed", 
-     ylab = "N extractable records", mgp = c(1.45, 0.7, -0.2))
+     ylab = "N extractable records", mgp = c(1.45, 0.5, 0), tcl = -0.2)
 polygon(x = c(0, 21, 21, 0, 0), y = c(-1, -1, 11, 11, 0), col = "grey70")
@@
 ## Variance in extracted statistics
+### By DOI
 ```{r}
@@
             "study_duration")
-mus <- cal[inout == 1, lapply(.SD, mean, na.rm = TRUE), by = .(DOI, observer),             .SDcols = varnms]
+mus <- cal[inout == 1, lapply(.SD, mean, na.rm = TRUE), by = .(DOI, observer),             .SDcols = varnms]  # mean variable val per paper per observer
+# variance and mean per paper across observers
 vars <- mus[, lapply(.SD, sd, na.rm = TRUE), by = DOI, .SDcols = varnms] 
@@
 
+# calr[DOI == "10.1007/s00442-005-0338-3" & feature == 2]
 ```
 
-```{r}
-par(mfcol = c(10, 4), mar = c(2, 2, 2, 2), mgp = c(1.5, 0.5, 0))
+### By feature (individual records) across DOIs
+```{r, results='hold'}
+CV <- function(x) sd(x) / mean(x)
+rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", "t_btwn_samp",
+          "study_duration")
+calr <- cal[, rnms, with = FALSE]
+# calr[feature != 99, ]
+# calr[feature != 99, lapply(.SD, mean), by = .(DOI, feature), 
+#      .SDcols = rnms[4:7]]
+# calr[feature != 99, .N, by = .(DOI, feature)]
+calr[feature != 99, lapply(.SD, CV), by = .(DOI, feature), 
+     .SDcols = rnms[4:7]][, lapply(.SD, mean, na.rm = TRUE), 
+                          .SDcols = rnms[4:7]]
 
-mus[, plot(plot_res), by = DOI]
+# par(mfcol = c(10, 4), mar = c(2, 2, 2, 2), mgp = c(1.5, 0.5, 0))
+# mus[, plot(plot_res), by = DOI]
 
@@
 
-Tim needs to fix paleo studies in calibration (and main set) giving too much credit to effective area.
+Tim needs to fix paleo studies in calibration (and main set) giving too much credit to effective area. [DONE]
 
-Others need to fix their study duration figures. Paul on paleo one in calibration, Choi, Chang, etc in main set.
+Others need to fix their study duration figures. Paul on paleo one in calibration, Choi, Chang, etc in main set.  [Ahmed remaining]
 
-For N-sites, count sites only once if there are repeated measures taken at them, e.g. Jonathan in 10.1111/j.1466-8238.2009.00454.x for bird counts
+For N-sites, count sites only once if there are repeated measures taken at them, e.g. Jonathan in 10.1111/j.1466-8238.2009.00454.x for bird counts [Chang, Ahmed remaining]

### 2016-04-05 Rmd/calibration.Rmd
changes to space-time.Rmd: 1. rules and code for perturbing uncertain variables replaced in favor of uncertainty analysis on 4 main variables for all records. 2. mean features values for calibration data merged with main dataset 3. 2 resampling/bootstrap variants run and plotted: 1.  perturbation of 4 main variables; 2. bootstrapping with replacement. 4. Main figure made on merged calibration/main dataset made. 5. Additional analyses appears towards end: a) code for calculating percent of records within different portions of space-time domain (still to be made); b) assessment of percentage of

@@
 
-```{r}
+```{r, warning = FALSE}
 # external fixes and recodes, bring back in
@@
 CV <- function(x) sd(x) / mean(x)
-rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", "t_btwn_samp",
-          "study_duration")
+rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", 
+          "sampled_area", "t_btwn_samp", "study_duration")
 calr <- cal[, rnms, with = FALSE]
@@
 calr[feature != 99, lapply(.SD, CV), by = .(DOI, feature), 
-     .SDcols = rnms[4:7]][, lapply(.SD, mean, na.rm = TRUE), 
-                          .SDcols = rnms[4:7]]
+     .SDcols = rnms[4:8]][, lapply(.SD, mean, na.rm = TRUE), 
+                          .SDcols = rnms[4:8]]
 

### 2016-04-05 Rmd/space-time.Rmd
changes to space-time.Rmd: 1. rules and code for perturbing uncertain variables replaced in favor of uncertainty analysis on 4 main variables for all records. 2. mean features values for calibration data merged with main dataset 3. 2 resampling/bootstrap variants run and plotted: 1.  perturbation of 4 main variables; 2. bootstrapping with replacement. 4. Main figure made on merged calibration/main dataset made. 5. Additional analyses appears towards end: a) code for calculating percent of records within different portions of space-time domain (still to be made); b) assessment of percentage of

--- 2015-12-19 Rmd/space-time.Rmd
@@
 ```{r, eval = FALSE}
-library(data.table)
-library(raster)
-library(sp)
-require(splancs)
-library(lmisc)
+library(ecoscales)
+library(readxl)
+library(stringdist)
+library(maptools)
+
 p_root <- set_base_path("ecoscales")
-p_dat <- full_path(p_root, "external/data")
-dat <- fread(full_path(p_dat, "merged-data.csv"), stringsAsFactors = FALSE)
+p_dat <- full_path(p_root, "external/data/result")
+p_calib <- full_path(p_root, "external/data/calibration")
+# dat <- as.data.table(read_excel(full_path(p_dat, "merged.xlsx")))
+dat <- as.data.table(read_excel(full_path(p_dat, "mergedf.xlsx")))
+cal <- data.table(read_excel(fp(p_calib, "merged_calibrationff.xlsx")))
+
 setnames(dat, "DOI/title", "DOI")
+dat[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
+dat[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
+
 dat <- dat[, names(dat)[1:20], with = FALSE]
-unique(dat$DOI)  # 139
+unique(dat$DOI)  # 138
 ```
 
-## Spatal and temporal data
+## Data prep
 
+### Merge full and calibration datasets
+```{r, eval = FALSE}
+dat2 <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
+                     t_btwn_samp, study_duration)])
+
+rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", 
+          "sampled_area", "t_btwn_samp", "study_duration")
+calr <- cal[, rnms, with = FALSE]
+for(j in names(calr)[c(2, 4:8)]) set(calr, j=j, value = as.numeric(calr[[j]]))
+
+# select down to columns of interest
+calr <- calr[feature != 99, lapply(.SD, mean), by = .(DOI, feature), 
+             .SDcols = rnms[4:8]]
+length(unique(calr$DOI)) + length(unique(dat$DOI))  # 148 papers providing dat
+
+calr[, samp_duration := NA]
+calr[, c("DOI", "feature") := NULL]
+setcolorder(calr, names(calr)[c(1:3, 6, 4:5)])
+
+datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
+```
+
+### Adjustments/fixes
+1.
+if `samp_duration != study_duration` and `t_btwn_samp == 0`, set `study_duration == samp_duration` (actually, no, in many cases there is a good reason to have to `study_duration != samp_duration` (see below))
+2.
+Prior to scaling the data in log10 space, we need to make one adjustment.
+Time between samples has many 0 values because many observations are simply one-offs.
+Set these to an arbitrarily large value which will represent clear separation on axis from high frequency studies.
+A function is defined here for this purpose, and will be applied after bootstrapping is done.
+
+```{r, eval=FALSE}
+# apply fixes (mostly to account for observer omissions)
+# 1. if samp_duration != study_duration and t_btwn_samp == 0, set 
+#    study_duration == samp_duration
+names(datf)
+datf[t_btwn_samp == 0 & (study_duration != samp_duration), ]  
+dat[t_btwn_samp == 0 & (study_duration != samp_duration), c(1, 3, 8:13), with=F]  
+# function to set t_btwn_samp to 365 * 100000 to indicate once-off studies 
+# when t_btwn_samp == 0.  
+to_infin_byond <- function(tbtwn) ifelse(tbtwn == 0, 365 * 10000, tbtwn)
+
+```
+
+### Bootstraps
+
+### With uncertainties
+
+Defined by the per variable CV based on uncertainty between observers.
+Here we are just going to use apply the variability to each variable for every observation (previous incarnation used uncertainty on just the variables listed as uncertain, with recalculation of dependent variables (`study_duration`, `sampled_area`) made after pertubation [see commit prior to second or third commit on 4 April for code]).
+
+```{r, eval=FALSE}
+p <- c(0.79, 0.99, 0.53, 1.22)
+
+# Function for bootstrapping with uncertainties
+resamp_func <- function(dat, p, iter) {
+  bout <- lapply(1:iter, function(x) {
+    if((x / 100) %in% 1:100) print(x)
+    if(length(p) < 4) stop("p must be length of 4", call. = FALSE)
+
+    dnew <- copy(datf[, .(plot_res, sampled_area, t_btwn_samp, study_duration)])
+    # modifiers
+    modvec <- sapply(p, function(y) {
+      mv <- runif(nrow(dnew), min = 1 - y, max = 1 + y)
+      mv
+    })
+
+    for(j in 1:ncol(modvec)) set(dnew, j = j, value = dnew[[j]] * modvec[, j])
+    dnew[plot_res <= 0, plot_res := 0.00000001]
+    dnew[sampled_area <= 0, sampled_area := 0.00000001]
+    dnew[study_duration <= 0, study_duration := 0.00000001]
+    dnew[t_btwn_samp < 0, t_btwn_samp := 0.00000001]
+
+    dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
+    dnew
+  })
+  return(bout)
+}
+
+# old function, retained here for easy recollection of first approach
+# full function for resampling
+# resamp_func <- function(dat, p, iter) {
+#   bout <- lapply(1:iter, function(x) {  # x <- 1
+#     if((x / 100) %in% 1:100) print(x)
+#     # amount to perturb
+#     a <- runif(1, min = 1 - p, max = 1 + p)  # plot_res perturb
+#     b <- runif(1, min = 1 - p, max = 1 + p)  # n_sites perturb
+#     d <- runif(1, min = 1 - p, max = 1 + p)  # samp_duration perturb
+#     e <- runif(1, min = 1 - p, max = 1 + p)  # t_btwn_samp perturb
+#     
+#     if(length(p) < 4) stop("p must be length of 4", call. = FALSE)
+# 
+#     
+#     # spatial variables
+#     dnew <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
+#                          t_btwn_samp, study_duration)])
+#     dnew[which(senst$res == 1), plot_res := plot_res * a]
+#     dnew[which(senst$n == 1), n_sites := n_sites * b]
+#     dnew[, sampled_area := (plot_res * n_sites) / 10000]
+# 
+#     # temporal variables
+#     dnew[which(senst$sampd == 1), samp_duration := samp_duration * d]
+#     dnew[which(senst$tbtwn == 1), t_btwn_samp := t_btwn_samp * e]
+#     dnew[, study_duration := {
+#       recalc_duration(study_duration, samp_duration, t_btwn_samp)
+#     }]
+#     dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
+#   })
+#   return(bout)
+# }
+
+# perturbed samples
+# p10 <- resamp_func(dat, 0.1, 1000)  # 10% perturb
+
+set.seed(1)
+bootperturb <- resamp_func(datf, p = p, iter = 1000)  # 10% perturb
+# bootperturb2 <- resamp_func(datf, p = p, iter = 1000)  # 10% perturb
+# all(bootperturb[[200]]$plot_res == bootperturb2[[200]]$plot_res)
+# rm(bootperturb2)
+
+```
+
+### Without uncertainties
+
+Good old-fashioned bootstrapping with replacement, no perturbation of value
+```{r, eval = FALSE}
+# bootstrap
+set.seed(1)
+bootsamp <- lapply(1:1000, function(x) {
+  if((x / 100) %in% 1:100) print(x)
+  dnew <- copy(datf[, .(plot_res, sampled_area, t_btwn_samp, study_duration)])
+  dnew <- dnew[sample(1:.N, size = .N, replace = TRUE), ]
+  dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
+  dnew
+})
+
+```
+
+## Plots
 ### Scaling in log10 space
-```{r}
-i <- 0.000001
-j <- rep(0, 15)
+```{r, eval = FALSE}
+i <- 0.0000001
+j <- rep(0, 16)
 for(k in 1:length(j)) {
@@
 
-# label axis for time
-taxis <- cbind.data.frame(
-  c("Sec", "Min", "Hour", "Day", "Month", "Year", "Decade", "Century", 
-    "Millenium", "10000"),  
-  c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 30, 365, 365 * 10, 
+# labels for temporal axes
+# return interval
+tlab1 <- c(expression("second"^-1), expression("minute"^-1),
+           expression("hour"^-1), expression("day"^-1), 
+           expression("week"^-1), expression("month"^-1),            
+           expression("year"^-1), expression("decade"^-1),
+           expression("century"^-1), 
+           expression("millenium"^-1), "once-off")# expression(infinity))
+taxis1 <- cbind.data.frame( 
+  c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 30, 365, 365 * 10, 
     365 * 100, 365 * 1000, 365 * 10000))
-taxis <- data.table(taxis)
-setnames(taxis, names(taxis), c("label", "days"))
-taxis[, logdays := log10(days)]
+taxis1 <- data.table(taxis1)
+setnames(taxis1, names(taxis1), "days")
+taxis1[, logdays := log10(days)]
 
-rdt <- data.table("scaleval" = j, "area" = j^2)
-rdt[, slog := log10(scaleval)]
-rdt[, rlog := log10(area)]
+taxis2 <- cbind.data.frame(
+  c("second", "minute", "hour", "day", "week", "month", "year", "decade",
+    "century", "millenium", "10 KA"),  
+  c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 30, 365, 365 * 10, 
+    365 * 100, 365 * 1000, 365 * 10000))
+taxis2 <- data.table(taxis2)
+setnames(taxis2, names(taxis2), c("label", "days"))
+taxis2[, logdays := log10(days)]
 
-adt <- data.table("area" = j)
-adt[, alog := log10(area)]
+# spatial
+# plot(1:10, xlab = expression(paste("mm"^2)), xaxt = "n")
+# axis(1, at = 1:8, labels = lab, las = 2)
+
+# plot resolution
+alab1 <- c(expression("0.01 cm"^2), expression("0.1 cm"^2), 
+           expression("1 cm"^2), expression("10 cm"^2),
+           expression("100 cm"^2), expression("1000 cm"^2), 
+           expression("1 m"^2), expression("10 m"^2),
+           expression("100 m"^2), expression("1000 m"^2), 
+           "1 ha", "10 ha", "100 ha", "1000 ha", "10000 ha")
+
+aaxis1 <- cbind.data.frame(c(0.000001, 0.00001, 0.0001, 0.001, 0.01, 0.1,
+                             1^2, 10, 10^2, 10^3, 10^4, 10^5, 10^6, 10^7, 10^8))
+aaxis1 <- data.table(aaxis1)
+setnames(aaxis1, names(aaxis1), "res")
+aaxis1[, logres := log10(res)]
+
+# sampled area
+alab2 <- c(expression("1 m"^2), expression("10 m"^2), expression("100 m"^2),
+           expression("1000 m"^2), expression("1 ha"), expression("10 ha"),
+           expression("100 ha"), expression("1000 ha"), 
+           expression(paste(10^4 , " ha")), expression(paste(10^5 , " ha")), 
+           expression(paste(10^6 , " ha")), expression(paste(10^7 , " ha")), 
+           expression(paste(10^8 , " ha")), expression(paste(10^9 , " ha")), 
+           expression(paste(10^10 , " ha")))
+
+aaxis2 <- cbind.data.frame(
+  c(0.1 / 10000, 10 / 10000, 1000 / 10000, 1, 10, 10^2, 10^3, 10^4, 10^5, 10^6,
+    10^7, 10^8, 10^9, 10^10))
+aaxis2 <- data.table(aaxis2)
+setnames(aaxis2, names(aaxis2), "area")
+aaxis2[, logarea := log10(area)]
 
@@
 
-### Resolution versus extent, gridded
-```{r,eval = FALSE}
-rres <- 0.25
-# # raster for resolution (x dimension) versus area (y dimension)
-# # arst <- raster(xmn = -6, xmx = 9, ymn = -5, ymx = 10)
-# rst <- raster(xmn = -5, xmx = 10, ymn = -5, ymx = 10)
-# res(rst) <- rres
-# # rst@crs <- CRS("")
-# rst[] <- 1
-# rst2 <- raster(xmn = 0, xmx = 10, ymn = 0, ymx = 10)
-# # res(rst2) <- c(0.25, 0.25)
-# res(rst2) <- rres
-# # rst@crs <- CRS("")
-# rst2[] <- 1
-# plot(rst2)
-
-resdat <- data.table(x = log10(dat$t_btwn_samp + 0.00001), 
-                     y = log10(dat$plot_res))
-# resdat[x < -5, x := -4.5]
-# resdat[y < -5, y := -4.5]
+### Perturbed boot-strap
+```{r, eval = FALSE}
+bootp_dt <- rbindlist(bootperturb)
+resdat <- data.table(x = log10(bootp_dt$t_btwn_samp), 
+                      y = log10(bootp_dt$plot_res))
 resdat$n <- rep(1, nrow(resdat))
+rx <- range(resdat$x)
+rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
+# rx[1] <- -5
+rx[2] <- 6.56
+resdat[x < rx[1], x := rx[1]]
+resdat[x > rx[2], x := rx[2]]
+# rx[2] <- 7.6
+ry <- range(aaxis1$logres)
+ry[1] <- -5  # set limits to 0.01 cm^2
+resdat[y < ry[1], y := ry[1]]
+resdat[y > ry[2], y := ry[2]]
 coordinates(resdat) <- ~x + y
-# plot(resdat$x, resdat$y)
-rx <- range(resdat$x)
-# rx <- c(floor(rx[1]), c(ceiling(rx[2])))
-rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-rxd <- length(seq(rx[1], rx[2], rres))  # 0.25))
-ry <- range(resdat$y)
-# ry <- c(floor(ry[1]), c(ceiling(ry[2])))
-ry <- round(ry / 0.5) * 0.5  # round to nearest 0.5
-ryd <- length(seq(ry[1], ry[2], rres))#0.25))
-
-# length(seq(-5, 4, 0.25))
-
-extdat <- data.table("x" = log10(dat$study_duration + 0.00001), 
-                     "y" = log10(dat$sampled_area + 0.00001)) 
-# extdat <- data.table("x" = log10(dat$study_span), 
-#                      "y" = log10(dat$sampled_area)) 
-# extdat[x < 0, x := 0]
-# extdat[y < 0, y := 0]  # setting minima to 1 day and 1 ha
-extdat$n <- rep(1, nrow(extdat))
-coordinates(extdat) <- ~x + y
-# plot(extdat$x, extdat$y)
-ex <- range(extdat$x)
-# ex <- c(floor(ex[1]), c(ceiling(ex[2])))
-ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
-ex[1] <- -5
-exd <- length(seq(ex[1], ex[2], rres))#0.25))
-ey <- range(extdat$y)
-# ey <- c(floor(ey[1]), c(ceiling(ey[2])))
-ey <- round(ey / 0.5) * 0.5  # round to nearest 0.5
-eyd <- length(seq(ey[1], ey[2], rres))#0.25))
-
-# set-up grid topologies
-rgrd <- GridTopology(cellcentre.offset = c(rx[1], ry[1]) + 0.125, 
-                     #cellsize = c(0.25, 0.25), cells.dim = c(rxd, ryd))
-                     cellsize = c(rres, rres), cells.dim = c(rxd, ryd))
-egrd <- GridTopology(cellcentre.offset = c(ex[1], ey[1]) + 0.125, 
-                     #cellsize = c(0.25, 0.25), cells.dim = c(exd, eyd))
-                     cellsize = c(rres, rres), cells.dim = c(exd, eyd))
-# seq(0, 10, 10 / 59)
-
-rst <- raster(xmn = rx[1], xmx = rx[2], ymn = ry[1], ymx = ry[2])
-# rst <- raster(xmn = min(resdat$x), xmx = ceiling(max(resdat$x)), 
-#               ymn = min(resdat$y), ymx = ceiling(max(resdat$y)))
-res(rst) <- rres
-# rst@crs <- CRS("")
-rst[] <- 1
-# plot(rst)
-rst2 <- raster(xmn = ex[1], xmx = ex[2], ymn = ey[1], ymx = ey[2])
-# rst2 <- raster(xmn = min(extdat$x), xmx = ceiling(max(extdat$x)), 
-#               ymn = min(extdat$y), ymx = ceiling(max(extdat$y)))
-res(rst2) <- rres
-# rst@crs <- CRS("")
-rst2[] <- 1
-# plot(rst2)
-
-# kernel densities
-# resolution
-ppoly <- as(extent(rst), "SpatialPolygons")
-crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
-kern1 <- spkernel2d(resdat, crds, h0 = 1, rgrd)
-kgrid1 <- as(SpatialGridDataFrame(rgrd, data = data.frame(kern1)),
-             "SpatialPixelsDataFrame")
 
 # extent
-ppoly <- as(extent(rst2), "SpatialPolygons")
-crds <- slot(slot(slot(ppoly, "polygons")[[1]], "Polygons")[[1]], "coords") 
-kern2 <- spkernel2d(extdat, crds, h0 = 1, egrd)
-kgrid2 <- SpatialGridDataFrame(egrd, data = data.frame(kern2))
+extdat <- data.table(x = log10(bootp_dt$study_duration), 
+                     y = log10(bootp_dt$sampled_area))
+extdat$n <- rep(1, nrow(extdat))
+ex <- range(taxis2$logdays)
+ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
+extdat[x < ex[1], x := ex[1]]
+extdat[x > ex[2], x := ex[2]]
+# ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
+# ex[2] <- 7.6
+ey <- range(aaxis2$logarea)
+extdat[y < ey[1], y := ey[1]]
+extdat[y > ey[2], y := ey[2]]
+coordinates(extdat) <- ~x + y
 
-# rasterize
-# space
-rr <- raster(kgrid1)
-rr[is.na(rr)] <- 0
-# time
-er <- raster(kgrid2)
-er[is.na(er)] <- 0
+# densities
+# resolution
+rres <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat, 1)
+rres <- (rres / cellStats(rres, sum)) * 100
+rng <- range(rres[is.finite(rres)])
+rng[2] <- 0.1
+bwidth <- (rng[2] - rng[1]) / 30
+brks <- seq(rng[1], rng[2], bwidth)
+brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
 
-# have a look
-par(mfrow = c(2, 1))
-plot(er, axes = FALSE, add = FALSE)
-arg <- list(at = taxis$logdays, labels = taxis$label)
-image(er, col = rev(terrain.colors(100)), axes = FALSE, xlab = "", ylab = "", 
-      xlim = c(-5, 6))
-axis(1, at = taxis$logdays, labels = taxis$label, las = 2)
-axis(2, at = -5:10, las = 2)
+# extent
+rext <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat, 1)
+rext <- (rext / cellStats(rext, sum)) * 100
+rng <- range(rext[is.finite(rext)])
+rng[2] <- 0.1
+bwidth <- (rng[2] - rng[1]) / 30
+brks2 <- seq(rng[1], rng[2], bwidth)
+brklabs2 <- round(brks2[seq(1, length(brks2), 6)], 2)
 
-# points(extdat$x, extdat$y)
-plot(rr)
-# points(resdat$x, resdat$y)
+# cols <- colorRampPalette(c("blue", "gold", "red"))
+# cols <- c(rev(terrain.colors(30))[1], cols(29))
 
-ers <- er
-rrs <- rr
+cxa = 0.7
+pdf("paper/figures/res_v_extent_pbstrap.pdf", width = 7, 
+    height = 2.5)
+par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
+    mgp = c(2, 0.5, 0))
+# image(rres, col = cols, breaks = brks,
+image(rres, col = rev(terrain.colors(30)), breaks = brks,
+     axes = FALSE, xlab = "", ylab = "")
+# plot(rres2, col = rev(terrain.colors(20)))
+axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
+     las = 2, tcl = -0.2, cex.axis = cxa)
+axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
+     cex.axis = cxa)
+# plot(rres2, axes = FALSE, legend.only = TRUE)
+aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
+              cex.axis = cxa, tcl = -0.1)
+plot(rres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
+image(rext, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
+      ylab = "", breaks = brks2)
+axis(1, at = taxis2$logdays[-5], labels = taxis2$label[-5], las = 2,
+     cex.axis = cxa, tcl = -0.2)
+axis(2, at = aaxis2$logarea, labels = alab2[-2], las = 2, cex.axis = cxa,
+     tcl = -0.2)
+aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
+              cex.axis = cxa, tcl = -0.1)
+plot(rext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
+dev.off()
+
 ```
 
-### Convert to rgb space and plot
+### Ordinary bootstrap
 ```{r, eval = FALSE}
+bootdt <- rbindlist(bootsamp)
+resdat2 <- data.table(x = log10(bootdt$t_btwn_samp), y = log10(bootdt$plot_res))
+resdat2$n <- rep(1, nrow(resdat2))
+rx <- range(resdat2$x)
+rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
+rx[2] <- 6.56
+resdat2[x < rx[1], x := rx[1]]
+resdat2[x > rx[2], x := rx[2]]
+ry <- range(aaxis1$logres)
+ry[1] <- -5  # set limits to 0.01 cm^2
+resdat2[y < ry[1], y := ry[1]]
+resdat2[y > ry[2], y := ry[2]]
+coordinates(resdat2) <- ~x + y
+
+# extent
+extdat2 <- data.table(x = log10(bootdt$study_duration), 
+                      y = log10(bootdt$sampled_area))
+extdat2$n <- rep(1, nrow(extdat2))
+ex <- range(taxis2$logdays)
+ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
+extdat2[x < ex[1], x := ex[1]]
+extdat2[x > ex[2], x := ex[2]]
+ey <- range(aaxis2$logarea)
+extdat2[y < ey[1], y := ey[1]]
+extdat2[y > ey[2], y := ey[2]]
+coordinates(extdat2) <- ~x + y
+
+# densities
+# resolution
+rres2 <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat2, 1)
+rres2 <- (rres2 / cellStats(rres2, sum)) * 100
+rng <- range(rres2[is.finite(rres2)])
+bwidth <- (rng[2] - rng[1]) / 30
+brks <- seq(rng[1], rng[2], bwidth)
+brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
+
+# extent
+rext2 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat2, 1)
+rext2 <- (rext2 / cellStats(rext2, sum)) * 100
+rng <- range(rext2[is.finite(rext2)])
+rng[2] <- 0.1
+bwidth <- (rng[2] - rng[1]) / 30
+brks2 <- seq(rng[1], rng[2], bwidth)
+brklabs2 <- brklabs #round(brks2[seq(1, length(brks2), 3)], 2)
+
+cxa = 0.7
+pdf("paper/figures/res_v_extent_bstrap.pdf", width = 7, 
+    height = 2.5)
+par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
+    mgp = c(2, 0.5, 0))
+# image(rres, col = cols, breaks = brks,
+image(rres2, col = rev(terrain.colors(30)), breaks = brks,
+     axes = FALSE, xlab = "", ylab = "")
+# plot(rres2, col = rev(terrain.colors(20)))
+axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
+     las = 2, tcl = -0.2, cex.axis = cxa)
+axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
+     cex.axis = cxa)
+# plot(rres2, axes = FALSE, legend.only = TRUE)
+aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
+              cex.axis = cxa, tcl = -0.1)
+plot(rres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
+
+image(rext2, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
+      ylab = "", breaks = brks2)
+axis(1, at = taxis2$logdays[-5], labels = taxis2$label[-5], las = 2,
+     cex.axis = cxa, tcl = -0.2)
+axis(2, at = aaxis2$logarea, labels = alab2[-2], las = 2, cex.axis = cxa,
+     tcl = -0.2)
+aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
+              cex.axis = cxa, tcl = -0.1)
+plot(rext2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
+dev.off()
+
+# datf[, plot(log10(study_duration), log10(sampled_area), xlim)]
+
+```
+
+### On primary dataset only
+```{r, eval = FALSE}
+resdat3 <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
+                      y = log10(datf$plot_res))
+resdat3$n <- rep(1, nrow(resdat3))
+rx <- range(resdat3$x)
+rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
+rx[2] <- 6.56
+resdat3[x < rx[1], x := rx[1]]
+resdat3[x > rx[2], x := rx[2]]
+ry <- range(aaxis1$logres)
+ry[1] <- -5  # set limits to 0.01 cm^2
+resdat3[y < ry[1], y := ry[1]]
+resdat3[y > ry[2], y := ry[2]]
+coordinates(resdat3) <- ~x + y
+
+# extent
+extdat3 <- data.table(x = log10(datf$study_duration), 
+                      y = log10(datf$sampled_area))
+extdat3$n <- rep(1, nrow(extdat3))
+ex <- range(taxis2$logdays)
+ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
+extdat3[x < ex[1], x := ex[1]]
+extdat3[x > ex[2], x := ex[2]]
+ey <- range(aaxis2$logarea)
+extdat3[y < ey[1], y := ey[1]]
+extdat3[y > ey[2], y := ey[2]]
+coordinates(extdat3) <- ~x + y
+
+# densities
+# resolution
+rres3 <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat3, 1)
+rres3 <- (rres3 / cellStats(rres3, sum)) * 100
+rng <- range(rres3[is.finite(rres3)])
+bwidth <- (rng[2] - rng[1]) / 30
+brks <- seq(rng[1], rng[2], bwidth)
+brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
+
+# extent
+rext3 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat3, 1)
+rext3 <- (rext3 / cellStats(rext3, sum)) * 100
+rng <- range(rext3[is.finite(rext3)])
+rng[2] <- 0.1
+bwidth <- (rng[2] - rng[1]) / 30
+brks2 <- seq(rng[1], rng[2], bwidth)
+brklabs2 <- brklabs #round(brks2[seq(1, length(brks2), 3)], 2)
+
+cxa = 0.7
+pdf("paper/figures/res_v_extent.pdf", width = 7, 
+    height = 2.5)
+par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
+    mgp = c(2, 0.5, 0))
+image(rres3, col = rev(terrain.colors(30)), breaks = brks,
+     axes = FALSE, xlab = "", ylab = "")
+axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
+     las = 2, tcl = -0.2, cex.axis = cxa)
+axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
+     cex.axis = cxa)
+
+aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
+              cex.axis = cxa, tcl = -0.1)
+plot(rres3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
+
+image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
+      ylab = "", breaks = brks2)
+axis(1, at = taxis2$logdays[-5], labels = taxis2$label[-5], las = 2,
+     cex.axis = cxa, tcl = -0.2)
+axis(2, at = aaxis2$logarea, labels = alab2[-2], las = 2, cex.axis = cxa,
+     tcl = -0.2)
+aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
+              cex.axis = cxa, tcl = -0.1)
+plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
+dev.off()
+
+# datf[, plot(log10(study_duration), log10(sampled_area), xlim)]
+
+```
+
+## Ancillary/supplementary plots and stats
+
+### Statistics
+```{r, eval=FALSE}
+length(which(dat$study_type %in% 
+               c("remote sensing", "other geographic data"))) / nrow(dat)
+dat[t_btwn_samp == 0, .N] / dat[, .N]  # 31% of samples are once-offs
+dat[study_duration == samp_duration, .N]
+
+par(mfrow = c(2, 2))
+datf[, {
+  hist(log10(plot_res), main = "Plot res")
+  hist(log10(t_btwn_samp), main = "Interval")
+  hist(log10(sampled_area), main = "Area")
+  hist(log10(study_duration), main = "Duration")
+}]
+
+```
+
+### Sensitivity
+Just for the four main variables. Rules:
+
++ If `plot_res` and/or `n_sites` uncertain, then `sampled_area` must be. Recalculate using random draw within the uncertain variable 
++ If `t_btwn_samp` and/or `samp_duration` uncertain, do the same as above, but we have no variable for n repeats for calculating `study_duration`.
+
+```{r, eval = FALSE}
+svars <- c("plot_res", "n_sites", "sampled_area", "samp_duration",
+           "t_btwn_samp", "study_duration")
+
+sens <- gsub("\\\\", ";", dat$sensitivity)
+sens <- gsub(", ", ";", gsub("; ", ";", sens))
+sens <- tolower(sens)
+sensl <- strsplit(sens, ";")
+unique(unlist(sensl))
+
+sens2 <- gsub("plo_", "plot_", sens)  # fix this one
+sens2 <- gsub("[[:blank:]]", "_", sens2)  # replace spaces with _
+
+corrtab <- cbind(c("^res", "plo_res", "sample_area", "samples_area", 
+                   "sampled area",
+                   "t_btwn_samples", "t_btwn_sample", "t_btwn_samp_samp", 
+                   "t_twn_samp", "t_btw_samp", "samp_duratiion", 
+                   "sampling_duration", "sample_duration", 
+                   "plot_resolution", "number_of_sites", 
+                   "time_between_sampling", "plot_size", 
+                   "time_between_samples", "samp_study", ":", " ", ";_"), 
+                 c(svars[c(1, 1, 3, 3, 3, 5, 5, 5, 5, 5, 4, 4, 4, 1, 2,
+                           5, 1, 5)], "samp;study", ";", ";", ";"))
+
+for(i in 1:nrow(corrtab)) {
+  sens2 <- gsub(corrtab[i, 1], corrtab[i, 2], sens2)
+}
+# gsub("[a-z]( )[a-z]", "_", sens[[228]])
+# sens[[228]]
+# sens2[is.na(sens2)]
+sens2[is.na(sens2)] <- "none"
+sensl2 <- lapply(1:length(sens2), function(x) strsplit(sens2[x], ";")[[1]])
+# unique(unlist(sensl))
+unique(unlist(sensl2))
+
+# create a sensitivity table
+senst <- do.call(cbind.data.frame, lapply(svars, function(x) {
+  v <- sapply(sensl2, function(y) ifelse(any(y == x), 1, 0))
+  # ifelse(is.na(v), 0, v)
+}))
+colnames(senst) <- c("res", "n", "sarea", "sampd", "tbtwn", "studyd")
+
+# add column for uncertainty about n repeats
+# senst$nrep <- rep(0, nrow(senst))
+
+# i <- 300:305
+# sensl2[i]
+# senst[i, ]
+
+# apply fixes (mostly to account for observer omissions)
+
+# 1. if samp_duration OR t_btwn_samp is uncertain, study_duration must also be. 
+id <- which((senst$sampd == 1 | senst$tbtwn == 1) & senst$studyd == 0)
+senst[id, "studyd"] <- 1
+
+# 2. if study_duration is uncertain but t_btwn and samp_duration are not, treat
+# both as uncertain
+# although it is possible that just n_repeats is uncertain, but more likely 
+# that observers didn't note this is sensitive. 
+id <- which(senst$tbtwn == 0 & senst$sampd == 0 & senst$studyd == 1)
+senst[id, c("sampd", "tbtwn")] <- 1
+
+# 3. if plot_res or n_sites is uncertain, then sampled_area must be
+id <- which((senst$res == 1 | senst$n == 1) & senst$sarea == 0)
+senst[id, "sarea"] <- 1
+
+# percentage of records that are uncertain
+colSums(senst) / nrow(senst) * 100
+
+```
+
+### Varying kernel sizes on resolution versus extent
+```{r,eval = FALSE}
+# resolution data
+# resdat <- data.table(x = log10(datf$t_btwn_samp), y = log10(datf$plot_res))
+# resdat$n <- rep(1, nrow(resdat))
+# coordinates(resdat) <- ~x + y
+# rx <- range(resdat$x)
+# # rx <- range(aaxis$logarea)
+# rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
+# # rx[1] <- -6.5  # set temporal interval to -6.5 to have null returns centered
+# # ry <- range(resdat$y)
+# ry <- range(aaxis$logarea)
+# ry <- round(ry / 0.5) * 0.5  # round to nearest 0.5
+# 
+# # extent data
+# # fix area rounded to zero in Jon's plots
+# extdat <- data.table("x" = log10(dat$study_duration), 
+#                      "y" = log10(dat$sampled_area)) 
+# # extdat[is.infinite(x), ]; extdat[is.infinite(y), ]
+# 
+# extdat$n <- rep(1, nrow(extdat))
+# coordinates(extdat) <- ~x + y
+# # ex <- range(extdat$x)
+# ex <- range(taxis2$logdays)
+# ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
+# # ex[1] <- -5
+# ey <- range(extdat$y)
+# ey <- round(ey / 0.5) * 0.5  # round to nearest 0.5
+
+# density rasters
+# vary the kernel size about them
+kernsl <- lapply(c(0.4, 0.6, 0.8, 1), function(x) {
+  rr <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat3, x)
+  er <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat3, x)
+  kerns <- list("res" = rr, "ext" = er)
+})
+
 # As separate plots
-png("paper/figures/res&extent2.png", width = 900, height = 450)
-par(mfrow = c(1, 2), oma = c(0, 0, 0, 0), mar = c(4, 4, 4, 4))
-plot(rrs, xlim = c(-5, 15), ylim = c(-5, 15), 
-     xlab = "Sample interval - log10(days)", 
-     ylab = "Plot resolution - log10(m^2)")
-plot(ers, box = FALSE, xlim = c(-5, 15), ylim = c(-5, 15), 
-     xlab = "Study duration - log10(days)", ylab = "Sampled area - log10(ha)")
+pdf("paper/figures/res_v_extent_ksize.pdf", width = 7, height = 12, )
+par(mfrow = c(4, 2), oma = c(2, 2, 0, 0), mar = c(4, 4, 4, 4))
+lapply(kernsl, function(x) {
+  # x <- kernsl[[1]]
+  image(x$res, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
+        ylab = "")#, ylim = range(aaxis1$logarea))
+  axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
+       las = 2, tcl = -0.2, cex.axis = cxa)
+  axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
+       cex.axis = cxa)  
+  aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
+                cex.axis = cxa, tcl = -0.1)
+  plot(x$res, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
+
+  image(x$ext, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
+        ylab = "")
+  axis(1, at = taxis2$logdays, labels = taxis2$label, las = 2)
+  axis(2, at = aaxis2$logarea, labels = alab2, las = 2)
+  aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
+                cex.axis = cxa, tcl = -0.1)
+  plot(x$ext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+       legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
+})
+par(xpd = NA)
+xx <- grconvertX(0.5, from = "ndc", to = "inches")
+ii <- c(0.4, 0.6, 0.8, 1)
+yy <- c(0.95, 0.75, 0.5, 0.25)
+# yy <- c(-1, -5, -10, -20)
+for(i in 1:length(ii)) {
+  y <- grconvertY(yy[i], from = "ndc", to = "inches")
+  # text(x = xx, y = y, labels = paste("kernel =", ii[i]))
+  # par(xpd = NA)
+  # mtext(text = paste("kernel =", ii[i]), side = 3, line = yy[i])
+}
 dev.off()
+
+# plot(rrs, xlim = c(-5, 15), ylim = c(-5, 15), 
+#      xlab = "Sample interval - log10(days)", 
+#      ylab = "Plot resolution - log10(m^2)")
+# plot(ers, box = FALSE, xlim = c(-5, 15), ylim = c(-5, 15), 
+#      xlab = "Study duration - log10(days)", ylab = "Sampled area - log10(ha)")
+# dev.off()
 ```

### 2016-04-05 Rmd/space-time.Rmd
1. Removed commented out code for old resampling with perturbation approach. 2. Cleaned up and compiled html output of space-time

@@
 # Space versus time analysis
-```{r, eval = FALSE}
+```{r, message = FALSE, warning=FALSE, results='hold'}
 library(ecoscales)
@@
 dat <- dat[, names(dat)[1:20], with = FALSE]
-unique(dat$DOI)  # 138
 ```
 
+`r length(unique(dat$DOI))` papers  main analysis
+
 ## Data prep
-
 ### Merge full and calibration datasets
-```{r, eval = FALSE}
+```{r, message=FALSE, warning=FALSE, results='hide'}
 dat2 <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
@@
 
-```{r, eval=FALSE}
+```{r, message=FALSE, warning=FALSE, results='hide'}
 # apply fixes (mostly to account for observer omissions)
@@
 
-### Bootstraps
+## Analyses
 
-### With uncertainties
+### Resampling with uncertainties
 
@@
 ```{r, eval=FALSE}
-p <- c(0.79, 0.99, 0.53, 1.22)
+p <- c(0.79, 0.99, 0.53, 1.22)  # from calibration.Rmd
 
@@
 
-# old function, retained here for easy recollection of first approach
-# full function for resampling
-# resamp_func <- function(dat, p, iter) {
-#   bout <- lapply(1:iter, function(x) {  # x <- 1
-#     if((x / 100) %in% 1:100) print(x)
-#     # amount to perturb
-#     a <- runif(1, min = 1 - p, max = 1 + p)  # plot_res perturb
-#     b <- runif(1, min = 1 - p, max = 1 + p)  # n_sites perturb
-#     d <- runif(1, min = 1 - p, max = 1 + p)  # samp_duration perturb
-#     e <- runif(1, min = 1 - p, max = 1 + p)  # t_btwn_samp perturb
-#     
-#     if(length(p) < 4) stop("p must be length of 4", call. = FALSE)
-# 
-#     
-#     # spatial variables
-#     dnew <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
-#                          t_btwn_samp, study_duration)])
-#     dnew[which(senst$res == 1), plot_res := plot_res * a]
-#     dnew[which(senst$n == 1), n_sites := n_sites * b]
-#     dnew[, sampled_area := (plot_res * n_sites) / 10000]
-# 
-#     # temporal variables
-#     dnew[which(senst$sampd == 1), samp_duration := samp_duration * d]
-#     dnew[which(senst$tbtwn == 1), t_btwn_samp := t_btwn_samp * e]
-#     dnew[, study_duration := {
-#       recalc_duration(study_duration, samp_duration, t_btwn_samp)
-#     }]
-#     dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
-#   })
-#   return(bout)
-# }
-
-# perturbed samples
-# p10 <- resamp_func(dat, 0.1, 1000)  # 10% perturb
-
 set.seed(1)
@@
 
-### Without uncertainties
+### Bootstrapping
 
@@
 
-### Perturbed boot-strap
+### Perturbed set
 ```{r, eval = FALSE}
@@
 
-### Ordinary bootstrap
+### Bootstrap
 ```{r, eval = FALSE}
@@
 
-### On primary dataset only
+### Primary dataset (main figure)
 ```{r, eval = FALSE}
@@
 
-```{r, eval = FALSE}
+```{r}
 svars <- c("plot_res", "n_sites", "sampled_area", "samp_duration",
@@
 sensl <- strsplit(sens, ";")
-unique(unlist(sensl))
+# unique(unlist(sensl))
 
@@
 # unique(unlist(sensl))
-unique(unlist(sensl2))
+# unique(unlist(sensl2))
 
@@
 # percentage of records that are uncertain
-colSums(senst) / nrow(senst) * 100
+round(colSums(senst) / nrow(senst) * 100, 1)
 
@@
 
-### Varying kernel sizes on resolution versus extent
+### Varying kernel sizes
 ```{r,eval = FALSE}
@@
   rr <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat3, x)
+  rr <- (rr / cellStats(rr, sum)) * 100
   er <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat3, x)
+  er <- (er / cellStats(er, sum)) * 100
   kerns <- list("res" = rr, "ext" = er)
@@
        cex.axis = cxa)  
-  aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
+  aargs <- list(mgp = c(3, 0.25, 0), #at = brklabs, labels = brklabs, 
                 cex.axis = cxa, tcl = -0.1)
   plot(x$res, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
+     legend.shrink = 0.9, #breaks = brks, 
+     col = rev(terrain.colors(30)))
 
@@
         ylab = "")
-  axis(1, at = taxis2$logdays, labels = taxis2$label, las = 2)
-  axis(2, at = aaxis2$logarea, labels = alab2, las = 2)
-  aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
+  axis(1, at = taxis2$logdays[-5], labels = taxis2$label[-5], las = 2,
+       cex.axis = cxa, tcl = -0.2)
+  axis(2, at = aaxis2$logarea, labels = alab2[-2], las = 2, cex.axis = cxa,
+       tcl = -0.2)
+  aargs <- list(mgp = c(3, 0.25, 0), #at = brklabs2, labels = brklabs2, 
                 cex.axis = cxa, tcl = -0.1)
   plot(x$ext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-       legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
+       legend.shrink = 0.9, #breaks = brks2, 
+       col = rev(terrain.colors(30)))
 })

### 2016-04-07 Rmd/space-time.Rmd
Intro paragraphs of paper drafted. Slight tweaks to select-titles.Rmd.

@@
 # return interval
-tlab1 <- c(expression("second"^-1), expression("minute"^-1),
+tlab1 <- c(expression(paste(NULL<="second"^-1)), expression("minute"^-1),
            expression("hour"^-1), expression("day"^-1), 
@@
 
-taxis2 <- cbind.data.frame(
-  c("second", "minute", "hour", "day", "week", "month", "year", "decade",
-    "century", "millenium", "10 KA"),  
-  c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 30, 365, 365 * 10, 
-    365 * 100, 365 * 1000, 365 * 10000))
+tlab2 <- c(expression(NULL<="second"), "minute", "hour", "day", "week", "month",
+           "year","decade", "century", "millenium", "10 KA")  
+taxis2 <- cbind.data.frame(c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 30,
+                             365, 365 * 10, 365 * 100, 365 * 1000, 365 * 10000))
 taxis2 <- data.table(taxis2)
-setnames(taxis2, names(taxis2), c("label", "days"))
+setnames(taxis2, names(taxis2), c("days"))
 taxis2[, logdays := log10(days)]
@@
 # plot resolution
-alab1 <- c(expression("0.01 cm"^2), expression("0.1 cm"^2), 
+alab1 <- c(expression(paste(NULL<="0.01 cm"^2)), expression("0.1 cm"^2), 
            expression("1 cm"^2), expression("10 cm"^2),
@@
            expression("100 m"^2), expression("1000 m"^2), 
-           "1 ha", "10 ha", "100 ha", "1000 ha", "10000 ha")
+           "1 ha", "10 ha", "100 ha", "1000 ha", 
+           expression(paste(NULL>="10000 ha")))
 
@@
 # sampled area
-alab2 <- c(expression("1 m"^2), expression("10 m"^2), expression("100 m"^2),
-           expression("1000 m"^2), expression("1 ha"), expression("10 ha"),
-           expression("100 ha"), expression("1000 ha"), 
+alab2 <- c(expression(paste(NULL<="1 m"^2)), expression("10 m"^2), 
+           expression("100 m"^2), expression("1000 m"^2), expression("1 ha"),
+           expression("10 ha"), expression("100 ha"), expression("1000 ha"), 
            expression(paste(10^4 , " ha")), expression(paste(10^5 , " ha")), 
@@
       ylab = "", breaks = brks2)
-axis(1, at = taxis2$logdays[-5], labels = taxis2$label[-5], las = 2,
+axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
      cex.axis = cxa, tcl = -0.2)
@@
 
+# temporal framing
+tempdat <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
+                      y = log10(datf$study_duration))
+tempdat$n <- rep(1, nrow(tempdat))
+tx <- range(tempdat$x)
+tx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
+tx[2] <- 6.56
+tempdat[x < rx[1], x := rx[1]]
+tempdat[x > rx[2], x := rx[2]]
+ty <- range(taxis2$logdays)
+ty <- c(floor(ty[1]), 5.56)#ceiling(ex[2]))
+tempdat[y < ty[1], y := ty[1]]
+tempdat[y > ty[2], y := ty[2]]
+coordinates(tempdat) <- ~x + y
+
+# spatial framing
+spatdat <- data.table(x = log10(datf$plot_res), y = log10(datf$sampled_area))
+spatdat$n <- rep(1, nrow(spatdat))
+sx <- range(aaxis1$logres)
+sx[1] <- -5  # set limits to 0.01 cm^2
+spatdat[x < sx[1], y := sx[1]]
+spatdat[x > sx[2], y := sx[2]]
+sy <- range(aaxis2$logarea)
+spatdat[y < sy[1], y := sy[1]]
+spatdat[y > sy[2], y := sy[2]]
+coordinates(spatdat) <- ~x + y
+
 # densities
@@
 
+# # temporal
+# tres <- kdensity(tx[1], tx[2], ty[1], ty[2], 0.1, tempdat, 1)
+# #spatial
+# sres <- kdensity(sx[1], sx[2], sy[1], sy[2], 0.1, spatdat, 1)
+# image(sres, col = rev(terrain.colors(30)))
+
 cxa = 0.7
@@
      cex.axis = cxa)
+mtext("Sampling interval", 1, cex = cxa, line = 3)
+mtext("Spatial resolution", 2, cex = cxa, line = 3)
 
@@
               cex.axis = cxa, tcl = -0.1)
+mtext("Temporal duration", 1, cex = cxa, line = 3)
+mtext("Spatial extent", 2, cex = cxa, line = 3)
+
 plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,

### 2016-04-07 paper/main.tex
Intro paragraphs of paper drafted. Slight tweaks to select-titles.Rmd.

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

\usepackage{scicite}
\usepackage{times}
\usepackage{graphicx}
\usepackage{lineno}

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

\title{The Spatial and Temporal Domains of Modern Ecology \\
OR Ecology's Spatio-temporal Domains\\
OR Space, Time, and Ecology}

\author
{Lyndon Estes$^{\ast1, 2}$, Labeeb Ahmed$^{3}$, Kelly Caylor$^{2}$, Jason Chang$^{3}$, \\
Jonathan Choi$^{4}$, Erle Ellis$^{3}$, Paul Elsen$^{4}$, and Tim Treur$^{4}$ \\
\\
\normalsize{$^{1}$Woodrow Wilson School, Princeton University, Princeton, NJ 08544, USA}\\
\normalsize{$^{2}$Civil and Environmental Engineering, Princeton University, Princeton, NJ 08544, USA}\\
\normalsize{$^{3}$Geography and Environmental Systems, University of Maryland Baltimore County,}\\
\normalsize{Baltimore, MD 21250, USA}\\
\normalsize{$^{4}$Ecology and Evolutionary Biology, Princeton University, Princeton, NJ 08544, USA}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  lestes@princeton.edu.}
}
% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\begin{document}

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
An abstract
%  An example of the style is the special\texttt{\{sciabstract\}} environment used to set up the abstract you see here.
\end{sciabstract}

\linenumbers
The scales at which ecosystems are observed plays a critical role in shaping our understanding of how they are structured and function \cite{levin_problem_1992,chave_problem_2013}.  Ecological patterns emerge within temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is the \emph{sine qua non} for understanding ecological phenomena \cite{levin_problem_1992}. Awareness of the importance of scale has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations (e.g. cite) to entire biomes (e.g. cite), while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad range of scales \cite{schneider_rise_2001,chave_problem_2013}.

Given the importance of multi-scale studies for providing ecological understanding, and the growing ability to undertake them, it is important to rigorously assess whether ecology is becoming a multi-scale discipline. One approach to answering this question is to quantify the spatial and temporal domains within which observations in ecological studies are collected. Observations provide the necessary means for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992,tilman_ecological_1989}, thus it stands to reason that the temporal and spatial range of ecological observations, and their density within different portions of those ranges, will shed light on modern ecology's progress towards a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}. In this study, we quantified the spatio-temporal domain of current ecological studies, using a representative sample of papers published between 2004-2014 in the top 30 ecological journals (by 2014 impact factor) to extract and quantify two key dimensions of spatial observation, resolution (grain) and total spatial extent, and their temporal corollaries, sampling interval and total temporal duration. We collected this information from 367 ecological observations \cite[defined here as non-manipulated, or ``natural'']{tilman_ecological_1989} reported within a 148 paper subset of 284 (add Tim's also) randomly selected articles .

%16 + 29 + 26 + 67 + 14 + tim + 10

\begin{enumerate}
 \item Intro/rationale 
% \begin{enumerate}
%   \item Understanding of natural patterns and processes depends upon the scales at which they are observed \cite{levin_problem_1992}(maybe   provide an example or two, could just come from Schneider)
%    \item Scale is an increasingly measured aspect of ecosystems \cite{schneider_rise_2001}
%    \item Given this importance, and the increasing awareness of scale in ecology, it is timely to consider the scales at which ecosystems are observed, as this provides insight into how our understanding of ecosystems might be affected (work in Levin's language about this). 
%    \end{enumerate}
   \item{What we did}
   \begin{enumerate}
    \item Examined representative sample of papers from top 30 ecology journals--random sample of all titles from 2005-2014. Selected studies having observation component, excluding experimental manipulations, purely theoretical studies, commentaries, etc. 
    \item Extracted information on the spatial resolution of the study, that is, the two-dimensional space that was observed in the study; the temporal analog, which was the interval between repeat observations; the total spatial extent covered by the observations; the total time period covered. 
    \item We recorded the scales that were actually covered by the observations, rather than potential scales that the collected data might represent--this would require information on autocorrelation length, which is not provided. 
    \subitem For spatial resolution, this was the finest unit representing complete spatial coverage, etc. etc. 
    \subitem Limitation for ocean studies, we did not study third dimension, volume (might be more relevant). 
   \end{enumerate}
   \item{What we found} 
   \begin{enumerate}
    \item We reviewed a total of X randomly selected papers, of which 148 met our inclusion criteria and from which 367 distinct records related to observations made were collected. [Say something about size of sample and amount of time collected???].  
    \item  Of these, the vast majority (X\%) were field-based studies, while Y\% involved automated collection of data via instrumentation, and just Z\% made use of remote sensing.   
    \item From a perspective of temporal resolution, the majority of observations made in ecology are not repeated.
    \item Spatial resolution - most observations are between 10cm$^2$ to 10m$^2$
    \item Spatial extent
    \item Temporal extent
   \end{enumerate}
   \item{Implications} 
   \begin{enumerate}
    \item Although limited sample size, insight into the scales at which science of ecology is making bulk of observations
    \item Despite recognized importance of making observation across scales \cite{levin_problem_1992}, this does not appear to be happening. 
    \item Technical capabilities for making observations are not achieving adoption, despite potential of remote sensing to provide multi-scale ecological data \cite{estes_predictive_2011,estes_habitat_2008} 
   \end{enumerate}

\end{enumerate}

Notes: Wheatley and Johnson (2009) citing arbitrariness of scales in wildlife studies (cited by Chave).

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/res_v_extent.pdf}
\vspace{-0.15 cm}
\caption{}
\label{afoto1}
\end{figure}

%\paragraph*{Displayed math.}

\bibliography{/Users/lestes/Dropbox/publications/fullbib}

\bibliographystyle{Science}

% Following is a new environment, {scilastnote}, that's defined in the
% preamble and that allows authors to add a reference at the end of the
% list that's not signaled in the text; such references are used in
% *Science* for acknowledgments of funding, help, etc.

%\begin{scilastnote}
%\item We've included in the template file \texttt{scifile.tex} a new
%environment, \texttt{\{scilastnote\}}, that generates a numbered final
%citation without a corresponding signal in the text.  This environment
%can be used to generate a final numbered reference containing
%acknowledgments, sources of funding, and the like, per {\it Science\/}
%style.
%\end{scilastnote}

% For your review copy (i.e., the file you initially send in for
% evaluation), you can use the {figure} environment and the
% \includegraphics command to stream your figures into the text, placing
% all figures at the end.  For the final, revised manuscript for
% acceptance and production, however, PostScript or other graphics
% should not be streamed into your compliled file.  Instead, set
% captions as simple paragraphs (with a \noindent tag), setting them
% off from the rest of the text with a \clearpage as shown  below, and
% submit figures as separate files according to the Art Department's
% instructions.

\clearpage

\end{document}

### 2016-04-07 Rmd/space-time.Rmd
Draft through initial results (basic statistics), new four panel main figure made (space-time.Rmd) and histogram of statistics, including summary of those statistics.

@@
 ```{r, eval=FALSE}
-p <- c(0.79, 0.99, 0.53, 1.22)  # from calibration.Rmd
+p <- c(0.64, 0.84, 0.53, 1.21)  # from calibration.Rmd
 
@@
 # sampled area
-alab2 <- c(expression(paste(NULL<="1 m"^2)), expression("10 m"^2), 
+alab2 <- c(expression(paste(NULL<="0.1 m"^2)), expression("1 m"^2),
+           expression("10 m"^2), 
            expression("100 m"^2), expression("1000 m"^2), expression("1 ha"),
@@
 aaxis2 <- cbind.data.frame(
-  c(0.1 / 10000, 10 / 10000, 1000 / 10000, 1, 10, 10^2, 10^3, 10^4, 10^5, 10^6,
-    10^7, 10^8, 10^9, 10^10))
+  c(0.1 / 10000, 1 / 10000, 10 / 10000, 100 / 10000, 1000 / 10000, 1, 10, 10^2,
+    10^3, 10^4, 10^5, 10^6, 10^7, 10^8, 10^9, 10^10))
 aaxis2 <- data.table(aaxis2)
@@
 rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-# rx[1] <- -5
+rx[1] <- -5
 rx[2] <- 6.56
@@
 rng <- range(rres[is.finite(rres)])
-rng[2] <- 0.1
+# rng[2] <- 0.1
 bwidth <- (rng[2] - rng[1]) / 30
@@
      cex.axis = cxa, tcl = -0.2)
-axis(2, at = aaxis2$logarea, labels = alab2[-2], las = 2, cex.axis = cxa,
+axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
      tcl = -0.2)
@@
 rng <- range(rext2[is.finite(rext2)])
-rng[2] <- 0.1
+# rng[2] <- 0.1
 bwidth <- (rng[2] - rng[1]) / 30
@@
 rng <- range(rext3[is.finite(rext3)])
-rng[2] <- 0.1
+# rng[2] <- 0.1
 bwidth <- (rng[2] - rng[1]) / 30
@@
 
-# # temporal
-# tres <- kdensity(tx[1], tx[2], ty[1], ty[2], 0.1, tempdat, 1)
-# #spatial
-# sres <- kdensity(sx[1], sx[2], sy[1], sy[2], 0.1, spatdat, 1)
-# image(sres, col = rev(terrain.colors(30)))
+# temporal
+tres <- kdensity(tx[1], tx[2], ty[1], ty[2], 0.1, tempdat, 1)
+tres <- (tres / cellStats(tres, sum)) * 100
+# spatial
+sres <- kdensity(sx[1], sx[2], sy[1], sy[2], 0.1, spatdat, 1)
+sres <- (sres / cellStats(sres, sum)) * 100
 
-cxa = 0.7
-pdf("paper/figures/res_v_extent.pdf", width = 7, 
-    height = 2.5)
-par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
-    mgp = c(2, 0.5, 0))
-image(rres3, col = rev(terrain.colors(30)), breaks = brks,
+# cxa = 0.7
+ll <- 3.5
+cxa = 0.8
+# pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
+pdf("paper/figures/kde4.pdf", width = 7, height = 5.5)
+par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
+    mgp = c(2, 0.25, 0))
+# resolutions
+# image(rres3, col = rev(terrain.colors(30)), breaks = brks,
+image(rres3, col = rev(terrain.colors(50)), 
      axes = FALSE, xlab = "", ylab = "")
-axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
+axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
      las = 2, tcl = -0.2, cex.axis = cxa)
@@
      cex.axis = cxa)
-mtext("Sampling interval", 1, cex = cxa, line = 3)
-mtext("Spatial resolution", 2, cex = cxa, line = 3)
-
+mtext("Sampling interval", 1, cex = cxa, line = ll)
+mtext("Spatial resolution", 2, cex = cxa, line = ll)
 aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
@@
 plot(rres3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
+     #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
+     legend.shrink = 0.9, col = rev(terrain.colors(50)))
 
-image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
-      ylab = "", breaks = brks2)
+# extents
+# image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
+image(rext3, col = rev(terrain.colors(50)), axes = FALSE, xlab = "", 
+      ylab = "")#, breaks = brks2)
 axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
      cex.axis = cxa, tcl = -0.2)
-axis(2, at = aaxis2$logarea, labels = alab2[-2], las = 2, cex.axis = cxa,
+axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
      tcl = -0.2)
@@
               cex.axis = cxa, tcl = -0.1)
-mtext("Temporal duration", 1, cex = cxa, line = 3)
-mtext("Spatial extent", 2, cex = cxa, line = 3)
+mtext("Temporal duration", 1, cex = cxa, line = ll)
+mtext("Spatial extent", 2, cex = cxa, line = ll)
+plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     #legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
+     legend.shrink = 0.9, col = rev(terrain.colors(50)))
 
-plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
+# spatial
+image(sres, col = rev(terrain.colors(100)), axes = FALSE, xlab = "", ylab = "")
+# points(spatdat$x, spatdat$y, cex = 0.1)
+axis(1, at = aaxis1$logres, labels = alab1, las = 2, tcl = -0.2, cex.axis = cxa)
+axis(2, at = aaxis2$logarea, labels = alab2, las = 2, tcl = -0.2, 
+     cex.axis = cxa)
+mtext("Spatial resolution", 1, cex = cxa, line = ll)
+mtext("Spatial extent", 2, cex = cxa, line = ll)
+aargs <- list(mgp = c(3, 0.25, 0), at = seq(0, 0.12, 0.02), 
+              labels = seq(0, 0.12, 0.02), cex.axis = cxa, tcl = -0.1)
+plot(sres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, col = rev(terrain.colors(100)))
+
+# temporal
+image(tres, col = rev(terrain.colors(50)), axes = FALSE, xlab = "", ylab = "")
+# points(spatdat$x, spatdat$y, cex = 0.1)
+axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
+     las = 2, tcl = -0.2, cex.axis = cxa)
+axis(2, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
+     cex.axis = cxa, tcl = -0.2)
+mtext("Sampling interval", 1, cex = cxa, line = ll)
+mtext("Temporal duration", 2, cex = cxa, line = ll)
+aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
+              cex.axis = cxa, tcl = -0.1)
+plot(tres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, col = rev(terrain.colors(100)))
 dev.off()
@@
 dat[t_btwn_samp == 0, .N] / dat[, .N]  # 31% of samples are once-offs
-dat[study_duration == samp_duration, .N]
+dat[study_duration == samp_duration, .N] / nrow(dat)
 
-par(mfrow = c(2, 2))
-datf[, {
-  hist(log10(plot_res), main = "Plot res")
-  hist(log10(t_btwn_samp), main = "Interval")
-  hist(log10(sampled_area), main = "Area")
-  hist(log10(study_duration), main = "Duration")
-}]
+hdat <- cbind("res" = log10(datf$plot_res), "ext" = log10(datf$sampled_area), 
+              "int" = log10(to_infin_byond(datf$t_btwn_samp)), 
+              "dur" = log10(datf$study_duration))
+hdat <- data.table(hdat)
+# set bounds for histograms
+hdat[res < aaxis1$logres[1], res := aaxis1$logres[1]]
+hdat[res > ry[2], res := ry[2]]
+hdat[ext < -6, ext := -6]  # set minimum extent -6 (0.01 m2)
+
+# hist(hdat$res, breaks = aaxis1$logres)
+# hist(log10(datf$plot_res), breaks = resbrks)
+# hist(log10(datf$plot_res), breaks = resbrks)
+
+pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
+par(mfrow = c(1, 4), mar = c(4, 2, 1, 1), oma = c(2, 3, 0, 0))
+# resolution
+h <- hist(hdat$res, breaks = aaxis1$logres, plot = FALSE)
+h$density <- h$counts / sum(h$counts) * 100
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
+     xlab = "Spatial resolution", ylim = c(0, 25), main = "", col = "red1")
+axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
+     tcl = -0.2, mgp = c(2, 0.25, -0.45))
+mtext("Percent of observations", side = 2, line = 3, cex = 0.8)
+
+# extent
+h <- hist(hdat$ext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
+h$density <- h$counts / sum(h$counts) * 100
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
+     xlab = "Spatial extent", ylim = c(0, 20), main = "", col = "red3")
+axis(1, at = c(-6, aaxis2$logarea), 
+     labels = c(expression(paste(NULL<="0.01m"^2)), expression(paste(0.1, m^2)), 
+                alab2[-1]), las = 2, cex.axis = cxa, 
+     tcl = -0.2, mgp = c(2, 0.25, -0.45))
+
+# interval
+tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
+h <- hist(hdat$int, breaks = taxis1$logdays, plot = FALSE)
+h$density <- h$counts / sum(h$counts) * 100
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
+     xlab = "Sampling interval", ylim = c(0, 30), main = "", col = "blue1")
+axis(1, at = tax, labels = tlab1[-10], las = 2, cex.axis = cxa, 
+     tcl = -0.2, mgp = c(2, 0.25, -0.45))
+
+# duration
+# tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
+h <- hist(hdat$dur, breaks = taxis2$logdays, plot = FALSE)
+h$density <- h$counts / sum(h$counts) * 100
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
+     xlab = "Temporal duration", ylim = c(0, 30), main = "", col = "blue4")
+axis(1, at = taxis2$logdays, labels = tlab2, las = 2, cex.axis = cxa, 
+     tcl = -0.2, mgp = c(2, 0.25, -0.45))
+dev.off()
+
+# stats
+axisl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
+              taxis2$logdays)
+hdatstat <- lapply(1:4, function(x) {
+  h <- hist(hdat[[x]], breaks = axisl[[x]], plot = FALSE)
+  h$density <- h$counts / sum(h$counts) * 100
+  h
+})
+
+# res
+l <- length(hdatstat[[1]]$density)
+sum(hdatstat[[1]]$density[1:6])  # 52 % between 10cm^2 and 1 m^2; 63% < 1m^2
+sum(hdatstat[[1]]$density[7:10]) # 25 % between 1 m^2 and 10000 m^2
+sum(hdatstat[[1]]$density[11:(l)]) # 12 % 1-100 ha
+# dat[plot_res >= 10000^2]
+aaxis1
+
+# extent
+l <- length(hdatstat[[2]]$density)
+sum(hdatstat[[2]]$density[1:7])  # 85 % < 10 ha
+sum(hdatstat[[2]]$density[1:2])  # 30 % < 1 m^2
+sum(hdatstat[[2]]$density[10:l]) # 7 % covered an area > 10 ha
+sum(hdatstat[[2]]$density[14:l]) # 7 % covered an area > 10 ha
+
+# interval
+l <- length(hdatstat[[3]]$density)
+sum(hdatstat[[3]]$density[l])  # 30% are once-offs
+sum(hdatstat[[3]]$density[1:3])  # 25 % < daily
+sum(hdatstat[[3]]$density[4:5])  # 20% daily up to monthly
+sum(hdatstat[[3]]$density[5:6])  # 35% monthly up to yearly
+sum(hdatstat[[3]]$density[7:9]) # 4 % yearly to decadal
+
+# duration
+l <- length(hdatstat[[4]]$density)
+sum(hdatstat[[4]]$density[1:3])  # 59 % up to 1 day
+sum(hdatstat[[4]]$density[4:6])  # 35% covered between 1 day and 1 year
+sum(hdatstat[[4]]$density[7:10]) # 6% year to century
 

### 2016-04-07 paper/main.tex
Draft through initial results (basic statistics), new four panel main figure made (space-time.Rmd) and histogram of statistics, including summary of those statistics.

@@
 Observations provide the necessary means for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992,tilman_ecological_1989}, thus it stands to reason that the temporal and spatial range of ecological observations, and their density within different portions of those ranges, will shed light on modern ecology's progress towards a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.
-In this study, we quantified the spatio-temporal domain of current ecological studies, using a representative sample of papers published between 2004-2014 in the top 30 ecological journals (by 2014 impact factor) to extract and quantify two key dimensions of spatial observation, resolution (grain) and total spatial extent, and their temporal corollaries, sampling interval and total temporal duration.
-We collected this information from 367 ecological observations \cite[defined here as non-manipulated, or ``natural'']{tilman_ecological_1989} reported within a 148 paper subset of 284 (add Tim's also) randomly selected articles .
+In this study, we quantified the spatio-temporal domain of current ecological studies, using a representative sample of papers published between 2004-2014 in the top 30 ecological journals (by 2014 impact factor) to measure two key dimensions of spatial observation, resolution (grain) and total spatial extent, and their temporal corollaries, sampling interval and total temporal duration.
+We collected this information from 367 ecological observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 148 paper subset of 299 randomly selected articles (1.4\% of all papers).
 
-%16 + 29 + 26 + 67 + 14 + tim + 10
+%16 + 29 + 26 + 67 + 14 + 37 + 10
+In terms of spatial resolution, here defined as the two-dimensional space in which all measurable features of a natural object were recorded (as opposed to sub-sampled), the majority 63\% were collected in plots having resolutions $<$1 m$^2$, while 25\% were collected within plots of 1 m$^2$, up to 1 ha, and the remaining 12\% in plots of $\geq$1 ha.
+The total spatial extent covered (the number of sampled sites multiplied by the spatial resolution) by 85\% of observations was $<$10 ha, while 31\% covered less than 1 m$^2$.
+Only 7\% covered an extent $\geq$10 ha, with just 1.1\% spanning areas $\geq$10 million ha.
+
+In the temporal dimensions, 30\% of the assessed observations were ``once-offs'' that were not repeated.
+High frequency observations (ranging between as or more frequent than once per second up to daily) comprised 25\% of observations, 20\% were made at daily to monthly time steps, while 35\% and 4\% were respectively made at monthly up to yearly and yearly to decadal intervals.
+The temporal duration of studies--the total amount of time the ecological feature was observed (the number of repeat observations multiplied by the effective sampling duration, SI)--was less than 1 day for 59\% of sampled observations,  35\% between 1 day and 1 year, and just 6\% covering greater than 1 year (including several paleoecological studies covering centuries).
+
+Relationships between scales (Figure 1)
+Vast majority of cases, plot resolution scales directly with the total extent of the study--these are field studies.  Those with
+
+Reflect nature of observation approaches. 
+Use of remote sensing still relatively rare in ecological studies--thus the synoptic view of ecosystems not being seen.
+
+High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled.
 
@@
    \begin{enumerate}
-    \item Examined representative sample of papers from top 30 ecology journals--random sample of all titles from 2005-2014.
+%    \item Examined representative sample of papers from top 30 ecology journals--random sample of all titles from 2005-2014.
 Selected studies having observation component, excluding experimental manipulations, purely theoretical studies, commentaries, etc.
-    \item Extracted information on the spatial resolution of the study, that is, the two-dimensional space that was observed in the study; the temporal analog, which was the interval between repeat observations; the total spatial extent covered by the observations; the total time period covered.
+%    \item Extracted information on the spatial resolution of the study, that is, the two-dimensional space that was observed in the study; the temporal analog, which was the interval between repeat observations; the total spatial extent covered by the observations; the total time period covered.
     \item We recorded the scales that were actually covered by the observations, rather than potential scales that the collected data might represent--this would require information on autocorrelation length, which is not provided.
@@
    \begin{enumerate}
-    \item We reviewed a total of X randomly selected papers, of which 148 met our inclusion criteria and from which 367 distinct records related to observations made were collected.
+%    \item We reviewed a total of X randomly selected papers, of which 148 met our inclusion criteria and from which 367 distinct records related to observations made were collected.
 [Say something about size of sample and amount of time collected???].
@@
 %\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/res_v_extent.pdf}
+\includegraphics[width=1\textwidth]{figures/kde4.pdf}
 \vspace{-0.15 cm}

### 2016-04-07 Rmd/space-time.Rmd
Histogram figure added to main text as Figure 1 (ties into text describing primary statistics). Histogram figure and main kde figure (Fig. 2) sub-panel labels added.

@@
      legend.shrink = 0.9, col = rev(terrain.colors(50)))
+mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05)
 
@@
      legend.shrink = 0.9, col = rev(terrain.colors(50)))
+mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05)
 
@@
      legend.shrink = 0.9, col = rev(terrain.colors(100)))
+mtext("C", side = 3, line = -1, cex = 0.8, adj = 0.05)
 
@@
      legend.shrink = 0.9, col = rev(terrain.colors(100)))
+mtext("D", side = 3, line = -1, cex = 0.8, adj = 0.05)
 dev.off()
@@
 mtext("Percent of observations", side = 2, line = 3, cex = 0.8)
+mtext("A", side = 3, line = -1.5, cex = 0.8, adj = 0.05)
 
@@
      tcl = -0.2, mgp = c(2, 0.25, -0.45))
+mtext("B", side = 3, line = -1.5, cex = 0.8, adj = 0.05)
 
@@
      tcl = -0.2, mgp = c(2, 0.25, -0.45))
+mtext("C", side = 3, line = -1.5, cex = 0.8, adj = 0.05)
 
@@
      tcl = -0.2, mgp = c(2, 0.25, -0.45))
+mtext("D", side = 3, line = -1.5, cex = 0.8, adj = 0.05)
+
 dev.off()

### 2016-04-07 paper/main.tex
Histogram figure added to main text as Figure 1 (ties into text describing primary statistics). Histogram figure and main kde figure (Fig. 2) sub-panel labels added.

@@
 %16 + 29 + 26 + 67 + 14 + 37 + 10
-In terms of spatial resolution, here defined as the two-dimensional space in which all measurable features of a natural object were recorded (as opposed to sub-sampled), the majority 63\% were collected in plots having resolutions $<$1 m$^2$, while 25\% were collected within plots of 1 m$^2$, up to 1 ha, and the remaining 12\% in plots of $\geq$1 ha.
-The total spatial extent covered (the number of sampled sites multiplied by the spatial resolution) by 85\% of observations was $<$10 ha, while 31\% covered less than 1 m$^2$.
+In terms of spatial resolution, here defined as the two-dimensional space in which all measurable features of a natural object were recorded (as opposed to sub-sampled), the majority 63\% were collected in plots having resolutions $<$1 m$^2$, while 25\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 12\% in plots of $\geq$1 ha (Fig.
+1A).
+The total spatial extent covered (the number of sampled sites multiplied by the spatial resolution) by 85\% of observations was $<$10 ha, while 31\% covered less than 1 m$^2$ (Fig.
+1B).
 Only 7\% covered an extent $\geq$10 ha, with just 1.1\% spanning areas $\geq$10 million ha.
 
-In the temporal dimensions, 30\% of the assessed observations were ``once-offs'' that were not repeated.
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\includegraphics[width=1\textwidth]{figures/hists.pdf}
+\vspace{-0.15 cm}
+\caption{Histograms of the spatial resolution (A) and extent (B), sampling interval (C) and temporal duration (D) of ecological observations collected from the surveyed ecological studies.
+Bars represent percentage of the 367 collected records falling within each bin.}
+\label{afoto1}
+\end{figure}
+
+In the temporal dimensions, 30\% of the assessed observations were ``once-offs'' that were not repeated (Fig.
+1C).
 High frequency observations (ranging between as or more frequent than once per second up to daily) comprised 25\% of observations, 20\% were made at daily to monthly time steps, while 35\% and 4\% were respectively made at monthly up to yearly and yearly to decadal intervals.
-The temporal duration of studies--the total amount of time the ecological feature was observed (the number of repeat observations multiplied by the effective sampling duration, SI)--was less than 1 day for 59\% of sampled observations,  35\% between 1 day and 1 year, and just 6\% covering greater than 1 year (including several paleoecological studies covering centuries).
+The temporal duration of studies--the total amount of time the ecological feature was observed (the number of repeat observations multiplied by the effective sampling duration, SI)--was less than 1 day for 59\% of sampled observations,  35\% between 1 day and 1 year, and just 6\% covering greater than 1 year (including several paleoecological studies covering centuries to millenia; Fig.
+1D).
 
-Relationships between scales (Figure 1)
+The juxtaposition of these scales of observations provides further insight into the spatio-temporal distribution of ecological observations (Fig. 2).  
 Vast majority of cases, plot resolution scales directly with the total extent of the study--these are field studies.  Those with
+
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\includegraphics[width=1\textwidth]{figures/kde4.pdf}
+\vspace{-0.15 cm}
+\caption{}
+\label{afoto1}
+\end{figure}
 
@@
 \begin{enumerate}
- \item Intro/rationale 
+% \item Intro/rationale 
 %    \begin{enumerate}
@@
     \item We recorded the scales that were actually covered by the observations, rather than potential scales that the collected data might represent--this would require information on autocorrelation length, which is not provided.
-    \subitem For spatial resolution, this was the finest unit representing complete spatial coverage, etc. etc. 
+%    \subitem For spatial resolution, this was the finest unit representing complete spatial coverage, etc. etc. 
     \subitem Limitation for ocean studies, we did not study third dimension, volume (might be more relevant). 
@@
     \item  Of these, the vast majority (X\%) were field-based studies, while Y\% involved automated collection of data via instrumentation, and just Z\% made use of remote sensing.   
-    \item From a perspective of temporal resolution, the majority of observations made in ecology are not repeated.
-    \item Spatial resolution - most observations are between 10cm$^2$ to 10m$^2$
-    \item Spatial extent
-    \item Temporal extent
+%    \item From a perspective of temporal resolution, the majority of observations made in ecology are not repeated.
+%    \item Spatial resolution - most observations are between 10cm$^2$ to 10m$^2$
+%    \item Spatial extent
+%    \item Temporal extent
    \end{enumerate}
@@
 Notes: Wheatley and Johnson (2009) citing arbitrariness of scales in wildlife studies (cited by Chave).
-
-\begin{figure}[!ht]
-%\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/kde4.pdf}
-\vspace{-0.15 cm}
-\caption{}
-\label{afoto1}
-\end{figure}
 

### 2016-04-08 paper/main.tex
Drafted through main results

@@
 \oddsidemargin 0.2cm
-\textwidth 16cm 
+\textwidth 16.5cm 
 \textheight 21cm
@@
 
-The juxtaposition of these scales of observations provides further insight into the spatio-temporal distribution of ecological observations (Fig. 2).  
-Vast majority of cases, plot resolution scales directly with the total extent of the study--these are field studies.  Those with
+The juxtaposition of these observational scales provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (illustrated using two-dimensional kernel density estimates of the log-transformed scale values; Fig.
+2).
+A comparison between the sampling interval and spatial resolution of ecological observations reveals three primary regions of observational concentration (Fig.
+1A), the greatest being once-off samples with resolutions of 100 cm$^2$ to 1 m$^2$, followed next by observations of daily to monthly frequency of slightly finer resolution domain (10 cm$^2$ to slightly less than 1 m$^2$), and then a smaller center of monthly to yearly measurements collected at scales of 0.1-10 ha.
+Almost no observations were collected with intervals greater than a decade, at all spatial resolutions, while high frequency (daily to sub-daily) observations were rare and collected at very fine resolutions ($<$1000 cm$^2$).
+Almost no observations with $\geq$1 m$^2$ resolution were collected with sub-second (effectively continuous) to hourly frequency, and the density of 10-1000 m$^2$ resolution, sub-daily to sub-monthly frequency observations was also low.
 
-\begin{figure}[!ht]
+\begin{figure}[h]
 %\begin{wrapfigure}{c}{1\textwidth}
@@
 
-Reflect nature of observation approaches. 
-Use of remote sensing still relatively rare in ecological studies--thus the synoptic view of ecosystems not being seen.
+In terms of temporal duration and spatial extent (Fig.
+1B), the highest frequency ecological observations span time periods of 1 hour to $<$1 month, and cover $\leq$1 m$^2$ of space.
+The duration-extent domain extent appears to reach its apex at one day of duration and 1000-10,000 ha of extent.
+The very few observations collected that cover $>$1,000,000 ha were of sub-second duration, as these were collected by satellites that make instantaneous measurements.
 
-High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled.
+Comparing the two spatial scales of measurements against one another shows that observational extent scales log-linearly with resolution (Fig.
+1C), revealing a nearly 1:1 correspondence up to 100 ha of both resolution and extent.
+This relationship indicates that the larger the plot size, the fewer the number of replicates, which is a characteristic of field-based studies because effort and cost typically increases with plot size \cite{kareiva_spatial_1988}.
+In our sample, 85\% of the observations were collected using manual, field-based methods.
+These stand in contrast to remotely sensed observations, which made up just 4\% of the sample, but which are characterized by plot resolutions that are very high relative to their spatial extent--these are the low density patches above the 1:1 line defined by the denser concentration of field studies, namely the patches at the following resolution/extent domains; 1-10 m$^2$/100-10,000 ha; 0.1-10 ha/10 million ha; 10-10,000 ha/100,000-10 billion ha.
 
-\begin{enumerate}
-% \item Intro/rationale 
-%    \begin{enumerate}
-%   \item Understanding of natural patterns and processes depends upon the scales at which they are observed \cite{levin_problem_1992}(maybe   provide an example or two, could just come from Schneider)
-%    \item Scale is an increasingly measured aspect of ecosystems \cite{schneider_rise_2001}
-%    \item Given this importance, and the increasing awareness of scale in ecology, it is timely to consider the scales at which ecosystems are observed, as this provides insight into how our understanding of ecosystems might be affected (work in Levin's language about this).
-%    \end{enumerate}
-   \item{What we did}
-   \begin{enumerate}
-%    \item Examined representative sample of papers from top 30 ecology journals--random sample of all titles from 2005-2014.
-Selected studies having observation component, excluding experimental manipulations, purely theoretical studies, commentaries, etc.
-%    \item Extracted information on the spatial resolution of the study, that is, the two-dimensional space that was observed in the study; the temporal analog, which was the interval between repeat observations; the total spatial extent covered by the observations; the total time period covered.
-    \item We recorded the scales that were actually covered by the observations, rather than potential scales that the collected data might represent--this would require information on autocorrelation length, which is not provided.
-%    \subitem For spatial resolution, this was the finest unit representing complete spatial coverage, etc. etc. 
-    \subitem Limitation for ocean studies, we did not study third dimension, volume (might be more relevant). 
-   \end{enumerate}
-   \item{What we found} 
-   \begin{enumerate}
-%    \item We reviewed a total of X randomly selected papers, of which 148 met our inclusion criteria and from which 367 distinct records related to observations made were collected.
-[Say something about size of sample and amount of time collected???].
-    \item  Of these, the vast majority (X\%) were field-based studies, while Y\% involved automated collection of data via instrumentation, and just Z\% made use of remote sensing.   
-%    \item From a perspective of temporal resolution, the majority of observations made in ecology are not repeated.
-%    \item Spatial resolution - most observations are between 10cm$^2$ to 10m$^2$
-%    \item Spatial extent
-%    \item Temporal extent
-   \end{enumerate}
-   \item{Implications} 
-   \begin{enumerate}
-    \item Although limited sample size, insight into the scales at which science of ecology is making bulk of observations
-    \item Despite recognized importance of making observation across scales \cite{levin_problem_1992}, this does not appear to be happening. 
-    \item Technical capabilities for making observations are not achieving adoption, despite potential of remote sensing to provide multi-scale ecological data \cite{estes_predictive_2011,estes_habitat_2008}
-   \end{enumerate}
+Comparing sampling interval to temporal duration shows that the bulk of once-off observations last between a minute to a week in duration (Fig.
+1D).
+For repeated observations, the greatest density of observations were those made at monthly to yearly intervals and an aggregate duration of hours to weeks.
+Very few repeat observations exceeded a total duration measured in years, and those that did were of daily to yearly frequency.
 
-\end{enumerate}
+Reflect nature of observation approaches. \\
+Use of remote sensing still relatively rare in ecological studies--thus the synoptic view of ecosystems not being seen.  \\
+High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled. \\
+Don't forget to mention automated approaches. 
+Limitation for ocean studies, we did not study third dimension, volume (might be more relevant).
+Notes: Wheatley and Johnson (2009) citing arbitrariness of scales in wildlife studies (cited by Chave).
 
-Notes: Wheatley and Johnson (2009) citing arbitrariness of scales in wildlife studies (cited by Chave).
+Although limited sample size, insight into the scales at which science of ecology is making bulk of observations
+Despite recognized importance of making observation across scales \cite{levin_problem_1992}, this does not appear to be happening. 
+Technical capabilities for making observations are not achieving adoption, despite potential of remote sensing to provide multi-scale ecological data \
 

### 2016-04-09 Rmd/space-time.Rmd
First draft complete, except for wrap-up paragraph

@@
 dat <- dat[, names(dat)[1:20], with = FALSE]
+
+unique(dat$study_type)
+dat[, study_type := tolower(study_type)]
+kwords <- c("field", "paleo", "remote", "automated", "other")
+for(i in 1:length(kwords)) dat[like(study_type, kwords[i]), st := i]
+dat[st == 5, .N] / nrow(dat)
+
 ```
@@
 ```{r, message=FALSE, warning=FALSE, results='hide'}
-dat2 <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
+dat2 <- copy(dat[, .(st, plot_res, n_sites, sampled_area, samp_duration,
                      t_btwn_samp, study_duration)])
-
-rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", 
+rnms <- c("st", "DOI", "feature", "observer", "plot_res", "n_sites", 
           "sampled_area", "t_btwn_samp", "study_duration")
 calr <- cal[, rnms, with = FALSE]
-for(j in names(calr)[c(2, 4:8)]) set(calr, j=j, value = as.numeric(calr[[j]]))
+for(j in names(calr)[c(3, 5:9)]) set(calr, j=j, value = as.numeric(calr[[j]]))
 
 # select down to columns of interest
+# st <- calr$study_type[which(calr$feature != 99)]
 calr <- calr[feature != 99, lapply(.SD, mean), by = .(DOI, feature), 
-             .SDcols = rnms[4:8]]
+             .SDcols = rnms[c(1, 5:9)]]
 length(unique(calr$DOI)) + length(unique(dat$DOI))  # 148 papers providing dat
 
+# combine full and calibration
 calr[, samp_duration := NA]
 calr[, c("DOI", "feature") := NULL]
-setcolorder(calr, names(calr)[c(1:3, 6, 4:5)])
-
+setcolorder(calr, names(calr)[c(1:4, 7, 5:6)])
 datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
@@
 # image(rres3, col = rev(terrain.colors(30)), breaks = brks,
-image(rres3, col = rev(terrain.colors(50)), 
-     axes = FALSE, xlab = "", ylab = "")
+image(rres3, col = rev(terrain.colors(50)), axes = FALSE, xlab = "", ylab = "")
 axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
@@
 mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05)
+# points(resdat3$x[which(datf$st == 4)], resdat3$y[which(datf$st == 4)], 
+#        pch = 20, col = "red", cex = 0.5)
+# points(resdat3$x[which(datf$st == 3)], resdat3$y[which(datf$st == 3)], 
+#        pch = 20, col = "blue", cex = 0.5)
+# points(resdat3$x[which(datf$st == 5)], resdat3$y[which(datf$st == 5)], 
+#        pch = 20, col = "blue", cex = 0.5)
 
@@
 
+image(rres3, col = rev(terrain.colors(100)))
+points(resdat3$x, resdat3$y, pch = 20, cex = 0.5)
 ```
@@
 
+cxa <- 0.9
 pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
-par(mfrow = c(1, 4), mar = c(4, 2, 1, 1), oma = c(2, 3, 0, 0))
+par(mfrow = c(1, 4), mar = c(4, 1, 1, 1), oma = c(2, 3, 0, 0))
 # resolution
@@
 sum(hdatstat[[3]]$density[l])  # 30% are once-offs
-sum(hdatstat[[3]]$density[1:3])  # 25 % < daily
+sum(hdatstat[[3]]$density[1:3])  # 24 % < daily
 sum(hdatstat[[3]]$density[4:5])  # 20% daily up to monthly
@@
 sum(hdatstat[[4]]$density[7:10]) # 6% year to century
+
+# datf[, plot(log10(n_sites), log10(plot_res))]
+dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
+dat[like(study_type, "field"), plot(log10(plot_res), log10(sampled_area), 
+                                    ylim = c(-10, 10))]
+
+dat[like(study_type, "remote"), 
+    points(log10(plot_res), log10(sampled_area), pch = 20, col = "red")]
+dat[like(study_type, "automated"), 
+    points(log10(plot_res), log10(sampled_area), pch = 20, col = "blue")]
+lines(aaxis1$logres, aaxis2$logarea[-16] )
+datf[, .N / nrow(datf) * 100, by = st]
+points(resdat3$x[which(datf$st == 4)], resdat3$y[which(datf$st == 4)])
+points(resdat3$x[which(datf$st == 1)], resdat3$y[which(datf$st == 1)])
 

### 2016-04-09 paper/main.tex
First draft complete, except for wrap-up paragraph

@@
 OR Ecology's Spatio-temporal Domains\\
-OR Space, Time, and Ecology}
+OR Space, Time, and Ecology \\
+OR Ecology: still a small-scale, field-based discipline}
 
@@
 \begin{sciabstract}
-An abstract
+A pithy and highly compelling abstract
 %  An example of the style is the special\texttt{\{sciabstract\}} environment used to set up the abstract you see here.
@@
 A comparison between the sampling interval and spatial resolution of ecological observations reveals three primary regions of observational concentration (Fig.
-1A), the greatest being once-off samples with resolutions of 100 cm$^2$ to 1 m$^2$, followed next by observations of daily to monthly frequency of slightly finer resolution domain (10 cm$^2$ to slightly less than 1 m$^2$), and then a smaller center of monthly to yearly measurements collected at scales of 0.1-10 ha.
+2A), the greatest being once-off samples with resolutions of 100 cm$^2$ to 1 m$^2$, followed next by observations of daily to monthly frequency of slightly finer resolution domain (10 cm$^2$ to slightly less than 1 m$^2$), and then a smaller center of monthly to yearly measurements collected at scales of 0.1-10 ha.
 Almost no observations were collected with intervals greater than a decade, at all spatial resolutions, while high frequency (daily to sub-daily) observations were rare and collected at very fine resolutions ($<$1000 cm$^2$).
@@
 
-\begin{figure}[h]
+\begin{figure}[!ht]
 %\begin{wrapfigure}{c}{1\textwidth}
@@
 \vspace{-0.15 cm}
-\caption{}
+\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by A) sampling interval and spatial resolution, B) temporal duration and spatial extent, C) spatial resolution and spatial extent, and D) sampling interval and temporal duration.
+Density estimates were applied to the log-transformed values of each observation dimension, and density estimates are rescaled to represent percentages.
+}
 \label{afoto1}
@@
 In terms of temporal duration and spatial extent (Fig.
-1B), the highest frequency ecological observations span time periods of 1 hour to $<$1 month, and cover $\leq$1 m$^2$ of space.
+2B), the highest frequency ecological observations span time periods of 1 hour to $<$1 month, and cover $\leq$1 m$^2$ of space.
 The duration-extent domain extent appears to reach its apex at one day of duration and 1000-10,000 ha of extent.
@@
 Comparing the two spatial scales of measurements against one another shows that observational extent scales log-linearly with resolution (Fig.
-1C), revealing a nearly 1:1 correspondence up to 100 ha of both resolution and extent.
+2C), revealing a nearly 1:1 correspondence up to 100 ha of both resolution and extent.
 This relationship indicates that the larger the plot size, the fewer the number of replicates, which is a characteristic of field-based studies because effort and cost typically increases with plot size \cite{kareiva_spatial_1988}.
@@
 Comparing sampling interval to temporal duration shows that the bulk of once-off observations last between a minute to a week in duration (Fig.
-1D).
+2D).
 For repeated observations, the greatest density of observations were those made at monthly to yearly intervals and an aggregate duration of hours to weeks.
@@
 
-Reflect nature of observation approaches. \\
-Use of remote sensing still relatively rare in ecological studies--thus the synoptic view of ecosystems not being seen.  \\
-High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled. \\
-Don't forget to mention automated approaches. 
-Limitation for ocean studies, we did not study third dimension, volume (might be more relevant).
-Notes: Wheatley and Johnson (2009) citing arbitrariness of scales in wildlife studies (cited by Chave).
+This analysis reveals that ecology is still primarily a field-based science that is necessarily focused on small spatial scales, both in terms of sampling resolution and extent of coverage (Fig.
+2C).
+The reason for this is primarily due to the surprisingly low use of remote sensing, which has been available for several decades and is increasingly easy to use.
+Yet, despite prominent and repeated calls for ecologists to adopt remote sensing because it provides a synoptic view that field measurements cannot \cite{turner_remote_2003,kerr_space_2003,pettorelli_satellite_2014}, and subsequent demonstration of its importance for multi-scale ecological studies \cite{estes_habitat_2008,estes_predictive_2011}, our results indicate that this observational method is dramatically under-utilized by ecologists, likely because of perceived (and in some cases real) logistical barriers \cite{pettorelli_satellite_2014}.
+[Something here about how this is falling; and in something about UAVs]
 
-Although limited sample size, insight into the scales at which science of ecology is making bulk of observations
-Despite recognized importance of making observation across scales \cite{levin_problem_1992}, this does not appear to be happening. 
-Technical capabilities for making observations are not achieving adoption, despite potential of remote sensing to provide multi-scale ecological data \
+These relatively narrow spatial domains of observation naturally raise a question about the degree to which ecologists have truly set about to understand phenomena across multiple scales \cite{levin_problem_1992}.
+Adding to this concern is the fact that the scales of observation we measured were not clearly reported in the majority of studies; the spatial resolution and extent of studies was uncertain in 64\% and 70\% of sampled cases (and even the number of spatial replicates was not clearly reported for 23\% of the sample), while uncertainty in the precise interval and study duration were uncertain in 43\% and 86\% of cases, respectively.
+Although sensitivity analyses show that this uncertainty does not alter the domains of observation (SI), this lack of clarity supports the idea that ecology as a discipline is still insufficiently attentive to the concept of scale \cite{chave_problem_2013,wheatley_factors_2009}.
+
+However, in the temporal domains there appears to a greater diversity of scales represented (Fig.
+1D).
+At least 70\% of sampled observations involved repeat assessment, and one-quarter of these were of high frequency.
+Automated surface-based sensors provided most of the very high frequency observations, and comprised 9\% of sampled observations, which shows that, unlike remote sensing, these technologies are being more broadly adopted in ecology.
+As with spatial resolution, finely resolved temporal data can be used to derive less frequent, aggregated observations (whereas long-interval data cannot be disaggregated to create high frequency observations), thereby enabling multi-scale analyses.
+
+Temporal duration was of course relatively narrow in most studies, but here we calculated the summed time of repeat observations, which is brief in most instances (e.g.
+for satellite imaging) and thus amounts to a short period of time.
+Frequent snapshots in time may be sufficient to capture the temporal dynamics of many phenomena (e.g.
+changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these our measurement of this domain may understate the temporal breadth of ecological observations [note: might need to show this, at least in supplementals].
+However, for more dynamic phenomena long periods of continuous observation may be more important for understanding dynamics than frequent repeats (e.g.
+behavioral ecology, can someone help here with a CITE and or a better example).
+For this reason we chose to capture the integrated observational duration, rather then the span of time between the first and last observation.
+
+[still need a wrap-up paragraph, e.g.
+ecology appears to not be producing the diversity of scales in observation that would allow it to develop and test hypothesis across scales, more so in spatial domains then in temporal.
+Technological advances are offering ever more opportunities to advance this, e.g.
+UAVs for high frequency, high resolution observation (Fig.
+2.A), cite Gaston paper in Frontiers, Dandois and Ellis, but the question is will discipline overcome the inertia for adopting in particular geospatial advances, etc, etc.]
+
+%Based on this analysis, ecology thus appears to have
+
+%High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled. \\
+%Limitation for ocean studies, we did not study third dimension, volume (might be more relevant).
+
+%Although limited sample size, insight into the scales at which science of ecology is making bulk of observations
 

### 2016-04-09 paper/main.tex
Added total title count and corrected sample percentage of this total

@@
 In this study, we quantified the spatio-temporal domain of current ecological studies, using a representative sample of papers published between 2004-2014 in the top 30 ecological journals (by 2014 impact factor) to measure two key dimensions of spatial observation, resolution (grain) and total spatial extent, and their temporal corollaries, sampling interval and total temporal duration.
-We collected this information from 367 ecological observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 148 paper subset of 299 randomly selected articles (1.4\% of all papers).
+We collected this information from 367 ecological observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 148 paper subset of 299 randomly selected articles (0.67\% of 44,346 papers).
 

### 2016-06-02 Rmd/calibration.Rmd
Full draft as of 1 June 2016

@@
     toc: yes
+    number_sections: true
+    toc_depth: 3
     fig_caption: true
@@
 
-```{r, eval=FALSE}
+```{r, eval=FALSE, echo = FALSE}
 # Read in and clean-up calibration data
-cal <- data.table(read_excel(fp(p_dat, "merged_calibration.xlsx")))
-cal[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
-cal[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
-names(cal)
-dropnm <-  c("V29")#, "V31", "journal", "exclusion reason", "sensitivity")
-cal[, c(dropnm) := NULL]
-
-# check data types
-for(j in names(cal)[8:24]) set(cal, j = j, value = as.numeric(cal[[j]]))
-
-# use NA in n_sites as indicator that study was excluded
-cal[, inout := ifelse(is.na(n_sites), 0, 1)]  
+# cal <- data.table(read_excel(fp(p_dat, "merged_calibration.xlsx")))
+# cal[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
+# cal[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
+# names(cal)
+# dropnm <-  c("V29")#, "V31", "journal", "exclusion reason", "sensitivity")
+# cal[, c(dropnm) := NULL]
+# 
+# # check data types
+# for(j in names(cal)[8:24]) set(cal, j = j, value = as.numeric(cal[[j]]))
+# 
+# # use NA in n_sites as indicator that study was excluded
+# cal[, inout := ifelse(is.na(n_sites), 0, 1)]  
 # write.csv(cal, file = fp(p_dat, "merged_calibrationf.csv"))
@@
 # cal <- data.table(read_excel(fp(p_dat, "merged_calibrationf.xlsx")))
-cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
+# cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
+cal <- data.table(read_excel(fp(p_dat, "merged_calibrationffff.xlsx")))
 names(cal)
-cal[, c("notes", "V29") := NULL]
+cal[, c("notes", "V30") := NULL]
 for(j in names(cal)[c(4:5, 10:24)]) {
@@
 
-Others need to fix their study duration figures. Paul on paleo one in calibration, Choi, Chang, etc in main set.  [Ahmed remaining]
+Others need to fix their study duration figures. Paul on paleo one in calibration, Choi, Chang, etc in main set.  [DONE]
 

### 2016-06-02 Rmd/space-time.Rmd
Full draft as of 1 June 2016

--- 2016-04-07 Rmd/space-time.Rmd
@@
 library(maptools)
+library(RColorBrewer)
 
@@
 # dat <- as.data.table(read_excel(full_path(p_dat, "merged.xlsx")))
-dat <- as.data.table(read_excel(full_path(p_dat, "mergedf.xlsx")))
-cal <- data.table(read_excel(fp(p_calib, "merged_calibrationff.xlsx")))
+dat <- as.data.table(read_excel(full_path(p_dat, "mergedfff.xlsx")))
+cal <- data.table(read_excel(fp(p_calib, "merged_calibrationffff.xlsx")))
+full <- data.table(read_excel(fp(p_dat, "full_less_Treuer.xlsx")))
 
@@
 dat[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
+full <- gsub("DOI:|doi:|DOI: |doi: |DOI ", "", full$DOI)
+full <- gsub("\\s", "", full)  # remove DOI and whitespace
 
 dat <- dat[, names(dat)[1:20], with = FALSE]
+
+unique(dat$study_type)
+dat[, study_type := tolower(study_type)]
+kwords <- c("field", "paleo", "remote", "automated", "other")
+for(i in 1:length(kwords)) dat[like(study_type, kwords[i]), st := i]
+
+# dat[st == 5, .N] / nrow(dat)
+
+# class(cal$DOI)
+# class(dat$DOI)
+# unique(cal$DOI) %in% unique(dat$DOI)
+# match(cal$DOI, dat$DOI)
+# match(dat$DOI, cal$DOI)
+# dat[dat$DOI %in% cal$DOI, DOI]
+# dat[cal$DOI %in% dat$DOI, ]
+# cal[DOI == dat[cal$DOI %in% dat$DOI, DOI][11]]
+# dat[DOI == dat[cal$DOI %in% dat$DOI, DOI][7]]
+
 ```
-
-`r length(unique(dat$DOI))` papers  main analysis
 
@@
 ```{r, message=FALSE, warning=FALSE, results='hide'}
-dat2 <- copy(dat[, .(plot_res, n_sites, sampled_area, samp_duration,
-                     t_btwn_samp, study_duration)])
-
-rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", 
-          "sampled_area", "t_btwn_samp", "study_duration")
+dat2 <- copy(dat[, .(st, DOI, study_type, plot_res, n_sites, sampled_area,
+                     samp_duration, t_btwn_samp, study_duration)])
+rnms <- c("st", "DOI", "study_type", "feature", "observer", "plot_res", 
+          "n_sites", "sampled_area", "t_btwn_samp", "study_duration")
 calr <- cal[, rnms, with = FALSE]
-for(j in names(calr)[c(2, 4:8)]) set(calr, j=j, value = as.numeric(calr[[j]]))
+for(j in names(calr)[c(4, 6:10)]) set(calr, j=j, value = as.numeric(calr[[j]]))
 
 # select down to columns of interest
+# st <- calr$study_type[which(calr$feature != 99)]
 calr <- calr[feature != 99, lapply(.SD, mean), by = .(DOI, feature), 
-             .SDcols = rnms[4:8]]
-length(unique(calr$DOI)) + length(unique(dat$DOI))  # 148 papers providing dat
+             .SDcols = rnms[c(1, 6:10)]]
+length(unique(calr$DOI)) + length(unique(dat$DOI))  # 140 papers providing dat
+unique(calr$DOI)
+calr_type <- data.table(cbind("DOI" = unique(calr$DOI), # calib field studies
+                              "study_type" = c(rep("field", 5), "paleo",
+                                               rep("field", 4))))
 
+# combine full and calibration
 calr[, samp_duration := NA]
-calr[, c("DOI", "feature") := NULL]
-setcolorder(calr, names(calr)[c(1:3, 6, 4:5)])
+calr[, "feature" := NULL] #c("DOI", "feature") := NULL]
+calr <- merge(calr, calr_type, by = "DOI")
+setcolorder(calr, names(calr)[c(2:1, 9, 3:5, 8, 6:7)])
+datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
 
-datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
+doisup <- c(unique(cal$DOI_data_source), unique(dat$DOI_data_source))
+doisup <- doisup[!is.na(doisup) & doisup != "NA" & doisup != "-"]
+doisup <- unlist(strsplit(doisup, ";"))
+
 ```
+
+`r length(unique(datf$DOI))` papers  main analysis
+
+`r length(unique(full)) + 20 + 57` papers main analysis (everyone's but Tim's)
+
+`r (length(unique(full)) + 20 + 57) / 42918 * 100` % of all papers since 2004
+
+`r length(unique(doisup))` additional papers or other publications tracked down
+
+`r nrow(datf)` records (ecological observations)
+
+<a href="#top">Back to top</a>
 
@@
 
+<a href="#top">Back to top</a>
+
 ## Analyses
-
 ### Resampling with uncertainties
@@
 ```{r, eval=FALSE}
-p <- c(0.64, 0.84, 0.53, 1.21)  # from calibration.Rmd
+# p <- c(0.64, 0.84, 0.53, 1.21)  # from calibration.Rmd
+p <- c(0.58, 0.82, 0.99, 1.24)  # from calibration.Rmd
 
@@
 
+<a href="#top">Back to top</a>
+
+```{r, eval = FALSE, echo = FALSE}
 ### Bootstrapping
-
-Good old-fashioned bootstrapping with replacement, no perturbation of value
-```{r, eval = FALSE}
+# Good old-fashioned bootstrapping with replacement, no perturbation of value
 # bootstrap
@@
 ```
+
+<a href="#top">Back to top</a>
 
@@
            expression("century"^-1), 
-           expression("millenium"^-1), "once-off")# expression(infinity))
+           expression("millenium"^-1), "unreplicated")# expression(infinity))
 taxis1 <- cbind.data.frame( 
@@
            "year","decade", "century", "millenium", "10 KA")  
-taxis2 <- cbind.data.frame(c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 30,
-                             365, 365 * 10, 365 * 100, 365 * 1000, 365 * 10000))
+taxis2 <- cbind.data.frame(c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 
+                             30, 365, 365 * 10, 365 * 100, 365 * 1000, 
+                             365 * 10000))
 taxis2 <- data.table(taxis2)
@@
 
+brkfun <- function(ext, ival, n) {
+  rng <- range(ext[is.finite(ext)])
+  bwidth <- (rng[2] - rng[1]) / n
+  brks <- seq(rng[1], rng[2], bwidth)
+  brklabs <- seq(0, round(rng[2], 2), ival)
+  list("brks" = brks, "labs" = brklabs)
+}
+cuts <- 80
+
 cxa = 0.7
-pdf("paper/figures/res_v_extent_pbstrap.pdf", width = 7, 
+pdf("paper/figures/res_v_extent_pbstrap2.pdf", width = 7, 
     height = 2.5)
@@
 # image(rres, col = cols, breaks = brks,
-image(rres, col = rev(terrain.colors(30)), breaks = brks,
+brks <- brkfun(rres, 0.01, cuts)  # breaks
+image(rres, col = rev(terrain.colors(cuts)), #breaks = brks,
      axes = FALSE, xlab = "", ylab = "")
@@
 # plot(rres2, axes = FALSE, legend.only = TRUE)
-aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
+aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
               cex.axis = cxa, tcl = -0.1)
 plot(rres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
-image(rext, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
-      ylab = "", breaks = brks2)
+     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
+brks <- brkfun(rext, 0.01, cuts)  # breaks
+image(rext, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
+      ylab = "")
 axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
@@
      tcl = -0.2)
-aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
+aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
               cex.axis = cxa, tcl = -0.1)
 plot(rext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
+     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
 dev.off()
-
 ```
 
-### Bootstrap
-```{r, eval = FALSE}
+<a href="#top">Back to top</a>
+
+```{r, eval = FALSE, echo = FALSE}
+### Bootstrap - deprecated for now
+
 bootdt <- rbindlist(bootsamp)
@@
 cxa = 0.7
-pdf("paper/figures/res_v_extent_bstrap.pdf", width = 7, 
+pdf("paper/figures/res_v_extent_bstrap2.pdf", width = 7, 
     height = 2.5)
@@
 # datf[, plot(log10(study_duration), log10(sampled_area), xlim)]
+```
 
-```
+<a href="#top">Back to top</a>
 
@@
 rres3 <- (rres3 / cellStats(rres3, sum)) * 100
-rng <- range(rres3[is.finite(rres3)])
-bwidth <- (rng[2] - rng[1]) / 30
-brks <- seq(rng[1], rng[2], bwidth)
-brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
+# rng <- range(rres3[is.finite(rres3)])
+# bwidth <- (rng[2] - rng[1]) / 30
+# brks <- seq(rng[1], rng[2], bwidth)
+# brklabs <- seq(0, 0.12, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
+# brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
 
@@
 rext3 <- (rext3 / cellStats(rext3, sum)) * 100
-rng <- range(rext3[is.finite(rext3)])
-# rng[2] <- 0.1
-bwidth <- (rng[2] - rng[1]) / 30
-brks2 <- seq(rng[1], rng[2], bwidth)
-brklabs2 <- brklabs #round(brks2[seq(1, length(brks2), 3)], 2)
+# rng <- range(rext3[is.finite(rext3)])
+# # rng[2] <- 0.1
+# bwidth <- (rng[2] - rng[1]) / 30
+# brks2 <- seq(rng[1], rng[2], bwidth)
+# brklabs2 <- brklabs #round(brks2[seq(1, length(brks2), 3)], 2)
 
@@
 
+brkfun <- function(ext, ival, n) {
+  rng <- range(ext[is.finite(ext)])
+  bwidth <- (rng[2] - rng[1]) / n
+  brks <- seq(rng[1], rng[2], bwidth)
+  brklabs <- seq(0, round(rng[2], 2), ival)
+  list("brks" = brks, "labs" = brklabs)
+}
+
 # cxa = 0.7
@@
 cxa = 0.8
+stps <- sapply(c(3, 4), function(x) which(datf$st == x))  # IDs rs/auto/paleo
+pchs <- list(1, "+")
+cexs <- c(0.4, 0.6)
+cuts <- 80
+
 # pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
-pdf("paper/figures/kde4.pdf", width = 7, height = 5.5)
+pdf("paper/figures/kde43.pdf", width = 7, height = 5.5)
 par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
@@
 # image(rres3, col = rev(terrain.colors(30)), breaks = brks,
-image(rres3, col = rev(terrain.colors(50)), 
-     axes = FALSE, xlab = "", ylab = "")
+image(rres3, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
+      ylab = "")
 axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
@@
 mtext("Spatial resolution", 2, cex = cxa, line = ll)
-aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
+brks <- brkfun(rres3, 0.01, cuts)  # breaks
+aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
               cex.axis = cxa, tcl = -0.1)
@@
      #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
-     legend.shrink = 0.9, col = rev(terrain.colors(50)))
+     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
+mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05)
+# for(i in 1:length(stps)) {
+#   points(resdat3$x[stps[[i]]], resdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
+# }
 
@@
 # image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
-image(rext3, col = rev(terrain.colors(50)), axes = FALSE, xlab = "", 
+image(rext3, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
       ylab = "")#, breaks = brks2)
@@
      tcl = -0.2)
-aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
+brks <- brkfun(rext3, 0.01, cuts)  # breaks
+aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
               cex.axis = cxa, tcl = -0.1)
@@
      #legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
-     legend.shrink = 0.9, col = rev(terrain.colors(50)))
+     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
+mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05)
+# for(i in 1:length(stps)) {
+#   points(extdat3$x[stps[[i]]], extdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
+# }
 
 # spatial
-image(sres, col = rev(terrain.colors(100)), axes = FALSE, xlab = "", ylab = "")
+image(sres, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
+      ylab = "")
 # points(spatdat$x, spatdat$y, cex = 0.1)
@@
 mtext("Spatial extent", 2, cex = cxa, line = ll)
-aargs <- list(mgp = c(3, 0.25, 0), at = seq(0, 0.12, 0.02), 
-              labels = seq(0, 0.12, 0.02), cex.axis = cxa, tcl = -0.1)
+brks <- brkfun(sres, 0.01, cuts)  # breaks
+aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
+              cex.axis = cxa, tcl = -0.1)
 plot(sres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, col = rev(terrain.colors(100)))
+     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
+mtext("C", side = 3, line = -1, cex = 0.8, adj = 0.05)
+# for(i in 1:length(stps)) {
+#   points(spatdat$x[stps[[i]]], spatdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
+# }
 
 # temporal
-image(tres, col = rev(terrain.colors(50)), axes = FALSE, xlab = "", ylab = "")
+image(tres, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", ylab = "")
 # points(spatdat$x, spatdat$y, cex = 0.1)
@@
 mtext("Temporal duration", 2, cex = cxa, line = ll)
-aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
+brks <- brkfun(tres, 0.02, cuts)  # breaks
+aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
               cex.axis = cxa, tcl = -0.1)
 plot(tres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, col = rev(terrain.colors(100)))
+     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
+mtext("D", side = 3, line = -1, cex = 0.8, adj = 0.05)
+# for(i in 1:length(stps)) {
+#   points(tempdat$x[stps[[i]]], tempdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
+# }
 dev.off()
@@
 
+# image(rres3, col = rev(terrain.colors(100)))
+# points(resdat3$x, resdat3$y, pch = 20, cex = 0.5)
+image(sres, col = rev(terrain.colors(100)))
+points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)
 ```
 
-## Ancillary/supplementary plots and stats
+<a href="#top">Back to top</a>
+
+## Stats and supplementary plots
 
@@
                c("remote sensing", "other geographic data"))) / nrow(dat)
-dat[t_btwn_samp == 0, .N] / dat[, .N]  # 31% of samples are once-offs
+length(which(dat$study_type %in% 
+               c("remote sensing", "other geographic data"))) / nrow(dat)
+dat[t_btwn_samp == 0, .N] / dat[, .N]  # 35% of samples are once-offs
 dat[study_duration == samp_duration, .N] / nrow(dat)
+dat[(study_duration == samp_duration) & t_btwn_samp != 0]
+dat[(study_duration != samp_duration) & t_btwn_samp == 0]
 
@@
 
-pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
-par(mfrow = c(1, 4), mar = c(4, 2, 1, 1), oma = c(2, 3, 0, 0))
+mga <- -0.34
+sfigl <- -1.2
+cxa <- 0.8
+cxl <- 0.9
+reds <- brewer.pal(9, name = "Reds")
+blues <- brewer.pal(9, name = "Blues")
+plot(1:2, pch = 20, col = blues[c(4, 6)])
+yl <- 35
+# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
+pdf("paper/figures/hists2.pdf", width = 5, height = 5)
+par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
 # resolution
@@
 h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
-     xlab = "Spatial resolution", ylim = c(0, 25), main = "", col = "red1")
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
+     tcl = -0.2, cex.axis = cxa, xlab = "Spatial resolution", 
+     ylim = c(0, yl - 10), main = "", col = reds[4], cex.lab = cxl)
 axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
-     tcl = -0.2, mgp = c(2, 0.25, -0.45))
-mtext("Percent of observations", side = 2, line = 3, cex = 0.8)
+     tcl = -0.2, mgp = c(2, 0.25, mga))
+mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
+mtext("A", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
 
@@
 h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
-     xlab = "Spatial extent", ylim = c(0, 20), main = "", col = "red3")
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
+     tcl = -0.2, cex.axis = cxa, xlab = "Spatial extent", ylim = c(0, yl - 10), 
+     main = "", col = reds[6], cex.lab = cxl)
 axis(1, at = c(-6, aaxis2$logarea), 
@@
                 alab2[-1]), las = 2, cex.axis = cxa,
-     tcl = -0.2, mgp = c(2, 0.25, -0.45))
+     tcl = -0.2, mgp = c(2, 0.25, mga))
+mtext("B", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
 
@@
 h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
-     xlab = "Sampling interval", ylim = c(0, 30), main = "", col = "blue1")
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
+     tcl = -0.2, cex.axis = cxa, xlab = "Sampling interval", ylim = c(0, yl), 
+     main = "", col = blues[4], cex.lab = cxl)
 axis(1, at = tax, labels = tlab1[-10], las = 2, cex.axis = cxa, 
-     tcl = -0.2, mgp = c(2, 0.25, -0.45))
+     tcl = -0.2, mgp = c(2, 0.25, -0.34))
+mtext("C", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
+mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
 
@@
 h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", 
-     xlab = "Temporal duration", ylim = c(0, 30), main = "", col = "blue4")
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
+     tcl = -0.2, cex.axis = cxa, xlab = "Temporal duration", ylim = c(0, yl), 
+     main = "", col = blues[6], cex.lab = cxl)
 axis(1, at = taxis2$logdays, labels = tlab2, las = 2, cex.axis = cxa, 
-     tcl = -0.2, mgp = c(2, 0.25, -0.45))
+     tcl = -0.2, mgp = c(2, 0.25, mga))
+mtext("D", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
+
 dev.off()
@@
 l <- length(hdatstat[[1]]$density)
-sum(hdatstat[[1]]$density[1:6])  # 52 % between 10cm^2 and 1 m^2; 63% < 1m^2
-sum(hdatstat[[1]]$density[7:10]) # 25 % between 1 m^2 and 10000 m^2
-sum(hdatstat[[1]]$density[11:(l)]) # 12 % 1-100 ha
+sum(hdatstat[[1]]$density[4:6])  # 55 % between 10cm^2 and 1 m^2
+sum(hdatstat[[1]]$density[1:6])  #  69% < 1m^2
+sum(hdatstat[[1]]$density[7:10]) # 22 % between 1 m^2 and 1 ha
+sum(hdatstat[[1]]$density[11:(l)]) # 9 % 1-10000 ha
 # dat[plot_res >= 10000^2]
-aaxis1
+
+# length(which(hdat$res >= -3 & hdat$res <= 0)) / length(hdat$res)
+# length(which(hdat$res <= 4)) / length(hdat$res)
+# sum(hdatstat[[1]]$counts[4:6]) / sum(hdatstat[[1]]$counts)
 
@@
 l <- length(hdatstat[[2]]$density)
-sum(hdatstat[[2]]$density[1:7])  # 85 % < 10 ha
-sum(hdatstat[[2]]$density[1:2])  # 30 % < 1 m^2
-sum(hdatstat[[2]]$density[10:l]) # 7 % covered an area > 10 ha
-sum(hdatstat[[2]]$density[14:l]) # 7 % covered an area > 10 ha
+sum(hdatstat[[2]]$density[1:2])  # 35 % < 1 m^2
+sum(hdatstat[[2]]$density[1:6])  # 82% < 1 ha
+sum(hdatstat[[2]]$density[1:7])  # 86% < 10 ha
+sum(hdatstat[[2]]$density[1:8])  # 89 % < 100 ha
+sum(hdatstat[[2]]$density[1:9])  # 92 % < 1000 ha
+sum(hdatstat[[2]]$density[8:9])  # 6 % 10-1000 ha
+sum(hdatstat[[2]]$density[8:l])  # 14% > 10 ha
+sum(hdatstat[[2]]$density[10:l]) # 8.4 % covered an area > 1000 ha
+sum(hdatstat[[2]]$density[12:l]) # 5.1 % covered an area > 100,000 ha
+sum(hdatstat[[2]]$density[13:l]) # 4.1 % covered an area > 1,000,000 ha
+
+# length(which(hdat$ext < -4)) / length(hdat$ext)
+# length(which(hdat$ext > 3)) / length(hdat$ext)
+# length(which(hdat$ext > 1 & hdat$ext <= 3)) / length(hdat$ext)
+# which(hdatstat[[2]]$breaks == log10(1000000))
+# length(hdatstat[[2]]$counts[7:8])
+# hdatstat[[2]]$mids[12]
 
@@
 l <- length(hdatstat[[3]]$density)
-sum(hdatstat[[3]]$density[l])  # 30% are once-offs
-sum(hdatstat[[3]]$density[1:3])  # 25 % < daily
-sum(hdatstat[[3]]$density[4:5])  # 20% daily up to monthly
-sum(hdatstat[[3]]$density[5:6])  # 35% monthly up to yearly
-sum(hdatstat[[3]]$density[7:9]) # 4 % yearly to decadal
+sum(hdatstat[[3]]$density[l])  # 36% are once-offs
+sum(hdatstat[[3]]$density[1:3])  # 17 % < daily
+sum(hdatstat[[3]]$density[1:4])  # 17 % < daily
+sum(hdatstat[[3]]$density[4:5])  # 19% daily up to monthly
+sum(hdatstat[[3]]$density[6])  # 23% monthly up to yearly
+sum(hdatstat[[3]]$density[7:9]) # 5 % yearly to decadal
 
@@
 l <- length(hdatstat[[4]]$density)
-sum(hdatstat[[4]]$density[1:3])  # 59 % up to 1 day
-sum(hdatstat[[4]]$density[4:6])  # 35% covered between 1 day and 1 year
-sum(hdatstat[[4]]$density[7:10]) # 6% year to century
+sum(hdatstat[[4]]$density[1:3])  # 69 % up to 1 day
+sum(hdatstat[[4]]$density[4:5])  # 19% covered between 1 day and 1 month
+sum(hdatstat[[4]]$density[6])  # 8% covered between 1 month and 1 year
+# sum(hdatstat[[4]]$density[4:6])  # 26% covered between 1 day and 1 year
+sum(hdatstat[[4]]$density[7:10]) # 5% year to century
+
+# datf[, plot(log10(n_sites), log10(plot_res))]
+# dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
+datf[like(study_type, "field"), plot(log10(plot_res), log10(sampled_area), 
+                                     ylim = c(-10, 10))]
+
+datf[like(study_type, "remote"), 
+     points(log10(plot_res), log10(sampled_area), pch = 20, col = "red")]
+datf[like(study_type, "automated"), 
+     points(log10(plot_res), log10(sampled_area), pch = 20, col = "blue")]
+lines(aaxis1$logres, aaxis2$logarea[-16] )
+datf[, .N / nrow(datf) * 100, by = st]
+
+datf[, plot(log10(t_btwn_samp), log10(study_duration), pch = 20)]
+datf[which(hdat[[3]] < -1.380211), 
+     points(log10(t_btwn_samp), log10(study_duration), pch = 20, col = "blue")]
+datf[like(study_type, "automated"), 
+     points(log10(t_btwn_samp), log10(study_duration), pch = 20, col = "red")]
+
+# points(resdat3$x[which(datf$st == 4)], resdat3$y[which(datf$st == 4)])
+# points(resdat3$x[which(datf$st == 1)], resdat3$y[which(datf$st == 1)])
+
+# types of observations
+# unique(datf$study_type)
+datf[like(study_type, "field"), .N] / nrow(datf)  # 80% field studies
+datf[like(study_type, "auto"), .N] / nrow(datf) # 12 % automated
+datf[like(study_type, "auto") & log10(t_btwn_samp) < -1.380211, .N] /   length(which(hdat[[3]] < -1.380211))  # 75%
+
+datf[like(study_type, "remote|other"), .N] / nrow(datf) # 6 % automated
+datf[like(study_type, "paleo"), .N] / nrow(datf) * 100 # 1 % automated
 
 ```
+
+<a href="#top">Back to top</a>
 
@@
 
+<a href="#top">Back to top</a>
+
 ### Varying kernel sizes
@@
 # vary the kernel size about them
-kernsl <- lapply(c(0.4, 0.6, 0.8, 1), function(x) {
+kernsl <- lapply(c(0.4, 0.7, 1), function(x) {
   rr <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat3, x)
@@
 
+cxa <- 1.25
+
 # As separate plots
-pdf("paper/figures/res_v_extent_ksize.pdf", width = 7, height = 12, )
-par(mfrow = c(4, 2), oma = c(2, 2, 0, 0), mar = c(4, 4, 4, 4))
-lapply(kernsl, function(x) {
+pdf("paper/figures/res_v_extent_ksize.pdf", width = 8, height = 10)
+par(mfrow = c(3, 2), oma = c(6, 3, 0, 0), mar = c(1, 5, 3, 4))
+lapply(1:4, function(i) {
+  x <- kernsl[[i]]
   # x <- kernsl[[1]]
-  image(x$res, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
+  image(x$res, col = rev(terrain.colors(80)), axes = FALSE, xlab = "", 
         ylab = "")#, ylim = range(aaxis1$logarea))
+  if(i == 3) {
     axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
          las = 2, tcl = -0.2, cex.axis = cxa)
+  }
   axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
@@
      legend.shrink = 0.9, #breaks = brks, 
-     col = rev(terrain.colors(30)))
+     col = rev(terrain.colors(0)))
 
-  image(x$ext, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
+  image(x$ext, col = rev(terrain.colors(80)), axes = FALSE, xlab = "", 
         ylab = "")
-  axis(1, at = taxis2$logdays[-5], labels = taxis2$label[-5], las = 2,
+  if(i == 3) {
+    axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
          cex.axis = cxa, tcl = -0.2)
-  axis(2, at = aaxis2$logarea, labels = alab2[-2], las = 2, cex.axis = cxa,
+  }
+  axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
        tcl = -0.2)
@@
        legend.shrink = 0.9, #breaks = brks2, 
-       col = rev(terrain.colors(30)))
+       col = rev(terrain.colors(80)))
 })
@@
 ```
+
+<a href="#top">Back to top</a>
+
+### Scale diversity within study
+```{r}
+# plot resolution
+doin <- datf[, .N, by = DOI]  # n obs/DOI
+resuni <- datf[, length(unique(plot_res)), by = DOI]  # unique resolutions
+resrng <- datf[, diff(range(plot_res)), by = DOI]
+rescv <- datf[DOI %in% resuni[V1 > 1, DOI], cv(plot_res), by = DOI]
+ressd <- datf[DOI %in% resuni[V1 > 1, DOI], log10(sd(plot_res)), by = DOI]
+resmu <- datf[DOI %in% resuni[V1 > 1, DOI], log10(mean(plot_res)), by = DOI]
+# plot(resuni$V1, resrng$V1)
+# plot(resmu$V1, ressd$V1)
+# plot(resmu$V1, rescv$V1)
+
+# sampling interval
+intuni <- datf[, length(unique(t_btwn_samp)), by = DOI]  # unique intervals
+intrng <- datf[, diff(range(t_btwn_samp)), by = DOI]
+intcv <- datf[DOI %in% intuni[V1 > 1, DOI], cv(t_btwn_samp), by = DOI]
+intsd <- datf[DOI %in% intuni[V1 > 1, DOI], log10(sd(t_btwn_samp)), by = DOI]
+intmu <- datf[DOI %in% intuni[V1 > 1, DOI], log10(mean(t_btwn_samp)), by = DOI]
+
+# Plot Resolution Statistics
+round(c("%" = length(which(resuni$V1 > 1)) / resrng[, .N],  # % > 1 res 
+        "medrng" = resrng[which(resuni$V1 > 1), median(V1)],  # median res rng
+        "murng" = resrng[which(resuni$V1 > 1), mean(V1)] / 10000,  # mu res rng
+        "rescv" = rescv[, mean(V1)], # mean resolution CV
+        "resmu" = resmu[, mean(V1)],
+        "muN" = resuni[, mean(V1)]), 2)  # mean number of resolutions
+
+# Plot Resolution Statistics
+round(c("%" = length(which(intuni$V1 > 1)) / intrng[, .N],  # % > 1 interval
+        "medrng" = intrng[which(intuni$V1 > 1), median(V1)],  # median int range
+        "murng" = intrng[which(intuni$V1 > 1), mean(V1)],  # mean interval range
+        "intcv" = intcv[, mean(V1)], # mean intolution CV
+        "intmu" = intmu[, mean(V1)],
+        "muN" = intuni[, mean(V1)]), 2)  # mean number of intolutions
+```
+
+<a href="#top">Back to top</a>

### 2016-06-02 paper/SI.tex
Full draft as of 1 June 2016

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% www.sciencemag.org/about/authors/prep/TeX_help/ .
% This package should properly format in-text
% reference calls and reference-list numbers.

\usepackage{scicite}
\usepackage{times}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{longtable}
%\usepackage{multicolumn}
\usepackage[skip=2pt,font=normalsize]{caption}

% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16.5cm 
\textheight 21cm
\footskip 1.0cm

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% If your reference list includes text notes as well as references,
% include the following line; otherwise, comment it out.

\renewcommand\refname{References and Notes}

% The following lines set up an environment for the last note in the
% reference list, which commonly includes acknowledgments of funding,
% help, etc.  It's intended for users of BibTeX or the {thebibliography}
% environment.  Users who are hand-coding their references at the end
% using a list environment such as {enumerate} can simply add another
% item at the end, and it will be numbered automatically.

\newcounter{lastnote}
\newenvironment{scilastnote}{%
\setcounter{lastnote}{\value{enumiv}}%
\addtocounter{lastnote}{+1}%
\begin{list}%
{\arabic{lastnote}.}
{\setlength{\leftmargin}{.22in}}
{\setlength{\labelsep}{.5em}}}
{\end{list}}

% Include your paper's title here

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\begin{document} 
\beginsupplement

\linenumbers
\baselineskip24pt

\section*{Supplementary Information}% Make the title.
\subsection*{Methods}% Make the title.
\vspace{-4pt}
%\maketitle 
\subsubsection*{Paper selection}
\vspace{-10pt}
We used the 2012 Web of Science impact factors to select the highest ranked ecology-themed journals that published studies with an observational component, excluding journals devoted to reviews, meta-analyses, laboratory, cellular, or experimental studies.  To select a representative sample of recent ecology studies, we downloaded the metadata for all papers published in the selected journals (Table S1) between 2004 and 2014. Our study involved six different observers (those reviewing the papers to extract the observational scales), each of whom was given a randomly selected batch of 500 titles.  A separate set of 20 randomly selected papers was also given to each observer, which formed the basis for calibrating observers' interpretation and extraction of scale-related information, as well as estimating between observer variance.

\centerline{
%\hspace{1cm} 
\parbox{11cm}{\noindent Table 1. The selected journals and their 2012 impact factors.}
}
\vspace{-12pt}
%\begin{longtable}[h]
\begin{longtable}{lr}
%\centering
%\caption{The selected journals and their 2012 impact factors.}\\
%\centering
%\begin{tabular}{lr}
  \hline
\textbf{Journal} & \textbf{Impact Factor} \\
\hline
\endhead 
  \hline
Ecology Letters & 17.95 \\ 
  Ecological Monographs & 8.09 \\ 
  Frontiers In Ecology And The Environment & 7.62 \\ 
  Global Ecology And Biogeography & 7.22 \\ 
  Global Change Biology & 6.91 \\ 
  Diversity And Distributions & 6.12 \\ 
  Methods In Ecology And Evolution & 5.92 \\ 
  Proceedings Of The Royal Society B-biological Sciences & 5.68 \\ 
  Journal Of Ecology & 5.43 \\ 
  Ecology & 5.17 \\ 
  Ecography & 5.12 \\ 
  Journal Of Biogeography & 4.86 \\ 
  Functional Ecology & 4.86 \\ 
  Journal Of Animal Ecology & 4.84 \\ 
  Journal Of Applied Ecology & 4.74 \\ 
  American Naturalist & 4.55 \\ 
  Conservation Biology & 4.36 \\ 
  Ecological Applications & 3.81 \\ 
  Biological Conservation & 3.79 \\ 
  Biogeosciences & 3.75 \\ 
  Bulletin Of The American Museum Of Natural History & 3.48 \\ 
  Biology Letters & 3.35 \\ 
  Oikos & 3.32 \\ 
  Behavioral Ecology & 3.22 \\ 
  Ecosystems & 3.17 \\ 
  Advances In Ecological Research & 3.08 \\ 
  Oecologia & 3.01 \\ 
  Landscape Ecology & 2.90 \\ 
  Agriculture Ecosystems \& Environment & 2.86 \\ 
  Ecological Economics & 2.85 \\ 
   \hline
%\end{tabular}
\end{longtable}
% Place your abstract within the special {sciabstract} environment.

\subsubsection*{Estimating observational scales}
\vspace{-10pt}

Each observer first reviewed the papers in the calibration set, and then commenced reviewing papers in their individual random draws, beginning at the top of the list and then proceeding until at least 20 eligible papers describing ecological observations were reviewed. In cases where the reviewed papers used observations that were described in another publication, we reviewed those source papers in order to extract the observational dimensions. We excluded papers that were opinion or perspectives pieces (unless they presented or used existing observational data), theoretical studies based on generated data, or those which were entirely based on experimental manipulations. We left out the latter category because our intent was to evaluate the domains for observations of natural systems, and we wanted to avoid the bias that would be imposed by the relatively narrow spatial and temporal scales of experiments \cite{kareiva_spatial_1988,tilman_global_2011}.

To record the spatio-temporal dimensions of ecological observations, we considered the observational scales to be those at which a total measurement of the feature(s) of interest was made, rather than the scales that might be represented by the observations.  As an example, if a paper reported data on plant species diversity collected using 10 1 m$^2$ sub-plots randomly placed within a larger plot of 100 m$^2$, of which 10 were placed, we recorded 1 m$^2$ as the spatial resolution and 100 m$^2$ as the total extent of the observation. While the 10 sub-plots might be representative of diversity at larger scales (e.g. the particular plant community), that relationship depends both on sample design (number of replicates, placement, grain/resolution) and the properties of the ecosystem \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002}. Most studies also did not quantify the broader scales represented by their observations, nor did they report dimensions that would allow these to be independently estimated, such as the maximum distance between sample plots, which could have been used to estimate the spatial extent represented by the sample \cite{palmer_scale_1994}.

Similarly, we did not try to quantify the extent to which values of individual observations (e.g. a single plot) of an ecological feature might be correlated with spatially or temporally adjacent values.  Autocorrelation is common to most ecological phenomena \cite{legendre_spatial_1993,collins_method_2000-1}, and it means that an observation is descriptive of the properties within a broader area or time around the point at which it is taken. For example, soil properties are typically measured by analysis of soil cores collected using an augur. The two-dimensional resolution of this measurement is equal to the area of the augur, but the measured soil properties are likely to be very similar, if not identical, to those found in the soil within a radius of at least a few to several hundreds of meters around the core location. However, the distance and strength of this correlation can vary substantially according to the ecological features being observed, and such correlation lengths were reported in almost no studies. Attempting to estimate these lengths, and add them to actual observations scales, would have added further uncertainty to estimates. Moreover, this concern only applies to spatially and temporally discontinuous (or discontiguous) observations, and since many of observations are collected using techniques that permit continuous measurements (e.g. weather data over time or satellite images over space), we elected to use a consistent standard for estimating dimensions.

In assessing spatial scales, our analysis only considered evaluated the Cartesian plane. We did not consider the z, or depth, dimension, which for certain aspects of ecology, this measure is of greater value (e.g. depth profiles in marine ecology), although in some cases (paleoecology) values extracted from the z-dimension provided a temporal measure that was used to calculated both the interval and the duration of the observation.

%Furthermore quantifying the actual domains of observation is important because highly dynamic spatial and temporal processes, such as the population ecology of riverine fish \cite{fausch_landscapes_2002} or the evolution of wildfires \cite{clements_observing_2007}, may require continuous observation to properly measure, and can vary substantially in the space or time between observations.

A fuller description of the project methodology is contained within the answers to the list of questions below. This set of Frequently Asked Questions (FAQ) was provided to each observer for initial study and reference, in order to ensure methodological consistency (see next section).

\noindent\underline{\textbf{General:}}
\def\Hitem{\item [Q\stepcounter{enumi}\arabic{enumi}.]} 
\def\Hsubitem{\item [Q\arabic{enumi}.\arabic{enumi}.]} 
\vspace{-0.5cm}
\begin{enumerate}
  \Hitem \emph{\textbf{What are the general inclusion/exclusion criteria for studies?}} Studies should be excluded from this analysis if they are: 1) opinion/perspectives pieces; 2) model-only studies, particularly theoretical models, which are not developed or tested against observed data; 3) if they are experimental manipulations (but if there is a study that has a mix of observational and experimental, record the observational treatments and exclude the manipulated treatments).

  \Hitem \emph{\textbf{What are the standard categories to be used for defining “Study type”?}} Define study type according to the following categories: Remote sensing, passive/automated data collection, other geographic data (e.g. non-remotely sensed GIS data), field/direct observation, or paleo-reconstruction (tree rings, charcoal cores, etc).

  \Hitem \emph{\textbf{What happens when the study draws on a separately published dataset as a key part of the methods?}} Track down the study describing the paper, and then record the DOI of that paper/those papers.

  \Hitem \emph{\textbf{What is the best unique identifier of a study I am reviewing?}} The DOI!

  \Hitem \emph{\textbf{What do I record for a time or space scale when it is not clearly reported in the paper, or when I am unsure? For example, in a paleo-ecological study looking at historical charcoal deposition, sediment cores were extracted from lakes, which the authors report as the number of samples. However, it is unclear how many sediment cores were drawn from each lake, and it is these which should be the number of spatial replicates.}} For these sorts of issues, we record that the scale in question is uncertain, and then your best estimate of the measure (e.g. you might assume that only 1 core was made per lake).

\hspace{-1cm}\underline{\textbf{Temporal scales:}}
\vspace{-0.5cm} 
  \Hitem \emph{\textbf{What is sampling duration, and how do we record it?}} How long an individual observation of an individual point in space took to make. Sampling duration multiplied by the number of repeats observations is used to calculate study duration. Often this value will not be reported, so you will have to use your best judgement, based on your knowledge of ecological methods, to approximate the duration. For example, for a field based method with intensive plot methods, if you can't estimate a plausible duration, assign a token 1 day. For remote sensing observations you can assume one second (the observations are effectively instantaneous).  
   \Hitem \emph{\textbf{What is sampling interval, and how do we record it?}} Sampling interval is the time that elapsed between repeat observations of the same point in space or individual organism. In many cases, observations will only be made one time--list a value of 0 for these.  
  \Hitem \emph{\textbf{What is study duration, and how do we calculate it?}} Study duration is the integral of sampling duration, or the time spent making one observation of the phenomenon in question. To clarify, duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample, nor the integral of time spent in observing all spatial replicates. 
  \Hitem \emph{\textbf{Should study duration be the total time spent sampling all sites or the amount of time spent sampling per site (e.g. for 5 minute point counts of birds at 10 sites each repeated twice, should we enter 100 minutes (5 minutes X 2 repeats X 10 sites) or 10 minutes (5 minutes X 2 repeats) for duration)?}} As stated above, duration is the total spent observing a single point in space, so in this case that would be 10 minutes (then converted to days, so 10 / (60 * 24))).
  \Hitem \emph{\textbf{How do we record study duration when there are no repeat observations?}} In these cases, study duration is equal to sampling duration. 
   \Hitem \emph{\textbf{How do I record sampling interval in cases where the interval is inconsistent?  For example, in a study were observations were observations were repeated in 1979, 1980, 1981, 1984, 2007, 2009?}} Find the time between each successive period, and then take the average of that (remember to convert to days!). If there are two or more sets of unevenly spaced days for each site/plot/measurement being taken in the study, then find the average interval for each, and average the averages.  
    \Hitem \emph{\textbf{How do you determine the time between samples for paleo-reconstructions?}} Use the minimum estimate for dating precision as the estimate of time between samples (e.g. 50 years in the study of European charcoal deposits DOI:10.1111/geb.12090). 
    \Hitem \emph{\textbf{How do you determine the sample duration (our third time category) for paleo-reconstructions?}} Similar to the previous question, the sampling duration is also the same as the minimum estimate of dating precision. The logic behind this is that in such cases, where a sediment or tree core or similar measurement is being collected, this effectively represents a continuous ``observation'', and the value associated with the minimum (or other reported) interval is typically an average (or another summary statistic like the maximum) of the amount accumulated. 
    \Hitem \emph{\textbf{What about sampling intervals and durations for instrument-collected observations?}} These are often similar in essence to the paleo-reconstruction case. Take the minimum temperature or daily rainfall recorded at a weather station, which require constant observation over 24 hours to report.  In such cases, the sampling interval and sampling duration are both 24 hours. On the other hand, automated logging systems will provide a series of high frequency observations that are collected instantaneously. In these cases the sampling duration should be $\leq$second, and the sampling interval should be the period between successive instantaneous measurements. 
    \Hitem \emph{\textbf{How do you handle the time between samples for a case where repeated samples are taken during a season, across several seasons (e.g.``we performed repeat bird counts every 10 days between March and June of 2005, 2006, and 2008'')?}} Since the sampling is focused on seasons, and presumably some season-specific phenomenon (e.g. breeding behavior), the reported values should be pegged to the season, not averaged across the full study span (the start and end dates of the study). So in this example, it would be 10 days.

\hspace{-1cm}\underline{\textbf{Spatial scales:}}
  \Hitem \emph{\textbf{What is sampling resolution, and how do I record it?}} This is the finest scale at which a complete measurement of every unit of the quantity of interest is recorded.  For example, if the measurement in question is a tree stem count, the resolution is determined by the size of the plot used to record every tree stem. Taking this example further, let's say a study reports a plot size of 100 x 100 m, but then goes on to report that they counted stems within a single 1 m wide transect within this larger plot.  In this case, the plot resolution is in fact 100 m x 1 m, or 100 m$^2$ (sampling resolution should be reported in m$^2$).  In another example, if the reported plot size was 20 x 20 m, but the authors in fact only measured a random selection of, say, grass stems on which they counted aphids within those plots, then use an estimate of the area of the grass stem as the sampling resolution \cite{gagic_agricultural_2012}.  
  \Hitem \emph{\textbf{What is study extent, and how do I record it?}} Broadly defined, study extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares. For studies in which the spatial replicates are not spatially contiguous (as with most field-based studies), this means plot resolution (see Q16) multiplied by number of plots. For spatially contiguous studies (e.g. those based on remote sensing imagery), it should be the total area covered by the imagery, i.e. pixel resolution multiplied by the number of pixels. However, only record the area analyzed by the authors. If they used a sub-sample of pixels, extent is the number of those pixels a multiplied by pixel resolution.

  \Hitem \emph{\textbf{How do you determine sampling resolution for paleo-reconstructions and other approaches where a sampling method is presumed to draw from a larger area (e.g. mammal traps, mist nets, etc)?}}  For sampling resolution, estimate the size of the sample actually taken, rather than the assumed catchment/shed area of the sample (e.g. the area of the corer used to take a sediment sample, rather than an estimate of the area that that sample is assumed to draw from), and then indicate that the plot resolution was uncertain. Related to this, you may also have to estimate the number of samples collected, as exemplified in a charcoal study of Europe where the number of lakes sampled was provided, not the number of cores per lake \cite{molinari_exploring_2013}.

  \Hitem \emph{\textbf{What about studies that sample individual organisms?}} If the study is making a total count of all organisms (let's say a mammal species) within a fixed plot size, or even a variable plot size from which an average plot size (and thus sampling resolution) can be estimated, then follow the procedures described in Q16-17. However, if the individual animals are the unit of measure (either because a sub-sample of them is being made within a defined plot, or because the observation is not contingent on being located within a plot (maybe a blood sample or body weight is being recorded, for example), then simply estimate two-dimensional area occupied by the animal, the number of animals sampled, and then calculate area from that. Occasionally individual animals might be recorded, but within the context of some natural feature, such as a nesting site where the survival of individual chicks is the measurement of interest \cite{roche_relative_2008-1}. In this case, an estimate of the nest area provides the sampling resolution. Another possible case is where individual animals are being tracked using radio or GPS collars. In these cases, use the estimate of the collared/tagged animal's body area, and then treat the number of reported locational observations as the number of spatial replicates, and then use those two values to calculate sampled area. The total GPS points collected from all individuals can be used if we can safely assume that each individual sampled did not cover the same exact location.  If the number of GPS points is not given, the number of fixes can be estimated from the duration during which individuals were collared and the recording interval.

\end{enumerate}

\subsubsection*{Calibration and consistency}
\vspace{-10pt}
Most studies did not explicitly report values for all four of the assessed scales, and thus interpretation and judgement had to be applied to develop reasonable estimates for their values. The FAQ provided the protocol we followed, and was initially developed following a consultative process prior to the commencement of review. We conducted an iterative process of calibration to ensure consistency and reliability of estimates. First, we used the calibration set to calculate between-observer variability with respect to paper selection/rejection and the estimation of scales. Based on this, the lead author reviewed individual records in each observer's calibration set, flagged values were the estimation procedure departed from the protocol, and returned these to observers for re-estimation, without providing an estimate of the actual value.  Instead, the relevant section of the protocol was highlighted, and further explanation and clarifying discussion undertaken as needed. Protocol language was adjusted for clarity during this process, and new items added to cover circumstances that had not been addressed by the initial version. The variability measures (see Results) were re-calculated after each iteration.

To ensure consistency within the main analysis, the lead author also reviewed each observer's results from their individual draw of papers and flagged values that appeared to deviate from the protocol for re-review by the observer. Re-reviewed values were re-inspected, and in some cases a secondary review of particular papers was undertaken to cross-check the estimated scales.

\subsubsection*{Estimating observational density}
\vspace{-10pt}
In presenting results (Fig. 1 and 2 in main text), we log-transformed (base 10) the observational scales because of the large range in values.  To reveal the densities of observations within two dimensions (Fig. 2 main text), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values, using a bandwidth of 1 onto a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations are concentrated. Bandwidths of varying resolutions were tested (see Supplementary Results section).

\subsubsection*{Sensitivity of results to observer and estimation error}
\vspace{-10pt}
To test the impact of between-observer inconsistencies on results, as well as the possibility of error by individual observers made in estimating the scales of actual observations scale, we conducted a sensitivity analysis. For each of the four dimensions, we performed a 1000-iteration resample of values, adding to each record a random amount of error at each iteration. The error values were constrained to fall within a maximum percentage of the record's value, which was determined by the coefficient of variation between observers' estimates for the calibration set.  We then combined the set of 1000 perturbed observations, log-transformed them, recalculated the kernel density estimates for the comparisons of the two resolution and extent dimensions (i.e. Fig. 2A and B in main text), and assessed the degree to which the domains of observational concentration differed from those in our main analysis (Fig. S2).

\subsection*{Results}
\vspace{-4pt}

\subsubsection*{Variability and consistency between observers}
\vspace{-10pt}

We assessed variability and consistency between observers along multiple dimensions. First, we assessed how reliablly observers agreed with respect to selecting or rejecting papers for review, using the R's Agreement package \cite{yu_agreement:_2012} to calculate Fleiss' kappa statistic \cite{fleiss_measuring_1971}, which was 0.72 (z = 12.5, p$<$0.000), indicating substantial and significant agreement between observers \cite{landis_measurement_1977}.

Second, we calculated the intra-class correlation coefficient \cite{bartko_intraclass_1966} to assess agreement between observers regarding the number of ecological observations that could be extracted from each paper (multiple ecological observations were reported in many studies, and we listed observations as separate records if they varied on one or more dimensions).  The coefficient, calculated using the IRR package \cite{gamer_irr:_2012} of R, was 0.70 (F = 15.4, p$<$0.0001, 95\% confidence interval = 0.54 - 0.85)

Finally, we calculated the coefficient of variation (CV) between observers' estimates of scales for each dimension, first across all six observers' mean scale estimates, and then as the average CV among observers' estimates of each individual record (Table S2).

\begin{table}[ht]
\captionsetup{width=0.725\linewidth}
\caption{Estimates of variability calculated from each observer's estimates of the spatial and temporal scales of ecological observations reported within a common set of the 20 papers used for calibration. Variability is expressed as the coefficient of variation (CV; standard deviation divided by mean multiplied by 100) between each observer's overall mean, and as the mean CV of observers' estimates for individual records. }
\vspace{-4 pt}
\begin{center}
\begin{tabular}{rrrrr}
\hline
& \multicolumn{2}{c}{\textbf{Spatial}} & \multicolumn{2}{c}{\textbf{Temporal}}\\\cline{2-5}
Value & Resolution & Extent & Interval & Duration \\ 
\hline
CV of overall mean & 54 & 82 & 54  & 110 \\
Mean of record-wise CV & 58 & 68 & 99 & 124 \\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}%

\subsubsection*{Sensitivity analysis}
We used the maximum uncertainty values for each dimension from the inter-observer variability analysis (Table S2) to determine the bounds of the random perturbations applied to each record in the sensitivity analysis. Kernel density estimates applied to the merged results of this analysis for the contrasts between spatial resolution and temporal intervals, and to spatial extent and temporal duration, suggest that the primary features observed in Figures 2A and 2B (main text) are largely unaltered by the addition of uncertainty to the estimates. The largest difference is in the greater density of second to sub-second duration observations across all spatial extents.

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=1\textwidth]{figures/res_v_extent_pbstrap2.pdf}
\vspace{-0.15 cm}
\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by sampling interval and spatial resolution (left) and temporal duration and spatial extent (right). Density estimates were applied to the log-transformed values of each observational dimension after being randomly perturbed within uncertainty bounds drawn from an assessment of between-observer variability. Density estimates are rescaled to represent percentages. }
\label{afoto1}
\end{figure}

Figure S2 indicates the effect that varying bandwidth has on the appearance of the kernel density estimates, as estimated for the same two contrasts shown in Figure S1. the smallest bandwidth (0.4) tested (Fig. S2 top) shows that the primary observational concentrations described in the main text are evident but harder to discern.  Most difficult to discern is the oblong concentration of observations that in Fig. 1A (main text) is bounded on the lower right at monthly to yearly intervals and 100-1000 m$^2$ resolutions and on the upper left by near-daily to monthly observations and 0.1-10 ha resolutions. With the smaller these appears as two separate patches (Fig. S2 top left), but with 0.7 bandwidth applied become coherent (Fig. S3 middle left).

\begin{figure}[!ht]
%\begin{wrapfigure}{c}{1\textwidth}
\includegraphics[width=0.8\textwidth]{figures/res_v_extent_ksize.pdf}
\vspace{10 pt}
\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by sampling interval and spatial resolution (left column) and temporal duration and spatial extent (right column), applied to log-transformed values of each observational dimension. Rows indicate the effects of selecting different bandwidths: 0.4 (top row); 0.7 (middle row); 1 (bottom row). }
\label{afoto1}
\end{figure}

%\subsection*{Rationale}

%\begin{figure}[!ht]
%%\begin{wrapfigure}{c}{1\textwidth}
%\includegraphics[width=1\textwidth]{figures/hists.pdf}
%\vspace{-0.15 cm}
%\caption{Histograms of the spatial resolution (A) and extent (B), sampling interval (C) and temporal duration (D) of ecological observations collected from the surveyed ecological studies. Bars represent percentage of the 367 collected records falling within each bin.}
%\label{afoto1}
%\end{figure}

\clearpage
\bibliography{/Users/lestes/Dropbox/publications/fullbib}

\bibliographystyle{Science}

% Following is a new environment, {scilastnote}, that's defined in the
% preamble and that allows authors to add a reference at the end of the
% list that's not signaled in the text; such references are used in
% *Science* for acknowledgments of funding, help, etc.

%\begin{scilastnote}
%\item We've included in the template file \texttt{scifile.tex} a new
%environment, \texttt{\{scilastnote\}}, that generates a numbered final
%citation without a corresponding signal in the text.  This environment
%can be used to generate a final numbered reference containing
%acknowledgments, sources of funding, and the like, per {\it Science\/}
%style.
%\end{scilastnote}

% For your review copy (i.e., the file you initially send in for
% evaluation), you can use the {figure} environment and the
% \includegraphics command to stream your figures into the text, placing
% all figures at the end.  For the final, revised manuscript for
% acceptance and production, however, PostScript or other graphics
% should not be streamed into your compliled file.  Instead, set
% captions as simple paragraphs (with a \noindent tag), setting them
% off from the rest of the text with a \clearpage as shown  below, and
% submit figures as separate files according to the Art Department's
% instructions.

\clearpage

\end{document}

### 2016-06-02 paper/main.tex
Full draft as of 1 June 2016

--- 2016-04-08 paper/main.tex
@@
 
-\usepackage{scicite}
+\usepackage[sort]{scicite}
 \usepackage{times}
@@
 
-\title{The Spatial and Temporal Domains of Modern Ecology \\
-OR Ecology's Spatio-temporal Domains\\
-OR Space, Time, and Ecology}
+\title{The Spatial and Temporal Domains of Modern Ecology }%\\
+%OR Ecology's Spatio-temporal Domains\\
+%OR Space, Time, and Ecology \\
+%OR Ecology: still a small-scale, field-based discipline}
 
@@
 {Lyndon Estes$^{\ast1, 2}$, Labeeb Ahmed$^{3}$, Kelly Caylor$^{2}$, Jason Chang$^{3}$, \\
-Jonathan Choi$^{4}$, Erle Ellis$^{3}$, Paul Elsen$^{4}$, and Tim Treur$^{4}$ \\
+Jonathan Choi$^{4}$, Erle Ellis$^{3}$, Paul R. Elsen$^{1}$, and Tim Treur$^{4}$ \\
 \\
@@
 \begin{sciabstract}
-An abstract
+Observational scales plays a critical role in shaping understanding of ecological pattern and process.
+To gain insight into the spatial and temporal domains being addressed by modern ecological science, we conducted a meta-analysis of papers published in 30 ecological journals between 2004 and 2014.
+We estimated the spatial resolution and extent and temporal interval and duration of 371 ecological observations reported within 140 papers (out of 364 randomly selected titles).
+Our analysis revealed that ecology remains a primarily field-based discipline that primarily collects data within narrow spatial and temporal domains.
+Ecologists still make limited use of methods (e.g.
+remote sensing, used for 6\% of observations) for observing larger-scale patterns and processes, although make greater use of those which permit high frequency or continuous temporal measurements (12\% of observations).
+Observational scales were not clearly reported in the majority of three dimensions, which, together with the narrow spatio-temporal ranges of observed scales suggests a limited ability to generalize the findings of many studies in modern ecological research.
+
 %  An example of the style is the special\texttt{\{sciabstract\}} environment used to set up the abstract you see here.
@@
 \linenumbers
-The scales at which ecosystems are observed plays a critical role in shaping our understanding of how they are structured and function \cite{levin_problem_1992,chave_problem_2013}.
-Ecological patterns emerge within temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is the \emph{sine qua non} for understanding ecological phenomena \cite{levin_problem_1992}.
-Awareness of the importance of scale has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations (e.g.
-cite) to entire biomes (e.g.
-cite), while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad range of scales \cite{schneider_rise_2001,chave_problem_2013}.
+The scales at which ecosystems are observed plays a critical role in shaping our understanding of how they are structured and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989-1}.
+Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is the \emph{sine qua non} for understanding ecological phenomena \cite{levin_problem_1992}.
+Awareness of the importance of scale has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad range of scales \cite{schneider_rise_2001,chave_problem_2013}.
 
-Given the importance of multi-scale studies for providing ecological understanding, and the growing ability to undertake them, it is important to rigorously assess whether ecology is becoming a multi-scale discipline.
-One approach to answering this question is to quantify the spatial and temporal domains within which observations in ecological studies are collected.
-Observations provide the necessary means for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992,tilman_ecological_1989}, thus it stands to reason that the temporal and spatial range of ecological observations, and their density within different portions of those ranges, will shed light on modern ecology's progress towards a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.
-In this study, we quantified the spatio-temporal domain of current ecological studies, using a representative sample of papers published between 2004-2014 in the top 30 ecological journals (by 2014 impact factor) to measure two key dimensions of spatial observation, resolution (grain) and total spatial extent, and their temporal corollaries, sampling interval and total temporal duration.
-We collected this information from 367 ecological observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 148 paper subset of 299 randomly selected articles (1.4\% of all papers).
+Given the centrality of scale to understanding ecology, and the growing ability to study ecological phenomena across a broad range of spatial and temporal scales, it is important to assess the scales at which ecological research is presently being conducted.
+To gain insight into this question, we quantified the spatial and temporal domains of empirical observations that were reported in a representative sample of studies published between 2004-2014 in the top 30 ecological journals (by 2014 impact factor).
+Empirical observations provide the necessary means for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, thus it stands to reason that the temporal and spatial distributions of ecological observations may be indicative of modern ecology's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.
+
+We characterized observational domains along two key spatial dimensions, resolution (grain) and total spatial extent, and their temporal corollaries, sampling interval and total temporal duration, and for each dimension calculated the actual, rather than representative, scales of observation (SI).
+Our review estimated the spatio-temporal scales of 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 364 randomly selected articles (0.85\% of total).
+In terms of spatial resolution, here defined as the two-dimensional space in which all measurable features of a natural object were recorded (as opposed to sub-sampled), the majority (69\%) of observations were collected in plots of $<$1 m$^2$ resolution, 22\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
+1A).
+The total spatial extent covered (the number of sampled sites multiplied by the spatial resolution) by 86\% of observations was $<$10 ha, while 35\% covered less than 1 m$^2$ (Fig.
+1B).
+Only 14\% covered an extent $\geq$10 ha, with just 4\% spanning areas $\geq$1 million ha.
 
 %16 + 29 + 26 + 67 + 14 + 37 + 10
-In terms of spatial resolution, here defined as the two-dimensional space in which all measurable features of a natural object were recorded (as opposed to sub-sampled), the majority 63\% were collected in plots having resolutions $<$1 m$^2$, while 25\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 12\% in plots of $\geq$1 ha (Fig.
-1A).
-The total spatial extent covered (the number of sampled sites multiplied by the spatial resolution) by 85\% of observations was $<$10 ha, while 31\% covered less than 1 m$^2$ (Fig.
-1B).
-Only 7\% covered an extent $\geq$10 ha, with just 1.1\% spanning areas $\geq$10 million ha.
 
@@
 %\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/hists.pdf}
+\includegraphics[width=1\textwidth]{figures/hists2.pdf}
 \vspace{-0.15 cm}
@@
 
-In the temporal dimensions, 30\% of the assessed observations were ``once-offs'' that were not repeated (Fig.
-1C).
-High frequency observations (ranging between as or more frequent than once per second up to daily) comprised 25\% of observations, 20\% were made at daily to monthly time steps, while 35\% and 4\% were respectively made at monthly up to yearly and yearly to decadal intervals.
-The temporal duration of studies--the total amount of time the ecological feature was observed (the number of repeat observations multiplied by the effective sampling duration, SI)--was less than 1 day for 59\% of sampled observations,  35\% between 1 day and 1 year, and just 6\% covering greater than 1 year (including several paleoecological studies covering centuries to millenia; Fig.
+In the temporal dimensions, 36\% of observations were not repeated (Fig.
+1C), 17\% were repeated at short intervals (sub-second to daily), 19\% at daily to monthly intervals, 23\% at monthly to yearly intervals, and 5\% at yearly to decadal intervals.
+The temporal duration of studies--defined here as the total amount of time the ecological feature was directly observed, and calculated as the number of repeat observations multiplied by the effective sampling duration (SI)--was less than one day for 69\% of sampled observations, between one day and one month for 19\% of observations, between one month and one year for 8\% of observations, and greater than one year for 5\% of observations (including several paleoecological studies covering centuries to millenia; Fig.
 1D).
 
-The juxtaposition of these observational scales provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (illustrated using two-dimensional kernel density estimates of the log-transformed scale values; Fig.
+The juxtaposition of these observational dimensions provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (Fig.
 2).
-A comparison between the sampling interval and spatial resolution of ecological observations reveals three primary regions of observational concentration (Fig.
-1A), the greatest being once-off samples with resolutions of 100 cm$^2$ to 1 m$^2$, followed next by observations of daily to monthly frequency of slightly finer resolution domain (10 cm$^2$ to slightly less than 1 m$^2$), and then a smaller center of monthly to yearly measurements collected at scales of 0.1-10 ha.
-Almost no observations were collected with intervals greater than a decade, at all spatial resolutions, while high frequency (daily to sub-daily) observations were rare and collected at very fine resolutions ($<$1000 cm$^2$).
-Almost no observations with $\geq$1 m$^2$ resolution were collected with sub-second (effectively continuous) to hourly frequency, and the density of 10-1000 m$^2$ resolution, sub-daily to sub-monthly frequency observations was also low.
+Contrasting spatial resolution with sampling interval reveals that the greatest density of observations consists of unreplicated samples collected in plots of 10 cm$^2$ to 1 m$^2$.
+Two smaller centers of observational concentration are also evident, the largest of which had plots that were also of 10 cm$^2$-1 m$^2$ and were revisited at daily to monthly intervals.
+A less dense, oblong concentration of observations defined on the lower right by monthly to yearly observations at 100-1000 m$^2$ and on the upper left by near-daily to monthly observations with 0.1-10 ha resolution is also evident.
+This lower right to upper left orientation marks the tradeoff between resolution and return that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-second intervals), high spatial resolution (0.1-100 cm$^2$) observations.
+This line demonstrates the inverse tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort and thus reduce sampling frequency \cite{kareiva_spatial_1988}.
 
-\begin{figure}[h]
+%Comparing the spatial and temporal resolutions of  interval to spatial resolution of ecological observations reveals three primary regions of observational concentration (Fig.
+2A), the greatest being unreplicated samples with resolutions of 100 cm$^2$ to 1 m$^2$, followed by observations of daily to monthly interval with slightly finer spatial resolutions (10 cm$^2$ to 1 m$^2$), and finally a smaller concentration of monthly to yearly measurements collected at spatial scales of 0.1-10 ha.
+
+\begin{figure}[!ht]
 %\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/kde4.pdf}
+\includegraphics[width=1\textwidth]{figures/kde43.pdf}
 \vspace{-0.15 cm}
-\caption{}
+\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by A) sampling interval and spatial resolution, B) temporal duration and spatial extent, C) spatial resolution and spatial extent, and D) sampling interval and temporal duration.
+Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.
+}
 \label{afoto1}
@@
 
-In terms of temporal duration and spatial extent (Fig.
-1B), the highest frequency ecological observations span time periods of 1 hour to $<$1 month, and cover $\leq$1 m$^2$ of space.
-The duration-extent domain extent appears to reach its apex at one day of duration and 1000-10,000 ha of extent.
-The very few observations collected that cover $>$1,000,000 ha were of sub-second duration, as these were collected by satellites that make instantaneous measurements.
+Contrasting temporal duration and spatial extent (Fig.
+2B) reveals that the bulk of ecological observations covered total time periods of one hour to $<$1 month and areas of $\leq$1 m$^2$.
+The next greatest concentration were observations centered on 1 hour duration and between 1 m$^2$ and 1 ha in extent.
+Very few observations' had spatial extents $>$1,000,000 ha, and because these were collected by satellites that make near-instantaneous measurements, they tended to have aggregate durations of $<$1 hour.
+In contrast, the longest periods of observation were provided by paleo-ecological or long-term weather stations, both of which produce continuous, point-scale measurements (i.e.
+soil cores and weather instrumentation).
 
-Comparing the two spatial scales of measurements against one another shows that observational extent scales log-linearly with resolution (Fig.
-1C), revealing a nearly 1:1 correspondence up to 100 ha of both resolution and extent.
-This relationship indicates that the larger the plot size, the fewer the number of replicates, which is a characteristic of field-based studies because effort and cost typically increases with plot size \cite{kareiva_spatial_1988}.
-In our sample, 85\% of the observations were collected using manual, field-based methods.
-These stand in contrast to remotely sensed observations, which made up just 4\% of the sample, but which are characterized by plot resolutions that are very high relative to their spatial extent--these are the low density patches above the 1:1 line defined by the denser concentration of field studies, namely the patches at the following resolution/extent domains; 1-10 m$^2$/100-10,000 ha; 0.1-10 ha/10 million ha; 10-10,000 ha/100,000-10 billion ha.
+Comparing the two spatial dimensions against one another shows that observational extent scales log-linearly with resolution (Fig.
+2C).
+Up to 100 ha, there is a nearly 1:1 correspondence between resolution and extent that results from the tendency for spatial replicates to decline with plot size (and thus effort) in field-based studies \cite{kareiva_spatial_1988}, which accounted for 80\% of the observations in our study.
+These stand in contrast to most remotely sensed observations, which comprised just 6.5\% of observations, and are characterized by spatial resolutions that are very fine relative to their extent.
+These constitute the majority of the low density patches (at 0.1-10 ha resolution and 1-10 million ha extent; 100-10,000 ha resolution and 1 million-10 billion ha extent) above the line defined by the denser concentration of field studies in Figure 2C.
 
-Comparing sampling interval to temporal duration shows that the bulk of once-off observations last between a minute to a week in duration (Fig.
+A similar linear relationship between sampling interval and temporal duration was not evident.
+For repeated observations, the greatest concentration was among those made at daily to yearly intervals with overall durations of hours to weeks (Fig.
+2D).
+Very few repeat observations exceeded a year in duration, and these were collected at daily to yearly intervals.
+The majority of unreplicated observations lasted between one minute and one day.
+
+This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI).
+Observations that were non-continuous in time or space (e.g.
+point-based field measurements) may effectively represent larger spatial scales than our estimates suggest, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997, palmer_scale_1994,cao_comparison_2002, legendre_spatial_1993,collins_method_2000-1}.
+This concern does not apply to spatially continuous observations, thus the net effect is that if we made estimates of effective spatial scale, the spatial resolution and extent estimates would be compressed from the lower end of their ranges.
+From a temporal perspective, our analysis likely underestimates the effective duration of many observations, particularly those where instantaneous, repeated measures of slow-changing ecological features were made.
+Snapshots in time may be sufficient to capture the temporal dynamics of such phenomena (e.g.
+changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these the total time elapsed between the first and last measurements (the span of observation) may more closely approximate effective duration.
+However, many studies focused on more dynamic phenomena, and for these long periods of continuous observation may be more important for understanding dynamics than frequent repeats.
+For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to directly measure fire behavior \cite{clements_observing_2007}.
+For this reason, and to provide a consistent standard for estimating duration, we recorded actual duration.
+
+Adding to our motivation to record actual observational scales was the fact that most studies provided insufficient methodological detail to judge whether sampling schemes were suitable matched to and representative of the phenomena they were measuring.
+The observational scales were not clearly reported in many studies, such that we had to estimate spatial resolution and extent in 62\% and 68\% of cases, respectively, and temporal interval and duration in 40\% and 86\% of cases.
+Although sensitivity analyses show that our general findings are robust to this uncertainty (SI), it would have likely confounded efforts to estimate effective scales.
+
+Despite these caveats, our results provide valuable insight into the spatial and temporal domains being addressed by modern ecological research.
+Our findings suggest that most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month).
+Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, suggesting that, despite well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989-1}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989-1}.
+Furthermore, the unclear documenting of observational scales suggests that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
+Taken together, the narrowness of scales and the ambiguity of their documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) suggests that much of the knowledge produced in ecology may not be generalizable \cite{margulies_ambiguous_2016,wheatley_factors_2009,wiens_spatial_1989-1}.
+
+The generally small spatial scales of observation is a consequence of the continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
+Despite early and repeated calls for ecologists to use remote sensing because it provides a synoptic view that field measurements cannot \cite{turner_remote_2003, kerr_space_2003, pettorelli_satellite_2014}, and subsequent demonstration of its importance for multi-scale studies \cite{estes_habitat_2008, estes_predictive_2011}, our results indicate this observational method is still under-utilized in ecology.
+In contrast, ecologists have more widely adopted methods that increase the temporal resolution of observations.
+Automated sensors were used to record 12\% of all observations, and accounted for most very high frequency measurements (intervals $\leq$1 hour).
+As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses (and many of the reviewed studies that used automated sensing aggregated the resulting observations before analyzing them), whereas it is extremely difficult, if not impossible, to disaggregate long-interval data.
+
+In the coming years, rapid technological advances should increase the concentration of ecological observations in currently under-represented domains.
+The growing numbers of high-resolution satellite sensors, together with new analytical platforms that provide free access to large volumes of pre-processed data and computational power \cite{googleearthengine}, will lower technical barriers that have so far prevented ecologists from adopting this observational technology  \cite{pettorelli_satellite_2014}.
+Similarly, the advent of unmanned systems offers the ability to measure ecological features at high spatial and temporal frequencies over large areas \cite{anderson_lightweight_2013}, which were scales that were previously impractical to access.
+The ever-falling cost of sensor technology and the ubiquity of cell phone connections also means that ecologists have a growing ability to make spatially dense, high frequency observations \cite{wolf_gsm-based_2012,collins_new_2006,porter_wireless_2005}.
+A greater attentiveness to scale in general, including more meticulous documentation of observed dimensions, may help to facilitate the spread of ecological research to sparsely studied scales, while improving transferability of knowledge within the discipline.
+
+%Plot resolution issue: need high resolution over large extents, because this relies on remote sensing, and to really capture details, need high resolution.  
+%likely because of perceived (and in some cases real) logistical barriers \cite{pettorelli_satellite_2014}.
+
+%These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989-1}.
+
+%Despite the understanding that ecological patterns and process change with scale \cite{levin_problem_1992}, the narrow distribution of observations within spatial domains calls into question the degree to which ecological phenomena are being investigated at multiple scales.
+Adding to this concern is the fact the scales of observation were not clearly reported in the majority of studies; the spatial resolution and extent of studies was uncertain in 64\% and 70\% of cases, respectively, while the precise interval and study duration were uncertain in 43\% and 86\% of cases.
+Although sensitivity analyses show that our conclusions are robust to this uncertainty (SI), the unclear documenting of observational scales suggest that they are not a primary methodological concern in ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
+%
+%On the other hand, we found a greater diversity of scales in the temporal domains of observations (Fig.
 1D).
-For repeated observations, the greatest density of observations were those made at monthly to yearly intervals and an aggregate duration of hours to weeks.
-Very few repeat observations exceeded a total duration measured in years, and those that did were of daily to yearly frequency.
+Over 70\% of sampled observations involved repeated assessments, and one-quarter of these were made with high frequency (intervals $\leq$1 day).
+Automated, surface-based sensors were used to record most of the very high frequency observations (intervals $\leq$1 hour), and comprised 9\% of all observations, suggesting that ecologists are adopting new technologies for collecting temporal data more rapidly than they are for spatial data.
+As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses, whereas it is extremely difficult, if not impossible, to disaggregate long-interval data.
+%
+%The temporal duration of most collected observations appears to be brief ($\leq$1year), which is due to the fact that we estimated the summed time of repeated observations, rather than the total time elapsed between the first and last measurements (the span of observation).
+In many cases, the actual observations are very brief, and thus the temporal duration is narrow.
+Snapshots in time may be sufficient to capture the temporal dynamics of many phenomena (e.g.
+changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these our measurement of this domain may understate the temporal breadth of ecological observations.
+However, for more dynamic phenomena, long periods of continuous observation may be more important for understanding dynamics than frequent repeats.
+For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe behavior and evolution of the active fire \cite{clements_observing_2007}.
+For this reason, and because duration is the temporal analog of spatial extent, we chose to record observational duration rather then span.
 
-Reflect nature of observation approaches. \\
-Use of remote sensing still relatively rare in ecological studies--thus the synoptic view of ecosystems not being seen.  \\
-High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled. \\
-Don't forget to mention automated approaches. 
-Limitation for ocean studies, we did not study third dimension, volume (might be more relevant).
-Notes: Wheatley and Johnson (2009) citing arbitrariness of scales in wildlife studies (cited by Chave).
+%Based on the observational domains recorded here, it appears that ecologists are still primarily collecting data at very fine spatial scales and over very narrow extents.
+Although the temporal duration of most ecological observations is also narrow, and a sizable number of observations are not repeated, greater effort is being devoted to capturing temporal dynamics.
+These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989-1}.
 
-Although limited sample size, insight into the scales at which science of ecology is making bulk of observations
-Despite recognized importance of making observation across scales \cite{levin_problem_1992}, this does not appear to be happening. 
-Technical capabilities for making observations are not achieving adoption, despite potential of remote sensing to provide multi-scale ecological data \
+%Based on this analysis, ecology thus appears to have
+
+%High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled. \\
+%Limitation for ocean studies, we did not study third dimension, volume (might be more relevant).
+
+%Although limited sample size, insight into the scales at which science of ecology is making bulk of observations
 
@@
 
-%\begin{scilastnote}
-%\item We've included in the template file \texttt{scifile.tex} a new
-%environment, \texttt{\{scilastnote\}}, that generates a numbered final
-%citation without a corresponding signal in the text.  This environment
-%can be used to generate a final numbered reference containing
-%acknowledgments, sources of funding, and the like, per {\it Science\/}
-%style.
-%\end{scilastnote}
-
 % For your review copy (i.e., the file you initially send in for
@@
 
+\begin{scilastnote}
+ \item This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G),
+\end{scilastnote}
+
 \clearpage

### 2016-07-07 Rmd/space-time.Rmd
Commit of all changes (analysis, text) up to 6 July, 2016 (just before incorporating Erle's edits).

@@
 full <- gsub("\\s", "", full)  # remove DOI and whitespace
+unique(full)
 
@@
 datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
-
 doisup <- c(unique(cal$DOI_data_source), unique(dat$DOI_data_source))
@@
 # rm(bootperturb2)
-
-```
-
-<a href="#top">Back to top</a>
-
-```{r, eval = FALSE, echo = FALSE}
-### Bootstrapping
-# Good old-fashioned bootstrapping with replacement, no perturbation of value
-# bootstrap
-set.seed(1)
-bootsamp <- lapply(1:1000, function(x) {
-  if((x / 100) %in% 1:100) print(x)
-  dnew <- copy(datf[, .(plot_res, sampled_area, t_btwn_samp, study_duration)])
-  dnew <- dnew[sample(1:.N, size = .N, replace = TRUE), ]
-  dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
-  dnew
-})
 
@@
 
-### Perturbed set
+### Figure 1
+
+```{r, eval=FALSE}
+length(which(datf$study_type %in% 
+               c("remote sensing", "other geographic data"))) / nrow(datf)
+datf[t_btwn_samp == 0, .N] / datf[, .N]  # 36% of samples are once-offs
+datf[study_duration == samp_duration, .N] / nrow(datf)
+datf[(study_duration == samp_duration) & t_btwn_samp != 0]
+datf[(study_duration != samp_duration) & t_btwn_samp == 0]
+
+hdat <- cbind("res" = log10(datf$plot_res), "ext" = log10(datf$sampled_area), 
+              "int" = log10(to_infin_byond(datf$t_btwn_samp)), 
+              "dur" = log10(datf$study_duration))
+hdat <- data.table(hdat)
+# set bounds for histograms
+hdat[res < aaxis1$logres[1], res := aaxis1$logres[1]]
+hdat[res > ry[2], res := ry[2]]
+hdat[ext < -6, ext := -6]  # set minimum extent -6 (0.01 m2)
+
+# histograms from bootstrap
+brksl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
+              taxis2$logdays)
+
+vs <- c("res", "ext", "int", "dur")
+hdat_bs <- lapply(1:length(bootperturb), function(x) {  # x <- 1
+  print(x)
+  DT <- bootperturb[[x]]
+  hdb <- cbind("res" = log10(DT$plot_res), "ext" = log10(DT$sampled_area), 
+               "int" = log10(to_infin_byond(DT$t_btwn_samp)), 
+               "dur" = log10(DT$study_duration))
+  hdb <- data.table(hdb)
+  hdb[res < aaxis1$logres[1], res := aaxis1$logres[1]]
+  hdb[res > ry[2], res := ry[2]]
+  hdb[, lapply(.SD, range)]
+  hdb[ext < -6, ext := -6]  # set minimum extent -6 (0.01 m2)
+  hists <- lapply(1:length(vs), function(y) {  # y <- 2
+    #print(y)
+    hdbv <- hdb[[vs[y]]]
+    hbrks <- brksl[[y]]
+    hdbv[hdbv < hbrks[1]] <- hbrks[1]
+    hdbv[hdbv > hbrks[length(hbrks)]] <- hbrks[length(hbrks)]
+    h <- hist(hdbv, breaks = hbrks, plot = FALSE)
+    h$density <- h$counts / sum(h$counts) * 100
+    h
+  })
+})
+
+hdat_bsdt <- lapply(1:length(vs), function(x) { # x <- 1
+  mat <- do.call(rbind, lapply(1:length(hdat_bs), function(y) {
+    hdat_bs[[y]][[x]]$density
+  }))
+  DT <- data.table(mat)
+})
+
+# hdat_bsdt <- lapply(1:length(vs), function(x) { # x <- 1
+#   mat <- do.call(rbind, lapply(1:length(hdat_bs), function(y) {
+#     hdat_bs[[y]][[x]]
+#   }))
+#   data.table(mat)
+# })
+
+# setnames(hdat_bsdt, names(hdat_bsdt), as.character(aaxis1$logres))
+
+# hist(hdat$res, breaks = aaxis1$logres)
+# hist(log10(datf$plot_res), breaks = resbrks)
+# hist(log10(datf$plot_res), breaks = resbrks)
+
+## Ordinary histograms
+mga <- -0.34
+sfigl <- -1.2
+cxa <- 0.8
+cxl <- 0.9
+reds <- brewer.pal(9, name = "Reds")
+blues <- brewer.pal(9, name = "Blues")
+plot(1:2, pch = 20, col = blues[c(4, 6)])
+yl <- 35
+# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
+pdf("paper/figures/hists3.pdf", width = 5, height = 5)
+par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
+# resolution
+h <- hist(hdat$res, breaks = aaxis1$logres, plot = FALSE)
+h$density <- h$counts / sum(h$counts) * 100
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
+     tcl = -0.2, cex.axis = cxa, xlab = "Spatial resolution", 
+     ylim = c(0, yl - 10), main = "", col = reds[4], cex.lab = cxl)
+axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
+     tcl = -0.2, mgp = c(2, 0.25, mga))
+mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
+mtext("A", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
+
+# extent
+h <- hist(hdat$ext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
+h$density <- h$counts / sum(h$counts) * 100
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
+     tcl = -0.2, cex.axis = cxa, xlab = "Spatial extent", ylim = c(0, yl - 10), 
+     main = "", col = reds[6], cex.lab = cxl)
+axis(1, at = c(-6, aaxis2$logarea), 
+     labels = c(expression(paste(NULL<="0.01m"^2)), expression(paste(0.1, m^2)),                 alab2[-1]), las = 2, cex.axis = cxa,
+     tcl = -0.2, mgp = c(2, 0.25, mga))
+mtext("B", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
+
+# interval
+tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
+h <- hist(hdat$int, breaks = taxis1$logdays, plot = FALSE)
+h$density <- h$counts / sum(h$counts) * 100
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
+     tcl = -0.2, cex.axis = cxa, xlab = "Sampling interval", ylim = c(0, yl), 
+     main = "", col = blues[4], cex.lab = cxl)
+axis(1, at = tax, labels = tlab1[-10], las = 2, cex.axis = cxa, 
+     tcl = -0.2, mgp = c(2, 0.25, -0.34))
+mtext("C", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
+mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
+
+# duration
+# tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
+h <- hist(hdat$dur, breaks = taxis2$logdays, plot = FALSE)
+h$density <- h$counts / sum(h$counts) * 100
+plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
+     tcl = -0.2, cex.axis = cxa, xlab = "Temporal duration", ylim = c(0, yl), 
+     main = "", col = blues[6], cex.lab = cxl)
+axis(1, at = taxis2$logdays, labels = tlab2, las = 2, cex.axis = cxa, 
+     tcl = -0.2, mgp = c(2, 0.25, mga))
+mtext("D", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
+
+dev.off()
+
+### Bootstrapped histograms (now main figure)
+qtf <- function(x) c(mean(x), quantile(x, probs = c(0.025, 0.975))) # quant f
+xlabs <- c("Spatial resolution", "Spatial extent", "Sampling interval", 
+           "Temporal duration")
+# axes <- list("aax1" = aaxis1$logres, "aax2" = c(-6, aaxis2$logarea), 
+#              "tax1" = taxis1$logdays, "tax2" = taxis2$logdays)
+axes2 <- list("aax1" = aaxis1$logres, "aax2" = c(-6, aaxis2$logarea), 
+              "tax1" = c(taxis2$logdays[-c(10:11)],
+                         mean(taxis2$logdays[c(10:11)])), 
+              "tax2" = taxis2$logdays)
+axlabs <- list(alab1, c(expression(paste(NULL<="0.01m"^2)),
+                        expression(paste(0.1, m^2)), alab2[-1]),
+               tlab1[-10], tlab2)
+mga <- -0.34
+sfigl <- -1.2
+cxa <- 0.8
+cxl <- 0.9
+reds <- brewer.pal(9, name = "Reds")
+blues <- brewer.pal(9, name = "Blues")
+# plot(1:2, pch = 20, col = blues[c(4, 6)])
+yl <- 35
+cols <- c(reds[c(4, 6)], blues[c(4, 6)])
+
+# calculate mean, 2.5th, 97.5th
+hdatstat_bt <- lapply(1:length(hdat_bsdt), function(x) {
+  h <- hdat_bsdt[[x]][, lapply(.SD, qtf)]
+  hmu <- unlist(h[1, ])
+  h2 <- unlist(h[2, ])
+  h98 <- unlist(h[3, ])
+  list("mu" = hmu, "p2" = h2, "p98" = h98)
+})
+names(hdatstat_bt) <- vs
+
+# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
+pdf("paper/figures/hists_bs.pdf", width = 5, height = 5)
+par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
+
+for(i in 1:length(hdatstat_bt)) {  # i <- 1
+  hvals <- hdatstat_bt[[i]]
+  axv <- brksl[[i]]
+  mids <- sapply(2:length(axv), function(x) mean(c(axv[x], axv[x - 1])))
+
+    # base plot
+  yl <- ifelse(i %in% c(1, 2), 30, 35) 
+  plot(c(axv[1], axv[length(axv)]), c(0, yl - 10), ylab = "", las = 2, 
+         xaxt = "n", xlab = xlabs[i], ylim = c(0, yl), main = "", 
+         col = cols[i], cex.lab = cxl, pch = "", bty = "l", axes = FALSE)
+
+  # create polygons to mimic histograms of varying width
+  sapply(2:length(axv), function(x) {  # x <- 2
+    px <- axv[c(x - 1, x)]
+    pxs <- c(rep(px[1], 2), rep(px[2], 2), px[1])
+    pys <- c(0, rep(hvals$mu[[x - 1]], 2), 0, 0)
+    polygon(pxs, pys, col = cols[i])
+  })
+  axis(1, at = axes2[[i]], labels = axlabs[[i]], las = 2, 
+       cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.25, mga))
+  axis(2, las = 2, cex.axis = cxa, tcl = -0.2, mgp = c(2, 0.25, mga))
+  if(i %in% c(1, 3)) {
+    mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
+  }
+  mtext(LETTERS[i], side = 3, line = sfigl, cex = 0.8, adj = 0.05)
+
+  # confidence intervals
+  points(mids, hvals$p2, pch = "-", cex = 1, col = "grey40")
+  points(mids, hvals$p98, pch = "-", cex = 1, col = "grey40")
+
+}
+dev.off()
+
+# stats
+# bootp_dt
+# axisl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
+#               taxis2$logdays)
+hdatstat <- lapply(1:4, function(x) {
+  h <- hist(hdat[[x]], breaks = brksl[[x]], plot = FALSE)
+  h$density <- h$counts / sum(h$counts) * 100
+  h
+})
+
+# res
+# hdatst <- hdatstat[[1]]$density
+hdatst <- hdatstat_bt$res$mu
+l <- length(hdatst)
+sum(hdatst[4:6])  # 53 % between 10cm^2 and 1 m^2
+sum(hdatst[1:6])  #  67% < 1m^2
+sum(hdatst[7:10]) # 24 % between 1 m^2 and 1 ha
+sum(hdatst[11:(l)]) # 9 % 1-10000 ha
+# dat[plot_res >= 10000^2]
+
+# length(which(hdat$res >= -3 & hdat$res <= 0)) / length(hdat$res)
+# length(which(hdat$res <= 4)) / length(hdat$res)
+# sum(hdatstat[[1]]$counts[4:6]) / sum(hdatstat[[1]]$counts)
+
+# extent
+# hdatst <- hdatstat[[2]]$density
+hdatst <- hdatstat_bt$ext$mu
+l <- length(hdatst)
+sum(hdatst[1:2])  # 36 % < 1 m^2
+sum(hdatst[1:6])  # 82% < 1 ha
+sum(hdatst[1:7])  # 86% < 10 ha
+sum(hdatst[1:8])  # 90 % < 100 ha
+sum(hdatst[1:9])  # 92 % < 1000 ha
+sum(hdatst[8:9])  # 6 % 10-1000 ha
+sum(hdatst[8:l])  # 14% > 10 ha
+sum(hdatst[10:l]) # 8.2 % covered an area > 1000 ha
+sum(hdatst[12:l]) # 5.1 % covered an area > 100,000 ha
+sum(hdatst[13:l]) # 3.9 % covered an area > 1,000,000 ha
+
+# length(which(hdat$ext < -4)) / length(hdat$ext)
+# length(which(hdat$ext > 3)) / length(hdat$ext)
+# length(which(hdat$ext > 1 & hdat$ext <= 3)) / length(hdat$ext)
+# which(hdatstat[[2]]$breaks == log10(1000000))
+# length(hdatstat[[2]]$counts[7:8])
+# hdatstat[[2]]$mids[12]
+
+# interval
+# hdatst <- hdatstat[[3]]$density
+hdatst <- hdatstat_bt$int$mu
+l <- length(hdatst)
+sum(hdatst[l])  # 36% are once-offs
+sum(hdatst[1:3])  # 15 % < daily
+sum(hdatst[1:4])  # 25 % < weekly
+sum(hdatst[4:5])  # 21% daily up to monthly
+sum(hdatst[6])  # 19% monthly up to yearly
+sum(hdatst[7:8]) # 8.2 % yearly to decadal
+
+# duration
+# hdatst <- hdatstat[[4]]$density
+hdatst <- hdatstat_bt$dur$mu
+l <- length(hdatst)
+sum(hdatst[1:3])  # 71 % up to 1 day
+sum(hdatst[4:5])  # 17% covered between 1 day and 1 month
+sum(hdatst[6])  # 7.3% covered between 1 month and 1 year
+# sum(hdatstat[[4]]$density[4:6])  # 26% covered between 1 day and 1 year
+sum(hdatst[7:10]) # 4.5% year to century
+
+# datf[, plot(log10(n_sites), log10(plot_res))]
+# dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
+datf[like(study_type, "field"), plot(log10(plot_res), log10(sampled_area), 
+                                     ylim = c(-10, 10))]
+
+datf[like(study_type, "remote"), 
+     points(log10(plot_res), log10(sampled_area), pch = 20, col = "red")]
+datf[like(study_type, "automated"), 
+     points(log10(plot_res), log10(sampled_area), pch = 20, col = "blue")]
+lines(aaxis1$logres, aaxis2$logarea[-16] )
+datf[, .N / nrow(datf) * 100, by = st]
+
+datf[, plot(log10(t_btwn_samp), log10(study_duration), pch = 20)]
+datf[which(hdat[[3]] < -1.380211), 
+     points(log10(t_btwn_samp), log10(study_duration), pch = 20, col = "blue")]
+datf[like(study_type, "automated"), 
+     points(log10(t_btwn_samp), log10(study_duration), pch = 20, col = "red")]
+
+# points(resdat3$x[which(datf$st == 4)], resdat3$y[which(datf$st == 4)])
+# points(resdat3$x[which(datf$st == 1)], resdat3$y[which(datf$st == 1)])
+
+# types of observations
+# unique(datf$study_type)
+datf[like(study_type, "field"), .N] / nrow(datf)  # 80% field studies
+datf[like(study_type, "auto"), .N] / nrow(datf) # 12 % automated
+datf[like(study_type, "auto") & log10(t_btwn_samp) < -1.380211, .N] /   length(which(hdat[[3]] < -1.380211))  # 75%
+
+datf[like(study_type, "remote|other"), .N] / nrow(datf) # 6 % automated
+datf[like(study_type, "paleo"), .N] / nrow(datf) * 100 # <1 % paleo
+
+```
+
+<a href="#top">Back to top</a>
+
+### Figure 2 
 ```{r, eval = FALSE}
 bootp_dt <- rbindlist(bootperturb)
-resdat <- data.table(x = log10(bootp_dt$t_btwn_samp), 
-                      y = log10(bootp_dt$plot_res))
-resdat$n <- rep(1, nrow(resdat))
-rx <- range(resdat$x)
-rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-rx[1] <- -5
-rx[2] <- 6.56
-resdat[x < rx[1], x := rx[1]]
-resdat[x > rx[2], x := rx[2]]
-# rx[2] <- 7.6
-ry <- range(aaxis1$logres)
-ry[1] <- -5  # set limits to 0.01 cm^2
-resdat[y < ry[1], y := ry[1]]
-resdat[y > ry[2], y := ry[2]]
-coordinates(resdat) <- ~x + y
-
-# extent
-extdat <- data.table(x = log10(bootp_dt$study_duration), 
-                     y = log10(bootp_dt$sampled_area))
-extdat$n <- rep(1, nrow(extdat))
-ex <- range(taxis2$logdays)
-ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
-extdat[x < ex[1], x := ex[1]]
-extdat[x > ex[2], x := ex[2]]
-# ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
-# ex[2] <- 7.6
-ey <- range(aaxis2$logarea)
-extdat[y < ey[1], y := ey[1]]
-extdat[y > ey[2], y := ey[2]]
-coordinates(extdat) <- ~x + y
-
-# densities
-# resolution
-rres <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat, 1)
-rres <- (rres / cellStats(rres, sum)) * 100
-rng <- range(rres[is.finite(rres)])
-# rng[2] <- 0.1
-bwidth <- (rng[2] - rng[1]) / 30
-brks <- seq(rng[1], rng[2], bwidth)
-brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
-
-# extent
-rext <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat, 1)
-rext <- (rext / cellStats(rext, sum)) * 100
-rng <- range(rext[is.finite(rext)])
-rng[2] <- 0.1
-bwidth <- (rng[2] - rng[1]) / 30
-brks2 <- seq(rng[1], rng[2], bwidth)
-brklabs2 <- round(brks2[seq(1, length(brks2), 6)], 2)
-
-# cols <- colorRampPalette(c("blue", "gold", "red"))
-# cols <- c(rev(terrain.colors(30))[1], cols(29))
-
-brkfun <- function(ext, ival, n) {
-  rng <- range(ext[is.finite(ext)])
-  bwidth <- (rng[2] - rng[1]) / n
-  brks <- seq(rng[1], rng[2], bwidth)
-  brklabs <- seq(0, round(rng[2], 2), ival)
-  list("brks" = brks, "labs" = brklabs)
-}
-cuts <- 80
-
-cxa = 0.7
-pdf("paper/figures/res_v_extent_pbstrap2.pdf", width = 7, 
-    height = 2.5)
-par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
-    mgp = c(2, 0.5, 0))
-# image(rres, col = cols, breaks = brks,
-brks <- brkfun(rres, 0.01, cuts)  # breaks
-image(rres, col = rev(terrain.colors(cuts)), #breaks = brks,
-     axes = FALSE, xlab = "", ylab = "")
-# plot(rres2, col = rev(terrain.colors(20)))
-axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
-     las = 2, tcl = -0.2, cex.axis = cxa)
-axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
-     cex.axis = cxa)
-# plot(rres2, axes = FALSE, legend.only = TRUE)
-aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
-              cex.axis = cxa, tcl = -0.1)
-plot(rres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
-brks <- brkfun(rext, 0.01, cuts)  # breaks
-image(rext, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
-      ylab = "")
-axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
-     cex.axis = cxa, tcl = -0.2)
-axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
-     tcl = -0.2)
-aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
-              cex.axis = cxa, tcl = -0.1)
-plot(rext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
-dev.off()
-```
-
-<a href="#top">Back to top</a>
-
-```{r, eval = FALSE, echo = FALSE}
-### Bootstrap - deprecated for now
-
-bootdt <- rbindlist(bootsamp)
-resdat2 <- data.table(x = log10(bootdt$t_btwn_samp), y = log10(bootdt$plot_res))
-resdat2$n <- rep(1, nrow(resdat2))
-rx <- range(resdat2$x)
-rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-rx[2] <- 6.56
-resdat2[x < rx[1], x := rx[1]]
-resdat2[x > rx[2], x := rx[2]]
-ry <- range(aaxis1$logres)
-ry[1] <- -5  # set limits to 0.01 cm^2
-resdat2[y < ry[1], y := ry[1]]
-resdat2[y > ry[2], y := ry[2]]
-coordinates(resdat2) <- ~x + y
-
-# extent
-extdat2 <- data.table(x = log10(bootdt$study_duration), 
-                      y = log10(bootdt$sampled_area))
-extdat2$n <- rep(1, nrow(extdat2))
-ex <- range(taxis2$logdays)
-ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
-extdat2[x < ex[1], x := ex[1]]
-extdat2[x > ex[2], x := ex[2]]
-ey <- range(aaxis2$logarea)
-extdat2[y < ey[1], y := ey[1]]
-extdat2[y > ey[2], y := ey[2]]
-coordinates(extdat2) <- ~x + y
-
-# densities
-# resolution
-rres2 <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat2, 1)
-rres2 <- (rres2 / cellStats(rres2, sum)) * 100
-rng <- range(rres2[is.finite(rres2)])
-bwidth <- (rng[2] - rng[1]) / 30
-brks <- seq(rng[1], rng[2], bwidth)
-brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
-
-# extent
-rext2 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat2, 1)
-rext2 <- (rext2 / cellStats(rext2, sum)) * 100
-rng <- range(rext2[is.finite(rext2)])
-# rng[2] <- 0.1
-bwidth <- (rng[2] - rng[1]) / 30
-brks2 <- seq(rng[1], rng[2], bwidth)
-brklabs2 <- brklabs #round(brks2[seq(1, length(brks2), 3)], 2)
-
-cxa = 0.7
-pdf("paper/figures/res_v_extent_bstrap2.pdf", width = 7, 
-    height = 2.5)
-par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
-    mgp = c(2, 0.5, 0))
-# image(rres, col = cols, breaks = brks,
-image(rres2, col = rev(terrain.colors(30)), breaks = brks,
-     axes = FALSE, xlab = "", ylab = "")
-# plot(rres2, col = rev(terrain.colors(20)))
-axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
-     las = 2, tcl = -0.2, cex.axis = cxa)
-axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
-     cex.axis = cxa)
-# plot(rres2, axes = FALSE, legend.only = TRUE)
-aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
-              cex.axis = cxa, tcl = -0.1)
-plot(rres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
-
-image(rext2, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
-      ylab = "", breaks = brks2)
-axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
-     cex.axis = cxa, tcl = -0.2)
-axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
-     tcl = -0.2)
-aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
-              cex.axis = cxa, tcl = -0.1)
-plot(rext2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
-dev.off()
-
-# datf[, plot(log10(study_duration), log10(sampled_area), xlim)]
-```
-
-<a href="#top">Back to top</a>
-
-### Primary dataset (main figure)
-```{r, eval = FALSE}
-resdat3 <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
-                      y = log10(datf$plot_res))
+
+# choose which dataset to use
+# kdat <- datf  # ordinary
+kdat <- bootp_dt  # bootstrap
+resdat3 <- data.table(x = log10(to_infin_byond(kdat$t_btwn_samp)), 
+                      y = log10(kdat$plot_res))
+resdat3 <- resdat3[round(x, 8) != 6.56229286, ]  # remove one-time observations
 resdat3$n <- rep(1, nrow(resdat3))
@@
 rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-rx[2] <- 6.56
+rx[1] <- -5
+#rx[2] <- 6.56
+rx[2] <- 5.57
 resdat3[x < rx[1], x := rx[1]]
@@
 # extent
-extdat3 <- data.table(x = log10(datf$study_duration), 
-                      y = log10(datf$sampled_area))
+extdat3 <- data.table(x = log10(kdat$study_duration), 
+                      y = log10(kdat$sampled_area))
 extdat3$n <- rep(1, nrow(extdat3))
@@
 # temporal framing
-tempdat <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
-                      y = log10(datf$study_duration))
+tempdat <- data.table(x = log10(to_infin_byond(kdat$t_btwn_samp)), 
+                      y = log10(kdat$study_duration))
+tempdat <- tempdat[round(x, 8) != 6.56229286, ]  # remove one-time observations
 tempdat$n <- rep(1, nrow(tempdat))
@@
 tx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-tx[2] <- 6.56
+#tx[2] <- 6.56
+tx[2] <- 5.57
 tempdat[x < rx[1], x := rx[1]]
@@
 # spatial framing
-spatdat <- data.table(x = log10(datf$plot_res), y = log10(datf$sampled_area))
+spatdat <- data.table(x = log10(kdat$plot_res), y = log10(kdat$sampled_area))
 spatdat$n <- rep(1, nrow(spatdat))
@@
 cxa = 0.8
-stps <- sapply(c(3, 4), function(x) which(datf$st == x))  # IDs rs/auto/paleo
+stps <- sapply(c(3, 4), function(x) which(kdat$st == x))  # IDs rs/auto/paleo
 pchs <- list(1, "+")
 cexs <- c(0.4, 0.6)
-cuts <- 80
+cuts <- 60
 
 # pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
-pdf("paper/figures/kde43.pdf", width = 7, height = 5.5)
+pdf("paper/figures/kde45.pdf", width = 7, height = 5.5)
 par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
@@
       ylab = "")
-axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
+axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
+# axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
      las = 2, tcl = -0.2, cex.axis = cxa)
@@
 # points(spatdat$x, spatdat$y, cex = 0.1)
-axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
+axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
+# axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
      las = 2, tcl = -0.2, cex.axis = cxa)
@@
 # points(resdat3$x, resdat3$y, pch = 20, cex = 0.5)
-image(sres, col = rev(terrain.colors(100)))
-points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)
-```
-
-<a href="#top">Back to top</a>
-
-## Stats and supplementary plots
-
-### Statistics
-```{r, eval=FALSE}
-length(which(dat$study_type %in% 
-               c("remote sensing", "other geographic data"))) / nrow(dat)
-length(which(dat$study_type %in% 
-               c("remote sensing", "other geographic data"))) / nrow(dat)
-dat[t_btwn_samp == 0, .N] / dat[, .N]  # 35% of samples are once-offs
-dat[study_duration == samp_duration, .N] / nrow(dat)
-dat[(study_duration == samp_duration) & t_btwn_samp != 0]
-dat[(study_duration != samp_duration) & t_btwn_samp == 0]
-
-hdat <- cbind("res" = log10(datf$plot_res), "ext" = log10(datf$sampled_area), 
-              "int" = log10(to_infin_byond(datf$t_btwn_samp)), 
-              "dur" = log10(datf$study_duration))
-hdat <- data.table(hdat)
-# set bounds for histograms
-hdat[res < aaxis1$logres[1], res := aaxis1$logres[1]]
-hdat[res > ry[2], res := ry[2]]
-hdat[ext < -6, ext := -6]  # set minimum extent -6 (0.01 m2)
-
-# hist(hdat$res, breaks = aaxis1$logres)
-# hist(log10(datf$plot_res), breaks = resbrks)
-# hist(log10(datf$plot_res), breaks = resbrks)
-
-mga <- -0.34
-sfigl <- -1.2
-cxa <- 0.8
-cxl <- 0.9
-reds <- brewer.pal(9, name = "Reds")
-blues <- brewer.pal(9, name = "Blues")
-plot(1:2, pch = 20, col = blues[c(4, 6)])
-yl <- 35
-# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
-pdf("paper/figures/hists2.pdf", width = 5, height = 5)
-par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
-# resolution
-h <- hist(hdat$res, breaks = aaxis1$logres, plot = FALSE)
-h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Spatial resolution", 
-     ylim = c(0, yl - 10), main = "", col = reds[4], cex.lab = cxl)
-axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
-     tcl = -0.2, mgp = c(2, 0.25, mga))
-mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
-mtext("A", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
-
-# extent
-h <- hist(hdat$ext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
-h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Spatial extent", ylim = c(0, yl - 10), 
-     main = "", col = reds[6], cex.lab = cxl)
-axis(1, at = c(-6, aaxis2$logarea), 
-     labels = c(expression(paste(NULL<="0.01m"^2)), expression(paste(0.1, m^2)), 
-                alab2[-1]), las = 2, cex.axis = cxa,
-     tcl = -0.2, mgp = c(2, 0.25, mga))
-mtext("B", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
-
-# interval
-tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
-h <- hist(hdat$int, breaks = taxis1$logdays, plot = FALSE)
-h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Sampling interval", ylim = c(0, yl), 
-     main = "", col = blues[4], cex.lab = cxl)
-axis(1, at = tax, labels = tlab1[-10], las = 2, cex.axis = cxa, 
-     tcl = -0.2, mgp = c(2, 0.25, -0.34))
-mtext("C", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
-mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
-
-# duration
-# tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
-h <- hist(hdat$dur, breaks = taxis2$logdays, plot = FALSE)
-h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Temporal duration", ylim = c(0, yl), 
-     main = "", col = blues[6], cex.lab = cxl)
-axis(1, at = taxis2$logdays, labels = tlab2, las = 2, cex.axis = cxa, 
-     tcl = -0.2, mgp = c(2, 0.25, mga))
-mtext("D", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
-
-dev.off()
-
-# stats
-axisl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
-              taxis2$logdays)
-hdatstat <- lapply(1:4, function(x) {
-  h <- hist(hdat[[x]], breaks = axisl[[x]], plot = FALSE)
-  h$density <- h$counts / sum(h$counts) * 100
-  h
-})
-
-# res
-l <- length(hdatstat[[1]]$density)
-sum(hdatstat[[1]]$density[4:6])  # 55 % between 10cm^2 and 1 m^2
-sum(hdatstat[[1]]$density[1:6])  #  69% < 1m^2
-sum(hdatstat[[1]]$density[7:10]) # 22 % between 1 m^2 and 1 ha
-sum(hdatstat[[1]]$density[11:(l)]) # 9 % 1-10000 ha
-# dat[plot_res >= 10000^2]
-
-# length(which(hdat$res >= -3 & hdat$res <= 0)) / length(hdat$res)
-# length(which(hdat$res <= 4)) / length(hdat$res)
-# sum(hdatstat[[1]]$counts[4:6]) / sum(hdatstat[[1]]$counts)
-
-# extent
-l <- length(hdatstat[[2]]$density)
-sum(hdatstat[[2]]$density[1:2])  # 35 % < 1 m^2
-sum(hdatstat[[2]]$density[1:6])  # 82% < 1 ha
-sum(hdatstat[[2]]$density[1:7])  # 86% < 10 ha
-sum(hdatstat[[2]]$density[1:8])  # 89 % < 100 ha
-sum(hdatstat[[2]]$density[1:9])  # 92 % < 1000 ha
-sum(hdatstat[[2]]$density[8:9])  # 6 % 10-1000 ha
-sum(hdatstat[[2]]$density[8:l])  # 14% > 10 ha
-sum(hdatstat[[2]]$density[10:l]) # 8.4 % covered an area > 1000 ha
-sum(hdatstat[[2]]$density[12:l]) # 5.1 % covered an area > 100,000 ha
-sum(hdatstat[[2]]$density[13:l]) # 4.1 % covered an area > 1,000,000 ha
-
-# length(which(hdat$ext < -4)) / length(hdat$ext)
-# length(which(hdat$ext > 3)) / length(hdat$ext)
-# length(which(hdat$ext > 1 & hdat$ext <= 3)) / length(hdat$ext)
-# which(hdatstat[[2]]$breaks == log10(1000000))
-# length(hdatstat[[2]]$counts[7:8])
-# hdatstat[[2]]$mids[12]
-
-# interval
-l <- length(hdatstat[[3]]$density)
-sum(hdatstat[[3]]$density[l])  # 36% are once-offs
-sum(hdatstat[[3]]$density[1:3])  # 17 % < daily
-sum(hdatstat[[3]]$density[1:4])  # 17 % < daily
-sum(hdatstat[[3]]$density[4:5])  # 19% daily up to monthly
-sum(hdatstat[[3]]$density[6])  # 23% monthly up to yearly
-sum(hdatstat[[3]]$density[7:9]) # 5 % yearly to decadal
-
-# duration
-l <- length(hdatstat[[4]]$density)
-sum(hdatstat[[4]]$density[1:3])  # 69 % up to 1 day
-sum(hdatstat[[4]]$density[4:5])  # 19% covered between 1 day and 1 month
-sum(hdatstat[[4]]$density[6])  # 8% covered between 1 month and 1 year
-# sum(hdatstat[[4]]$density[4:6])  # 26% covered between 1 day and 1 year
-sum(hdatstat[[4]]$density[7:10]) # 5% year to century
-
-# datf[, plot(log10(n_sites), log10(plot_res))]
-# dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
-datf[like(study_type, "field"), plot(log10(plot_res), log10(sampled_area), 
-                                     ylim = c(-10, 10))]
-
-datf[like(study_type, "remote"), 
-     points(log10(plot_res), log10(sampled_area), pch = 20, col = "red")]
-datf[like(study_type, "automated"), 
-     points(log10(plot_res), log10(sampled_area), pch = 20, col = "blue")]
-lines(aaxis1$logres, aaxis2$logarea[-16] )
-datf[, .N / nrow(datf) * 100, by = st]
-
-datf[, plot(log10(t_btwn_samp), log10(study_duration), pch = 20)]
-datf[which(hdat[[3]] < -1.380211), 
-     points(log10(t_btwn_samp), log10(study_duration), pch = 20, col = "blue")]
-datf[like(study_type, "automated"), 
-     points(log10(t_btwn_samp), log10(study_duration), pch = 20, col = "red")]
-
-# points(resdat3$x[which(datf$st == 4)], resdat3$y[which(datf$st == 4)])
-# points(resdat3$x[which(datf$st == 1)], resdat3$y[which(datf$st == 1)])
-
-# types of observations
-# unique(datf$study_type)
-datf[like(study_type, "field"), .N] / nrow(datf)  # 80% field studies
-datf[like(study_type, "auto"), .N] / nrow(datf) # 12 % automated
-datf[like(study_type, "auto") & log10(t_btwn_samp) < -1.380211, .N] /   length(which(hdat[[3]] < -1.380211))  # 75%
-
-datf[like(study_type, "remote|other"), .N] / nrow(datf) # 6 % automated
-datf[like(study_type, "paleo"), .N] / nrow(datf) * 100 # 1 % automated
-
+# image(sres, col = rev(terrain.colors(100)))
+# points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)
 ```
@@
 ```{r,eval = FALSE}
-# resolution data
-# resdat <- data.table(x = log10(datf$t_btwn_samp), y = log10(datf$plot_res))
-# resdat$n <- rep(1, nrow(resdat))
-# coordinates(resdat) <- ~x + y
-# rx <- range(resdat$x)
-# # rx <- range(aaxis$logarea)
-# rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-# # rx[1] <- -6.5  # set temporal interval to -6.5 to have null returns centered
-# # ry <- range(resdat$y)
-# ry <- range(aaxis$logarea)
-# ry <- round(ry / 0.5) * 0.5  # round to nearest 0.5
-# 
-# # extent data
-# # fix area rounded to zero in Jon's plots
-# extdat <- data.table("x" = log10(dat$study_duration), 
-#                      "y" = log10(dat$sampled_area)) 
-# # extdat[is.infinite(x), ]; extdat[is.infinite(y), ]
-# 
-# extdat$n <- rep(1, nrow(extdat))
-# coordinates(extdat) <- ~x + y
-# # ex <- range(extdat$x)
-# ex <- range(taxis2$logdays)
-# ex <- round(ex / 0.5) * 0.5  # round to nearest 0.5
-# # ex[1] <- -5
-# ey <- range(extdat$y)
-# ey <- round(ey / 0.5) * 0.5  # round to nearest 0.5
 
@@
 # As separate plots
-pdf("paper/figures/res_v_extent_ksize.pdf", width = 8, height = 10)
+pdf("paper/figures/res_v_extent_ksize2.pdf", width = 8, height = 10)
 par(mfrow = c(3, 2), oma = c(6, 3, 0, 0), mar = c(1, 5, 3, 4))
-lapply(1:4, function(i) {
+lapply(1:3, function(i) {  # i <- 3
   x <- kernsl[[i]]
@@
      legend.shrink = 0.9, #breaks = brks, 
-     col = rev(terrain.colors(0)))
+     col = rev(terrain.colors(80)))
 
@@
 
+## Tried and deprecated
+
+Code not shown in HTML
+
 ### Scale diversity within study
-```{r}
+```{r, eval=FALSE, echo=FALSE}
 # plot resolution
@@
 <a href="#top">Back to top</a>
+
+### Normal bootstrapping
+
+```{r, eval = FALSE, echo = FALSE}
+### Bootstrapping
+# Good old-fashioned bootstrapping with replacement, no perturbation of value
+# bootstrap
+set.seed(1)
+bootsamp <- lapply(1:1000, function(x) {
+  if((x / 100) %in% 1:100) print(x)
+  dnew <- copy(datf[, .(plot_res, sampled_area, t_btwn_samp, study_duration)])
+  dnew <- dnew[sample(1:.N, size = .N, replace = TRUE), ]
+  dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
+  dnew
+})
+
+```
+
+<a href="#top">Back to top</a>
+
+### Normal bootstrapping plot
+```{r, eval = FALSE, echo = FALSE}
+### Bootstrap - deprecated for now
+
+bootdt <- rbindlist(bootsamp)
+resdat2 <- data.table(x = log10(bootdt$t_btwn_samp), y = log10(bootdt$plot_res))
+resdat2$n <- rep(1, nrow(resdat2))
+rx <- range(resdat2$x)
+rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
+rx[2] <- 6.56
+resdat2[x < rx[1], x := rx[1]]
+resdat2[x > rx[2], x := rx[2]]
+ry <- range(aaxis1$logres)
+ry[1] <- -5  # set limits to 0.01 cm^2
+resdat2[y < ry[1], y := ry[1]]
+resdat2[y > ry[2], y := ry[2]]
+coordinates(resdat2) <- ~x + y
+
+# extent
+extdat2 <- data.table(x = log10(bootdt$study_duration), 
+                      y = log10(bootdt$sampled_area))
+extdat2$n <- rep(1, nrow(extdat2))
+ex <- range(taxis2$logdays)
+ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
+extdat2[x < ex[1], x := ex[1]]
+extdat2[x > ex[2], x := ex[2]]
+ey <- range(aaxis2$logarea)
+extdat2[y < ey[1], y := ey[1]]
+extdat2[y > ey[2], y := ey[2]]
+coordinates(extdat2) <- ~x + y
+
+# densities
+# resolution
+rres2 <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat2, 1)
+rres2 <- (rres2 / cellStats(rres2, sum)) * 100
+rng <- range(rres2[is.finite(rres2)])
+bwidth <- (rng[2] - rng[1]) / 30
+brks <- seq(rng[1], rng[2], bwidth)
+brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
+
+# extent
+rext2 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat2, 1)
+rext2 <- (rext2 / cellStats(rext2, sum)) * 100
+rng <- range(rext2[is.finite(rext2)])
+# rng[2] <- 0.1
+bwidth <- (rng[2] - rng[1]) / 30
+brks2 <- seq(rng[1], rng[2], bwidth)
+brklabs2 <- brklabs #round(brks2[seq(1, length(brks2), 3)], 2)
+
+cxa = 0.7
+pdf("paper/figures/res_v_extent_bstrap2.pdf", width = 7, 
+    height = 2.5)
+par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
+    mgp = c(2, 0.5, 0))
+# image(rres, col = cols, breaks = brks,
+image(rres2, col = rev(terrain.colors(30)), breaks = brks,
+     axes = FALSE, xlab = "", ylab = "")
+# plot(rres2, col = rev(terrain.colors(20)))
+axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
+     las = 2, tcl = -0.2, cex.axis = cxa)
+axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
+     cex.axis = cxa)
+# plot(rres2, axes = FALSE, legend.only = TRUE)
+aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
+              cex.axis = cxa, tcl = -0.1)
+plot(rres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
+
+image(rext2, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
+      ylab = "", breaks = brks2)
+axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
+     cex.axis = cxa, tcl = -0.2)
+axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
+     tcl = -0.2)
+aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
+              cex.axis = cxa, tcl = -0.1)
+plot(rext2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
+dev.off()
+
+# datf[, plot(log10(study_duration), log10(sampled_area), xlim)]
+```
+
+<a href="#top">Back to top</a>
+
+### Polygon version of plots
+
+```{r, eval = FALSE, echo = FALSE}
+stype <- data.table(datf[, study_type])
+stype[like(V1, "remote|other"), newnm := "remote"]
+stype[like(V1, "field|paleo"), newnm := "field"]
+stype[like(V1, "automated"), newnm := "automated"]
+
+tbtwn <- datf$t_btwn_samp
+# which(tbtwn < (0.1 / (60 * 60 * 24)) & tbtwn > 0)
+tbtwn <- ifelse(tbtwn == 0, 0.1 / (60 * 60 * 24), tbtwn)
+tempdat2 <- data.table(x = log10(tbtwn), y = log10(datf$study_duration))
+tempdat2[x == min(x), x := -5.5]
+tempdat2[x > -5.5 & x < tx[1], x := tx[1]]
+# tempdat2[y == min(y)]
+tempdat2[y > ty[2], y := ty[2]]
+tx2 <- round(range(tempdat2$x), 1)
+ty2 <- round(range(tempdat2$y), 1)
+
+# tempdat2 <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
+#                       y = log10(datf$study_duration))
+spatdat2 <- data.table(x = log10(datf$plot_res / 10000), # convert to ha
+                       y = log10(datf$sampled_area))
+sx2 <- round(range(spatdat2$x), 1)
+sx2[2] <- 4.8
+sy2 <- round(range(spatdat2$y), 1)
+
+# x1 = interval, x2 = study duration
+# y1 = plot res, y2 = extent
+
+pgons <- lapply(1:nrow(tempdat2), function(i) {  # i <- 6
+  print(i)
+  x <- c(rep(tempdat2[i, x], 2), rep(tempdat2[i, y], 2), tempdat2[i, x])
+  y <- c(spatdat2[i, x], rep(spatdat2[i, y], 2), rep(spatdat2[i, x], 2))
+  Polygons(list(Polygon(cbind(x, y))), ID = as.character(i))
+})
+spgons <- SpatialPolygons(pgons)
+spgons$type <- stype$newnm
+
+xl <- range(c(tx2, ty2))
+yl <- range(c(sx2, sy2))
+
+# not compelling, polygonized version
+apha <- 0.05
+par(mar = rep(2, 4))
+plot(spgons, lty = 0, axes = TRUE, xlim = xl, ylim = yl)
+plot(spgons[spgons$type == "field", ], add = TRUE, 
+     col = rgb(1, 0, 0, alpha = apha), border = "transparent")
+plot(spgons[spgons$type == "remote", ], add = TRUE, 
+     col = rgb(0, 1, 0, alpha = apha), border = "transparent")
+plot(spgons[spgons$type == "automated", ], add = TRUE, 
+     col = rgb(0, 0, 1, alpha = apha), border = "transparent")
+
+# rasterized version less compelling
+r <- raster(extent(c(range(c(tx2, ty2)), range(sx2, sy2))))
+res(r) <- 0.1
+r[] <- 1
+plot(r)
+rpts <- rasterToPoints(r, spatial = TRUE)
+spct <- sapply(over(rpts, geometry(spgons), returnList = TRUE), length)
+rpts$n <- spct
+rptsgrd <- SpatialPixelsDataFrame(rpts, rpts@data)
+rptsr <- raster(rptsgrd, layer = "n")
+plot(rptsr)
+
+```

### 2016-07-07 paper/SI.tex
Commit of all changes (analysis, text) up to 6 July, 2016 (just before incorporating Erle's edits).

@@
 Our study involved six different observers (those reviewing the papers to extract the observational scales), each of whom was given a randomly selected batch of 500 titles.
-A separate set of 20 randomly selected papers was also given to each observer, which formed the basis for calibrating observers' interpretation and extraction of scale-related information, as well as estimating between observer variance.
+A separate set of 20 papers was also randomly selected, and this set was given to all observers to review, in order to 1) calibrate the interpretations and extraction of scale-related information between observers, and 2) to estimate between-observer variance.
 
@@
 a single plot) of an ecological feature might be correlated with spatially or temporally adjacent values.
-Autocorrelation is common to most ecological phenomena \cite{legendre_spatial_1993,collins_method_2000-1}, and it means that an observation is descriptive of the properties within a broader area or time around the point at which it is taken.
+Autocorrelation is common to most ecological phenomena \cite{legendre_spatial_1993,collins_method_2000-1}, and it means that an observation can be representative of the properties within a broader area or time around the point at which it is taken.
 For example, soil properties are typically measured by analysis of soil cores collected using an augur.
@@
 
-In assessing spatial scales, our analysis only considered evaluated the Cartesian plane.
-We did not consider the z, or depth, dimension, which for certain aspects of ecology, this measure is of greater value (e.g.
-depth profiles in marine ecology), although in some cases (paleoecology) values extracted from the z-dimension provided a temporal measure that was used to calculated both the interval and the duration of the observation.
+In assessing spatial scales, our analysis only considered the Cartesian plane.
+We did not calculate the z, or depth, dimension, although this dimension is of greater importance for certain sub-disciplines of ecology (e.g.
+depth profiles in marine ecology).
+In some cases (primarily paleoecological studies), values extracted from the z-dimension provided temporal information that was used to calculate both the interval and the duration of the observation.
 
@@
 However, if the individual animals are the unit of measure (either because a sub-sample of them is being made within a defined plot, or because the observation is not contingent on being located within a plot (maybe a blood sample or body weight is being recorded, for example), then simply estimate two-dimensional area occupied by the animal, the number of animals sampled, and then calculate area from that.
-Occasionally individual animals might be recorded, but within the context of some natural feature, such as a nesting site where the survival of individual chicks is the measurement of interest \cite{roche_relative_2008-1}.
+Occasionally individual animals might be recorded, but within the context of some natural feature, such as a nesting site where the survival of individual chicks is the measurement of interest \cite{roche_relative_2008}.
 In this case, an estimate of the nest area provides the sampling resolution.
@@
 Most studies did not explicitly report values for all four of the assessed scales, and thus interpretation and judgement had to be applied to develop reasonable estimates for their values.
-The FAQ provided the protocol we followed, and was initially developed following a consultative process prior to the commencement of review.
+The FAQ provided the protocol we followed, and was initially developed following consultation between observers prior to the commencement of review.
 We conducted an iterative process of calibration to ensure consistency and reliability of estimates.
 First, we used the calibration set to calculate between-observer variability with respect to paper selection/rejection and the estimation of scales.
-Based on this, the lead author reviewed individual records in each observer's calibration set, flagged values were the estimation procedure departed from the protocol, and returned these to observers for re-estimation, without providing an estimate of the actual value.
+Based on this, the lead author reviewed individual records in each observer's calibration set, flagged values where the estimation procedure departed from the protocol, and returned these to observers for re-estimation, without providing an estimate of the actual value.
 Instead, the relevant section of the protocol was highlighted, and further explanation and clarifying discussion undertaken as needed.
@@
 
-\subsubsection*{Estimating observational density}
+\subsubsection*{Accounting for estimation uncertainty}
+\vspace{-10pt}
+Two major sources of uncertainty affected our estimation of observational scales: 1) unclear documentation of observational scales in the reviewed studies, 2) variation between observers in estimating observational scales.
+We estimated and accounted for these uncertainties in several ways.
+First, we estimated the degree of between-observer variability based on each observer's results from reviewing the 20-paper calibration set.
+We calculated how well observers agreed regarding paper inclusion/exclusion, how many extractable observations there were per included paper, and what the coefficient of variation was across all observers' estimates of scale.
+We also recorded when observations were reported in a study with an unclear or missing scale value, which we used to assess a dimension-specific estimate of uncertainty.
+
+We used the between-observer coefficients of variation for each dimension as the basis for randomly perturbing the scale values for each of the 371 observations over 1000 iterations.
+For each observation at each iteration, we perturbed its scale value in each dimension by randomly selecting a percentage value p that fell between 100 + y and 100 - y, where y was the between-observer coefficient of variation (expressed as a percentage) for a given observational dimension, and multiplying that observation's scale value by that proportion.
+This perturbed set of observations provided a basis for estimating uncertainty within our extracted set of observational scales.
+
+\subsubsection*{Scale metrics}
 \vspace{-10pt}
 In presenting results (Fig.
-1 and 2 in main text), we log-transformed (base 10) the observational scales because of the large range in values.
+1 and 2 in main text), we log-transformed (base 10) the observational scales within each of the 1000 perturbed ensemble members in order to account for the large range in scale values.
+We calculated histograms for each observational dimension from each of the log-transformed ensemble members, calculating the mean percentage density estimate for each bin across all 1000 histograms, as well as the upper and lower 2.5th percentile values for each bin (Fig.
+1 main text).
+
 To reveal the densities of observations within two dimensions (Fig.
-2 main text), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values, using a bandwidth of 1 onto a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations are concentrated.
-Bandwidths of varying resolutions were tested (see Supplementary Results section).
-
-\subsubsection*{Sensitivity of results to observer and estimation error}
-\vspace{-10pt}
-To test the impact of between-observer inconsistencies on results, as well as the possibility of error by individual observers made in estimating the scales of actual observations scale, we conducted a sensitivity analysis.
-For each of the four dimensions, we performed a 1000-iteration resample of values, adding to each record a random amount of error at each iteration.
-The error values were constrained to fall within a maximum percentage of the record's value, which was determined by the coefficient of variation between observers' estimates for the calibration set.
-We then combined the set of 1000 perturbed observations, log-transformed them, recalculated the kernel density estimates for the comparisons of the two resolution and extent dimensions (i.e.
-Fig.
-2A and B in main text), and assessed the degree to which the domains of observational concentration differed from those in our main analysis (Fig.
-S2).
+2 main text), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values across all ensemble members, using a bandwidth of 1 onto a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations are concentrated.
+Bandwidths of varying resolutions were tested on kernel density estimates were tested on kernel density estimates of sampling interval versus plot resolution (see Supplementary Results section).
 
@@
 We assessed variability and consistency between observers along multiple dimensions.
-First, we assessed how reliablly observers agreed with respect to selecting or rejecting papers for review, using the R's Agreement package \cite{yu_agreement:_2012} to calculate Fleiss' kappa statistic \cite{fleiss_measuring_1971}, which was 0.72 (z = 12.5, p$<$0.000), indicating substantial and significant agreement between observers \cite{landis_measurement_1977}.
+First, we assessed how reliably observers agreed with respect to selecting or rejecting papers for review, using the R's Agreement package \cite{yu_agreement:_2012} to calculate Fleiss' kappa statistic \cite{fleiss_measuring_1971}, which was 0.72 (z = 12.5, p$<$0.000), indicating substantial and significant agreement between observers \cite{landis_measurement_1977}.
 
@@
 
-\subsubsection*{Sensitivity analysis}
-We used the maximum uncertainty values for each dimension from the inter-observer variability analysis (Table S2) to determine the bounds of the random perturbations applied to each record in the sensitivity analysis.
-Kernel density estimates applied to the merged results of this analysis for the contrasts between spatial resolution and temporal intervals, and to spatial extent and temporal duration, suggest that the primary features observed in Figures 2A and 2B (main text) are largely unaltered by the addition of uncertainty to the estimates.
-The largest difference is in the greater density of second to sub-second duration observations across all spatial extents.
+\subsubsection*{Between-observer coefficient of variation}
+We used the maximum uncertainty values for each dimension from the inter-observer variability analysis (Table S2) to determine the bounds of the random perturbations applied to each record over the 1000 iteration resample.
 
-\begin{figure}[!ht]
-%\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/res_v_extent_pbstrap2.pdf}
-\vspace{-0.15 cm}
-\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by sampling interval and spatial resolution (left) and temporal duration and spatial extent (right).
+%\begin{figure}[!ht]
+%%\begin{wrapfigure}{c}{1\textwidth}
+%\includegraphics[width=1\textwidth]{figures/res_v_extent_pbstrap2.pdf}
+%\vspace{-0.15 cm}
+%\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by sampling interval and spatial resolution (left) and temporal duration and spatial extent (right).
 Density estimates were applied to the log-transformed values of each observational dimension after being randomly perturbed within uncertainty bounds drawn from an assessment of between-observer variability.
@@
 }
-\label{afoto1}
-\end{figure}
-
-Figure S2 indicates the effect that varying bandwidth has on the appearance of the kernel density estimates, as estimated for the same two contrasts shown in Figure S1.
-the smallest bandwidth (0.4) tested (Fig.
+%\label{afoto1}
+%\end{figure}
+%
+Figure S1 indicates the effect that varying bandwidth has on the appearance of the kernel density estimates.
+The smallest bandwidth (0.4) tested (Fig.
 S2 top) shows that the primary observational concentrations described in the main text are evident but harder to discern.
@@
 1A (main text) is bounded on the lower right at monthly to yearly intervals and 100-1000 m$^2$ resolutions and on the upper left by near-daily to monthly observations and 0.1-10 ha resolutions.
-With the smaller these appears as two separate patches (Fig.
-S2 top left), but with 0.7 bandwidth applied become coherent (Fig.
+With the smaller bandwidth this concentration appears as two separate patches (Fig.
+S2 top left), but with 0.7 bandwidth applied becomes coherent (Fig.
 S3 middle left).
@@
 %\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=0.8\textwidth]{figures/res_v_extent_ksize.pdf}
+\includegraphics[width=0.9\textwidth]{figures/res_v_extent_ksize2.pdf}
 \vspace{10 pt}

### 2016-07-07 paper/main.tex
Commit of all changes (analysis, text) up to 6 July, 2016 (just before incorporating Erle's edits).

@@
 {Lyndon Estes$^{\ast1, 2}$, Labeeb Ahmed$^{3}$, Kelly Caylor$^{2}$, Jason Chang$^{3}$, \\
-Jonathan Choi$^{4}$, Erle Ellis$^{3}$, Paul R. Elsen$^{1}$, and Tim Treur$^{4}$ \\
+Jonathan Choi$^{4}$, Erle Ellis$^{3}$, Paul R. Elsen$^{1}$, and Tim Treuer$^{4}$ \\
 \\
@@
 \begin{sciabstract}
-Observational scales plays a critical role in shaping understanding of ecological pattern and process.
+Observational scales play a critical role in shaping understanding of ecological pattern and process.
 To gain insight into the spatial and temporal domains being addressed by modern ecological science, we conducted a meta-analysis of papers published in 30 ecological journals between 2004 and 2014.
-We estimated the spatial resolution and extent and temporal interval and duration of 371 ecological observations reported within 140 papers (out of 364 randomly selected titles).
-Our analysis revealed that ecology remains a primarily field-based discipline that primarily collects data within narrow spatial and temporal domains.
-Ecologists still make limited use of methods (e.g.
-remote sensing, used for 6\% of observations) for observing larger-scale patterns and processes, although make greater use of those which permit high frequency or continuous temporal measurements (12\% of observations).
-Observational scales were not clearly reported in the majority of three dimensions, which, together with the narrow spatio-temporal ranges of observed scales suggests a limited ability to generalize the findings of many studies in modern ecological research.
+We estimated two spatial dimensions (resolution and extent) and two temporal dimensions (interval and duration) of 371 ecological observations reported within 140 papers (out of 364 randomly selected titles).
+Our analysis revealed that ecology largely remains a field-based discipline that primarily collects observations within narrow spatial and temporal domains.
+Ecologists still make limited use of methods for observing larger-scale patterns and processes (e.g.
+remote sensing, used for 6\% of observations), but make greater use of those which permit high frequency or temporally continuous \emph{in situ} measurements (12\% of observations).
+Observational scales were not clearly reported in the majority of three dimensions, which, together with the narrow spatio-temporal ranges of observed scales, suggests a limited ability to generalize the findings of many studies.
 
@@
 \linenumbers
-The scales at which ecosystems are observed plays a critical role in shaping our understanding of how they are structured and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989-1}.
+The scales at which ecosystems are observed plays a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989-1}.
 Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is the \emph{sine qua non} for understanding ecological phenomena \cite{levin_problem_1992}.
@@
 
-We characterized observational domains along two key spatial dimensions, resolution (grain) and total spatial extent, and their temporal corollaries, sampling interval and total temporal duration, and for each dimension calculated the actual, rather than representative, scales of observation (SI).
-Our review estimated the spatio-temporal scales of 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 364 randomly selected articles (0.85\% of total).
-In terms of spatial resolution, here defined as the two-dimensional space in which all measurable features of a natural object were recorded (as opposed to sub-sampled), the majority (69\%) of observations were collected in plots of $<$1 m$^2$ resolution, 22\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
+We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, sampling interval and total temporal duration.
+We calculated the actual (as opposed to representative; SI) scales of observation for these four dimensions from 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 364 randomly selected articles (0.85\% of total).
+To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI).
+We constructed histograms for each dimension from the mean of the perturbed ensembles, and estimated 95\% confidence intervals for each histogram bin (Fig.
+1).
+We constructed  kernel density estimates from the full resampled ensemble in order to assess observational distributions within different juxtapositions of the four space-time dimensions (Fig.
+2).
+
+In terms of spatial resolution (the two-dimensional space in which all measurable features of a natural object(s) were recorded, as opposed to sub-sampled), the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
 1A).
-The total spatial extent covered (the number of sampled sites multiplied by the spatial resolution) by 86\% of observations was $<$10 ha, while 35\% covered less than 1 m$^2$ (Fig.
+The total spatial extent (the number of sampled sites multiplied by the spatial resolution) of 86\% of observations was $<$10 ha, while 36\% covered less than 1 m$^2$ (Fig.
 1B).
@@
 %\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/hists2.pdf}
+\includegraphics[width=1\textwidth]{figures/hists_bs.pdf}
 \vspace{-0.15 cm}
 \caption{Histograms of the spatial resolution (A) and extent (B), sampling interval (C) and temporal duration (D) of ecological observations collected from the surveyed ecological studies.
-Bars represent percentage of the 367 collected records falling within each bin.}
+Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval.
+}
 \label{afoto1}
@@
 In the temporal dimensions, 36\% of observations were not repeated (Fig.
-1C), 17\% were repeated at short intervals (sub-second to daily), 19\% at daily to monthly intervals, 23\% at monthly to yearly intervals, and 5\% at yearly to decadal intervals.
-The temporal duration of studies--defined here as the total amount of time the ecological feature was directly observed, and calculated as the number of repeat observations multiplied by the effective sampling duration (SI)--was less than one day for 69\% of sampled observations, between one day and one month for 19\% of observations, between one month and one year for 8\% of observations, and greater than one year for 5\% of observations (including several paleoecological studies covering centuries to millenia; Fig.
+1C), 15\% were repeated at short intervals (sub-second to daily), 21\% at daily to monthly intervals, 19\% at monthly to yearly intervals, and 8\% at yearly to decadal intervals.
+The temporal duration of studies (the total amount of time the ecological feature was directly observed, calculated as the number of repeat observations multiplied by the effective sampling duration; SI) was less than one day for 71\% of sampled observations, between one day and one month for 17\% of observations, between one month and one year for 7\% of observations, and greater than one year for 5\% of observations (including several paleoecological studies covering centuries to millennia; Fig.
 1D).
 
-The juxtaposition of these observational dimensions provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (Fig.
+Juxtaposing the observational dimensions provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (Fig.
 2).
-Contrasting spatial resolution with sampling interval reveals that the greatest density of observations consists of unreplicated samples collected in plots of 10 cm$^2$ to 1 m$^2$.
-Two smaller centers of observational concentration are also evident, the largest of which had plots that were also of 10 cm$^2$-1 m$^2$ and were revisited at daily to monthly intervals.
-A less dense, oblong concentration of observations defined on the lower right by monthly to yearly observations at 100-1000 m$^2$ and on the upper left by near-daily to monthly observations with 0.1-10 ha resolution is also evident.
-This lower right to upper left orientation marks the tradeoff between resolution and return that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-second intervals), high spatial resolution (0.1-100 cm$^2$) observations.
-This line demonstrates the inverse tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort and thus reduce sampling frequency \cite{kareiva_spatial_1988}.
+Contrasting spatial resolution with sampling interval reveals that the majority of temporally replicated observations (the 36\% that were unreplicated were excluded because they have no interval value) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig.
+2A).
+A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident.
+This lower right to upper left orientation reflects the tradeoff between resolution and return interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations.
+This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
 
@@
 %\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/kde43.pdf}
+\includegraphics[width=1\textwidth]{figures/kde45.pdf}
 \vspace{-0.15 cm}
@@
 
-Contrasting temporal duration and spatial extent (Fig.
-2B) reveals that the bulk of ecological observations covered total time periods of one hour to $<$1 month and areas of $\leq$1 m$^2$.
-The next greatest concentration were observations centered on 1 hour duration and between 1 m$^2$ and 1 ha in extent.
-Very few observations' had spatial extents $>$1,000,000 ha, and because these were collected by satellites that make near-instantaneous measurements, they tended to have aggregate durations of $<$1 hour.
-In contrast, the longest periods of observation were provided by paleo-ecological or long-term weather stations, both of which produce continuous, point-scale measurements (i.e.
-soil cores and weather instrumentation).
+Contrasting temporal duration and spatial extent (for all observations) reveals that the bulk of ecological observations covered total time periods of one minute to $<$1 month and areas of $\leq$1 m$^2$ (Fig.
+2B).
+The next greatest concentration were instantaneous ($\leq$1 second) observations of $<$1 m$^2$ extent, followed by a fainter concentration centered on 1 hour duration and between 1 m$^2$ and 1 ha in extent.
+Very few observations had spatial extents $>$1,000,000 ha, and because these were collected by satellites that make near-instantaneous measurements, they tended to have aggregate durations of $\leq$1 hour.
+In contrast, the longest periods of observation were provided by paleo-ecological or long-term weather stations, both of which produce continuous, point-scale measurements (e.g.
+sediment cores and weather instrumentation).
 
-Comparing the two spatial dimensions against one another shows that observational extent scales log-linearly with resolution (Fig.
+Comparing the two spatial dimensions against one another (for all observations) reveals a log-linear relationship wherein resolution strongly constrains extent, such that the ratio between extent and resolution falls as the latter increases (Fig.
 2C).
-Up to 100 ha, there is a nearly 1:1 correspondence between resolution and extent that results from the tendency for spatial replicates to decline with plot size (and thus effort) in field-based studies \cite{kareiva_spatial_1988}, which accounted for 80\% of the observations in our study.
-These stand in contrast to most remotely sensed observations, which comprised just 6.5\% of observations, and are characterized by spatial resolutions that are very fine relative to their extent.
-These constitute the majority of the low density patches (at 0.1-10 ha resolution and 1-10 million ha extent; 100-10,000 ha resolution and 1 million-10 billion ha extent) above the line defined by the denser concentration of field studies in Figure 2C.
+As with interval, this pattern indicates that spatial replication diminishes with plot size \cite{kareiva_spatial_1988} in field-based studies, which accounted for 80\% of observations.
+This stands in contrast to the 6.5\% of observations that were made with remote sensing, which have fine spatial resolutions relative to their large extents.
+These comprise the majority of observations in the four low density patches (e.g.
+0.1-10 ha resolution and 1-10 million ha extent) falling above the denser line that terminates at 100 ha resolution and 1000 ha extent.
 
-A similar linear relationship between sampling interval and temporal duration was not evident.
-For repeated observations, the greatest concentration was among those made at daily to yearly intervals with overall durations of hours to weeks (Fig.
+A similar linear relationship between sampling interval and temporal duration was not evident (unreplicated observations were excluded because they lacked interval values).
+For repeated observations, the greatest concentration was among those made at daily to yearly intervals with overall durations of hours to months (Fig.
 2D).
+A second concentration of sub-monthly to yearly observations lasting one second or less is also evident.
 Very few repeat observations exceeded a year in duration, and these were collected at daily to yearly intervals.
-The majority of unreplicated observations lasted between one minute and one day.
 
@@
 Observations that were non-continuous in time or space (e.g.
-point-based field measurements) may effectively represent larger spatial scales than our estimates suggest, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997, palmer_scale_1994,cao_comparison_2002, legendre_spatial_1993,collins_method_2000-1}.
-This concern does not apply to spatially continuous observations, thus the net effect is that if we made estimates of effective spatial scale, the spatial resolution and extent estimates would be compressed from the lower end of their ranges.
+point-based field measurements) may effectively represent larger scales than our estimates suggest, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997, palmer_scale_1994,cao_comparison_2002, legendre_spatial_1993,collins_method_2000-1}.
+From a spatial perspective, this concern does not apply to contiguous sampling schemes, which were primarily based on remote sensing and cover the largest area, thus the net effect is that this analysis may underestimate the effective areas of smaller observational extents.
 From a temporal perspective, our analysis likely underestimates the effective duration of many observations, particularly those where instantaneous, repeated measures of slow-changing ecological features were made.
@@
 However, many studies focused on more dynamic phenomena, and for these long periods of continuous observation may be more important for understanding dynamics than frequent repeats.
-For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to directly measure fire behavior \cite{clements_observing_2007}.
+For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to directly measure fire behavior \cite{clements_observing_2007}, except patterns resolving themselves at scales greater than 12-24 hours.
 For this reason, and to provide a consistent standard for estimating duration, we recorded actual duration.
@@
 The observational scales were not clearly reported in many studies, such that we had to estimate spatial resolution and extent in 62\% and 68\% of cases, respectively, and temporal interval and duration in 40\% and 86\% of cases.
-Although sensitivity analyses show that our general findings are robust to this uncertainty (SI), it would have likely confounded efforts to estimate effective scales.
+Although the confidence intervals around our scale estimates (Fig.
+1) suggests that our findings are robust to this level of uncertainty (SI), it would have likely confounded efforts to estimate effective scales.
 
@@
 Our findings suggest that most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month).
-Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, suggesting that, despite well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989-1}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989-1}.
+Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, suggesting that, despite the well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989-1}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989-1}.
 Furthermore, the unclear documenting of observational scales suggests that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
-Taken together, the narrowness of scales and the ambiguity of their documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) suggests that much of the knowledge produced in ecology may not be generalizable \cite{margulies_ambiguous_2016,wheatley_factors_2009,wiens_spatial_1989-1}.
+Taken together, this narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) suggests that the findings associated with many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989-1}.
 
 The generally small spatial scales of observation is a consequence of the continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
-Despite early and repeated calls for ecologists to use remote sensing because it provides a synoptic view that field measurements cannot \cite{turner_remote_2003, kerr_space_2003, pettorelli_satellite_2014}, and subsequent demonstration of its importance for multi-scale studies \cite{estes_habitat_2008, estes_predictive_2011}, our results indicate this observational method is still under-utilized in ecology.
-In contrast, ecologists have more widely adopted methods that increase the temporal resolution of observations.
+Despite early and repeated calls for ecologists to use remote sensing because it provides a synoptic view that field measurements cannot \cite{turner_remote_2003, kerr_space_2003, pettorelli_satellite_2014}, and subsequent demonstration of its importance for multi-scale studies \cite{estes_habitat_2008, estes_predictive_2011}, our results indicate this observational method is has not yet been widely adopted in for ecological research.
+There may two reasons for the low penetration of remote sensing into ecology.
+First, remote sensing can be a challenging method for ecologists to learn, many of whom may not have access to appropriate training \cite{pettorelli_satellite_2014}.
+Second, satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.
+
+In contrast to remote sensing, ecologists have made broader use of technologies that increase the temporal resolution of observations.
 Automated sensors were used to record 12\% of all observations, and accounted for most very high frequency measurements (intervals $\leq$1 hour).
-As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses (and many of the reviewed studies that used automated sensing aggregated the resulting observations before analyzing them), whereas it is extremely difficult, if not impossible, to disaggregate long-interval data.
+As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses (and many of the reviewed studies that used automated sensing aggregated the resulting observations before analyzing them), whereas longer-interval data (e.g.
+annual biomass accumulation) cannot be disaggregated into shorter interval measurements (e.g.
+weekly biomass accumulation) without interpolation, which turns data into modeled, rather than direct, observations.
 
@@
 \begin{scilastnote}
- \item This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G),
+ \item This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G).
 \end{scilastnote}

### 2016-07-07 paper/main.tex
Final draft with Erle's edits. Still left: full review bibliography

@@
 {\arabic{lastnote}.}
-{\setlength{\leftmargin}{.22in}}
+{\setlength{\leftmargin}{.3in}}
 {\setlength{\labelsep}{.5em}}}
@@
 remote sensing, used for 6\% of observations), but make greater use of those which permit high frequency or temporally continuous \emph{in situ} measurements (12\% of observations).
-Observational scales were not clearly reported in the majority of three dimensions, which, together with the narrow spatio-temporal ranges of observed scales, suggests a limited ability to generalize the findings of many studies.
+Resolution, extent, and duration were not clearly reported for the majority of observations, which, together with the narrow spatio-temporal ranges of observed scales, suggests a limited ability to generalize the findings of many studies.
 
@@
 We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, sampling interval and total temporal duration.
-We calculated the actual (as opposed to representative; SI) scales of observation for these four dimensions from 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 364 randomly selected articles (0.85\% of total).
+We calculated the actual, as opposed to representative (SI), scales for these four dimensions from 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 364 randomly selected articles (from 42,918 total).
 To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI).
@@
 
+\noindent \textbf{\emph{Uncertainties in quantifying observational scales}}\\
 This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI).
@@
 Although the confidence intervals around our scale estimates (Fig.
-1) suggests that our findings are robust to this level of uncertainty (SI), it would have likely confounded efforts to estimate effective scales.
+1) suggests that our findings are robust to this level of uncertainty, it would have likely confounded efforts to estimate effective scales.
 
+\noindent \textbf{\emph{Insights into the scales of modern ecology}}\\
 Despite these caveats, our results provide valuable insight into the spatial and temporal domains being addressed by modern ecological research.
-Our findings suggest that most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month).
-Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, suggesting that, despite the well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989-1}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989-1}.
-Furthermore, the unclear documenting of observational scales suggests that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
-Taken together, this narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) suggests that the findings associated with many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989-1}.
+Our results show that most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month).
+Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, indicating that, despite the well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989-1}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989-1}.
+Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
+Taken together, this narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) suggests that the ecological understanding drawn from many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989-1}, a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}.
 
 The generally small spatial scales of observation is a consequence of the continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
-Despite early and repeated calls for ecologists to use remote sensing because it provides a synoptic view that field measurements cannot \cite{turner_remote_2003, kerr_space_2003, pettorelli_satellite_2014}, and subsequent demonstration of its importance for multi-scale studies \cite{estes_habitat_2008, estes_predictive_2011}, our results indicate this observational method is has not yet been widely adopted in for ecological research.
-There may two reasons for the low penetration of remote sensing into ecology.
+Despite early and repeated calls for ecologists to use remote sensing because it provides a synoptic view that field measurements cannot \cite{turner_remote_2003, kerr_space_2003, pettorelli_satellite_2014}, and subsequent demonstration of its importance for multi-scale studies \cite{estes_habitat_2008, estes_predictive_2011}, our results indicate this method has not yet been widely adopted in ecological research.
+Two reasons lie behind this slow uptake.
 First, remote sensing can be a challenging method for ecologists to learn, many of whom may not have access to appropriate training \cite{pettorelli_satellite_2014}.
@@
 Similarly, the advent of unmanned systems offers the ability to measure ecological features at high spatial and temporal frequencies over large areas \cite{anderson_lightweight_2013}, which were scales that were previously impractical to access.
-The ever-falling cost of sensor technology and the ubiquity of cell phone connections also means that ecologists have a growing ability to make spatially dense, high frequency observations \cite{wolf_gsm-based_2012,collins_new_2006,porter_wireless_2005}.
+The ever-falling cost of sensor technology and the ubiquity of cell phones also means that ecologists, together with a growing army of citizen-scientists, have the unprecedented ability to make spatially dense, high frequency observations over large areas \cite{wolf_gsm-based_2012,collins_new_2006,porter_wireless_2005,dickinson_current_2012}.
 A greater attentiveness to scale in general, including more meticulous documentation of observed dimensions, may help to facilitate the spread of ecological research to sparsely studied scales, while improving transferability of knowledge within the discipline.
@@
 \item This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G).
+Erle Ellis, Jason Chang, and Labeeb Ahmed of the GLOBE Project (http://globe.umbc.edu) were supported by the U.S.
+National Science Foundation (1125210).
 \end{scilastnote}

### 2016-08-25 paper/SI.tex
Final draft with Erle's edits. Still left: full review bibliography
Updated ms to partially match Nature EcoEvo format, mainly with fuller Nature-style abstract.

@@
 
-\renewcommand\refname{References and Notes}
+%\renewcommand\refname{References and Notes}
 
@@
 We left out the latter category because our intent was to evaluate the domains for observations of natural systems, and we wanted to avoid the bias that would be imposed by the relatively narrow spatial and temporal scales of experiments \cite{kareiva_spatial_1988,tilman_global_2011}.
+A bibliography of the reviewed papers follows the References and Notes section below.
 
@@
 \begin{enumerate}
-  \Hitem \emph{\textbf{What are the general inclusion/exclusion criteria for studies?}} Studies should be excluded from this analysis if they are: 1) opinion/perspectives pieces; 2) model-only studies, particularly theoretical models, which are not developed or tested against observed data; 3) if they are experimental manipulations (but if there is a study that has a mix of observational and experimental, record the observational treatments and exclude the manipulated treatments).
+  \Hitem \emph{\textbf{What are the general inclusion/exclusion criteria for studies?}} Studies should be excluded from this analysis if they are: 1) opinion/perspectives pieces; 2) book reviews; 3) model-only studies, particularly theoretical models, which are not developed or tested against observed data; 4) if they are experimental manipulations (but if there is a study that has a mix of observational and experimental, record the observational treatments and exclude the manipulated treatments).
 
@@
 We calculated how well observers agreed regarding paper inclusion/exclusion, how many extractable observations there were per included paper, and what the coefficient of variation was across all observers' estimates of scale.
-We also recorded when observations were reported in a study with an unclear or missing scale value, which we used to assess a dimension-specific estimate of uncertainty.
+We also recorded when observations were reported in a study with an unclear or missing scale value.
 
@@
 
-\bibliographystyle{Science}
-
-% Following is a new environment, {scilastnote}, that's defined in the
-% preamble and that allows authors to add a reference at the end of the
-% list that's not signaled in the text; such references are used in
-% *Science* for acknowledgments of funding, help, etc.
-
-%\begin{scilastnote}
-%\item We've included in the template file \texttt{scifile.tex} a new
-%environment, \texttt{\{scilastnote\}}, that generates a numbered final
-%citation without a corresponding signal in the text.  This environment
-%can be used to generate a final numbered reference containing
-%acknowledgments, sources of funding, and the like, per {\it Science\/}
-%style.
-%\end{scilastnote}
-
-% For your review copy (i.e., the file you initially send in for
-% evaluation), you can use the {figure} environment and the
-% \includegraphics command to stream your figures into the text, placing
-% all figures at the end.  For the final, revised manuscript for
-% acceptance and production, however, PostScript or other graphics
-% should not be streamed into your compliled file.  Instead, set
-% captions as simple paragraphs (with a \noindent tag), setting them
-% off from the rest of the text with a \clearpage as shown  below, and
-% submit figures as separate files according to the Art Department's
-% instructions.
-
-\clearpage
+\bibliographystyle{naturemag}
 

### 2016-08-25 paper/main.tex
Updated ms to partially match Nature EcoEvo format, mainly with fuller Nature-style abstract.

@@
 {Lyndon Estes$^{\ast1, 2}$, Labeeb Ahmed$^{3}$, Kelly Caylor$^{2}$, Jason Chang$^{3}$, \\
-Jonathan Choi$^{4}$, Erle Ellis$^{3}$, Paul R. Elsen$^{1}$, and Tim Treuer$^{4}$ \\
+Jonathan J. Choi$^{4}$, Erle Ellis$^{3}$, Paul R. Elsen$^{1}$, and Tim Treuer$^{4}$ \\
 \\
@@
 
-\begin{sciabstract}
-Observational scales play a critical role in shaping understanding of ecological pattern and process.
+%\begin{sciabstract}
+\textbf{Observational scale shapes our understanding of ecosystems \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
+Ecological patterns emerge at temporal and spatial scales that may be coarser or finer than the processes that shape them, thus investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992}.
+Given the centrality of scale to ecology, and the growing ability to observe ecological phenomena beyond a field perspective, it is important to understand the scales at which modern ecological research is conducted.
+To reveal these, we conducted a meta-analysis of recent (2004-2014) ecological research, in which we estimated the spatial resolution and extent, and temporal interval and duration, of reported ecological observations.
+Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
+Ecologists still make limited use of remote sensing (6\% of observations), which is necessary for observing larger-scale patterns and processes \cite{turner_remote_2003,pettorelli_satellite_2014}, but make slightly greater use of technologies that capture high frequency or temporally continuous point-based measurements (12\% of observations).
+Resolution, extent, and duration were not clearly reported in most reviewed studies.
+This indicates inattentiveness to scale, which, together with the narrow spatio-temporal domains of observations, suggests a limited ability to generalize the knowledge gained in many ecological studies, which is critical for developing an ability to predict ecological responses in an era of increasing global change \cite{levin_problem_1992}.
+An emerging wave of observing and analytical technologies (e.g.
+nano-satellites, unmanned systems, open image archives, and high-power processing platforms) now makes it dramatically easier and cheaper to collect high-quality, ecologically meaningful measurements at scales inaccessible to field scientists.
+Combined with a greater attention to scale in research design, these advances can facilitate the spread of ecological research to sparsely observed scales, while improving the transferability of ecological knowledge.}
+
+%Observational scales play a critical role in shaping our understanding of ecological pattern and process.
 To gain insight into the spatial and temporal domains being addressed by modern ecological science, we conducted a meta-analysis of papers published in 30 ecological journals between 2004 and 2014.
-We estimated two spatial dimensions (resolution and extent) and two temporal dimensions (interval and duration) of 371 ecological observations reported within 140 papers (out of 364 randomly selected titles).
+We estimated two spatial dimensions (resolution and extent) and two temporal dimensions (interval and duration) of 371 ecological observations reported within 140 papers (out of 346 randomly selected titles).
 Our analysis revealed that ecology largely remains a field-based discipline that primarily collects observations within narrow spatial and temporal domains.
@@
 %  An example of the style is the special\texttt{\{sciabstract\}} environment used to set up the abstract you see here.
-\end{sciabstract}
+%\end{sciabstract}
 
 \linenumbers
-The scales at which ecosystems are observed plays a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989-1}.
+The scales at which ecosystems are observed plays a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
 Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is the \emph{sine qua non} for understanding ecological phenomena \cite{levin_problem_1992}.
@@
 Given the centrality of scale to understanding ecology, and the growing ability to study ecological phenomena across a broad range of spatial and temporal scales, it is important to assess the scales at which ecological research is presently being conducted.
-To gain insight into this question, we quantified the spatial and temporal domains of empirical observations that were reported in a representative sample of studies published between 2004-2014 in the top 30 ecological journals (by 2014 impact factor).
+To gain insight into this question, we quantified the spatial and temporal domains of empirical observations that were reported in a representative sample of studies published between 2004-2014 in the top 30 ecological journals (by impact factor).
 Empirical observations provide the necessary means for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, thus it stands to reason that the temporal and spatial distributions of ecological observations may be indicative of modern ecology's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.
@@
 We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, sampling interval and total temporal duration.
-We calculated the actual, as opposed to representative (SI), scales for these four dimensions from 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 364 randomly selected articles (from 42,918 total).
+We calculated the actual, as opposed to representative (SI), scales for these four dimensions from 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 346 randomly selected articles (from 42,918 total).
+An additional 56 papers that were cited as the source of observations in the selected papers were also reviewed.
 To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI).
@@
 Our results show that most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month).
-Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, indicating that, despite the well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989-1}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989-1}.
+Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, indicating that, despite the well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989}.
 Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
-Taken together, this narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) suggests that the ecological understanding drawn from many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989-1}, a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}.
+Taken together, this narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) suggests that the ecological understanding drawn from many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}, a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}.
 
@@
 
-%These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989-1}.
+%These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989}.
 
@@
 %Based on the observational domains recorded here, it appears that ecologists are still primarily collecting data at very fine spatial scales and over very narrow extents.
-Although the temporal duration of most ecological observations is also narrow, and a sizable number of observations are not repeated, greater effort is being devoted to capturing temporal dynamics.
-These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989-1}.
+Although then temporal duration of most ecological observations is also narrow, and a sizable number of observations are not repeated, greater effort is being devoted to capturing temporal dynamics.
+These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989}.
 
@@
 %\paragraph*{Displayed math.} 
+\bibliography{/Users/lestes/Dropbox/publications/fullbib}
+\bibliographystyle{naturemag}
 
-\bibliography{/Users/lestes/Dropbox/publications/fullbib}
-
-\bibliographystyle{Science}
+\noindent \textbf{Acknowledgements} This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G).
+Erle Ellis, Jason Chang, and Labeeb Ahmed of the GLOBE Project (http://globe.umbc.edu) were supported by the U.S.
+National Science Foundation (1125210).
 
@@
 
-\begin{scilastnote}
-\item This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G).
+%\begin{scilastnote}
+%\item This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G).
 Erle Ellis, Jason Chang, and Labeeb Ahmed of the GLOBE Project (http://globe.umbc.edu) were supported by the U.S.
 National Science Foundation (1125210).
-\end{scilastnote}
+%\end{scilastnote}
 

### 2017-04-05 Rmd/calibration.Rmd
Initial steps in revision of analysis, including updated calibration and main data sets (addition of new effective extent variable, change of variable names:
sampled_area -> act_ext
eff_ext
study_duration -> act_dur
eff_dur
samp_duration -> samp_dur

Analyses revised through calibration set and data merging of main analysis

@@
 # cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
-cal <- data.table(read_excel(fp(p_dat, "merged_calibrationffff.xlsx")))
+cal <- data.table(read_excel(fp(p_dat, "merged_calibrationfffff.xlsx")))
 names(cal)
-cal[, c("notes", "V30") := NULL]
-for(j in names(cal)[c(4:5, 10:24)]) {
+
+cal[, c("notes", "V31") := NULL]
+nuvars <- c("inout", "feature", "study_year", "st", "plot_res", "n_sites",
+            "act_ext", "eff_ext", "samp_dur", "t_btwn_samp", "act_dur", 
+            "eff_dur", paste0("comp", 1:3), paste0("struc", 1:3), 
+            paste0("func", 1:3))
+for(j in nuvars) {
   set(cal, j = j, value = as.numeric(cal[[j]]))
@@
 ```{r}
-varnms <- c("plot_res", "n_sites", "sampled_area", "t_btwn_samp", 
-            "study_duration")
+varnms <- c("plot_res", "n_sites", "act_ext", "eff_ext", "t_btwn_samp", 
+            "act_dur", "eff_dur")
 mus <- cal[inout == 1, lapply(.SD, mean, na.rm = TRUE), by = .(DOI, observer),             .SDcols = varnms]  # mean variable val per paper per observer
@@
 
-par(mfrow = c(3, 2), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
+par(mfrow = c(3, 3), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
 for(i in varnms) {
@@
 ```{r, results='hold'}
-CV <- function(x) sd(x) / mean(x)
+CV <- function(x) sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)
 rnms <- c("DOI", "feature", "observer", "plot_res", "n_sites", 
-          "sampled_area", "t_btwn_samp", "study_duration")
+          "act_ext", "eff_ext", "t_btwn_samp", "act_dur", "eff_dur")
 calr <- cal[, rnms, with = FALSE]
+
 # calr[feature != 99, ]
@@
 calr[feature != 99, lapply(.SD, CV), by = .(DOI, feature), 
-     .SDcols = rnms[4:8]][, lapply(.SD, mean, na.rm = TRUE), 
-                          .SDcols = rnms[4:8]]
+     .SDcols = rnms[4:10]][, lapply(.SD, mean, na.rm = TRUE), 
+                          .SDcols = rnms[4:10]]
 
@@
 ```
-
-Tim needs to fix paleo studies in calibration (and main set) giving too much credit to effective area. [DONE]
-
-Others need to fix their study duration figures. Paul on paleo one in calibration, Choi, Chang, etc in main set.  [DONE]
-
-For N-sites, count sites only once if there are repeated measures taken at them, e.g. Jonathan in 10.1111/j.1466-8238.2009.00454.x for bird counts [Chang, Ahmed remaining]

### 2017-04-05 Rmd/space-time.Rmd
Initial steps in revision of analysis, including updated calibration and main data sets (addition of new effective extent variable, change of variable names:
sampled_area -> act_ext
eff_ext
study_duration -> act_dur
eff_dur
samp_duration -> samp_dur

Analyses revised through calibration set and data merging of main analysis

@@
 # dat <- as.data.table(read_excel(full_path(p_dat, "merged.xlsx")))
-dat <- as.data.table(read_excel(full_path(p_dat, "mergedfff.xlsx")))
-cal <- data.table(read_excel(fp(p_calib, "merged_calibrationffff.xlsx")))
-full <- data.table(read_excel(fp(p_dat, "full_less_Treuer.xlsx")))
+dat <- as.data.table(read_excel(full_path(p_dat, "mergedfffff.xlsx")))
+dato <- as.data.table(read_excel(full_path(p_dat, "archive/mergedfff.xlsx")))
+cal <- data.table(read_excel(fp(p_calib, "merged_calibrationfffff.xlsx")))
+full <- data.table(read_excel(fp(p_dat, "full_no_treuer_toss.xlsx")))
+# full <- data.table(read_excel(fp(p_dat, "full_less_Treuer.xlsx")))
+
+setnames(dato, "DOI/title", "DOI")
+dato[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
+dato[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
 
@@
 dat[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
+cal[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
 full <- gsub("DOI:|doi:|DOI: |doi: |DOI ", "", full$DOI)
 full <- gsub("\\s", "", full)  # remove DOI and whitespace
-unique(full)
-
-dat <- dat[, names(dat)[1:20], with = FALSE]
-
-unique(dat$study_type)
+
+# subset main results and define study type
+dat <- dat[, names(dat)[1:22], with = FALSE]
 dat[, study_type := tolower(study_type)]
-kwords <- c("field", "paleo", "remote", "automated", "other")
-for(i in 1:length(kwords)) dat[like(study_type, kwords[i]), st := i]
+# kwords <- c("field", "paleo", "remote", "automated", "other")
+tst <- dat[, study_type]
+
+kwords <- cbind(c("field", "paleo", "remote", "automated", "other"), 
+                c("field", "paleo", "remote", "automated", "other"))
+# for(i in 1:length(kwords)) dat[like(study_type, kwords[i]), st := i]
+for(i in 1:nrow(kwords)) {
+  dat[like(study_type, kwords[i, 1]), study_type := kwords[i, 2]]
+}
+# cbind(tst, dat[, study_type])
 
@@
 
-# class(cal$DOI)
-# class(dat$DOI)
+kwords <- cbind(c("field", "direct", "paleo", "remote", "automated", "other"),
+                c("field", "field", "paleo", "remote", "automated", "other"))
+cal[, study_type := tolower(study_type)]
+tst <- cal[, study_type]
+for(i in 1:nrow(kwords)) {
+  cal[like(study_type, kwords[i, 1]), study_type := kwords[i, 2]]
+}
+# cbind(tst, cal[, study_type])
+
+# unique(dat$DOI)
+# a <- unique(dato$DOI)
+# b <- unique(dat$DOI)
+# notin <- sort(a)[!sort(a) %in% sort(b)]
+# dato[DOI %in% notin, 1:5, with = FALSE]
+# dat[observer == "Ahmed" & journal == "Oecologia"]
 # unique(cal$DOI) %in% unique(dat$DOI)
@@
 # dat[DOI == dat[cal$DOI %in% dat$DOI, DOI][7]]
+# dat[observer == "Treuer", unique(DOI)]  # 19
 
@@
 ```{r, message=FALSE, warning=FALSE, results='hide'}
-dat2 <- copy(dat[, .(st, DOI, study_type, plot_res, n_sites, sampled_area,
-                     samp_duration, t_btwn_samp, study_duration)])
-rnms <- c("st", "DOI", "study_type", "feature", "observer", "plot_res", 
-          "n_sites", "sampled_area", "t_btwn_samp", "study_duration")
+# dat2 <- copy(dat[, .(st, DOI, study_type, plot_res, n_sites, act_ext, eff_ext,
+#                      samp_dur, t_btwn_samp, act_dur, eff_dur)])
+# rnms <- c("st", "DOI", "study_type", "feature", "observer", "plot_res", 
+#           "n_sites", "act_ext", "eff_ext", "samp_dur", "t_btwn_samp",
+#           "act_dur", "eff_dur")
+dat2 <- copy(dat[, .(DOI, study_type, plot_res, n_sites, act_ext, eff_ext,
+                     samp_dur, t_btwn_samp, act_dur, eff_dur)])
+rnms <- c("DOI", "study_type", "feature", "observer", "plot_res", 
+          "n_sites", "act_ext", "eff_ext", "samp_dur", "t_btwn_samp",
+          "act_dur", "eff_dur")
 calr <- cal[, rnms, with = FALSE]
-for(j in names(calr)[c(4, 6:10)]) set(calr, j=j, value = as.numeric(calr[[j]]))
-
+selind <- c(3, 5:length(rnms))
+for(j in names(calr)[selind]) set(calr, j=j, value = as.numeric(calr[[j]]))
+
+# str(calr)
 # select down to columns of interest
 # st <- calr$study_type[which(calr$feature != 99)]
-calr <- calr[feature != 99, lapply(.SD, mean), by = .(DOI, feature), 
-             .SDcols = rnms[c(1, 6:10)]]
-length(unique(calr$DOI)) + length(unique(dat$DOI))  # 140 papers providing dat
-unique(calr$DOI)
-calr_type <- data.table(cbind("DOI" = unique(calr$DOI), # calib field studies
-                              "study_type" = c(rep("field", 5), "paleo",
-                                               rep("field", 4))))
+calr_type <- calr[feature != 99, { 
+  list(unique(study_type[!is.na(study_type)]))
+}, by = .(DOI, feature)][, list("study_type" = unique(V1)), by = DOI]
+
+# calr[DOI == "10.1007/s00442-005-0338-3" & feature == 2, 
+#      mean(eff_dur, na.rm = T)]
+calr <- calr[feature != 99, {
+  lapply(.SD, function(x) mean(x, na.rm = TRUE))
+},  by = .(DOI, feature), .SDcols = rnms[c(5:ncol(calr))]]
+length(unique(calr$DOI)) + length(unique(dat$DOI))  # 134 papers providing dat
+# unique(calr$DOI)
+
+# calr_type <- data.table(cbind("DOI" = unique(calr$DOI), # calib field studies
+#                               "study_type" = c(rep("field", 5), "paleo",
+#                                                rep("field", 4))))
 
 # combine full and calibration
-calr[, samp_duration := NA]
+# calr[, samp_dur := NA]
+calr[, samp_dur := NA]
 calr[, "feature" := NULL] #c("DOI", "feature") := NULL]
 calr <- merge(calr, calr_type, by = "DOI")
-setcolorder(calr, names(calr)[c(2:1, 9, 3:5, 8, 6:7)])
+setcolorder(calr, names(dat2))
 datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
@@
 doisup <- unlist(strsplit(doisup, ";"))
-
 ```
@@
 
-`r length(unique(full)) + 20 + 57` papers main analysis (everyone's but Tim's)
-
-`r (length(unique(full)) + 20 + 57) / 42918 * 100` % of all papers since 2004
+`r length(unique(full)) + (57 - 19)` papers main analysis (adding in Tim's rejected number separately)
+
+`r (length(unique(full)) + (57 - 19)) / 42918 * 100` % of all papers since 2004
 
@@
 ### Adjustments/fixes
-1.
-if `samp_duration != study_duration` and `t_btwn_samp == 0`, set `study_duration == samp_duration` (actually, no, in many cases there is a good reason to have to `study_duration != samp_duration` (see below))
+1. if `samp_dur != act_dur` and `t_btwn_samp == 0`, set `act_dur == samp_dur` (actually, no, in many cases there is a good reason to have to `act_dur != samp_dur` (see below))
 2.
@@
 # apply fixes (mostly to account for observer omissions)
-# 1. if samp_duration != study_duration and t_btwn_samp == 0, set 
-#    study_duration == samp_duration
+# 1. if samp_dur != act_dur and t_btwn_samp == 0, set 
+#    act_dur == samp_dur
 names(datf)
-datf[t_btwn_samp == 0 & (study_duration != samp_duration), ]  
-dat[t_btwn_samp == 0 & (study_duration != samp_duration), c(1, 3, 8:13), with=F]  
+datf[t_btwn_samp == 0 & (act_dur != samp_dur), ]  
+dat[t_btwn_samp == 0 & (act_dur != samp_dur), c(1, 3, 8:13), with=F]  
 # function to set t_btwn_samp to 365 * 100000 to indicate once-off studies 
@@
 Defined by the per variable CV based on uncertainty between observers.
-Here we are just going to use apply the variability to each variable for every observation (previous incarnation used uncertainty on just the variables listed as uncertain, with recalculation of dependent variables (`study_duration`, `sampled_area`) made after pertubation [see commit prior to second or third commit on 4 April for code]).
+Here we are just going to use apply the variability to each variable for every observation (previous incarnation used uncertainty on just the variables listed as uncertain, with recalculation of dependent variables (`act_dur`, `act_ext`) made after pertubation [see commit prior to second or third commit on 4 April for code]).
 
@@
 
-    dnew <- copy(datf[, .(plot_res, sampled_area, t_btwn_samp, study_duration)])
+    dnew <- copy(datf[, .(plot_res, act_ext, t_btwn_samp, act_dur)])
     # modifiers
@@
     dnew[plot_res <= 0, plot_res := 0.00000001]
-    dnew[sampled_area <= 0, sampled_area := 0.00000001]
-    dnew[study_duration <= 0, study_duration := 0.00000001]
+    dnew[act_ext <= 0, act_ext := 0.00000001]
+    dnew[act_dur <= 0, act_dur := 0.00000001]
     dnew[t_btwn_samp < 0, t_btwn_samp := 0.00000001]
@@
 datf[t_btwn_samp == 0, .N] / datf[, .N]  # 36% of samples are once-offs
-datf[study_duration == samp_duration, .N] / nrow(datf)
-datf[(study_duration == samp_duration) & t_btwn_samp != 0]
-datf[(study_duration != samp_duration) & t_btwn_samp == 0]
-
-hdat <- cbind("res" = log10(datf$plot_res), "ext" = log10(datf$sampled_area), 
+datf[act_dur == samp_dur, .N] / nrow(datf)
+datf[(act_dur == samp_dur) & t_btwn_samp != 0]
+datf[(act_dur != samp_dur) & t_btwn_samp == 0]
+
+hdat <- cbind("res" = log10(datf$plot_res), "ext" = log10(datf$act_ext), 
               "int" = log10(to_infin_byond(datf$t_btwn_samp)), 
-              "dur" = log10(datf$study_duration))
+              "dur" = log10(datf$act_dur))
 hdat <- data.table(hdat)
@@
   DT <- bootperturb[[x]]
-  hdb <- cbind("res" = log10(DT$plot_res), "ext" = log10(DT$sampled_area), 
+  hdb <- cbind("res" = log10(DT$plot_res), "ext" = log10(DT$act_ext), 
                "int" = log10(to_infin_byond(DT$t_btwn_samp)), 
-               "dur" = log10(DT$study_duration))
+               "dur" = log10(DT$act_dur))
   hdb <- data.table(hdb)
@@
 # dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
-datf[like(study_type, "field"), plot(log10(plot_res), log10(sampled_area), 
+datf[like(study_type, "field"), plot(log10(plot_res), log10(act_ext), 
                                      ylim = c(-10, 10))]
@@
 datf[like(study_type, "remote"), 
-     points(log10(plot_res), log10(sampled_area), pch = 20, col = "red")]
+     points(log10(plot_res), log10(act_ext), pch = 20, col = "red")]
 datf[like(study_type, "automated"), 
-     points(log10(plot_res), log10(sampled_area), pch = 20, col = "blue")]
+     points(log10(plot_res), log10(act_ext), pch = 20, col = "blue")]
 lines(aaxis1$logres, aaxis2$logarea[-16] )
@@
 
-datf[, plot(log10(t_btwn_samp), log10(study_duration), pch = 20)]
+datf[, plot(log10(t_btwn_samp), log10(act_dur), pch = 20)]
 datf[which(hdat[[3]] < -1.380211), 
-     points(log10(t_btwn_samp), log10(study_duration), pch = 20, col = "blue")]
+     points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "blue")]
 datf[like(study_type, "automated"), 
-     points(log10(t_btwn_samp), log10(study_duration), pch = 20, col = "red")]
+     points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "red")]
 
@@
 # extent
-extdat3 <- data.table(x = log10(kdat$study_duration), 
-                      y = log10(kdat$sampled_area))
+extdat3 <- data.table(x = log10(kdat$act_dur), 
+                      y = log10(kdat$act_ext))
 extdat3$n <- rep(1, nrow(extdat3))
@@
 tempdat <- data.table(x = log10(to_infin_byond(kdat$t_btwn_samp)), 
-                      y = log10(kdat$study_duration))
+                      y = log10(kdat$act_dur))
 tempdat <- tempdat[round(x, 8) != 6.56229286, ]  # remove one-time observations
@@
 # spatial framing
-spatdat <- data.table(x = log10(kdat$plot_res), y = log10(kdat$sampled_area))
+spatdat <- data.table(x = log10(kdat$plot_res), y = log10(kdat$act_ext))
 spatdat$n <- rep(1, nrow(spatdat))
@@
 
-# datf[, plot(log10(study_duration), log10(sampled_area), xlim)]
+# datf[, plot(log10(act_dur), log10(act_ext), xlim)]
 
@@
 
-+ If `plot_res` and/or `n_sites` uncertain, then `sampled_area` must be. Recalculate using random draw within the uncertain variable 
-+ If `t_btwn_samp` and/or `samp_duration` uncertain, do the same as above, but we have no variable for n repeats for calculating `study_duration`.
++ If `plot_res` and/or `n_sites` uncertain, then `act_ext` must be. Recalculate using random draw within the uncertain variable 
++ If `t_btwn_samp` and/or `samp_dur` uncertain, do the same as above, but we have no variable for n repeats for calculating `act_dur`.
 
 ```{r}
-svars <- c("plot_res", "n_sites", "sampled_area", "samp_duration",
-           "t_btwn_samp", "study_duration")
+svars <- c("plot_res", "n_sites", "act_ext", "samp_dur",
+           "t_btwn_samp", "act_dur")
 
@@
 
-# 1. if samp_duration OR t_btwn_samp is uncertain, study_duration must also be. 
+# 1. if samp_dur OR t_btwn_samp is uncertain, act_dur must also be. 
 id <- which((senst$sampd == 1 | senst$tbtwn == 1) & senst$studyd == 0)
@@
 
-# 2. if study_duration is uncertain but t_btwn and samp_duration are not, treat
+# 2. if act_dur is uncertain but t_btwn and samp_dur are not, treat
 # both as uncertain
@@
 
-# 3. if plot_res or n_sites is uncertain, then sampled_area must be
+# 3. if plot_res or n_sites is uncertain, then act_ext must be
 id <- which((senst$res == 1 | senst$n == 1) & senst$sarea == 0)
@@
   if((x / 100) %in% 1:100) print(x)
-  dnew <- copy(datf[, .(plot_res, sampled_area, t_btwn_samp, study_duration)])
+  dnew <- copy(datf[, .(plot_res, act_ext, t_btwn_samp, act_dur)])
   dnew <- dnew[sample(1:.N, size = .N, replace = TRUE), ]
@@
 # extent
-extdat2 <- data.table(x = log10(bootdt$study_duration), 
-                      y = log10(bootdt$sampled_area))
+extdat2 <- data.table(x = log10(bootdt$act_dur), 
+                      y = log10(bootdt$act_ext))
 extdat2$n <- rep(1, nrow(extdat2))
@@
 
-# datf[, plot(log10(study_duration), log10(sampled_area), xlim)]
+# datf[, plot(log10(act_dur), log10(act_ext), xlim)]
 ```
@@
 tbtwn <- ifelse(tbtwn == 0, 0.1 / (60 * 60 * 24), tbtwn)
-tempdat2 <- data.table(x = log10(tbtwn), y = log10(datf$study_duration))
+tempdat2 <- data.table(x = log10(tbtwn), y = log10(datf$act_dur))
 tempdat2[x == min(x), x := -5.5]
@@
 # tempdat2 <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
-#                       y = log10(datf$study_duration))
+#                       y = log10(datf$act_dur))
 spatdat2 <- data.table(x = log10(datf$plot_res / 10000), # convert to ha
-                       y = log10(datf$sampled_area))
+                       y = log10(datf$act_ext))
 sx2 <- round(range(spatdat2$x), 1)

### 2017-04-06 Rmd/space-time.Rmd
Plots revised through histogram (Figure 1).

@@
 library(readxl)
-library(stringdist)
+# library(stringdist)
 library(maptools)
@@
 dat <- as.data.table(read_excel(full_path(p_dat, "mergedfffff.xlsx")))
-dato <- as.data.table(read_excel(full_path(p_dat, "archive/mergedfff.xlsx")))
+# dato <- as.data.table(read_excel(full_path(p_dat, "archive/mergedfff.xlsx")))
 cal <- data.table(read_excel(fp(p_calib, "merged_calibrationfffff.xlsx")))
@@
 
-setnames(dato, "DOI/title", "DOI")
-dato[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
-dato[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
-
+# setnames(dato, "DOI/title", "DOI")
+# dato[, DOI := gsub("DOI:|doi:|DOI: |doi: |DOI ", "", DOI)]
+# dato[, DOI := gsub("\\s", "", DOI)]  # remove DOI and whitespace
+
+# clean up DOIs
 setnames(dat, "DOI/title", "DOI")
@@
 # kwords <- c("field", "paleo", "remote", "automated", "other")
-tst <- dat[, study_type]
+# tst <- dat[, study_type]
 
@@
 cal[, study_type := tolower(study_type)]
-tst <- cal[, study_type]
+# tst <- cal[, study_type]
 for(i in 1:nrow(kwords)) {
@@
                      samp_dur, t_btwn_samp, act_dur, eff_dur)])
+selind <- 3:ncol(dat2)
+for(j in names(dat2)[selind]) set(dat2, j=j, value = as.numeric(dat2[[j]]))
 rnms <- c("DOI", "study_type", "feature", "observer", "plot_res", 
@@
 datf[t_btwn_samp == 0 & (act_dur != samp_dur), ]  
-dat[t_btwn_samp == 0 & (act_dur != samp_dur), c(1, 3, 8:13), with=F]  
+dat[t_btwn_samp == 0 & (act_dur != samp_dur), c(1, 3, 8:13), with = F]  
 # function to set t_btwn_samp to 365 * 100000 to indicate once-off studies 
@@
 # p <- c(0.64, 0.84, 0.53, 1.21)  # from calibration.Rmd
-p <- c(0.58, 0.82, 0.99, 1.24)  # from calibration.Rmd
+# p <- c(0.58, 0.82, 0.99, 1.24)  # from calibration.Rmd
+##     plot_res   n_sites   act_ext   eff_ext t_btwn_samp  act_dur   eff_dur
+## 1: 0.5769693 0.3821107 0.7196559 0.4049972     1.01134 1.253045 0.5284913
+p <- c(0.58, 0.72, 0.41, 1.01, 1.25, 0.53)  # from calibration.Rmd
 
 # Function for bootstrapping with uncertainties
-resamp_func <- function(dat, p, iter) {
-  bout <- lapply(1:iter, function(x) {
+vars <- c("plot_res", "act_ext", "eff_ext", "t_btwn_samp", "act_dur", 
+          "eff_dur")
+
+resamp_func <- function(dat, vars, p, iter) {
+  bout <- lapply(1:iter, function(x) {  # x <- 1
     if((x / 100) %in% 1:100) print(x)
-    if(length(p) < 4) stop("p must be length of 4", call. = FALSE)
-
-    dnew <- copy(datf[, .(plot_res, act_ext, t_btwn_samp, act_dur)])
+    if(length(p) < length(var)) stop("p must be length of var", call. = FALSE)
+
+    # dnew <- copy(datf[, .(plot_res, act_ext, eff_ext, t_btwn_samp, act_dur)])
+    # dnew <- copy(datf[, vars, with = FALSE])
+    dnew <- copy(dat[, vars, with = FALSE])
     # modifiers
@@
     })
-
+    # plot(dnew[[j]] * modvec[, j], dnew[[j]])
+
+    # perturb and set any values < 0 to very small amount
     for(j in 1:ncol(modvec)) set(dnew, j = j, value = dnew[[j]] * modvec[, j])
@@
     dnew[act_ext <= 0, act_ext := 0.00000001]
+    dnew[eff_ext <= 0, eff_ext := 0.00000001]
     dnew[act_dur <= 0, act_dur := 0.00000001]
+    dnew[eff_dur <= 0, eff_dur := 0.00000001]
     dnew[t_btwn_samp < 0, t_btwn_samp := 0.00000001]
@@
 set.seed(1)
-bootperturb <- resamp_func(datf, p = p, iter = 1000)  # 10% perturb
+bootperturb <- resamp_func(datf, vars = vars, p = p, iter = 1000) # 10% perturb
 # bootperturb2 <- resamp_func(datf, p = p, iter = 1000)  # 10% perturb
@@
 # rm(bootperturb2)
-
+```
+
+```{r, echo = FALSE, eval=FALSE}
+# check
+par(mfrow = c(2, 3))
+for(i in sample(1:1000, 6)) {
+  plot(datf[, eff_ext], datf[, eff_ext] - bootperturb[[i]][, eff_ext], 
+       pch = 20)
+}
 ```
@@
 
-# sampled area
+# effective extent
+### fix this here to have label for 10^-6 (so that limit isn't hard-coded below)
 alab2 <- c(expression(paste(NULL<="0.1 m"^2)), expression("1 m"^2),
@@
 
+### fix this here to have value for 10^-6 (so that limit isn't hard-coded below)
 aaxis2 <- cbind.data.frame(
@@
 aaxis2[, logarea := log10(area)]
-
 ```
@@
 ```{r, eval=FALSE}
-length(which(datf$study_type %in% 
-               c("remote sensing", "other geographic data"))) / nrow(datf)
-datf[t_btwn_samp == 0, .N] / datf[, .N]  # 36% of samples are once-offs
+# length(which(datf$study_type %in% c("remote", "other"))) / nrow(datf)
+datf[t_btwn_samp == 0, .N] / datf[, .N]  # 38% of samples are once-offs
 datf[act_dur == samp_dur, .N] / nrow(datf)
@@
 
-hdat <- cbind("res" = log10(datf$plot_res), "ext" = log10(datf$act_ext), 
+hdat <- cbind("res" = log10(datf$plot_res), "aext" = log10(datf$act_ext),
+              "ext" = log10(datf$eff_ext), 
               "int" = log10(to_infin_byond(datf$t_btwn_samp)), 
-              "dur" = log10(datf$act_dur))
+              "adur" = log10(datf$act_dur), "dur" = log10(datf$eff_dur))
+# NA data checks
+# hdat[is.na(rowSums(hdat)), ]
+# dat[DOI %in% datf[which(is.na(rowSums(hdat))), DOI], ]
+# dat[DOI %in% datf[which(is.na(hdat[, 6])), DOI]]
+# dat[DOI %in% datf[which(is.na(hdat[, 3])), DOI]]
+# calr[DOI %in% datf[which(is.na(hdat[, 3])), DOI] & is.na(eff_ext)]
+
 hdat <- data.table(hdat)
+
 # set bounds for histograms
-hdat[res < aaxis1$logres[1], res := aaxis1$logres[1]]
-hdat[res > ry[2], res := ry[2]]
-hdat[ext < -6, ext := -6]  # set minimum extent -6 (0.01 m2)
+# first set variable limits
+vlr <- range(aaxis1$logres)  #for resolution
+### fix in scaling section (so that min extent limit isn't hard-coded)
+vle <- c(-6, range(aaxis2$logarea)[2])  ### fix
+
+# then apply limits to main (unperturbed) set of variables
+hdat[res < vlr[1], res := vlr[1]]  # set min res to 0.01 cm"^2 
+hdat[res > vlr[2], res := vlr[2]]  # set max res to 10000 ha 
+hdat[ext < vle[1], ext := vle[1]]  # set minimum extent -6 (0.01 m2) 
+hdat[ext > vle[2], ext := vle[2]]  # set maximum extent 10 (10^10 ha)
+# hdat[aext < vle[1], aext := vle[1]]  # set minimum extent -6 (0.01 m2) 
+# hdat[aext > vle[2], aext := vle[2]]  # set maximum extent 10 (10^10 ha)
 
 # histograms from bootstrap
+### fix in scaling section (so that min extent limit isn't hard-coded)
 brksl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
@@
   DT <- bootperturb[[x]]
-  hdb <- cbind("res" = log10(DT$plot_res), "ext" = log10(DT$act_ext), 
+  hdb <- cbind("res" = log10(DT$plot_res), "ext" = log10(DT$eff_ext), 
                "int" = log10(to_infin_byond(DT$t_btwn_samp)), 
-               "dur" = log10(DT$act_dur))
+               "dur" = log10(DT$eff_dur))
   hdb <- data.table(hdb)
-  hdb[res < aaxis1$logres[1], res := aaxis1$logres[1]]
-  hdb[res > ry[2], res := ry[2]]
-  hdb[, lapply(.SD, range)]
-  hdb[ext < -6, ext := -6]  # set minimum extent -6 (0.01 m2)
+  hdb[res < vlr[1], res := vlr[1]]  # set min res to 0.01 cm"^2 
+  hdb[res > vlr[2], res := vlr[2]]  # set max res to 10000 ha 
+  # hdb[, lapply(.SD, range, na.rm = TRUE)]
+  ### fix in scaling section (so that this limit isn't hard-coded)
+  hdb[ext < vle[1], ext := vle[1]]  # set minimum extent -6 (0.01 m2)
+  hdb[ext > vle[2], ext := vle[2]]  # set maximum extent 10 (10^10 ha)
   hists <- lapply(1:length(vs), function(y) {  # y <- 2
@@
 plot(1:2, pch = 20, col = blues[c(4, 6)])
-yl <- 35
+yl <- c(25, 40)
 # pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
-pdf("paper/figures/hists3.pdf", width = 5, height = 5)
+pdf("paper/figures/hists.pdf", width = 5, height = 5)
 par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
@@
      tcl = -0.2, cex.axis = cxa, xlab = "Spatial resolution", 
-     ylim = c(0, yl - 10), main = "", col = reds[4], cex.lab = cxl)
+     ylim = c(0, yl[1]), main = "", col = reds[4], cex.lab = cxl)
 axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
@@
 # extent
+### fix in scaling section (so that this limit isn't hard-coded)
 h <- hist(hdat$ext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
+# h <- hist(hdat$aext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
 h$density <- h$counts / sum(h$counts) * 100
 plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Spatial extent", ylim = c(0, yl - 10), 
+     tcl = -0.2, cex.axis = cxa, xlab = "Spatial extent", ylim = c(0, yl[1]), 
      main = "", col = reds[6], cex.lab = cxl)
+### fix in scaling section (so that this limit isn't hard-coded)
 axis(1, at = c(-6, aaxis2$logarea),  
@@
 plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Sampling interval", ylim = c(0, yl), 
+     tcl = -0.2, cex.axis = cxa, xlab = "Sampling interval", ylim = c(0, yl[2]), 
      main = "", col = blues[4], cex.lab = cxl)
@@
 plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Temporal duration", ylim = c(0, yl), 
+     tcl = -0.2, cex.axis = cxa, xlab = "Temporal duration", ylim = c(0, yl[2]), 
      main = "", col = blues[6], cex.lab = cxl)
@@
 #              "tax1" = taxis1$logdays, "tax2" = taxis2$logdays)
+### fix in scaling section (so that this limit isn't hard-coded)
 axes2 <- list("aax1" = aaxis1$logres, "aax2" = c(-6, aaxis2$logarea), 
@@
               "tax2" = taxis2$logdays)
+### fix in scaling section (so that this limit isn't hard-coded)
 axlabs <- list(alab1, c(expression(paste(NULL<="0.01m"^2)),
@@
     # base plot
-  yl <- ifelse(i %in% c(1, 2), 30, 35) 
-  plot(c(axv[1], axv[length(axv)]), c(0, yl - 10), ylab = "", las = 2, 
+  yl <- ifelse(i %in% c(1, 2), 25, 40) 
+  plot(c(axv[1], axv[length(axv)]), c(0, yl), ylab = "", las = 2, 
          xaxt = "n", xlab = xlabs[i], ylim = c(0, yl), main = "", 
@@
 #               taxis2$logdays)
-hdatstat <- lapply(1:4, function(x) {
-  h <- hist(hdat[[x]], breaks = brksl[[x]], plot = FALSE)
-  h$density <- h$counts / sum(h$counts) * 100
-  h
-})
+# hdatstat <- lapply(1:4, function(x) {
+#   h <- hist(hdat[[x]], breaks = brksl[[x]], plot = FALSE)
+#   h$density <- h$counts / sum(h$counts) * 100
+#   h
+# })
 
@@
 
+##############
+# Ended here 6/4/2017 
+##############
 # extent

### 2017-05-06 Rmd/calibration.Rmd
Various updates towards revision, including new versions of figure 3

@@
 # cal <- data.table(read_excel(fp(p_dat, "merged_calibrationff.xlsx")))
-cal <- data.table(read_excel(fp(p_dat, "merged_calibrationfffff.xlsx")))
+cal <- data.table(read_excel(fp(p_dat, "merged_calibrationffffff.xlsx")))
 names(cal)
@@
 
-par(mfrow = c(3, 3), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
-for(i in varnms) {
-  mu <- musmu[, get(i)]
-  sdev <- vars[, get(i)]
-#   print(paste(i, ":", sdev))
-  print(paste(i, ":", round(mu, 2), ":", round(sdev, 2), ":", 
-              round(sdev / mu, 2)))
-  # plot(mu, sdev / mu, xlab = "mean", ylab = "CV", main = i)
-  plot(mu, sdev, xlab = "mean", ylab = "sd", main = i)
-}
+# par(mfrow = c(3, 3), mar = c(3, 3, 2, 2), mgp = c(1.5, 0.5, 0))
+# for(i in varnms) {
+#   mu <- musmu[, get(i)]
+#   sdev <- vars[, get(i)]
+# #   print(paste(i, ":", sdev))
+#   print(paste(i, ":", round(mu, 2), ":", round(sdev, 2), ":", 
+#               round(sdev / mu, 2)))
+#   # plot(mu, sdev / mu, xlab = "mean", ylab = "CV", main = i)
+#   plot(mu, sdev, xlab = "mean", ylab = "sd", main = i)
+# }
 
 vars[, lapply(.SD, mean, na.rm = TRUE), .SDcols = varnms]
-colMeans(sapply(varnms, function(i) {
+round(colMeans(sapply(varnms, function(i) {
   mu <- musmu[, get(i)]
@@
   sdev / mu
-}), na.rm = TRUE)
+}), na.rm = TRUE), 3)
 
@@
 # calr[feature != 99, .N, by = .(DOI, feature)]
+mymu <- function(x) round(mean(x, na.rm = TRUE), 3)
 calr[feature != 99, lapply(.SD, CV), by = .(DOI, feature), 
-     .SDcols = rnms[4:10]][, lapply(.SD, mean, na.rm = TRUE), 
+     .SDcols = rnms[4:10]][, lapply(.SD, mymu), 
                           .SDcols = rnms[4:10]]

### 2017-05-06 Rmd/space-time.Rmd
Various updates towards revision, including new versions of figure 3

@@
 library(RColorBrewer)
+library(viridis)
 
@@
 # dat <- as.data.table(read_excel(full_path(p_dat, "merged.xlsx")))
-dat <- as.data.table(read_excel(full_path(p_dat, "mergedfffff.xlsx")))
+dat <- as.data.table(read_excel(full_path(p_dat, "mergedffffff.xlsx")))
 # dato <- as.data.table(read_excel(full_path(p_dat, "archive/mergedfff.xlsx")))
-cal <- data.table(read_excel(fp(p_calib, "merged_calibrationfffff.xlsx")))
+cal <- data.table(read_excel(fp(p_calib, "merged_calibrationffffff.xlsx")))
 full <- data.table(read_excel(fp(p_dat, "full_no_treuer_toss.xlsx")))
@@
 datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
+datf[act_dur > eff_dur]
 doisup <- c(unique(cal$DOI_data_source), unique(dat$DOI_data_source))
@@
 ### Adjustments/fixes
-1. if `samp_dur != act_dur` and `t_btwn_samp == 0`, set `act_dur == samp_dur` (actually, no, in many cases there is a good reason to have to `act_dur != samp_dur` (see below))
+1.
+One could argue that if `samp_dur != act_dur` and `t_btwn_samp == 0`, set `act_dur == samp_dur`, but in a small number of cases there is a good reason to have to `act_dur != samp_dur` (e.g.
+DOI 10.1890/08-0611.1 when farmers were asked about their seasonal management practices--time period of interview was clearly != to information being collected))
 2.
@@
 datf[t_btwn_samp == 0 & (act_dur != samp_dur), ]  
-dat[t_btwn_samp == 0 & (act_dur != samp_dur), c(1, 3, 8:13), with = F]  
+dat[t_btwn_samp == 0 & (act_dur != samp_dur), c(1, 3:4, 9:15), with = FALSE]  
 # function to set t_btwn_samp to 365 * 100000 to indicate once-off studies 
 # when t_btwn_samp == 0.  
-to_infin_byond <- function(tbtwn) ifelse(tbtwn == 0, 365 * 10000, tbtwn)
-
+dat[act_dur < 1 & t_btwn_samp > 1000, ]
+dat[act_dur > 1 & t_btwn_samp == 0, c(1, 3:4, 9:15), with = FALSE]
 ```
@@
 ## 1: 0.5769693 0.3821107 0.7196559 0.4049972     1.01134 1.253045 0.5284913
-p <- c(0.58, 0.72, 0.41, 1.01, 1.25, 0.53)  # from calibration.Rmd
-
-# Function for bootstrapping with uncertainties
-vars <- c("plot_res", "act_ext", "eff_ext", "t_btwn_samp", "act_dur", 
-          "eff_dur")
-
+p <- c(0.58, 0.75, 0.50, 1.01, 1.23, 0.59)  # from calibration.Rmd
+
+# Functions for bootstrapping with uncertainties
+# sets unreplicated observations to arbitrarily large number
+to_infin_byond <- function(tbtwn) ifelse(tbtwn == 0, 365 * 10000, tbtwn)
+
+# bootstrapping function
 resamp_func <- function(dat, vars, p, iter) {
@@
 
+# Run bootstrap
+vars <- c("plot_res", "act_ext", "eff_ext", "t_btwn_samp", "act_dur", "eff_dur")
 set.seed(1)
@@
 }
+dev.off()
 ```
@@
 # effective extent
-### fix this here to have label for 10^-6 (so that limit isn't hard-coded below)
-alab2 <- c(expression(paste(NULL<="0.1 m"^2)), expression("1 m"^2),
-           expression("10 m"^2), 
+alab2 <- c(expression(paste(NULL<="0.01 m"^2)), expression(paste("0.1 m"^2)), 
+           expression("1 m"^2), expression("10 m"^2), 
            expression("100 m"^2), expression("1000 m"^2), expression("1 ha"),
@@
 
-### fix this here to have value for 10^-6 (so that limit isn't hard-coded below)
 aaxis2 <- cbind.data.frame(
-  c(0.1 / 10000, 1 / 10000, 10 / 10000, 100 / 10000, 1000 / 10000, 1, 10, 10^2,
-    10^3, 10^4, 10^5, 10^6, 10^7, 10^8, 10^9, 10^10))
+  c(0.01 / 10000, 0.1 / 10000, 1 / 10000, 10 / 10000, 100 / 10000, 1000 / 10000,
+    1, 10, 10^2, 10^3, 10^4, 10^5, 10^6, 10^7, 10^8, 10^9, 10^10))
 aaxis2 <- data.table(aaxis2)
@@
 
+Histograms for each of the four main dimensions.
 ```{r, eval=FALSE}
 # length(which(datf$study_type %in% c("remote", "other"))) / nrow(datf)
-datf[t_btwn_samp == 0, .N] / datf[, .N]  # 38% of samples are once-offs
+datf[t_btwn_samp == 0, .N] / datf[, .N]  # 37% of samples are once-offs
 datf[act_dur == samp_dur, .N] / nrow(datf)
@@
 # first set variable limits
-vlr <- range(aaxis1$logres)  #for resolution
-### fix in scaling section (so that min extent limit isn't hard-coded)
-vle <- c(-6, range(aaxis2$logarea)[2])  ### fix
+vlr <- range(aaxis1$logres)  # for resolution
+vle <- range(aaxis2$logarea)  # for extent
 
@@
 # histograms from bootstrap
-### fix in scaling section (so that min extent limit isn't hard-coded)
-brksl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
-              taxis2$logdays)
-
+brksl <- list(aaxis1$logres, aaxis2$logarea, taxis1$logdays, taxis2$logdays)
+
+# Assign bootstrapped values to bins, making necessary adjustments as needed 
 vs <- c("res", "ext", "int", "dur")
@@
   # hdb[, lapply(.SD, range, na.rm = TRUE)]
-  ### fix in scaling section (so that this limit isn't hard-coded)
   hdb[ext < vle[1], ext := vle[1]]  # set minimum extent -6 (0.01 m2)
@@
 
+# convert density values for each dimension to data.tables
 hdat_bsdt <- lapply(1:length(vs), function(x) { # x <- 1
@@
 blues <- brewer.pal(9, name = "Blues")
-plot(1:2, pch = 20, col = blues[c(4, 6)])
+# plot(1:2, pch = 20, col = blues[c(4, 6)])
 yl <- c(25, 40)
@@
 # extent
-### fix in scaling section (so that this limit isn't hard-coded)
-h <- hist(hdat$ext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
+h <- hist(hdat$ext, breaks = aaxis2$logarea, plot = FALSE)
 # h <- hist(hdat$aext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
@@
      main = "", col = reds[6], cex.lab = cxl)
-### fix in scaling section (so that this limit isn't hard-coded)
-axis(1, at = c(-6, aaxis2$logarea),  
-     labels = c(expression(paste(NULL<="0.01m"^2)), expression(paste(0.1, m^2)),                 alab2[-1]), las = 2, cex.axis = cxa,
+axis(1, at = aaxis2$logarea,
+     labels = alab2,
+     # labels = c(expression(paste(NULL<="0.01m"^2)), expression(paste(0.1, m^2)),                 alab2[-1]), 
+     las = 2, cex.axis = cxa,
      tcl = -0.2, mgp = c(2, 0.25, mga))
@@
 plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Temporal duration", ylim = c(0, yl[2]), 
-     main = "", col = blues[6], cex.lab = cxl)
+     tcl = -0.2, cex.axis = cxa, xlab = "Temporal duration", ylim = c(0, yl[2]),      main = "", col = blues[6], cex.lab = cxl)
 axis(1, at = taxis2$logdays, labels = tlab2, las = 2, cex.axis = cxa, 
@@
 mtext("D", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
-
 dev.off()
@@
 #              "tax1" = taxis1$logdays, "tax2" = taxis2$logdays)
-### fix in scaling section (so that this limit isn't hard-coded)
-axes2 <- list("aax1" = aaxis1$logres, "aax2" = c(-6, aaxis2$logarea), 
+axes2 <- list("aax1" = aaxis1$logres, "aax2" = aaxis2$logarea, 
               "tax1" = c(taxis2$logdays[-c(10:11)],
@@
               "tax2" = taxis2$logdays)
-### fix in scaling section (so that this limit isn't hard-coded)
-axlabs <- list(alab1, c(expression(paste(NULL<="0.01m"^2)),
-                        expression(paste(0.1, m^2)), alab2[-1]),
-               tlab1[-10], tlab2)
+# axlabs <- list(alab1, c(expression(paste(NULL<="0.01m"^2)),
+#                         expression(paste(0.1, m^2)), alab2[-c(1:2)]),
+axlabs <- list(alab1, alab2, tlab1[-10], tlab2)
 mga <- -0.34
@@
 # })
-
+```
+
+Statistics associated with Figure 1: percentage of observations falling within different scale ranges, and associated with different observational methods
+```{r}
 # res
@@
 
-##############
-# Ended here 6/4/2017 
-##############
 # extent
@@
                                      ylim = c(-10, 10))]
-
 datf[like(study_type, "remote"), 
@@
      points(log10(plot_res), log10(act_ext), pch = 20, col = "blue")]
-lines(aaxis1$logres, aaxis2$logarea[-16] )
+lines(aaxis1$logres, aaxis2$logarea[-c(16:17)] )
 datf[, .N / nrow(datf) * 100, by = st]
@@
 datf[like(study_type, "auto"), .N] / nrow(datf) # 12 % automated
-datf[like(study_type, "auto") & log10(t_btwn_samp) < -1.380211, .N] /   length(which(hdat[[3]] < -1.380211))  # 75%
-
-datf[like(study_type, "remote|other"), .N] / nrow(datf) # 6 % automated
+datf[like(study_type, "auto") & log10(t_btwn_samp) < -1.380211, .N] / 
+  length(which(hdat[["int"]] < -1.380211))  # 78 % high frequency obs automated 
+datf[like(study_type, "remote|other"), .N] / nrow(datf) # 7 % remote
 datf[like(study_type, "paleo"), .N] / nrow(datf) * 100 # <1 % paleo
@@
                       y = log10(kdat$plot_res))
-resdat3 <- resdat3[round(x, 8) != 6.56229286, ]  # remove one-time observations
-resdat3$n <- rep(1, nrow(resdat3))
+resdat3 <- resdat3[round(x, 8) != 6.56229286]  # remove one-time observations
+resdat3[, n := 1]  # set up count column
+# resdat3$n <- rep(1, nrow(resdat3))
 rx <- range(resdat3$x)
 rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-rx[1] <- -5
+rx[1] <- -5  # set min interval to less than second
 #rx[2] <- 6.56
-rx[2] <- 5.57
+rx[2] <- 5.57  # set max interval to millenium
 resdat3[x < rx[1], x := rx[1]]
@@
 # extent
-extdat3 <- data.table(x = log10(kdat$act_dur), 
-                      y = log10(kdat$act_ext))
-extdat3$n <- rep(1, nrow(extdat3))
+extdat3 <- data.table(x = log10(kdat$eff_dur),#log10(kdat$act_dur), 
+                      y = log10(kdat$eff_ext))#log10(kdat$act_ext))
+extdat3 <- extdat3[!is.na(x) & !is.na(y)] # cut out null values
+# extdat3$n <- rep(1, nrow(extdat3))
+extdat3[, n := 1]  # set up count column
 ex <- range(taxis2$logdays)
-ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
+ex <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
 extdat3[x < ex[1], x := ex[1]]
@@
 tempdat <- data.table(x = log10(to_infin_byond(kdat$t_btwn_samp)), 
-                      y = log10(kdat$act_dur))
+                      y = log10(kdat$eff_dur))#kdat$act_dur))
+# tt <- data.table(x = log10(kdat$t_btwn_samp), y = log10(kdat$eff_dur))
+# range(tt$x - tempdat$x)
+tempdat <- tempdat[!is.na(x) & !is.na(y)] # cut out null values
 tempdat <- tempdat[round(x, 8) != 6.56229286, ]  # remove one-time observations
-tempdat$n <- rep(1, nrow(tempdat))
+# tempdat$n <- rep(1, nrow(tempdat))
+tempdat[, n := 1]
 tx <- range(tempdat$x)
@@
 # spatial framing
-spatdat <- data.table(x = log10(kdat$plot_res), y = log10(kdat$act_ext))
-spatdat$n <- rep(1, nrow(spatdat))
+spatdat <- data.table(x = log10(kdat$plot_res), y = log10(kdat$eff_ext))#act_ext))
+spatdat <- spatdat[!is.na(x) & !is.na(y)] # cut out null values
+# spatdat$n <- rep(1, nrow(spatdat))
+spatdat[, n := 1]
 sx <- range(aaxis1$logres)
@@
 cexs <- c(0.4, 0.6)
-cuts <- 60
+cuts <- 40
+bump <- 3
+cols <- inferno(cuts + bump)[-c(2:(bump + 1))]
+# plot(1:cuts, col = cols, pch = 20)
 
 # pdf("paper/figures/res_v_extent.pdf", width = 7, height = 2.5)
-pdf("paper/figures/kde45.pdf", width = 7, height = 5.5)
+pdf("paper/figures/kde45.pdf", width = 7, height = 6)
 par(mfrow = c(2, 2), mar = c(6, 6, 0.3, 3), oma = c(0, 0, 0.5, 1), 
@@
 # image(rres3, col = rev(terrain.colors(30)), breaks = brks,
-image(rres3, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
-      ylab = "")
+image(rres3, axes = FALSE, xlab = "", ylab = "", #rev(terrain.colors(cuts))) 
+      col = cols)
 axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
-# axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
      las = 2, tcl = -0.2, cex.axis = cxa)
@@
      #legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
-     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
-mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05)
+     legend.shrink = 0.9, col = cols)#inferno(cuts))
+     #col = rev(terrain.colors(cuts)))
+mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
 # for(i in 1:length(stps)) {
@@
 # image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
-image(rext3, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
-      ylab = "")#, breaks = brks2)
+image(rext3, axes = FALSE, xlab = "", ylab = "", col = cols)
 axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
@@
 plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     #legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
-     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
-mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05)
-# for(i in 1:length(stps)) {
-#   points(extdat3$x[stps[[i]]], extdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
-# }
+     legend.shrink = 0.9, col = cols)
+mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
 
 # spatial
-image(sres, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", 
-      ylab = "")
-# points(spatdat$x, spatdat$y, cex = 0.1)
+image(sres, col = cols, axes = FALSE, xlab = "", ylab = "") 
 axis(1, at = aaxis1$logres, labels = alab1, las = 2, tcl = -0.2, cex.axis = cxa)
@@
 plot(sres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
-mtext("C", side = 3, line = -1, cex = 0.8, adj = 0.05)
+     legend.shrink = 0.9, col = cols)
+mtext("C", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
 # for(i in 1:length(stps)) {
@@
 # temporal
-image(tres, col = rev(terrain.colors(cuts)), axes = FALSE, xlab = "", ylab = "")
-# points(spatdat$x, spatdat$y, cex = 0.1)
+image(tres, axes = FALSE, xlab = "", ylab = "", col = cols)
 axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
-# axis(1, at = taxis1$logdays[-c(5, 10)], labels = tlab1[-c(5, 10)], 
      las = 2, tcl = -0.2, cex.axis = cxa)
@@
 mtext("Temporal duration", 2, cex = cxa, line = ll)
-brks <- brkfun(tres, 0.02, cuts)  # breaks
+# brks <- brkfun(tres, 0.02, 100)  # breaks
+brks <- brkfun(tres, 0.01, cuts)  # breaks
 aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
@@
 plot(tres, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, col = rev(terrain.colors(cuts)))
-mtext("D", side = 3, line = -1, cex = 0.8, adj = 0.05)
-# for(i in 1:length(stps)) {
-#   points(tempdat$x[stps[[i]]], tempdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
-# }
+     legend.shrink = 0.9, col = cols)
+mtext("D", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
 dev.off()
-
-# datf[, plot(log10(act_dur), log10(act_ext), xlim)]
 
@@
 <a href="#top">Back to top</a>
+
+### Figure 3
+```{r}
+datf[, plot(log10(eff_ext), log10(act_ext))]
+datf[, plot(log10(act_dur), log10(eff_dur))]
+datf[log10(eff_dur) < log10(act_dur)]
+calr[log10(eff_dur) < log10(act_dur)]
+dat[DOI == "10.1007/s10021-013-9684-x"]
+cal[DOI == "10.1007/s00442-010-1713-2"]
+dat[DOI %in% datf[log10(eff_ext) < log10(act_ext), DOI]]
+dat[DOI %in% datf[log10(eff_ext) < log10(act_ext), DOI]]
+datf[log10(eff_ext) < log10(act_ext), ]
+datf[DOI == "10.1111/j.1466-8238.2011.00658.x", 
+     .(log10(round(eff_ext)), log10(round(act_ext)))]
+datf[DOI == "10.1111/j.1466-8238.2011.00658.x", 
+     .(round(eff_ext), round(act_ext))]
+
+# actual versus effective duration
+kdat <- bootp_dt  # bootstrap
+# tempdat2 <- data.table(x = log10(kdat$act_dur), y = log10(kdat$eff_dur))
+tempdat2 <- data.table(x = log10(kdat$eff_dur), y = log10(kdat$act_dur))
+dinds <- kdat[, which(t_btwn_samp == 365 * 10000)]  # drop unreplicated obs
+tempdat2 <- tempdat2[-dinds, ]
+tempdat2 <- tempdat2[!is.na(x) & !is.na(y)] # cut out null values
+tempdat2[, n := 1]  # set up count column
+tx <- range(taxis2$logdays)
+tx <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
+tempdat2[x < tx[1], x := tx[1]]
+tempdat2[x > tx[2], x := tx[2]]
+ty <- range(taxis2$logdays)
+ty <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
+tempdat2[y < tx[1], y := tx[1]]
+tempdat2[y > tx[2], y := tx[2]]
+coordinates(tempdat2) <- ~x + y
+
+# extent
+# extdat4 <- data.table(x = log10(kdat$act_ext), y = log10(kdat$eff_ext))
+extdat4 <- data.table(x = log10(kdat$eff_ext), y = log10(kdat$act_ext))
+extdat4 <- extdat4[!is.na(x) & !is.na(y)] # cut out null values
+extdat4[, n := 1]  # set up count column
+ex <- range(aaxis2$logarea)
+extdat4[x < ex[1], x := ex[1]]
+extdat4[x > ex[2], x := ex[2]]
+ey <- range(aaxis2$logarea)
+extdat4[y < ey[1], y := ey[1]]
+extdat4[y > ey[2], y := ey[2]]
+coordinates(extdat4) <- ~x + y
+
+# densities
+# temporal 
+tres2 <- kdensity(tx[1], tx[2], ty[1], ty[2], 0.1, tempdat2, 1)
+tres2 <- (tres2 / cellStats(tres2, sum)) * 100
+
+# extent
+rext4 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat4, 1)
+rext4 <- (rext4 / cellStats(rext4, sum)) * 100
+
+# cxa = 0.7
+ll <- 3.5
+cxa = 0.7
+stps <- sapply(c(3, 4), function(x) which(kdat$st == x))  # IDs rs/auto/paleo
+pchs <- list(1, "+")
+cexs <- c(0.4, 0.6)
+cuts <- 40
+bump <- 3
+cols <- inferno(cuts + bump)[-c(2:(bump + 1))]
+# plot(1:cuts, col = cols, pch = 20)
+
+pdf("paper/figures/act_v_eff.pdf", width = 7, height = 3)
+par(mfrow = c(1, 2), mar = c(4.5, 4.5, 0.3, 3), oma = c(0, 0, 0.2, 1), 
+    mgp = c(2, 0.25, 0))
+
+mtl <- 3
+# spatial extents
+image(rext4, axes = FALSE, xlab = "", ylab = "", col = cols)
+axis(1, at = aaxis2$logarea, labels = alab2, las = 2,
+     cex.axis = cxa, tcl = -0.2)
+axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
+     tcl = -0.2)
+brks <- brkfun(rext4, 0.01, cuts)  # breaks
+aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
+              cex.axis = cxa, tcl = -0.1)
+mtext("Actual extent", 2, cex = cxa, line = mtl)
+mtext("Effective extent", 1, cex = cxa, line = mtl)
+plot(rext4, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, col = cols)
+mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
+
+# temporal extents
+image(tres2, axes = FALSE, xlab = "", ylab = "", col = cols)
+axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
+     las = 2, tcl = -0.2, cex.axis = cxa)
+axis(2, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], las = 2, tcl = -0.2, 
+     cex.axis = cxa)
+mtext("Actual duration", 2, cex = cxa, line = mtl)
+mtext("Effective duration", 1, cex = cxa, line = mtl)
+brks <- brkfun(tres2, 0.01, cuts)  # breaks
+aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
+              cex.axis = cxa, tcl = -0.1)
+plot(tres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+     legend.shrink = 0.9, col = cols)
+     #col = rev(terrain.colors(cuts)))
+mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
+dev.off()
+
+# par(mfrow = c(1, 2))
+# datf[, {
+#   # sel = which()
+#   eext = log10(eff_ext)
+#   aext = log10(act_ext)
+#   edur = log10(eff_dur[t_btwn_samp != 365 * 10000])
+#   adur = log10(act_dur[t_btwn_samp != 365 * 10000])
+#   plot(aext, eext - aext)
+#   plot(adur, edur - adur)
+#   length(edur)
+# }]
+
+# Bootstrap version
+vs1 <- c("eff_ext", "act_ext", "eff_dur", "act_dur")
+vs2 <- c("eext", "aext", "edur", "adur")
+hdat_bs2 <- lapply(1:length(bootperturb), function(x) {  # x <- 1
+  # print(x)
+  DT <- bootperturb[[x]]
+  edb <- DT[, lapply(.SD, log10), .SDcol = c("eff_ext", "act_ext")]
+  # edb <- edb[eff_ext >= act_ext, ]  # drop eff_dur < act
+  setnames(edb, c("eff_ext", "act_ext"), c("eext", "aext"))
+  edb[, dif := eext - aext]
+  # elm <- edb[, lm(dif ~ aext)]
+  # preds <- predict(elm, newdata = data.frame(aext = extdum))
+  # plot(extdum, preds)
+
+  tdb <- DT[t_btwn_samp != 365 * 10000, lapply(.SD, log10), 
+            .SDcol = c("eff_dur", "act_dur")]
+  # tdb <- tdb[eff_dur >= act_dur, ]  # drop eff_dur < act
+  setnames(tdb, c("eff_dur", "act_dur"), c("edur", "adur"))
+  print(paste(edb[, .N], tdb[, .N]))
+  tdb[, dif := edur - adur]
+  # tlm <- tdb[, summary(lm(dif ~ adur))]
+  return(list("edb" = edb, "tdb" = tdb, #"elm" = elm, "tlm" = tlm, 
+              "Ns" = c(edb[, .N], tdb[, .N])))
+})
+
+# check number of drops due to eff < act
+# hdat_bs2_ns <- do.call(rbind, lapply(hdat_bs2, function(x) x$Ns))
+# summary(bootperturb[[1]][, .N] - hdat_bs2_ns[, 1])
+# summary(bootperturb[[1]][t_btwn_samp != 365 * 10000, .N] - hdat_bs2_ns[, 2])
+
+# extract regression coefficients
+# hdat_bs2_elm <- do.call(rbind, lapply(hdat_bs2, function(x) coef(x$elm)[, 1]))
+# hdat_bs2_tlm <- do.call(rbind, lapply(hdat_bs2, function(x) coef(x$tlm)[, 1]))
+
+# calculate box statistics
+# Extent differences
+eint <- aaxis2$logarea
+edif_int <- rbindlist(lapply(hdat_bs2, function(y) {  # y <- hdat_bs2[[2]]
+  dd <- copy(y$edb[!is.na(aext)])
+  dd[aext < ex[1], aext := ex[1]]  # set to min
+  # dd[, list("int" = findInterval(aext, eint), "dif" = dif)]
+  dd[, int := findInterval(aext, eint)]
+  # dd[, list("dif" = mean(dif, na.rm = TRUE), "N" = .N), by = int][order(int)]
+}))
+
+ints <- sort(edif_int[, unique(int)])
+ebox <- lapply(ints, function(x) edif_int[int == x, lmisc::box_stats(dif)])
+pctobs <- sapply(ints, function(x) edif_int[int == x, .N])
+ewgt <- pctobs / sum(pctobs)
+ewgt * 100
+ealphs <- (ewgt - min(ewgt)) / diff(range(ewgt))
+
+# temporal differences
+dint <- taxis2$logdays[-5]
+tx <- range(dint)
+ddif_int <- rbindlist(lapply(hdat_bs2, function(y) {  # y <- hdat_bs2[[2]]
+  dd <- copy(y$tdb[!is.na(adur)])
+  dd[adur < tx[1], adur := tx[1]]  # set to min
+  dd[adur > tx[2], adur := tx[2]]  # set to max
+  dd[, int := findInterval(adur, dint)]
+}))
+
+dints <- sort(ddif_int[, unique(int)])
+tbox <- lapply(dints, function(x) ddif_int[int == x, lmisc::box_stats(dif)])
+pctobs <- sapply(dints, function(x) ddif_int[int == x, .N])
+dwgt <- pctobs / sum(pctobs)
+dalphs <- (dwgt - min(dwgt)) / diff(range(dwgt))
+
+pdf("paper/figures/act_v_eff_diff.pdf", width = 7, height = 4)
+par(mfrow = c(1, 2), mar = c(4, 0, 1, 0.75), oma = c(0, 3, 0, 0))
+bx <- rowMeans(cbind(eint[-length(eint)], eint[-1]))
+cols <- sapply(ealphs, function(x) rgb(0, 0.1, 1, alpha = x))
+# plot(1:length(cols), col = lcols, pch = 20)
+plot(ex, c(0, 14), pch = "", xlab = "", ylab = "", xaxt = "n", yaxt = "n",
+     mgp = c(3, 0.5, 0), tcl = -0.2, xaxs = "i")
+# grid(lty = 1, col="gray")
+# polygon(c(ex, rev(ex), ex[1]), c(-1, -1, 15, 15, -1), col = "grey90")
+for(i in 1:length(ebox)) {
+  boxplot_v(bx[i], y = ebox[[i]], n = 1.75, whiskhgt = 1.5, bfill = cols[i])
+} 
+axis(1, at = eint, labels = alab2, las = 2, tcl = -0.2, cex.axis = cxa, 
+     mgp = c(3, 0.5, 0))
+axis(2, at = seq(0, 14, 2), las = 2, tcl = -0.2, cex.axis = cxa, 
+     mgp = c(3, 0.5, 0))
+mtext("Actual Extent", side = 1, line = 3)
+mtext("Difference Magnitude:", side = 2, line = 2)
+mtext("Effective minus Actual", side = 2, line = 1)
+# xs <- seq(2, 10, by = (10 - 2) / (length(ealphs)))
+# ealphss <- sort(ealphs, decreasing = TRUE)
+# lcols <- sapply(ealphss, function(x) rgb(0, 0.1, 1, alpha = x))
+# rect_shade(xs, y = c(13, 14), fillcol = lcols, linecol = "grey")
+# text(xs, rep(13, length(xs)), sort(round(ewgt * 100), decreasing = TRUE),
+#      srt = 90)
+
+bx <- rowMeans(cbind(dint[-length(dint)], dint[-1]))
+bd <- abs(dint[-length(dint)] - dint[-1])
+cols <- sapply(dalphs, function(x) rgb(1, 0, 0.1, alpha = x))
+# plot(1:length(cols), col = cols, pch = 20)
+plot(tx, c(0, 14), pch = "", xlab = "", ylab = "", xaxt = "n", 
+     mgp = c(3, 0.5, 0), tcl = -0.2, yaxt = "n")
+for(i in 1:length(tbox)) {
+  boxplot_v(bx[i], y = tbox[[i]], n = 0.9, inhgt = bd[i] * 10, whiskhgt = 0.5, 
+            bfill = cols[i])
+} 
+axis(1, at = dint, labels = tlab2[-c(5)], las = 2, tcl = -0.2, cex.axis = cxa, 
+     mgp = c(3, 0.5, 0))
+axis(2, at = seq(0, 14, 2), labels = rep("", 8), las = 2, tcl = -0.2, 
+     cex.axis = cxa, mgp = c(3, 0.5, 0))
+mtext("Actual Duration", side = 1, line = 3)
+dev.off()
+
+# # Calculate predictions from ensemble of regression coefficients
+# extdum <- seq(-10, 10, 1)
+# extpred <- do.call(rbind, lapply(1:nrow(hdat_bs2_elm), function(x) {
+#   hdat_bs2_elm[x, 1] + hdat_bs2_elm[x, 2] * extdum
+# }))
+# extconf <- apply(extpred, 2, function(x) c(mean(x), quantile(x, c(0.025, 0.975))))
+# durdum <- seq(-5, 5, 1)
+# durpred <- lapply(1:nrow(hdat_bs2_tlm), function(x) {
+#   hdat_bs2_tlm[x, 1] + hdat_bs2_tlm[x, 2] * durdum
+# })
+# 
+# plot(extmu[, 1], extmu[, 2], col = "grey", pch = 20)
+# lines(extdum, extconf[1, ])
+# lines(extdum, extconf[2, ])
+# 
+# 
+# datf[, plot(log10(act_ext), log10(eff_ext - act_ext))]
+# 
+# log10(0.1 - 0.05)
+# log10(0.1) - log10(0.05)
+# a <- seq(0.1, 10, 0.1)^10
+# mv <- runif(length(a), min = 1 - 0.4, max = 1 + 0.4)
+# b <- sapply(1:length(a), function(x) a[x] * mv[x])
+# plot(a, b)
+# cbind(log10(a), log10(b), log10(a) - log10(b), a, b, log10(a - b))
+# plot(log10(a), log10(b))
+# 
+# datf2 <- copy(datf[, .(act_ext, eff_ext)])
+# chkd <- datf2[order(act_ext), {
+#    ae = round(act_ext, 6)
+#    ee = round(eff_ext, 2) 
+#    aext = round(log10(act_ext), 2)
+#    eext = round(log10(eff_ext), 2)
+#    dif = round(ee - ae, 6)
+#    dlog = round(log10(eff_ext) - log10(act_ext), 2)
+#    logd = round(log10(eff_ext - act_ext), 2)
+#    list("ae" = ae, "ee" = ee, "aext" = aext, "eext" = eext, 
+#         "dif" = dif, "dlog" = dlog, "logd" = logd)
+# }]
+# chkd[1:100, ]
+# chkd[, plot(aext, logd)]
+# 
+# plot(durmu[, 1], durmu[, 2])
+# plot(extmu[, 1], extmu[, 2])
+
+```
+
+## Supplemental Analyses
 

### 2017-05-06 paper/SI.tex
Various updates towards revision, including new versions of figure 3

@@
 \vspace{-0.5cm} 
+   \Hitem \emph{\textbf{What is sampling interval, and how do we record it?}} Sampling interval is the time that elapsed between repeat observations of the same point in space or individual organism.
+In many cases, observations will only be made one time--list a value of 0 for these.
   \Hitem \emph{\textbf{What is sampling duration, and how do we record it?}} How long an individual observation of an individual point in space took to make.
@@
 For remote sensing observations you can assume one second (the observations are effectively instantaneous).
-   \Hitem \emph{\textbf{What is sampling interval, and how do we record it?}} Sampling interval is the time that elapsed between repeat observations of the same point in space or individual organism.
-In many cases, observations will only be made one time--list a value of 0 for these.
-  \Hitem \emph{\textbf{What is study duration, and how do we calculate it?}} Study duration is the integral of sampling duration, or the time spent making one observation of the phenomenon in question.
-To clarify, duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample, nor the integral of time spent in observing all spatial replicates.
+   \Hitem \emph{\textbf{What is the effective study duration, and how do estimate it?}} The effective study duration is the total period of time over which the phenomenon of interest was observed.
+More specifically, in the case of repeated observations, this is the total time that elapsed between when the first and last observations at a given point were made.
+For once-off (unreplicated) observations, this time is equivalent to the sample duration.
+However, there may be cases where once-off observations may have a longer effective duration then the sample duration.
+For example, consider a study that counts occurrences of pollinators over three years, using transects that are located in different locations within the broader study area during each year \cite{rollin_differences_2013-1}.
+The observations are therefore not strictly temporally replicated, but the authors control for year of collection in their subsequent analysis to avoid confounding effects.
+In this case, we can consider the effective duration to be three years, as the temporal information is encoded in the analysis.
+  \Hitem \emph{\textbf{What is actual study duration, and how do we calculate it?}} The actual study duration is the integral of sampling duration, or the time spent making one observation of the phenomenon in question.
+To clarify, duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample (the effective duration), nor the integral of time spent in observing all spatial replicates.
   \Hitem \emph{\textbf{Should study duration be the total time spent sampling all sites or the amount of time spent sampling per site (e.g.
@@
 In another example, if the reported plot size was 20 x 20 m, but the authors in fact only measured a random selection of, say, grass stems on which they counted aphids within those plots, then use an estimate of the area of the grass stem as the sampling resolution \cite{gagic_agricultural_2012}.
-  \Hitem \emph{\textbf{What is study extent, and how do I record it?}} Broadly defined, study extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares.
+  \Hitem \emph{\textbf{What is study extent, and how do I record it?}}  Broadly defined, actual study extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares.
+For studies in which the spatial replicates are not spatially contiguous (as with most field-based studies), this means plot resolution (see Q16) multiplied by number of plots.
+For spatially contiguous studies (e.g., those based on remote sensing imagery), it should be the total area covered by the imagery, i.e., pixel resolution multiplied by the number of pixels.
+However, only record the area analyzed by the authors.
+If they used a sub-sample of pixels, the actual extent is the number of those pixels multiplied by pixel resolution.
+
+Effective survey extent is defined as the total extent bounding all spatial replicates, divided by 10,000 to convert to hectares.
+For studies in which spatial replicates are not spatially contiguous, this means the area of the minimum convex polygon bounding all spatial replicates.
+For example, if the study is conducted in a national park, the effective survey extent would be the area of the national park; if the study is conducted in three national parks, the effective survey extent would be the sum of the areas of all national parks.
+To calculate the effective survey extent, use the area of the study area/region given in the paper; when the area is not given, but when the survey region is given by name (e.g., Joshua Tree National Park or United States), look up the area through an online search and convert to hectares.
+When the area is not named, but a map is given, use an appropriate digitizing platform with a suitable map-providing backend (e.g.
+Google Earth Pro, QGIS with OpenLayers plugin) to navigate to the region and delineate a minimum convex polygon surrounding the plots/transects/sampling units to calculate the area in hectares.
+When studies focus on portions of survey regions (e.g., mangroves in a coastal National Park), attempt to delineate the focal portions (the mangroves) and not the larger survey region (entire National Park) using Google Earth Pro.
+
+For spatially contiguous studies (e.g.
+those based on satellite imagery), the effective survey extent is the total area covered by the imagery (in such cases, effective survey extent equals actual survey extent).
+However, if spatially contiguous studies only use a sub-sample of pixels, the effective survey extent is the area of the minimum convex polygon bounding all pixels (calculated following the methods above).
+
+For studies that record individual, mobile organisms as the units of observation, use the minimum polygon surrounding the outermost observations of the complete space-time dataset (i.e.
+observations from all individuals and times) to define extent.
+
+%But I'm noticing now that this isn't updated to include how to handle when the study unit is animals.
+I guess we could just add a sentence saying that the effective extent is the surface area of an average-sized individual of the animal species, or something along those lines.
+Let me know what you think.
+
+%  Broadly defined, study extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares.
 For studies in which the spatial replicates are not spatially contiguous (as with most field-based studies), this means plot resolution (see Q16) multiplied by number of plots.

### 2017-05-06 paper/main.tex
Various updates towards revision, including new versions of figure 3

@@
 \usepackage{lineno}
+\usepackage{color}
 
@@
 To reveal these, we conducted a meta-analysis of recent (2004-2014) ecological research, in which we estimated the spatial resolution and extent, and temporal interval and duration, of reported ecological observations.
-Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
+\textcolor{blue}{Replace this text with more specific summary of observed scales:} \textcolor{red}{Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}}.
 Ecologists still make limited use of remote sensing (6\% of observations), which is necessary for observing larger-scale patterns and processes \cite{turner_remote_2003,pettorelli_satellite_2014}, but make slightly greater use of technologies that capture high frequency or temporally continuous point-based measurements (12\% of observations).
@@
 
-We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, sampling interval and total temporal duration.
-We calculated the actual, as opposed to representative (SI), scales for these four dimensions from 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 346 randomly selected articles (from 42,918 total).
-An additional 56 papers that were cited as the source of observations in the selected papers were also reviewed.
+We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, sampling interval and  temporal duration (Table 1).
+Here spatial extent refers to the area within the perimeter described by the outer-most sample plots, and duration refers to the total time elapsed between the collection of the first and last observations a time series.
+We also calculated two additional measures, actual extent and actual duration, which respetively
+
+We calculated the actual, as opposed to representative (SI), scales for these four dimensions from 375 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 134 paper subset of 346 randomly selected articles (from 42,918 total).
+An additional 59 papers that were cited as the source of observations in the selected papers were also reviewed.
+
 To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI).

### 2017-05-11 Rmd/calibration.Rmd
Revision through new figure 3, and updates to text. Culling out old code that does not appear or is used in analysis (e.g. normal bootstrapping code, polygon version of plots)

@@
 cal <- data.table(read_excel(fp(p_dat, "merged_calibrationffffff.xlsx")))
-names(cal)
+# names(cal)
 

### 2017-05-11 Rmd/space-time.Rmd
Revision through new figure 3, and updates to text. Culling out old code that does not appear or is used in analysis (e.g. normal bootstrapping code, polygon version of plots)

@@
 ## 1: 0.5769693 0.3821107 0.7196559 0.4049972     1.01134 1.253045 0.5284913
-p <- c(0.58, 0.75, 0.50, 1.01, 1.23, 0.59)  # from calibration.Rmd
+##    plot_res n_sites  act_ext  eff_ext t_btwn_samp  act_dur  eff_dur
+##       0.494   0.457    0.749    0.499       0.475    1.093    0.338
+##    plot_res n_sites  act_ext  eff_ext t_btwn_samp  act_dur  eff_dur
+##       0.577   0.382     0.72    0.405       1.045    1.261    0.643
+p <- c(0.58, 0.72, 0.41, 1.05, 1.26, 0.64)  # from calibration.Rmd
 
@@
 
-# hdat_bsdt <- lapply(1:length(vs), function(x) { # x <- 1
-#   mat <- do.call(rbind, lapply(1:length(hdat_bs), function(y) {
-#     hdat_bs[[y]][[x]]
-#   }))
-#   data.table(mat)
-# })
-
-# setnames(hdat_bsdt, names(hdat_bsdt), as.character(aaxis1$logres))
-
-# hist(hdat$res, breaks = aaxis1$logres)
-# hist(log10(datf$plot_res), breaks = resbrks)
-# hist(log10(datf$plot_res), breaks = resbrks)
-
-## Ordinary histograms
-mga <- -0.34
-sfigl <- -1.2
-cxa <- 0.8
-cxl <- 0.9
-reds <- brewer.pal(9, name = "Reds")
-blues <- brewer.pal(9, name = "Blues")
-# plot(1:2, pch = 20, col = blues[c(4, 6)])
-yl <- c(25, 40)
-# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
-pdf("paper/figures/hists.pdf", width = 5, height = 5)
-par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
-# resolution
-h <- hist(hdat$res, breaks = aaxis1$logres, plot = FALSE)
-h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Spatial resolution", 
-     ylim = c(0, yl[1]), main = "", col = reds[4], cex.lab = cxl)
-axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
-     tcl = -0.2, mgp = c(2, 0.25, mga))
-mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
-mtext("A", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
-
-# extent
-h <- hist(hdat$ext, breaks = aaxis2$logarea, plot = FALSE)
-# h <- hist(hdat$aext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
-h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Spatial extent", ylim = c(0, yl[1]), 
-     main = "", col = reds[6], cex.lab = cxl)
-axis(1, at = aaxis2$logarea,
-     labels = alab2,
-     # labels = c(expression(paste(NULL<="0.01m"^2)), expression(paste(0.1, m^2)),                 alab2[-1]), 
-     las = 2, cex.axis = cxa,
-     tcl = -0.2, mgp = c(2, 0.25, mga))
-mtext("B", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
-
-# interval
-tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
-h <- hist(hdat$int, breaks = taxis1$logdays, plot = FALSE)
-h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Sampling interval", ylim = c(0, yl[2]), 
-     main = "", col = blues[4], cex.lab = cxl)
-axis(1, at = tax, labels = tlab1[-10], las = 2, cex.axis = cxa, 
-     tcl = -0.2, mgp = c(2, 0.25, -0.34))
-mtext("C", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
-mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
-
-# duration
-# tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
-h <- hist(hdat$dur, breaks = taxis2$logdays, plot = FALSE)
-h$density <- h$counts / sum(h$counts) * 100
-plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
-     tcl = -0.2, cex.axis = cxa, xlab = "Temporal duration", ylim = c(0, yl[2]),      main = "", col = blues[6], cex.lab = cxl)
-axis(1, at = taxis2$logdays, labels = tlab2, las = 2, cex.axis = cxa, 
-     tcl = -0.2, mgp = c(2, 0.25, mga))
-mtext("D", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
-dev.off()
-
 ### Bootstrapped histograms (now main figure)
@@
               "tax2" = taxis2$logdays)
-# axlabs <- list(alab1, c(expression(paste(NULL<="0.01m"^2)),
-#                         expression(paste(0.1, m^2)), alab2[-c(1:2)]),
 axlabs <- list(alab1, alab2, tlab1[-10], tlab2)
@@
 
-# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
-pdf("paper/figures/hists_bs.pdf", width = 5, height = 5)
+# pdf("paper/figures/hists_bs.pdf", width = 5, height = 5)
+png("paper/figures/hists_bs.png", width = 5, height = 5, res = 600, 
+    units = "in")
 par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
@@
 Statistics associated with Figure 1: percentage of observations falling within different scale ranges, and associated with different observational methods
-```{r}
+```{r, eval = FALSE}
 # res
@@
      points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "red")]
-
-# points(resdat3$x[which(datf$st == 4)], resdat3$y[which(datf$st == 4)])
-# points(resdat3$x[which(datf$st == 1)], resdat3$y[which(datf$st == 1)])
 
@@
 rres3 <- (rres3 / cellStats(rres3, sum)) * 100
-# rng <- range(rres3[is.finite(rres3)])
-# bwidth <- (rng[2] - rng[1]) / 30
-# brks <- seq(rng[1], rng[2], bwidth)
-# brklabs <- seq(0, 0.12, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
-# brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
 
@@
 rext3 <- (rext3 / cellStats(rext3, sum)) * 100
-# rng <- range(rext3[is.finite(rext3)])
-# # rng[2] <- 0.1
-# bwidth <- (rng[2] - rng[1]) / 30
-# brks2 <- seq(rng[1], rng[2], bwidth)
-# brklabs2 <- brklabs #round(brks2[seq(1, length(brks2), 3)], 2)
 
@@
 ### Figure 3
-```{r}
-datf[, plot(log10(eff_ext), log10(act_ext))]
-datf[, plot(log10(act_dur), log10(eff_dur))]
-datf[log10(eff_dur) < log10(act_dur)]
-calr[log10(eff_dur) < log10(act_dur)]
-dat[DOI == "10.1007/s10021-013-9684-x"]
-cal[DOI == "10.1007/s00442-010-1713-2"]
-dat[DOI %in% datf[log10(eff_ext) < log10(act_ext), DOI]]
-dat[DOI %in% datf[log10(eff_ext) < log10(act_ext), DOI]]
-datf[log10(eff_ext) < log10(act_ext), ]
-datf[DOI == "10.1111/j.1466-8238.2011.00658.x", 
-     .(log10(round(eff_ext)), log10(round(act_ext)))]
-datf[DOI == "10.1111/j.1466-8238.2011.00658.x", 
-     .(round(eff_ext), round(act_ext))]
-
+```{r, eval=FALSE}
 # actual versus effective duration
@@
 
+# par(mfrow = c(1, 2))
+# datf[, {
+#   # sel = which()
+#   eext = log10(eff_ext)
+#   aext = log10(act_ext)
+#   edur = log10(eff_dur[t_btwn_samp != 365 * 10000])
+#   adur = log10(act_dur[t_btwn_samp != 365 * 10000])
+#   plot(aext, eext - aext)
+#   plot(adur, edur - adur)
+#   length(edur)
+# }]
+
+# Difference in effective versus actual extent/duration against actual ext/dur
+hdat_bs2 <- lapply(1:length(bootperturb), function(x) {  # x <- 1
+  print(x)
+  DT <- bootperturb[[x]]
+  edb <- DT[, lapply(.SD, log10), .SDcol = c("eff_ext", "act_ext")]
+  # edb <- edb[eff_ext >= act_ext, ]  # drop eff_dur < act
+  setnames(edb, c("eff_ext", "act_ext"), c("eext", "aext"))
+  edb[, dif := eext - aext]
+
+  tdb <- DT[t_btwn_samp != 365 * 10000, lapply(.SD, log10), 
+            .SDcol = c("eff_dur", "act_dur")]
+  # tdb <- tdb[eff_dur >= act_dur, ]  # drop eff_dur < act
+  setnames(tdb, c("eff_dur", "act_dur"), c("edur", "adur"))
+  # print(paste(edb[, .N], tdb[, .N]))
+  tdb[, dif := edur - adur]
+  return(list("edb" = edb, "tdb" = tdb, #"elm" = elm, "tlm" = tlm, 
+              "Ns" = c(edb[, .N], tdb[, .N])))
+})
+
+# check number of drops due to eff < act
+# hdat_bs2_ns <- do.call(rbind, lapply(hdat_bs2, function(x) x$Ns))
+# summary(bootperturb[[1]][, .N] - hdat_bs2_ns[, 1])
+# summary(bootperturb[[1]][t_btwn_samp != 365 * 10000, .N] - hdat_bs2_ns[, 2])
+
+# extract regression coefficients
+# hdat_bs2_elm <- do.call(rbind, lapply(hdat_bs2, function(x) coef(x$elm)[, 1]))
+# hdat_bs2_tlm <- do.call(rbind, lapply(hdat_bs2, function(x) coef(x$tlm)[, 1]))
+
+# calculate box statistics
+# Extent differences
+eint <- aaxis2$logarea
+edif_int <- rbindlist(lapply(hdat_bs2, function(y) {  # y <- hdat_bs2[[2]]
+  dd <- copy(y$edb[!is.na(aext)])
+  dd[aext < ex[1], aext := ex[1]]  # set to min
+  # dd[, list("int" = findInterval(aext, eint), "dif" = dif)]
+  dd[, int := findInterval(aext, eint, rightmost.closed = TRUE)]
+  # dd[, list("dif" = mean(dif, na.rm = TRUE), "N" = .N), by = int][order(int)]
+}))
+ints <- sort(edif_int[, unique(int)])
+ebox <- lapply(ints, function(x) edif_int[int == x, lmisc::box_stats(dif)])
+pctobs <- sapply(ints, function(x) edif_int[int == x, .N])
+ewgt <- pctobs / sum(pctobs)
+# ealphs <- (ewgt - min(ewgt)) / diff(range(ewgt))
+ewgtint <- c(0, 1, 5, 10, 15, 20, 25) / 100  # weight intervals
+ewgttxt <- c("  0", "  1", "  5", seq(10, 25, 5))
+ecols <- findInterval(ewgt, ewgtint)  # corresponding color index
+
+# temporal differences
+dint <- taxis2$logdays[-5]
+tx <- range(dint)
+ddif_int <- rbindlist(lapply(hdat_bs2, function(y) {  # y <- hdat_bs2[[2]]
+  dd <- copy(y$tdb[!is.na(adur)])
+  dd[adur < tx[1], adur := tx[1]]  # set to min
+  dd[adur > tx[2], adur := tx[2]]  # set to max
+  dd[, int := findInterval(adur, dint, rightmost.closed = TRUE)]  # include high
+}))
+dints <- sort(ddif_int[, unique(int)])
+tbox <- lapply(dints, function(x) ddif_int[int == x, lmisc::box_stats(dif)])
+pctobs <- sapply(dints, function(x) ddif_int[int == x, .N])
+dwgt <- pctobs / sum(pctobs)
+# dalphs <- (dwgt - min(dwgt)) / diff(range(dwgt))
+dwgtint <- c(0, 1, 5, 10, 15, 20, 25, 30) / 100  # weight intervals
+dwgttxt <- c("  0", "  1", "  5", seq(10, 30, 5))
+dcols <- findInterval(dwgt, dwgtint)
+
+# y grid line function
+ylines <- function(yr, by, lty = 1, col = "grey") {
+  abline(h = seq.int(from = yr[1], to = yr[2], by = 2), lty = lty, col = col)
+}
+
+# pdf("paper/figures/act_v_eff_diff.pdf", width = 7, height = 4)
+png("paper/figures/act_v_eff_diff.png", width = 7, height = 4, res = 600, 
+    units = "in")
+par(mfrow = c(1, 2), mar = c(4, 0, 1, 0.75), oma = c(0, 3, 0, 0))
+bx <- rowMeans(cbind(eint[-length(eint)], eint[-1]))
+# cols <- sapply(ealphs, function(x) rgb(0, 0.1, 1, alpha = x))
+cols <- brewer.pal(length(ewgtint) - 1, "Blues")
+plot(ex, c(0, 14), pch = "", xlab = "", ylab = "", xaxt = "n", yaxt = "n",
+     mgp = c(3, 0.5, 0), tcl = -0.2, xaxs = "i")
+polygon(c(ex, rev(ex), ex[1]), c(-1, -1, 15, 15, -1), col = "grey90")
+ylines(yr = c(0, 14), by = 2, col = "grey95")
+for(i in 1:length(ebox)) {
+  boxplot_v(bx[i], y = ebox[[i]], n = 1.75, whiskhgt = 1.5, 
+            bfill = cols[ecols][i])
+} 
+axis(1, at = eint, labels = alab2, las = 2, tcl = -0.2, cex.axis = cxa, 
+     mgp = c(3, 0.5, 0))
+axis(2, at = seq(0, 14, 2), las = 2, tcl = -0.2, cex.axis = cxa, 
+     mgp = c(3, 0.5, 0))
+mtext("Actual Extent", side = 1, line = 3, cex = 0.9)
+mtext("Difference Between", side = 2, line = 2.1, cex = 0.9)
+mtext("Effective and Actual (Decades)", side = 2, line = 1.4, cex = 0.9)
+lrng <- c(mean(ex), max(ex) - 0.05 * max(ex))
+xs <- seq(lrng[1], lrng[2], by = (diff(lrng)) / (length(cols)))
+rect_shade(xs, y = c(11.5, 12.5), fillcol = cols, linecol = "grey30")
+text(xs, rep(10.5, length(xs)), ewgttxt, srt = 90, col = "grey30")
+text(mean(xs), 13, "% of observations", col = "grey30")
+mtext("A", side = 3, line = -1, cex = 1, adj = 0.05)
+
+# duration
+bx <- rowMeans(cbind(dint[-length(dint)], dint[-1]))
+bd <- abs(dint[-length(dint)] - dint[-1])
+# cols <- sapply(dalphs, function(x) rgb(1, 0, 0.1, alpha = x))
+cols <- brewer.pal(length(dwgtint) - 1, "Reds")
+plot(tx, c(0, 14), pch = "", xlab = "", ylab = "", xaxt = "n", xaxs = "i",
+     mgp = c(3, 0.5, 0), tcl = -0.2, yaxt = "n")
+polygon(c(tx, rev(tx), tx[1]), c(-1, -1, 15, 15, -1), col = "grey90")
+ylines(yr = c(0, 14), by = 2, col = "grey95")
+for(i in 1:length(tbox)) {
+  boxplot_v(bx[i], y = tbox[[i]], n = 0.9, inhgt = bd[i] * 10, whiskhgt = 0.5, 
+            bfill = cols[dcols][i])
+} 
+axis(1, at = dint, labels = tlab2[-c(5)], las = 2, tcl = -0.2, cex.axis = cxa, 
+     mgp = c(3, 0.5, 0))
+axis(2, at = seq(0, 14, 2), labels = rep("", 8), las = 2, tcl = -0.2, 
+     cex.axis = cxa, mgp = c(3, 0.5, 0))
+mtext("Actual Duration", side = 1, line = 3, cex = 0.9)
+lrng <- c(mean(tx), max(tx) - 0.05 * max(tx))
+xs <- seq(lrng[1], lrng[2], by = (diff(lrng)) / (length(cols)))
+rect_shade(xs, y = c(11.5, 12.5), fillcol = cols, linecol = "grey30")
+text(xs, rep(10.5, length(xs)), dwgttxt, srt = 90, col = "grey30")
+text(mean(xs), 13, "% of observations", col = "grey30")
+mtext("B", side = 3, line = -1, cex = 1, adj = 0.05)
+dev.off()
+
+# Kernel density version
 pdf("paper/figures/act_v_eff.pdf", width = 7, height = 3)
@@
 
-# par(mfrow = c(1, 2))
-# datf[, {
-#   # sel = which()
-#   eext = log10(eff_ext)
-#   aext = log10(act_ext)
-#   edur = log10(eff_dur[t_btwn_samp != 365 * 10000])
-#   adur = log10(act_dur[t_btwn_samp != 365 * 10000])
-#   plot(aext, eext - aext)
-#   plot(adur, edur - adur)
-#   length(edur)
-# }]
-
-# Bootstrap version
-vs1 <- c("eff_ext", "act_ext", "eff_dur", "act_dur")
-vs2 <- c("eext", "aext", "edur", "adur")
-hdat_bs2 <- lapply(1:length(bootperturb), function(x) {  # x <- 1
-  # print(x)
-  DT <- bootperturb[[x]]
-  edb <- DT[, lapply(.SD, log10), .SDcol = c("eff_ext", "act_ext")]
-  # edb <- edb[eff_ext >= act_ext, ]  # drop eff_dur < act
-  setnames(edb, c("eff_ext", "act_ext"), c("eext", "aext"))
-  edb[, dif := eext - aext]
-  # elm <- edb[, lm(dif ~ aext)]
-  # preds <- predict(elm, newdata = data.frame(aext = extdum))
-  # plot(extdum, preds)
-
-  tdb <- DT[t_btwn_samp != 365 * 10000, lapply(.SD, log10), 
-            .SDcol = c("eff_dur", "act_dur")]
-  # tdb <- tdb[eff_dur >= act_dur, ]  # drop eff_dur < act
-  setnames(tdb, c("eff_dur", "act_dur"), c("edur", "adur"))
-  print(paste(edb[, .N], tdb[, .N]))
-  tdb[, dif := edur - adur]
-  # tlm <- tdb[, summary(lm(dif ~ adur))]
-  return(list("edb" = edb, "tdb" = tdb, #"elm" = elm, "tlm" = tlm, 
-              "Ns" = c(edb[, .N], tdb[, .N])))
-})
-
-# check number of drops due to eff < act
-# hdat_bs2_ns <- do.call(rbind, lapply(hdat_bs2, function(x) x$Ns))
-# summary(bootperturb[[1]][, .N] - hdat_bs2_ns[, 1])
-# summary(bootperturb[[1]][t_btwn_samp != 365 * 10000, .N] - hdat_bs2_ns[, 2])
-
-# extract regression coefficients
-# hdat_bs2_elm <- do.call(rbind, lapply(hdat_bs2, function(x) coef(x$elm)[, 1]))
-# hdat_bs2_tlm <- do.call(rbind, lapply(hdat_bs2, function(x) coef(x$tlm)[, 1]))
-
-# calculate box statistics
-# Extent differences
-eint <- aaxis2$logarea
-edif_int <- rbindlist(lapply(hdat_bs2, function(y) {  # y <- hdat_bs2[[2]]
-  dd <- copy(y$edb[!is.na(aext)])
-  dd[aext < ex[1], aext := ex[1]]  # set to min
-  # dd[, list("int" = findInterval(aext, eint), "dif" = dif)]
-  dd[, int := findInterval(aext, eint)]
-  # dd[, list("dif" = mean(dif, na.rm = TRUE), "N" = .N), by = int][order(int)]
-}))
-
-ints <- sort(edif_int[, unique(int)])
-ebox <- lapply(ints, function(x) edif_int[int == x, lmisc::box_stats(dif)])
-pctobs <- sapply(ints, function(x) edif_int[int == x, .N])
-ewgt <- pctobs / sum(pctobs)
-ewgt * 100
-ealphs <- (ewgt - min(ewgt)) / diff(range(ewgt))
-
-# temporal differences
-dint <- taxis2$logdays[-5]
-tx <- range(dint)
-ddif_int <- rbindlist(lapply(hdat_bs2, function(y) {  # y <- hdat_bs2[[2]]
-  dd <- copy(y$tdb[!is.na(adur)])
-  dd[adur < tx[1], adur := tx[1]]  # set to min
-  dd[adur > tx[2], adur := tx[2]]  # set to max
-  dd[, int := findInterval(adur, dint)]
-}))
-
-dints <- sort(ddif_int[, unique(int)])
-tbox <- lapply(dints, function(x) ddif_int[int == x, lmisc::box_stats(dif)])
-pctobs <- sapply(dints, function(x) ddif_int[int == x, .N])
-dwgt <- pctobs / sum(pctobs)
-dalphs <- (dwgt - min(dwgt)) / diff(range(dwgt))
-
-pdf("paper/figures/act_v_eff_diff.pdf", width = 7, height = 4)
-par(mfrow = c(1, 2), mar = c(4, 0, 1, 0.75), oma = c(0, 3, 0, 0))
-bx <- rowMeans(cbind(eint[-length(eint)], eint[-1]))
-cols <- sapply(ealphs, function(x) rgb(0, 0.1, 1, alpha = x))
-# plot(1:length(cols), col = lcols, pch = 20)
-plot(ex, c(0, 14), pch = "", xlab = "", ylab = "", xaxt = "n", yaxt = "n",
-     mgp = c(3, 0.5, 0), tcl = -0.2, xaxs = "i")
-# grid(lty = 1, col="gray")
-# polygon(c(ex, rev(ex), ex[1]), c(-1, -1, 15, 15, -1), col = "grey90")
-for(i in 1:length(ebox)) {
-  boxplot_v(bx[i], y = ebox[[i]], n = 1.75, whiskhgt = 1.5, bfill = cols[i])
-} 
-axis(1, at = eint, labels = alab2, las = 2, tcl = -0.2, cex.axis = cxa, 
-     mgp = c(3, 0.5, 0))
-axis(2, at = seq(0, 14, 2), las = 2, tcl = -0.2, cex.axis = cxa, 
-     mgp = c(3, 0.5, 0))
-mtext("Actual Extent", side = 1, line = 3)
-mtext("Difference Magnitude:", side = 2, line = 2)
-mtext("Effective minus Actual", side = 2, line = 1)
-# xs <- seq(2, 10, by = (10 - 2) / (length(ealphs)))
-# ealphss <- sort(ealphs, decreasing = TRUE)
-# lcols <- sapply(ealphss, function(x) rgb(0, 0.1, 1, alpha = x))
-# rect_shade(xs, y = c(13, 14), fillcol = lcols, linecol = "grey")
-# text(xs, rep(13, length(xs)), sort(round(ewgt * 100), decreasing = TRUE),
-#      srt = 90)
-
-bx <- rowMeans(cbind(dint[-length(dint)], dint[-1]))
-bd <- abs(dint[-length(dint)] - dint[-1])
-cols <- sapply(dalphs, function(x) rgb(1, 0, 0.1, alpha = x))
-# plot(1:length(cols), col = cols, pch = 20)
-plot(tx, c(0, 14), pch = "", xlab = "", ylab = "", xaxt = "n", 
-     mgp = c(3, 0.5, 0), tcl = -0.2, yaxt = "n")
-for(i in 1:length(tbox)) {
-  boxplot_v(bx[i], y = tbox[[i]], n = 0.9, inhgt = bd[i] * 10, whiskhgt = 0.5, 
-            bfill = cols[i])
-} 
-axis(1, at = dint, labels = tlab2[-c(5)], las = 2, tcl = -0.2, cex.axis = cxa, 
-     mgp = c(3, 0.5, 0))
-axis(2, at = seq(0, 14, 2), labels = rep("", 8), las = 2, tcl = -0.2, 
-     cex.axis = cxa, mgp = c(3, 0.5, 0))
-mtext("Actual Duration", side = 1, line = 3)
-dev.off()
-
-# # Calculate predictions from ensemble of regression coefficients
-# extdum <- seq(-10, 10, 1)
-# extpred <- do.call(rbind, lapply(1:nrow(hdat_bs2_elm), function(x) {
-#   hdat_bs2_elm[x, 1] + hdat_bs2_elm[x, 2] * extdum
-# }))
-# extconf <- apply(extpred, 2, function(x) c(mean(x), quantile(x, c(0.025, 0.975))))
-# durdum <- seq(-5, 5, 1)
-# durpred <- lapply(1:nrow(hdat_bs2_tlm), function(x) {
-#   hdat_bs2_tlm[x, 1] + hdat_bs2_tlm[x, 2] * durdum
-# })
-# 
-# plot(extmu[, 1], extmu[, 2], col = "grey", pch = 20)
-# lines(extdum, extconf[1, ])
-# lines(extdum, extconf[2, ])
-# 
-# 
-# datf[, plot(log10(act_ext), log10(eff_ext - act_ext))]
-# 
-# log10(0.1 - 0.05)
-# log10(0.1) - log10(0.05)
-# a <- seq(0.1, 10, 0.1)^10
-# mv <- runif(length(a), min = 1 - 0.4, max = 1 + 0.4)
-# b <- sapply(1:length(a), function(x) a[x] * mv[x])
-# plot(a, b)
-# cbind(log10(a), log10(b), log10(a) - log10(b), a, b, log10(a - b))
-# plot(log10(a), log10(b))
-# 
-# datf2 <- copy(datf[, .(act_ext, eff_ext)])
-# chkd <- datf2[order(act_ext), {
-#    ae = round(act_ext, 6)
-#    ee = round(eff_ext, 2) 
-#    aext = round(log10(act_ext), 2)
-#    eext = round(log10(eff_ext), 2)
-#    dif = round(ee - ae, 6)
-#    dlog = round(log10(eff_ext) - log10(act_ext), 2)
-#    logd = round(log10(eff_ext - act_ext), 2)
-#    list("ae" = ae, "ee" = ee, "aext" = aext, "eext" = eext, 
-#         "dif" = dif, "dlog" = dlog, "logd" = logd)
-# }]
-# chkd[1:100, ]
-# chkd[, plot(aext, logd)]
-# 
-# plot(durmu[, 1], durmu[, 2])
-# plot(extmu[, 1], extmu[, 2])
-
 ```
@@
 ```{r}
-svars <- c("plot_res", "n_sites", "act_ext", "samp_dur",
-           "t_btwn_samp", "act_dur")
-
+svars <- c("plot_res", "n_sites", "act_ext", "eff_ext", "samp_dur",
+           "t_btwn_samp", "act_dur", "eff_dur")
 sens <- gsub("\\\\", ";", dat$sensitivity)
@@
 
-sens2 <- gsub("plo_", "plot_", sens)  # fix this one
+sens2 <- gsub("plo_|plot-", "plot_", sens)  # fix this one
 sens2 <- gsub("[[:blank:]]", "_", sens2)  # replace spaces with _
-
-corrtab <- cbind(c("^res", "plo_res", "sample_area", "samples_area", 
-                   "sampled area",
+sens2 <- gsub("-", "_", sens2)  # replace - with _
+
+corrtab <- cbind(c("^res", "plo_res", "plot_resolution", "plot_size", 
+                   "sample_area", "samples_area", "sampled area", "study_area",
+                   "sampled_area",
                    "t_btwn_samples", "t_btwn_sample", "t_btwn_samp_samp", 
-                   "t_twn_samp", "t_btw_samp", "samp_duratiion", 
-                   "sampling_duration", "sample_duration", 
-                   "plot_resolution", "number_of_sites", 
-                   "time_between_sampling", "plot_size", 
-                   "time_between_samples", "samp_study", ":", " ", ";_"), 
-                 c(svars[c(1, 1, 3, 3, 3, 5, 5, 5, 5, 5, 4, 4, 4, 1, 2,
-                           5, 1, 5)], "samp;study", ";", ";", ";"))
-
+                   "t_twn_samp", "t_btw_samp", "time_between_sampling", 
+                   "time_between_samples", 
+                   "samp_duratiion", "sampling_duration", "sample_duration", 
+                   "number_of_sites", 
+                   "study_span", "sutdy_span", "study_dur", "study_soan",
+                   "samp_study", ":", " ", ";_", 
+                   "duration", "dure", 
+                   "extent"), 
+                 c(rep("plot_res", 4), rep("act_ext", 5), rep("t_btwn_samp", 7),                    rep("samp_dur", 3), "n_sites", rep("eff_dur", 4), 
+                   "samp;study", ";", ";", ";", 
+                   "dur", "dur", "ext"))
+
+# replace
 for(i in 1:nrow(corrtab)) {
@@
 # unique(unlist(sensl))
-# unique(unlist(sensl2))
+unique(unlist(sensl2))
 
@@
 }))
-colnames(senst) <- c("res", "n", "sarea", "sampd", "tbtwn", "studyd")
-
-# add column for uncertainty about n repeats
-# senst$nrep <- rep(0, nrow(senst))
-
-# i <- 300:305
-# sensl2[i]
-# senst[i, ]
+colnames(senst) <- c("res", "n", "aext", "eext", "sampd", "tbtwn", "actdur", 
+                     "effdur")
 
 # apply fixes (mostly to account for observer omissions)
-
 # 1. if samp_dur OR t_btwn_samp is uncertain, act_dur must also be. 
-id <- which((senst$sampd == 1 | senst$tbtwn == 1) & senst$studyd == 0)
-senst[id, "studyd"] <- 1
+id <- which((senst$sampd == 1 | senst$tbtwn == 1) & senst$actdur == 0)
+senst[id, "actdur"] <- 1
 
@@
 # that observers didn't note this is sensitive. 
-id <- which(senst$tbtwn == 0 & senst$sampd == 0 & senst$studyd == 1)
+id <- which(senst$tbtwn == 0 & senst$sampd == 0 & senst$actdur == 1)
 senst[id, c("sampd", "tbtwn")] <- 1
@@
 # 3. if plot_res or n_sites is uncertain, then act_ext must be
-id <- which((senst$res == 1 | senst$n == 1) & senst$sarea == 0)
-senst[id, "sarea"] <- 1
+id <- which((senst$res == 1 | senst$n == 1) & senst$aext == 0)
+senst[id, "aext"] <- 1
 
@@
 ```{r,eval = FALSE}
+# redo range variables
+rx <- range(resdat3$x)
+rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
+rx[1] <- -5  # set min interval to less than second
+rx[2] <- 5.57  # set max interval to millenium
+ex <- range(taxis2$logdays)
+ex <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
+ey <- range(aaxis2$logarea)
 
@@
 
+# plot parameters
 cxa <- 1.25
+cuts <- 40
+bump <- 3
+cols <- inferno(cuts + bump)[-c(2:(bump + 1))]
 
@@
 par(mfrow = c(3, 2), oma = c(6, 3, 0, 0), mar = c(1, 5, 3, 4))
-lapply(1:3, function(i) {  # i <- 3
-  x <- kernsl[[i]]
-  # x <- kernsl[[1]]
-  image(x$res, col = rev(terrain.colors(80)), axes = FALSE, xlab = "", 
-        ylab = "")#, ylim = range(aaxis1$logarea))
+for(i in 1:3) {  # i <- 3
+  x <- kernsl[[i]] # x <- kernsl[[1]]
+  image(x$res, col = cols, axes = FALSE, xlab = "", ylab = "")
   if(i == 3) {
@@
   plot(x$res, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, #breaks = brks, 
-     col = rev(terrain.colors(80)))
-
-  image(x$ext, col = rev(terrain.colors(80)), axes = FALSE, xlab = "", 
-        ylab = "")
+     legend.shrink = 0.9, col = cols)#breaks = brks,
+
+  image(x$ext, col = cols, axes = FALSE, xlab = "", ylab = "")
   if(i == 3) {
@@
   plot(x$ext, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-       legend.shrink = 0.9, #breaks = brks2, 
-       col = rev(terrain.colors(80)))
-})
-par(xpd = NA)
-xx <- grconvertX(0.5, from = "ndc", to = "inches")
-ii <- c(0.4, 0.6, 0.8, 1)
-yy <- c(0.95, 0.75, 0.5, 0.25)
-# yy <- c(-1, -5, -10, -20)
-for(i in 1:length(ii)) {
-  y <- grconvertY(yy[i], from = "ndc", to = "inches")
+       legend.shrink = 0.9, col = cols)#breaks = brks2, 
+}
+# # par(xpd = NA)
+# for(j in c("in", "dev", "ndc", "nfc", "npc", "nic")) {
+#   xx <- grconvertX(0.5, from = "user", to = j)
+#   ii <- c(0.4, 0.6, 0.8, 1)
+#   yy <- c(0.95, 0.75, 0.5, 0.25)
+#   # yy <- c(-1, -5, -10, -20)
+#   for(i in 1:length(ii)) {
+#     y <- grconvertY(yy[i], from = "user", to = j)
 #     text(x = xx, y = y, labels = paste("kernel =", ii[i]))
-  # par(xpd = NA)
-  # mtext(text = paste("kernel =", ii[i]), side = 3, line = yy[i])
-}
+#     # par(xpd = NA)
+#     # mtext(text = paste("kernel =", ii[i]), side = 3, line = yy[i])
+#   }
+# }
 dev.off()
@@
 <a href="#top">Back to top</a>
-
-### Polygon version of plots
-
-```{r, eval = FALSE, echo = FALSE}
-stype <- data.table(datf[, study_type])
-stype[like(V1, "remote|other"), newnm := "remote"]
-stype[like(V1, "field|paleo"), newnm := "field"]
-stype[like(V1, "automated"), newnm := "automated"]
-
-tbtwn <- datf$t_btwn_samp
-# which(tbtwn < (0.1 / (60 * 60 * 24)) & tbtwn > 0)
-tbtwn <- ifelse(tbtwn == 0, 0.1 / (60 * 60 * 24), tbtwn)
-tempdat2 <- data.table(x = log10(tbtwn), y = log10(datf$act_dur))
-tempdat2[x == min(x), x := -5.5]
-tempdat2[x > -5.5 & x < tx[1], x := tx[1]]
-# tempdat2[y == min(y)]
-tempdat2[y > ty[2], y := ty[2]]
-tx2 <- round(range(tempdat2$x), 1)
-ty2 <- round(range(tempdat2$y), 1)
-
-# tempdat2 <- data.table(x = log10(to_infin_byond(datf$t_btwn_samp)), 
-#                       y = log10(datf$act_dur))
-spatdat2 <- data.table(x = log10(datf$plot_res / 10000), # convert to ha
-                       y = log10(datf$act_ext))
-sx2 <- round(range(spatdat2$x), 1)
-sx2[2] <- 4.8
-sy2 <- round(range(spatdat2$y), 1)
-
-# x1 = interval, x2 = study duration
-# y1 = plot res, y2 = extent
-
-pgons <- lapply(1:nrow(tempdat2), function(i) {  # i <- 6
-  print(i)
-  x <- c(rep(tempdat2[i, x], 2), rep(tempdat2[i, y], 2), tempdat2[i, x])
-  y <- c(spatdat2[i, x], rep(spatdat2[i, y], 2), rep(spatdat2[i, x], 2))
-  Polygons(list(Polygon(cbind(x, y))), ID = as.character(i))
-})
-spgons <- SpatialPolygons(pgons)
-spgons$type <- stype$newnm
-
-xl <- range(c(tx2, ty2))
-yl <- range(c(sx2, sy2))
-
-# not compelling, polygonized version
-apha <- 0.05
-par(mar = rep(2, 4))
-plot(spgons, lty = 0, axes = TRUE, xlim = xl, ylim = yl)
-plot(spgons[spgons$type == "field", ], add = TRUE, 
-     col = rgb(1, 0, 0, alpha = apha), border = "transparent")
-plot(spgons[spgons$type == "remote", ], add = TRUE, 
-     col = rgb(0, 1, 0, alpha = apha), border = "transparent")
-plot(spgons[spgons$type == "automated", ], add = TRUE, 
-     col = rgb(0, 0, 1, alpha = apha), border = "transparent")
-
-# rasterized version less compelling
-r <- raster(extent(c(range(c(tx2, ty2)), range(sx2, sy2))))
-res(r) <- 0.1
-r[] <- 1
-plot(r)
-rpts <- rasterToPoints(r, spatial = TRUE)
-spct <- sapply(over(rpts, geometry(spgons), returnList = TRUE), length)
-rpts$n <- spct
-rptsgrd <- SpatialPixelsDataFrame(rpts, rpts@data)
-rptsr <- raster(rptsgrd, layer = "n")
-plot(rptsr)
-
-```

### 2017-05-11 paper/main.tex
Revision through new figure 3, and updates to text. Culling out old code that does not appear or is used in analysis (e.g. normal bootstrapping code, polygon version of plots)

--- 2016-08-25 paper/main.tex
@@
 \author
-{Lyndon Estes$^{\ast1, 2}$, Labeeb Ahmed$^{3}$, Kelly Caylor$^{2}$, Jason Chang$^{3}$, \\
-Jonathan J. Choi$^{4}$, Erle Ellis$^{3}$, Paul R. Elsen$^{1}$, and Tim Treuer$^{4}$ \\
+{Lyndon Estes$^{\ast1, 2, 3}$, Paul R. Elsen$^{1, 4}$, Tim Treuer$^{5}$, Labeeb Ahmed$^{6}$, \\
+Kelly Caylor$^{1, 7}$, Jason Chang$^{6}$, Jonathan J. Choi$^{5}$, and Erle Ellis$^{6}$ \\
 \\
@@
 \normalsize{$^{2}$Civil and Environmental Engineering, Princeton University, Princeton, NJ 08544, USA}\\
-\normalsize{$^{3}$Geography and Environmental Systems, University of Maryland Baltimore County,}\\
+\normalsize{$^{3}$Graduate School of Geography, Clark University, Worcester MA 01610, USA}\\
+\normalsize{$^{4}$Department of Environmental Science, Policy, and Management,}\\
+\normalsize{University of California Berkeley, Berkeley, CA 94720, USA}\\
+\normalsize{$^{5}$Ecology and Evolutionary Biology, Princeton University, Princeton, NJ 08544, USA}\\
+\normalsize{$^{6}$Geography and Environmental Systems, University of Maryland Baltimore County,}\\ 
 \normalsize{Baltimore, MD 21250, USA}\\
-\normalsize{$^{4}$Ecology and Evolutionary Biology, Princeton University, Princeton, NJ 08544, USA}\\
+\normalsize{$^{7}$Earth Research Institute, University of California Santa Barbara, Santa Barbara, CA 93106, USA}\\
 \\
-\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  lestes@princeton.edu.}
-}
+\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  LEstes@clarku.edu.}}
 % Include the date command, but leave its argument blank.
@@
 %\begin{sciabstract}
-\textbf{Observational scale shapes our understanding of ecosystems \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
-Ecological patterns emerge at temporal and spatial scales that may be coarser or finer than the processes that shape them, thus investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992}.
-Given the centrality of scale to ecology, and the growing ability to observe ecological phenomena beyond a field perspective, it is important to understand the scales at which modern ecological research is conducted.
-To reveal these, we conducted a meta-analysis of recent (2004-2014) ecological research, in which we estimated the spatial resolution and extent, and temporal interval and duration, of reported ecological observations.
-Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
-Ecologists still make limited use of remote sensing (6\% of observations), which is necessary for observing larger-scale patterns and processes \cite{turner_remote_2003,pettorelli_satellite_2014}, but make slightly greater use of technologies that capture high frequency or temporally continuous point-based measurements (12\% of observations).
+\section*{Abstract}
+%Ecologists have long known that the observational scales play a critical role in shaping our understanding of ecosystems, and that observations spanning multiple scales of space and time are needed to effectively characterize ecological phenomena.
+\textbf{In order to properly understand ecological phenomena, it is necessary to quantify their behavior over the range of spatial and temporal scales at which they manifest.
+Ecology has been concerned with this need since the early 1990s, and the ability to collect multi-scaled ecological observations has grown rapidly since then.
+Characterizing the spatial and temporal domains of modern ecological observations can therefore provide important insight into the field's progress in understanding towards a more comprehensive understanding of ecosystem behaviour.
+To characterize these domains, we conducted a meta-analysis of recent (2004-2014) ecological studies, in which we quantified four primary dimensions of their reported observations: plot resolution, sampling interval, effective duration (time between start and end of temporal replicates), and effective extent (area enclosed by spatial replicates).
+We also estimated the \emph{actual} extent and duration, which respectively represent the summed area and time covered by spatial and temporal replicates.
+\textcolor{blue}{Replace this text with more specific summary of observed scales:} \textcolor{red}{Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}}}.
+% Ecologists still make limited use of remote sensing (6\% of observations), which is necessary for observing larger-scale patterns and processes \cite{turner_remote_2003,pettorelli_satellite_2014}, but make slightly greater use of technologies that capture high frequency or temporally continuous point-based measurements (12\% of observations).
 Resolution, extent, and duration were not clearly reported in most reviewed studies.
@@
 \linenumbers
+\vspace{10pt}
 The scales at which ecosystems are observed plays a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
 Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is the \emph{sine qua non} for understanding ecological phenomena \cite{levin_problem_1992}.
-Awareness of the importance of scale has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad range of scales \cite{schneider_rise_2001,chave_problem_2013}.
+This awareness has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad range of scales \cite{schneider_rise_2001, chave_problem_2013}.
 
-Given the centrality of scale to understanding ecology, and the growing ability to study ecological phenomena across a broad range of spatial and temporal scales, it is important to assess the scales at which ecological research is presently being conducted.
+Given this awareness of the centrality of scale to understanding ecology, and the growing ability to study ecological phenomena across a broad range of spatial and temporal scales, it is important to assess the scales at which current ecological research is conducted.
 To gain insight into this question, we quantified the spatial and temporal domains of empirical observations that were reported in a representative sample of studies published between 2004-2014 in the top 30 ecological journals (by impact factor).
@@
 
-We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, sampling interval and total temporal duration.
-We calculated the actual, as opposed to representative (SI), scales for these four dimensions from 371 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 140 paper subset of 346 randomly selected articles (from 42,918 total).
-An additional 56 papers that were cited as the source of observations in the selected papers were also reviewed.
+We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, interval and duration (Table 1).
+Here resolution is the area of an individual spatial replicate, or the two-dimensional space in which all measurable features of a natural object(s) were recorded (as opposed to sub-sampled), while extent is the area enclosed by the outer-most spatial replicates, or, if the system or habitat being sampled was distinct from its surrounding matrix (e.g.
+forest patches in grassland habitats), the summed area of sampled patches (see SI for full definition).
+Interval refers to the average time elapsed between individual temporal replicates, and duration is the time elapsed between the first and last temporal replicates, or, in the case of temporally unreplicated observations, the estimated time spent collecting the observation (SI).
+We also calculated two additional metrics, the integrated area of spatial replicates (i.e.
+resolution multiplied by number of replicates) and the summed observational time of all temporal replicates.
+We estimated these additional dimensions to evaluate the degree to which the actual scales of ecological observations differ from those they ostensibly represent, and therefore refer to them as the \emph{actual} extent and \emph{actual} duration.
+
+\begin{table}[htp]
+\caption{The dimensions of ecological observations estimated in this meta-analysis.}
+\begin{center}
+\begin{tabular}{ll}
+\hline
+\textbf{Dimension} & \textbf{Description} \\
+\hline
+Resolution & Area (m$^2$) of an individual spatial replicate (plot) \\
+Extent & Area (ha) encompassed by all spatial replicates \\
+Actual extent & Summed area (ha) of all spatial replicates\\
+Interval & Time elapsed (days) between successive temporal replicates \\
+Duration & Time elapsed (days) between first and last temporal replicates\\
+Actual duration & Summed observational time (days) of all temporal replicates\\
+\hline
+\end{tabular}
+\end{center}
+\label{default}
+\end{table}%
+
+We calculated these dimensions from 379 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 134 paper subset of 348 randomly selected articles (from 42,918 total).
+An additional 62 papers that were cited as the source of observations in the selected papers were also reviewed.
+
 To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI).
@@
 
-In terms of spatial resolution (the two-dimensional space in which all measurable features of a natural object(s) were recorded, as opposed to sub-sampled), the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
+In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
 1A).

### 2017-05-11 Rmd/space-time.Rmd
Updated previous commit to actually finalize getting rid of all deprecated (tried but used) analyses

@@
 <a href="#top">Back to top</a>
-
-## Tried and deprecated
-
-Code not shown in HTML
-
-### Scale diversity within study
-```{r, eval=FALSE, echo=FALSE}
-# plot resolution
-doin <- datf[, .N, by = DOI]  # n obs/DOI
-resuni <- datf[, length(unique(plot_res)), by = DOI]  # unique resolutions
-resrng <- datf[, diff(range(plot_res)), by = DOI]
-rescv <- datf[DOI %in% resuni[V1 > 1, DOI], cv(plot_res), by = DOI]
-ressd <- datf[DOI %in% resuni[V1 > 1, DOI], log10(sd(plot_res)), by = DOI]
-resmu <- datf[DOI %in% resuni[V1 > 1, DOI], log10(mean(plot_res)), by = DOI]
-# plot(resuni$V1, resrng$V1)
-# plot(resmu$V1, ressd$V1)
-# plot(resmu$V1, rescv$V1)
-
-# sampling interval
-intuni <- datf[, length(unique(t_btwn_samp)), by = DOI]  # unique intervals
-intrng <- datf[, diff(range(t_btwn_samp)), by = DOI]
-intcv <- datf[DOI %in% intuni[V1 > 1, DOI], cv(t_btwn_samp), by = DOI]
-intsd <- datf[DOI %in% intuni[V1 > 1, DOI], log10(sd(t_btwn_samp)), by = DOI]
-intmu <- datf[DOI %in% intuni[V1 > 1, DOI], log10(mean(t_btwn_samp)), by = DOI]
-
-# Plot Resolution Statistics
-round(c("%" = length(which(resuni$V1 > 1)) / resrng[, .N],  # % > 1 res 
-        "medrng" = resrng[which(resuni$V1 > 1), median(V1)],  # median res rng
-        "murng" = resrng[which(resuni$V1 > 1), mean(V1)] / 10000,  # mu res rng
-        "rescv" = rescv[, mean(V1)], # mean resolution CV
-        "resmu" = resmu[, mean(V1)],
-        "muN" = resuni[, mean(V1)]), 2)  # mean number of resolutions
-
-# Plot Resolution Statistics
-round(c("%" = length(which(intuni$V1 > 1)) / intrng[, .N],  # % > 1 interval
-        "medrng" = intrng[which(intuni$V1 > 1), median(V1)],  # median int range
-        "murng" = intrng[which(intuni$V1 > 1), mean(V1)],  # mean interval range
-        "intcv" = intcv[, mean(V1)], # mean intolution CV
-        "intmu" = intmu[, mean(V1)],
-        "muN" = intuni[, mean(V1)]), 2)  # mean number of intolutions
-```
-
-<a href="#top">Back to top</a>
-
-### Normal bootstrapping
-
-```{r, eval = FALSE, echo = FALSE}
-### Bootstrapping
-# Good old-fashioned bootstrapping with replacement, no perturbation of value
-# bootstrap
-set.seed(1)
-bootsamp <- lapply(1:1000, function(x) {
-  if((x / 100) %in% 1:100) print(x)
-  dnew <- copy(datf[, .(plot_res, act_ext, t_btwn_samp, act_dur)])
-  dnew <- dnew[sample(1:.N, size = .N, replace = TRUE), ]
-  dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
-  dnew
-})
-
-```
-
-<a href="#top">Back to top</a>
-
-### Normal bootstrapping plot
-```{r, eval = FALSE, echo = FALSE}
-### Bootstrap - deprecated for now
-
-bootdt <- rbindlist(bootsamp)
-resdat2 <- data.table(x = log10(bootdt$t_btwn_samp), y = log10(bootdt$plot_res))
-resdat2$n <- rep(1, nrow(resdat2))
-rx <- range(resdat2$x)
-rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-rx[2] <- 6.56
-resdat2[x < rx[1], x := rx[1]]
-resdat2[x > rx[2], x := rx[2]]
-ry <- range(aaxis1$logres)
-ry[1] <- -5  # set limits to 0.01 cm^2
-resdat2[y < ry[1], y := ry[1]]
-resdat2[y > ry[2], y := ry[2]]
-coordinates(resdat2) <- ~x + y
-
-# extent
-extdat2 <- data.table(x = log10(bootdt$act_dur), 
-                      y = log10(bootdt$act_ext))
-extdat2$n <- rep(1, nrow(extdat2))
-ex <- range(taxis2$logdays)
-ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
-extdat2[x < ex[1], x := ex[1]]
-extdat2[x > ex[2], x := ex[2]]
-ey <- range(aaxis2$logarea)
-extdat2[y < ey[1], y := ey[1]]
-extdat2[y > ey[2], y := ey[2]]
-coordinates(extdat2) <- ~x + y
-
-# densities
-# resolution
-rres2 <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat2, 1)
-rres2 <- (rres2 / cellStats(rres2, sum)) * 100
-rng <- range(rres2[is.finite(rres2)])
-bwidth <- (rng[2] - rng[1]) / 30
-brks <- seq(rng[1], rng[2], bwidth)
-brklabs <- seq(0, 0.1, 0.02)#round(brks[seq(1, length(brks), 5)], 2)
-
-# extent
-rext2 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat2, 1)
-rext2 <- (rext2 / cellStats(rext2, sum)) * 100
-rng <- range(rext2[is.finite(rext2)])
-# rng[2] <- 0.1
-bwidth <- (rng[2] - rng[1]) / 30
-brks2 <- seq(rng[1], rng[2], bwidth)
-brklabs2 <- brklabs #round(brks2[seq(1, length(brks2), 3)], 2)
-
-cxa = 0.7
-pdf("paper/figures/res_v_extent_bstrap2.pdf", width = 7, 
-    height = 2.5)
-par(mfrow = c(1, 2), mar = c(4, 4, 0.3, 3), oma = c(0, 0, 0, 1), 
-    mgp = c(2, 0.5, 0))
-# image(rres, col = cols, breaks = brks,
-image(rres2, col = rev(terrain.colors(30)), breaks = brks,
-     axes = FALSE, xlab = "", ylab = "")
-# plot(rres2, col = rev(terrain.colors(20)))
-axis(1, at = taxis1$logdays[-c(5, 9, 10)], labels = tlab1[-c(5, 9, 10)], 
-     las = 2, tcl = -0.2, cex.axis = cxa)
-axis(2, at = aaxis1$logres[-1], labels = alab1[-1], las = 2, tcl = -0.2, 
-     cex.axis = cxa)
-# plot(rres2, axes = FALSE, legend.only = TRUE)
-aargs <- list(mgp = c(3, 0.25, 0), at = brklabs, labels = brklabs, 
-              cex.axis = cxa, tcl = -0.1)
-plot(rres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, breaks = brks, col = rev(terrain.colors(30)))
-
-image(rext2, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
-      ylab = "", breaks = brks2)
-axis(1, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
-     cex.axis = cxa, tcl = -0.2)
-axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
-     tcl = -0.2)
-aargs <- list(mgp = c(3, 0.25, 0), at = brklabs2, labels = brklabs2, 
-              cex.axis = cxa, tcl = -0.1)
-plot(rext2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, breaks = brks2, col = rev(terrain.colors(30)))
-dev.off()
-
-# datf[, plot(log10(act_dur), log10(act_ext), xlim)]
-```
-
-<a href="#top">Back to top</a>

### 2017-05-14 Rmd/space-time.Rmd
Paper revised through end of text describing Figure 2.

@@
 aaxis2[, logarea := log10(area)]
+```
+```{r, eval = FALSE, echo = FALSE}
+save(aaxis1, aaxis2, taxis1, taxis2, alab1, alab2, tlab1, tlab2, 
+     file = fp(p_dat, "derived/dimbreaks.rda"))
 ```
@@
 ```
+```{r, echo = FALSE, eval = FALSE}
+save(hdat, hdatstat_bt, file = fp(p_dat, "derived/hdatstat.rda"))
+```
 
 Statistics associated with Figure 1: percentage of observations falling within different scale ranges, and associated with different observational methods
-```{r, eval = FALSE}
+```{r, echo = FALSE}
+load(fp(p_dat, "derived/hdatstat.rda"))
+load(fp(p_dat, "derived/dimbreaks.rda"))
+```
+```{r}
 # res
-# hdatst <- hdatstat[[1]]$density
+# sum(hdatst[4:6])  # 53 % between 10cm^2 and 1 m^2
+# sum(hdatst[1:6])  #  67% < 1m^2
+# sum(hdatst[7:10]) # 24 % between 1 m^2 and 1 ha
+# sum(hdatst[11:(l)]) # 9 % 1-10000 ha
+# dat[plot_res >= 10000^2]
 hdatst <- hdatstat_bt$res$mu
 l <- length(hdatst)
-sum(hdatst[4:6])  # 53 % between 10cm^2 and 1 m^2
-sum(hdatst[1:6])  #  67% < 1m^2
-sum(hdatst[7:10]) # 24 % between 1 m^2 and 1 ha
-sum(hdatst[11:(l)]) # 9 % 1-10000 ha
-# dat[plot_res >= 10000^2]
+hds <- c("10cm^2 - 1 m^2", "<1m^2", "1 m^2-1 ha", "1-10000 ha")
+inds <- list(4:6, 1:6, 7:10, 11:l)
+sapply(1:length(inds), function(x) {
+  v <- round(sum(hdatst[inds[[x]]]), 1)
+  names(v) <- hds[x]
+  v
+})
 
@@
 l <- length(hdatst)
-sum(hdatst[1:2])  # 36 % < 1 m^2
-sum(hdatst[1:6])  # 82% < 1 ha
-sum(hdatst[1:7])  # 86% < 10 ha
-sum(hdatst[1:8])  # 90 % < 100 ha
-sum(hdatst[1:9])  # 92 % < 1000 ha
-sum(hdatst[8:9])  # 6 % 10-1000 ha
-sum(hdatst[8:l])  # 14% > 10 ha
-sum(hdatst[10:l]) # 8.2 % covered an area > 1000 ha
-sum(hdatst[12:l]) # 5.1 % covered an area > 100,000 ha
-sum(hdatst[13:l]) # 3.9 % covered an area > 1,000,000 ha
-
-# length(which(hdat$ext < -4)) / length(hdat$ext)
-# length(which(hdat$ext > 3)) / length(hdat$ext)
-# length(which(hdat$ext > 1 & hdat$ext <= 3)) / length(hdat$ext)
-# which(hdatstat[[2]]$breaks == log10(1000000))
-# length(hdatstat[[2]]$counts[7:8])
-# hdatstat[[2]]$mids[12]
+# sum(hdatst[1:2])  # 36 % < 1 m^2
+# sum(hdatst[1:6])  # 82% < 1 ha
+# sum(hdatst[1:7])  # 86% < 10 ha
+# sum(hdatst[1:8])  # 90 % < 100 ha
+# sum(hdatst[1:9])  # 92 % < 1000 ha
+# sum(hdatst[8:9])  # 6 % 10-1000 ha
+# sum(hdatst[8:l])  # 14% > 10 ha
+# sum(hdatst[10:l]) # 8.2 % covered an area > 1000 ha
+# sum(hdatst[12:l]) # 5.1 % covered an area > 100,000 ha
+# sum(hdatst[13:l]) # 3.9 % covered an area > 1,000,000 ha
+hds <- c("<1 ha", "<10 ha", "<100 ha", "<1000ha", "10-1K ha", ">10ha", "1K-1MK",
+         ">1Kha", ">100Kha", ">1Mha")
+inds <- list(1:6, 1:7, 1:8, 1:9, 8:9, 8:l, 10:12, 10:l, 12:l, 13:l)
+sapply(1:length(inds), function(x) {
+  v <- round(sum(hdatst[inds[[x]]]), 1)
+  names(v) <- hds[x]
+  v
+})
 
@@
 l <- length(hdatst)
-sum(hdatst[l])  # 36% are once-offs
-sum(hdatst[1:3])  # 15 % < daily
-sum(hdatst[1:4])  # 25 % < weekly
-sum(hdatst[4:5])  # 21% daily up to monthly
-sum(hdatst[6])  # 19% monthly up to yearly
-sum(hdatst[7:8]) # 8.2 % yearly to decadal
+# sum(hdatst[l])  # 36% are once-offs
+# sum(hdatst[1:3])  # 15 % < daily
+# sum(hdatst[1:4])  # 25 % < weekly
+# sum(hdatst[4:5])  # 21% daily up to monthly
+# sum(hdatst[6])  # 19% monthly up to yearly
+# sum(hdatst[7:8]) # 8.2 % yearly to decadal
+hds <- c("norep", "<daily", "<weekly",  "day-mo", "mo-yr", "yr-dec")
+inds <- list(l, 1:3, 1:4, 4:5, 6, 7:8)
+sapply(1:length(inds), function(x) {
+  v <- round(sum(hdatst[inds[[x]]]), 1)
+  names(v) <- hds[x]
+  v
+})
 
@@
 l <- length(hdatst)
-sum(hdatst[1:3])  # 71 % up to 1 day
-sum(hdatst[4:5])  # 17% covered between 1 day and 1 month
-sum(hdatst[6])  # 7.3% covered between 1 month and 1 year
-# sum(hdatstat[[4]]$density[4:6])  # 26% covered between 1 day and 1 year
-sum(hdatst[7:10]) # 4.5% year to century
+# sum(hdatst[1:3])  # 71 % up to 1 day
+# sum(hdatst[4:5])  # 17% covered between 1 day and 1 month
+# sum(hdatst[6])  # 7.3% covered between 1 month and 1 year
+# # sum(hdatstat[[4]]$density[4:6])  # 26% covered between 1 day and 1 year
+# sum(hdatst[7:10]) # 4.5% year to century
+hds <- c("<=daily", "day-mo", "mo-yr", "yr-dec", "yr-cent")
+inds <- list(1:3, 4:5, 6, 7, 8:10)
+sapply(1:length(inds), function(x) {
+  v <- round(sum(hdatst[inds[[x]]]), 1)
+  names(v) <- hds[x]
+  v
+})
 
@@
 # dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
-datf[like(study_type, "field"), plot(log10(plot_res), log10(act_ext), 
-                                     ylim = c(-10, 10))]
-datf[like(study_type, "remote"), 
-     points(log10(plot_res), log10(act_ext), pch = 20, col = "red")]
-datf[like(study_type, "automated"), 
-     points(log10(plot_res), log10(act_ext), pch = 20, col = "blue")]
-lines(aaxis1$logres, aaxis2$logarea[-c(16:17)] )
-datf[, .N / nrow(datf) * 100, by = st]
-
-datf[, plot(log10(t_btwn_samp), log10(act_dur), pch = 20)]
-datf[which(hdat[[3]] < -1.380211), 
-     points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "blue")]
-datf[like(study_type, "automated"), 
-     points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "red")]
+# datf[like(study_type, "field"), plot(log10(plot_res), log10(act_ext), 
+#                                      ylim = c(-10, 10))]
+# datf[like(study_type, "remote"), 
+#      points(log10(plot_res), log10(act_ext), pch = 20, col = "red")]
+# datf[like(study_type, "automated"), 
+#      points(log10(plot_res), log10(act_ext), pch = 20, col = "blue")]
+# lines(aaxis1$logres, aaxis2$logarea[-c(16:17)] )
+# datf[, .N / nrow(datf) * 100, by = st]
+
+# datf[, plot(log10(t_btwn_samp), log10(act_dur), pch = 20)]
+# datf[which(hdat[[3]] < -1.380211), 
+#      points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "blue")]
+# datf[like(study_type, "automated"), 
+#      points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "red")]
 
@@
 # unique(datf$study_type)
-datf[like(study_type, "field"), .N] / nrow(datf)  # 80% field studies
-datf[like(study_type, "auto"), .N] / nrow(datf) # 12 % automated
-datf[like(study_type, "auto") & log10(t_btwn_samp) < -1.380211, .N] / 
-  length(which(hdat[["int"]] < -1.380211))  # 78 % high frequency obs automated 
-datf[like(study_type, "remote|other"), .N] / nrow(datf) # 7 % remote
-datf[like(study_type, "paleo"), .N] / nrow(datf) * 100 # <1 % paleo
-
+datf[, round(.N / nrow(datf) * 100, 2), by = study_type]
+round(datf[study_type == "automated" & log10(t_btwn_samp) < -1.380211, .N] / 
+        hdat[int < -1.380211, .N] * 100, 2)
 ```
@@
      cex.axis = cxa)
-mtext("Sampling interval", 1, cex = cxa, line = ll)
-mtext("Spatial resolution", 2, cex = cxa, line = ll)
+mtext("Interval", 1, cex = cxa, line = ll)
+mtext("Resolution", 2, cex = cxa, line = ll)
 brks <- brkfun(rres3, 0.01, cuts)  # breaks
@@
               cex.axis = cxa, tcl = -0.1)
-mtext("Temporal duration", 1, cex = cxa, line = ll)
-mtext("Spatial extent", 2, cex = cxa, line = ll)
+mtext("Duration", 1, cex = cxa, line = ll)
+mtext("Extent", 2, cex = cxa, line = ll)
 plot(rext3, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
@@
      cex.axis = cxa)
-mtext("Spatial resolution", 1, cex = cxa, line = ll)
-mtext("Spatial extent", 2, cex = cxa, line = ll)
+mtext("Resolution", 1, cex = cxa, line = ll)
+mtext("Extent", 2, cex = cxa, line = ll)
 brks <- brkfun(sres, 0.01, cuts)  # breaks
@@
      cex.axis = cxa, tcl = -0.2)
-mtext("Sampling interval", 1, cex = cxa, line = ll)
-mtext("Temporal duration", 2, cex = cxa, line = ll)
+mtext("Interval", 1, cex = cxa, line = ll)
+mtext("Duration", 2, cex = cxa, line = ll)
 # brks <- brkfun(tres, 0.02, 100)  # breaks
@@
 # image(rres3, col = rev(terrain.colors(100)))
+# image(rext3, col = rev(terrain.colors(100)))
+# datf[, points(log10(eff_dur), log10(eff_ext), pch = 20, cex = 0.5)]
+# datf[like(study_type, "remote"), {
+#   points(log10(eff_dur), log10(eff_ext), pch = 20, cex = 0.5, col = "blue")
+# }]
 # points(resdat3$x, resdat3$y, pch = 20, cex = 0.5)
 # image(sres, col = rev(terrain.colors(100)))
+# datf[, points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5)]
+# datf[like(study_type, "remote"), {
+#   points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5, col = "blue")
+# }]
+# datf[like(study_type, "field"), {
+#   points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5, col = "red")
+#   abline(lm(log10(plot_res) ~ log10(eff_ext)))
+# }]
 # points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)

### 2017-05-14 paper/main.tex
Paper revised through end of text describing Figure 2.

@@
 
-We calculated these dimensions from 379 discrete observations (defined here as data collected from non-experimentally manipulated, or ``natural'' \cite{tilman_ecological_1989}, systems) reported within a 134 paper subset of 348 randomly selected articles (from 42,918 total).
+We calculated these dimensions from 379 discrete observations reported within a 134 paper subset of 348 randomly selected articles (from 42,918 total).
 An additional 62 papers that were cited as the source of observations in the selected papers were also reviewed.
+We confined our analysis to observations made of ``natural'' \cite{tilman_ecological_1989}, or non-experimentally manipulated, systems, given that the inclusion of experiments could have skewed our assessments towards the relatively fine scales at which such studies typically focus.
 
@@
 1A).
-The total spatial extent (the number of sampled sites multiplied by the spatial resolution) of 86\% of observations was $<$10 ha, while 36\% covered less than 1 m$^2$ (Fig.
+The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 42\% 1000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig.
 1B).
-Only 14\% covered an extent $\geq$10 ha, with just 4\% spanning areas $\geq$1 million ha.
 
@@
 \vspace{-0.15 cm}
-\caption{Histograms of the spatial resolution (A) and extent (B), sampling interval (C) and temporal duration (D) of ecological observations collected from the surveyed ecological studies.
+\caption{Histograms of the resolution (A), extent (B), interval (C), and duration (D) of observations collected from the surveyed ecological studies.
 Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval.
@@
 
-In the temporal dimensions, 36\% of observations were not repeated (Fig.
-1C), 15\% were repeated at short intervals (sub-second to daily), 21\% at daily to monthly intervals, 19\% at monthly to yearly intervals, and 8\% at yearly to decadal intervals.
-The temporal duration of studies (the total amount of time the ecological feature was directly observed, calculated as the number of repeat observations multiplied by the effective sampling duration; SI) was less than one day for 71\% of sampled observations, between one day and one month for 17\% of observations, between one month and one year for 7\% of observations, and greater than one year for 5\% of observations (including several paleoecological studies covering centuries to millennia; Fig.
+In the temporal dimensions, 37\% of observations were not repeated (Fig.
+1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, and 8\% at yearly to decadal intervals.
+Duration was one day or less for 31\% of sampled observations, while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 10\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig.
 1D).
 
-Juxtaposing the observational dimensions provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (Fig.
+Juxtaposing these observational dimensions provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (Fig.
 2).
-Contrasting spatial resolution with sampling interval reveals that the majority of temporally replicated observations (the 36\% that were unreplicated were excluded because they have no interval value) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig.
+Contrasting resolution with interval reveals that the majority of temporally replicated observations (the 37\% that were unreplicated were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig.
 2A).
 A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident.
-This lower right to upper left orientation reflects the tradeoff between resolution and return interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations.
+This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations.
 This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
@@
 \vspace{-0.15 cm}
-\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by A) sampling interval and spatial resolution, B) temporal duration and spatial extent, C) spatial resolution and spatial extent, and D) sampling interval and temporal duration.
-Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.
-}
+\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by A) interval and resolution (of temporally replicated observations only), B) duration and extent, C) resolution and extent, and D) interval and duration (of temporally replicated observations).
+Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.}
 \label{afoto1}
@@
 
-Contrasting temporal duration and spatial extent (for all observations) reveals that the bulk of ecological observations covered total time periods of one minute to $<$1 month and areas of $\leq$1 m$^2$ (Fig.
+Contrasting duration and extent (for all observations) reveals two primary domains of observational concentration.
+The first consists of observations spanning one month to one decade in time and 10-1000 ha in space, while the second is defined by observations of one year to several decades that cover 10,000 to 1,000,000 ha (Fig.
+2B).
+Three other notable, but lesser areas of concentration are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($<$1 day) of either 1-10 ha or 10,000-1,000,000 ha.
+
+% the bulk of ecological observations span one month to one decade in time and covered total time periods of one minute to $<$1 month and areas of $\leq$1 m$^2$ (Fig.
 2B).
@@
 
-Comparing the two spatial dimensions against one another (for all observations) reveals a log-linear relationship wherein resolution strongly constrains extent, such that the ratio between extent and resolution falls as the latter increases (Fig.
+Comparing the two spatial dimensions against one another (for all observations) shows a primary concentration of observations with 10 cm$^2$ to 100 m$^2$ resolution that have extents ranging between slightly over 1,000 to nearly 1,000,000 ha (Fig.
+2C).
+The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha.
+These three concentrations suggest a tendency for observational extent to increase with resolution, which is a relationship that becomes more pronounced in the (less densely observed) portion of the domain where resolutions $\geq$100 m$^2$.
+
+%reveals a log-linear relationship wherein resolution strongly constrains extent, such that the ratio between extent and resolution falls as the latter increases (Fig.
 2C).
@@
 
-A similar linear relationship between sampling interval and temporal duration was not evident (unreplicated observations were excluded because they lacked interval values).
+A similar tendency for duration to increase with interval was also evident amongst temporally replicated observations (Fig.
+2D), where the majority of observations were repeated at daily to decadal intervals and spanned  $\geq$ 1 month to $<$100 years.
+The orientation of this concentration shows that observations lasting one year to one decade tend to have corresponding intervals, while those lasting one month to one year have daily to monthly intervals.
+This suggests that most long-duration observations have just a single temporal replicate.
+The fainter densities of observations having sub-daily intervals shows that relatively few high frequency, long duration ecological measurements are undertaken.
+
+%A similar linear relationship between sampling interval and temporal duration was not evident (unreplicated observations were excluded because they lacked interval values).
 For repeated observations, the greatest concentration was among those made at daily to yearly intervals with overall durations of hours to months (Fig.

### 2017-05-15 Rmd/space-time.Rmd
Added Figure 3 to text and associated text.  Edited Figure 3 to add percetage values for each bin atop whiskers.

@@
 #   points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5, col = "blue")
+#   abline(lm(log10(eff_ext) ~ log10(plot_res)))
 # }]
@@
 #   points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5, col = "red")
-#   abline(lm(log10(plot_res) ~ log10(eff_ext)))
+#   abline(lm(log10(eff_ext) ~ log10(plot_res)))
 # }]
@@
   dd[aext < ex[1], aext := ex[1]]  # set to min
-  # dd[, list("int" = findInterval(aext, eint), "dif" = dif)]
+  # dd[eext < ex[1], eext := ex[1]]  # set to min
   dd[, int := findInterval(aext, eint, rightmost.closed = TRUE)]
+  # dd[, int := findInterval(eext, eint, rightmost.closed = TRUE)]
   # dd[, list("dif" = mean(dif, na.rm = TRUE), "N" = .N), by = int][order(int)]
@@
 ewgt <- pctobs / sum(pctobs)
-# ealphs <- (ewgt - min(ewgt)) / diff(range(ewgt))
+# round(cumsum(ewgt), 2)
 ewgtint <- c(0, 1, 5, 10, 15, 20, 25) / 100  # weight intervals
@@
 dwgt <- pctobs / sum(pctobs)
-# dalphs <- (dwgt - min(dwgt)) / diff(range(dwgt))
+# round(cumsum(dwgt), 2)
+
 dwgtint <- c(0, 1, 5, 10, 15, 20, 25, 30) / 100  # weight intervals
@@
             bfill = cols[ecols][i])
+  text(bx[i], ebox[[i]][5] + 0.15, round(ewgt[i] * 100, 1), cex = 0.3)
 } 
@@
 mtext("Actual Extent", side = 1, line = 3, cex = 0.9)
-mtext("Difference Between", side = 2, line = 2.1, cex = 0.9)
-mtext("Effective and Actual (Decades)", side = 2, line = 1.4, cex = 0.9)
+mtext("Difference between Extent/Duration", side = 2, line = 2.2, cex = 0.9)
+mtext("and Actual Extent/Duration (Decades)", side = 2, line = 1.4, cex = 0.9)
 lrng <- c(mean(ex), max(ex) - 0.05 * max(ex))
@@
 bd <- abs(dint[-length(dint)] - dint[-1])
-# cols <- sapply(dalphs, function(x) rgb(1, 0, 0.1, alpha = x))
 cols <- brewer.pal(length(dwgtint) - 1, "Reds")
@@
 for(i in 1:length(tbox)) {
-  boxplot_v(bx[i], y = tbox[[i]], n = 0.9, inhgt = bd[i] * 10, whiskhgt = 0.5, 
+  boxplot_v(bx[i], y = tbox[[i]], n = 0.9, inhgt = bd[i] * 10, whiskhgt = 0.75, 
             bfill = cols[dcols][i])
+  text(bx[i], tbox[[i]][5] + 0.15, round(dwgt[i] * 100, 1), cex = 0.3)
+
 } 

### 2017-05-15 paper/main.tex
Added Figure 3 to text and associated text.  Edited Figure 3 to add percetage values for each bin atop whiskers.

@@
 
-%reveals a log-linear relationship wherein resolution strongly constrains extent, such that the ratio between extent and resolution falls as the latter increases (Fig.
-2C).
-As with interval, this pattern indicates that spatial replication diminishes with plot size \cite{kareiva_spatial_1988} in field-based studies, which accounted for 80\% of observations.
-This stands in contrast to the 6.5\% of observations that were made with remote sensing, which have fine spatial resolutions relative to their large extents.
-These comprise the majority of observations in the four low density patches (e.g.
-0.1-10 ha resolution and 1-10 million ha extent) falling above the denser line that terminates at 100 ha resolution and 1000 ha extent.
-
 A similar tendency for duration to increase with interval was also evident amongst temporally replicated observations (Fig.
@@
 This suggests that most long-duration observations have just a single temporal replicate.
-The fainter densities of observations having sub-daily intervals shows that relatively few high frequency, long duration ecological measurements are undertaken.
+The low densities of observations having sub-daily intervals shows that relatively few high frequency, long duration ecological measurements are undertaken.
 
-%A similar linear relationship between sampling interval and temporal duration was not evident (unreplicated observations were excluded because they lacked interval values).
-For repeated observations, the greatest concentration was among those made at daily to yearly intervals with overall durations of hours to months (Fig.
-2D).
-A second concentration of sub-monthly to yearly observations lasting one second or less is also evident.
-Very few repeat observations exceeded a year in duration, and these were collected at daily to yearly intervals.
+To provide further insight into the spatio-temporal domains of ecological observations, we also evaluated the differences in scale between extent and actual extent and between duration and actual duration.
+To make these comparisons, we calculated the magnitudes of difference (decades) between each dimension and its actual equivalent, and examined how these varied as a function of the scale of the actual measurement (Fig.
+3).
+The majority (80\%) of the assessed observations had actual extents of $\leq$1 ha, which on average was 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).
+Actual extent converged with extent above 100,000 ha, but this applies to just 5\% of observations.
+The actual duration of 65\% of observations is $<$ 1 day, which is on average 3 to nearly 9 orders of magnitude shorter than the time span covered by temporal replicates (Fig.
+3B).
+The two duration measures were effectively the same for the 16\% of observations exceeding one month of actual duration.
+
+\begin{figure}[ht]
+\includegraphics[width=1\textwidth]{figures/act_v_eff_diff.png}
+\vspace{-0.2 cm}
+\caption{The difference between \emph{actual} extent (the summed area of spatial replicates) and extent (A) and \emph{actual} duration (the summed sampling duration across temporal replicates) and duration (B).
+Differences are expressed as decades, or how many orders of magnitude greater the extent or duration is then the actual extent or duration, and are summarized (as box plots, with circle in box representing the mean and line the median) in bins representing different levels of actual extent/duration.
+The percentages of observations falling within each bin are indicated by the color of the inter-quartile and the numeric value above the upper whisker.}
+\label{afoto1}
+\end{figure}
 
@@
 \noindent \textbf{\emph{Insights into the scales of modern ecology}}\\
+%These results indicate that for most ecological observations are collected with spatially discontinuous replicates that sample very small portions of the area in which they are distributed.
+
 Despite these caveats, our results provide valuable insight into the spatial and temporal domains being addressed by modern ecological research.

### 2017-05-24 Rmd/space-time.Rmd
Updated analysis of annual change in observing methods and scales. Revision of main text to accomodate this change, as well as addition of related figures to SI.  Fixes on explanation of methods/FAQ in SI.

@@
 library(viridis)
+# library(zyp)
 
 p_root <- set_base_path("ecoscales")
-p_dat <- full_path(p_root, "external/data/result")
-p_calib <- full_path(p_root, "external/data/calibration")
+p_dat <- fp(p_root, "external/data/result")
+p_calib <- fp(p_root, "external/data/calibration")
+p_fig <- fp(p_root, "paper/figures")
+
 # dat <- as.data.table(read_excel(full_path(p_dat, "merged.xlsx")))
@@
                      samp_dur, t_btwn_samp, act_dur, eff_dur)])
-selind <- 3:ncol(dat2)
+# selind <- 3:ncol(dat2)
+selind <- 4:ncol(dat2)
 for(j in names(dat2)[selind]) set(dat2, j=j, value = as.numeric(dat2[[j]]))
@@
 
+### Study type by year
+
+Calculate types of study by year, as total and percent
+```{r, warning = FALSE, message = FALSE, fig.height=3}
+# One Oecologia study listed as 2004 (by Endnote) was actually 2003 
+# Set it to 2004
+stypes <- copy(dat)  # new dataset out of dat (which retain year info)
+stypes[study_year < 2004, study_year := 2004]
+
+# bring in study_type from calr set, grabbing year from original cal dt
+calyr <- cal[DOI %in% calr[, unique(DOI)], unique(study_year), by = DOI]
+calyr[, V1 := as.numeric(V1)]
+calyr <- calyr[, list("study_year" = mean(V1, na.rm = TRUE)), by = DOI]
+calyr <- merge(calyr, calr[, list(DOI, study_type)], by = "DOI")
+
+# bind
+stypes <- rbind(stypes[, list(DOI, study_year, study_type)], calyr)
+na_effext <- datf[is.na(eff_ext), unique(DOI)]  # ID NA eff_ext values
+
+# Estimate and plot study type by year
+stypeyr <- stypes[, list(.N, study_type), by = study_year][, {
+  list("tot" = N, "ct" = .N, "prop" = .N / N * 100)
+}, by = .(study_type, study_year)][order(study_year)]
+stypeyr <- stypeyr[, lapply(.SD, mean), .SDcols = c("tot", "ct", "prop"), 
+                   by = .(study_type, study_year)]
+```
+```{r, eval = FALSE}
+obs_types <- c("remote", "field", "auto")
+obs_types2 <- c("Remote sensing", "Field", "Automatic")
+cols <- c("orange", "green4", "blue4")
+png(fp(p_fig, "stype_by_yr.png"), width = 7, height = 2.75, res = 300, 
+    units = "in")
+par(mfrow = c(1, 3), oma = c(0, 3, 0, 0), mar = c(4, 1, 2, 1))
+for(i in 1:3) {  # i <- 1
+  plot(c(2004, 2014), c(0, 100), xlab = "", ylab = "", pch = "")
+  stypeyr[like(study_type, obs_types[i]), {
+    points(study_year, prop, col = cols[i], pch = 20)
+    yr = as.numeric(study_year)
+    slm = lm(prop ~ yr, weights = tot)  # weighted regression
+    slms = summary(slm)
+    abline(slm, col = cols[i])
+    cf = round(coef(slm)[1:2], 4)
+    pf = round(coef(slms)[8], 2)
+    rf = round(slms$adj.r.squared, 2)
+    cfa = paste0(cf[2], "% year")
+    text(x = 2006, y = 100, labels = substitute(x^-1, list(x = cfa)), 
+         cex = 0.9)
+    text(x = 2011.5, y = 100, cex = 0.9,
+         labels = substitute(R^2~"="~x~"; p <"~y, list(x = rf, y = pf)))
+    if(i == 1) mtext("percent", side = 2, line = 2.5, cex = 0.8)
+    if(i == 2) mtext("year", side = 1, line = 2.5, cex = 0.8)
+    mtext(obs_types2[i], side = 3, cex = 0.8)
+  }]
+}
+dev.off()
+```
+
+Extrapolating percentage of studies using remote sensing to 2017
+```{r}
+preds <- stypeyr[like(study_type, "remote"), {
+  slm = lm(prop ~ study_year, weights = tot)
+  predict(slm, newdata = data.frame("study_year" = 2004:2017))
+}]
+names(preds) <- 2004:2017
+```
+
+Percent of remote sensing studies by year
+```{r}
+round(preds, 1)
+```
+
+Average annual percent of studies using remote sensing 2004-2014, and projected increases in this percent through 2017 (2004-2017) 
+```{r}
+# murs2014 <- stypeyr[study_type == "remote", round(mean(prop), 2)]
+murs2014 <- stypeyr[, sum(ct[study_type == "remote"]) / sum(ct) * 100]
+pctinc <- mean(preds) / mean(preds[1:(length(preds) - 3)])
+
+pcttab <- cbind("2014" = murs2014, "pct increase" = (pctinc - 1) * 100,
+                "2017" = murs2014 * pctinc)
+round(pcttab, 1)
+```
+
+```{r}
+types <- datf[, .N, by = study_type]
+rs2017 <- mean(preds)  # avg predicted % using RS through 2004-2017
+rs2014 <- mean(preds[1:(length(preds) - 3)])  # avg predicted % RS 2004-2014
+# rs2017 / rs2014
+
+# translate to N more RS studies would have been selected given those %ages
+rsinc <- rs2017 / rs2014 * types[study_type == "remote", N] - 
+                 types[study_type == "remote", N]
+
+remote <- datf[study_type == "remote", eff_ext]
+notremote <- datf[study_type != "remote", eff_ext]
+# a <- (length(remote) + rsinc) / (nrow(datf))
+# b <- length(remote) / nrow(datf)
+# a / b
+# ((length(remote) + rsinc) / (length(notremote)) / length(b))
+
+# (rsext * rs2017 + nonrsext * (1 - rs2017)) / 
+#   (rsext * rs2014 + nonrsext * (1 - rs2014))
+
+set.seed(1)
+ext2017 <- unlist(lapply(1:1000, function(x) {  # x = 1
+  ind <- sample(1:length(remote), size = round(rsinc))
+  mean(c(remote, remote[ind], notremote), na.rm = TRUE)
+}))
+
+dec1417 <- log10(mean(ext2017, na.rm = TRUE) / 
+                   datf[, mean(eff_ext, na.rm = TRUE)])
+qtiles1417 <- log10(quantile(ext2017, c(0.025, 0.975), na.rm = TRUE) / 
+                      datf[, mean(eff_ext, na.rm = TRUE)])
+
+pctdiff1417 <- mean(ext2017, na.rm = TRUE) / datf[, mean(eff_ext, na.rm = TRUE)]
+pctqtiles1417 <- quantile(ext2017, c(0.025, 0.975), na.rm = TRUE) / 
+                      datf[, mean(eff_ext, na.rm = TRUE)]
+
+cbind("decade" = round(dec1417, 4), "dec2.5" = round(qtiles1417[1], 4),
+      "dec97.5" = round(qtiles1417[2], 4),
+      "pctdiff" = round((pctdiff1417 - 1) * 100, 1),
+      "pctdiff2.5" = round((pctqtiles1417[1] - 1) * 100, 1), 
+      "pctdiff97.5" = round((pctqtiles1417[2] - 1) * 100, 1))
+```
+
+### Change in dimensions by year
+```{r}
+# Examines and plot trends in dimensions
+doiyr <- rbind(unique(dat[, .(DOI, study_year)]), 
+               unique(cal[, .(DOI, study_year)]))
+doiyr <- doiyr[study_year %in% 2004:2014]
+datf2 <- copy(datf)
+datf2 <- merge(doiyr, datf2, by = "DOI")  # collect year of publication
+# datf2[, N := .N, by = study_year]
+```
+```{r, eval = FALSE}
+vars <- c("plot_res", "eff_ext", "t_btwn_samp", "eff_dur")
+varlabs <- c("Resolution", "Extent", "Interval", "Duration")
+labcols <- c(brewer.pal(8, "Reds")[c(7, 7)], brewer.pal(8, "Blues")[c(7, 7)])
+png(fp(p_fig, "dim_by_yr.png"), width = 7, height = 6, res = 300, 
+    units = "in")
+par(mfrow = c(2, 2), mar = c(1, 3, 1, 1), oma = c(2, 3, 0, 0))
+for(i in 1:length(varlabs)) {  # i <- 1
+  if(i == 3) {
+    DT <- copy(datf2[t_btwn_samp != 0, ])
+  } else {
+    DT <- copy(datf2)
+  }
+  DT[, list(mean(get(vars[i]), na.rm = TRUE), .N), by = study_year][, {
+    dim = log10(V1)
+    yr = as.numeric(study_year)
+    if(i %in% 1:2) rng = c(0, 10)
+    if(i %in% 3:4) rng = c(0, 5)
+    if(i %in% 3:4) {
+      xlabs = seq(2004, 2014, 2)
+    } else {
+      xlabs = rep("", 6)
+    }
+    plot(yr, dim, ylim = rng, xaxt = "n", pch = 20, col = labcols[i], 
+         main = varlabs[i], mgp = c(2, 0.5, 0), tcl = -0.2, ylab = "")
+    axis(1, at = seq(2004, 2014, 2), labels = xlabs, tcl = -0.2, 
+         mgp = c(2, 0.5, 0))
+    slm = lm(dim ~ yr, weights = N)
+    slms = summary(slm)
+    abline(slm, col = labcols[i])
+    cf = round(coef(slm)[1:2], 3)
+    pf = round(coef(slms)[8], 2)
+    rf = round(slms$adj.r.squared, 2)
+    text(x = 2006.25, y = rng[2] * 0.97, cex = 0.9,
+         labels = substitute(x~year^-1, list(x = cf[2])))
+    text(x = 2011.5, y = rng[2] * 0.97, cex = 0.9,
+         labels = substitute(R^2~"="~x~"; p <"~y, list(x = rf, y = pf)))
+    if(i == 1) mtext(expression("log10 m"^2), side = 2, line = 2)
+    if(i == 2) mtext("log10 ha", side = 2, line = 2)
+    if(i %in% 3:4) mtext("log10 days", side = 2, line = 2)
+  }]
+}
+dev.off()
+```
+
+Percent increase in average extent by year
+```{r}
+preds <- datf2[, list(mean(eff_ext, na.rm = TRUE), .N), by = study_year][, {
+  dim = log10(V1)
+  yr = as.numeric(study_year)
+  slm = lm(dim ~ yr, weights = N)
+  predict(slm, newdata = data.frame("yr" = 2004:2017))
+}]
+# preds[14] - preds[13]
+
+# projected percent increase in extent 2014-2017
+pctinc <- mean(preds) / mean(preds[1:(length(preds) - 3)])
+muext2014 <- datf2[, mean(eff_ext, na.rm = TRUE)]
+pcttab <- cbind("2014" = log10(muext2014), "pct increase" = (pctinc - 1) * 100,
+                "2017" = log10(muext2014 * pctinc), 
+                "decade" = log10((pctinc * muext2014) / muext2014))
+round(pcttab, 3)
+```
+
 ### Varying kernel sizes

### 2017-05-24 paper/SI.tex
Updated analysis of annual change in observing methods and scales. Revision of main text to accomodate this change, as well as addition of related figures to SI.  Fixes on explanation of methods/FAQ in SI.

@@
 
-To record the spatio-temporal dimensions of ecological observations, we considered the observational scales to be those at which a total measurement of the feature(s) of interest was made, rather than the scales that might be represented by the observations.
+We recorded six primary dimensions of ecological observations, three related to space and three related to time.
+The space-related dimensions were resolution, extent, and actual extent.
+Here extent was primarily defined as the area falling within a perimeter defined by the outermost spatial replicates, while actual extent was defined as the summed area of all sample plots (i.e.
+N * resolution, where N is the number of spatial replicates, which we also recorded), or the area that ecologists observe in practice.
+In assessing spatial scales, our analysis only considered the Cartesian plane.
+We did not calculate the z, or depth, dimension, although this dimension is of greater importance for certain sub-disciplines of ecology (e.g.
+depth profiles in marine ecology).
+In some cases (primarily paleoecological studies), values extracted from the z-dimension provided temporal information that was used to calculate both the interval and the duration of the observation.
+
+For time dimensions, we extracted information related to interval, duration, and actual duration.
+Duration was defined as the time between the first and last temporal replicate, whereas actual duration quantifies the amount of time spent observing a particular location, which we calculated by multiplying sampling duration (the time spent collecting a single temporal replicate) by the number of temporal replicates.
+
+A full definition of all dimensions and how they were recorded is contained within the answers to the list of questions below.
+This set of Frequently Asked Questions (FAQ) was provided to each observer for initial study and reference, in order to ensure methodological consistency (see next section).
+
+%To record the spatio-temporal dimensions of ecological observations, we considered the observational scales to be those at which a total measurement of the feature(s) of interest was made, rather than the scales that might be represented by the observations.
 As an example, if a paper reported data on plant species diversity collected using 10 1 m$^2$ sub-plots randomly placed within a larger plot of 100 m$^2$, of which 10 were placed, we recorded 1 m$^2$ as the spatial resolution and 100 m$^2$ as the total extent of the observation.
-While the 10 sub-plots might be representative of diversity at larger scales (e.g.
+
+%While the 10 sub-plots might be representative of diversity at larger scales (e.g.
 the particular plant community), that relationship depends both on sample design (number of replicates, placement, grain/resolution) and the properties of the ecosystem \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002}.
@@
 
-Similarly, we did not try to quantify the extent to which values of individual observations (e.g.
+%Similarly, we did not try to quantify the extent to which values of individual observations (e.g.
 a single plot) of an ecological feature might be correlated with spatially or temporally adjacent values.
@@
 
-In assessing spatial scales, our analysis only considered the Cartesian plane.
-We did not calculate the z, or depth, dimension, although this dimension is of greater importance for certain sub-disciplines of ecology (e.g.
-depth profiles in marine ecology).
-In some cases (primarily paleoecological studies), values extracted from the z-dimension provided temporal information that was used to calculate both the interval and the duration of the observation.
-
 %Furthermore quantifying the actual domains of observation is important because highly dynamic spatial and temporal processes, such as the population ecology of riverine fish \cite{fausch_landscapes_2002} or the evolution of wildfires \cite{clements_observing_2007}, may require continuous observation to properly measure, and can vary substantially in the space or time between observations.
-
-A fuller description of the project methodology is contained within the answers to the list of questions below.
-This set of Frequently Asked Questions (FAQ) was provided to each observer for initial study and reference, in order to ensure methodological consistency (see next section).
 
@@
 \vspace{-0.5cm} 
-   \Hitem \emph{\textbf{What is sampling interval, and how do we record it?}} Sampling interval is the time that elapsed between repeat observations of the same point in space or individual organism.
+   \Hitem \emph{\textbf{What is interval, and how do we record it?}} Interval is the time that elapsed between repeated observations of the same point in space or individual organism.
 In many cases, observations will only be made one time--list a value of 0 for these.
@@
 For remote sensing observations you can assume one second (the observations are effectively instantaneous).
-   \Hitem \emph{\textbf{What is the effective study duration, and how do estimate it?}} The effective study duration is the total period of time over which the phenomenon of interest was observed.
-More specifically, in the case of repeated observations, this is the total time that elapsed between when the first and last observations at a given point were made.
+   \Hitem \emph{\textbf{What is the study duration, and how do estimate it?}} The study duration (or, simply, duration) is the total period of time over which the phenomenon of interest was observed.
+More specifically, in the case of repeated observations, this is the total time that elapsed between the first and last observations at a given point in space (or of the same individual organism or community) were made.
 For once-off (unreplicated) observations, this time is equivalent to the sample duration.
-However, there may be cases where once-off observations may have a longer effective duration then the sample duration.
+However, there may be cases where once-off observations may have a longer duration then the sample duration.
 For example, consider a study that counts occurrences of pollinators over three years, using transects that are located in different locations within the broader study area during each year \cite{rollin_differences_2013-1}.
@@
 In this case, we can consider the effective duration to be three years, as the temporal information is encoded in the analysis.
-  \Hitem \emph{\textbf{What is actual study duration, and how do we calculate it?}} The actual study duration is the integral of sampling duration, or the time spent making one observation of the phenomenon in question.
-To clarify, duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample (the effective duration), nor the integral of time spent in observing all spatial replicates.
-  \Hitem \emph{\textbf{Should study duration be the total time spent sampling all sites or the amount of time spent sampling per site (e.g.
-for 5 minute point counts of birds at 10 sites each repeated twice, should we enter 100 minutes (5 minutes X 2 repeats X 10 sites) or 10 minutes (5 minutes X 2 repeats) for duration)?}} As stated above, duration is the total spent observing a single point in space, so in this case that would be 10 minutes (then converted to days, so 10 / (60 * 24))).
-  \Hitem \emph{\textbf{How do we record study duration when there are no repeat observations?}} In these cases, study duration is equal to sampling duration. 
-   \Hitem \emph{\textbf{How do I record sampling interval in cases where the interval is inconsistent?
+  \Hitem \emph{\textbf{What is \emph{actual duration}, and how do we calculate it?}} The actual duration is the integral of sampling duration, or the time spent making one observation of the phenomenon in question.
+To clarify, actual duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample (duration), nor the integral of time spent in observing all spatial replicates.
+  \Hitem \emph{\textbf{Should actual duration be the total time spent sampling all sites or the amount of time spent sampling per site (e.g.
+for 5 minute point counts of birds at 10 sites each repeated twice, should we enter 100 minutes (5 minutes X 2 repeats X 10 sites) or 10 minutes (5 minutes X 2 repeats) for duration)?}} As stated above, actual duration is the total spent observing a single point in space, so in this case that would be 10 minutes (then converted to days, so 10 / (60 * 24))).
+  \Hitem \emph{\textbf{How do we record duration and actual duration when there are no repeat observations?}} In these cases, study duration is equal to sampling duration. 
+   \Hitem \emph{\textbf{How do I record interval in cases where the interval is inconsistent?
 For example, in a study were observations were observations were repeated in 1979, 1980, 1981, 1984, 2007, 2009?}} Find the time between each successive period, and then take the average of that (remember to convert to days!).
 If there are two or more sets of unevenly spaced days for each site/plot/measurement being taken in the study, then find the average interval for each, and average the averages.
-    \Hitem \emph{\textbf{How do you determine the time between samples for paleo-reconstructions?}} Use the minimum estimate for dating precision as the estimate of time between samples (e.g.
+    \Hitem \emph{\textbf{How do you determine the interval for paleo-reconstructions?}} Use the minimum estimate for dating precision as the estimate of time between samples (e.g.
 50 years in the study of European charcoal deposits DOI:10.1111/geb.12090).
@@
 The logic behind this is that in such cases, where a sediment or tree core or similar measurement is being collected, this effectively represents a continuous ``observation'', and the value associated with the minimum (or other reported) interval is typically an average (or another summary statistic like the maximum) of the amount accumulated.
-    \Hitem \emph{\textbf{What about sampling intervals and durations for instrument-collected observations?}} These are often similar in essence to the paleo-reconstruction case.
+    \Hitem \emph{\textbf{What about intervals and durations for instrument-collected, or automatically-collected, observations?}} These are often similar in essence to the paleo-reconstruction case.
 Take the minimum temperature or daily rainfall recorded at a weather station, which require constant observation over 24 hours to report.
-In such cases, the sampling interval and sampling duration are both 24 hours.
+In such cases, the interval and sampling duration are both 24 hours.
 On the other hand, automated logging systems will provide a series of high frequency observations that are collected instantaneously.
-In these cases the sampling duration should be $\leq$second, and the sampling interval should be the period between successive instantaneous measurements.
-    \Hitem \emph{\textbf{How do you handle the time between samples for a case where repeated samples are taken during a season, across several seasons (e.g.``we performed repeat bird counts every 10 days between March and June of 2005, 2006, and 2008'')?}} Since the sampling is focused on seasons, and presumably some season-specific phenomenon (e.g.
-breeding behavior), the reported values should be pegged to the season, not averaged across the full study span (the start and end dates of the study).
+In these cases the sampling duration should be $\leq$second, and the interval should be the period between successive instantaneous measurements.
+    \Hitem \emph{\textbf{How do you treat interval for a case where repeated samples are taken during a season, across several seasons (e.g.``we performed repeat bird counts every 10 days between March and June of 2005, 2006, and 2008'')?}} Since the sampling is focused on seasons, and presumably some season-specific phenomenon (e.g.
+breeding behavior), the reported values should be pegged to the season, not averaged across the duration (the start and end dates of the study).
 So in this example, it would be 10 days.
@@
 \hspace{-1cm}\underline{\textbf{Spatial scales:}}
-  \Hitem \emph{\textbf{What is sampling resolution, and how do I record it?}} This is the finest scale at which a complete measurement of every unit of the quantity of interest is recorded.
+  \Hitem \emph{\textbf{What is resolution, and how do I record it?}} This is the finest scale at which a complete measurement of every unit of the quantity of interest is recorded.
 For example, if the measurement in question is a tree stem count, the resolution is determined by the size of the plot used to record every tree stem.
@@
 In another example, if the reported plot size was 20 x 20 m, but the authors in fact only measured a random selection of, say, grass stems on which they counted aphids within those plots, then use an estimate of the area of the grass stem as the sampling resolution \cite{gagic_agricultural_2012}.
-  \Hitem \emph{\textbf{What is study extent, and how do I record it?}}  Broadly defined, actual study extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares.
-For studies in which the spatial replicates are not spatially contiguous (as with most field-based studies), this means plot resolution (see Q16) multiplied by number of plots.
-For spatially contiguous studies (e.g., those based on remote sensing imagery), it should be the total area covered by the imagery, i.e., pixel resolution multiplied by the number of pixels.
-However, only record the area analyzed by the authors.
-If they used a sub-sample of pixels, the actual extent is the number of those pixels multiplied by pixel resolution.
 
-Effective survey extent is defined as the total extent bounding all spatial replicates, divided by 10,000 to convert to hectares.
-For studies in which spatial replicates are not spatially contiguous, this means the area of the minimum convex polygon bounding all spatial replicates.
+  \Hitem \emph{\textbf{What is study extent, and how do I record it?}} Study extent (or, simply, extent) is defined as the total extent bounding all spatial replicates, divided by 10,000 to convert to hectares.
+For studies in which spatial replicates are not spatially contiguous, this means the area of the minimum polygon bounding all spatial replicates.
 For example, if the study is conducted in a national park, the effective survey extent would be the area of the national park; if the study is conducted in three national parks, the effective survey extent would be the sum of the areas of all national parks.
@@
 Google Earth Pro, QGIS with OpenLayers plugin) to navigate to the region and delineate a minimum convex polygon surrounding the plots/transects/sampling units to calculate the area in hectares.
-When studies focus on portions of survey regions (e.g., mangroves in a coastal National Park), attempt to delineate the focal portions (the mangroves) and not the larger survey region (entire National Park) using Google Earth Pro.
+When studies focus on portions of survey regions that are clearly distinct from their surroundings (e.g., mangroves in a coastal National Park), try to delineate the focal portions (the mangroves) and not the larger survey region (entire National Park) using Google Earth Pro or a similar application with digitizing and area estimation capabilities.
 
 For spatially contiguous studies (e.g.
-those based on satellite imagery), the effective survey extent is the total area covered by the imagery (in such cases, effective survey extent equals actual survey extent).
-However, if spatially contiguous studies only use a sub-sample of pixels, the effective survey extent is the area of the minimum convex polygon bounding all pixels (calculated following the methods above).
+those based on satellite imagery), the extent is the total area covered by the imagery (in such cases, extent equals actual extent), but only records the area of imagery analyzed by the authors (e.g.
+if the study area required four Landsat scenes to cover, but covered only the inner quarter of each image, report the extent as the summed area of the four quarters).
+However, if spatially contiguous studies only use a sub-sample of pixels, extent is the area of the minimum convex polygon bounding all pixels (calculated following the methods above).
 
@@
 observations from all individuals and times) to define extent.
+
+  \Hitem \emph{\textbf{What is \emph{actual extent}, and how do I record it?}}  Actual extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares.
+For studies in which the spatial replicates are not spatially contiguous (as with most field-based studies), this means resolution (see Q17) multiplied by number of plots.
+For spatially contiguous studies (e.g., those based on remote sensing imagery), it should be the total area covered by the imagery, i.e., pixel resolution multiplied by the number of pixels.
+However, as with extent, only record the area analyzed by the authors.
+If they used a sub-sample of pixels, the actual extent is the number of those pixels multiplied by pixel resolution.
 
@@
 
-  \Hitem \emph{\textbf{What about studies that sample individual organisms?}} If the study is making a total count of all organisms (let's say a mammal species) within a fixed plot size, or even a variable plot size from which an average plot size (and thus sampling resolution) can be estimated, then follow the procedures described in Q16-17.
-However, if the individual animals are the unit of measure (either because a sub-sample of them is being made within a defined plot, or because the observation is not contingent on being located within a plot (maybe a blood sample or body weight is being recorded, for example), then simply estimate two-dimensional area occupied by the animal, the number of animals sampled, and then calculate area from that.
+  \Hitem \emph{\textbf{What about studies that sample individual organisms?}} If the study is making a total count of all organisms (let's say a mammal species) within a fixed plot size, or even a variable plot size from which an average plot size (and thus sampling resolution) can be estimated, then follow the procedures described in Q17.
+However, if the individual animals are the unit of measure (either because a sub-sample of them is being made within a defined plot, or because the observation is not contingent on being located within a plot (maybe a blood sample or body weight is being recorded, for example), then simply estimate two-dimensional area occupied by the animal as the plot resolution, and the number of sampled animals provide the spatial replicates (for calculating actual extent).
 Occasionally individual animals might be recorded, but within the context of some natural feature, such as a nesting site where the survival of individual chicks is the measurement of interest \cite{roche_relative_2008}.
 In this case, an estimate of the nest area provides the sampling resolution.
-Another possible case is where individual animals are being tracked using radio or GPS collars.
-In these cases, use the estimate of the collared/tagged animal's body area, and then treat the number of reported locational observations as the number of spatial replicates, and then use those two values to calculate sampled area.
-The total GPS points collected from all individuals can be used if we can safely assume that each individual sampled did not cover the same exact location.
+In cases where individual animals are tracked using radio or GPS collars, to calculate actual extent, use the number of locational fixes as the quantity of spatial replicates and the animal's two-dimensional body area.
 If the number of GPS points is not given, the number of fixes can be estimated from the duration during which individuals were collared and the recording interval.
@@
 \subsubsection*{Between-observer coefficient of variation}
+\vspace{-10pt}
 We used the maximum uncertainty values for each dimension from the inter-observer variability analysis (Table S2) to determine the bounds of the random perturbations applied to each record over the 1000 iteration resample.
@@
 %
+
+\subsubsection*{Trends in observational methods by year}
+\vspace{-10pt}
+
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\includegraphics[width=1\textwidth]{figures/stype_by_yr.png}
+\vspace{10 pt}
+\caption{Trends in percentage of observing methods by year of publication.
+The coefficient of a weighted (by number of studies in each year) linear regression fit to the annual percentages of observations made with remote sensing (left), field methods (center), and automated sensors (right) is presented at the top of each plot, as well as the regression coefficient of determination and p-value.
+}
+\label{afoto1}
+\end{figure}
+
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\center{\includegraphics[width=0.8\textwidth]{figures/dim_by_yr.png}}
+\vspace{5 pt}
+\caption{Trends in observational scales by year of publication.
+The coefficient of a weighted (by number of studies in each year) linear regression, fit to the logarithm (base 10) of the mean scale values (calculated for each publication year) for the six assessed dimensions ispresented at the top of each plot, as well as the model coefficient of determination and p-value.
+}
+\label{afoto1}
+\end{figure}
+%\clearpage
+
+\subsubsection*{Choice of bandwidth in kernel density estimation}
+\vspace{-10pt}
 Figure S1 indicates the effect that varying bandwidth has on the appearance of the kernel density estimates.

### 2017-05-24 paper/main.tex
Updated analysis of annual change in observing methods and scales. Revision of main text to accomodate this change, as well as addition of related figures to SI.  Fixes on explanation of methods/FAQ in SI.

@@
 
-\noindent \textbf{\emph{Uncertainties in quantifying observational scales}}\\
-This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI).
+\noindent \textbf{\emph{Observational methods}}\\
+We classified the method used to collect each observation into several broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing, other geographic data, and paleo-reconstruction approaches.
+Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.3\%, and paleo-reconstruction and other geographic data each for less than 1\%.
+Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance.
+Automated sensing methods showed no trend (SI).
+
+\noindent \textbf{\emph{Potential biases and uncertainties in quantifying observational scales}}\\
+%This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI).
 Observations that were non-continuous in time or space (e.g.
@@
 
-Adding to our motivation to record actual observational scales was the fact that most studies provided insufficient methodological detail to judge whether sampling schemes were suitable matched to and representative of the phenomena they were measuring.
+%Adding to our motivation to record actual observational scales was the fact that 
+There were several potential methodological aspects that could have influenced our assessment of ecology's spatial and temporal domains.
+The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate, rather than simply record, these values for most observations (specifically, in 63, 60, and 67\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration).
+The inevitable errors caused by estimation may have biased our overall findings.
+However, we attempted to quantify this error by assessing inter-rater disagreement and resampling techniques.
+The resulting confidence intervals (Fig.
+1) suggest that it was unlikely that estimation errors unduly influenced our findings.
+
+Another potential source of bias was the rule set we used to estimate observational scales.
+
+Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates.
+Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find that remote sensing was used for 7.7\% of observations (a 22\% increase), which would increase the mean effective extent by 17.4\% (95\% CI = -1.3-67\%), or 0.07 orders of magnitude, above that observed here (see SI for details of calculation).
+Further support for such trend can be found by averaging dimensions by publication year (SI), which reveals that observation extent increased by 0.253 orders of magnitude per year between 2004-2014 (R$^2$ = 0.2.5, p$<$0.07).
+This somewhat clearer trend also suggests that including more recent studies would show observational extent to be somewhat larger, although in this case by a more modest 5.5\% (0.02 orders of magnitude).
+
+%Additional sources of uncertainty: rules for extracting scales; diversity of study types and ecological methods.
+
+%Most studies provided insufficient methodological detail to judge whether sampling schemes were suitable matched to and representative of the phenomena they were measuring.
 The observational scales were not clearly reported in many studies, such that we had to estimate spatial resolution and extent in 62\% and 68\% of cases, respectively, and temporal interval and duration in 40\% and 86\% of cases.
@@
 
-Despite these caveats, our results provide valuable insight into the spatial and temporal domains being addressed by modern ecological research.
+the continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
+
+satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.
+
+Adapt and tone down, retain these points:
+Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
+(Scale is not a concern that cuts across the entire discipline of ecology).
+Narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) 
+Ecological understanding drawn from many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}, 
+a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}.
+
+Our results provide valuable insight into the spatial and temporal domains being addressed by modern ecological research.
 Our results show that most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month).

### 2017-05-25 Rmd/space-time.Rmd
Finalized (or nearly) SI, and main text revised up to final section, and final part of abstract.

@@
 # dev.off()
+
 ```

### 2017-05-25 paper/SI.tex
Finalized (or nearly) SI, and main text revised up to final section, and final part of abstract.

@@
 \usepackage[skip=2pt,font=normalsize]{caption}
+\usepackage[LGRgreek]{mathastext}
+\usepackage{multirow}
 
@@
 {\end{quote}}
+
+\usepackage{array}
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 \vspace{-10pt}
-Most studies did not explicitly report values for all four of the assessed scales, and thus interpretation and judgement had to be applied to develop reasonable estimates for their values.
+Most studies did not explicitly report values for all the assessed scales, and thus interpretation and judgement had to be applied to develop reasonable estimates for their values.
 The FAQ provided the protocol we followed, and was initially developed following consultation between observers prior to the commencement of review.
@@
 We estimated and accounted for these uncertainties in several ways.
-First, we estimated the degree of between-observer variability based on each observer's results from reviewing the 20-paper calibration set.
+First, we calculated the degree of between-observer variability based on each observer's results from reviewing the 20-paper calibration set.
 We calculated how well observers agreed regarding paper inclusion/exclusion, how many extractable observations there were per included paper, and what the coefficient of variation was across all observers' estimates of scale.
-We also recorded when observations were reported in a study with an unclear or missing scale value.
+We also recorded when observations were reported in any study (calibration set or otherwise) with an unclear or missing scale value.
 
 We used the between-observer coefficients of variation for each dimension as the basis for randomly perturbing the scale values for each of the 371 observations over 1000 iterations.
-For each observation at each iteration, we perturbed its scale value in each dimension by randomly selecting a percentage value p that fell between 100 + y and 100 - y, where y was the between-observer coefficient of variation (expressed as a percentage) for a given observational dimension, and multiplying that observation's scale value by that proportion.
+For each observation at each iteration, we perturbed its scale value in each dimension by randomly selecting (from a uniform distribution) a percentage value p that fell between 100 + y and 100 - y, where y was the between-observer coefficient of variation (expressed as a percentage) for a given observational dimension, and multiplying that observation's scale value by that proportion.
 This perturbed set of observations provided a basis for estimating uncertainty within our extracted set of observational scales.
@@
 1 and 2 in main text), we log-transformed (base 10) the observational scales within each of the 1000 perturbed ensemble members in order to account for the large range in scale values.
-We calculated histograms for each observational dimension from each of the log-transformed ensemble members, calculating the mean percentage density estimate for each bin across all 1000 histograms, as well as the upper and lower 2.5th percentile values for each bin (Fig.
+We calculated histograms for the four primary observational dimensions from each of the log-transformed ensemble members, calculating the mean percentage density estimate for each bin across all 1000 histograms, as well as the upper and lower 2.5th percentile values for each bin (Fig.
 1 main text).
@@
 To reveal the densities of observations within two dimensions (Fig.
-2 main text), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values across all ensemble members, using a bandwidth of 1 onto a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations are concentrated.
+2 main text), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values across all ensemble members, using a bandwidth of 1 on a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations are concentrated.
 Bandwidths of varying resolutions were tested on kernel density estimates were tested on kernel density estimates of sampling interval versus plot resolution (see Supplementary Results section).
+
+To compare the differences between actual extent and extent and actual duration and duration (Fig. 3 main text) we calculated the magnitude of difference (decade) between each pair as:
+
+\begin{equation}
+decade = log_{10}(x) - log_{10}(y) = log_{10}\left(\frac{x}{y}\right)
+\end{equation}
+
+\noindent Where x is either extent or duration and y is its actual counterpart.
+We then evaluated how the magnitudes of difference varied with increasing values of actual extent/duration, using box plots to summarize decades within the same bins used for extent and duration (Fig.
+1B and D main text).
+Decades were calculated for each pair for all bootstrap replicates.
+The whiskers in the summary box plots (Fig.
+3 main text) fall slightly below 0 in some bins, even though extent and duration can never be smaller than their actual value.
+This is an artifact of the perturbation applied to estimated values, which in some instances caused extent/duration values to fall below their actual value.
+We retained these discrepancies rather than forcing such cases to zero to avoid biasing the summary statistics.
+
+\subsubsection*{Trends in methods and scale}
+To assess the degree to which observational methods changed over the course of the 10 years worth of studies we surveyed, we calculated the percentage of observations made using remote sensing, general field methods, and automated \emph{in situ} methods, and fit a linear regression between these percentages and publication year, weighting the regression by the total number of observations in each year.
+We performed the same analysis for the four primary dimensions (resolution, extent, interval, and duration) in order to assess whether there were any trends in observational scales.
 
@@
 Second, we calculated the intra-class correlation coefficient \cite{bartko_intraclass_1966} to assess agreement between observers regarding the number of ecological observations that could be extracted from each paper (multiple ecological observations were reported in many studies, and we listed observations as separate records if they varied on one or more dimensions).
-The coefficient, calculated using the IRR package \cite{gamer_irr:_2012} of R, was 0.70 (F = 15.4, p$<$0.0001, 95\% confidence interval = 0.54 - 0.85)
+The coefficient, calculated using the IRR package \cite{gamer_irr:_2012} of R, was 0.71 (F = 15.4, p$<$0.0001, 95\% confidence interval = 0.54 - 0.85)
 
-Finally, we calculated the coefficient of variation (CV) between observers' estimates of scales for each dimension, first across all six observers' mean scale estimates, and then as the average CV among observers' estimates of each individual record (Table S2).
+Finally, we calculated the coefficient of variation (CV) between observers' estimates of scales for each dimension, first across all observers' mean scale estimates, and then as the average CV among observers' estimates of each individual record (Table S2).
+CV values were estimated from all six observers for resolution, actual extent, and interval, from five observers (all but Choi) for duration, and from three observers (Elsen, Estes, and Treuer) for extent.
+The smaller numbers for duration and extent reflect that initial efforts were focused on estimating resolution, interval, and actual extent and actual duration, thus fewer observers were available to assess extent and extent.
+
+\subsubsection*{Between-observer coefficient of variation}
+\vspace{-10pt}
+We used the maximum uncertainty values for each dimension from the inter-observer variability analysis (Table S2) to determine the bounds of the random perturbations applied to each record over the 1000 iteration resample.
 
 \begin{table}[ht]
-\captionsetup{width=0.725\linewidth}
+\captionsetup{width=1\linewidth}
 \caption{Estimates of variability calculated from each observer's estimates of the spatial and temporal scales of ecological observations reported within a common set of the 20 papers used for calibration.
@@
 \begin{center}
-\begin{tabular}{rrrrr}
+\begin{tabular}{rR{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}R{1.5cm}}
 \hline
-& \multicolumn{2}{c}{\textbf{Spatial}} & \multicolumn{2}{c}{\textbf{Temporal}}\\\cline{2-5}
-Value & Resolution & Extent & Interval & Duration \\ 
+& \multicolumn{3}{c}{\textbf{Spatial}} & \multicolumn{3}{c}{\textbf{Temporal}}\\\cline{2-7}
+%Value & Resolution & Extent & \multirow{2}{*}{Actual} & Interval & Duration & Actual Duration \\
+Value & Resolution & Extent & Actual Extent & Interval & Duration & Actual Duration \\
 \hline
-CV of overall mean & 54 & 82 & 54  & 110 \\
-Mean of record-wise CV & 58 & 68 & 99 & 124 \\
+CV of overall mean & 50 & 50 & 75 & 48 & 34 & 109 \\
+Mean of record-wise CV & 57 & 41 & 72 & 104 & 64 & 124 \\
 \hline
@@
 \end{table}%
-
-\subsubsection*{Between-observer coefficient of variation}
-\vspace{-10pt}
-We used the maximum uncertainty values for each dimension from the inter-observer variability analysis (Table S2) to determine the bounds of the random perturbations applied to each record over the 1000 iteration resample.
 
@@
 
-\subsubsection*{Trends in observational methods by year}
+\subsubsection*{Trends in observational methods and scales by year}
 \vspace{-10pt}
+Figures S1 and S2 respectively show how the frequency of ecological observing methods and the average scale in each dimensions varies by publication year (2004-2014).
 
@@
 
-\begin{figure}[!ht]
+\begin{figure}[ht]
 %\begin{wrapfigure}{c}{1\textwidth}
@@
 \vspace{-10pt}
-Figure S1 indicates the effect that varying bandwidth has on the appearance of the kernel density estimates.
+Figure S3 indicates the effect that varying bandwidth has on the appearance of the kernel density estimates.
 The smallest bandwidth (0.4) tested (Fig.
-S2 top) shows that the primary observational concentrations described in the main text are evident but harder to discern.
-Most difficult to discern is the oblong concentration of observations that in Fig.
-1A (main text) is bounded on the lower right at monthly to yearly intervals and 100-1000 m$^2$ resolutions and on the upper left by near-daily to monthly observations and 0.1-10 ha resolutions.
+S3 top) shows that the primary observational concentrations described in the main text are evident, but tend to be devolved into separate concentrations.
+An example is the oblong concentration of observations that in Fig.
+1A (and in the lower left panel in Fig.
+S3) is bounded on the lower right at monthly to yearly intervals and 100-1000 m$^2$ resolutions and on the upper left by near-daily to monthly observations and 0.1-10 ha resolutions.
 With the smaller bandwidth this concentration appears as two separate patches (Fig.
@@
 S3 middle left).
+This increasing separation of points into separate clusters is expected as bandwidth is reduced.
 
@@
 \vspace{10 pt}
-\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by sampling interval and spatial resolution (left column) and temporal duration and spatial extent (right column), applied to log-transformed values of each observational dimension.
+\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by sampling interval and spatial resolution (left column) and duration and extent (right column), applied to log-transformed values of each observational dimension.
 Rows indicate the effects of selecting different bandwidths: 0.4 (top row); 0.7 (middle row); 1 (bottom row).

### 2017-05-25 paper/main.tex
Finalized (or nearly) SI, and main text revised up to final section, and final part of abstract.

@@
 The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate, rather than simply record, these values for most observations (specifically, in 63, 60, and 67\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration).
-The inevitable errors caused by estimation may have biased our overall findings.
+The inevitable estimation errors may have biased our overall findings.
 However, we attempted to quantify this error by assessing inter-rater disagreement and resampling techniques.
@@
 
-Another potential source of bias was the rule set we used to estimate observational scales.
+Another potential source of bias was the rule set we used to estimate observational scales, chiefly through our definition of resolution as the smallest area in which all features of interest were measured.
+We made this choice because we wanted to use a consistent standard.
+However, some papers reported plot resolutions wherein sub-samples were made within the plot area, thus the average resolution (and actual extent) calculated from our estimates is likely to be smaller than it would be if we had simply recorded reported resolution in all such cases.
+In addition to this, our decision to exclude experimentally manipulated observations is also likely to have influenced the distributions of all assessed dimensions.
+For example, the average values of estimated resolutions would likely be smaller if we had included experiments.
 
 Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates.
-Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find that remote sensing was used for 7.7\% of observations (a 22\% increase), which would increase the mean effective extent by 17.4\% (95\% CI = -1.3-67\%), or 0.07 orders of magnitude, above that observed here (see SI for details of calculation).
-Further support for such trend can be found by averaging dimensions by publication year (SI), which reveals that observation extent increased by 0.253 orders of magnitude per year between 2004-2014 (R$^2$ = 0.2.5, p$<$0.07).
-This somewhat clearer trend also suggests that including more recent studies would show observational extent to be somewhat larger, although in this case by a more modest 5.5\% (0.02 orders of magnitude).
+Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find that remote sensing was used for 7.7\% of observations (a 22\% increase), which would increase the mean extent by 17.4\% (95\% CI = -1.3-67\%), or 0.07 orders of magnitude, above that observed here (see SI for details of calculation).
+Further support for such a trend can be found by averaging dimensions by publication year (SI), which reveals that observational extent increased by 0.253 orders of magnitude per year between 2004-2014 (R$^2$ = 0.2.5, p$<$0.07).
+This somewhat clearer trend also suggests that including more recent studies would show extent to be somewhat larger, although in this case by a more modest 5.5\% (0.02 orders of magnitude).
 
@@
 \noindent \textbf{\emph{Insights into the scales of modern ecology}}\\
+The continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
+
 %These results indicate that for most ecological observations are collected with spatially discontinuous replicates that sample very small portions of the area in which they are distributed.
 
-the continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
-
-satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.
+Satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.
 

### 2017-06-12 Rmd/space-time.Rmd
Everything through final discussion section revised

@@
 datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
-datf[act_dur > eff_dur]
+
+# drop one calibration study only done by one observer
+datf <- datf[DOI != "10.1016/j.agee.2013.11.021", ]
+
+# DOIs of additional studies pointed to as data sources
 doisup <- c(unique(cal$DOI_data_source), unique(dat$DOI_data_source))
@@
 doisup <- unlist(strsplit(doisup, ";"))
+
 ```
@@
 
+# randomly reorders two variables, for reassignment in case of physically 
+# unrealistic values
+rvord <- function(a, b) {
+  v <- c(a, b)
+  v <- v[sample(1:2)]
+  v
+}
+
+# function for parse-eval in data.table i
+psr <- function(a) eval(parse(text = a))
+
 # bootstrapping function
-resamp_func <- function(dat, vars, p, iter) {
+resamp_func <- function(dat, p, iter) {
+  vars <- c("plot_res", "act_ext", "eff_ext", "t_btwn_samp", "act_dur", "eff_dur")
   bout <- lapply(1:iter, function(x) {  # x <- 1
@@
       mv <- runif(nrow(dnew), min = 1 - y, max = 1 + y)
+      mv[mv < 0] <- 0.00000001  # can't have reduction of more than 100%
       mv
     })
-    # plot(dnew[[j]] * modvec[, j], dnew[[j]])
 
@@
     for(j in 1:ncol(modvec)) set(dnew, j = j, value = dnew[[j]] * modvec[, j])
-    dnew[plot_res <= 0, plot_res := 0.00000001]
-    dnew[act_ext <= 0, act_ext := 0.00000001]
-    dnew[eff_ext <= 0, eff_ext := 0.00000001]
-    dnew[act_dur <= 0, act_dur := 0.00000001]
-    dnew[eff_dur <= 0, eff_dur := 0.00000001]
-    dnew[t_btwn_samp < 0, t_btwn_samp := 0.00000001]
-
+
+    # correct physically impossible results 
+    dnew[, plot_res := plot_res / 10000]  # set plot_res to ha
+    dnew[plot_res > act_ext, plot_res := act_ext]  # plot_res can't be > act_ext
+    dnew[plot_res > eff_ext, plot_res := eff_ext]  # or eff_ext
+    dnew[act_ext > eff_ext, act_ext := eff_ext] # act_ext can't be > eff_ext
+    dnew[t_btwn_samp > eff_dur, t_btwn_samp := eff_dur]  # int can't be > eff_dur
+    dnew[act_dur > eff_dur, act_dur := eff_dur]  # act_dur can't be > eff_dur
+    dnew[, plot_res := plot_res * 10000]  # set plot_res back to m2
+
+    # set unreplicated to high value for first figure 
     dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
@@
 # Run bootstrap
-vars <- c("plot_res", "act_ext", "eff_ext", "t_btwn_samp", "act_dur", "eff_dur")
 set.seed(1)
-bootperturb <- resamp_func(datf, vars = vars, p = p, iter = 1000) # 10% perturb
+bootperturb <- resamp_func(datf, p = p, iter = 1000) 
 # bootperturb2 <- resamp_func(datf, p = p, iter = 1000)  # 10% perturb
@@
 
-tlab2 <- c(expression(NULL<="second"), "minute", "hour", "day", "week", "month",
-           "year","decade", "century", "millenium", "10 KA")  
+tlab2 <- c(expression(NULL<="second"), "minute", "hour", "day", "week", 
+           "month","year","decade", "century", "millenium", "10 KA")  
 taxis2 <- cbind.data.frame(c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 
@@
            expression("1 cm"^2), expression("10 cm"^2),
-           expression("100 cm"^2), expression("1000 cm"^2), 
+           expression("0.01 m"^2), expression("0.1 m"^2), 
            expression("1 m"^2), expression("10 m"^2),
@@
 qtf <- function(x) c(mean(x), quantile(x, probs = c(0.025, 0.975))) # quant f
-xlabs <- c("Spatial resolution", "Spatial extent", "Sampling interval", 
-           "Temporal duration")
+xlabs <- c("Resolution", "Extent", "Interval", "Duration")
 # axes <- list("aax1" = aaxis1$logres, "aax2" = c(-6, aaxis2$logarea), 
@@
 # res
-# sum(hdatst[4:6])  # 53 % between 10cm^2 and 1 m^2
-# sum(hdatst[1:6])  #  67% < 1m^2
-# sum(hdatst[7:10]) # 24 % between 1 m^2 and 1 ha
-# sum(hdatst[11:(l)]) # 9 % 1-10000 ha
-# dat[plot_res >= 10000^2]
 hdatst <- hdatstat_bt$res$mu
@@
 l <- length(hdatst)
-# sum(hdatst[1:2])  # 36 % < 1 m^2
-# sum(hdatst[1:6])  # 82% < 1 ha
-# sum(hdatst[1:7])  # 86% < 10 ha
-# sum(hdatst[1:8])  # 90 % < 100 ha
-# sum(hdatst[1:9])  # 92 % < 1000 ha
-# sum(hdatst[8:9])  # 6 % 10-1000 ha
-# sum(hdatst[8:l])  # 14% > 10 ha
-# sum(hdatst[10:l]) # 8.2 % covered an area > 1000 ha
-# sum(hdatst[12:l]) # 5.1 % covered an area > 100,000 ha
-# sum(hdatst[13:l]) # 3.9 % covered an area > 1,000,000 ha
 hds <- c("<1 ha", "<10 ha", "<100 ha", "<1000ha", "10-1K ha", ">10ha", "1K-1MK",
@@
 l <- length(hdatst)
-# sum(hdatst[l])  # 36% are once-offs
-# sum(hdatst[1:3])  # 15 % < daily
-# sum(hdatst[1:4])  # 25 % < weekly
-# sum(hdatst[4:5])  # 21% daily up to monthly
-# sum(hdatst[6])  # 19% monthly up to yearly
-# sum(hdatst[7:8]) # 8.2 % yearly to decadal
 hds <- c("norep", "<daily", "<weekly",  "day-mo", "mo-yr", "yr-dec")
@@
 l <- length(hdatst)
-# sum(hdatst[1:3])  # 71 % up to 1 day
-# sum(hdatst[4:5])  # 17% covered between 1 day and 1 month
-# sum(hdatst[6])  # 7.3% covered between 1 month and 1 year
-# # sum(hdatstat[[4]]$density[4:6])  # 26% covered between 1 day and 1 year
-# sum(hdatst[7:10]) # 4.5% year to century
 hds <- c("<=daily", "day-mo", "mo-yr", "yr-dec", "yr-cent")
@@
 # range(tt$x - tempdat$x)
+# tempdat[x != log10(365 * 10000)]
 tempdat <- tempdat[!is.na(x) & !is.na(y)] # cut out null values
 tempdat <- tempdat[round(x, 8) != 6.56229286, ]  # remove one-time observations
+
+tempdat[x > y, .N] / tempdat[, .N]
+
 # tempdat$n <- rep(1, nrow(tempdat))
@@
 tx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-#tx[2] <- 6.56
-tx[2] <- 5.57
+tx[2] <- 6.6  # just above 10 KA
+# tx[2] <- 5.56
 tempdat[x < rx[1], x := rx[1]]
@@
 ty <- range(taxis2$logdays)
-ty <- c(floor(ty[1]), 5.56)#ceiling(ex[2]))
+# ty <- c(floor(ty[1]), 5.56)#ceiling(ex[2]))
+ty <- c(floor(ty[1]), 6.6)  # just above 10 KA
 tempdat[y < ty[1], y := ty[1]]
@@
 image(tres, axes = FALSE, xlab = "", ylab = "", col = cols)
-axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
-     las = 2, tcl = -0.2, cex.axis = cxa)
-axis(2, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
+for(i in 1:2) {  # using taxis2 values for both axes because same
+  axis(i, at = taxis2$logdays[-5], labels = tlab2[-5], las = 2,
        cex.axis = cxa, tcl = -0.2)
+}
+# axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
+#      las = 2, tcl = -0.2, cex.axis = cxa)
 mtext("Interval", 1, cex = cxa, line = ll)
@@
 # points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)
+
+# bootp_dt[!t_btwn_samp == 365 * 10000, {
+#   mean(t_btwn_samp[eff_dur > 30 & eff_dur < 365])
+# }]
+# bootp_dt[!t_btwn_samp == 365 * 10000, {
+#   mean(t_btwn_samp[eff_dur > 365 & eff_dur <= (365 * 10)])
+# }]
+
 ```
@@
 text(mean(xs), 13, "% of observations", col = "grey30")
-mtext("A", side = 3, line = -1, cex = 1, adj = 0.05)
+mtext("A", side = 3, line = -1, cex = 1, adj = 0.02)
 
@@
 text(mean(xs), 13, "% of observations", col = "grey30")
-mtext("B", side = 3, line = -1, cex = 1, adj = 0.05)
+mtext("B", side = 3, line = -1, cex = 1, adj = 0.02)
 dev.off()
@@
 ### Varying kernel sizes
-```{r,eval = FALSE}
+```{r, eval = FALSE}
 # redo range variables
@@
 
+datf[t_btwn_samp != 0, plot(log10(t_btwn_samp), log10(eff_ext))]
+datf[, plot(log10(plot_res), log10(eff_dur))]
+datf[, plot(log10(eff_dur), log10(eff_ext))]
+datf[t_btwn_samp == 0, points(log10(eff_dur), log10(eff_ext), col = "red")]
+datf[t_btwn_samp == 0 & log10(eff_dur) > 4]
+
 ```

### 2017-06-12 paper/main.tex
Everything through final discussion section revised

--- 2017-05-24 paper/main.tex
@@
 \usepackage{color}
+\usepackage{multirow}
 
@@
 \textbf{In order to properly understand ecological phenomena, it is necessary to quantify their behavior over the range of spatial and temporal scales at which they manifest.
-Ecology has been concerned with this need since the early 1990s, and the ability to collect multi-scaled ecological observations has grown rapidly since then.
-Characterizing the spatial and temporal domains of modern ecological observations can therefore provide important insight into the field's progress in understanding towards a more comprehensive understanding of ecosystem behaviour.
+Ecology has been concerned with this need for decades, and the ability to collect ecological observations across multiple scales has grown rapidly since then.
+Characterizing the spatial and temporal domains (the distribution of observations within one or more scale spectra) of modern ecological observations can therefore provide important insight into the field's progress towards a more comprehensive understanding of ecosystem behaviour.
 To characterize these domains, we conducted a meta-analysis of recent (2004-2014) ecological studies, in which we quantified four primary dimensions of their reported observations: plot resolution, sampling interval, effective duration (time between start and end of temporal replicates), and effective extent (area enclosed by spatial replicates).
@@
 The scales at which ecosystems are observed plays a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
-Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is the \emph{sine qua non} for understanding ecological phenomena \cite{levin_problem_1992}.
-This awareness has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad range of scales \cite{schneider_rise_2001, chave_problem_2013}.
+Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992}.
+This awareness has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad and increasing range of scales \cite{schneider_rise_2001, chave_problem_2013}.
 
-Given this awareness of the centrality of scale to understanding ecology, and the growing ability to study ecological phenomena across a broad range of spatial and temporal scales, it is important to assess the scales at which current ecological research is conducted.
-To gain insight into this question, we quantified the spatial and temporal domains of empirical observations that were reported in a representative sample of studies published between 2004-2014 in the top 30 ecological journals (by impact factor).
-Empirical observations provide the necessary means for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, thus it stands to reason that the temporal and spatial distributions of ecological observations may be indicative of modern ecology's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.
+%One of the best known and most comprehensive assessments of scale in ecological research ident
 
-We characterized observational domains along two key spatial dimensions, resolution (grain) and extent, and their temporal corollaries, interval and duration (Table 1).
-Here resolution is the area of an individual spatial replicate, or the two-dimensional space in which all measurable features of a natural object(s) were recorded (as opposed to sub-sampled), while extent is the area enclosed by the outer-most spatial replicates, or, if the system or habitat being sampled was distinct from its surrounding matrix (e.g.
+Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) analysis of ecological research scales was published in 1989 \cite{tilman_ecological_1989}, it is both timely and important to assess the scales of contemporary ecological investigation.
+To address this need, we quantified the spatial and temporal domains of empirical observations that were reported within recently (2004-2014) published ecological studies (here domain means the distribution of observations within the spectrum of one or more scale dimensions\footnote{This definition differs slightly from Wiens' \cite{wiens_spatial_1989}, who defined ``domain of scale'' as ``a portion of the scale spectrum within which process-pattern relationships are consistent regardless of scale.''}).
+Empirical observations are critical for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, therefore the spatio-temporal domains of observations provide an important indicator of the field's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.
+
+%This question may be best addressed by evaluating the scales of empirical observations.
+Empirical observations are critical for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, therefore the spatio-temporal domains of observations provide an important indicator of the field's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.
+
+Our analysis focused on two dimensions of spatial scale, resolution (grain) and extent, and two of temporal scale, interval and duration (Table 1).
+Resolution is the area of an individual spatial replicate within which a complete measurement (as opposed to a sub-sample) of the feature of interest was made.
+Extent is the area enclosed by the outer-most spatial replicates, or, if the system or habitat being sampled was distinct from its surrounding matrix (e.g.
 forest patches in grassland habitats), the summed area of sampled patches (see SI for full definition).
-Interval refers to the average time elapsed between individual temporal replicates, and duration is the time elapsed between the first and last temporal replicates, or, in the case of temporally unreplicated observations, the estimated time spent collecting the observation (SI).
-We also calculated two additional metrics, the integrated area of spatial replicates (i.e.
-resolution multiplied by number of replicates) and the summed observational time of all temporal replicates.
-We estimated these additional dimensions to evaluate the degree to which the actual scales of ecological observations differ from those they ostensibly represent, and therefore refer to them as the \emph{actual} extent and \emph{actual} duration.
+Interval refers to the average time elapsed between individual temporal replicates.
+Duration measures the time elapsed between the first and last temporal replicates, or, in the case of temporally unreplicated observations, the estimated time spent collecting the observation (SI).
+We also assessed observational scales within two additional dimensions, \emph{actual} extent (the integrated area of spatial replicates) and \emph{actual} duration (the summed observational time of temporal replicates).
+We evaluated these additional dimensions to gain insight into how much the actual scales of observation (i.e.
+how much space and time is covered by the measurement) differ from those they are intended to represent.
+
+%For example, consider a hypothetical study reporting estimates of herbaceous plant cover in a forest understory, in which the authors reported that they placed 10 400 m$^2$ plots throughout the forest, and within each plot randomly placed 10 1 m$^2$ quadrats, within which they estimated percent herbaceous plant cover.
+In this example, we consider 1 m$^2$ to be the resolution.
 
 \begin{table}[htp]
-\caption{The dimensions of ecological observations estimated in this meta-analysis.}
+\caption{The scale dimensions of ecological observations assessed in this meta-analysis.}
 \begin{center}
-\begin{tabular}{ll}
+\begin{tabular}{llll}
 \hline
-\textbf{Dimension} & \textbf{Description} \\
-\hline
-Resolution & Area (m$^2$) of an individual spatial replicate (plot) \\
-Extent & Area (ha) encompassed by all spatial replicates \\
-Actual extent & Summed area (ha) of all spatial replicates\\
-Interval & Time elapsed (days) between successive temporal replicates \\
-Duration & Time elapsed (days) between first and last temporal replicates\\
-Actual duration & Summed observational time (days) of all temporal replicates\\
+\multicolumn{2}{c}{\textbf{Component}} & \textbf{Units} & \textbf{Description} \\\hline\hline
+\multirow{3}*{Spatial} & Resolution & m$^2$ & Area of an individual spatial replicate (plot) \\%\cmidrule(l){2-3}
+   & Extent & ha & Area  encompassed by all spatial replicates \\
+   & Actual extent & ha & Summed area of all spatial replicates\\\hline
+\multirow{3}*{Temporal} & Interval & days & Time elapsed between successive temporal replicates \\
+   & Duration & days & Time elapsed between first and last temporal replicates\\
+   & Actual duration & days & Summed observational time of all temporal replicates\\
 \hline
@@
 
-We calculated these dimensions from 379 discrete observations reported within a 134 paper subset of 348 randomly selected articles (from 42,918 total).
+%We extracted these scale dimensions from 379 observations that were reported in 134 papers that were published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals. 
+%
+
+Our analysis was based on a review of 348 papers randomly selected from 42,918 published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals.
+We extracted scale data from 379 observations of ``natural'' (i.e.
+non-experimentally manipulated) ecological features that were reported within 134 of the reviewed papers (plus an additional 62 that these cited as the source of observations).
+We excluded experiments because they tend to be of limited extent, duration, and resolution due to their higher logistical costs \cite{tilman_ecological_1989, kareiva_spatial_1988}, and would therefore likely bias our findings towards finer scales.
+
+%In contrast, cost is less likely to dictate the choice of observational scale in studies of natural systems (considering, for example, that many global-scale, multi-decadal, satellite-derived datasets are free \cite[e.g.][]{tucker_nasas_2004,friedl_modis_2010}), therefore an assessment confined to such studies may provide a less biased picture of ecology's scale domains.
+
+%We calculated these from 379 discrete observations reported within a 134 paper subset of 348
+%randomly selected articles (from 42,918 total). An additional 62 papers that were cited as the
+%source of observations in the selected papers were also reviewed.
+We confined our analysis to observations of “natural” (7), or non-experimentally manipulated, phenomena.
+We made this choice because experiments are costly and therefore tend to be fine-scaled in both space and time (7, 9), which could unduly bias our analysis.
+%
+%We extracted these scale dimensions from 379 observations reported within 134 papers selected from the top 30 (based on impact factor) ecology-themed journals.
+The papers were a subset of 348 that we reviewed in total, which we selected randomly from the 42,918 studies published between 2004-2014 .
 An additional 62 papers that were cited as the source of observations in the selected papers were also reviewed.
-We confined our analysis to observations made of ``natural'' \cite{tilman_ecological_1989}, or non-experimentally manipulated, systems, given that the inclusion of experiments could have skewed our assessments towards the relatively fine scales at which such studies typically focus.
+We confined our analysis to observations of ``natural'' \cite{tilman_ecological_1989}, or non-experimentally manipulated, phenomena.
+We made this choice because experiments are costly and therefore tend to be fine-scaled in both space and time \cite{tilman_ecological_1989,kareiva_spatial_1988}, which could unduly bias our analysis.
 
@@
 1).
-We constructed kernel density estimates from the full resampled ensemble in order to assess observational distributions within different juxtapositions of the four space-time dimensions (Fig.
+We constructed kernel density estimates from the full resampled ensemble in order to assess observational distributions within different juxtapositions of the four primary (resolution, extent, interval, duration) space-time dimensions (Fig.
 2).
 
+\noindent \textbf{\emph{Domains}}\\
 In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
@@
 %\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/hists_bs.pdf}
+\includegraphics[width=1\textwidth]{figures/hists_bs.png}
 \vspace{-0.15 cm}
@@
 
-Juxtaposing these observational dimensions provides further insight into the spatio-temporal distribution of ecological observations and the domains in which they are concentrated (Fig.
+Juxtaposing these observational dimensions provides further insight into the spatio-temporal domains of ecological observations (Fig.
 2).
@@
 \vspace{-0.15 cm}
-\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by A) interval and resolution (of temporally replicated observations only), B) duration and extent, C) resolution and extent, and D) interval and duration (of temporally replicated observations).
+\caption{Kernel density estimates of observational densities within the domains defined by A) interval and resolution (of temporally replicated observations only), B) duration and extent, C) resolution and extent, and D) interval and duration (of temporally replicated observations).
 Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.}
@@
 2C).
-The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha.
+The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha.
 These three concentrations suggest a tendency for observational extent to increase with resolution, which is a relationship that becomes more pronounced in the (less densely observed) portion of the domain where resolutions $\geq$100 m$^2$.
@@
 A similar tendency for duration to increase with interval was also evident amongst temporally replicated observations (Fig.
-2D), where the majority of observations were repeated at daily to decadal intervals and spanned  $\geq$ 1 month to $<$100 years.
-The orientation of this concentration shows that observations lasting one year to one decade tend to have corresponding intervals, while those lasting one month to one year have daily to monthly intervals.
-This suggests that most long-duration observations have just a single temporal replicate.
+2D), with the majority of repeated at daily to decadal intervals and spanning $\geq$ 1 month to 1 decade.
+The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting one year to one decade tend to have yearly to decadal intervals.
 The low densities of observations having sub-daily intervals shows that relatively few high frequency, long duration ecological measurements are undertaken.
 
-To provide further insight into the spatio-temporal domains of ecological observations, we also evaluated the differences in scale between extent and actual extent and between duration and actual duration.
-To make these comparisons, we calculated the magnitudes of difference (decades) between each dimension and its actual equivalent, and examined how these varied as a function of the scale of the actual measurement (Fig.
+To provide further insight into observational domains, we also evaluated the degree to which the scales \emph{represented} by observations differed from those that were actually observed.
+To make this assessment, we first log$_10$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, in order to calculate the magnitude of difference (or decade) between each pair of dimensions for each observation.
+We then compared how these magnitudes varied in relation to the scale of the actual dimension (Fig.
 3).
-The majority (80\%) of the assessed observations had actual extents of $\leq$1 ha, which on average was 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).
+This comparison showed that a majority (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).
 Actual extent converged with extent above 100,000 ha, but this applies to just 5\% of observations.
-The actual duration of 65\% of observations is $<$ 1 day, which is on average 3 to nearly 9 orders of magnitude shorter than the time span covered by temporal replicates (Fig.
+The actual duration of 64\% of observations was $<$ 1 day, which on average was 3 to nearly 9 orders of magnitude shorter than the time span covered by temporal replicates (Fig.
 3B).
-The two duration measures were effectively the same for the 16\% of observations exceeding one month of actual duration.
+The two duration measures were approximately equal for the 16\% of observations exceeding one month of actual duration.
 
@@
 Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance.
-Automated sensing methods showed no trend (SI).
+Automated sensing methods showed no trend over time (SI).
 
-\noindent \textbf{\emph{Potential biases and uncertainties in quantifying observational scales}}\\
+\noindent \textbf{\emph{Potential biases and uncertainties in quantifying scales}}\\
+There were several potential methodological aspects that could have influenced our assessment of ecology's spatial and temporal domains.
+The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate--rather than simply record--these values for most observations (specifically, in 63, 60, and 67\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration, respectively).
+The inevitable estimation errors may have biased our overall findings.
+However, we attempted to quantify this error by assessing inter-rater disagreement and incorporating this uncertainty into our resampling methodology.
+The resulting confidence intervals (Fig.
+1) suggest that it was unlikely that estimation errors unduly influenced our findings.
+
+Another potential source of bias lies within our scale-estimation protocols, chiefly with respect to our rule for estimating resolution (the smallest areal unit of \emph{complete} measurement).
+We selected this definition for the sake of consistency, but some papers reported resolution as a larger area in which sub-samples were taken.
+For these, our estimates are finer than what the studies' authors apparently considered to be plot resolution.
+Additionally, our domain estimates would presumably be somewhat different if we had included experimentally manipulated observations.
+For example, average resolution and duration would likely be finer \cite{tilman_ecological_1989,kareiva_spatial_1988}.
+
+Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates.
+Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average (see SI for details of calculation).
+Further evidence for this trend lies within the extent values themselves, which increased 0.25 orders of magnitude per year between 2004-2014 (R$^2$ = 0.25, p$<$0.07).
+This somewhat clearer trend also suggests that including more recent studies would show extent to be somewhat larger, although in this case by a more modest 5.5\% (0.02 orders of magnitude).
+
+%Additional sources of uncertainty: rules for extracting scales; diversity of study types and ecological methods.
+
+%Most studies provided insufficient methodological detail to judge whether sampling schemes were suitable matched to and representative of the phenomena they were measuring.
+The observational scales were not clearly reported in many studies, such that we had to estimate spatial resolution and extent in 62\% and 68\% of cases, respectively, and temporal interval and duration in 40\% and 86\% of cases.
+Although the confidence intervals around our scale estimates (Fig.
+1) suggests that our findings are robust to this level of uncertainty, it would have likely confounded efforts to estimate effective scales.
+
 %This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI).
@@
 %Adding to our motivation to record actual observational scales was the fact that
-There were several potential methodological aspects that could have influenced our assessment of ecology's spatial and temporal domains.
-The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate, rather than simply record, these values for most observations (specifically, in 63, 60, and 67\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration).
-The inevitable errors caused by estimation may have biased our overall findings.
-However, we attempted to quantify this error by assessing inter-rater disagreement and resampling techniques.
-The resulting confidence intervals (Fig.
-1) suggest that it was unlikely that estimation errors unduly influenced our findings.
 
-Another potential source of bias was the rule set we used to estimate observational scales.
+\noindent \textbf{Broader implications}\\
 
-Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates.
-Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find that remote sensing was used for 7.7\% of observations (a 22\% increase), which would increase the mean effective extent by 17.4\% (95\% CI = -1.3-67\%), or 0.07 orders of magnitude, above that observed here (see SI for details of calculation).
-Further support for such trend can be found by averaging dimensions by publication year (SI), which reveals that observation extent increased by 0.253 orders of magnitude per year between 2004-2014 (R$^2$ = 0.2.5, p$<$0.07).
-This somewhat clearer trend also suggests that including more recent studies would show observational extent to be somewhat larger, although in this case by a more modest 5.5\% (0.02 orders of magnitude).
+Our results show that modern ecological observations most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month).
 
-%Additional sources of uncertainty: rules for extracting scales; diversity of study types and ecological methods.
+\begin{enumerate}
+\item General patterns of scales 
+  \subitem{-} General patterns
+  \subitem{-} The continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
+  \subitem{-} Notable holes: high frequency (sub-daily, daily), high resolution
 
-%Most studies provided insufficient methodological detail to judge whether sampling schemes were suitable matched to and representative of the phenomena they were measuring.
-The observational scales were not clearly reported in many studies, such that we had to estimate spatial resolution and extent in 62\% and 68\% of cases, respectively, and temporal interval and duration in 40\% and 86\% of cases.
-Although the confidence intervals around our scale estimates (Fig.
-1) suggests that our findings are robust to this level of uncertainty, it would have likely confounded efforts to estimate effective scales.
+   \subitem{-} Duration hasn't improved much since Tilman (check this).
+   \subitem{-} Evaluate how duration has improved since then.
+387/749 of the papers he looked at from the journal Ecology (1977-1984) were mainly experimental or had an experimental component.
+99 papers were mainly theoretical or conceptual.
+Remainder observational.
+We could compare our results to that here.
+Of his 623 field based studies (including 180 experiments), $\sim$41\% lasted less than one year,  $\sim$85\% less than three years, most studies $>$10 years and all $>$50 years were paleoecology/chronosequences.
+  \subitem{-} Maybe mention here (or probably better later in the discussion), that our decision has the unfortunate consequence of not being directly comparable to previous assessments that included experiments.
+More of just a caveat or note for the reader who is familiar with the earlier work and their summary statistics
+  \subitem But there do seem to be trends increasing extent (and duration, but need to put this above), perhaps due to increasing use of remote sensing 
+\item Scales are not well-documented 
+  \subitem Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
+(Scale is not a concern that cuts across the entire discipline of ecology).
 
-\noindent \textbf{\emph{Insights into the scales of modern ecology}}\\
+\item Represented scale is very different from actual scale of observations. 
+  \subitem In heeding Levin's call to think at multiple and appropriate ecological scales, we also need to give pause to think about whether our data are overextrapolated/overleveraged.
+  \subitem How over-leveraged/under-leveraged might depend on the auto-correlation properties of feature being observed. Maybe use example of fire process
+    \subitem Maybe we should go further here to explain why this is an important (and overlooked) assessment to make (actual versus effective comparison)
+\item Where to from here?  What do we need to improve?   
+  \subitem Satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.
+
+\end{enumerate}
+
+%In contrast, cost is less likely to dictate the choice of observational scale in studies of natural systems (considering, for example, that many global-scale, multi-decadal, satellite-derived datasets are free \cite[e.g.][]{tucker_nasas_2004,friedl_modis_2010}), therefore an assessment confined to such studies may provide a less biased picture of ecology's scale domains.
+
 %These results indicate that for most ecological observations are collected with spatially discontinuous replicates that sample very small portions of the area in which they are distributed.
 
-the continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
-
-satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.
-
-Adapt and tone down, retain these points:
-Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
-(Scale is not a concern that cuts across the entire discipline of ecology).
 Narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) 

### 2017-06-13 Rmd/space-time.Rmd
Figure revisions, initial paragraph of discussion section.

@@
 
-# dat[st == 5, .N] / nrow(dat)
-
 kwords <- cbind(c("field", "direct", "paleo", "remote", "automated", "other"),
@@
 }
-# cbind(tst, cal[, study_type])
-
-# unique(dat$DOI)
-# a <- unique(dato$DOI)
-# b <- unique(dat$DOI)
-# notin <- sort(a)[!sort(a) %in% sort(b)]
-# dato[DOI %in% notin, 1:5, with = FALSE]
-# dat[observer == "Ahmed" & journal == "Oecologia"]
-# unique(cal$DOI) %in% unique(dat$DOI)
-# match(cal$DOI, dat$DOI)
-# match(dat$DOI, cal$DOI)
-# dat[dat$DOI %in% cal$DOI, DOI]
-# dat[cal$DOI %in% dat$DOI, ]
-# cal[DOI == dat[cal$DOI %in% dat$DOI, DOI][11]]
-# dat[DOI == dat[cal$DOI %in% dat$DOI, DOI][7]]
-# dat[observer == "Treuer", unique(DOI)]  # 19
 
@@
 ```{r, message=FALSE, warning=FALSE, results='hide'}
-# dat2 <- copy(dat[, .(st, DOI, study_type, plot_res, n_sites, act_ext, eff_ext,
-#                      samp_dur, t_btwn_samp, act_dur, eff_dur)])
-# rnms <- c("st", "DOI", "study_type", "feature", "observer", "plot_res", 
-#           "n_sites", "act_ext", "eff_ext", "samp_dur", "t_btwn_samp",
-#           "act_dur", "eff_dur")
 dat2 <- copy(dat[, .(DOI, study_type, plot_res, n_sites, act_ext, eff_ext,
@@
 l <- length(hdatst)
-hds <- c("<1 ha", "<10 ha", "<100 ha", "<1000ha", "10-1K ha", ">10ha", "1K-1MK",
-         ">1Kha", ">100Kha", ">1Mha")
-inds <- list(1:6, 1:7, 1:8, 1:9, 8:9, 8:l, 10:12, 10:l, 12:l, 13:l)
+hds <- c("<1 ha", "<10 ha", "<100 ha", "<1000ha", "<10Kha", "10-1k ha", "1k-10k",  
+         "10k-100k", "100k-1Mk", ">1Kha", ">100Kha", ">1Mha")
+inds <- list(1:6, 1:7, 1:8, 1:9, 1:10, 8:9, 10, 11, 12, 10:l, 12:l, 13:l)
 sapply(1:length(inds), function(x) {
@@
 # resdat3$n <- rep(1, nrow(resdat3))
-rx <- range(resdat3$x)
-rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-rx[1] <- -5  # set min interval to less than second
+ilim <- range(resdat3$x)
+ilim <- round(ilim / 0.5) * 0.5  # round to nearest 0.5
+ilim[1] <- -5  # set min interval to less than second
 #rx[2] <- 6.56
-rx[2] <- 5.57  # set max interval to millenium
-resdat3[x < rx[1], x := rx[1]]
-resdat3[x > rx[2], x := rx[2]]
-ry <- range(aaxis1$logres)
-ry[1] <- -5  # set limits to 0.01 cm^2
-resdat3[y < ry[1], y := ry[1]]
-resdat3[y > ry[2], y := ry[2]]
+ilim[2] <- 5.57  # set max interval to millenium
+resdat3[x < ilim[1], x := ilim[1]]
+resdat3[x > ilim[2], x := ilim[2]]
+rlim <- range(aaxis1$logres)
+rlim[1] <- -5  # set limits to 0.01 cm^2
+resdat3[y < rlim[1], y := rlim[1]]
+resdat3[y > rlim[2], y := rlim[2]]
 coordinates(resdat3) <- ~x + y
@@
 extdat3[, n := 1]  # set up count column
-ex <- range(taxis2$logdays)
-ex <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
-extdat3[x < ex[1], x := ex[1]]
-extdat3[x > ex[2], x := ex[2]]
-ey <- range(aaxis2$logarea)
-extdat3[y < ey[1], y := ey[1]]
-extdat3[y > ey[2], y := ey[2]]
+dlim <- range(taxis2$logdays)
+dlim <- c(floor(dlim[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
+extdat3[x < dlim[1], x := dlim[1]]
+extdat3[x > dlim[2], x := dlim[2]]
+elim <- range(aaxis2$logarea)
+extdat3[y < elim[1], y := elim[1]]
+extdat3[y > elim[2], y := elim[2]]
 coordinates(extdat3) <- ~x + y
 
-# temporal framing
+# temporal framing (interval-duration)
 tempdat <- data.table(x = log10(to_infin_byond(kdat$t_btwn_samp)), 
                       y = log10(kdat$eff_dur))#kdat$act_dur))
-# tt <- data.table(x = log10(kdat$t_btwn_samp), y = log10(kdat$eff_dur))
-# range(tt$x - tempdat$x)
-# tempdat[x != log10(365 * 10000)]
 tempdat <- tempdat[!is.na(x) & !is.na(y)] # cut out null values
@@
 
-tempdat[x > y, .N] / tempdat[, .N]
-
 # tempdat$n <- rep(1, nrow(tempdat))
 tempdat[, n := 1]
-tx <- range(tempdat$x)
-tx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
-tx[2] <- 6.6  # just above 10 KA
-# tx[2] <- 5.56
-tempdat[x < rx[1], x := rx[1]]
-tempdat[x > rx[2], x := rx[2]]
-ty <- range(taxis2$logdays)
+# tx <- range(tempdat$x)
+# tx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
+tempdat[x < ilim[1], x := ilim[1]]
+tempdat[x > ilim[2], x := ilim[2]]
+dlimty <- range(taxis2$logdays)
 # ty <- c(floor(ty[1]), 5.56)#ceiling(ex[2]))
-ty <- c(floor(ty[1]), 6.6)  # just above 10 KA
-tempdat[y < ty[1], y := ty[1]]
-tempdat[y > ty[2], y := ty[2]]
+# ty <- c(floor(ty[1]), 6.6)  # just above 10 KA
+tempdat[y < elim[1], y := elim[1]]
+tempdat[y > elim[2], y := elim[2]]
 coordinates(tempdat) <- ~x + y
 
-# spatial framing
+# spatial framing (res-extent)
 spatdat <- data.table(x = log10(kdat$plot_res), y = log10(kdat$eff_ext))#act_ext))
@@
 spatdat[, n := 1]
-sx <- range(aaxis1$logres)
+# sx <- range(aaxis1$logres)
 sx[1] <- -5  # set limits to 0.01 cm^2
-spatdat[x < sx[1], y := sx[1]]
-spatdat[x > sx[2], y := sx[2]]
-sy <- range(aaxis2$logarea)
-spatdat[y < sy[1], y := sy[1]]
-spatdat[y > sy[2], y := sy[2]]
+spatdat[x < rlim[1], y := rlim[1]]
+spatdat[x > rlim[2], y := rlim[2]]
+# sy <- range(aaxis2$logarea)
+spatdat[y < elim[1], y := elim[1]]
+spatdat[y > elim[2], y := elim[2]]
 coordinates(spatdat) <- ~x + y
 
+# # interval-extent
+# iextdat <- data.table(x = log10(kdat$t_btwn_samp), y = log10(kdat$eff_ext))
+# iextdat <- iextdat[!is.na(x) & !is.na(y)] # cut out null values
+# iextdat <- iextdat[round(x, 8) != 6.56229286, ]  # remove one-time observations
+# iextdat[, n := 1]
+# iextdat[x < ilim[1], x := ilim[1]]
+# iextdat[x > ilim[2], x := ilim[2]]
+# iextdat[y < elim[1], y := elim[1]]
+# iextdat[y > elim[2], y := elim[2]]
+# coordinates(iextdat) <- ~x + y
+# 
+# # resolution-duration
+# rdurdat <- data.table(x = log10(kdat$eff_dur), y = log10(kdat$plot_res))
+# rdurdat <- rdurdat[!is.na(x) & !is.na(y)] # cut out null values
+# rdurdat[, n := 1]
+# rdurdat[x < dlim[1], x := dlim[1]]
+# rdurdat[x > dlim[2], x := dlim[2]]
+# rdurdat[y < rlim[1], y := rlim[1]]
+# rdurdat[y > rlim[2], y := rlim[2]]
+# coordinates(rdurdat) <- ~x + y
+
 # densities
-# resolution
-rres3 <- kdensity(rx[1], rx[2], ry[1], ry[2], 0.1, resdat3, 1)
+# interval-resolution
+rres3 <- kdensity(ilim[1], ilim[2], rlim[1], rlim[2], 0.1, resdat3, 1)
 rres3 <- (rres3 / cellStats(rres3, sum)) * 100
 
-# extent
-rext3 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat3, 1)
+# duration-extent
+rext3 <- kdensity(dlim[1], dlim[2], elim[1], elim[2], 0.1, extdat3, 1)
 rext3 <- (rext3 / cellStats(rext3, sum)) * 100
 
-# temporal
-tres <- kdensity(tx[1], tx[2], ty[1], ty[2], 0.1, tempdat, 1)
+# interval-duration
+tres <- kdensity(ilim[1], ilim[2], dlim[1], dlim[2], 0.1, tempdat, 1)
 tres <- (tres / cellStats(tres, sum)) * 100
-# spatial
-sres <- kdensity(sx[1], sx[2], sy[1], sy[2], 0.1, spatdat, 1)
+
+# resolution-extent
+sres <- kdensity(rlim[1], rlim[2], elim[1], elim[2], 0.1, spatdat, 1)
 sres <- (sres / cellStats(sres, sum)) * 100
+
+# # interval-extent
+# iext <- kdensity(ilim[1], ilim[2], elim[1], elim[2], 0.1, iextdat, 1)
+# iext <- (iext / cellStats(iext, sum)) * 100
+# 
+# # resolution-duration
+# rdur <- kdensity(dlim[1], dlim[2], rlim[1], rlim[2], 0.1, rdurdat, 1)
+# rdur <- (rdur / cellStats(rdur, sum)) * 100
 
@@
 }
-# axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
-#      las = 2, tcl = -0.2, cex.axis = cxa)
 mtext("Interval", 1, cex = cxa, line = ll)
@@
 
-# image(rres3, col = rev(terrain.colors(100)))
-# image(rext3, col = rev(terrain.colors(100)))
-# datf[, points(log10(eff_dur), log10(eff_ext), pch = 20, cex = 0.5)]
-# datf[like(study_type, "remote"), {
-#   points(log10(eff_dur), log10(eff_ext), pch = 20, cex = 0.5, col = "blue")
-# }]
-# points(resdat3$x, resdat3$y, pch = 20, cex = 0.5)
-# image(sres, col = rev(terrain.colors(100)))
-# datf[, points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5)]
-# datf[like(study_type, "remote"), {
-#   points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5, col = "blue")
-#   abline(lm(log10(eff_ext) ~ log10(plot_res)))
-# }]
-# datf[like(study_type, "field"), {
-#   points(log10(plot_res), log10(eff_ext), pch = 20, cex = 0.5, col = "red")
-#   abline(lm(log10(eff_ext) ~ log10(plot_res)))
-# }]
-# points(spatdat$x, spatdat$y, pch = 20, cex = 0.5)
-
-# bootp_dt[!t_btwn_samp == 365 * 10000, {
-#   mean(t_btwn_samp[eff_dur > 30 & eff_dur < 365])
-# }]
-# bootp_dt[!t_btwn_samp == 365 * 10000, {
-#   mean(t_btwn_samp[eff_dur > 365 & eff_dur <= (365 * 10)])
-# }]
+# image(rdur, axes = FALSE, xlab = "", ylab = "", col = cols)
+# plot(rdur, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+#      legend.shrink = 0.9, col = cols)
 
@@
 # actual versus effective duration
-kdat <- bootp_dt  # bootstrap
-# tempdat2 <- data.table(x = log10(kdat$act_dur), y = log10(kdat$eff_dur))
-tempdat2 <- data.table(x = log10(kdat$eff_dur), y = log10(kdat$act_dur))
-dinds <- kdat[, which(t_btwn_samp == 365 * 10000)]  # drop unreplicated obs
-tempdat2 <- tempdat2[-dinds, ]
-tempdat2 <- tempdat2[!is.na(x) & !is.na(y)] # cut out null values
-tempdat2[, n := 1]  # set up count column
-tx <- range(taxis2$logdays)
-tx <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
-tempdat2[x < tx[1], x := tx[1]]
-tempdat2[x > tx[2], x := tx[2]]
-ty <- range(taxis2$logdays)
-ty <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
-tempdat2[y < tx[1], y := tx[1]]
-tempdat2[y > tx[2], y := tx[2]]
-coordinates(tempdat2) <- ~x + y
-
-# extent
-# extdat4 <- data.table(x = log10(kdat$act_ext), y = log10(kdat$eff_ext))
-extdat4 <- data.table(x = log10(kdat$eff_ext), y = log10(kdat$act_ext))
-extdat4 <- extdat4[!is.na(x) & !is.na(y)] # cut out null values
-extdat4[, n := 1]  # set up count column
+# kdat <- bootp_dt  # bootstrap
+# # tempdat2 <- data.table(x = log10(kdat$act_dur), y = log10(kdat$eff_dur))
+# tempdat2 <- data.table(x = log10(kdat$eff_dur), y = log10(kdat$act_dur))
+# dinds <- kdat[, which(t_btwn_samp == 365 * 10000)]  # drop unreplicated obs
+# tempdat2 <- tempdat2[-dinds, ]
+# tempdat2 <- tempdat2[!is.na(x) & !is.na(y)] # cut out null values
+# tempdat2[, n := 1]  # set up count column
+# tx <- range(taxis2$logdays)
+# tx <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
+# tempdat2[x < tx[1], x := tx[1]]
+# tempdat2[x > tx[2], x := tx[2]]
+# ty <- range(taxis2$logdays)
+# ty <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
+# tempdat2[y < tx[1], y := tx[1]]
+# tempdat2[y > tx[2], y := tx[2]]
+# coordinates(tempdat2) <- ~x + y
+# 
+# # extent
+# # extdat4 <- data.table(x = log10(kdat$act_ext), y = log10(kdat$eff_ext))
+# extdat4 <- data.table(x = log10(kdat$eff_ext), y = log10(kdat$act_ext))
+# extdat4 <- extdat4[!is.na(x) & !is.na(y)] # cut out null values
+# extdat4[, n := 1]  # set up count column
+# ex <- range(aaxis2$logarea)
+# extdat4[x < ex[1], x := ex[1]]
+# extdat4[x > ex[2], x := ex[2]]
+# ey <- range(aaxis2$logarea)
+# extdat4[y < ey[1], y := ey[1]]
+# extdat4[y > ey[2], y := ey[2]]
+# coordinates(extdat4) <- ~x + y
+# 
+# # densities
+# # temporal 
+# tres2 <- kdensity(tx[1], tx[2], ty[1], ty[2], 0.1, tempdat2, 1)
+# tres2 <- (tres2 / cellStats(tres2, sum)) * 100
+# 
+# # extent
+# rext4 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat4, 1)
+# rext4 <- (rext4 / cellStats(rext4, sum)) * 100
+# 
+# # cxa = 0.7
+
+# plot params
 ex <- range(aaxis2$logarea)
-extdat4[x < ex[1], x := ex[1]]
-extdat4[x > ex[2], x := ex[2]]
 ey <- range(aaxis2$logarea)
-extdat4[y < ey[1], y := ey[1]]
-extdat4[y > ey[2], y := ey[2]]
-coordinates(extdat4) <- ~x + y
-
-# densities
-# temporal 
-tres2 <- kdensity(tx[1], tx[2], ty[1], ty[2], 0.1, tempdat2, 1)
-tres2 <- (tres2 / cellStats(tres2, sum)) * 100
-
-# extent
-rext4 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat4, 1)
-rext4 <- (rext4 / cellStats(rext4, sum)) * 100
-
-# cxa = 0.7
 ll <- 3.5
@@
 cols <- inferno(cuts + bump)[-c(2:(bump + 1))]
-# plot(1:cuts, col = cols, pch = 20)
-
-# par(mfrow = c(1, 2))
-# datf[, {
-#   # sel = which()
-#   eext = log10(eff_ext)
-#   aext = log10(act_ext)
-#   edur = log10(eff_dur[t_btwn_samp != 365 * 10000])
-#   adur = log10(act_dur[t_btwn_samp != 365 * 10000])
-#   plot(aext, eext - aext)
-#   plot(adur, edur - adur)
-#   length(edur)
-# }]
+# # plot(1:cuts, col = cols, pch = 20)
+# 
+# # par(mfrow = c(1, 2))
+# # datf[, {
+# #   # sel = which()
+# #   eext = log10(eff_ext)
+# #   aext = log10(act_ext)
+# #   edur = log10(eff_dur[t_btwn_samp != 365 * 10000])
+# #   adur = log10(act_dur[t_btwn_samp != 365 * 10000])
+# #   plot(aext, eext - aext)
+# #   plot(adur, edur - adur)
+# #   length(edur)
+# # }]
 
@@
 })
-
-# check number of drops due to eff < act
-# hdat_bs2_ns <- do.call(rbind, lapply(hdat_bs2, function(x) x$Ns))
-# summary(bootperturb[[1]][, .N] - hdat_bs2_ns[, 1])
-# summary(bootperturb[[1]][t_btwn_samp != 365 * 10000, .N] - hdat_bs2_ns[, 2])
-
-# extract regression coefficients
-# hdat_bs2_elm <- do.call(rbind, lapply(hdat_bs2, function(x) coef(x$elm)[, 1]))
-# hdat_bs2_tlm <- do.call(rbind, lapply(hdat_bs2, function(x) coef(x$tlm)[, 1]))
 
@@
 
-# Kernel density version
-pdf("paper/figures/act_v_eff.pdf", width = 7, height = 3)
-par(mfrow = c(1, 2), mar = c(4.5, 4.5, 0.3, 3), oma = c(0, 0, 0.2, 1), 
-    mgp = c(2, 0.25, 0))
-
-mtl <- 3
-# spatial extents
-image(rext4, axes = FALSE, xlab = "", ylab = "", col = cols)
-axis(1, at = aaxis2$logarea, labels = alab2, las = 2,
-     cex.axis = cxa, tcl = -0.2)
-axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
-     tcl = -0.2)
-brks <- brkfun(rext4, 0.01, cuts)  # breaks
-aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
-              cex.axis = cxa, tcl = -0.1)
-mtext("Actual extent", 2, cex = cxa, line = mtl)
-mtext("Effective extent", 1, cex = cxa, line = mtl)
-plot(rext4, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, col = cols)
-mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
-
-# temporal extents
-image(tres2, axes = FALSE, xlab = "", ylab = "", col = cols)
-axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
-     las = 2, tcl = -0.2, cex.axis = cxa)
-axis(2, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], las = 2, tcl = -0.2, 
-     cex.axis = cxa)
-mtext("Actual duration", 2, cex = cxa, line = mtl)
-mtext("Effective duration", 1, cex = cxa, line = mtl)
-brks <- brkfun(tres2, 0.01, cuts)  # breaks
-aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
-              cex.axis = cxa, tcl = -0.1)
-plot(tres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-     legend.shrink = 0.9, col = cols)
-     #col = rev(terrain.colors(cuts)))
-mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
-dev.off()
-
+# # Kernel density version
+# pdf("paper/figures/act_v_eff.pdf", width = 7, height = 3)
+# par(mfrow = c(1, 2), mar = c(4.5, 4.5, 0.3, 3), oma = c(0, 0, 0.2, 1), 
+#     mgp = c(2, 0.25, 0))
+# 
+# mtl <- 3
+# # spatial extents
+# image(rext4, axes = FALSE, xlab = "", ylab = "", col = cols)
+# axis(1, at = aaxis2$logarea, labels = alab2, las = 2,
+#      cex.axis = cxa, tcl = -0.2)
+# axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
+#      tcl = -0.2)
+# brks <- brkfun(rext4, 0.01, cuts)  # breaks
+# aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
+#               cex.axis = cxa, tcl = -0.1)
+# mtext("Actual extent", 2, cex = cxa, line = mtl)
+# mtext("Effective extent", 1, cex = cxa, line = mtl)
+# plot(rext4, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+#      legend.shrink = 0.9, col = cols)
+# mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
+# 
+# # temporal extents
+# image(tres2, axes = FALSE, xlab = "", ylab = "", col = cols)
+# axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
+#      las = 2, tcl = -0.2, cex.axis = cxa)
+# axis(2, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], las = 2, tcl = -0.2, 
+#      cex.axis = cxa)
+# mtext("Actual duration", 2, cex = cxa, line = mtl)
+# mtext("Effective duration", 1, cex = cxa, line = mtl)
+# brks <- brkfun(tres2, 0.01, cuts)  # breaks
+# aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
+#               cex.axis = cxa, tcl = -0.1)
+# plot(tres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
+#      legend.shrink = 0.9, col = cols)
+#      #col = rev(terrain.colors(cuts)))
+# mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
+# dev.off()
 ```
@@
 }
-# # par(xpd = NA)
-# for(j in c("in", "dev", "ndc", "nfc", "npc", "nic")) {
-#   xx <- grconvertX(0.5, from = "user", to = j)
-#   ii <- c(0.4, 0.6, 0.8, 1)
-#   yy <- c(0.95, 0.75, 0.5, 0.25)
-#   # yy <- c(-1, -5, -10, -20)
-#   for(i in 1:length(ii)) {
-#     y <- grconvertY(yy[i], from = "user", to = j)
-#     text(x = xx, y = y, labels = paste("kernel =", ii[i]))
-#     # par(xpd = NA)
-#     # mtext(text = paste("kernel =", ii[i]), side = 3, line = yy[i])
-#   }
-# }
 dev.off()
-
-# plot(rrs, xlim = c(-5, 15), ylim = c(-5, 15), 
-#      xlab = "Sample interval - log10(days)", 
-#      ylab = "Plot resolution - log10(m^2)")
-# plot(ers, box = FALSE, xlim = c(-5, 15), ylim = c(-5, 15), 
-#      xlab = "Study duration - log10(days)", ylab = "Sampled area - log10(ha)")
-# dev.off()
 

### 2017-06-13 paper/main.tex
Figure revisions, initial paragraph of discussion section.

@@
 
-Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) analysis of ecological research scales was published in 1989 \cite{tilman_ecological_1989}, it is both timely and important to assess the scales of contemporary ecological investigation.
+Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) meta-analyses of ecological research scales were published nearly 30 years ago \cite{tilman_ecological_1989,kareiva_spatial_1988}, it is both timely and important to assess the scales of contemporary ecological investigation.
 To address this need, we quantified the spatial and temporal domains of empirical observations that were reported within recently (2004-2014) published ecological studies (here domain means the distribution of observations within the spectrum of one or more scale dimensions\footnote{This definition differs slightly from Wiens' \cite{wiens_spatial_1989}, who defined ``domain of scale'' as ``a portion of the scale spectrum within which process-pattern relationships are consistent regardless of scale.''}).
@@
 Our analysis was based on a review of 348 papers randomly selected from 42,918 published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals.
-We extracted scale data from 379 observations of ``natural'' (i.e.
-non-experimentally manipulated) ecological features that were reported within 134 of the reviewed papers (plus an additional 62 that these cited as the source of observations).
-We excluded experiments because they tend to be of limited extent, duration, and resolution due to their higher logistical costs \cite{tilman_ecological_1989, kareiva_spatial_1988}, and would therefore likely bias our findings towards finer scales.
+We extracted scale data from 378 observations of ``natural'' (i.e.
+non-experimentally manipulated) ecological features that were reported within 133 of the reviewed papers (plus an additional 62 that these cited as the source of observations).
+We excluded experiments because they tend to be of limited extent, duration, and resolution due to their higher logistical costs \cite{tilman_ecological_1989, kareiva_spatial_1988}, and would therefore likely bias our findings towards finer scales, while minimizing the impact that new observing methods (e.g.
+satellite imaging, wireless sensing) may have had in expanding the scales of ecological investigation \cite{turner_remote_2003,pettorelli_satellite_2014,porter_wireless_2005}.
 
@@
 1A).
-The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 42\% 1000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig.
+The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 11\% 1-10,000 ha, 19\% 10,000-100,000 ha, 27\% 100,000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig.
 1B).
@@
 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, and 8\% at yearly to decadal intervals.
-Duration was one day or less for 31\% of sampled observations, while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 10\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig.
+Duration was one day or less for 31\% of sampled observations, while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 9\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig.
 1D).
@@
 
-A similar tendency for duration to increase with interval was also evident amongst temporally replicated observations (Fig.
-2D), with the majority of repeated at daily to decadal intervals and spanning $\geq$ 1 month to 1 decade.
+The majority of temporally replicated observations have daily to decadal intervals and durations of $\geq$ 1 month to 1 decade.
 The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting one year to one decade tend to have yearly to decadal intervals.
@@
 This comparison showed that a majority (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).
-Actual extent converged with extent above 100,000 ha, but this applies to just 5\% of observations.
+Actual extent converged with extent $\geq$1 100,000 ha, but this applies to just 5\% of observations.
 The actual duration of 64\% of observations was $<$ 1 day, which on average was 3 to nearly 9 orders of magnitude shorter than the time span covered by temporal replicates (Fig.
@@
 We classified the method used to collect each observation into several broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing, other geographic data, and paleo-reconstruction approaches.
-Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.3\%, and paleo-reconstruction and other geographic data each for less than 1\%.
+Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.4\%, and paleo-reconstruction and other geographic data each for less than 1\% each.
 Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance.
@@
 There were several potential methodological aspects that could have influenced our assessment of ecology's spatial and temporal domains.
-The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate--rather than simply record--these values for most observations (specifically, in 63, 60, and 67\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration, respectively).
+The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate--rather than simply record--these values for most observations (specifically, in 63, 60, and 69\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration, respectively).
 The inevitable estimation errors may have biased our overall findings.
@@
 \noindent \textbf{Broader implications}\\
+Our results suggest that the spatial and temporal domains of modern ecological observations are fairly narrow, and are collected primarily with conventional field techniques.
+Spatially, the majority of observations have grains of $<1$ m$^2$ and extents of $<$10,000 ha (Fig 1A;B).
+In the temporal domains, most observations are either un-replicated snapshots, or of low frequency ($\geq$1 month interval; Fig.
+1C) and relatively short duration ($leq$1 year; Fig.
+1D; 2D).
+Comparisons between observational dimensions revealed that larger extents and are associated with larger plot sizes (Fig.
+2C), while longer durations are associated with longer intervals (Fig.
+2D).
+The latter association presumably reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of field-based data collection methods.
+This same tradeoff is also responsible for the inverse relationship between resolution and interval that is the most obvious feature within that domain space (Fig.
+1A).
 
-Our results show that modern ecological observations most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month).
+Comparing these general findings to those of earlier related studies, it appears that modern ecology's observational scales have not changed greatly in the past few decades.
+
+These general results suggest that the scale domains of modern ecology have not changed greatly since the 1980s.  Porter's study.
+
+Comparing these primary tendencies to earlier efforts, it seems that not much has changed in ecological
+
+Taken collectively, these
 
@@
   \subitem{-} Notable holes: high frequency (sub-daily, daily), high resolution 
-
+   \subitem{-} Change since Tilman/Kareiva
    \subitem{-} Duration hasn't improved much since Tilman (check this).

### 2017-06-20 Rmd/space-time.Rmd
Full revision, including discussion and NatEcolEvol style abstract.

@@
 p_fig <- fp(p_root, "paper/figures")
+p_til <- fp(p_root, "external/data/misc")
 
@@
 ```{r, eval=FALSE}
-# p <- c(0.64, 0.84, 0.53, 1.21)  # from calibration.Rmd
-# p <- c(0.58, 0.82, 0.99, 1.24)  # from calibration.Rmd
-##     plot_res   n_sites   act_ext   eff_ext t_btwn_samp  act_dur   eff_dur
-## 1: 0.5769693 0.3821107 0.7196559 0.4049972     1.01134 1.253045 0.5284913
-##    plot_res n_sites  act_ext  eff_ext t_btwn_samp  act_dur  eff_dur
-##       0.494   0.457    0.749    0.499       0.475    1.093    0.338
-##    plot_res n_sites  act_ext  eff_ext t_btwn_samp  act_dur  eff_dur
-##       0.577   0.382     0.72    0.405       1.045    1.261    0.643
 p <- c(0.58, 0.72, 0.41, 1.05, 1.26, 0.64)  # from calibration.Rmd
@@
 bootperturb <- resamp_func(datf, p = p, iter = 1000) 
-# bootperturb2 <- resamp_func(datf, p = p, iter = 1000)  # 10% perturb
-# all(bootperturb[[200]]$plot_res == bootperturb2[[200]]$plot_res)
-# rm(bootperturb2)
 ```
@@
 
-# spatial
-# plot(1:10, xlab = expression(paste("mm"^2)), xaxt = "n")
-# axis(1, at = 1:8, labels = lab, las = 2)
-
 # plot resolution
@@
 ### Figure 1 
-
 Histograms for each of the four main dimensions.
@@
 
-# length(which(hdat$res >= -3 & hdat$res <= 0)) / length(hdat$res)
-# length(which(hdat$res <= 4)) / length(hdat$res)
-# sum(hdatstat[[1]]$counts[4:6]) / sum(hdatstat[[1]]$counts)
-
 # extent
@@
 l <- length(hdatst)
-hds <- c("norep", "<daily", "<weekly",  "day-mo", "mo-yr", "yr-dec")
-inds <- list(l, 1:3, 1:4, 4:5, 6, 7:8)
+hds <- c("norep", "<daily", "<weekly", "<minute", "min-hour", 
+         "hr-day", "day-wk", "wk-mo", "mo-yr", "yr-dec", "dec-cen", "cen-mil")
+inds <- list(l, 1:3, 1:4, 1, 2, 3, 4, 5, 6, 7, 8, 9)
 sapply(1:length(inds), function(x) {
@@
 })
-
-# datf[, plot(log10(n_sites), log10(plot_res))]
-# dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
-# datf[like(study_type, "field"), plot(log10(plot_res), log10(act_ext), 
-#                                      ylim = c(-10, 10))]
-# datf[like(study_type, "remote"), 
-#      points(log10(plot_res), log10(act_ext), pch = 20, col = "red")]
-# datf[like(study_type, "automated"), 
-#      points(log10(plot_res), log10(act_ext), pch = 20, col = "blue")]
-# lines(aaxis1$logres, aaxis2$logarea[-c(16:17)] )
-# datf[, .N / nrow(datf) * 100, by = st]
-
-# datf[, plot(log10(t_btwn_samp), log10(act_dur), pch = 20)]
-# datf[which(hdat[[3]] < -1.380211), 
-#      points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "blue")]
-# datf[like(study_type, "automated"), 
-#      points(log10(t_btwn_samp), log10(act_dur), pch = 20, col = "red")]
 
@@
 mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
-# for(i in 1:length(stps)) {
-#   points(resdat3$x[stps[[i]]], resdat3$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
-# }
 
 # extents
-# image(rext3, col = rev(terrain.colors(30)), axes = FALSE, xlab = "", 
 image(rext3, axes = FALSE, xlab = "", ylab = "", col = cols)
@@
 mtext("C", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
-# for(i in 1:length(stps)) {
-#   points(spatdat$x[stps[[i]]], spatdat$y[stps[[i]]], pch=pchs[[i]], cex=cexs[i])
-# }
 
@@
 
+save(bootp_dt, file = fp(p_dat, "derived/bootresults.rda"))
 ```
@@
 ### Figure 3
+
+Actual versus effective duration/extent
 ```{r, eval=FALSE}
-# actual versus effective duration
-# kdat <- bootp_dt  # bootstrap
-# # tempdat2 <- data.table(x = log10(kdat$act_dur), y = log10(kdat$eff_dur))
-# tempdat2 <- data.table(x = log10(kdat$eff_dur), y = log10(kdat$act_dur))
-# dinds <- kdat[, which(t_btwn_samp == 365 * 10000)]  # drop unreplicated obs
-# tempdat2 <- tempdat2[-dinds, ]
-# tempdat2 <- tempdat2[!is.na(x) & !is.na(y)] # cut out null values
-# tempdat2[, n := 1]  # set up count column
-# tx <- range(taxis2$logdays)
-# tx <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
-# tempdat2[x < tx[1], x := tx[1]]
-# tempdat2[x > tx[2], x := tx[2]]
-# ty <- range(taxis2$logdays)
-# ty <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
-# tempdat2[y < tx[1], y := tx[1]]
-# tempdat2[y > tx[2], y := tx[2]]
-# coordinates(tempdat2) <- ~x + y
-# 
-# # extent
-# # extdat4 <- data.table(x = log10(kdat$act_ext), y = log10(kdat$eff_ext))
-# extdat4 <- data.table(x = log10(kdat$eff_ext), y = log10(kdat$act_ext))
-# extdat4 <- extdat4[!is.na(x) & !is.na(y)] # cut out null values
-# extdat4[, n := 1]  # set up count column
-# ex <- range(aaxis2$logarea)
-# extdat4[x < ex[1], x := ex[1]]
-# extdat4[x > ex[2], x := ex[2]]
-# ey <- range(aaxis2$logarea)
-# extdat4[y < ey[1], y := ey[1]]
-# extdat4[y > ey[2], y := ey[2]]
-# coordinates(extdat4) <- ~x + y
-# 
-# # densities
-# # temporal 
-# tres2 <- kdensity(tx[1], tx[2], ty[1], ty[2], 0.1, tempdat2, 1)
-# tres2 <- (tres2 / cellStats(tres2, sum)) * 100
-# 
-# # extent
-# rext4 <- kdensity(ex[1], ex[2], ey[1], ey[2], 0.1, extdat4, 1)
-# rext4 <- (rext4 / cellStats(rext4, sum)) * 100
-# 
-# # cxa = 0.7
 
@@
 cols <- inferno(cuts + bump)[-c(2:(bump + 1))]
-# # plot(1:cuts, col = cols, pch = 20)
-# 
-# # par(mfrow = c(1, 2))
-# # datf[, {
-# #   # sel = which()
-# #   eext = log10(eff_ext)
-# #   aext = log10(act_ext)
-# #   edur = log10(eff_dur[t_btwn_samp != 365 * 10000])
-# #   adur = log10(act_dur[t_btwn_samp != 365 * 10000])
-# #   plot(aext, eext - aext)
-# #   plot(adur, edur - adur)
-# #   length(edur)
-# # }]
 
@@
   DT <- bootperturb[[x]]
+
+  # extent measures log10 transformed
   edb <- DT[, lapply(.SD, log10), .SDcol = c("eff_ext", "act_ext")]
-  # edb <- edb[eff_ext >= act_ext, ]  # drop eff_dur < act
   setnames(edb, c("eff_ext", "act_ext"), c("eext", "aext"))
-  edb[, dif := eext - aext]
-
+  edb[, dif := eext - aext]  # difference between extent and actual extent
+
+  # duration measures log10 transformed, unreplicated dropped
   tdb <- DT[t_btwn_samp != 365 * 10000, lapply(.SD, log10), 
             .SDcol = c("eff_dur", "act_dur")]
-  # tdb <- tdb[eff_dur >= act_dur, ]  # drop eff_dur < act
   setnames(tdb, c("eff_dur", "act_dur"), c("edur", "adur"))
   # print(paste(edb[, .N], tdb[, .N]))
-  tdb[, dif := edur - adur]
+  tdb[, dif := edur - adur]  # difference between duration and actual duration
   return(list("edb" = edb, "tdb" = tdb, #"elm" = elm, "tlm" = tlm, 
@@
 
-# # Kernel density version
-# pdf("paper/figures/act_v_eff.pdf", width = 7, height = 3)
-# par(mfrow = c(1, 2), mar = c(4.5, 4.5, 0.3, 3), oma = c(0, 0, 0.2, 1), 
-#     mgp = c(2, 0.25, 0))
-# 
-# mtl <- 3
-# # spatial extents
-# image(rext4, axes = FALSE, xlab = "", ylab = "", col = cols)
-# axis(1, at = aaxis2$logarea, labels = alab2, las = 2,
-#      cex.axis = cxa, tcl = -0.2)
-# axis(2, at = aaxis2$logarea, labels = alab2, las = 2, cex.axis = cxa,
-#      tcl = -0.2)
-# brks <- brkfun(rext4, 0.01, cuts)  # breaks
-# aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
-#               cex.axis = cxa, tcl = -0.1)
-# mtext("Actual extent", 2, cex = cxa, line = mtl)
-# mtext("Effective extent", 1, cex = cxa, line = mtl)
-# plot(rext4, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-#      legend.shrink = 0.9, col = cols)
-# mtext("A", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
-# 
-# # temporal extents
-# image(tres2, axes = FALSE, xlab = "", ylab = "", col = cols)
-# axis(1, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], 
-#      las = 2, tcl = -0.2, cex.axis = cxa)
-# axis(2, at = taxis1$logdays[-c(5)], labels = tlab1[-c(5)], las = 2, tcl = -0.2, 
-#      cex.axis = cxa)
-# mtext("Actual duration", 2, cex = cxa, line = mtl)
-# mtext("Effective duration", 1, cex = cxa, line = mtl)
-# brks <- brkfun(tres2, 0.01, cuts)  # breaks
-# aargs <- list(mgp = c(3, 0.25, 0), at = brks$labs, labels = brks$labs, 
-#               cex.axis = cxa, tcl = -0.1)
-# plot(tres2, legend.only = TRUE, axis.args = aargs, legend.width = 1.25,
-#      legend.shrink = 0.9, col = cols)
-#      #col = rev(terrain.colors(cuts)))
-# mtext("B", side = 3, line = -1, cex = 0.8, adj = 0.05, col = "white")
-# dev.off()
-```
-
-## Supplemental Analyses
+```
+
+## Additional Analyses
 
@@
 notremote <- datf[study_type != "remote", eff_ext]
-# a <- (length(remote) + rsinc) / (nrow(datf))
-# b <- length(remote) / nrow(datf)
-# a / b
-# ((length(remote) + rsinc) / (length(notremote)) / length(b))
-
-# (rsext * rs2017 + nonrsext * (1 - rs2017)) / 
-#   (rsext * rs2014 + nonrsext * (1 - rs2014))
 
@@
 
+### Scales in earlier studies
+#### Duration from Tilman, 1989
+```{r, echo=FALSE}
+load(fp(p_dat, "derived/bootresults.rda"))
+```
+```{r}
+std <- function(x) (x - min(x)) / diff(range(x))
+
+# Tilman (1989), Figure 6.1 (visually interpreted)
+# Data extracted from figure using http://arohatgi.info/WebPlotDigitizer/app/
+tilman <- fread(fp(p_til, "tilman1989.csv"))
+
+# Find n and proportion of non-experimental observations per year
+# find mean/median value of data > 100 years in our observed data
+gt100 <- bootp_dt[eff_dur > (365 * 100), {
+  list("med" = median(eff_dur) / 365, "mu" = mean(eff_dur) / 365) 
+}]
+
+# find weights and proportion/year
+tilman_noexp <- tilman[, {
+  a = round(623 * prop[fig == 1] * (1 / sum(prop[fig == 1])))
+  b = round(180 * prop[fig == 2] * (1 / sum(prop[fig == 2])))
+  padj = a - b
+  w = padj / sum(padj)
+  list("yrmu" = yrmu[fig == 1], "n" = padj, "prop" = w)
+}]
+
+# median and mean per year, assuming 150 years for studies > 100
+tilmedmu <- tilman_noexp[, rep(yrmu, n), by = n][, {
+  list("med" = median(V1), "mean" = mean(V1))
+}]
+
+# test additional upper bin years
+gt1002 <- unname(c(200, 500, 1000, unlist(gt100)))
+tilmedmu2 <- rbindlist(lapply(gt1002, function(x) {
+  DT2 = copy(tilman_noexp)[yrmu == 150, yrmu := x]
+  DT2[, rep(yrmu, n), by = n][, list("med" = median(V1), "mean" = mean(V1))]
+}))
+
+durmedmu <- bootp_dt[, {
+  list("med" = median(eff_dur[t_btwn_samp != (365 * 10000)], na.rm = TRUE) / 365, 
+       "mean" = mean(eff_dur[t_btwn_samp != (365 * 10000)], na.rm = TRUE) / 365, 
+       "medall" = median(eff_dur, na.rm = TRUE) / 365, 
+       "meanall" = mean(eff_dur, na.rm = TRUE) / 365)
+}]
+
+# And calculate only for years less than 100
+tilmedmu_lt100 <- tilman_noexp[yrmu < 150, rep(yrmu, n), by = n][, {
+  list("med" = median(V1), "mean" = mean(V1))
+}]
+
+durmedmu_lt100 <- bootp_dt[eff_dur < (365 * 100), {
+  list("med" = median(eff_dur[t_btwn_samp != (365 * 10000)], na.rm = TRUE) / 365, 
+       "mean" = mean(eff_dur[t_btwn_samp != (365 * 10000)], na.rm = TRUE) / 365, 
+       "medall" = median(eff_dur, na.rm = TRUE) / 365, 
+       "meanall" = mean(eff_dur, na.rm = TRUE) / 365)
+}]
+
+```
+
+Median, mean duration from Tilman (1989)
+```{r}
+round(rbind(tilmedmu, tilmedmu2), 2)
+```
+
+Tilman with studies > 100 years dropped
+```{r}
+round(tilmedmu_lt100, 2)
+```
+
+Tilman (1989) duration of observational studies by bin
+```{r}
+tilman_noexp[, {
+  a = round(prop * 100, 1)
+  names(a) = tilman[fig == 1, yr]
+  rbind(a, n)
+}]
+```
+
+Median, mean duration from this study
+```{r}
+round(durmedmu, 2)
+```
+
+This study with years > 100 dropped
+```{r}
+round(durmedmu_lt100, 2)
+```
+
+#### Kareiva and Anderson (1989)
+
+Resolution versus this study
+```{r}
+# helper func
+q80 <- function(x) quantile(x, 0.8, na.rm = TRUE)
+q90 <- function(x) quantile(x, 0.9, na.rm = TRUE)
+q95 <- function(x) quantile(x, 0.95, na.rm = TRUE)
+
+kareiva <- fread(fp(p_til, "kareiva_anderson1989.csv"))
+a <- kareiva[, {
+  list("med" = median(res^2), "mean" = mean(res^2), "80" = q80(res^2), 
+       "90" = q90(res^2), "95" = q95(res^2))
+}]
+b <- bootp_dt[, {
+  list("med" = median(plot_res), "mean" = mean(plot_res), "80" = q80(plot_res), 
+       "90" = q90(plot_res), "95" = q95(plot_res))
+}]
+study <- c("Kareiva & Anderson, 1989", "This study")
+cbind(study, round(rbind(a, b), 2))
+```
+
+#### Porter et al (2005)
+Interval
+```{r}
+porter <- fread(fp(p_til, "porter.csv"))
+porter <- porter[sample == 1]
+porter[, ext := ext^2 / 10000]
+
+study <- c("Porter et al, 1989", "This study")
+a <- porter[int > 0, {
+  list("med" = median(int), "mean" = mean(int), "80" = q80(int), "90" = q90(int),
+       "95" = q95(int))
+}]
+
+b <- bootp_dt[t_btwn_samp != (365 * 10000), {
+  list("med" = median(t_btwn_samp), "mean" = mean(t_btwn_samp), 
+       "80" = q80(t_btwn_samp), "90" = q90(t_btwn_samp), "95" = q95(t_btwn_samp))
+}]
+cbind(study, round(rbind(a, b)))
+```
+
+Extent
+```{r}
+study <- c("Porter et al, 1989", "This study", "Ratio")
+a <- porter[, {
+  list("med" = median(ext), "mean" = mean(ext), "80" = q80(ext), 
+       "90" = q90(ext), "95" = q95(ext))
+}] 
+b <- bootp_dt[, {
+  list("med" = median(eff_ext, na.rm = TRUE), 
+       "mean" = mean(eff_ext, na.rm = TRUE), "80" = q80(eff_ext), 
+       "90" = q90(eff_ext), "95" = q95(eff_ext))
+}]
+cbind(study, round(rbind(a, b, b / a)))
+```
+
+### Share of remote sensing, automated, and paleo observations
+```{r}
+
+aext <- 10^4
+datf[, {
+  a = length(which(act_ext > aext))
+  b1 = length(which((act_ext > aext) & like(study_type, "remote")))
+  b2 = length(which((act_ext > aext) & like(study_type, "other")))
+  round(c("remote" = b1 / a * 100, "other" = b2 / a * 100), 1)
+}]
+
+adur <- 365
+datf[, {
+  a = length(which(act_dur > adur))
+  b1 = length(which((act_dur > adur) & like(study_type, "auto")))
+  b2 = length(which((act_dur > adur) & like(study_type, "paleo")))
+  round(c("automated" = b1 / a * 100, "paleo" = b2 / a * 100), 1)
+}]
+
+datf[, {
+  a = length(which(act_dur <= adur))
+  b1 = length(which((act_dur <= adur) & like(study_type, "auto")))
+  b2 = length(which((act_dur <= adur) & like(study_type, "paleo")))
+  round(c("automated" = b1 / a * 100, "paleo" = b2 / a * 100), 1)
+}]
+
+```
+
 ### Varying kernel sizes

### 2017-06-20 paper/SI.tex
Everything through final discussion section revised
Full revision, including discussion and NatEcolEvol style abstract.

@@
 
+\subsubsection*{Extracting and analyzing data from earlier meta-analyses}
+
+419 observational studies lasting $\leq100 years; longer studies were excluded because the upper duration bound was not published)
+
+reviewed in this study
+
 \subsection*{Results}
@@
 CV of overall mean & 50 & 50 & 75 & 48 & 34 & 109 \\
-Mean of record-wise CV & 57 & 41 & 72 & 104 & 64 & 124 \\
+Mean of record-wise CV & 57 & 41 & 72 & 104 & 64 & 126 \\
 \hline

### 2017-06-20 paper/main.tex
Full revision, including discussion and NatEcolEvol style abstract.

@@
 %Ecologists have long known that the observational scales play a critical role in shaping our understanding of ecosystems, and that observations spanning multiple scales of space and time are needed to effectively characterize ecological phenomena.
-\textbf{In order to properly understand ecological phenomena, it is necessary to quantify their behavior over the range of spatial and temporal scales at which they manifest.
-Ecology has been concerned with this need for decades, and the ability to collect ecological observations across multiple scales has grown rapidly since then.
-Characterizing the spatial and temporal domains (the distribution of observations within one or more scale spectra) of modern ecological observations can therefore provide important insight into the field's progress towards a more comprehensive understanding of ecosystem behaviour.
-To characterize these domains, we conducted a meta-analysis of recent (2004-2014) ecological studies, in which we quantified four primary dimensions of their reported observations: plot resolution, sampling interval, effective duration (time between start and end of temporal replicates), and effective extent (area enclosed by spatial replicates).
-We also estimated the \emph{actual} extent and duration, which respectively represent the summed area and time covered by spatial and temporal replicates.
-\textcolor{blue}{Replace this text with more specific summary of observed scales:} \textcolor{red}{Here we show that ecology remains a largely field-based discipline that makes observations within generally narrow spatial and temporal domains, despite the well-established literature on the importance of scale \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}}}.
-% Ecologists still make limited use of remote sensing (6\% of observations), which is necessary for observing larger-scale patterns and processes \cite{turner_remote_2003,pettorelli_satellite_2014}, but make slightly greater use of technologies that capture high frequency or temporally continuous point-based measurements (12\% of observations).
-Resolution, extent, and duration were not clearly reported in most reviewed studies.
-This indicates inattentiveness to scale, which, together with the narrow spatio-temporal domains of observations, suggests a limited ability to generalize the knowledge gained in many ecological studies, which is critical for developing an ability to predict ecological responses in an era of increasing global change \cite{levin_problem_1992}.
-An emerging wave of observing and analytical technologies (e.g.
-nano-satellites, unmanned systems, open image archives, and high-power processing platforms) now makes it dramatically easier and cheaper to collect high-quality, ecologically meaningful measurements at scales inaccessible to field scientists.
-Combined with a greater attention to scale in research design, these advances can facilitate the spread of ecological research to sparsely observed scales, while improving the transferability of ecological knowledge.}
+\textbf{To properly understand ecological phenomena, it is necessary to observe them across a range of spatial and temporal scales.
+Ecologists first raised this point in the 1980s, since when the ability to collect multi-scale observations has grown rapidly.
+To assess modern ecology's progress in addressing scale, we analyzed the resolution, extent, interval, and duration of observations in 348 studies published between 2004-2014.
+We found that the scale domains of observations remain fairly narrow, and are collected primarily with conventional field techniques.
+Compared to prior meta-analyses from the 1980s and early 2000s, observational durations and resolutions remain largely unchanged, but intervals have become finer and extents larger.
+Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, raising questions about observational representativeness.
+Adding to this concern, observational scales were not clearly reported in most studies, suggesting that scale often is a minor consideration.
+Journals can help mitigate this problem by adopting scale reporting standards, which can further spur ecologists to adopt rapidly improving observational technologies, and thereby close key gaps in current observational domains.}
+
+%as well as \emph{actual} extent (the summed resolution of spatial replicates) and \emph{actual} duration (integrated sample duration of temporal replicates).
+Our results suggest that the scale domains of modern ecological observations remain fairly narrow, and are collected primarily with conventional field techniques.
+A comparison with results from prior earlier, scale-focused meta-analyses revealed that observational durations and resolutions are largely unchanged since the 1980s, but intervals have become finer and extents larger, the latter driven by steadily increasing use of remote sensing.
+%Most observations had resolutions of $\leq$1 m$^2$ (67\%) and extents $\leq$10,000 ha (53\%).
+Temporally, most observations were either unreplicated (37\%) or had intervals $\geq$1 month (26\%), while 64\% of observations had durations of $\leq$1 year.
+Juxtaposing dimensions revealed several notable ``holes'' (domains with few to no observations), particularly in the realms of high frequency (sub-daily to daily intervals) observations with fine to moderate resolutions ($\geq$1 m$^2$ up to 100 ha) and $\geq$decadal durations.
+These holes relate to tradeoffs between observational frequency and resolution, as well as duration, that are characteristic of field-based data collection, which was used for 80\% of observations.
+An analysis of values reported in the most comparable prior scale meta-analyses revealed that observational durations and resolutions are largely unchanged since the 1980s, but intervals have become finer and extents larger.
+Our data also showed an internal trend of increasing observational extent, coinciding with a 1.3\% per year increase in remote sensing-based observations.
+Despite this trend, the \emph{actual} extent of 81\% of observations was 4-8 orders of magnitude smaller than extent.
+Similarly, the \emph{actual} duration of 67\% of observations was 3-9 orders of magnitude shorter than duration.
+These large discrepancies raise questions about the representativeness of observations, which may in part be addressed by examining the degree to which auto-correlation processes close the gap between actual and represented scales of observation.
+Adding to this concern is the fact that most observations did not clearly report observational scales, suggesting that scale remains a minor consideration throughout much of ecology.}
 
@@
 
+\noindent \textbf{Results}\\
 \noindent \textbf{\emph{Domains}}\\
@@
 
-Juxtaposing these observational dimensions provides further insight into the spatio-temporal domains of ecological observations (Fig.
+Juxtaposing these observational dimensions provides further insight into the spatio-temporal domains of observations (Fig.
 2).
-Contrasting resolution with interval reveals that the majority of temporally replicated observations (the 37\% that were unreplicated were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig.
+Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig.
 2A).
@@
 To provide further insight into observational domains, we also evaluated the degree to which the scales \emph{represented} by observations differed from those that were actually observed.
-To make this assessment, we first log$_10$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, in order to calculate the magnitude of difference (or decade) between each pair of dimensions for each observation.
+To make this assessment, we first log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, in order to calculate the magnitude of difference (or decade) between each pair of dimensions for each observation.
 We then compared how these magnitudes varied in relation to the scale of the actual dimension (Fig.
@@
 Further evidence for this trend lies within the extent values themselves, which increased 0.25 orders of magnitude per year between 2004-2014 (R$^2$ = 0.25, p$<$0.07).
-This somewhat clearer trend also suggests that including more recent studies would show extent to be somewhat larger, although in this case by a more modest 5.5\% (0.02 orders of magnitude).
+This somewhat clearer trend also suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).
 
@@
 point-based field measurements) may effectively represent larger scales than our estimates suggest, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997, palmer_scale_1994,cao_comparison_2002, legendre_spatial_1993,collins_method_2000-1}.
-From a spatial perspective, this concern does not apply to contiguous sampling schemes, which were primarily based on remote sensing and cover the largest area, thus the net effect is that this analysis may underestimate the effective areas of smaller observational extents.
-From a temporal perspective, our analysis likely underestimates the effective duration of many observations, particularly those where instantaneous, repeated measures of slow-changing ecological features were made.
-Snapshots in time may be sufficient to capture the temporal dynamics of such phenomena (e.g.
-changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these the total time elapsed between the first and last measurements (the span of observation) may more closely approximate effective duration.
-However, many studies focused on more dynamic phenomena, and for these long periods of continuous observation may be more important for understanding dynamics than frequent repeats.
-For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to directly measure fire behavior \cite{clements_observing_2007}, except patterns resolving themselves at scales greater than 12-24 hours.
-For this reason, and to provide a consistent standard for estimating duration, we recorded actual duration.
 
 %Adding to our motivation to record actual observational scales was the fact that 
-
-\noindent \textbf{Broader implications}\\
-Our results suggest that the spatial and temporal domains of modern ecological observations are fairly narrow, and are collected primarily with conventional field techniques.
-Spatially, the majority of observations have grains of $<1$ m$^2$ and extents of $<$10,000 ha (Fig 1A;B).
+\noindent \textbf{Discussion}\\
+\noindent \textbf{\emph{Insights into the scale domains of modern ecology}}\\
+Our results suggest that the scale domains of modern ecological observations are fairly narrow, and are collected primarily with conventional field techniques.
+Spatially, the majority of observations have grains of $\leq$1m$^2$ and extents of $\leq$10,000 ha (Fig 1A;B).
 In the temporal domains, most observations are either un-replicated snapshots, or of low frequency ($\geq$1 month interval; Fig.
-1C) and relatively short duration ($leq$1 year; Fig.
+1C) and relatively short duration ($\leq$1 year; Fig.
 1D; 2D).
-Comparisons between observational dimensions revealed that larger extents and are associated with larger plot sizes (Fig.
+Contrasting observational dimensions reveals that larger extents are associated with larger plot sizes (Fig.
 2C), while longer durations are associated with longer intervals (Fig.
 2D).
-The latter association presumably reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of field-based data collection methods.
-This same tradeoff is also responsible for the inverse relationship between resolution and interval that is the most obvious feature within that domain space (Fig.
-1A).
+The latter association presumably reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of traditional field-based observation.
+The same tradeoff is also responsible for the inverse relationship between resolution and interval that dominates that domain space (Fig.
+2A).
+As a result of these tradeoffs, there are notable ``holes'' in the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($\geq$1 m$^2$ up to 100 ha; Fig.
+2A) and 2) decadal or longer durations (Fig.
+2B).
 
-Comparing these general findings to those of earlier related studies, it appears that modern ecology's observational scales have not changed greatly in the past few decades.
+%there are prominent holes in the scale domains of ecology, notably in the spaces defined by observations having 1) daily to sub-daily intervals and resolutions $\geq$1 m$^2$ up to 100 ha; 2) decadal or longer durations and extents $\geq$10,000,000 ha, such that few observations are collected at 1) high spatial and temporal resolution (1 m$^2$, 2) over large areas for long time periods.
 
-These general results suggest that the scale domains of modern ecology have not changed greatly since the 1980s.  Porter's study.
+Have these domains changed since the seminal papers on scale first began to appear in the late 1980s \cite{wiens_spatial_1989, levin_problem_1992, tilman_ecological_1989}?
+A comprehensive answer to this question is not possible, absent a similar study targeting earlier literature, but an analysis of results presented in three prior studies provides partial insight.
+The first and most comprehensive dataset consists of duration values extracted by Tilman \cite{tilman_ecological_1989} from 623 studies published between 1977-1987 in the journal \emph{Ecology}.
+The average duration of the most comparable subset of those values (n=419; see SI) was 3.6 years, compared to 3.3 years for observations in our sample (or 5.1 if temporally un-replicated observations are excluded).
+The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who present the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986.
+The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,510,247 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (197 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
+The third dataset is provided by Porter et al \cite{porter_wireless_2005}, who compared the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}).
+The mean interval was 178 days, compared to 684 in our sample, but the 80th percentile value in our study was 169 compared to 329 in the earlier study.
+Extent in our sample was substantially larger according to multiple summary statistics, including the mean (115,098,283 ha in our study versus 368,403 ha), median (5,120 ha versus 9 ha), and 95th percentile (45,104,594 ha versus 136,000 ha).
 
-Comparing these primary tendencies to earlier efforts, it seems that not much has changed in ecological
+Although limited due to methodological differences (e.g.
+a focus on experiments versus un-manipulated systems; SI), these comparisons suggest that the duration and, less clearly, resolution of ecological observations have changed little in the past 30 years, but observational frequency and particularly extent have increased.
+The growth in observational extent is also reflected by the positive year-on-year trend within our own dataset, which corresponds to the increasing use of remote sensing (SI).
 
-Taken collectively, these
+However, even though observational extent is increasing, there remains for most observations a large gulf between the area that is actually sampled and that which the spatial replicates purportedly represent (Fig.
+3A).
+A similarly large discrepancy is also evident between the time spanned by repeated observations and the time that is spent actually observing a phenomenon (Fig.
+3B).
+These differences between the actual and \emph{representative} scales of observation have implications for ecological understanding, as the unobserved portions of space and time may contain important patterns and processes that are not captured by replicates, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002, legendre_spatial_1993,collins_method_2000}.
+Brief, infrequent snapshots, or fine-grained, spatially sparse replicates, may be sufficient to characterize many phenomena (as an example with respect to temporal replication, annual changes in tree cover are well-represented by low frequency satellite imaging \cite{hansen_high-resolution_2013}), but may be inadequate for more dynamic phenomena.
+For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe fire behavior \cite{clements_observing_2007}.
+To capture such behavior, long periods of continuous observation may be more important for understanding dynamics than frequent repeats.
 
-\begin{enumerate}
-\item General patterns of scales 
-  \subitem{-} General patterns
-  \subitem{-} The continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
-  \subitem{-} Notable holes: high frequency (sub-daily, daily), high resolution 
-   \subitem{-} Change since Tilman/Kareiva
-   \subitem{-} Duration hasn't improved much since Tilman (check this).
-   \subitem{-} Evaluate how duration has improved since then.
-387/749 of the papers he looked at from the journal Ecology (1977-1984) were mainly experimental or had an experimental component.
-99 papers were mainly theoretical or conceptual.
-Remainder observational.
-We could compare our results to that here.
-Of his 623 field based studies (including 180 experiments), $\sim$41\% lasted less than one year,  $\sim$85\% less than three years, most studies $>$10 years and all $>$50 years were paleoecology/chronosequences.
-  \subitem{-} Maybe mention here (or probably better later in the discussion), that our decision has the unfortunate consequence of not being directly comparable to previous assessments that included experiments.
-More of just a caveat or note for the reader who is familiar with the earlier work and their summary statistics
-  \subitem But there do seem to be trends increasing extent (and duration, but need to put this above), perhaps due to increasing use of remote sensing 
-\item Scales are not well-documented 
-  \subitem Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
-(Scale is not a concern that cuts across the entire discipline of ecology).
+It is therefore important to examine whether the phenomena being observed are adequately captured by the design of replicates.
+Our methods suggest one possible procedure for assessing the representativeness of replicates: 1) measure auto-correlation (spatial or serial) in the replicates, 2) add the autocorrelation length to the replicate area/duration, 3) calculate an autocorrelation-adjusted actual extent/duration, and 4) plot where it falls between actual extent/duration and extent/duration.
+If the adjusted actual value is substantially closer to the representative value, then this may provide greater confidence in the adequacy of the sampling scheme.
+If not, then alternative sampling methods may be used to close this gap.
+For example, remote sensing provides wall-to-wall spatial coverage of a study area, erasing the difference between actual extent and extent.
+Similarly, the interval of high-resolution (a desirable image property because it allows individual ecological features to be more effectively discerned  \cite{dark_modifiable_2007, hay_comparison_2003}) imaging is now approaching daily to sub-daily scales \cite{drusch_sentinel-2:_2012, hand_startup_2015}, allowing improved representation of temporal dynamics.
+For phenomena that can't be measured from space, either because they are not visible or because they require continuous observation, new approaches for \emph{in situ} or near-surface observations (e.g.
+low-cost wireless sensors \cite{wolf_gsm-based_2012, collins_new_2006, porter_wireless_2005}, citizen observers \cite{dickinson_current_2012}, and autonomous vehicles \cite{anderson_lightweight_2013}) can be used to increase the spatial and temporal coverage of observations.
 
-\item Represented scale is very different from actual scale of observations. 
-  \subitem In heeding Levin's call to think at multiple and appropriate ecological scales, we also need to give pause to think about whether our data are overextrapolated/overleveraged.
-  \subitem How over-leveraged/under-leveraged might depend on the auto-correlation properties of feature being observed. Maybe use example of fire process
-    \subitem Maybe we should go further here to explain why this is an important (and overlooked) assessment to make (actual versus effective comparison)
-\item Where to from here?  What do we need to improve?   
-  \subitem Satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.
+The aforementioned insights regarding modern observational domains must be tempered by the uncertainty within our own scale estimates, as detailed in the preceding section.
+However, most of this uncertainty is attributable to unclear reporting of scale values in the majority of papers we reviewed (a problem also noted in geography studies \cite{margulies_ambiguous_2016}).
+This tendency towards vague documentation offers one final insight, which is that, despite decades of accumulated knowledge regarding the importance of scale in ecology \cite{levin_problem_1992, wiens_spatial_1989, chave_problem_2013, wheatley_factors_2009}, scale appears to remain a low priority throughout much of the discipline.
+Beyond contributing to the broader problem of scientific reproducibility \cite{baker_1500_2016}, inattentiveness to scale increases the risk that observations inadequately represent the phenomenon of interest, thereby limiting the generalizability of any derived ecological knowledge \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}.
+To mitigate this problem, we recommend that ecological journals require authors to quantify and clearly report the values of resolution, extent, interval, and duration.
 
-\end{enumerate}
+\noindent \textbf{\emph{Looking forward}}\\
+Our study suggests that the concept of scale has yet to fully permeate the discipline of ecology.
+Evidence for this assertion lies in the continued narrowness of ecology's observational scale domains and the poor documentation of scale dimensions in the literature.
+However, the increasing extent of ecological observations, enabled by remote sensing and presumably motivated by many ecologists' appreciation of scale-related issues, suggests that ecology's scale domains are gradually changing.
+In the coming years, the accelerating gains in technology and analytical methods will allow researchers new and unprecedented capabilities to peer into, and thus close, the prominent holes in observational scale domains.
+A renewed, discipline-wide focus on scale's importance, including the adoption of stricter scale-reporting standards by journals, will help spur ecologists to address these gaps, while fostering the improved transferability of knowledge within the discipline.
 
@@
 %These results indicate that for most ecological observations are collected with spatially discontinuous replicates that sample very small portions of the area in which they are distributed.
-
-Narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) 
-Ecological understanding drawn from many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}, 
-a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}.
-
-Our results provide valuable insight into the spatial and temporal domains being addressed by modern ecological research.
-Our results show that most observations are collected at small spatial scales, are either unrepeated or relatively infrequent ($\geq$1 month interval), and in aggregate cover relatively narrow periods of time ($\leq$1 month).
-Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, indicating that, despite the well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989}.
-Furthermore, the unclear documenting of observational scales implies that scale is not a primary concern in much ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
-Taken together, this narrowness and poor documentation (a tendency that is also evident in the geographical sciences \cite{margulies_ambiguous_2016}) suggests that the ecological understanding drawn from many of these observations may have limited generalizability \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}, a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}.
-
-The generally small spatial scales of observation is a consequence of the continued dominance of field-based research, and the limited use of methods that allow larger areas to be comprehensively observed, such as remote sensing.
-Despite early and repeated calls for ecologists to use remote sensing because it provides a synoptic view that field measurements cannot \cite{turner_remote_2003, kerr_space_2003, pettorelli_satellite_2014}, and subsequent demonstration of its importance for multi-scale studies \cite{estes_habitat_2008, estes_predictive_2011}, our results indicate this method has not yet been widely adopted in ecological research.
-Two reasons lie behind this slow uptake.
-First, remote sensing can be a challenging method for ecologists to learn, many of whom may not have access to appropriate training \cite{pettorelli_satellite_2014}.
-Second, satellite observations typically have lower information content than field measurements for a given location, and in many cases only provide proxy measures for the ecological features of interest, such as forest understorey structure \cite{estes_remote_2010}, thereby making ecologists less inclined to use the technology \cite{turner_remote_2003}.
-
-In contrast to remote sensing, ecologists have made broader use of technologies that increase the temporal resolution of observations.
-Automated sensors were used to record 12\% of all observations, and accounted for most very high frequency measurements (intervals $\leq$1 hour).
-As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses (and many of the reviewed studies that used automated sensing aggregated the resulting observations before analyzing them), whereas longer-interval data (e.g.
+%
+%Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, indicating that, despite the well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989}.
+%
+%a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}. 
+%
+%As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses (and many of the reviewed studies that used automated sensing aggregated the resulting observations before analyzing them), whereas longer-interval data (e.g.
 annual biomass accumulation) cannot be disaggregated into shorter interval measurements (e.g.
@@
 
-In the coming years, rapid technological advances should increase the concentration of ecological observations in currently under-represented domains.
+%In the coming years, rapid technological advances should increase the concentration of ecological observations in currently under-represented domains.
 The growing numbers of high-resolution satellite sensors, together with new analytical platforms that provide free access to large volumes of pre-processed data and computational power \cite{googleearthengine}, will lower technical barriers that have so far prevented ecologists from adopting this observational technology  \cite{pettorelli_satellite_2014}.
@@
 A greater attentiveness to scale in general, including more meticulous documentation of observed dimensions, may help to facilitate the spread of ecological research to sparsely studied scales, while improving transferability of knowledge within the discipline.
-
-%Plot resolution issue: need high resolution over large extents, because this relies on remote sensing, and to really capture details, need high resolution.  
-%likely because of perceived (and in some cases real) logistical barriers \cite{pettorelli_satellite_2014}.
-
-%These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989}.
-
-%Despite the understanding that ecological patterns and process change with scale \cite{levin_problem_1992}, the narrow distribution of observations within spatial domains calls into question the degree to which ecological phenomena are being investigated at multiple scales.
-Adding to this concern is the fact the scales of observation were not clearly reported in the majority of studies; the spatial resolution and extent of studies was uncertain in 64\% and 70\% of cases, respectively, while the precise interval and study duration were uncertain in 43\% and 86\% of cases.
-Although sensitivity analyses show that our conclusions are robust to this uncertainty (SI), the unclear documenting of observational scales suggest that they are not a primary methodological concern in ecological research \cite{chave_problem_2013, wheatley_factors_2009}.
-%
-%On the other hand, we found a greater diversity of scales in the temporal domains of observations (Fig.
-1D).
-Over 70\% of sampled observations involved repeated assessments, and one-quarter of these were made with high frequency (intervals $\leq$1 day).
-Automated, surface-based sensors were used to record most of the very high frequency observations (intervals $\leq$1 hour), and comprised 9\% of all observations, suggesting that ecologists are adopting new technologies for collecting temporal data more rapidly than they are for spatial data.
-As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses, whereas it is extremely difficult, if not impossible, to disaggregate long-interval data.
-%
-%The temporal duration of most collected observations appears to be brief ($\leq$1year), which is due to the fact that we estimated the summed time of repeated observations, rather than the total time elapsed between the first and last measurements (the span of observation).
-In many cases, the actual observations are very brief, and thus the temporal duration is narrow.
-Snapshots in time may be sufficient to capture the temporal dynamics of many phenomena (e.g.
-changes in vegetation cover \cite{hansen_high-resolution_2013}), and for these our measurement of this domain may understate the temporal breadth of ecological observations.
-However, for more dynamic phenomena, long periods of continuous observation may be more important for understanding dynamics than frequent repeats.
-For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe behavior and evolution of the active fire \cite{clements_observing_2007}.
-For this reason, and because duration is the temporal analog of spatial extent, we chose to record observational duration rather then span.
-
-%Based on the observational domains recorded here, it appears that ecologists are still primarily collecting data at very fine spatial scales and over very narrow extents.
-Although then temporal duration of most ecological observations is also narrow, and a sizable number of observations are not repeated, greater effort is being devoted to capturing temporal dynamics.
-These findings suggest that ecological understanding is shaped by studies of relatively fine-scaled patterns, and because they are small in extent may be less generalizable \cite{wiens_spatial_1989}.
-
-%Based on this analysis, ecology thus appears to have
-
-%High frequency, high resolution sampling also rare--gap into which UAS is only just allowing to be filled. \\
-%Limitation for ocean studies, we did not study third dimension, volume (might be more relevant).
-
-%Although limited sample size, insight into the scales at which science of ecology is making bulk of observations
 

### 2017-06-22 Rmd/space-time.Rmd
Near final submission (JJC, PE, TT, JC edits)

@@
 length(unique(calr$DOI)) + length(unique(dat$DOI))  # 134 papers providing dat
-# unique(calr$DOI)
-
-# calr_type <- data.table(cbind("DOI" = unique(calr$DOI), # calib field studies
-#                               "study_type" = c(rep("field", 5), "paleo",
-#                                                rep("field", 4))))
 
@@
 # return interval
-tlab1 <- c(expression(paste(NULL<="second"^-1)), expression("minute"^-1),
-           expression("hour"^-1), expression("day"^-1), 
-           expression("week"^-1), expression("month"^-1),            
-           expression("year"^-1), expression("decade"^-1),
-           expression("century"^-1), 
-           expression("millenium"^-1), "unreplicated")# expression(infinity))
+tlab1 <- c("second", "minute", "hour", "day", "week", "month", "year", "decade",
+           "century", "millenium", "unreplicated")
 taxis1 <- cbind.data.frame( 
@@
 
-tlab2 <- c(expression(NULL<="second"), "minute", "hour", "day", "week", 
-           "month","year","decade", "century", "millenium", "10 KA")  
+tlab2 <- c("second", "minute", "hour", "day", "week", 
+           "month","year","decade", "century", "millennium", "10 KA")  
 taxis2 <- cbind.data.frame(c(1 / (24 * 60 * 60), 1 / (24 * 60), 1 / 24, 1, 7, 
@@
 
-datf[t_btwn_samp != 0, plot(log10(t_btwn_samp), log10(eff_ext))]
-datf[, plot(log10(plot_res), log10(eff_dur))]
-datf[, plot(log10(eff_dur), log10(eff_ext))]
-datf[t_btwn_samp == 0, points(log10(eff_dur), log10(eff_ext), col = "red")]
-datf[t_btwn_samp == 0 & log10(eff_dur) > 4]
-
-```
-
+```
 <a href="#top">Back to top</a>
+
+```{r, eval=FALSE, echo=FALSE}
+# check effect of averaging by DOI
+datf[, {
+ list("a" = log10(mean(plot_res)), "b" = log10(mean(eff_ext, na.rm = TRUE)))
+}, by = DOI][, list("a" = mean(a), "b" = mean(b, na.rm = TRUE))]
+datf[, list("a" = mean(log10(plot_res)), "b" = mean(log10(eff_ext), na.rm = T))]
+
+```

### 2017-06-22 paper/SI.tex
Near final submission (JJC, PE, TT, JC edits)

@@
 \section*{Supplementary Information}% Make the title.
+
+The following text provides additional information on this study's Methods and Results.
+The full manuscript, data, and code used to analyze the data are available on an accompanying github repository (https://github.com/ecohydro/ecoscales).
+\textbf{Note to editors and reviewers: this repository is currently private, but will be transitioned to a public repository upon acceptance of manuscript.}
+
 \subsection*{Methods}% Make the title.
@@
 Our study involved six different observers (those reviewing the papers to extract the observational scales), each of whom was given a randomly selected batch of 500 titles.
-A separate set of 20 papers was also randomly selected, and this set was given to all observers to review, in order to 1) calibrate the interpretations and extraction of scale-related information between observers, and 2) to estimate between-observer variance.
+A separate set of 20 papers was also randomly selected, and this set was given to all observers to review independently, in order to 1) calibrate the interpretations and extraction of scale-related information between observers, and 2) to estimate between-observer variance.
 
@@
 Sampling duration multiplied by the number of repeats observations is used to calculate study duration.
-Often this value will not be reported, so you will have to use your best judgement, based on your knowledge of ecological methods, to approximate the duration.
-For example, for a field based method with intensive plot methods, if you can't estimate a plausible duration, assign a token 1 day.
-For remote sensing observations you can assume one second (the observations are effectively instantaneous).
-   \Hitem \emph{\textbf{What is the study duration, and how do estimate it?}} The study duration (or, simply, duration) is the total period of time over which the phenomenon of interest was observed.
-More specifically, in the case of repeated observations, this is the total time that elapsed between the first and last observations at a given point in space (or of the same individual organism or community) were made.
+Often this value will not be reported, so you will have to use your best judgement, based on your knowledge of ecological methods, to approximate the sampling duration.
+For example, for a field based method with intensive plot methods, if you can't estimate a plausible sampling duration, assign a token 1 day.
+For remote sensing observations you can assume one second (although the observations are effectively instantaneous).
+   \Hitem \emph{\textbf{What is duration, and how do we estimate it?}} The duration is the total period of time over which the phenomenon of interest was observed.
+More specifically, in the case of repeated observations, this is the total time that elapsed between the first and last observations at a given point in space (or of the same individual organism or community).
 For once-off (unreplicated) observations, this time is equivalent to the sample duration.
-However, there may be cases where once-off observations may have a longer duration then the sample duration.
+However, there may be cases where once-off observations may have a longer duration than the sample duration.
 For example, consider a study that counts occurrences of pollinators over three years, using transects that are located in different locations within the broader study area during each year \cite{rollin_differences_2013-1}.
@@
 On the other hand, automated logging systems will provide a series of high frequency observations that are collected instantaneously.
-In these cases the sampling duration should be $\leq$second, and the interval should be the period between successive instantaneous measurements.
+In these cases the sampling duration should be a token one second (to keep consistent with remote sensing [Q7]), and the interval should be the period between successive instantaneous measurements.
     \Hitem \emph{\textbf{How do you treat interval for a case where repeated samples are taken during a season, across several seasons (e.g.``we performed repeat bird counts every 10 days between March and June of 2005, 2006, and 2008'')?}} Since the sampling is focused on seasons, and presumably some season-specific phenomenon (e.g.
@@
 
-  \Hitem \emph{\textbf{What is study extent, and how do I record it?}} Study extent (or, simply, extent) is defined as the total extent bounding all spatial replicates, divided by 10,000 to convert to hectares.
+  \Hitem \emph{\textbf{What is extent, and how do I record it?}} Extent is defined as the total area enclosed within a perimeter defined by the outermost spatial replicates, divided by 10,000 to convert to hectares.
 For studies in which spatial replicates are not spatially contiguous, this means the area of the minimum polygon bounding all spatial replicates.
-For example, if the study is conducted in a national park, the effective survey extent would be the area of the national park; if the study is conducted in three national parks, the effective survey extent would be the sum of the areas of all national parks.
-To calculate the effective survey extent, use the area of the study area/region given in the paper; when the area is not given, but when the survey region is given by name (e.g., Joshua Tree National Park or United States), look up the area through an online search and convert to hectares.
-When the area is not named, but a map is given, use an appropriate digitizing platform with a suitable map-providing backend (e.g.
-Google Earth Pro, QGIS with OpenLayers plugin) to navigate to the region and delineate a minimum convex polygon surrounding the plots/transects/sampling units to calculate the area in hectares.
-When studies focus on portions of survey regions that are clearly distinct from their surroundings (e.g., mangroves in a coastal National Park), try to delineate the focal portions (the mangroves) and not the larger survey region (entire National Park) using Google Earth Pro or a similar application with digitizing and area estimation capabilities.
+To calculate the effective survey extent, use the area of the study area/region given in the paper; when the area is not given, but when the survey region is given by name (e.g., Joshua Tree National Park or United States), look up the area through an online search and convert to hectares, provided the observations were collected throughout the named survey region.
+When the area is not named, but a map is given (or if the area is named and the map of replicates shows that the replicates were located in a smaller area of the named region), use an appropriate digitizing platform with a suitable map-providing backend (e.g.
+Google Earth Pro, QGIS with OpenLayers plugin) to navigate to the region and delineate a minimum convex polygon surrounding the replicates (plots/transects/other sampling units) to calculate the area in hectares.
+There is an exception to this rule, however, for studies that focus on features that are clearly distinct and functionally isolated from their surroundings.
+Examples of such cases might be mangrove forests patches within in a coastal National Park, or populations of a rare species confined to three disjunct protected areas.
+In these cases, the extent will be the summed area of the sampled units (the summed area of sampled mangrove stands, and the summed areas of the three protected areas).
+In other words, try to delineate the focal portions and not the larger survey region.
 
 For spatially contiguous studies (e.g.
-those based on satellite imagery), the extent is the total area covered by the imagery (in such cases, extent equals actual extent), but only records the area of imagery analyzed by the authors (e.g.
+those based on satellite imagery), the extent is the total area covered by the imagery (in such cases, extent equals actual extent), but only record the area of imagery analyzed by the authors (e.g.
 if the study area required four Landsat scenes to cover, but covered only the inner quarter of each image, report the extent as the summed area of the four quarters).
-However, if spatially contiguous studies only use a sub-sample of pixels, extent is the area of the minimum convex polygon bounding all pixels (calculated following the methods above).
+However, if spatially contiguous studies only use a sub-sample of pixels, extent is the area of the polygon enclosing the outermost pixels (calculated following the methods above).
 
@@
 
-  \Hitem \emph{\textbf{How do you determine sampling resolution for paleo-reconstructions and other approaches where a sampling method is presumed to draw from a larger area (e.g.
+  \Hitem \emph{\textbf{How do you determine resolution for paleo-reconstructions and other approaches where a sampling method is presumed to draw from a larger area (e.g.
 mammal traps, mist nets, etc)?}}  For sampling resolution, estimate the size of the sample actually taken, rather than the assumed catchment/shed area of the sample (e.g.
@@
 \vspace{-10pt}
-Two major sources of uncertainty affected our estimation of observational scales: 1) unclear documentation of observational scales in the reviewed studies, 2) variation between observers in estimating observational scales.
+Two major and related sources of uncertainty affected our estimation of observational scales: 1) unclear documentation of observational scales in the reviewed studies, 2) variation between observers in estimating observational scales (largely in cases where scales were not explicitly reported).
 We estimated and accounted for these uncertainties in several ways.
@@
 For each observation at each iteration, we perturbed its scale value in each dimension by randomly selecting (from a uniform distribution) a percentage value p that fell between 100 + y and 100 - y, where y was the between-observer coefficient of variation (expressed as a percentage) for a given observational dimension, and multiplying that observation's scale value by that proportion.
+In certain cases, the perturbation resulted in values for some certain dimensions that were not physically possible (e.g.
+interval longer or actual duration longer than duration; actual extent larger than extent), in which case we adjusted the perturbed values for the dimension that should be the smallest to equal that of the one that should be largest (e.g.
+we set interval to equal the perturbed duration value).
 This perturbed set of observations provided a basis for estimating uncertainty within our extracted set of observational scales.
@@
 \subsubsection*{Extracting and analyzing data from earlier meta-analyses}
-
-419 observational studies lasting $\leq100 years; longer studies were excluded because the upper duration bound was not published)
-
-reviewed in this study
+To compare the results of our analysis with the observational scales of earlier ecological studies, we used graph capture software\footnote{http://arohatgi.info/WebPlotDigitizer/app/} to extract the data values from Figure 6.1 in Tilman \cite{tilman_ecological_1989}, Figure 1 in Kareiva and Anderson \cite{kareiva_spatial_1988}, and Figure 2 in Porter et al \cite{porter_crop_2005}.
+To maintain as much comparability as possible with our inclusion criteria, we excluded experiment studies in Tilman's data, as well as the values of any studies having duration values greater than 100 years (no upper time bound was provided for these), leaving duration values for 419 (out of 623) studies.
+As Tilman presented duration values as a histogram, we calculated the mean duration across all studies as the weighted (by number of observations per bin) mean of bin center-point values (i.e.
+the weighted mean of the bin means).
+We also excluded four (of 29) observation values from Porter et al's data on observational extent and frequency, which, in contrast to the other 25, were not randomly selected.
+Porter et al also used irregular scales for both their X (frequency) and Y (extent) axis, therefore we had to visually estimate the scale values for each data point after graphical extraction, and converted their extent values (in km) to hectares and their frequency values to intervals.
+Kareiva and Anderson presented resolution as plot diameters (m), which we squared to make comparable to our resolution metric.
 
@@
 
-Second, we calculated the intra-class correlation coefficient \cite{bartko_intraclass_1966} to assess agreement between observers regarding the number of ecological observations that could be extracted from each paper (multiple ecological observations were reported in many studies, and we listed observations as separate records if they varied on one or more dimensions).
+Second, we calculated the intra-class correlation coefficient \cite{bartko_intraclass_1966} to assess agreement between observers regarding the number of ecological observations that could be extracted from each paper (multiple ecological observations were reported in many studies; we listed observations as separate records if they measured different features, or more rarely, varied substantially on one or more dimensions).
 The coefficient, calculated using the IRR package \cite{gamer_irr:_2012} of R, was 0.71 (F = 15.4, p$<$0.0001, 95\% confidence interval = 0.54 - 0.85)
@@
 \caption{Trends in observational scales by year of publication.
-The coefficient of a weighted (by number of studies in each year) linear regression, fit to the logarithm (base 10) of the mean scale values (calculated for each publication year) for the six assessed dimensions ispresented at the top of each plot, as well as the model coefficient of determination and p-value.
+The coefficient of a weighted (by number of studies in each year) linear regression, fit to the logarithm (base 10) of the mean scale values (calculated for each publication year) for the six assessed dimensions is presented at the top of each plot, as well as the model coefficient of determination and p-value.
 }
@@
 With the smaller bandwidth this concentration appears as two separate patches (Fig.
-S2 top left), but with 0.7 bandwidth applied becomes coherent (Fig.
+S3 top left), but with 0.7 bandwidth applied becomes coherent (Fig.
 S3 middle left).
 This increasing separation of points into separate clusters is expected as bandwidth is reduced.
+However, this assessment indicates that the choice of bandwidth does not alter the interpretation of the results, as the domains of highest and lowest concentrations remain the same.
 

### 2017-06-22 paper/main.tex
Near final submission (JJC, PE, TT, JC edits)

@@
 \textbf{To properly understand ecological phenomena, it is necessary to observe them across a range of spatial and temporal scales.
-Ecologists first raised this point in the 1980s, since when the ability to collect multi-scale observations has grown rapidly.
+Ecologists first raised this point in the 1980s, and since then the ability to collect multi-scale observations has grown rapidly.
 To assess modern ecology's progress in addressing scale, we analyzed the resolution, extent, interval, and duration of observations in 348 studies published between 2004-2014.
-We found that the scale domains of observations remain fairly narrow, and are collected primarily with conventional field techniques.
+We found that the scale domains of observations are fairly narrow, and are collected primarily with conventional field techniques.
+In the spatial domain, most observations have resolutions of $\leq$1 m2 and extents of $\leq$10,000 ha.
+In the temporal domain, most observations were either unreplicated or of low frequency ($\geq$1 month interval), and were made over relatively short durations ($\leq$1 year).
 Compared to prior meta-analyses from the 1980s and early 2000s, observational durations and resolutions remain largely unchanged, but intervals have become finer and extents larger.
-Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, raising questions about observational representativeness.
-Adding to this concern, observational scales were not clearly reported in most studies, suggesting that scale often is a minor consideration.
-Journals can help mitigate this problem by adopting scale reporting standards, which can further spur ecologists to adopt rapidly improving observational technologies, and thereby close key gaps in current observational domains.}
+Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, raising concerns about observational representativeness.
+Adding to these concerns, scales were not clearly reported in most studies, suggesting that it is a minor consideration.
+Journals can help mitigate this problem by implementing scale reporting standards, which can spur ecologists to more rapidly adopt new observational technologies, and thereby close key gaps in current observational domains.}
 
@@
 
-Our analysis focused on two dimensions of spatial scale, resolution (grain) and extent, and two of temporal scale, interval and duration (Table 1).
+Our analysis focused on two dimensions of spatial scale, resolution (grain) and extent, and two of temporal scale, interval and duration (Table 1, and see SI for full definitions).
 Resolution is the area of an individual spatial replicate within which a complete measurement (as opposed to a sub-sample) of the feature of interest was made.
 Extent is the area enclosed by the outer-most spatial replicates, or, if the system or habitat being sampled was distinct from its surrounding matrix (e.g.
-forest patches in grassland habitats), the summed area of sampled patches (see SI for full definition).
+forest patches in grassland habitats), the summed area of sampled patches.
 Interval refers to the average time elapsed between individual temporal replicates.
-Duration measures the time elapsed between the first and last temporal replicates, or, in the case of temporally unreplicated observations, the estimated time spent collecting the observation (SI).
+Duration measures the time elapsed between the first and last temporal replicates, or, in the case of temporally unreplicated observations, the estimated time spent collecting the observation.
 We also assessed observational scales within two additional dimensions, \emph{actual} extent (the integrated area of spatial replicates) and \emph{actual} duration (the summed observational time of temporal replicates).
@@
 1A).
-The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 11\% 1-10,000 ha, 19\% 10,000-100,000 ha, 27\% 100,000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig.
+The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 11\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig.
 1B).
@@
 The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha.
-These three concentrations suggest a tendency for observational extent to increase with resolution, which is a relationship that becomes more pronounced in the (less densely observed) portion of the domain where resolutions $\geq$100 m$^2$.
+These three concentrations suggest a tendency for observational extent to increase with resolution, which is a relationship that becomes more pronounced in the less densely observed portion of the domain where resolutions $\geq$100 m$^2$.
 
-The majority of temporally replicated observations have daily to decadal intervals and durations of $\geq$ 1 month to 1 decade.
+Most temporally replicated observations have daily to decadal intervals and durations of $\geq$ 1 month to 1 decade.
 The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting one year to one decade tend to have yearly to decadal intervals.
@@
 3).
-This comparison showed that a majority (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).
-Actual extent converged with extent $\geq$1 100,000 ha, but this applies to just 5\% of observations.
+This comparison showed that most (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).
+Actual extent converged with extent $\geq$100,000 ha, but this applies to just 5\% of observations.
 The actual duration of 64\% of observations was $<$ 1 day, which on average was 3 to nearly 9 orders of magnitude shorter than the time span covered by temporal replicates (Fig.
@@
 There were several potential methodological aspects that could have influenced our assessment of ecology's spatial and temporal domains.
-The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate--rather than simply record--these values for most observations (specifically, in 63, 60, and 69\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration, respectively).
+The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate, rather than simply record, these values for most observations (specifically, in 63\%, 60\%, and 69\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration, respectively).
 The inevitable estimation errors may have biased our overall findings.
-However, we attempted to quantify this error by assessing inter-rater disagreement and incorporating this uncertainty into our resampling methodology.
+However, we attempted to quantify and account for this error by assessing inter-rater disagreement and incorporating this uncertainty into our resampling methodology.
 The resulting confidence intervals (Fig.
@@
 Our results suggest that the scale domains of modern ecological observations are fairly narrow, and are collected primarily with conventional field techniques.
-Spatially, the majority of observations have grains of $\leq$1m$^2$ and extents of $\leq$10,000 ha (Fig 1A;B).
+Spatially, the majority of observations have grains of $\leq$1 m$^2$ and extents of $\leq$10,000 ha (Fig 1A;B).
 In the temporal domains, most observations are either un-replicated snapshots, or of low frequency ($\geq$1 month interval; Fig.
@@
 Have these domains changed since the seminal papers on scale first began to appear in the late 1980s \cite{wiens_spatial_1989, levin_problem_1992, tilman_ecological_1989}?
-A comprehensive answer to this question is not possible, absent a similar study targeting earlier literature, but an analysis of results presented in three prior studies provides partial insight.
+A comprehensive answer to this question would require a similar study focused on earlier literature, but an analysis of results presented in three prior studies provides partial insight.
 The first and most comprehensive dataset consists of duration values extracted by Tilman \cite{tilman_ecological_1989} from 623 studies published between 1977-1987 in the journal \emph{Ecology}.
@@
 The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who present the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986.
-The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,510,247 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (197 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
+The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,510,247 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
 The third dataset is provided by Porter et al \cite{porter_wireless_2005}, who compared the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}).
-The mean interval was 178 days, compared to 684 in our sample, but the 80th percentile value in our study was 169 compared to 329 in the earlier study.
+The mean interval was 178 days, compared to 684 days in our sample, but the 80th percentile value in our study was 169 days compared to 329 days in theirs.
 Extent in our sample was substantially larger according to multiple summary statistics, including the mean (115,098,283 ha in our study versus 368,403 ha), median (5,120 ha versus 9 ha), and 95th percentile (45,104,594 ha versus 136,000 ha).
@@
 Although limited due to methodological differences (e.g.
-a focus on experiments versus un-manipulated systems; SI), these comparisons suggest that the duration and, less clearly, resolution of ecological observations have changed little in the past 30 years, but observational frequency and particularly extent have increased.
-The growth in observational extent is also reflected by the positive year-on-year trend within our own dataset, which corresponds to the increasing use of remote sensing (SI).
+a focus on experiments versus un-manipulated systems), these comparisons suggest that the duration and, less clearly, resolution of ecological observations have changed little in the past 30 years, but observational frequency and particularly extent have increased.
+The growth in observational extent is also reflected by the positive year-on-year trend within our own dataset, which corresponds to the increasing use of remote sensing (see figures in SI).
 
@@
 For example, remote sensing provides wall-to-wall spatial coverage of a study area, erasing the difference between actual extent and extent.
-Similarly, the interval of high-resolution (a desirable image property because it allows individual ecological features to be more effectively discerned  \cite{dark_modifiable_2007, hay_comparison_2003}) imaging is now approaching daily to sub-daily scales \cite{drusch_sentinel-2:_2012, hand_startup_2015}, allowing improved representation of temporal dynamics.
-For phenomena that can't be measured from space, either because they are not visible or because they require continuous observation, new approaches for \emph{in situ} or near-surface observations (e.g.
+Furthermore, the interval of high-resolution imaging (higher resolution is preferred in images as it allows individual features to be better discerned \cite{dark_modifiable_2007, hay_comparison_2003}) is now approaching daily to sub-daily scales \cite{drusch_sentinel-2:_2012, hand_startup_2015}, allowing improved representation of spatial and temporal dynamics.
+For phenomena that can't be measured from space, either because they are not visible or because they require continuous observation, new approaches for collecting \emph{in situ} or near-surface observations (e.g.
 low-cost wireless sensors \cite{wolf_gsm-based_2012, collins_new_2006, porter_wireless_2005}, citizen observers \cite{dickinson_current_2012}, and autonomous vehicles \cite{anderson_lightweight_2013}) can be used to increase the spatial and temporal coverage of observations.

### 2017-06-28 paper/main.tex
Final re-submission version of manuscript. Revised package structure so that input data will install as part of R package.  Corrected mistake in one perturbation parameter (for extent, which had uncertainty value slightly too low), re-rain analyses, updated figures in text (no substantive change).

@@
 1A).
-The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 11\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 16\% $>$1,000,000 ha (Fig.
+The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 12\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig.
 1B).
@@
 In the temporal dimensions, 37\% of observations were not repeated (Fig.
-1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, and 8\% at yearly to decadal intervals.
+1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals.
 Duration was one day or less for 31\% of sampled observations, while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 9\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig.
@@
 
-To provide further insight into observational domains, we also evaluated the degree to which the scales \emph{represented} by observations differed from those that were actually observed.
-To make this assessment, we first log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, in order to calculate the magnitude of difference (or decade) between each pair of dimensions for each observation.
-We then compared how these magnitudes varied in relation to the scale of the actual dimension (Fig.
+For our final analysis of observational domains, we evaluated the differences between the scales represented by the extent and duration dimensions and the scales that ecological observations actually cover.
+To make this assessment, we log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, yielding the magnitude of difference between each pair of dimensions for each observation.
+We plotted these values (summarized in box plots) against their corresponding actual extent/duration values to evaluate whether these differences vary with scale (Fig.
 3).
-This comparison showed that most (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the quantified extent (Fig 3.A).
-Actual extent converged with extent $\geq$100,000 ha, but this applies to just 5\% of observations.
-The actual duration of 64\% of observations was $<$ 1 day, which on average was 3 to nearly 9 orders of magnitude shorter than the time span covered by temporal replicates (Fig.
+This analysis showed that most (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the primary measure of extent (Fig 3A).
+These differences declined linearily with actual extent, becoming approximately equal for the 5\% of observations having $\geq$100,000 ha actual extent.
+The actual duration of 64\% of observations was $<$1 day, which on average was 3 to nearly 9 orders of magnitude shorter than the time span covered by the conventional metric of duration (Fig.
 3B).
-The two duration measures were approximately equal for the 16\% of observations exceeding one month of actual duration.
+As with extent, the magnitude of difference declined steadily with actual duration, effectively disappearing for the 16\% of observations with $\geq$1 month of actual duration.
 
@@
 \vspace{-0.2 cm}
-\caption{The difference between \emph{actual} extent (the summed area of spatial replicates) and extent (A) and \emph{actual} duration (the summed sampling duration across temporal replicates) and duration (B).
-Differences are expressed as decades, or how many orders of magnitude greater the extent or duration is then the actual extent or duration, and are summarized (as box plots, with circle in box representing the mean and line the median) in bins representing different levels of actual extent/duration.
+\caption{The difference between extent and \emph{actual} extent (the summed area of spatial replicates) (A) and duration and \emph{actual} duration (the summed sampling duration across temporal replicates) (B).
+Difference values are expressed in terms of how many orders of magnitude larger (or longer) extent (duration) is than actual extent (actual duration), and are summarized (as box plots, with circle in box representing the mean and line the median) in bins representing increasing scales of actual extent/duration.
 The percentages of observations falling within each bin are indicated by the color of the inter-quartile and the numeric value above the upper whisker.}
@@
 The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who present the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986.
-The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,510,247 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
+The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,496,069 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
 The third dataset is provided by Porter et al \cite{porter_wireless_2005}, who compared the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}).
 The mean interval was 178 days, compared to 684 days in our sample, but the 80th percentile value in our study was 169 days compared to 329 days in theirs.
-Extent in our sample was substantially larger according to multiple summary statistics, including the mean (115,098,283 ha in our study versus 368,403 ha), median (5,120 ha versus 9 ha), and 95th percentile (45,104,594 ha versus 136,000 ha).
+Extent in our sample was substantially larger according to multiple summary statistics, including the mean (114,965,072 ha in our study versus 368,403 ha), median (5,051 ha versus 9 ha), and 95th percentile (46,424,808 ha versus 136,000 ha).
 
@@
 
-It is therefore important to examine whether the phenomena being observed are adequately captured by the design of replicates.
-Our methods suggest one possible procedure for assessing the representativeness of replicates: 1) measure auto-correlation (spatial or serial) in the replicates, 2) add the autocorrelation length to the replicate area/duration, 3) calculate an autocorrelation-adjusted actual extent/duration, and 4) plot where it falls between actual extent/duration and extent/duration.
-If the adjusted actual value is substantially closer to the representative value, then this may provide greater confidence in the adequacy of the sampling scheme.
-If not, then alternative sampling methods may be used to close this gap.
+It is therefore important to examine whether the scales of the phenomena being observed are adequately captured by the design of replicates.
+Our methods suggest one possible procedure for assessing the \emph{scale representativeness} of replicates: 1) measure auto-correlation (spatial or serial) in the replicates, 2) add the autocorrelation length to the replicate area/duration, 3) calculate an autocorrelation-adjusted actual extent/duration, and 4) plot where it falls between actual extent/duration and extent/duration.
+The distance between the adjusted actual value and the ostensible value can provide a measure of how well the replicates represent the intended scale of observation.
+If the gap remains large, then alternative sampling methods may be used to close it.
 For example, remote sensing provides wall-to-wall spatial coverage of a study area, erasing the difference between actual extent and extent.

### 2017-06-28 paper/main.tex
Small text tweak to line 108-9 of final submitted version

@@
 3B).
-As with extent, the magnitude of difference declined steadily with actual duration, effectively disappearing for the 16\% of observations with $\geq$1 month of actual duration.
+As with extent, the average magnitude of difference declined steadily with actual duration, becoming effectively equal for the 16\% of observations with $\geq$1 month of actual duration.
 

### 2017-06-28 paper/SI.tex
Final submitted version (amended from previous to revise institutional affiliation and add links to code and data for reviewers)

@@
 \usepackage{multirow}
+\usepackage[
+  breaklinks=true,
+  colorlinks=true,
+  linkcolor=blue,anchorcolor=blue,
+  citecolor=blue,filecolor=blue,
+  menucolor=blue,pagecolor=blue,
+  urlcolor=blue]{hyperref}
 
@@
 The following text provides additional information on this study's Methods and Results.
-The full manuscript, data, and code used to analyze the data are available on an accompanying github repository (https://github.com/ecohydro/ecoscales).
-\textbf{Note to editors and reviewers: this repository is currently private, but will be transitioned to a public repository upon acceptance of manuscript.}
+The full manuscript, data, and code are available on an accompanying \href{https://github.com/ecohydro/ecoscales}{github repository}.
+\textbf{Note to editors and reviewers: this repository is currently private, but will be made public upon acceptance of manuscript.
+In the interim, the code and data can be downloaded from the following links}:
+\href{https://www.dropbox.com/s/4r9h8mn1wa4kjsn/calibration.html?dl=0}{Calibration analysis}; \href{https://www.dropbox.com/s/n0zq5zufu96ktfl/calibration-set.xlsx?dl=0
+}{Calibration dataset}; \href{https://www.dropbox.com/s/bzni8dhqy8un4fz/space-time.html?dl=0
+}{Main analysis}; \href{https://www.dropbox.com/s/vujjo2x4f5kwqkb/main-results.xlsx?dl=0
+}{Main dataset}
 

### 2017-06-28 paper/main.tex
Final submitted version (amended from previous to revise institutional affiliation and add links to code and data for reviewers)

@@
 \author
-{Lyndon Estes$^{\ast1, 2, 3}$, Paul R. Elsen$^{1, 4}$, Tim Treuer$^{5}$, Labeeb Ahmed$^{6}$, \\
-Kelly Caylor$^{1, 7}$, Jason Chang$^{6}$, Jonathan J. Choi$^{5}$, and Erle Ellis$^{6}$ \\
+{Lyndon Estes$^{\ast1, 2, 3}$, Paul R. Elsen$^{2, 4}$, Tim Treuer$^{5}$, Labeeb Ahmed$^{6}$, \\
+Kelly Caylor$^{3, 7}$, Jason Chang$^{6}$, Jonathan J. Choi$^{5}$, and Erle Ellis$^{6}$ \\
 \\
-\normalsize{$^{1}$Woodrow Wilson School, Princeton University, Princeton, NJ 08544, USA}\\
-\normalsize{$^{2}$Civil and Environmental Engineering, Princeton University, Princeton, NJ 08544, USA}\\
-\normalsize{$^{3}$Graduate School of Geography, Clark University, Worcester MA 01610, USA}\\
+\normalsize{$^{1}$Graduate School of Geography, Clark University, Worcester MA 01610, USA}\\
+\normalsize{$^{2}$Woodrow Wilson School, Princeton University, Princeton, NJ 08544, USA}\\
+\normalsize{$^{3}$Civil and Environmental Engineering, Princeton University, Princeton, NJ 08544, USA}\\
 \normalsize{$^{4}$Department of Environmental Science, Policy, and Management,}\\

### 2017-11-19 R/data.R
Added new vignette for figure 1 in main analysis. Moved resamp_func and to_infinity_beyond functions to package functions. Added documentation for lazy load datasets

@@ -1,0 +1,34 @@
+#' Combined scale results
+#' @name datf
+#' @docType data
+#' @description Cleaned and combined calibration and main scale datasets
+#' @author Estes, Treuer, Elsen, Chang, Choi, Ahmed, Caylor, Ellis
+#' @keywords data
+#' @examples
+#' data(datf)
+NULL
+
+#' Combined scale results resampled with uncertainty
+#' @name bootperturb
+#' @docType data
+#' @description Results of combined calibration and main scale datasets, 
+#' resampled 1000 times with random error introduced within bounds of 
+#' uncertainty determined from analysis of calibration dataset
+#' @author Estes, Treuer, Elsen, Chang, Choi, Ahmed, Caylor, Ellis
+#' @keywords data
+#' @examples
+#' data(bootperturb)
+NULL
+
+#' Combined scale results
+#' @name dimbreaks
+#' @docType data
+#' @description Contains multiple look-up tables for plotting and labelling 
+#' figures in log10 spaces (aaxis1, aaxis2, taxis1, taxis2, alab1, alab2, tlab1, 
+#' tlab2) 
+#' @author Estes, Treuer, Elsen, Chang, Choi, Ahmed, Caylor, Ellis
+#' @keywords data
+#' @examples
+#' data(datf)
+NULL
+

### 2017-11-19 R/resamp_func.R

@@ -1,0 +1,44 @@
+#' Resampling with perturbation
+#' @description Randomly perturbs scale values within specified uncertainty 
+#' bounds
+#' @param dat data.table of combined calibration and main scale results 
+#' @param p Vector of uncertainty bounds, one per assessed dimension
+#' @param iter Number of resampling iterations
+#' @return List of data.table equal to length iter
+#' @example 
+#' data(datf)
+#' set.seed(1)
+#' bootperturb <- resamp_func(datf, p = p, iter = 1000)
+#' @export
+resamp_func <- function(dat, p, iter) {
+  vars <- c("plot_res", "act_ext", "eff_ext", "t_btwn_samp", "act_dur", 
+            "eff_dur")
+  bout <- lapply(1:iter, function(x) {  # x <- 1
+    if((x / 100) %in% 1:100) print(x)
+    if(length(p) < length(var)) stop("p must be length of var", call. = FALSE)
+    dnew <- copy(dat[, vars, with = FALSE])
+    # modifiers
+    modvec <- sapply(p, function(y) {
+      mv <- runif(nrow(dnew), min = 1 - y, max = 1 + y)
+      mv[mv < 0] <- 0.00000001  # can't have reduction of more than 100%
+      mv
+    })
+    
+    # perturb and set any values < 0 to very small amount
+    for(j in 1:ncol(modvec)) set(dnew, j = j, value = dnew[[j]] * modvec[, j])
+    
+    # correct physically impossible results 
+    dnew[, plot_res := plot_res / 10000]  # set plot_res to ha
+    dnew[plot_res > act_ext, plot_res := act_ext]  # plot_res can't be > act_ext
+    dnew[plot_res > eff_ext, plot_res := eff_ext]  # or eff_ext
+    dnew[act_ext > eff_ext, act_ext := eff_ext] # act_ext can't be > eff_ext
+    dnew[t_btwn_samp > eff_dur, t_btwn_samp := eff_dur]  # int can't be > eff_dur
+    dnew[act_dur > eff_dur, act_dur := eff_dur]  # act_dur can't be > eff_dur
+    dnew[, plot_res := plot_res * 10000]  # set plot_res back to m2
+    
+    # set unreplicated to high value for first figure 
+    dnew[, t_btwn_samp := to_infin_byond(t_btwn_samp)]  # set high
+    cbind("type" = dat[, study_type], dnew)
+  })
+  return(bout)
+}

### 2017-11-19 R/utilities.R

@@ -1,0 +1,5 @@
+#' Sets interal of non-replicated samples to 365 * 10000 days
+#' @param tbtwn Vector of interval values
+#' @return Vector
+#' @export
+to_infin_byond <- function(tbtwn) ifelse(tbtwn == 0, 365 * 10000, tbtwn)

### 2017-11-20 R/resamp_func.R
Instituted checks at bottom of figure 1 vignette, and implemented caching of histogram calculations on bootperturb

@@ -5,7 +5,7 @@
 #' @param p Vector of uncertainty bounds, one per assessed dimension
 #' @param iter Number of resampling iterations
 #' @return List of data.table equal to length iter
-#' @example 
+#' @examples 
 #' data(datf)
 #' set.seed(1)
 #' bootperturb <- resamp_func(datf, p = p, iter = 1000)

### 2017-11-21 R/kdat_setup.R
Created new vignette for kde analysis and resulting Figure 2. Created functions and added much of the repeated code into lapply's

@@ -1,0 +1,29 @@
+#' Prepares scale data for kernel density estimates
+#' @description Processes two variables from input scale data.table for KDE
+#' @param dat Input data.table (an rbind of bootperturb)
+#' @param x Name of variable in dat that will form x axis 
+#' @param x Name of variable in dat that will form y axis 
+#' @param xlim Two element vector setting limits for x axis
+#' @param ylim Two element vector setting limits for y axis
+#' @param drop.oneoff Should unrepeated observations be dropped (default=FALSE)
+#' @param oneoff.var Axis containing unrepeated observations ("x" or "y")
+#' @return SpatialPointsDataFrame
+#' @examples 
+#' odt <- kdat_setup(kdat, x = "plot_res", y = "t_btwn_samp", xlim = c(-5, 8), 
+#'                   ylim = c(-5, 5.57), drop.oneoff = TRUE)  
+#' @export
+kdat_setup <- function(dat, x, y, xlim, ylim, drop.oneoff = FALSE, 
+                       oneoff.var = "y") {
+  odat <- dat[, list("x" = log10(get(x)), "y" = log10(get(y)))]
+  if(drop.oneoff == TRUE) {
+    odat <- odat[round(get(oneoff.var), 8) != 6.56229286]  # remove one-time obs
+  }
+  odat <- na.omit(odat)   # drop NA values
+  odat[, n := 1]  # set up count column
+  odat[x < xlim[1], x := xlim[1]]  # confine x to min
+  odat[x > xlim[2], x := xlim[2]]  # confine x to max
+  odat[y < ylim[1], y := ylim[1]]  # confine y to min
+  odat[y > ylim[2], y := ylim[2]]  # confine y to max
+  coordinates(odat) <- ~x + y  # to spdf
+  odat
+}

### 2017-11-21 R/kdensity.R

@@ -8,6 +8,10 @@
 #' @param pts SpatialPointsDataFrame showing distribution of observations
 #' @param kwidth Kernel smoothing width 
 #' @return Raster of density values
+#' @examples 
+#' kder <- kdensity(xmn = -5, xmx = 5.56, ymn = -5, ymx = 5.57, dxdy = 0.1, 
+#'                  pts = odt, kwidth = 1)
+#' kder <- (kder / cellStats(kder, sum)) * 100  # convert to percentage
 #' @export
 kdensity <- function(xmn, xmx, ymn, ymx, dxdy, pts, kwidth) {
   

### 2017-11-21 R/resamp_func.R

@@ -32,7 +32,7 @@ resamp_func <- function(dat, p, iter) {
     dnew[plot_res > act_ext, plot_res := act_ext]  # plot_res can't be > act_ext
     dnew[plot_res > eff_ext, plot_res := eff_ext]  # or eff_ext
     dnew[act_ext > eff_ext, act_ext := eff_ext] # act_ext can't be > eff_ext
-    dnew[t_btwn_samp > eff_dur, t_btwn_samp := eff_dur]  # int can't be > eff_dur
+    dnew[t_btwn_samp > eff_dur, t_btwn_samp := eff_dur] # int can't be > eff_dur
     dnew[act_dur > eff_dur, act_dur := eff_dur]  # act_dur can't be > eff_dur
     dnew[, plot_res := plot_res * 10000]  # set plot_res back to m2
     

### 2017-11-24 R/kdat_setup.R
Added initial calculations for mean and median of different forms of data collection, tested, then altered--added in mean/median calculation to kdat_setup.

@@ -14,16 +14,21 @@
 #' @export
 kdat_setup <- function(dat, x, y, xlim, ylim, drop.oneoff = FALSE, 
                        oneoff.var = "y") {
-  odat <- dat[, list("x" = log10(get(x)), "y" = log10(get(y)))]
+  odat <- dat[, list("x" = get(x), "y" = get(y))]
   if(drop.oneoff == TRUE) {
-    odat <- odat[round(get(oneoff.var), 8) != 6.56229286]  # remove one-time obs
+    odat <- odat[get(oneoff.var) != 365 * 10000]  # remove one-time obs
   }
   odat <- na.omit(odat)   # drop NA values
-  odat[, n := 1]  # set up count column
-  odat[x < xlim[1], x := xlim[1]]  # confine x to min
-  odat[x > xlim[2], x := xlim[2]]  # confine x to max
-  odat[y < ylim[1], y := ylim[1]]  # confine y to min
-  odat[y > ylim[2], y := ylim[2]]  # confine y to max
-  coordinates(odat) <- ~x + y  # to spdf
-  odat
+
+  mu <- odat[, lapply(.SD, mean)][, lapply(.SD, log10)]  # mean
+  med <- odat[, lapply(.SD, median)][, lapply(.SD, log10)]  # median
+  o <- odat[, lapply(.SD, log10)]  # convert to log10
+
+  o[, n := 1]  # set up count column
+  o[x < xlim[1], x := xlim[1]]  # confine x to min
+  o[x > xlim[2], x := xlim[2]]  # confine x to max
+  o[y < ylim[1], y := ylim[1]]  # confine y to min
+  o[y > ylim[2], y := ylim[2]]  # confine y to max
+  coordinates(o) <- ~x + y  # to spdf
+  list("dat" = o, "mu" = mu, "med" = med)
 }

### 2017-11-24 R/kdensity.R
Updated main kde and stats calcs to reflect new changes, updated figure 2, added trim function to kdensity function because of raster NAs along some margins.

@@ -36,8 +36,7 @@ kdensity <- function(xmn, xmx, ymn, ymx, dxdy, pts, kwidth) {
   kern <- spkernel2d(pts, crds, h0 = kwidth, grd)
   kgrid <- as(SpatialGridDataFrame(grd, data = data.frame(kern)),
               "SpatialPixelsDataFrame")
-
-  krast <- raster(kgrid)
+  krast <- trim(raster(kgrid))
   return(krast)
 }
 

### 2017-11-27 R/box_plots.R
Started on vignette for figure 3. Added box_stats functions from lmisc to package to prevent dependence on lmisc

@@ -1,0 +1,145 @@
+#' Calculates boxplot statistics
+#' @description Calculates the usual boxplot statistics plus mean
+#' @param x Vector of values from which to calculate statistics
+#' @param whiskers A two-element vector of the percentiles for the lower and upper whiskers (c(0.025, 0.975))
+#' @param weighted TRUE or FALSE (default) for weighted quantiles and mean
+#' @param weight.opts A list of options to be passed to wtd.quantile
+#' @param na.rm Should the quantile and mean functions exclude NA values in x? Default is TRUE, else FALSE
+#' @return A vector containing the 2.5, 25, 50, 75, 97.5th percentile values and the mean
+#' @note Function transplanted from lmisc package
+#' @details If weighted is TRUE, the wtd.quantile and wtd.mean functions from Hmisc are called, and a weights
+#' vector equal in length to x is needed, passed as the minimum necessary parameter for calculating weighted 
+#' quantiles. A named list should be provided where the name for each element matches the argument name for 
+#' Hmisc::wtd.quantile and Hmisd::wtd.mean. Other possible arguments include type and normwt.
+#' @examples 
+#' set.seed(234)
+#' x <- rnorm(20)
+#' w <- sample(1:10, size = 20, replace = TRUE)
+#' box_stats(x)
+#' box_stats(x, weighted = TRUE, weight.opts = list("weights" = w))
+#' @export
+box_stats <- function(x, whiskers = c(0.025, 0.975), weighted = FALSE, weight.opts = NULL, na.rm = TRUE) {
+  p <- c(whiskers[1], 0.25, 0.5, 0.75, whiskers[2])
+  if(weighted == FALSE) {
+    bs <- c(quantile(x, p, na.rm = na.rm), "mu" = mean(x, na.rm = na.rm))
+  } else if(weighted == TRUE) {
+      if(is.null(weight.opts)) stop("Parameters for weights are needed, and optionallym type, and normwt")
+      w <- weight.opts
+      if(!"weights" %in% names(w)) stop("You are missing, or have incorrectly named, the weights vector")
+      if(!"type" %in% names(w)) w$type <- "quantile"
[…]

### 2017-11-27 R/rect_shade.R
Added rect_shade function from lmisc to package

@@ -1,0 +1,46 @@
+#' Create adjacent rectangular polygons
+#' @description Generic function to create adjacent rectangular polygons from vectors of coordinates passed to 
+#' it
+#' @param x Vector of x coordinates, arranged in ascending order
+#' @param y Vector of y coordinates, arrange in acending order
+#' @param fillcol A 1 to n - 1 length vector of colors for filling. n = length of x or y (the longest)
+#' @param linecol A 1 to n - 1 length vector of polygon border colors. n = length of x or y (the longest)
+#' @return A polygon "chain": segmented polygons with interior vertical x or horizontal y axis
+#' @note Function transplanted from lmisc package
+#' @details This is modeled on the rect_coords function used by flex_legend, but the resulting polygon 
+#' can be of arbitrary shape. It can be used to make either a horizontal (i.e. a sequence of polygons
+#' separated on the X-axis, with the long dimension filling the Y-axis) or vertical chain of polygons (the
+#' opposite). To make the former, provide a vector giving the X-coordinates separating each polygon, and a 2 
+#' parameter vector specifying the lower and upper Y coordinates. For the opposite case, give the left and 
+#' right X coordinates to X, and a vector of Y coordinates to Y. 
+#' @examples
+#' a <- c(-50, -25, 0, 25, 50)
+#' b <- c(-1, 21)
+#' plot(c(-100, 100), c(-1, 22), pch = "")
+#' rect_shade(b, a)
+#' 
+#' plot(c(-100, 100), c(-1, 22), pch = "")
+#' rect_shade(a, b)
+#' 
+#' cols <- sapply(c(0.8, 0.5, 0.2, 0.1), function(x) rgb(1, 0, 0, alpha = x))
+#' plot(c(-1, 22), c(-100, 100), pch = "")
+#' # rect_shade(b, a, fillcol = cols)
+#' rect_shade(b, a, fillcol = cols, linecol = cols)
+#' @export
+
+rect_shade <- function(x, y, fillcol = "grey", linecol = "black", lwd = 1) {
+  vlengths <- c(length(x), length(y))
+  xydir <- which.max(vlengths)
+  if(length(fillcol) != (vlengths[xydir] - 1)) fillcol <- rep(fillcol[1], vlengths[xydir] - 1)
[…]

### 2017-12-01 R/rect_shade.R
Edits to figure 3; added missing publication years for certain records in calibration dataset; rect_shade committed.

@@ -1,6 +1,6 @@
 #' Create adjacent rectangular polygons
-#' @description Generic function to create adjacent rectangular polygons from vectors of coordinates passed to 
-#' it
+#' @description Generic function to create adjacent rectangular polygons from 
+#' vectors of coordinates passed to it
 #' @param x Vector of x coordinates, arranged in ascending order
 #' @param y Vector of y coordinates, arrange in acending order
 #' @param fillcol A 1 to n - 1 length vector of colors for filling. n = length of x or y (the longest)
@@ -31,15 +31,21 @@
 rect_shade <- function(x, y, fillcol = "grey", linecol = "black", lwd = 1) {
   vlengths <- c(length(x), length(y))
   xydir <- which.max(vlengths)
-  if(length(fillcol) != (vlengths[xydir] - 1)) fillcol <- rep(fillcol[1], vlengths[xydir] - 1)
-  if(length(linecol) != (vlengths[xydir] - 1)) linecol <- rep(linecol, vlengths[xydir] - 1)
+  if(length(fillcol) != (vlengths[xydir] - 1)) {
+    fillcol <- rep(fillcol[1], vlengths[xydir] - 1)
+  }
+  if(length(linecol) != (vlengths[xydir] - 1)) {
+    linecol <- rep(linecol, vlengths[xydir] - 1)
+  }
   rects <- lapply(1:(vlengths[xydir] - 1), function(j) {
     if(xydir == 1) {
-      pols <- polygon(x = c(x[j], rep(x[j + 1], 2), x[j]), y = c(y[1], y[1], y[2], y[2]), col = fillcol[j], 
+      pols <- polygon(x = c(x[j], rep(x[j + 1], 2), x[j]), 
+                      y = c(y[1], y[1], y[2], y[2]), col = fillcol[j], 
                       border = linecol[j], lwd = lwd)
     }
     if(xydir == 2) {
-      pols <- polygon(x = c(x[1], x[1], x[2], x[2]), y = c(y[j], rep(y[j + 1], 2), y[j]), col = fillcol[j], 
+      pols <- polygon(x = c(x[1], x[1], x[2], x[2]), 
+                      y = c(y[j], rep(y[j + 1], 2), y[j]), col = fillcol[j], 
                       border = linecol[j], lwd = lwd)
     }
   })

### 2017-12-01 R/box_plots.R
Cleaned up data-prep vignette to tidy up code on merging calibration and main dateset.  Revised figures1-3 vignettes thereafter. Removed box_plots imported from lmisc as ended up not using it.

@@ -1,145 +1,0 @@
-#' Calculates boxplot statistics
-#' @description Calculates the usual boxplot statistics plus mean
-#' @param x Vector of values from which to calculate statistics
-#' @param whiskers A two-element vector of the percentiles for the lower and upper whiskers (c(0.025, 0.975))
-#' @param weighted TRUE or FALSE (default) for weighted quantiles and mean
-#' @param weight.opts A list of options to be passed to wtd.quantile
-#' @param na.rm Should the quantile and mean functions exclude NA values in x? Default is TRUE, else FALSE
-#' @return A vector containing the 2.5, 25, 50, 75, 97.5th percentile values and the mean
-#' @note Function transplanted from lmisc package
-#' @details If weighted is TRUE, the wtd.quantile and wtd.mean functions from Hmisc are called, and a weights
-#' vector equal in length to x is needed, passed as the minimum necessary parameter for calculating weighted 
-#' quantiles. A named list should be provided where the name for each element matches the argument name for 
-#' Hmisc::wtd.quantile and Hmisd::wtd.mean. Other possible arguments include type and normwt.
-#' @examples 
-#' set.seed(234)
-#' x <- rnorm(20)
-#' w <- sample(1:10, size = 20, replace = TRUE)
-#' box_stats(x)
-#' box_stats(x, weighted = TRUE, weight.opts = list("weights" = w))
-#' @export
-box_stats <- function(x, whiskers = c(0.025, 0.975), weighted = FALSE, weight.opts = NULL, na.rm = TRUE) {
-  p <- c(whiskers[1], 0.25, 0.5, 0.75, whiskers[2])
-  if(weighted == FALSE) {
-    bs <- c(quantile(x, p, na.rm = na.rm), "mu" = mean(x, na.rm = na.rm))
-  } else if(weighted == TRUE) {
-      if(is.null(weight.opts)) stop("Parameters for weights are needed, and optionallym type, and normwt")
-      w <- weight.opts
-      if(!"weights" %in% names(w)) stop("You are missing, or have incorrectly named, the weights vector")
-      if(!"type" %in% names(w)) w$type <- "quantile"
[…]

### 2017-12-12 paper/main.tex
First round of edits for paper main text through Table 1 in intro.

@@
 
-% Place your abstract within the special {sciabstract} environment.
-
-%\begin{sciabstract}
 \section*{Abstract}
-%Ecologists have long known that the observational scales play a critical role in shaping our understanding of ecosystems, and that observations spanning multiple scales of space and time are needed to effectively characterize ecological phenomena.
 \textbf{To properly understand ecological phenomena, it is necessary to observe them across a range of spatial and temporal scales.
 Ecologists first raised this point in the 1980s, and since then the ability to collect multi-scale observations has grown rapidly.
-To assess modern ecology's progress in addressing scale, we analyzed the resolution, extent, interval, and duration of observations in 348 studies published between 2004-2014.
+To assess modern ecology's progress in addressing scale, we analyzed the resolution, extent, interval, and duration of observations in 348 observational studies published between 2004-2014.
 We found that the scale domains of observations are fairly narrow, and are collected primarily with conventional field techniques.
@@
 Compared to prior meta-analyses from the 1980s and early 2000s, observational durations and resolutions remain largely unchanged, but intervals have become finer and extents larger.
-Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, raising concerns about observational representativeness.
+Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, revealing unobserved portions of space and time not truly captured by replicates and raising concerns about observational comprehensiveness.
 Adding to these concerns, scales were not clearly reported in most studies, suggesting that it is a minor consideration.
 Journals can help mitigate this problem by implementing scale reporting standards, which can spur ecologists to more rapidly adopt new observational technologies, and thereby close key gaps in current observational domains.}
-
-%as well as \emph{actual} extent (the summed resolution of spatial replicates) and \emph{actual} duration (integrated sample duration of temporal replicates).
-Our results suggest that the scale domains of modern ecological observations remain fairly narrow, and are collected primarily with conventional field techniques.
-A comparison with results from prior earlier, scale-focused meta-analyses revealed that observational durations and resolutions are largely unchanged since the 1980s, but intervals have become finer and extents larger, the latter driven by steadily increasing use of remote sensing.
-%Most observations had resolutions of $\leq$1 m$^2$ (67\%) and extents $\leq$10,000 ha (53\%).
-Temporally, most observations were either unreplicated (37\%) or had intervals $\geq$1 month (26\%), while 64\% of observations had durations of $\leq$1 year.
-Juxtaposing dimensions revealed several notable ``holes'' (domains with few to no observations), particularly in the realms of high frequency (sub-daily to daily intervals) observations with fine to moderate resolutions ($\geq$1 m$^2$ up to 100 ha) and $\geq$decadal durations.
-These holes relate to tradeoffs between observational frequency and resolution, as well as duration, that are characteristic of field-based data collection, which was used for 80\% of observations.
-An analysis of values reported in the most comparable prior scale meta-analyses revealed that observational durations and resolutions are largely unchanged since the 1980s, but intervals have become finer and extents larger.
-Our data also showed an internal trend of increasing observational extent, coinciding with a 1.3\% per year increase in remote sensing-based observations.
-Despite this trend, the \emph{actual} extent of 81\% of observations was 4-8 orders of magnitude smaller than extent.
-Similarly, the \emph{actual} duration of 67\% of observations was 3-9 orders of magnitude shorter than duration.
-These large discrepancies raise questions about the representativeness of observations, which may in part be addressed by examining the degree to which auto-correlation processes close the gap between actual and represented scales of observation.
-Adding to this concern is the fact that most observations did not clearly report observational scales, suggesting that scale remains a minor consideration throughout much of ecology.}
-
-%Observational scales play a critical role in shaping our understanding of ecological pattern and process.
-To gain insight into the spatial and temporal domains being addressed by modern ecological science, we conducted a meta-analysis of papers published in 30 ecological journals between 2004 and 2014.
-We estimated two spatial dimensions (resolution and extent) and two temporal dimensions (interval and duration) of 371 ecological observations reported within 140 papers (out of 346 randomly selected titles).
-Our analysis revealed that ecology largely remains a field-based discipline that primarily collects observations within narrow spatial and temporal domains.
-Ecologists still make limited use of methods for observing larger-scale patterns and processes (e.g.
-remote sensing, used for 6\% of observations), but make greater use of those which permit high frequency or temporally continuous \emph{in situ} measurements (12\% of observations).
-Resolution, extent, and duration were not clearly reported for the majority of observations, which, together with the narrow spatio-temporal ranges of observed scales, suggests a limited ability to generalize the findings of many studies.
-
-%  An example of the style is the special\texttt{\{sciabstract\}} environment used to set up the abstract you see here.
-%\end{sciabstract}
 
@@
 
-%One of the best known and most comprehensive assessments of scale in ecological research ident
-
 Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) meta-analyses of ecological research scales were published nearly 30 years ago \cite{tilman_ecological_1989,kareiva_spatial_1988}, it is both timely and important to assess the scales of contemporary ecological investigation.
-To address this need, we quantified the spatial and temporal domains of empirical observations that were reported within recently (2004-2014) published ecological studies (here domain means the distribution of observations within the spectrum of one or more scale dimensions\footnote{This definition differs slightly from Wiens' \cite{wiens_spatial_1989}, who defined ``domain of scale'' as ``a portion of the scale spectrum within which process-pattern relationships are consistent regardless of scale.''}).
-Empirical observations are critical for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, therefore the spatio-temporal domains of observations provide an important indicator of the field's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.
-
-%This question may be best addressed by evaluating the scales of empirical observations.
+To address this need, we quantified the spatial and temporal domains (here domain means the distribution of observations within the spectrum of one or more scale dimensions\footnote{This definition differs slightly from Wiens' \cite{wiens_spatial_1989}, who defined ``domain of scale'' as ``a portion of the scale spectrum within which process-pattern relationships are consistent regardless of scale.''}) of empirical observations (defined here as ecological observations collected under un-controlled or non-manipulated conditions) that were reported within recently (2004-2014) published ecological studies.
 Empirical observations are critical for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, therefore the spatio-temporal domains of observations provide an important indicator of the field's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.
@@
 Duration measures the time elapsed between the first and last temporal replicates, or, in the case of temporally unreplicated observations, the estimated time spent collecting the observation.
-We also assessed observational scales within two additional dimensions, \emph{actual} extent (the integrated area of spatial replicates) and \emph{actual} duration (the summed observational time of temporal replicates).
+We also assessed observational scales within two additional dimensions, \emph{actual} extent (the summed area of spatial replicates) and \emph{actual} duration (the summed observational time of temporal replicates).
 We evaluated these additional dimensions to gain insight into how much the actual scales of observation (i.e.
-how much space and time is covered by the measurement) differ from those they are intended to represent.
-
-%For example, consider a hypothetical study reporting estimates of herbaceous plant cover in a forest understory, in which the authors reported that they placed 10 400 m$^2$ plots throughout the forest, and within each plot randomly placed 10 1 m$^2$ quadrats, within which they estimated percent herbaceous plant cover.
-In this example, we consider 1 m$^2$ to be the resolution.
+how much space and time is covered by the measurement) differ from the scales that the observations are intended to cover.
+This difference may contain important information about how effectively ecological observations characterize ecological phenomena.
+First, the odds that observations do not adequately represent the phenomenon of interest may increase (due to inappropriate sampling) as the gap between the actual and intended observational scales increases.
+Second (and related), a larger gap may increase the likelihood of collecting censored data (the values collected by observations are not fully known), as the frequent complexity, non-linearity, and non-randomness of ecosystems is such that phenomena may resolve themselves in the space or time between observations.
 
@@
 \multicolumn{2}{c}{\textbf{Component}} & \textbf{Units} & \textbf{Description} \\\hline\hline
-\multirow{3}*{Spatial} & Resolution & m$^2$ & Area of an individual spatial replicate (plot) \\%\cmidrule(l){2-3}
+\multirow{3}*{Spatial} & Resolution & m$^2$ & Area of an individual spatial replicate (e.g. plot) \\%\cmidrule(l){2-3}
    & Extent & ha & Area  encompassed by all spatial replicates \\

### 2017-12-15 paper/main.tex
Edits to main text. Updated figure 1 and figure 1 code to include line indicated unreplicated space in panel D (duration)

@@
 Compared to prior meta-analyses from the 1980s and early 2000s, observational durations and resolutions remain largely unchanged, but intervals have become finer and extents larger.
-Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, revealing unobserved portions of space and time not truly captured by replicates and raising concerns about observational comprehensiveness.
+Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, revealing portions of space and time not truly captured by replicates and raising concerns about observational comprehensiveness.
 Adding to these concerns, scales were not clearly reported in most studies, suggesting that it is a minor consideration.
@@
 We evaluated these additional dimensions to gain insight into how much the actual scales of observation (i.e.
-how much space and time is covered by the measurement) differ from the scales that the observations are intended to cover.
+how much space and time is covered by the measurement) differ from the scales that the observations are explicitly or implicitly intended to represent.
 This difference may contain important information about how effectively ecological observations characterize ecological phenomena.
-First, the odds that observations do not adequately represent the phenomenon of interest may increase (due to inappropriate sampling) as the gap between the actual and intended observational scales increases.
-Second (and related), a larger gap may increase the likelihood of collecting censored data (the values collected by observations are not fully known), as the frequent complexity, non-linearity, and non-randomness of ecosystems is such that phenomena may resolve themselves in the space or time between observations.
+First, an increasing gap between actual and intended observational scale inherently implies greater interpolation or extrapolation of observed measurements, raising the odds of over-leveraging data.
+Second (and relatedly), since natural systems are frequently complex, non-linear, and non-random \cite{levin_ecosystems_1998,pringle_spatial_2017,rietkerk_regular_2008}, a larger gap may increase the likelihood of encountering unanticipated data challenges such as censored data (\emph{sensu} \cite{efron_efficiency_1977}), as phenomena may resolve themselves in the space or time between observations.
 
@@
 
-%In contrast, cost is less likely to dictate the choice of observational scale in studies of natural systems (considering, for example, that many global-scale, multi-decadal, satellite-derived datasets are free \cite[e.g.][]{tucker_nasas_2004,friedl_modis_2010}), therefore an assessment confined to such studies may provide a less biased picture of ecology's scale domains.
-
-%We calculated these from 379 discrete observations reported within a 134 paper subset of 348
-%randomly selected articles (from 42,918 total). An additional 62 papers that were cited as the
-%source of observations in the selected papers were also reviewed.
-We confined our analysis to observations of “natural” (7), or non-experimentally manipulated, phenomena.
-We made this choice because experiments are costly and therefore tend to be fine-scaled in both space and time (7, 9), which could unduly bias our analysis.
-%
-%We extracted these scale dimensions from 379 observations reported within 134 papers selected from the top 30 (based on impact factor) ecology-themed journals.
-The papers were a subset of 348 that we reviewed in total, which we selected randomly from the 42,918 studies published between 2004-2014 .
-An additional 62 papers that were cited as the source of observations in the selected papers were also reviewed.
-We confined our analysis to observations of ``natural'' \cite{tilman_ecological_1989}, or non-experimentally manipulated, phenomena.
-We made this choice because experiments are costly and therefore tend to be fine-scaled in both space and time \cite{tilman_ecological_1989,kareiva_spatial_1988}, which could unduly bias our analysis.
-
 To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI).
@@
 1B).
+Bar widths in (C-D) indicate differences in scale between x-axis labels.
 
@@
 %\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/kde45.pdf}
+\includegraphics[width=1\textwidth]{../vignettes/figures/fig2.png}
 \vspace{-0.15 cm}
 \caption{Kernel density estimates of observational densities within the domains defined by A) interval and resolution (of temporally replicated observations only), B) duration and extent, C) resolution and extent, and D) interval and duration (of temporally replicated observations).
-Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.}
+Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.
+Letters in the plots denote the median values of different observational methods (f=field observations; a = automated sensing; r = remote sensing; p = paleo-observations).
+The grey shaded areas represent physically impossible domains (intervals greater than duration and resolutions greater than extent).}
 \label{afoto1}
@@
 Three other notable, but lesser areas of concentration are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($<$1 day) of either 1-10 ha or 10,000-1,000,000 ha.
-
-% the bulk of ecological observations span one month to one decade in time and covered total time periods of one minute to $<$1 month and areas of $\leq$1 m$^2$ (Fig.
-2B).
-The next greatest concentration were instantaneous ($\leq$1 second) observations of $<$1 m$^2$ extent, followed by a fainter concentration centered on 1 hour duration and between 1 m$^2$ and 1 ha in extent.
-Very few observations had spatial extents $>$1,000,000 ha, and because these were collected by satellites that make near-instantaneous measurements, they tended to have aggregate durations of $\leq$1 hour.
-In contrast, the longest periods of observation were provided by paleo-ecological or long-term weather stations, both of which produce continuous, point-scale measurements (e.g.
-sediment cores and weather instrumentation).
 
@@
 
-%Additional sources of uncertainty: rules for extracting scales; diversity of study types and ecological methods.
-
-%Most studies provided insufficient methodological detail to judge whether sampling schemes were suitable matched to and representative of the phenomena they were measuring.
-The observational scales were not clearly reported in many studies, such that we had to estimate spatial resolution and extent in 62\% and 68\% of cases, respectively, and temporal interval and duration in 40\% and 86\% of cases.
-Although the confidence intervals around our scale estimates (Fig.
-1) suggests that our findings are robust to this level of uncertainty, it would have likely confounded efforts to estimate effective scales.
-
-%This analysis did not necessarily quantify the scales \emph{represented} by ecological observations (SI).
-Observations that were non-continuous in time or space (e.g.
-point-based field measurements) may effectively represent larger scales than our estimates suggest, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997, palmer_scale_1994,cao_comparison_2002, legendre_spatial_1993,collins_method_2000-1}.
-
-%Adding to our motivation to record actual observational scales was the fact that 
 \noindent \textbf{Discussion}\\
@@
 2B).
-
-%there are prominent holes in the scale domains of ecology, notably in the spaces defined by observations having 1) daily to sub-daily intervals and resolutions $\geq$1 m$^2$ up to 100 ha; 2) decadal or longer durations and extents $\geq$10,000,000 ha, such that few observations are collected at 1) high spatial and temporal resolution (1 m$^2$, 2) over large areas for long time periods.
 
@@
 The distance between the adjusted actual value and the ostensible value can provide a measure of how well the replicates represent the intended scale of observation.
-If the gap remains large, then alternative sampling methods may be used to close it.
+Though increasing spatial or temporal coverage may not always be the goal of a study (e.g., when spatial or temporal autocorrelation is a measure of interest), if the gap between actual and ostensible values remains large, then alternative sampling methods may be used to close it.
 For example, remote sensing provides wall-to-wall spatial coverage of a study area, erasing the difference between actual extent and extent.
@@
 
-%In contrast, cost is less likely to dictate the choice of observational scale in studies of natural systems (considering, for example, that many global-scale, multi-decadal, satellite-derived datasets are free \cite[e.g.][]{tucker_nasas_2004,friedl_modis_2010}), therefore an assessment confined to such studies may provide a less biased picture of ecology's scale domains.
-
-%These results indicate that for most ecological observations are collected with spatially discontinuous replicates that sample very small portions of the area in which they are distributed.
-%
-%Very little research is conducted at high spatial and temporal resolutions over large areas or for long time periods, indicating that, despite the well-established understanding of the importance of multi-scale assessments for understanding ecological patterns and processes \cite{levin_problem_1992,wiens_spatial_1989}, efforts focused on larger scales are still relatively sparse within the discipline \cite{levin_problem_1992,wiens_spatial_1989}.
-%
-%a concern that has been previously noted due to ecology's geographical bias towards anthropogenically undisturbed and temperate ecosystems \cite{martin_mapping_2012}. 
-%
-%As with spatial data, finely resolved temporal data can be aggregated to facilitate multi-scale analyses (and many of the reviewed studies that used automated sensing aggregated the resulting observations before analyzing them), whereas longer-interval data (e.g.
-annual biomass accumulation) cannot be disaggregated into shorter interval measurements (e.g.
-weekly biomass accumulation) without interpolation, which turns data into modeled, rather than direct, observations.
-
-%In the coming years, rapid technological advances should increase the concentration of ecological observations in currently under-represented domains.
-The growing numbers of high-resolution satellite sensors, together with new analytical platforms that provide free access to large volumes of pre-processed data and computational power \cite{googleearthengine}, will lower technical barriers that have so far prevented ecologists from adopting this observational technology  \cite{pettorelli_satellite_2014}.
-Similarly, the advent of unmanned systems offers the ability to measure ecological features at high spatial and temporal frequencies over large areas \cite{anderson_lightweight_2013}, which were scales that were previously impractical to access.
-The ever-falling cost of sensor technology and the ubiquity of cell phones also means that ecologists, together with a growing army of citizen-scientists, have the unprecedented ability to make spatially dense, high frequency observations over large areas \cite{wolf_gsm-based_2012,collins_new_2006,porter_wireless_2005,dickinson_current_2012}.
-A greater attentiveness to scale in general, including more meticulous documentation of observed dimensions, may help to facilitate the spread of ecological research to sparsely studied scales, while improving transferability of knowledge within the discipline.
-
-%\paragraph*{Displayed math.} 
 \bibliography{/Users/lestes/Dropbox/publications/fullbib}
@@
 National Science Foundation (1125210).
-
-% Following is a new environment, {scilastnote}, that's defined in the
-% preamble and that allows authors to add a reference at the end of the
-% list that's not signaled in the text; such references are used in
-% *Science* for acknowledgments of funding, help, etc.
-
-% For your review copy (i.e., the file you initially send in for
-% evaluation), you can use the {figure} environment and the
-% \includegraphics command to stream your figures into the text, placing
-% all figures at the end.  For the final, revised manuscript for
-% acceptance and production, however, PostScript or other graphics
-% should not be streamed into your compliled file.  Instead, set
-% captions as simple paragraphs (with a \noindent tag), setting them
-% off from the rest of the text with a \clearpage as shown  below, and
-% submit figures as separate files according to the Art Department's
-% instructions.
-
-%\begin{scilastnote}
-%\item This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G).
-Erle Ellis, Jason Chang, and Labeeb Ahmed of the GLOBE Project (http://globe.umbc.edu) were supported by the U.S.
-National Science Foundation (1125210).
-%\end{scilastnote}
-
 \clearpage

### 2017-12-17 paper/main.tex
Removed cache statements in fig3, additional-analyses vignettes, some edits to calibration (remove commented out code), checking results in that. Edits in manuscript

@@
 \noindent \textbf{Results}\\
+\noindent \textbf{\emph{Observational methods}}\\
+To account for potential differences in scales related to methodology, we classified each observation according to the following broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing/other geographic data, and paleo-reconstruction approaches.
+Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.9\%, and paleo-reconstruction for less than 0.8\%.
+
 \noindent \textbf{\emph{Domains}}\\
@@
 1B).
-Bar widths in (C-D) indicate differences in scale between x-axis labels.
 
@@
 
-\begin{figure}[!ht]
-%\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/hists_bs.png}
-\vspace{-0.15 cm}
-\caption{Histograms of the resolution (A), extent (B), interval (C), and duration (D) of observations collected from the surveyed ecological studies.
-Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval.
-}
-\label{afoto1}
-\end{figure}
-
 In the temporal dimensions, 37\% of observations were not repeated (Fig.
 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals.
-Duration was one day or less for 31\% of sampled observations, while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 9\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig.
+Duration was one day or less for 31\% of sampled observations (due to lack of temporal replication), while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 9\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig.
 1D).
@@
 A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident.
+Different observational methods occupied substantially different portions of the domain space, as indicated by the locations of the median resolution and interval values for each:  field observations had a resolution of 0.1-1 m$^2$ and were repeated at monthly intervals, whereas remote observations had coarser resolutions (1000 m$^2$) but finer repeat intervals ($\sim$1 day).
+Paleo-reconstructions and automated sensing techniques were both finely resolved (10 cm$^2$ to 0.01 m$^2$), but automated approaches had hourly-daily intervals while paleo-reconstructions had multi-decadal intervals.
+The complete domain space of each observational type is detailed in the SI.
+
 This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations.
@@
 2A), the greatest being unreplicated samples with resolutions of 100 cm$^2$ to 1 m$^2$, followed by observations of daily to monthly interval with slightly finer spatial resolutions (10 cm$^2$ to 1 m$^2$), and finally a smaller concentration of monthly to yearly measurements collected at spatial scales of 0.1-10 ha.
-
-\begin{figure}[!ht]
-%\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{../vignettes/figures/fig2.png}
-\vspace{-0.15 cm}
-\caption{Kernel density estimates of observational densities within the domains defined by A) interval and resolution (of temporally replicated observations only), B) duration and extent, C) resolution and extent, and D) interval and duration (of temporally replicated observations).
-Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.
-Letters in the plots denote the median values of different observational methods (f=field observations; a = automated sensing; r = remote sensing; p = paleo-observations).
-The grey shaded areas represent physically impossible domains (intervals greater than duration and resolutions greater than extent).}
-\label{afoto1}
-\end{figure}
 
@@
 
-\begin{figure}[ht]
-\includegraphics[width=1\textwidth]{figures/act_v_eff_diff.png}
-\vspace{-0.2 cm}
-\caption{The difference between extent and \emph{actual} extent (the summed area of spatial replicates) (A) and duration and \emph{actual} duration (the summed sampling duration across temporal replicates) (B).
-Difference values are expressed in terms of how many orders of magnitude larger (or longer) extent (duration) is than actual extent (actual duration), and are summarized (as box plots, with circle in box representing the mean and line the median) in bins representing increasing scales of actual extent/duration.
-The percentages of observations falling within each bin are indicated by the color of the inter-quartile and the numeric value above the upper whisker.}
-\label{afoto1}
-\end{figure}
-
-\noindent \textbf{\emph{Observational methods}}\\
-We classified the method used to collect each observation into several broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing, other geographic data, and paleo-reconstruction approaches.
+%\noindent \textbf{\emph{Observational methods}}\\
+%We classified the method used to collect each observation into several broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing, other geographic data, and paleo-reconstruction approaches.
 Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.4\%, and paleo-reconstruction and other geographic data each for less than 1\% each.
-Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance.
-Automated sensing methods showed no trend over time (SI).
 
@@
 Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates.
+
+Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance.
+Automated sensing methods showed no trend over time (SI).
+
 Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average (see SI for details of calculation).
@@
 
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\includegraphics[width=1\textwidth]{../vignettes/figures/fig1.png}
+\vspace{-0.15 cm}
+\caption{Histograms of the resolution (A), extent (B), interval (C), and duration (D) of observations collected from the surveyed ecological studies.
+Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval.
+Bar widths in C-D indicate differences in scale between x-axis labels.
+The grey vertical line in D indicates that the majority ($>$95\%) of observations of $\leq$1 day duration were temporally unreplicated.}
+\label{afoto1}
+\end{figure}
+
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\includegraphics[width=1\textwidth]{../vignettes/figures/fig2.png}
+\vspace{-0.15 cm}
+\caption{Kernel density estimates of observational densities within the domains defined by A) interval and resolution (of temporally replicated observations only), B) duration and extent, C) resolution and extent, and D) interval and duration (of temporally replicated observations).
+Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.
+Letters in the plots denote the median values of different observational methods (f=field observations; a = automated sensing; r = remote sensing; p = paleo-observations).
+The grey shaded areas represent physically impossible domains (intervals greater than duration and resolutions greater than extent).}
+\label{afoto1}
+\end{figure}
+
+\begin{figure}[ht]
+\includegraphics[width=1\textwidth]{../vignettes/figures/fig3.png}
+\vspace{-0.2 cm}
+\caption{The difference between extent and \emph{actual} extent (the summed area of spatial replicates) (A) and duration and \emph{actual} duration (the summed sampling duration across temporal replicates) (B).
+Difference values are expressed in terms of how many orders of magnitude larger (or longer) extent (duration) is than actual extent (actual duration), and are summarized (as box plots, with circle in box representing the mean and line the median) in bins representing increasing scales of actual extent/duration.
+The percentages of observations falling within each bin are indicated by the color of the inter-quartile and the numeric value above the upper whisker.
+The grey vertical line in D indicates that the majority ($>$95\%) of observations of $\leq$1 day duration were temporally unreplicated.}
+\label{afoto1}
+\end{figure}
+
 \end{document}

### 2017-12-17 R/data.R
Added documentation for kderla, kderls, and hdatstat_bt to documentation. incremented package

@@ -32,3 +32,37 @@ NULL
 #' data(datf)
 NULL
 
+#' Histograms of scale dimensions
+#' @name hdatstat_bt
+#' @docType data
+#' @description A list containing mean, 2.5, and 97.5 percentile values 
+#' resulting from histograms applied to bootperturb for each of the four primary 
+#' dimensions (used in making Figure 1), divided by observation type.
+#' @author Estes, Treuer, Elsen, Chang, Choi, Ahmed, Caylor, Ellis
+#' @keywords data
+#' @examples
+#' data(hdatstat_bt)
+NULL
+
+#' Kernel density estimates for bivariate dimensions comparisons
+#' @name kderla
+#' @docType data
+#' @description A list of kernel density rasters and various plotting 
+#' parameters, calculated across all observations and by observational type, for
+#' Figure 2.
+#' @author Estes, Treuer, Elsen, Chang, Choi, Ahmed, Caylor, Ellis
+#' @keywords data
+#' @examples
+#' data(kderla)
+NULL
+
+#' Kernel density estimates for bandwidth sensitivity estimation
+#' @name kderls
+#' @docType data
+#' @description A list of kernel density rasters calculated for bandwidth 
+#' sensitivity analysis
+#' @author Estes, Treuer, Elsen, Chang, Choi, Ahmed, Caylor, Ellis
+#' @keywords data
+#' @examples
+#' data(kderls)
+NULL

### 2017-12-18 paper/SI.tex
After check of calibration dataset in June last commit and again rebuilding calibration vignette in the current version of master, found data didn't change much. There were a few places in TableS2 in SI where values were of at 2nd or 3rd decimal place. Updated those. Values that went into perturbing results did not change. Updated and re-ran through Figure 1 vignette
Added new observational methods figures to SI, updated all figure references in SI.

@@
 \hline
-CV of overall mean & 50 & 50 & 75 & 48 & 34 & 109 \\
-Mean of record-wise CV & 57 & 41 & 72 & 104 & 64 & 126 \\
+CV of overall mean & 50 & 50 & 78 & 49 & 36 & 109 \\
+Mean of record-wise CV & 58 & 41 & 72 & 105 & 64 & 126 \\
 \hline
@@
 
+\subsubsection*{The domains of observational methods}
+\vspace{-10pt}
+
+Fig.
+\ref{obshists} details the distributions of observations by method along the four primary dimensions, while Fig.
+\ref{obskde} reveals the densities of observations by method within the dimensional juxtapositions shown in Fig.
+2 in the main text.
+
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\includegraphics[width=1\textwidth]{../vignettes/figures/fig1s.png}
+\vspace{-20 pt}
+\caption{Histograms of the resolution, extent, interval, and duration of observational scales, shown for all observational methods (top row) and for the four main observational methods (rows 2-5).
+Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval.
+Bar widths for interval and duration indicate differences in scale between x-axis labels.
+}
+\label{obshists}
+\end{figure}
+
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\includegraphics[width=1\textwidth]{../vignettes/figures/fig2s.png}
+\vspace{-20 pt}
+\caption{Kernel density estimates across all observational types (top row) and for each of the four primary observational types (rows 2-5).
+Density estimates for resolution (x axis) versus interval (y axis) are presented in the first column, duration (x) versus interval (y) in the second column, resolution (x) versus extent (y) in the third column, and duration (x) versus extent (y) in the fourth column.
+Density estimates were applied to the log-transformed values of each observational dimension, and were rescaled to represent percentages.
+The grey shaded areas represent physically impossible domains (e.g.
+intervals greater than duration).}
+\label{obskde}
+\end{figure}
+
 \subsubsection*{Trends in observational methods and scales by year}
 \vspace{-10pt}
-Figures S1 and S2 respectively show how the frequency of ecological observing methods and the average scale in each dimensions varies by publication year (2004-2014).
+Figures \ref{type_by_yr} and \ref{sc_by_yr} respectively show how the frequency of ecological observing methods and the average scale in each dimensions varies by publication year (2004-2014).
 
@@
 }
-\label{afoto1}
+\label{type_by_yr}
 \end{figure}
@@
 }
-\label{afoto1}
+\label{sc_by_yr}
 \end{figure}
@@
 \vspace{-10pt}
-Figure S3 indicates the effect that varying bandwidth has on the appearance of the kernel density estimates.
+Figure \ref{ksens} indicates the effect that varying bandwidth has on the appearance of the kernel density estimates.
 The smallest bandwidth (0.4) tested (Fig.
-S3 top) shows that the primary observational concentrations described in the main text are evident, but tend to be devolved into separate concentrations.
+\ref{ksens} top) shows that the primary observational concentrations described in the main text are evident, but tend to be devolved into separate concentrations.
 An example is the oblong concentration of observations that in Fig.
 1A (and in the lower left panel in Fig.
-S3) is bounded on the lower right at monthly to yearly intervals and 100-1000 m$^2$ resolutions and on the upper left by near-daily to monthly observations and 0.1-10 ha resolutions.
+\ref{ksens}) is bounded on the lower right at monthly to yearly intervals and 100-1000 m$^2$ resolutions and on the upper left by near-daily to monthly observations and 0.1-10 ha resolutions.
 With the smaller bandwidth this concentration appears as two separate patches (Fig.
-S3 top left), but with 0.7 bandwidth applied becomes coherent (Fig.
-S3 middle left).
+\ref{ksens} top left), but with 0.7 bandwidth applied becomes coherent (Fig.
+\ref{ksens} middle left).
 This increasing separation of points into separate clusters is expected as bandwidth is reduced.
@@
 %\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=0.9\textwidth]{figures/res_v_extent_ksize2.pdf}
+\includegraphics[width=0.9\textwidth]{../vignettes/figures/res_v_extent_ksize2.png}
 \vspace{10 pt}
@@
 }
-\label{afoto1}
+\label{ksens}
 \end{figure}

### 2017-12-18 paper/SI.tex
Revised supplemental figure titles, added in all figures and caption to SI. Updated gitignore to ignore all non RMd vignette components

@@
 The following text provides additional information on this study's Methods and Results.
-The full manuscript, data, and code are available on an accompanying \href{https://github.com/ecohydro/ecoscales}{github repository}.
+The full manuscript, data, and code are available on an accompanying \href{https://github.com/agroimpacts/ecoscales}{github repository}.
 \textbf{Note to editors and reviewers: this repository is currently private, but will be made public upon acceptance of manuscript.
-In the interim, the code and data can be downloaded from the following links}:
-\href{https://www.dropbox.com/s/4r9h8mn1wa4kjsn/calibration.html?dl=0}{Calibration analysis}; \href{https://www.dropbox.com/s/n0zq5zufu96ktfl/calibration-set.xlsx?dl=0
-}{Calibration dataset}; \href{https://www.dropbox.com/s/bzni8dhqy8un4fz/space-time.html?dl=0
-}{Main analysis}; \href{https://www.dropbox.com/s/vujjo2x4f5kwqkb/main-results.xlsx?dl=0
-}{Main dataset}
+In the interim, the entire project folder can be downloaded from the following \href{https://www.dropbox.com/sh/1l9em6p0wyabwr3/AADhCpFvqjv6y7VbL5tgB3Iqa?dl=0}{link}.}
 
@@
 However, there may be cases where once-off observations may have a longer duration than the sample duration.
-For example, consider a study that counts occurrences of pollinators over three years, using transects that are located in different locations within the broader study area during each year \cite{rollin_differences_2013-1}.
+For example, consider a study that counts occurrences of pollinators over three years, using transects that are located in different locations within the broader study area during each year \cite{rollin_differences_2013}.
 The observations are therefore not strictly temporally replicated, but the authors control for year of collection in their subsequent analysis to avoid confounding effects.
@@
 2 in the main text.
-
-\begin{figure}[!ht]
-%\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{../vignettes/figures/fig1s.png}
-\vspace{-20 pt}
-\caption{Histograms of the resolution, extent, interval, and duration of observational scales, shown for all observational methods (top row) and for the four main observational methods (rows 2-5).
-Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval.
-Bar widths for interval and duration indicate differences in scale between x-axis labels.
-}
-\label{obshists}
-\end{figure}
-
-\begin{figure}[!ht]
-%\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{../vignettes/figures/fig2s.png}
-\vspace{-20 pt}
-\caption{Kernel density estimates across all observational types (top row) and for each of the four primary observational types (rows 2-5).
-Density estimates for resolution (x axis) versus interval (y axis) are presented in the first column, duration (x) versus interval (y) in the second column, resolution (x) versus extent (y) in the third column, and duration (x) versus extent (y) in the fourth column.
-Density estimates were applied to the log-transformed values of each observational dimension, and were rescaled to represent percentages.
-The grey shaded areas represent physically impossible domains (e.g.
-intervals greater than duration).}
-\label{obskde}
-\end{figure}
+Fig.
+\ref{obsdiff} shows the how the differences between actual extent and extent  and actual duration and duration vary by observational type.
 
@@
 Figures \ref{type_by_yr} and \ref{sc_by_yr} respectively show how the frequency of ecological observing methods and the average scale in each dimensions varies by publication year (2004-2014).
-
-\begin{figure}[!ht]
-%\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=1\textwidth]{figures/stype_by_yr.png}
-\vspace{10 pt}
-\caption{Trends in percentage of observing methods by year of publication.
-The coefficient of a weighted (by number of studies in each year) linear regression fit to the annual percentages of observations made with remote sensing (left), field methods (center), and automated sensors (right) is presented at the top of each plot, as well as the regression coefficient of determination and p-value.
-}
-\label{type_by_yr}
-\end{figure}
-
-\begin{figure}[ht]
-%\begin{wrapfigure}{c}{1\textwidth}
-\center{\includegraphics[width=0.8\textwidth]{figures/dim_by_yr.png}}
-\vspace{5 pt}
-\caption{Trends in observational scales by year of publication.
-The coefficient of a weighted (by number of studies in each year) linear regression, fit to the logarithm (base 10) of the mean scale values (calculated for each publication year) for the six assessed dimensions is presented at the top of each plot, as well as the model coefficient of determination and p-value.
-}
-\label{sc_by_yr}
-\end{figure}
-%\clearpage
 
@@
 \begin{figure}[!ht]
+\centering
+\includegraphics[width=0.9\textwidth]{../vignettes/figures/figS1.png}
+\vspace{-20 pt}
+\caption{Histograms of the resolution, extent, interval, and duration of observational scales, shown for all observational methods (top row) and for the four main observational methods (rows 2-5).
+Bars represent the average percentages for each bin realized after 1000 perturbed resamples, while grey bars indicate the 95\% confidence interval.
+Bar widths for interval and duration indicate differences in scale between x-axis labels.
+}
+\label{obshists}
+\end{figure}
+
+\begin{figure}[!ht]
 %\begin{wrapfigure}{c}{1\textwidth}
-\includegraphics[width=0.9\textwidth]{../vignettes/figures/res_v_extent_ksize2.png}
+\includegraphics[width=1\textwidth]{../vignettes/figures/figS2.png}
+\vspace{-20 pt}
+\caption{Kernel density estimates across all observational methods (top row) and for each of the four primary observational methods (rows 2-5).
+Density estimates for resolution (x axis) versus interval (y axis) are presented in the first column, duration (x) versus interval (y) in the second column, resolution (x) versus extent (y) in the third column, and duration (x) versus extent (y) in the fourth column.
+Density estimates were applied to the log-transformed values of each observational dimension, and were rescaled to represent percentages.
+The grey shaded areas represent physically impossible domains (e.g.
+intervals greater than duration).}
+\label{obskde}
+\end{figure}
+
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\centering
+\includegraphics[width=0.7\textwidth]{../vignettes/figures/figS3.png}
+\vspace{-5 pt}
+\caption{The difference between extent and \emph{actual} extent and duration and \emph{actual} duration, as calculated across all observational methods (top row) and for each of the four primary observational methods.
+Difference values are expressed in terms of how many orders of magnitude larger (or longer) extent (duration) is than actual extent (actual duration), and are summarized (as box plots, with circle in box representing the mean and line the median) in bins representing increasing scales of actual extent/duration.
+The percentages of observations falling within each bin are indicated by the color of the inter-quartile and the numeric value above the upper whisker.}
+\label{obsdiff}
+\end{figure}
+
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\includegraphics[width=1\textwidth]{../vignettes/figures/figS4.png}
+\vspace{10 pt}
+\caption{Trends in percentage of observing methods by year of publication.
+The coefficient of a weighted (by number of studies in each year) linear regression fit to the annual percentages of observations made with remote sensing (left), field methods (center), and automated sensors (right) is presented at the top of each plot, as well as the regression coefficient of determination and p-value.
+}
+\label{type_by_yr}
+\end{figure}
+
+\begin{figure}[ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\center{\includegraphics[width=0.8\textwidth]{../vignettes/figures/figS5.png}}
+\vspace{5 pt}
+\caption{Trends in observational scales by year of publication.
+The coefficient of a weighted (by number of studies in each year) linear regression, fit to the logarithm (base 10) of the mean scale values (calculated for each publication year) for the six assessed dimensions is presented at the top of each plot, as well as the model coefficient of determination and p-value.
+}
+\label{sc_by_yr}
+\end{figure}
+%\clearpage
+
+\begin{figure}[!ht]
+%\begin{wrapfigure}{c}{1\textwidth}
+\includegraphics[width=0.9\textwidth]{../vignettes/figures/figS6.png}
 \vspace{10 pt}

### 2017-12-18 paper/main.tex
Revised supplemental figure titles, added in all figures and caption to SI. Updated gitignore to ignore all non RMd vignette components

@@
 
-This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations.
+%This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations.
 This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.

### 2017-12-18 paper/main.tex
Revising main text to include information on observational type distributions, and updated figure 1 vignette to produce percentage of observations per bin by observations type

@@
 \noindent \textbf{\emph{Domains}}\\
-In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
+In terms of resolution, the majority (67\%) of observations (across all methods) were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
 1A).
+These distributions primarily reflect those of field observations, the dominant observational methodology.
+Examining the distributions for each observational method (Fig.
+S1 in SI) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer ($<$0.1 m$^2$) than field observations, while remote observations were generally coarser ($>$100 m$^2$).
+
 The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 12\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig.
 1B).
+As with resolution, the extent of automated observations tended to be finer
 

### 2017-12-18 paper/main.tex
Editing main text through spatio-temporal domains. Updarted figure 1 vignette to calculate observational type percents by bins with unreplicated values removed.

--- 2017-12-18 paper/main.tex
@@
 \noindent \textbf{\emph{Observational methods}}\\
-To account for potential differences in scales related to methodology, we classified each observation according to the following broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing/other geographic data, and paleo-reconstruction approaches.
+To account for potential differences in scales related to methodology, we classified each observation according to the following broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing/other geographic data (hereafter remote observations), and paleo-reconstruction approaches.
 Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.9\%, and paleo-reconstruction for less than 0.8\%.
 
-\noindent \textbf{\emph{Domains}}\\
-In terms of resolution, the majority (67\%) of observations were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
+\noindent \textbf{\emph{Distributions within the four primary dimensions}}\\
+In terms of resolution, the majority (67\%) of observations (across all methods) were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
 1A).
+These distributions primarily reflect those of field observations, the dominant observational methodology.
+Examining the distributions for each observational method (Fig.
+S1 in SI) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $<$0.1 m$^2$) than field observations (47\% $<$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $>$100 m$^2$).
+
 The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 12\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig.
 1B).
+As with resolution, the extent covered by automated sensing methods tended to be smaller (52\% $<$100 ha) than those of field observations (31\% $<$100 ha), while all but 4\% of remote observations covered areas $\geq$10,000 ha (as did the small number of paleo-reconstructions).
 
@@
 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals.
+With respect to temporally replicated observation (Fig.
+S1 in SI), automated sensing techniques had the finest intervals (61\% $\leq$1 day; 100\% $\leq$1 year), followed by remote observation (37\% $\leq$1 day; 78\% $\leq$1 year), field observations (17\% $\leq$1 day; 86\% $\leq$1 year), and paleo-reconstructions (21\% $\leq$1 decade).
+
 Duration was one day or less for 31\% of sampled observations (due to lack of temporal replication), while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 9\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig.
 1D).
+Paleo-reconstructions naturally had the longest duration (67\% $>$ 1 decade), while just $\sim$40\% of field, automated, and remote observations had durations of 1 year or longer.
 
-Juxtaposing these observational dimensions provides further insight into the spatio-temporal domains of observations (Fig.
+\noindent \textbf{\emph{Spatial and temporal domains}}\\
+Juxtaposing these observational dimensions provides further insight into the spatial and temporal domains of observations (Fig.
 2).
@@
 2A).
-A less dense, oblong concentration of observations bounded on the lower right by monthly to yearly observations at 100 m$^2$ and on the upper left by near-daily to monthly observations with 1-10 ha resolution is also evident.
-Different observational methods occupied substantially different portions of the domain space, as indicated by the locations of the median resolution and interval values for each:  field observations had a resolution of 0.1-1 m$^2$ and were repeated at monthly intervals, whereas remote observations had coarser resolutions (1000 m$^2$) but finer repeat intervals ($\sim$1 day).
-Paleo-reconstructions and automated sensing techniques were both finely resolved (10 cm$^2$ to 0.01 m$^2$), but automated approaches had hourly-daily intervals while paleo-reconstructions had multi-decadal intervals.
-The complete domain space of each observational type is detailed in the SI.
+A less dense, oblong concentration of observations bounded on the upper left by monthly to yearly observations at 100 m$^2$ resolution and on the lower right by near-daily to monthly observations with 1-10 ha resolution is also evident.
+The four observational methods occupied substantially different portions of the domain space, as indicated by the locations of their median values (and illustrated further in Fig.
+S2 in the SI): the median domain of field observations was between 0.1-1 m$^2$ of resolution with a monthly interval, whereas remote observations had coarser median resolutions (1000 m$^2$) but finer median intervals ($\sim$1 day).
+Paleo-reconstructions and automated sensing techniques were both finely resolved (medians from 10 cm$^2$ to 0.01 m$^2$), but automated approaches had hourly-daily median intervals compared to multi-decadal for paleo-reconstructions.
 
@@
 
-Contrasting duration and extent (for all observations) reveals two primary domains of observational concentration.
-The first consists of observations spanning one month to one decade in time and 10-1000 ha in space, while the second is defined by observations of one year to several decades that cover 10,000 to 1,000,000 ha (Fig.
+Comparing the interval and duration of temporally replicated observations showed most observations had daily to decadal intervals and durations of $\geq$1 month up to 1 decade (Fig.
 2B).
-Three other notable, but lesser areas of concentration are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($<$1 day) of either 1-10 ha or 10,000-1,000,000 ha.
+The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting one year to one decade tend to have yearly to decadal intervals.
+This tendency is reflected in the median domain locations of the primary observational methods: automated sensing had the finest median interval (hour-day) and shortest duration (month-year), followed by remote sensing (median interval slightly greater than one day and median duration of 1 year), field observations (median monthly interval and duration just over 1 year), and finally paleo-reconstructions (median interval 1 decade and millennial duration).
+The low densities of observations having sub-daily intervals shows that relatively few high frequency, long duration ecological measurements are undertaken; this position
 
-Comparing the two spatial dimensions against one another (for all observations) shows a primary concentration of observations with 10 cm$^2$ to 100 m$^2$ resolution that have extents ranging between slightly over 1,000 to nearly 1,000,000 ha (Fig.
+Contrasting the two spatial dimensions against one another (for all observations) shows a primary concentration of observations with 10 cm$^2$ to 100 m$^2$ resolution with extents ranging from just over 1,000 to nearly 1,000,000 ha (Fig.
 2C).
 The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha.
-These three concentrations suggest a tendency for observational extent to increase with resolution, which is a relationship that becomes more pronounced in the less densely observed portion of the domain where resolutions $\geq$100 m$^2$.
+These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Fig.
+S2 in SI) of automated (0.01 m$^$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.
+Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a result that likely reflects the very low sample size for this observational type.
 
-Most temporally replicated observations have daily to decadal intervals and durations of $\geq$ 1 month to 1 decade.
-The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting one year to one decade tend to have yearly to decadal intervals.
-The low densities of observations having sub-daily intervals shows that relatively few high frequency, long duration ecological measurements are undertaken.
+Two primary domains of observational concentration are revealed by juxtaposing duration and extent (across all observations).
+The first consists of observations spanning one month to one decade in time and 10-1,000 ha in space, while the second is defined by observations of one year to several decades that cover 10,000 to 1,000,000 ha (Fig.
+2D).
+Three other notable, but lesser areas of concentration are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($<$1 day) of either 1-10 ha or 10,000-1,000,000 ha.
+The median observation (1 year duration, 100 ha extent) from automated sensing lies near the center of the second major concentration, while the median extents of remote (1-10 million ha) and field observations bound the primary concentration at its upper and lower extents, while the median duration of both observational types falls between 1 month to 1 year.
 
+\noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
 For our final analysis of observational domains, we evaluated the differences between the scales represented by the extent and duration dimensions and the scales that ecological observations actually cover.
 To make this assessment, we log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, yielding the magnitude of difference between each pair of dimensions for each observation.
-We plotted these values (summarized in box plots) against their corresponding actual extent/duration values to evaluate whether these differences vary with scale (Fig.
+We plotted these values (summarized in box plots) against their corresponding extent/duration values to evaluate whether these differences vary with scale (Fig.
 3).

### 2017-12-19 paper/main.tex
Edited main text through difference magnitudes. Made more edits to figure S3 code, fixing too wide boxes in paleo duration. Added return to tops in a couple other vignettes

@@
 These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Fig.
-S2 in SI) of automated (0.01 m$^$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.
+S2 in SI) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.
 Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a result that likely reflects the very low sample size for this observational type.
@@
 To make this assessment, we log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, yielding the magnitude of difference between each pair of dimensions for each observation.
-We plotted these values (summarized in box plots) against their corresponding extent/duration values to evaluate whether these differences vary with scale (Fig.
+We plotted these values (summarized in box plots) against their corresponding extent/duration values to evaluate whether these differences varied with scale (Fig.
 3).
-This analysis showed that most (81\%) of the assessed observations had actual extents of $\leq$1 ha, which on average were 4 to nearly 8 orders of magnitude smaller than the primary measure of extent (Fig 3A).
-These differences declined linearily with actual extent, becoming approximately equal for the 5\% of observations having $\geq$100,000 ha actual extent.
-The actual duration of 64\% of observations was $<$1 day, which on average was 3 to nearly 9 orders of magnitude shorter than the time span covered by the conventional metric of duration (Fig.
+Extent was on average 5.6 orders of magnitude larger than actual extent, and this difference increased with extent, reaching a maximum of 8 orders of magnitude between 100 million and 1 billion ha of extent (Fig.
+3A).
+The difference fell to 3 orders of magnitude at 10 billion ha, a domain that was covered by $<$2\% of observations that were primarily collected with remote sensing.
+Remote observations had the smallest mean difference magnitude (1.9), whereas differences for the other three methods averaged at least 5.7 orders of magnitude larger (Fig.
+S3 in SI).
+
+The magnitude of difference between the duration and actual duration of observations was somewhat smaller, averaging 3.4 across all observations, ranging from just under 2 for the shortest durations (hour-day) to over 4 for observations lasting 1 decade to 1 century (Fig.
 3B).
-As with extent, the average magnitude of difference declined steadily with actual duration, becoming effectively equal for the 16\% of observations with $\geq$1 month of actual duration.
-
-%\noindent \textbf{\emph{Observational methods}}\\
-%We classified the method used to collect each observation into several broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing, other geographic data, and paleo-reconstruction approaches.
-Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.4\%, and paleo-reconstruction and other geographic data each for less than 1\% each.
+As with extent, the difference fell substantially for the longest durations (century to 10,000 years), as these domains were covered by paleo-reconstructions (Fig.
+S3 in SI), which have effectively no difference between duration and actual duration because the coring techniques they are typically based on capture continuous temporal records.
+The mean difference magnitudes for the other three observing methods ranged from just over 3 for field and automated sensing methods to nearly 6 for remote observations (Fig.
+S3 in SI).
 

### 2017-12-19 paper/SI.tex
Revising main text to include information on observational type distributions, and updated figure 1 vignette to produce percentage of observations per bin by observations type
Finished revision to results in main text, moved trend analysis discussion mostly to SI

@@
 The following text provides additional information on this study's Methods and Results.
-The full manuscript, data, and code are available on an accompanying \href{https://github.com/agroimpacts/ecoscales}{github repository}.
+The full manuscript, data, and code for all calculations are available on an accompanying \href{https://github.com/agroimpacts/ecoscales}{github repository}.
 \textbf{Note to editors and reviewers: this repository is currently private, but will be made public upon acceptance of manuscript.
@@
 \vspace{-10pt}
-Figures \ref{type_by_yr} and \ref{sc_by_yr} respectively show how the frequency of ecological observing methods and the average scale in each dimensions varies by publication year (2004-2014).
+To evaluate the potential impact that excluding studies from 2015-2017 would have on our findings, we analyzed the trends in 1) ecological observing methods, and 2) typical scales of ecological observations over the 10 year period.
+
+We used linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period.
+The results (Fig.
+\ref{type_by_yr}) suggest an increase (by 1.3\% per year) in the use of remote sensing and a corresponding decline in field methods (by the same percentage) during the 2004-2014 time period, although neither trend is statistically significant (remote observation regression, p$<$0.12; field observation regression, p$<$0.18).
+Automated sensing methods showed no visible trend.
+
+If the trend towards increasing use of remote sensing between 2004-2014 was real and not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average.
+
+Of course, we cannot attach confidence to a trend in observational extent extrapolated from a non-significant trend in observational type.
+However, some additional evidence in support of this possible trend lies within the extent values themselves, which appeared to increase by 0.25 orders of magnitude per year (Fig.
+\ref{sc_by_yr}) between 2004-2014 (R$^2$ = 0.25, p$<$0.07).
+This somewhat clearer (albeit still not significant at the customary p$<$0.05 threshold) trend suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).
+
+The trend calculations described above can be found in the ``Additional Analyses" vignette in the accompanying \href{https://github.com/agroimpacts/ecoscales}{R package/code repository}.
 
@@
 \includegraphics[width=0.9\textwidth]{../vignettes/figures/figS1.png}
-\vspace{-20 pt}
+\vspace{-5 pt}
 \caption{Histograms of the resolution, extent, interval, and duration of observational scales, shown for all observational methods (top row) and for the four main observational methods (rows 2-5).

### 2017-12-19 paper/main.tex
Finished revision to results in main text, moved trend analysis discussion mostly to SI

@@
 3A).
-The difference fell to 3 orders of magnitude at 10 billion ha, a domain that was covered by $<$2\% of observations that were primarily collected with remote sensing.
-Remote observations had the smallest mean difference magnitude (1.9), whereas differences for the other three methods averaged at least 5.7 orders of magnitude larger (Fig.
-S3 in SI).
+The difference fell to 3 orders of magnitude at 10 billion ha, a domain that was covered by the $<$2\% of observations that were primarily collected with remote sensing.
+Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (see Fig.
+S3 in SI for method-specific plots).
 
@@
 As with extent, the difference fell substantially for the longest durations (century to 10,000 years), as these domains were covered by paleo-reconstructions (Fig.
-S3 in SI), which have effectively no difference between duration and actual duration because the coring techniques they are typically based on capture continuous temporal records.
-The mean difference magnitudes for the other three observing methods ranged from just over 3 for field and automated sensing methods to nearly 6 for remote observations (Fig.
-S3 in SI).
+S3 in SI), which have effectively no difference between duration and actual duration because coring techniques capture continuous temporal records.
+The mean difference magnitudes for the other three observing methods ranged from just over 3 (field observations and automated sensing) to nearly 6 (remote observations).
 
@@
 
+The relatively small sample size of our study may also bias these findings.
+Although we used a randomized title draw to obtain a representative sample, we reviewed just 0.8\% of papers published during the 10 year period.
+Our sample may therefore under- or over-represent observational coverage in certain scale domains, particularly for analyses broken down by observational method.
+The case where this is of most concern is for paleo-reconstructions, the small sample size of which resulted in a likely over-estimate of the typical extent of such observations (e.g.
+Fig.
+2B, C; however, the interval and duration findings are probably more representative).
+
 Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates.
-
-Using linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period, the use of remote sensing appeared to increase by 1.3\% per year from 2004-2014 (R$^2$ = 0.25, p$<$0.12), and field methods declined by the same percentage (R$^2$ = 0.1, p$<$0.18), although both slopes failed to meet the customary threshold for statistical significance.
-Automated sensing methods showed no trend over time (SI).
-
-Indeed, if the trend towards increasing use of remote sensing between 2004-2014 was not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average (see SI for details of calculation).
-Further evidence for this trend lies within the extent values themselves, which increased 0.25 orders of magnitude per year between 2004-2014 (R$^2$ = 0.25, p$<$0.07).
-This somewhat clearer trend also suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).
+To evaluate this possibility, we used linear regressions to examine whether the relative frequency of observing methods changed during the 10 year study period, and whether our findings regarding scale domains might be different if we included more recent studies.
+Although our sample size was too small to assign statistical significance, we found a possible positive trend in the use of remote observations and a corresponding decline in field observations over the 10 year period.
+If these trends were not spurious, they suggest that including studies from 2015-2017 would result in a somewhat larger relative sample of remote observations, which would in turn increase slightly the mean extent of ecological observations (see SI for details of trend analyses).
 

### 2017-12-19 paper/SI.tex
More edits to SI on trends section, and additional edits to main text

@@
 \subsubsection*{Trends in methods and scale}
-To assess the degree to which observational methods changed over the course of the 10 years worth of studies we surveyed, we calculated the percentage of observations made using remote sensing, general field methods, and automated \emph{in situ} methods, and fit a linear regression between these percentages and publication year, weighting the regression by the total number of observations in each year.
-We performed the same analysis for the four primary dimensions (resolution, extent, interval, and duration) in order to assess whether there were any trends in observational scales.
+To evaluate the potential impact that excluding studies from 2015-2017 would have on our findings, we analyzed the trends in 1) ecological observing methods, and 2) typical scales of ecological observations over the 10 year period.
+To undertake the former assessment, we calculated the percentage of observations made using remote sensing, general field methods, and automated \emph{in situ} methods, and fit a linear regression between these percentages and publication year, weighting the regression by the total number of observations in each year.
+For the second analysis, we applied the same regression analysis to the four primary dimensions (resolution, extent, interval, and duration) in order to assess whether there were any trends in observational scales.
+
+The regressions and resulting code for trend extrapolations can be found in the ``Additional Analyses" vignette in the accompanying \href{https://github.com/agroimpacts/ecoscales}{R package/code repository}.
 
 \subsubsection*{Extracting and analyzing data from earlier meta-analyses}
-To compare the results of our analysis with the observational scales of earlier ecological studies, we used graph capture software\footnote{http://arohatgi.info/WebPlotDigitizer/app/} to extract the data values from Figure 6.1 in Tilman \cite{tilman_ecological_1989}, Figure 1 in Kareiva and Anderson \cite{kareiva_spatial_1988}, and Figure 2 in Porter et al \cite{porter_crop_2005}.
+To compare the results of our analysis with the observational scales of earlier ecological studies, we used graph capture software\footnote{http://arohatgi.info/WebPlotDigitizer/app/} to extract the data values from Fig.
+6.1 in Tilman \cite{tilman_ecological_1989}, Fig.
+1 in Kareiva and Anderson \cite{kareiva_spatial_1988}, and Fig.
+2 in Porter et al \cite{porter_crop_2005}.
 To maintain as much comparability as possible with our inclusion criteria, we excluded experiment studies in Tilman's data, as well as the values of any studies having duration values greater than 100 years (no upper time bound was provided for these), leaving duration values for 419 (out of 623) studies.
-As Tilman presented duration values as a histogram, we calculated the mean duration across all studies as the weighted (by number of observations per bin) mean of bin center-point values (i.e.
+Since Tilman presented duration values as a histogram, we calculated the mean duration across all studies as the weighted (by number of observations per bin) mean of bin center-point values (i.e.
 the weighted mean of the bin means).
@@
 Kareiva and Anderson presented resolution as plot diameters (m), which we squared to make comparable to our resolution metric.
+
+Calculations of scale values from these studies can be found in the ``Additional Analyses" vignette in the accompanying \href{https://github.com/agroimpacts/ecoscales}{R package/code repository}.
 
@@
 
-\subsubsection*{Trends in observational methods and scales by year}
+\subsubsection*{Trends in methods and scales}
 \vspace{-10pt}
-To evaluate the potential impact that excluding studies from 2015-2017 would have on our findings, we analyzed the trends in 1) ecological observing methods, and 2) typical scales of ecological observations over the 10 year period.
-
-We used linear regression (weighted by the number of observations per publication year) to assess whether the relative frequency of observing methods changed during the 10 year study period.
-The results (Fig.
+The results of weighted linear regression applied to observational types by year (Fig.
 \ref{type_by_yr}) suggest an increase (by 1.3\% per year) in the use of remote sensing and a corresponding decline in field methods (by the same percentage) during the 2004-2014 time period, although neither trend is statistically significant (remote observation regression, p$<$0.12; field observation regression, p$<$0.18).
@@
 Of course, we cannot attach confidence to a trend in observational extent extrapolated from a non-significant trend in observational type.
-However, some additional evidence in support of this possible trend lies within the extent values themselves, which appeared to increase by 0.25 orders of magnitude per year (Fig.
+However, some additional evidence in support of this possible trend lies within the extent values themselves.
+The regression analyses applied to dimensions shows an apparent increase in extent by 0.25 orders of magnitude per year (Fig.
 \ref{sc_by_yr}) between 2004-2014 (R$^2$ = 0.25, p$<$0.07).
 This somewhat clearer (albeit still not significant at the customary p$<$0.05 threshold) trend suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).
-
-The trend calculations described above can be found in the ``Additional Analyses" vignette in the accompanying \href{https://github.com/agroimpacts/ecoscales}{R package/code repository}.
 

### 2017-12-19 paper/main.tex
More edits to SI on trends section, and additional edits to main text

@@
 
-%This lower right to upper left orientation reflects the tradeoff between resolution and interval that is typical of satellite imaging \cite{estes_platform_2016}, and stands in contrast to the upper right to lower left line that stretches between this concentration and the high frequency (minute-hour intervals), high spatial resolution (0.1-100 cm$^2$) observations.
-This line demonstrates the opposite tradeoff that occurs with field-based observations, where larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
-
 %Comparing the spatial and temporal resolutions of  interval to spatial resolution of ecological observations reveals three primary regions of observational concentration (Fig.
@@
 2B).
-The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting one year to one decade tend to have yearly to decadal intervals.
+The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting 1 year to 1 decade tend to have yearly to decadal intervals.
 This tendency is reflected in the median domain locations of the primary observational methods: automated sensing had the finest median interval (hour-day) and shortest duration (month-year), followed by remote sensing (median interval slightly greater than one day and median duration of 1 year), field observations (median monthly interval and duration just over 1 year), and finally paleo-reconstructions (median interval 1 decade and millennial duration).
@@
 1C) and relatively short duration ($\leq$1 year; Fig.
-1D; 2D).
+1D).
+
 Contrasting observational dimensions reveals that larger extents are associated with larger plot sizes (Fig.
-2C), while longer durations are associated with longer intervals (Fig.
-2D).
-The latter association presumably reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of traditional field-based observation.
-The same tradeoff is also responsible for the inverse relationship between resolution and interval that dominates that domain space (Fig.
-2A).
-As a result of these tradeoffs, there are notable ``holes'' in the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($\geq$1 m$^2$ up to 100 ha; Fig.
+2C), while longer durations tend to be associated with longer intervals (Fig.
+2B).
+The latter association primarily reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of traditional field-based observation, the dominant observational mode, but is also reflected by the relative domain locations of the four main observational methods, suggesting that high observational frequency poses a durational cost across observational methods.
+A similar cost tradeoff is also evident in for the inverse relationship between resolution and interval that dominates that domain space (Fig.
+2A), which again primarily relates to characteristics of field-based observations, in which larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
+Less obvious is the opposite tradeoff that occurs with remote observation (but shown in Fig.
+S2 in SI), where finer resolution is desired to increase detail, but typically comes at the cost of longer intervals \cite{estes_platform_2016}.
+
+As a result of these aforementioned tradeoffs, there are several notable gaps in the spatio-temporal domains of ecological observations, which line in the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($\geq$1 m$^2$ up to 100 ha; Fig.
 2A) and 2) decadal or longer durations (Fig.
 2B).
+Another hole is evident in the high to moderate resolution, large extent (1 million-10 billion ha) domain (Fig.
+2C).
 
@@
 The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who present the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986.
-The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,496,069 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
+The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,496,070 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
 The third dataset is provided by Porter et al \cite{porter_wireless_2005}, who compared the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}).
@@
 a focus on experiments versus un-manipulated systems), these comparisons suggest that the duration and, less clearly, resolution of ecological observations have changed little in the past 30 years, but observational frequency and particularly extent have increased.
-The growth in observational extent is also reflected by the positive year-on-year trend within our own dataset, which corresponds to the increasing use of remote sensing (see figures in SI).
+A weak positive trend in our own data suggests that average extent of ecological observations is steadily increasing (Fig.
+S5 in SI), and likely corresponds to the increasing use of remote observation technologies (Fig.
+S4 in SI).
 
-However, even though observational extent is increasing, there remains for most observations a large gulf between the area that is actually sampled and that which the spatial replicates purportedly represent (Fig.
+However, even though observational extent appears to be increasing, there remains for most observations a large gulf between the area that is actually sampled and that which the spatial replicates purportedly represent (Fig.
 3A).
-A similarly large discrepancy is also evident between the time spanned by repeated observations and the time that is spent actually observing a phenomenon (Fig.
+A similarly large discrepancy is also evident between the time spans covered by repeated observations and the time that is spent actually observing a phenomenon (Fig.
 3B).

### 2017-12-19 paper/SI.tex
Basically finished revisions to main manuscript, including responses to reviewers (not versioned). Still need to make one last check through SI.

@@
 However, some additional evidence in support of this possible trend lies within the extent values themselves.
-The regression analyses applied to dimensions shows an apparent increase in extent by 0.25 orders of magnitude per year (Fig.
+The regression analyses applied to dimensions shows an apparent increase in extent of 0.25 orders of magnitude per year (Fig.
 \ref{sc_by_yr}) between 2004-2014 (R$^2$ = 0.25, p$<$0.07).
-This somewhat clearer (albeit still not significant at the customary p$<$0.05 threshold) trend suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).
+This weak trend suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).
 

### 2017-12-19 paper/main.tex
Basically finished revisions to main manuscript, including responses to reviewers (not versioned). Still need to make one last check through SI.

@@
 The scales at which ecosystems are observed plays a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
-Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992}.
+Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992,sandel_scale_2009}.
 This awareness has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad and increasing range of scales \cite{schneider_rise_2001, chave_problem_2013}.
 
-Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) meta-analyses of ecological research scales were published nearly 30 years ago \cite{tilman_ecological_1989,kareiva_spatial_1988}, it is both timely and important to assess the scales of contemporary ecological investigation.
+Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) meta-analyses \cite{tilman_ecological_1989,kareiva_spatial_1988} of ecological research scales were published nearly 30 years ago (but see \cite{porter_crop_2005,sandel_scale_2009} for more recent reviews), it is both timely and important to assess the scales of contemporary ecological investigation.
 To address this need, we quantified the spatial and temporal domains (here domain means the distribution of observations within the spectrum of one or more scale dimensions\footnote{This definition differs slightly from Wiens' \cite{wiens_spatial_1989}, who defined ``domain of scale'' as ``a portion of the scale spectrum within which process-pattern relationships are consistent regardless of scale.''}) of empirical observations (defined here as ecological observations collected under un-controlled or non-manipulated conditions) that were reported within recently (2004-2014) published ecological studies.
@@
 \noindent \textbf{\emph{Distributions within the four primary dimensions}}\\
-In terms of resolution, the majority (67\%) of observations (across all methods) were collected in plots of $<$1 m$^2$ resolution, 24\% were collected within plots of 1 m$^2$ up to 1 ha, and the remaining 9\% in plots of $\geq$1 ha (Fig.
+In terms of resolution, spatial replicates for the majority (67\%) of observations (across all methods) were $<$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $\geq$1 ha (Fig.
 1A).
@@
 This tendency is reflected in the median domain locations of the primary observational methods: automated sensing had the finest median interval (hour-day) and shortest duration (month-year), followed by remote sensing (median interval slightly greater than one day and median duration of 1 year), field observations (median monthly interval and duration just over 1 year), and finally paleo-reconstructions (median interval 1 decade and millennial duration).
-The low densities of observations having sub-daily intervals shows that relatively few high frequency, long duration ecological measurements are undertaken; this position
 
@@
 Three other notable, but lesser areas of concentration are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($<$1 day) of either 1-10 ha or 10,000-1,000,000 ha.
-The median observation (1 year duration, 100 ha extent) from automated sensing lies near the center of the second major concentration, while the median extents of remote (1-10 million ha) and field observations bound the primary concentration at its upper and lower extents, while the median duration of both observational types falls between 1 month to 1 year.
+The median observation (1 year duration, 100 ha extent) from automated sensing lies near the center of the second major concentration, while the median extents of field (1-10,000 ha) and remote (1-10 million ha) observations bound the primary concentration at its upper and lower extents, while the median duration of both observational types falls between 1 month to 1 year.
 
@@
 We selected this definition for the sake of consistency, but some papers reported resolution as a larger area in which sub-samples were taken.
-For these, our estimates are finer than what the studies' authors apparently considered to be plot resolution.
-Additionally, our domain estimates would presumably be somewhat different if we had included experimentally manipulated observations.
+For these, our estimates are finer than what the studies' authors apparently considered to be the resolution of spatial replicates.
+Our domain estimates would presumably be somewhat different if we had included experimentally manipulated observations.
 For example, average resolution and duration would likely be finer \cite{tilman_ecological_1989,kareiva_spatial_1988}.
+Additionally, the token 1 second (SI) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) mean that we underestimated actual duration-duration differences for this observational method (Fig.
+S3).
+However, the relatively low proportion of remote observations suggests that the impact on our overall findings was negligible (Fig.
+3).
 
@@
 To evaluate this possibility, we used linear regressions to examine whether the relative frequency of observing methods changed during the 10 year study period, and whether our findings regarding scale domains might be different if we included more recent studies.
-Although our sample size was too small to assign statistical significance, we found a possible positive trend in the use of remote observations and a corresponding decline in field observations over the 10 year period.
-If these trends were not spurious, they suggest that including studies from 2015-2017 would result in a somewhat larger relative sample of remote observations, which would in turn increase slightly the mean extent of ecological observations (see SI for details of trend analyses).
+Although our sample size was too small to assign statistical significance, we found a possible positive trend in the use of remote observations and a corresponding decline in field observations over the study period.
+If these trends were not spurious, they suggest that including studies from 2015-2017 would result in a somewhat larger relative sample of remote observations, which would in turn slightly increase the mean extent of ecological observations (see SI for details of trend analyses).
 
@@
 
-Contrasting observational dimensions reveals that larger extents are associated with larger plot sizes (Fig.
+Contrasting observational dimensions reveals that larger extents are associated with larger spatial replicates (Fig.
 2C), while longer durations tend to be associated with longer intervals (Fig.
@@
 A similar cost tradeoff is also evident in for the inverse relationship between resolution and interval that dominates that domain space (Fig.
-2A), which again primarily relates to characteristics of field-based observations, in which larger plot sizes demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
+2A), which again primarily relates to characteristics of field-based observations, in which larger spatial replicates demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
 Less obvious is the opposite tradeoff that occurs with remote observation (but shown in Fig.
@@
 
-As a result of these aforementioned tradeoffs, there are several notable gaps in the spatio-temporal domains of ecological observations, which line in the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($\geq$1 m$^2$ up to 100 ha; Fig.
+As a result of these aforementioned tradeoffs, there are several notable gaps in the spatio-temporal domains of ecological observations, which lie in the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($\geq$1 m$^2$ up to 100 ha; Fig.
 2A) and 2) decadal or longer durations (Fig.
@@
 a focus on experiments versus un-manipulated systems), these comparisons suggest that the duration and, less clearly, resolution of ecological observations have changed little in the past 30 years, but observational frequency and particularly extent have increased.
-A weak positive trend in our own data suggests that average extent of ecological observations is steadily increasing (Fig.
+A weak positive trend in our own data also suggests that average extent of ecological observations is steadily increasing (Fig.
 S5 in SI), and likely corresponds to the increasing use of remote observation technologies (Fig.
@@
 It is therefore important to examine whether the scales of the phenomena being observed are adequately captured by the design of replicates.
-Our methods suggest one possible procedure for assessing the \emph{scale representativeness} of replicates: 1) measure auto-correlation (spatial or serial) in the replicates, 2) add the autocorrelation length to the replicate area/duration, 3) calculate an autocorrelation-adjusted actual extent/duration, and 4) plot where it falls between actual extent/duration and extent/duration.
-The distance between the adjusted actual value and the ostensible value can provide a measure of how well the replicates represent the intended scale of observation.
+Our methods suggest one possible procedure for assessing the \emph{scale representativeness} of replicates, which is to i) calculate the autocorrelation (spatial or temporal) within the observations (using a semi-variogram, for example), ii) find the threshold distance (or time) below which a suitably strong correlation (e.g.
+r=0.7) will exist between neighboring sampled values, iii) add that distance (or time) to the sample resolution (or duration), and iv) recalculate actual extent (or duration) using the adjusted resolution (or sampling duration).
+The difference between this autocorrelation-adjusted actual extent (or actual duration) and extent (or duration) may provide a useful additional measure of how well the replicates represent the intended scale of observation.
 Though increasing spatial or temporal coverage may not always be the goal of a study (e.g., when spatial or temporal autocorrelation is a measure of interest), if the gap between actual and ostensible values remains large, then alternative sampling methods may be used to close it.
@@
 To mitigate this problem, we recommend that ecological journals require authors to quantify and clearly report the values of resolution, extent, interval, and duration.
+Fortunately, some journals already appear to be implementing such policies.
+For example, \emph{Global Ecology and Biogeography} now requires information on the spatial, temporal and taxonomic scale of studies to be in the abstract (a policy adopted in early 2016).
 
@@
 Letters in the plots denote the median values of different observational methods (f=field observations; a = automated sensing; r = remote sensing; p = paleo-observations).
-The grey shaded areas represent physically impossible domains (intervals greater than duration and resolutions greater than extent).}
+The grey shaded areas represent physically impossible domains (intervals greater than duration and resolutions greater than extent).
+Density values below the lower 3rd percentile fall within the darkest portion of the color gradient.}
 \label{afoto1}

### 2017-12-20 paper/SI.tex
Finished editing the SI

@@
 The following text provides additional information on this study's Methods and Results.
-The full manuscript, data, and code for all calculations are available on an accompanying \href{https://github.com/agroimpacts/ecoscales}{github repository}.
+The full manuscript, data, and code for all calculations are available as part of an R package that can be installed from an accompanying \href{https://github.com/agroimpacts/ecoscales}{GitHub repository}.
 \textbf{Note to editors and reviewers: this repository is currently private, but will be made public upon acceptance of manuscript.
@@
 We used the 2012 Web of Science impact factors to select the highest ranked ecology-themed journals that published studies with an observational component, excluding journals devoted to reviews, meta-analyses, laboratory, cellular, or experimental studies.
-To select a representative sample of recent ecology studies, we downloaded the metadata for all papers published in the selected journals (Table S1) between 2004 and 2014.
+To select a representative sample of recent ecology studies, we downloaded the metadata for all papers published in the selected journals (Table \ref{tabs1}) between 2004 and 2014.
 Our study involved six different observers (those reviewing the papers to extract the observational scales), each of whom was given a randomly selected batch of 500 titles.
-A separate set of 20 papers was also randomly selected, and this set was given to all observers to review independently, in order to 1) calibrate the interpretations and extraction of scale-related information between observers, and 2) to estimate between-observer variance.
+A separate set of 20 papers was also randomly selected, and was given to all observers to review independently in order to 1) calibrate the interpretations and extraction of scale-related information between observers, and 2) to estimate between-observer variance.
 
-\centerline{
-%\hspace{1cm} 
-\parbox{11cm}{\noindent Table 1. The selected journals and their 2012 impact factors.}
-}
+%\centerline{
+%%\hspace{1cm} 
+%\parbox{11cm}{\noindent Table S1. The selected journals and their 2012 impact factors.}
+%}
 \vspace{-12pt}
@@
 %\centering
-%\caption{The selected journals and their 2012 impact factors.}\\
+\caption{The selected journals and their 2012 impact factors.}\\
 %\centering
@@
 %\end{tabular}
+\label{tabs1}
 \end{longtable}
@@
 The space-related dimensions were resolution, extent, and actual extent.
-Here extent was primarily defined as the area falling within a perimeter defined by the outermost spatial replicates, while actual extent was defined as the summed area of all sample plots (i.e.
+Here extent was primarily defined as the area falling within a perimeter defined by the outermost spatial replicates, while actual extent was defined as the summed area of all spatial replicates (i.e.
 N * resolution, where N is the number of spatial replicates, which we also recorded), or the area that ecologists observe in practice.
@@
 A full definition of all dimensions and how they were recorded is contained within the answers to the list of questions below.
-This set of Frequently Asked Questions (FAQ) was provided to each observer for initial study and reference, in order to ensure methodological consistency (see next section).
+This set of Frequently Asked Questions (FAQ) was provided to each observer for initial study and reference, and adapted as necessary during the course of the study, in order to ensure methodological consistency (see next section).
 
@@
 
+\subsubsection*{Scale FAQ}
+\vspace{-10pt}
 \noindent\underline{\textbf{General:}}
@@
 \begin{enumerate}
-  \Hitem \emph{\textbf{What are the general inclusion/exclusion criteria for studies?}} Studies should be excluded from this analysis if they are: 1) opinion/perspectives pieces; 2) book reviews; 3) model-only studies, particularly theoretical models, which are not developed or tested against observed data; 4) if they are experimental manipulations (but if there is a study that has a mix of observational and experimental, record the observational treatments and exclude the manipulated treatments).
+  \Hitem \emph{\textbf{What are the general inclusion/exclusion criteria for studies?}} Studies should be excluded from this analysis if they are: 1) opinion/perspectives pieces; 2) book reviews; 3) model-only studies, particularly theoretical models that are not developed or tested against observations; 4) if they are experimental manipulations (but if a study has a mix of observational and experimental treatments, record the former but exclude the latter).
 
-  \Hitem \emph{\textbf{What are the standard categories to be used for defining “Study type”?}} Define study type according to the following categories: Remote sensing, passive/automated data collection, other geographic data (e.g.
-non-remotely sensed GIS data), field/direct observation, or paleo-reconstruction (tree rings, charcoal cores, etc).
+  \Hitem \emph{\textbf{What are the standard categories to be used for defining observational method?}} Define observational method according to the following categories: Remote sensing or other geographic data (e.g.
+non-remotely sensed GIS data), passive/automated data collection, field/direct observation, or paleo-reconstruction (tree rings, charcoal cores, etc).
 
@@
   \Hitem \emph{\textbf{What is sampling duration, and how do we record it?}} How long an individual observation of an individual point in space took to make.
-Sampling duration multiplied by the number of repeats observations is used to calculate study duration.
-Often this value will not be reported, so you will have to use your best judgement, based on your knowledge of ecological methods, to approximate the sampling duration.
-For example, for a field based method with intensive plot methods, if you can't estimate a plausible sampling duration, assign a token 1 day.
+Sampling duration multiplied by the number of repeat observations is used to calculate \emph{actual duration} (see Q9).
+This value will often not be reported, so you will have to use your best judgement, based on your knowledge of ecological methods, to approximate the sampling duration.
+For example, for a field based method with intensive plot methods, if not enough information is provided to estimate a plausible sampling duration, assign a token 1 day.
 For remote sensing observations you can assume one second (although the observations are effectively instantaneous).
@@
 More specifically, in the case of repeated observations, this is the total time that elapsed between the first and last observations at a given point in space (or of the same individual organism or community).
-For once-off (unreplicated) observations, this time is equivalent to the sample duration.
-However, there may be cases where once-off observations may have a longer duration than the sample duration.
+For once-off (unreplicated) observations, this time is equivalent to the sampling duration.
+However, there may be cases where once-off observations have a longer duration than the sample duration.
 For example, consider a study that counts occurrences of pollinators over three years, using transects that are located in different locations within the broader study area during each year \cite{rollin_differences_2013}.
@@
 In this case, we can consider the effective duration to be three years, as the temporal information is encoded in the analysis.
-  \Hitem \emph{\textbf{What is \emph{actual duration}, and how do we calculate it?}} The actual duration is the integral of sampling duration, or the time spent making one observation of the phenomenon in question.
-To clarify, actual duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample (duration), nor the integral of time spent in observing all spatial replicates.
+  \Hitem \emph{\textbf{What is \emph{actual duration}, and how do we calculate it?}} The actual duration is the integral of sampling duration (see Q7), or the time spent making one observation of the phenomenon in question.
+To clarify, actual duration is the total time spent sampling/observing a single point in space--not the span of time between first and last sample (duration), nor the integral of time spent in observing all spatial replicates (see Q10).
   \Hitem \emph{\textbf{Should actual duration be the total time spent sampling all sites or the amount of time spent sampling per site (e.g.
-for 5 minute point counts of birds at 10 sites each repeated twice, should we enter 100 minutes (5 minutes X 2 repeats X 10 sites) or 10 minutes (5 minutes X 2 repeats) for duration)?}} As stated above, actual duration is the total spent observing a single point in space, so in this case that would be 10 minutes (then converted to days, so 10 / (60 * 24))).
-  \Hitem \emph{\textbf{How do we record duration and actual duration when there are no repeat observations?}} In these cases, study duration is equal to sampling duration. 
+for 5 minute point counts of birds at 10 sites each repeated twice, should we enter 100 minutes (5 minutes X 2 repeats X 10 sites) or 10 minutes (5 minutes X 2 repeats) for duration)?}} As stated in Q9, actual duration is the total spent observing a single point in space, so in this case that would be 10 minutes (then converted to days, so 10 / (60 * 24))).
+  \Hitem \emph{\textbf{How do we record duration and actual duration when there are no repeat observations?}} In these cases, duration and actual duration are both equal to sampling duration. 
    \Hitem \emph{\textbf{How do I record interval in cases where the interval is inconsistent?
-For example, in a study were observations were observations were repeated in 1979, 1980, 1981, 1984, 2007, 2009?}} Find the time between each successive period, and then take the average of that (remember to convert to days!).
+For example, in a study where observations were repeated in 1979, 1980, 1981, 1984, 2007, 2009?}} Find the time between each successive period, and then take the average of that (remember to convert to days!).
 If there are two or more sets of unevenly spaced days for each site/plot/measurement being taken in the study, then find the average interval for each, and average the averages.
     \Hitem \emph{\textbf{How do you determine the interval for paleo-reconstructions?}} Use the minimum estimate for dating precision as the estimate of time between samples (e.g.
-50 years in the study of European charcoal deposits DOI:10.1111/geb.12090).
-    \Hitem \emph{\textbf{How do you determine the sample duration (our third time category) for paleo-reconstructions?}} Similar to the previous question, the sampling duration is also the same as the minimum estimate of dating precision.
+50 years in the study of European charcoal deposits \cite{molinari_exploring_2013}).
+    \Hitem \emph{\textbf{How do you determine the sample duration (our third time category) for paleo-reconstructions?}} Similar to the answer to the previous question, the sampling duration is also the same as the minimum estimate of dating precision.
 The logic behind this is that in such cases, where a sediment or tree core or similar measurement is being collected, this effectively represents a continuous ``observation'', and the value associated with the minimum (or other reported) interval is typically an average (or another summary statistic like the maximum) of the amount accumulated.
-    \Hitem \emph{\textbf{What about intervals and durations for instrument-collected, or automatically-collected, observations?}} These are often similar in essence to the paleo-reconstruction case.
+    \Hitem \emph{\textbf{What about intervals and durations for instrument-collected, or automated sensor-collected, observations?}} These are often similar to the paleo-reconstruction case.
 Take the minimum temperature or daily rainfall recorded at a weather station, which require constant observation over 24 hours to report.
 In such cases, the interval and sampling duration are both 24 hours.
-On the other hand, automated logging systems will provide a series of high frequency observations that are collected instantaneously.
+On the other hand, automated logging systems often provide a series of high frequency observations that are collected instantaneously.
 In these cases the sampling duration should be a token one second (to keep consistent with remote sensing [Q7]), and the interval should be the period between successive instantaneous measurements.
-    \Hitem \emph{\textbf{How do you treat interval for a case where repeated samples are taken during a season, across several seasons (e.g.``we performed repeat bird counts every 10 days between March and June of 2005, 2006, and 2008'')?}} Since the sampling is focused on seasons, and presumably some season-specific phenomenon (e.g.
+    \Hitem \emph{\textbf{How do you treat interval for a case where repeated samples are taken during a season, across several seasons (e.g.``we performed repeat bird counts every 10 days between March and June of 2005, 2006, and 2008'')?}} Since the sampling is focused on seasons, and presumably interested in some season-dependent phenomenon (e.g.
 breeding behavior), the reported values should be pegged to the season, not averaged across the duration (the start and end dates of the study).
-So in this example, it would be 10 days.
+So in this example it would be 10 days.
 
@@
 
-%But I'm noticing now that this isn't updated to include how to handle when the study unit is animals.
-I guess we could just add a sentence saying that the effective extent is the surface area of an average-sized individual of the animal species, or something along those lines.
-Let me know what you think.
-
-%  Broadly defined, study extent is the sampling resolution multiplied by the number of spatial replicates, divided by 10,000 to convert to hectares.
-For studies in which the spatial replicates are not spatially contiguous (as with most field-based studies), this means plot resolution (see Q16) multiplied by number of plots.
-For spatially contiguous studies (e.g.
-those based on remote sensing imagery), it should be the total area covered by the imagery, i.e.
-pixel resolution multiplied by the number of pixels.
-However, only record the area analyzed by the authors.
-If they used a sub-sample of pixels, extent is the number of those pixels a multiplied by pixel resolution.
-
     \Hitem \emph{\textbf{How do you determine resolution for paleo-reconstructions and other approaches where a sampling method is presumed to draw from a larger area (e.g.
@@
 To ensure consistency within the main analysis, the lead author also reviewed each observer's results from their individual draw of papers and flagged values that appeared to deviate from the protocol for re-review by the observer.
-Re-reviewed values were re-inspected, and in some cases a secondary review of particular papers was undertaken to cross-check the estimated scales.
+Revised values were re-inspected, and in some cases a secondary review of particular papers was undertaken to cross-check the estimated scales.
 
@@
 
-We used the between-observer coefficients of variation for each dimension as the basis for randomly perturbing the scale values for each of the 371 observations over 1000 iterations.
+We used the between-observer coefficients of variation for each dimension as the basis for randomly perturbing the scale values for each of the 378 observations over 1,000 iterations.
 For each observation at each iteration, we perturbed its scale value in each dimension by randomly selecting (from a uniform distribution) a percentage value p that fell between 100 + y and 100 - y, where y was the between-observer coefficient of variation (expressed as a percentage) for a given observational dimension, and multiplying that observation's scale value by that proportion.
-In certain cases, the perturbation resulted in values for some certain dimensions that were not physically possible (e.g.
-interval longer or actual duration longer than duration; actual extent larger than extent), in which case we adjusted the perturbed values for the dimension that should be the smallest to equal that of the one that should be largest (e.g.
+In certain cases, the perturbation resulted in values that were not physically possible (e.g.
+interval or actual duration longer than duration; actual extent larger than extent), in which case we adjusted the perturbed values for the dimension that should be the smallest to equal that of the one that should be largest (e.g.
 we set interval to equal the perturbed duration value).
@@
 In presenting results (Fig.
-1 and 2 in main text), we log-transformed (base 10) the observational scales within each of the 1000 perturbed ensemble members in order to account for the large range in scale values.
-We calculated histograms for the four primary observational dimensions from each of the log-transformed ensemble members, calculating the mean percentage density estimate for each bin across all 1000 histograms, as well as the upper and lower 2.5th percentile values for each bin (Fig.
+1 and 2 in main text), we log-transformed (base 10) the observational scales within each of the 1,000 perturbed ensemble members in order to account for the large range in scale values.
+We calculated histograms for the four primary observational dimensions from each of the log-transformed ensemble members, calculating the mean percentage density estimate for each bin across all 1,000 histograms, as well as the upper and lower 2.5th percentile values for each bin (Fig.
 1 main text).
@@
 To reveal the densities of observations within two dimensions (Fig.
-2 main text), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values across all ensemble members, using a bandwidth of 1 on a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations are concentrated.
-Bandwidths of varying resolutions were tested on kernel density estimates were tested on kernel density estimates of sampling interval versus plot resolution (see Supplementary Results section).
+2 main text), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values across all ensemble members, using a bandwidth of 1 on a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations were concentrated.
+Bandwidths of varying resolutions were tested on kernel density estimates of sampling interval versus plot resolution to test how sensitive our results were to the bandwidth value (see Supplementary Results section).
 
@@
 \noindent Where x is either extent or duration and y is its actual counterpart.
-We then evaluated how the magnitudes of difference varied with increasing values of actual extent/duration, using box plots to summarize decades within the same bins used for extent and duration (Fig.
+We then evaluated how the magnitudes of difference varied with increasing values of extent/duration, using box plots to summarize decades within the same bins used to summarize the frequency distributions of the extent and duration of observations (Fig.
 1B and D main text).
 Decades were calculated for each pair for all bootstrap replicates.
-The whiskers in the summary box plots (Fig.
-3 main text) fall slightly below 0 in some bins, even though extent and duration can never be smaller than their actual value.
-This is an artifact of the perturbation applied to estimated values, which in some instances caused extent/duration values to fall below their actual value.
-We retained these discrepancies rather than forcing such cases to zero to avoid biasing the summary statistics.
 
@@
 To undertake the former assessment, we calculated the percentage of observations made using remote sensing, general field methods, and automated \emph{in situ} methods, and fit a linear regression between these percentages and publication year, weighting the regression by the total number of observations in each year.
-For the second analysis, we applied the same regression analysis to the four primary dimensions (resolution, extent, interval, and duration) in order to assess whether there were any trends in observational scales.
+For the second analysis, we applied the same regression approach to the four primary dimensions (resolution, extent, interval, and duration), in order to assess whether there were any trends in observational scales.
 
@@
 We assessed variability and consistency between observers along multiple dimensions.
-First, we assessed how reliably observers agreed with respect to selecting or rejecting papers for review, using the R's Agreement package \cite{yu_agreement:_2012} to calculate Fleiss' kappa statistic \cite{fleiss_measuring_1971}, which was 0.72 (z = 12.5, p$<$0.000), indicating substantial and significant agreement between observers \cite{landis_measurement_1977}.
+First, we assessed tghe degree to which observers agreed with respect to selecting or rejecting papers for review, using R's Agreement package \cite{yu_agreement:_2012} to calculate Fleiss' kappa statistic \cite{fleiss_measuring_1971}, which was 0.72 (z = 12.5, p$<$0.000), indicating substantial and significant agreement between observers \cite{landis_measurement_1977}.
 
@@
 
-Finally, we calculated the coefficient of variation (CV) between observers' estimates of scales for each dimension, first across all observers' mean scale estimates, and then as the average CV among observers' estimates of each individual record (Table S2).
+Finally, we calculated the coefficient of variation (CV) between observers' estimates of scales for each dimension, first across all observers' mean scale estimates, and then as the average CV among observers' estimates of each individual record (Table \ref{tabs2}).
 CV values were estimated from all six observers for resolution, actual extent, and interval, from five observers (all but Choi) for duration, and from three observers (Elsen, Estes, and Treuer) for extent.
-The smaller numbers for duration and extent reflect that initial efforts were focused on estimating resolution, interval, and actual extent and actual duration, thus fewer observers were available to assess extent and extent.
+The smaller numbers for duration and extent reflect the fact that initial efforts were focused on estimating resolution, interval, and actual extent and actual duration.
+Fewer observers were available for subsequent efforts to assess duration and extent.
 
@@
 \vspace{-10pt}
-We used the maximum uncertainty values for each dimension from the inter-observer variability analysis (Table S2) to determine the bounds of the random perturbations applied to each record over the 1000 iteration resample.
+We used the maximum uncertainty values for each dimension from the inter-observer variability analysis (Table \ref{tabs2}) to determine the bounds of the random perturbations applied to each record over the 1000 iteration resample.
 
@@
 \captionsetup{width=1\linewidth}
-\caption{Estimates of variability calculated from each observer's estimates of the spatial and temporal scales of ecological observations reported within a common set of the 20 papers used for calibration.
-Variability is expressed as the coefficient of variation (CV; standard deviation divided by mean multiplied by 100) between each observer's overall mean, and as the mean CV of observers' estimates for individual records.
-}
+\caption{The between-observer variability of estimates of the spatial and temporal scales of ecological observations reported within the set of 20 papers used for calibration.
+Variability is expressed as the coefficient of variation (CV; standard deviation divided by mean multiplied by 100) between each observer's overall mean, and as the mean CV of observers' estimates for individual records.}
 \vspace{-4 pt}
@@
 \end{center}
-\label{default}
+\label{tabs2}
 \end{table}%
-
-%\begin{figure}[!ht]
-%%\begin{wrapfigure}{c}{1\textwidth}
-%\includegraphics[width=1\textwidth]{figures/res_v_extent_pbstrap2.pdf}
-%\vspace{-0.15 cm}
-%\caption{Two-dimensional kernel density estimates of observational densities within the domains defined by sampling interval and spatial resolution (left) and temporal duration and spatial extent (right).
-Density estimates were applied to the log-transformed values of each observational dimension after being randomly perturbed within uncertainty bounds drawn from an assessment of between-observer variability.
-Density estimates are rescaled to represent percentages.
-}
-%\label{afoto1}
-%\end{figure}
-%
 
@@
 Fig.
-\ref{obshists} details the distributions of observations by method along the four primary dimensions, while Fig.
-\ref{obskde} reveals the densities of observations by method within the dimensional juxtapositions shown in Fig.
+\ref{obshists} details the distributions of observations separated by observational method along the four primary dimensions.
+Fig.
+\ref{obskde} shows the densities of observations by observational method within the dimensional juxtapositions shown in Fig.
 2 in the main text.
@@
 \vspace{-10pt}
-The results of weighted linear regression applied to observational types by year (Fig.
-\ref{type_by_yr}) suggest an increase (by 1.3\% per year) in the use of remote sensing and a corresponding decline in field methods (by the same percentage) during the 2004-2014 time period, although neither trend is statistically significant (remote observation regression, p$<$0.12; field observation regression, p$<$0.18).
+The results of the weighted linear regression applied to observational types by year (Fig.
+\ref{type_by_yr}) suggest an increase (by 1.3\% per year) in the use of remote sensing and a corresponding decline in field methods (by the same percentage) during the 2004-2014 time period, although neither trend is statistically significant (p$<$0.12 for regression of remote observations; p$<$0.18 for field observation regression).
 Automated sensing methods showed no visible trend.
 
-If the trend towards increasing use of remote sensing between 2004-2014 was real and not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average.
-
-Of course, we cannot attach confidence to a trend in observational extent extrapolated from a non-significant trend in observational type.
-However, some additional evidence in support of this possible trend lies within the extent values themselves.
-The regression analyses applied to dimensions shows an apparent increase in extent of 0.25 orders of magnitude per year (Fig.
+If this trend towards increasing use of remote sensing between 2004-2014 was real and not spurious, we can project that a repeated study applied to papers published between 2004-2017 would find remote sensing used for 7.7\% of observations (a 22\% increase), which would increase mean extent by 17.4\% (95\% CI = -1.3-67\%;  or 0.07 orders of magnitude) above the 2004-2014 average.
+We cannot attach much confidence to this projections given that it derived from a non-significant trend; however, some support for this projection lies within the regression analysis applied to extent values themselves, which showed an apparent increase in extent of 0.25 orders of magnitude per year (Fig.
 \ref{sc_by_yr}) between 2004-2014 (R$^2$ = 0.25, p$<$0.07).
 This weak trend suggests that including more recent studies would have increased mean extent, but by a more modest 5.5\% (0.02 orders of magnitude).
+The other three dimensions did not show any clear year-on-year trends.
 
@@
 The smallest bandwidth (0.4) tested (Fig.
-\ref{ksens} top) shows that the primary observational concentrations described in the main text are evident, but tend to be devolved into separate concentrations.
-An example is the oblong concentration of observations that in Fig.
-1A (and in the lower left panel in Fig.
-\ref{ksens}) is bounded on the lower right at monthly to yearly intervals and 100-1000 m$^2$ resolutions and on the upper left by near-daily to monthly observations and 0.1-10 ha resolutions.
+\ref{ksens} top) shows the same primary observational concentrations revealed in Fig.
+1 in the main text, but these tend to be divided into separate sub-concentrations.
+An example is the oblong concentration of observations in Fig.
+2A (and in the lower left panel in Fig.
+\ref{ksens}) that is bounded on the upper left at monthly to yearly intervals and 100-1000 m$^2$ resolutions, and on the lower right by near-daily to monthly observations and 0.1-10 ha resolutions.
 With the smaller bandwidth this concentration appears as two separate patches (Fig.
@@
 \ref{ksens} middle left).
-This increasing separation of points into separate clusters is expected as bandwidth is reduced.
-However, this assessment indicates that the choice of bandwidth does not alter the interpretation of the results, as the domains of highest and lowest concentrations remain the same.
+It is expected that density clusters will increasingly separate as bandwidth is reduced.
+Nevertheless, this assessment indicates that the choice of bandwidth does not alter our primary interpretations, as the domains of highest and lowest concentrations remain the same regardless of bandwidth.
 

### 2017-12-22 paper/main.tex
Edited affiliations and made a few wording/typo changes suggested by Paul

@@
 \author
-{Lyndon Estes$^{\ast1, 2, 3}$, Paul R. Elsen$^{2, 4}$, Tim Treuer$^{5}$, Labeeb Ahmed$^{6}$, \\
-Kelly Caylor$^{3, 7}$, Jason Chang$^{6}$, Jonathan J. Choi$^{5}$, and Erle Ellis$^{6}$ \\
+{Lyndon Estes$^{\ast1, 2}$, Paul R. Elsen$^{3, 4}$, Tim Treuer$^{4}$, Labeeb Ahmed$^{5}$, \\
+Kelly Caylor$^{6, 7}$, Jason Chang$^{5}$, Jonathan J. Choi$^{5}$, and Erle Ellis$^{6}$ \\
 \\
@@
 \normalsize{$^{2}$Woodrow Wilson School, Princeton University, Princeton, NJ 08544, USA}\\
-\normalsize{$^{3}$Civil and Environmental Engineering, Princeton University, Princeton, NJ 08544, USA}\\
-\normalsize{$^{4}$Department of Environmental Science, Policy, and Management,}\\
+\normalsize{$^{3}$Department of Environmental Science, Policy, and Management,}\\
 \normalsize{University of California Berkeley, Berkeley, CA 94720, USA}\\
-\normalsize{$^{5}$Ecology and Evolutionary Biology, Princeton University, Princeton, NJ 08544, USA}\\
-\normalsize{$^{6}$Geography and Environmental Systems, University of Maryland Baltimore County,}\\ 
+\normalsize{$^{4}$Ecology and Evolutionary Biology, Princeton University, Princeton, NJ 08544, USA}\\
+\normalsize{$^{5}$Geography and Environmental Systems, University of Maryland Baltimore County,}\\ 
 \normalsize{Baltimore, MD 21250, USA}\\
-\normalsize{$^{7}$Earth Research Institute, University of California Santa Barbara, Santa Barbara, CA 93106, USA}\\
-\\
+\normalsize{$^{6}$Department of Geography, University of California Santa Barbara, Santa Barbara, CA 93106, USA}\\
+\normalsize{$^{7}$Bren School of Environmental Science and Management,}\\
+\normalsize{University of California Santa Barbara, Santa Barbara, CA 93106, USA}\\
 \normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  LEstes@clarku.edu.}}
@@
 \vspace{10pt}
-The scales at which ecosystems are observed plays a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
+The scales at which ecosystems are observed play a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
 Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992,sandel_scale_2009}.
@@
 Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) meta-analyses \cite{tilman_ecological_1989,kareiva_spatial_1988} of ecological research scales were published nearly 30 years ago (but see \cite{porter_crop_2005,sandel_scale_2009} for more recent reviews), it is both timely and important to assess the scales of contemporary ecological investigation.
-To address this need, we quantified the spatial and temporal domains (here domain means the distribution of observations within the spectrum of one or more scale dimensions\footnote{This definition differs slightly from Wiens' \cite{wiens_spatial_1989}, who defined ``domain of scale'' as ``a portion of the scale spectrum within which process-pattern relationships are consistent regardless of scale.''}) of empirical observations (defined here as ecological observations collected under un-controlled or non-manipulated conditions) that were reported within recently (2004-2014) published ecological studies.
+To address this need, we quantified the spatial and temporal domains of empirical observations that were reported within recently (2004-2014) published ecological studies.
+We define domain as the distribution of observations within the spectrum of one or more scale dimensions\footnote{This definition differs slightly from Wiens' \cite{wiens_spatial_1989}, who defined ``domain of scale'' as ``a portion of the scale spectrum within which process-pattern relationships are consistent regardless of scale.''}, and empirical observations as ecological observations collected under un-controlled or non-manipulated conditions.
 Empirical observations are critical for developing and testing the models that explain why ecological patterns vary in time and space \cite{levin_problem_1992, tilman_ecological_1989}, therefore the spatio-temporal domains of observations provide an important indicator of the field's progress towards achieving a holistic, predictive understanding of ecosystems \cite{chave_problem_2013,levin_problem_1992}.
@@
 The latter association primarily reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of traditional field-based observation, the dominant observational mode, but is also reflected by the relative domain locations of the four main observational methods, suggesting that high observational frequency poses a durational cost across observational methods.
-A similar cost tradeoff is also evident in for the inverse relationship between resolution and interval that dominates that domain space (Fig.
+A similar cost tradeoff is also evident in the inverse relationship between resolution and interval that dominates that domain space (Fig.
 2A), which again primarily relates to characteristics of field-based observations, in which larger spatial replicates demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
@@
 
-As a result of these aforementioned tradeoffs, there are several notable gaps in the spatio-temporal domains of ecological observations, which lie in the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($\geq$1 m$^2$ up to 100 ha; Fig.
+As a result of these tradeoffs, there are several notable gaps in the spatio-temporal domains of ecological observations, which lie in the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($\geq$1 m$^2$ up to 100 ha; Fig.
 2A) and 2) decadal or longer durations (Fig.

### 2017-12-22 paper/main.tex
Redid sensitivity analysis vignette in full

@@
 3).
-Extent was on average 5.6 orders of magnitude larger than actual extent, and this difference increased with extent, reaching a maximum of 8 orders of magnitude between 100 million and 1 billion ha of extent (Fig.
+Extent was on average 5.6 orders of magnitude larger than actual extent, and this difference increased with extent, reaching a maximum of 8.3 orders of magnitude between 100 million and 1 billion ha of extent (Fig.
 3A).
-The difference fell to 3 orders of magnitude at 10 billion ha, a domain that was covered by the $<$2\% of observations that were primarily collected with remote sensing.
+The difference fell to 3 orders of magnitude between 1-10 billion ha, a domain that was covered by the $<$2\% of observations that were primarily collected with remote sensing.
 Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (see Fig.

### 2017-12-22 paper/main.tex
Main and SI submissions, pacakge incremented to 0.4.0

@@
 Examining the distributions for each observational method (Fig.
-S1 in SI) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $<$0.1 m$^2$) than field observations (47\% $<$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $>$100 m$^2$).
+S1 in SI) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $<$0.1 m$^2$) than field observations (47\% $<$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $\geq$100 m$^2$).
 
-The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 12\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig.
+The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 11\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig.
 1B).
@@
 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals.
-With respect to temporally replicated observation (Fig.
+With respect to temporally replicated observations (Fig.
 S1 in SI), automated sensing techniques had the finest intervals (61\% $\leq$1 day; 100\% $\leq$1 year), followed by remote observation (37\% $\leq$1 day; 78\% $\leq$1 year), field observations (17\% $\leq$1 day; 86\% $\leq$1 year), and paleo-reconstructions (21\% $\leq$1 decade).
@@
 The four observational methods occupied substantially different portions of the domain space, as indicated by the locations of their median values (and illustrated further in Fig.
-S2 in the SI): the median domain of field observations was between 0.1-1 m$^2$ of resolution with a monthly interval, whereas remote observations had coarser median resolutions (1000 m$^2$) but finer median intervals ($\sim$1 day).
-Paleo-reconstructions and automated sensing techniques were both finely resolved (medians from 10 cm$^2$ to 0.01 m$^2$), but automated approaches had hourly-daily median intervals compared to multi-decadal for paleo-reconstructions.
-
-%Comparing the spatial and temporal resolutions of  interval to spatial resolution of ecological observations reveals three primary regions of observational concentration (Fig.
-2A), the greatest being unreplicated samples with resolutions of 100 cm$^2$ to 1 m$^2$, followed by observations of daily to monthly interval with slightly finer spatial resolutions (10 cm$^2$ to 1 m$^2$), and finally a smaller concentration of monthly to yearly measurements collected at spatial scales of 0.1-10 ha.
+S2 in the SI): the median domain of field observations was between 0.1-1 m$^2$ of resolution with a monthly interval, whereas remote observations had a coarser median resolution (1000 m$^2$) but finer median interval ($\sim$1 day).
+Paleo-reconstructions and automated sensing techniques were both finely resolved (median between10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.
 
@@
 
-Contrasting the two spatial dimensions against one another (for all observations) shows a primary concentration of observations with 10 cm$^2$ to 100 m$^2$ resolution with extents ranging from just over 1,000 to nearly 1,000,000 ha (Fig.
+Contrasting the two spatial dimensions against one another (for all observations) shows a primary concentration of observations of 10 cm$^2$ to nearly 100 m$^2$ resolution with extents ranging from $\sim$1,000 to nearly 1,000,000 ha (Fig.
 2C).
-The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha.
+The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $<$10 ha extent.
 These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Fig.
@@
 Two primary domains of observational concentration are revealed by juxtaposing duration and extent (across all observations).
-The first consists of observations spanning one month to one decade in time and 10-1,000 ha in space, while the second is defined by observations of one year to several decades that cover 10,000 to 1,000,000 ha (Fig.
+The first consists of observations spanning just over one month to one decade in time and 10-1,000 ha in space, while the second is defined by observations of one year to several decades that cover 10,000 to 1,000,000 ha (Fig.
 2D).
 Three other notable, but lesser areas of concentration are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($<$1 day) of either 1-10 ha or 10,000-1,000,000 ha.
-The median observation (1 year duration, 100 ha extent) from automated sensing lies near the center of the second major concentration, while the median extents of field (1-10,000 ha) and remote (1-10 million ha) observations bound the primary concentration at its upper and lower extents, while the median duration of both observational types falls between 1 month to 1 year.
+The median observation from automated sensing (1 year duration, 100 ha extent) lies near the center of the first major concentration, while the median extents of field (1-10,000 ha) and remote (1-10 million ha) observations bound the second major concentration at its upper and lower extents, with the median duration of both observational types falling between 1 month to 1 year.
 
@@
 Fig.
-2B, C; however, the interval and duration findings are probably more representative).
+2B, C; however, the interval and duration values are probably more representative).
 

### 2018-02-01 paper/SI.tex
First cut edits to shape paper to NEE formatting for publications (methods last, but reviewers' final suggestions).

@@
 The full manuscript, data, and code for all calculations are available as part of an R package that can be installed from an accompanying \href{https://github.com/agroimpacts/ecoscales}{GitHub repository}.
-\textbf{Note to editors and reviewers: this repository is currently private, but will be made public upon acceptance of manuscript.
-In the interim, the entire project folder can be downloaded from the following \href{https://www.dropbox.com/sh/1l9em6p0wyabwr3/AADhCpFvqjv6y7VbL5tgB3Iqa?dl=0}{link}.}
 
 \subsection*{Methods}% Make the title.
-\vspace{-4pt}
-%\maketitle 
-\subsubsection*{Paper selection}
-\vspace{-10pt}
-We used the 2012 Web of Science impact factors to select the highest ranked ecology-themed journals that published studies with an observational component, excluding journals devoted to reviews, meta-analyses, laboratory, cellular, or experimental studies.
-To select a representative sample of recent ecology studies, we downloaded the metadata for all papers published in the selected journals (Table \ref{tabs1}) between 2004 and 2014.
-Our study involved six different observers (those reviewing the papers to extract the observational scales), each of whom was given a randomly selected batch of 500 titles.
-A separate set of 20 papers was also randomly selected, and was given to all observers to review independently in order to 1) calibrate the interpretations and extraction of scale-related information between observers, and 2) to estimate between-observer variance.
 
-%\centerline{
-%%\hspace{1cm} 
-%\parbox{11cm}{\noindent Table S1. The selected journals and their 2012 impact factors.}
-%}
 \vspace{-12pt}
@@
 % Place your abstract within the special {sciabstract} environment.
-
-\subsubsection*{Estimating observational scales}
-\vspace{-10pt}
-
-Each observer first reviewed the papers in the calibration set, and then commenced reviewing papers in their individual random draws, beginning at the top of the list and then proceeding until at least 20 eligible papers describing ecological observations were reviewed.
-In cases where the reviewed papers used observations that were described in another publication, we reviewed those source papers in order to extract the observational dimensions.
-We excluded papers that were opinion or perspectives pieces (unless they presented or used existing observational data), theoretical studies based on generated data, or those which were entirely based on experimental manipulations.
-We left out the latter category because our intent was to evaluate the domains for observations of natural systems, and we wanted to avoid the bias that would be imposed by the relatively narrow spatial and temporal scales of experiments \cite{kareiva_spatial_1988,tilman_global_2011}.
-A bibliography of the reviewed papers follows the References and Notes section below.
-
-We recorded six primary dimensions of ecological observations, three related to space and three related to time.
-The space-related dimensions were resolution, extent, and actual extent.
-Here extent was primarily defined as the area falling within a perimeter defined by the outermost spatial replicates, while actual extent was defined as the summed area of all spatial replicates (i.e.
-N * resolution, where N is the number of spatial replicates, which we also recorded), or the area that ecologists observe in practice.
-In assessing spatial scales, our analysis only considered the Cartesian plane.
-We did not calculate the z, or depth, dimension, although this dimension is of greater importance for certain sub-disciplines of ecology (e.g.
-depth profiles in marine ecology).
-In some cases (primarily paleoecological studies), values extracted from the z-dimension provided temporal information that was used to calculate both the interval and the duration of the observation.
-
-For time dimensions, we extracted information related to interval, duration, and actual duration.
-Duration was defined as the time between the first and last temporal replicate, whereas actual duration quantifies the amount of time spent observing a particular location, which we calculated by multiplying sampling duration (the time spent collecting a single temporal replicate) by the number of temporal replicates.
-
-A full definition of all dimensions and how they were recorded is contained within the answers to the list of questions below.
-This set of Frequently Asked Questions (FAQ) was provided to each observer for initial study and reference, and adapted as necessary during the course of the study, in order to ensure methodological consistency (see next section).
-
-%To record the spatio-temporal dimensions of ecological observations, we considered the observational scales to be those at which a total measurement of the feature(s) of interest was made, rather than the scales that might be represented by the observations.
-As an example, if a paper reported data on plant species diversity collected using 10 1 m$^2$ sub-plots randomly placed within a larger plot of 100 m$^2$, of which 10 were placed, we recorded 1 m$^2$ as the spatial resolution and 100 m$^2$ as the total extent of the observation.
-
-%While the 10 sub-plots might be representative of diversity at larger scales (e.g.
-the particular plant community), that relationship depends both on sample design (number of replicates, placement, grain/resolution) and the properties of the ecosystem \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002}.
-Most studies also did not quantify the broader scales represented by their observations, nor did they report dimensions that would allow these to be independently estimated, such as the maximum distance between sample plots, which could have been used to estimate the spatial extent represented by the sample \cite{palmer_scale_1994}.
-
-%Similarly, we did not try to quantify the extent to which values of individual observations (e.g.
-a single plot) of an ecological feature might be correlated with spatially or temporally adjacent values.
-Autocorrelation is common to most ecological phenomena \cite{legendre_spatial_1993,collins_method_2000-1}, and it means that an observation can be representative of the properties within a broader area or time around the point at which it is taken.
-For example, soil properties are typically measured by analysis of soil cores collected using an augur.
-The two-dimensional resolution of this measurement is equal to the area of the augur, but the measured soil properties are likely to be very similar, if not identical, to those found in the soil within a radius of at least a few to several hundreds of meters around the core location.
-However, the distance and strength of this correlation can vary substantially according to the ecological features being observed, and such correlation lengths were reported in almost no studies.
-Attempting to estimate these lengths, and add them to actual observations scales, would have added further uncertainty to estimates.
-Moreover, this concern only applies to spatially and temporally discontinuous (or discontiguous) observations, and since many of observations are collected using techniques that permit continuous measurements (e.g.
-weather data over time or satellite images over space), we elected to use a consistent standard for estimating dimensions.
-
-%Furthermore quantifying the actual domains of observation is important because highly dynamic spatial and temporal processes, such as the population ecology of riverine fish \cite{fausch_landscapes_2002} or the evolution of wildfires \cite{clements_observing_2007}, may require continuous observation to properly measure, and can vary substantially in the space or time between observations.
 
@@
 
-\subsubsection*{Calibration and consistency}
-\vspace{-10pt}
-Most studies did not explicitly report values for all the assessed scales, and thus interpretation and judgement had to be applied to develop reasonable estimates for their values.
-The FAQ provided the protocol we followed, and was initially developed following consultation between observers prior to the commencement of review.
-We conducted an iterative process of calibration to ensure consistency and reliability of estimates.
-First, we used the calibration set to calculate between-observer variability with respect to paper selection/rejection and the estimation of scales.
-Based on this, the lead author reviewed individual records in each observer's calibration set, flagged values where the estimation procedure departed from the protocol, and returned these to observers for re-estimation, without providing an estimate of the actual value.
-Instead, the relevant section of the protocol was highlighted, and further explanation and clarifying discussion undertaken as needed.
-Protocol language was adjusted for clarity during this process, and new items added to cover circumstances that had not been addressed by the initial version.
-The variability measures (see Results) were re-calculated after each iteration.
-
-To ensure consistency within the main analysis, the lead author also reviewed each observer's results from their individual draw of papers and flagged values that appeared to deviate from the protocol for re-review by the observer.
-Revised values were re-inspected, and in some cases a secondary review of particular papers was undertaken to cross-check the estimated scales.
-
-\subsubsection*{Accounting for estimation uncertainty}
-\vspace{-10pt}
-Two major and related sources of uncertainty affected our estimation of observational scales: 1) unclear documentation of observational scales in the reviewed studies, 2) variation between observers in estimating observational scales (largely in cases where scales were not explicitly reported).
-We estimated and accounted for these uncertainties in several ways.
-First, we calculated the degree of between-observer variability based on each observer's results from reviewing the 20-paper calibration set.
-We calculated how well observers agreed regarding paper inclusion/exclusion, how many extractable observations there were per included paper, and what the coefficient of variation was across all observers' estimates of scale.
-We also recorded when observations were reported in any study (calibration set or otherwise) with an unclear or missing scale value.
-
-We used the between-observer coefficients of variation for each dimension as the basis for randomly perturbing the scale values for each of the 378 observations over 1,000 iterations.
-For each observation at each iteration, we perturbed its scale value in each dimension by randomly selecting (from a uniform distribution) a percentage value p that fell between 100 + y and 100 - y, where y was the between-observer coefficient of variation (expressed as a percentage) for a given observational dimension, and multiplying that observation's scale value by that proportion.
-In certain cases, the perturbation resulted in values that were not physically possible (e.g.
-interval or actual duration longer than duration; actual extent larger than extent), in which case we adjusted the perturbed values for the dimension that should be the smallest to equal that of the one that should be largest (e.g.
-we set interval to equal the perturbed duration value).
-This perturbed set of observations provided a basis for estimating uncertainty within our extracted set of observational scales.
-
-\subsubsection*{Scale metrics}
-\vspace{-10pt}
-In presenting results (Fig.
-1 and 2 in main text), we log-transformed (base 10) the observational scales within each of the 1,000 perturbed ensemble members in order to account for the large range in scale values.
-We calculated histograms for the four primary observational dimensions from each of the log-transformed ensemble members, calculating the mean percentage density estimate for each bin across all 1,000 histograms, as well as the upper and lower 2.5th percentile values for each bin (Fig.
-1 main text).
-
-To reveal the densities of observations within two dimensions (Fig.
-2 main text), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values across all ensemble members, using a bandwidth of 1 on a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations were concentrated.
-Bandwidths of varying resolutions were tested on kernel density estimates of sampling interval versus plot resolution to test how sensitive our results were to the bandwidth value (see Supplementary Results section).
-
-To compare the differences between actual extent and extent and actual duration and duration (Fig. 3 main text) we calculated the magnitude of difference (decade) between each pair as:
-
-\begin{equation}
-decade = log_{10}(x) - log_{10}(y) = log_{10}\left(\frac{x}{y}\right)
-\end{equation}
-
-\noindent Where x is either extent or duration and y is its actual counterpart.
-We then evaluated how the magnitudes of difference varied with increasing values of extent/duration, using box plots to summarize decades within the same bins used to summarize the frequency distributions of the extent and duration of observations (Fig.
-1B and D main text).
-Decades were calculated for each pair for all bootstrap replicates.
-
-\subsubsection*{Trends in methods and scale}
-To evaluate the potential impact that excluding studies from 2015-2017 would have on our findings, we analyzed the trends in 1) ecological observing methods, and 2) typical scales of ecological observations over the 10 year period.
-To undertake the former assessment, we calculated the percentage of observations made using remote sensing, general field methods, and automated \emph{in situ} methods, and fit a linear regression between these percentages and publication year, weighting the regression by the total number of observations in each year.
-For the second analysis, we applied the same regression approach to the four primary dimensions (resolution, extent, interval, and duration), in order to assess whether there were any trends in observational scales.
-
-The regressions and resulting code for trend extrapolations can be found in the ``Additional Analyses" vignette in the accompanying \href{https://github.com/agroimpacts/ecoscales}{R package/code repository}.
-
-\subsubsection*{Extracting and analyzing data from earlier meta-analyses}
-To compare the results of our analysis with the observational scales of earlier ecological studies, we used graph capture software\footnote{http://arohatgi.info/WebPlotDigitizer/app/} to extract the data values from Fig.
-6.1 in Tilman \cite{tilman_ecological_1989}, Fig.
-1 in Kareiva and Anderson \cite{kareiva_spatial_1988}, and Fig.
-2 in Porter et al \cite{porter_crop_2005}.
-To maintain as much comparability as possible with our inclusion criteria, we excluded experiment studies in Tilman's data, as well as the values of any studies having duration values greater than 100 years (no upper time bound was provided for these), leaving duration values for 419 (out of 623) studies.
-Since Tilman presented duration values as a histogram, we calculated the mean duration across all studies as the weighted (by number of observations per bin) mean of bin center-point values (i.e.
-the weighted mean of the bin means).
-We also excluded four (of 29) observation values from Porter et al's data on observational extent and frequency, which, in contrast to the other 25, were not randomly selected.
-Porter et al also used irregular scales for both their X (frequency) and Y (extent) axis, therefore we had to visually estimate the scale values for each data point after graphical extraction, and converted their extent values (in km) to hectares and their frequency values to intervals.
-Kareiva and Anderson presented resolution as plot diameters (m), which we squared to make comparable to our resolution metric.
-
-Calculations of scale values from these studies can be found in the ``Additional Analyses" vignette in the accompanying \href{https://github.com/agroimpacts/ecoscales}{R package/code repository}.
-
-\subsection*{Results}
+s\subsection*{Results}
 \vspace{-4pt}
@@
 We assessed variability and consistency between observers along multiple dimensions.
-First, we assessed tghe degree to which observers agreed with respect to selecting or rejecting papers for review, using R's Agreement package \cite{yu_agreement:_2012} to calculate Fleiss' kappa statistic \cite{fleiss_measuring_1971}, which was 0.72 (z = 12.5, p$<$0.000), indicating substantial and significant agreement between observers \cite{landis_measurement_1977}.
+First, we assessed the degree to which observers agreed with respect to selecting or rejecting papers for review, using R's Agreement package \cite{yu_agreement:_2012} to calculate Fleiss' kappa statistic \cite{fleiss_measuring_1971}, which was 0.72 (z = 12.5, p$<$0.000), indicating substantial and significant agreement between observers \cite{landis_measurement_1977}.
 

### 2018-02-01 paper/main.tex
First cut edits to shape paper to NEE formatting for publications (methods last, but reviewers' final suggestions).

@@
 \footskip 1.0cm
+
+\usepackage[
+  breaklinks=true,
+  colorlinks=true,
+  linkcolor=blue,anchorcolor=blue,
+  citecolor=blue,filecolor=blue,
+  menucolor=blue,pagecolor=blue,
+  urlcolor=blue]{hyperref}
 
@@
 
-\baselineskip24pt
+\baselineskip12pt
 
@@
 \section*{Abstract}
-\textbf{To properly understand ecological phenomena, it is necessary to observe them across a range of spatial and temporal scales.
-Ecologists first raised this point in the 1980s, and since then the ability to collect multi-scale observations has grown rapidly.
+\textbf{To understand ecological phenomena, it is necessary to observe their behavior across multiple spatial and temporal scales.
+This need was first highlighted in the 1980s, and since then technology has opened previously inaccessible scales to observation.
+Have there been corresponding changes in the scales observed by modern ecologists?
+To help answer this question, we analyzed the resolution, extent, interval, and duration of observations (excluding experiments) in 348 studies published between 2004-2014.
+We found that observational scales were generally narrow because ecologists still primarily use conventional field techniques.
+In the spatial domain, most observations had resolutions $\leq$1 m$^2$ and extents $\leq$10,000 ha.
+In the temporal domain, most observations were either unreplicated or infrequently repeated ($\geq$1 month interval), and $\leq$1 year in duration.
+Compared to studies conducted prior to 2004, observational durations and resolutions appear largely unchanged, but intervals have become finer and extents larger.
+We also found a large gulf between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, raising concerns about observational comprehensiveness.
+Furthermore, most studies did not clearly report scale, suggesting that it remains a minor concern.
+We suggest that ecologists can obtain better understanding of the scales represented by observations by incorporating autocorrelation measures, while journals can promote attentiveness to scale by implementing scale reporting standards.}
+
+%To fully understand ecological phenomena, we must observe them across multiple spatial and temporal scales.
+Beginning in the 1980s, several prominent ecological studies to this need beginning in the 1980s, and since then the ability to collect multi-scale observations has grown rapidly.
 To assess modern ecology's progress in addressing scale, we analyzed the resolution, extent, interval, and duration of observations in 348 observational studies published between 2004-2014.
-We found that the scale domains of observations are fairly narrow, and are collected primarily with conventional field techniques.
-In the spatial domain, most observations have resolutions of $\leq$1 m2 and extents of $\leq$10,000 ha.
-In the temporal domain, most observations were either unreplicated or of low frequency ($\geq$1 month interval), and were made over relatively short durations ($\leq$1 year).
-Compared to prior meta-analyses from the 1980s and early 2000s, observational durations and resolutions remain largely unchanged, but intervals have become finer and extents larger.
-Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, revealing portions of space and time not truly captured by replicates and raising concerns about observational comprehensiveness.
-Adding to these concerns, scales were not clearly reported in most studies, suggesting that it is a minor consideration.
+We found that most observations are collected with conventional field techniques, resulting in generally narrow observational scale domains.
+In the spatial domain, most observations have resolutions of $\leq$1 m$^2$ and extents of $\leq$10,000 ha.
+In the temporal domain, most observations were either unreplicated or of low frequency ($\geq$1 month interval), and had relatively short durations ($\leq$1 year).
+Compared to studies conducted prior to the early 2000s, observational durations and resolutions remain largely unchanged, but intervals have become finer and extents larger.
+Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, revealing portions of space and time not truly captured by replicates, which raises concerns about observational comprehensiveness.
+Adding to these concerns, scales were not clearly reported in most studies, suggesting that scale is a minor consideration.
 Journals can help mitigate this problem by implementing scale reporting standards, which can spur ecologists to more rapidly adopt new observational technologies, and thereby close key gaps in current observational domains.}
 
-\linenumbers
+%\linenumbers
 \vspace{10pt}
+\noindent \textbf{Introduction}
+\vspace{5pt}
+\\
 The scales at which ecosystems are observed play a critical role in shaping our understanding of their structure and function \cite{levin_problem_1992,chave_problem_2013,wiens_spatial_1989}.
@@
 
-Our analysis focused on two dimensions of spatial scale, resolution (grain) and extent, and two of temporal scale, interval and duration (Table 1, and see SI for full definitions).
-Resolution is the area of an individual spatial replicate within which a complete measurement (as opposed to a sub-sample) of the feature of interest was made.
-Extent is the area enclosed by the outer-most spatial replicates, or, if the system or habitat being sampled was distinct from its surrounding matrix (e.g.
-forest patches in grassland habitats), the summed area of sampled patches.
-Interval refers to the average time elapsed between individual temporal replicates.
-Duration measures the time elapsed between the first and last temporal replicates, or, in the case of temporally unreplicated observations, the estimated time spent collecting the observation.
+Our study focused on two dimensions of spatial scale, resolution (grain) and extent, and two of temporal scale, interval and duration (Table 1, and see Methods for full definitions).
 We also assessed observational scales within two additional dimensions, \emph{actual} extent (the summed area of spatial replicates) and \emph{actual} duration (the summed observational time of temporal replicates).
 We evaluated these additional dimensions to gain insight into how much the actual scales of observation (i.e.
-how much space and time is covered by the measurement) differ from the scales that the observations are explicitly or implicitly intended to represent.
+how much space and time is covered by the measurement) differ from the scales that the observations ostensibly represent.
 This difference may contain important information about how effectively ecological observations characterize ecological phenomena.
-First, an increasing gap between actual and intended observational scale inherently implies greater interpolation or extrapolation of observed measurements, raising the odds of over-leveraging data.
-Second (and relatedly), since natural systems are frequently complex, non-linear, and non-random \cite{levin_ecosystems_1998,pringle_spatial_2017,rietkerk_regular_2008}, a larger gap may increase the likelihood of encountering unanticipated data challenges such as censored data (\emph{sensu} \cite{efron_efficiency_1977}), as phenomena may resolve themselves in the space or time between observations.
+First, an increasing gap between actual and ostensible observational scale implies greater interpolation or extrapolation of observed measurements, raising the odds of over-leveraging data.
+Second (and relatedly), since natural systems are frequently complex, non-linear, and non-random \cite{levin_ecosystems_1998,pringle_spatial_2017,rietkerk_regular_2008}, a larger gap may increase the likelihood of encountering unanticipated data challenges such as censoring (\emph{sensu} \cite{efron_efficiency_1977}), as phenomena may resolve themselves in the space or time between observations.
 
@@
 
-%We extracted these scale dimensions from 379 observations that were reported in 134 papers that were published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals. 
-%
+We undertook three analyses to characterize the scale domains of modern ecological observations.
+We first evaluated the distribution of observations (domain) within each of the four primary dimensions.
+Next, we examined observational domains within four different two-dimensional juxtapostions (interval versus resolution; interval versus duration; extent versus resolution; extent versus duration).
+Lastly, we quantified the differences between actual and ostensible scales of observation (i.e.
+between actual extent and extent, and actual duration and duration), and examined how these differences varied within the domain of the corresponding ostensible scale.
 
-Our analysis was based on a review of 348 papers randomly selected from 42,918 published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals.
-We extracted scale data from 378 observations of ``natural'' (i.e.
+\vspace{10pt}
+\noindent \textbf{Results}
+\vspace{5pt}
+\\
+We reviewed 348 papers randomly selected from 42,918 titles published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals.
+We extracted and analyzed scale data from 378 observations of ``natural'' (i.e.
 non-experimentally manipulated) ecological features that were reported within 133 of the reviewed papers (plus an additional 62 that these cited as the source of observations).
-We excluded experiments because they tend to be of limited extent, duration, and resolution due to their higher logistical costs \cite{tilman_ecological_1989, kareiva_spatial_1988}, and would therefore likely bias our findings towards finer scales, while minimizing the impact that new observing methods (e.g.
-satellite imaging, wireless sensing) may have had in expanding the scales of ecological investigation \cite{turner_remote_2003,pettorelli_satellite_2014,porter_wireless_2005}.
+The majority of sampled observations were collected using manual field methods (80\%), followed by automated \emph{in situ} sensing techniques (12.4\%), remote sensing (6.9\%), and paleo-reconstruction ($<$0.8\%).
 
-To account for uncertainty in the estimation of observational dimensions due to 1) unclear methodological description in the reviewed papers, and 2) observer interpretation, we conducted a resampling analysis (n=1000) in which scale values were randomly perturbed within the bounds of estimated inter-observer variation (SI).
-We constructed histograms for each dimension from the mean of the perturbed ensembles, and estimated 95\% confidence intervals for each histogram bin (Fig.
-1).
-We constructed kernel density estimates from the full resampled ensemble in order to assess observational distributions within different juxtapositions of the four primary (resolution, extent, interval, duration) space-time dimensions (Fig.
-2).
-
-\noindent \textbf{Results}\\
-\noindent \textbf{\emph{Observational methods}}\\
-To account for potential differences in scales related to methodology, we classified each observation according to the following broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing/other geographic data (hereafter remote observations), and paleo-reconstruction approaches.
-Field methods were used for 80\% of observations, automated sensing for 12.4\%, remote sensing for 6.9\%, and paleo-reconstruction for less than 0.8\%.
-
+\vspace{5pt}
 \noindent \textbf{\emph{Distributions within the four primary dimensions}}\\
@@
 These distributions primarily reflect those of field observations, the dominant observational methodology.
-Examining the distributions for each observational method (Fig.
-S1 in SI) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $<$0.1 m$^2$) than field observations (47\% $<$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $\geq$100 m$^2$).
+Examining the distributions for each observational method (Supplementary Figure 1) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $<$0.1 m$^2$) than field observations (47\% $<$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $\geq$100 m$^2$).
 
@@
 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals.
-With respect to temporally replicated observations (Fig.
-S1 in SI), automated sensing techniques had the finest intervals (61\% $\leq$1 day; 100\% $\leq$1 year), followed by remote observation (37\% $\leq$1 day; 78\% $\leq$1 year), field observations (17\% $\leq$1 day; 86\% $\leq$1 year), and paleo-reconstructions (21\% $\leq$1 decade).
+With respect to temporally replicated observations (Supplementary Figure 1), automated sensing techniques had the finest intervals (61\% $\leq$1 day; 100\% $\leq$1 year), followed by remote observation (37\% $\leq$1 day; 78\% $\leq$1 year), field observations (17\% $\leq$1 day; 86\% $\leq$1 year), and paleo-reconstructions (21\% $\leq$1 decade).
 
@@
 
+\vspace{5pt}
 \noindent \textbf{\emph{Spatial and temporal domains}}\\
@@
 A less dense, oblong concentration of observations bounded on the upper left by monthly to yearly observations at 100 m$^2$ resolution and on the lower right by near-daily to monthly observations with 1-10 ha resolution is also evident.
-The four observational methods occupied substantially different portions of the domain space, as indicated by the locations of their median values (and illustrated further in Fig.
-S2 in the SI): the median domain of field observations was between 0.1-1 m$^2$ of resolution with a monthly interval, whereas remote observations had a coarser median resolution (1000 m$^2$) but finer median interval ($\sim$1 day).
-Paleo-reconstructions and automated sensing techniques were both finely resolved (median between10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.
+The four observational methods occupied substantially different portions of the domain space, as indicated by the locations of their median values (and illustrated further in Supplementary Figure 2): the median domain of field observations was between 0.1-1 m$^2$ of resolution with a monthly interval, whereas remote observations had a coarser median resolution (1,000 m$^2$) but finer median interval ($\sim$1 day).
+Paleo-reconstructions and automated sensing techniques were both finely resolved (median between 10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.
 
@@
 The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $<$10 ha extent.
-These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Fig.
-S2 in SI) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.
+These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Supplementary Figure 2) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.
 Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a result that likely reflects the very low sample size for this observational type.
@@
 
+\vspace{5pt}
 \noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
@@
 The difference fell to 3 orders of magnitude between 1-10 billion ha, a domain that was covered by the $<$2\% of observations that were primarily collected with remote sensing.
-Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (see Fig.
-S3 in SI for method-specific plots).
+Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (see Supplementary Figure 3 for method-specific plots).
 
@@
 3B).
-As with extent, the difference fell substantially for the longest durations (century to 10,000 years), as these domains were covered by paleo-reconstructions (Fig.
-S3 in SI), which have effectively no difference between duration and actual duration because coring techniques capture continuous temporal records.
+As with extent, the difference fell substantially for the longest durations (century to 10,000 years), as these domains were covered by paleo-reconstructions (Supplementary Figure 3 in SI), which have effectively no difference between duration and actual duration because coring techniques capture continuous temporal records.
 The mean difference magnitudes for the other three observing methods ranged from just over 3 (field observations and automated sensing) to nearly 6 (remote observations).
 
+\vspace{5pt}
 \noindent \textbf{\emph{Potential biases and uncertainties in quantifying scales}}\\
@@
 The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate, rather than simply record, these values for most observations (specifically, in 63\%, 60\%, and 69\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration, respectively).
-The inevitable estimation errors may have biased our overall findings.
-However, we attempted to quantify and account for this error by assessing inter-rater disagreement and incorporating this uncertainty into our resampling methodology.
+The inevitable estimation errors may have biased our findings.
+However, we attempted to quantify and account for this error by assessing between-observer variability and incorporating this uncertainty into our resampling methodology (Supplementary Results).
 The resulting confidence intervals (Fig.
@@
 For example, average resolution and duration would likely be finer \cite{tilman_ecological_1989,kareiva_spatial_1988}.
-Additionally, the token 1 second (SI) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) mean that we underestimated actual duration-duration differences for this observational method (Fig.
-S3).
+Additionally, the token 1 second (Supplemental Results) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) mean that we underestimated actual duration-duration differences for this observational method (Supplementary Figure 3).
 However, the relatively low proportion of remote observations suggests that the impact on our overall findings was negligible (Fig.
@@
 
-The relatively small sample size of our study may also bias these findings.
+It is also possible that our findings misrepresent the characteristic scales of ecological observations, due to sampling error.
 Although we used a randomized title draw to obtain a representative sample, we reviewed just 0.8\% of papers published during the 10 year period.
@@
 
-\noindent \textbf{Discussion}\\
+\vspace{10pt}
+\noindent \textbf{Discussion}
+\vspace{5pt}
+\\
 \noindent \textbf{\emph{Insights into the scale domains of modern ecology}}\\
@@
 2A), which again primarily relates to characteristics of field-based observations, in which larger spatial replicates demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
-Less obvious is the opposite tradeoff that occurs with remote observation (but shown in Fig.
-S2 in SI), where finer resolution is desired to increase detail, but typically comes at the cost of longer intervals \cite{estes_platform_2016}.
+Less obvious is the opposite tradeoff that occurs with remote observation (Supplementary Figure 2), where finer resolution is desired to increase detail, but typically comes at the cost of longer intervals \cite{estes_platform_2016}.
 
@@
 The first and most comprehensive dataset consists of duration values extracted by Tilman \cite{tilman_ecological_1989} from 623 studies published between 1977-1987 in the journal \emph{Ecology}.
-The average duration of the most comparable subset of those values (n=419; see SI) was 3.6 years, compared to 3.3 years for observations in our sample (or 5.1 if temporally un-replicated observations are excluded).
-The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who present the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986.
-The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,496,070 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
+The mean duration of the most comparable subset of those values (n=419; see SI) was 3.6 years, versus 3.3 years in our sample (or 5.1 if temporally un-replicated observations are excluded).
+The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who extracted resolutions from 97 community ecology experiments published in \emph{Ecology} between 1980-1986.
+The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,479,465 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
 The third dataset is provided by Porter et al \cite{porter_wireless_2005}, who compared the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}).
 The mean interval was 178 days, compared to 684 days in our sample, but the 80th percentile value in our study was 169 days compared to 329 days in theirs.
-Extent in our sample was substantially larger according to multiple summary statistics, including the mean (114,965,072 ha in our study versus 368,403 ha), median (5,051 ha versus 9 ha), and 95th percentile (46,424,808 ha versus 136,000 ha).
+Extent in our sample was substantially larger according to multiple summary statistics, including the mean (368,403 ha versus 114,965,072 ha in our study), median (9 ha versus 5,051 ha), and 95th percentile (136,000 ha versus 46,424,808 ha; this value is smaller than the mean, which is skewed by a small number of very large extent observations).
 
@@
 
-However, even though observational extent appears to be increasing, there remains for most observations a large gulf between the area that is actually sampled and that which the spatial replicates purportedly represent (Fig.
+However, even though observational extent appears to be increasing, there remains for most observations a large gulf between the area that is actually sampled and that which the spatial replicates ostensibly represent (Fig.
 3A).
@@
 3B).
-These differences between the actual and \emph{representative} scales of observation have implications for ecological understanding, as the unobserved portions of space and time may contain important patterns and processes that are not captured by replicates, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002, legendre_spatial_1993,collins_method_2000}.
+These differences between the actual and ostensible scales of observation have implications for ecological understanding, as the unobserved portions of space and time may contain important patterns and processes that are not captured by replicates, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002, legendre_spatial_1993,collins_method_2000}.
 Brief, infrequent snapshots, or fine-grained, spatially sparse replicates, may be sufficient to characterize many phenomena (as an example with respect to temporal replication, annual changes in tree cover are well-represented by low frequency satellite imaging \cite{hansen_high-resolution_2013}), but may be inadequate for more dynamic phenomena.
@@
 
+\vspace{5pt}
 \noindent \textbf{\emph{Looking forward}}\\
@@
 
+\vspace{10pt}
+\noindent \textbf{Methods}
+\vspace{5pt}
+\\
+\noindent \textbf{\emph{Paper selection and review}}\\
+We used the 2012 Web of Science impact factors to select the highest ranked ecology-themed journals that published studies with an observational component, excluding journals devoted to reviews, meta-analyses, laboratory, cellular, or experimental studies.
+To select a representative sample of recent ecology studies, we downloaded the metadata for all papers published in the 30 selected journals (Supplementary Table 1) between 2004 and 2014.
+Our study involved six different observers (those reviewing the papers to extract the observational scales), each of whom was given a randomly selected batch of 500 titles.
+A separate set of 20 papers was also randomly selected, and was given to all observers to review independently in order to 1) calibrate the interpretations and extraction of scale-related information between observers, and 2) to estimate between-observer variance.
+
+Each observer first reviewed the papers in the calibration set, and then commenced reviewing papers in their individual random draws, beginning at the top of the list and then proceeding until at least 20 eligible papers describing ecological observations were reviewed.
+In cases where the reviewed papers used observations that were described in another publication, we reviewed those source papers in order to extract the observational dimensions.
+We excluded papers that were opinion or perspectives pieces (unless they presented or used existing observational data), or theoretical studies based on generated data.
+We also did not collect scale information from papers (or the relevant parts of papers) describing experimental manipulations, because experiments tend to be of limited extent, duration, and resolution due to their higher logistical costs \cite{tilman_ecological_1989, kareiva_spatial_1988}.
+Including data from experiments would therefore likely have biased our findings towards finer scales, while minimizing the impact that new observing methods (e.g.
+satellite imaging, wireless sensing) may have had in expanding the scales of ecological investigation \cite{turner_remote_2003,pettorelli_satellite_2014,porter_wireless_2005}.
+A bibliography of the reviewed papers appears in the Supplementary Information.
+
+\vspace{5pt}
+\noindent \textbf{\emph{Estimating observational scales}}\\
+We recorded six primary dimensions of ecological observations, three related to space and three related to time.
+The space-related dimensions were resolution, extent, and actual extent.
+Here extent was primarily defined as the area falling within a perimeter defined by the outermost spatial replicates, while actual extent was  the summed area of all spatial replicates (i.e.
+N * resolution, where N is the number of spatial replicates, which we also recorded), or the area that ecologists observe in practice.
+In assessing spatial scales, our analysis only considered the Cartesian plane; we did not calculate the z, or depth, dimension, although this dimension is of greater importance for certain sub-disciplines of ecology (e.g.
+depth profiles in marine ecology).
+In some cases (primarily paleoecological studies), values extracted from the z-dimension provided temporal information that was used to calculate both the interval and the duration of the observation.
+
+For time dimensions, we extracted information related to observational interval, duration, and actual duration.
+Duration was defined as the time between the first and last temporal replicate, whereas actual duration quantifies the amount of time spent observing a particular location, which we calculated by multiplying sampling duration (the time spent collecting a single temporal replicate) by the number of temporal replicates.
+
+A full definition of all dimensions and how they were recorded is contained within a list of Frequently Asked Questions (FAQ; see Supplementary Methods), which was provided to each observer for initial study and reference, and adapted as necessary during the course of the study, in order to ensure methodological consistency.
+
+To account for potential differences in scales related to methodology, we classified each observation according to the following broad categories, which were field methods (manual \emph{in situ} data collection), automated (\emph{in situ}) sensing, remote sensing/other geographic data (hereafter remote observations), and paleo-reconstruction approaches.
+We also recorded when observations were reported in any study with an unclear or missing scale value.
+
+\vspace{5pt}
+\noindent \textbf{\emph{Calibration and consistency}}\\
+Most studies did not explicitly report values for all the assessed scales, and thus interpretation and judgement had to be applied to develop reasonable estimates for their values.
+The FAQ (Supplementary Methods) provided the protocol we followed, and was initially developed following consultation between observers prior to the commencement of review.
+We conducted an iterative process of calibration to ensure consistency and reliability of estimates.
+First, we used the calibration set to calculate between-observer variability with respect to paper selection/rejection and the estimation of scales.
+Based on this, the lead author reviewed individual records in each observer's calibration set, flagged values where the estimation procedure departed from the protocol, and returned these to observers for re-estimation, without providing an estimate of the actual value.
+Instead, the relevant section of the protocol was highlighted, and further explanation and clarifying discussion undertaken as needed.
+Protocol language was adjusted for clarity during this process, and new items added to cover circumstances that had not been addressed by the initial version.
+The variability measures were re-calculated after each iteration.
+
+To ensure consistency within the main analysis, the lead author also reviewed each observer's results from their individual draw of papers and flagged values that appeared to deviate from the protocol for re-review by the observer.
+Revised values were re-inspected, and in some cases a secondary review of particular papers was undertaken to cross-check the estimated scales.
+
+\vspace{5pt}
+\noindent \textbf{\emph{Scale estimation uncertainty}}\\
+Two major and related sources of uncertainty affected our estimation of observational scales: 1) unclear documentation of observational scales in the reviewed studies; 2) variation between observers in estimating observational scales (largely in cases where scales were not explicitly reported).
+To account for these uncertainties, we first quantified the between-observer variability in scale estimates (expressed as the coefficient of variation [CV]), which was constructed from each observer's final reported calibration set results.
+We then used the coefficients of variation for each dimension as the basis for randomly perturbing, over the course of 1000 iterations, the scale values for each of the sampled observations.
+For each observational dimension at each iteration, we perturbed its observer-estimated scale value by 1) randomly selecting (from a uniform distribution) a percentage value \emph{p} that fell between 100 + y and 100 - y (where y was the dimension-specific CV, expressed as a percentage), then 2) multiplied the scale value by the corresponding proportion (\emph{p} / 100).
+The perturbation occasionally resulted in physically impossible values (e.g.
+interval or actual duration longer than duration; actual extent larger than extent); in these cases we adjusted the perturbed values within the dimension that should be smallest to equal the value in the dimension that should be largest (e.g.
+we would set interval to equal duration, and not vice versa).
+We used the resulting set of perturbed observations to quantify uncertainty within our scale estimates.
+
+In addition to scale estimation CV, we also examined how well observers agreed regarding paper inclusion/exclusion, and how many extractable observations there were per included paper (see Supplementary Results).
+
+\vspace{5pt}
+\noindent \textbf{\emph{Analyses}}\\
+To characterize the scale domains of observations, we first log-transformed (base-10) the scale values within the 1,000 member perturbed ensemble, in order to account for the large range in values.
+To examine the distributions of observational scales within individual dimensions (Fig.
+1), we first constructed relative frequency histograms for each of the 1000 transformed ensemble members for each dimension, and then plotted the bin means across all members, as well as the upper and lower 2.5th percentile values for each bin.
+This produced a histogram of observational scales within each dimension that accounted for scale estimation uncertainty.
+
+To evaluate the distributions of observations within two scale dimensions (Fig.
+2), we used the splancs package \cite{rowlingson_splancs:_1993} of R \cite{r_development_core_team_r:_2011} to calculate a kernel density estimate of the log-transformed values across all ensemble members, using a bandwidth of 1 on a 0.1 resolution image to provide a smoothed result that served to more effectively highlight domains in which ecological observations were concentrated.
+Bandwidths of varying resolutions were tested on kernel density estimates of sampling interval versus plot resolution to test how sensitive our results were to the bandwidth value (see Supplementary Results section).
+
+To compare the differences between actual extent and extent and actual duration and duration (Fig. 3) we calculated the magnitude of difference (decade) between each pair as:
+
+\begin{equation}
+decade = log_{10}(x) - log_{10}(y) = log_{10}\left(\frac{x}{y}\right)
+\end{equation}
+
+\noindent Where x is either extent or duration and y is its actual counterpart.
+We then evaluated how the magnitudes of difference varied with increasing values of extent/duration, using box plots to summarize decades within the same bins used to summarize the frequency distributions of the extent and duration of observations (Fig.
+1B and D).
+Decades were calculated for each pair for all bootstrap replicates.
+
+\vspace{5pt}
+\noindent \textbf{\emph{Trends in methods and scale}}\\
+To evaluate the potential impact that excluding studies from 2015-2017 would have on our findings, we analyzed the trends in 1) ecological observing methods, and 2) typical scales of ecological observations over the 10 year period.
+To undertake the former assessment, we calculated the percentage of observations made using remote sensing, general field methods, and automated \emph{in situ} methods, and fit a linear regression between these percentages and publication year, weighting the regression by the total number of observations in each year.
+For the second analysis, we applied the same regression approach to the four primary dimensions (resolution, extent, interval, and duration), in order to assess whether there were any trends in observational scales.
+
+The regressions and resulting code for trend extrapolations can be found in the ``Additional Analyses" vignette in the accompanying \href{https://github.com/agroimpacts/ecoscales}{R package/code repository}.
+
+\vspace{5pt}
+\noindent \textbf{\emph{Extracting and analyzing data from earlier meta-analyses}}\\
+To compare the results of our analysis with the observational scales of earlier ecological studies, we used graph capture software\footnote{http://arohatgi.info/WebPlotDigitizer/app/} to extract the data values from Fig.
+6.1 in Tilman \cite{tilman_ecological_1989}, Fig.
+1 in Kareiva and Anderson \cite{kareiva_spatial_1988}, and Fig.
+2 in Porter et al \cite{porter_crop_2005}.
+To maintain as much comparability as possible with our inclusion criteria, we excluded experiment studies in Tilman's data, as well as the values of any studies having duration values greater than 100 years (no upper time bound was provided for these), leaving duration values for 419 (out of 623) studies.
+Since Tilman presented duration values as a histogram, we calculated the mean duration across all studies as the weighted (by number of observations per bin) mean of bin center-point values (i.e.
+the weighted mean of the bin means).
+We also excluded four (of 29) observation values from Porter et al's data on observational extent and frequency, which, in contrast to the other 25, were not randomly selected.
+Porter et al also used irregular scales for both their X (frequency) and Y (extent) axis, therefore we had to visually estimate the scale values for each data point after graphical extraction, and converted their extent values (in km) to hectares and their frequency values to intervals.
+Kareiva and Anderson presented resolution as plot diameters (m), which we squared to make comparable to our resolution metric.
+
+Calculations of scale values from these studies can be found in the ``Additional Analyses" vignette in the accompanying \href{https://github.com/agroimpacts/ecoscales}{R package/code repository}.
+
+\vspace{10pt}
+\noindent \textbf{Data availability}
+\vspace{5pt}
+\\
+The full data and code supporting this manuscript is available online at \href{https://github.com/agroimpacts/ecoscales}{https://github.com/agro\\impacts/ecoscales}.
+
+\vspace{10pt}
+\noindent \textbf{Acknowledgements} 
+\vspace{5pt}
+\\
+This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G).
+Erle Ellis, Jason Chang, and Labeeb Ahmed of the GLOBE Project (http://globe.umbc.edu) were supported by the U.S.
+National Science Foundation (1125210).
+Joshua Daskin, Jacob Socolar, and Charlotte Chang provided crucial help in developing and testing the sampling methodology.
+
+\vspace{10pt}
+\noindent \textbf{Author Contributions} 
+\vspace{5pt}
+\\
+L.E.
+and K.C.
+conceived the study.
+L.E., T.E., P.E., J.C., and E.E.
+helped design the sampling protocol.
+L.E., T.E., P.E., J.J.C., L.A, and J.C.
+extracted and analyzed scale data.
+L.E.
+performed analyses and wrote the manuscript, with assistance, comments, and edits provided by all authors.
+
+\vspace{10pt}
+\noindent \textbf{Declaration of Competing Financial Interests} 
+\vspace{5pt}
+\\
+The authors declare no competing financial interests.
+
 \bibliography{/Users/lestes/Dropbox/publications/fullbib}
@@
 
-\noindent \textbf{Acknowledgements} This work was supported by funds from the Princeton Environmental Institute Grand Challenges program and the NASA New Investigator Program (NNX15AC64G).
-Erle Ellis, Jason Chang, and Labeeb Ahmed of the GLOBE Project (http://globe.umbc.edu) were supported by the U.S.
-National Science Foundation (1125210).
 \clearpage

### 2018-02-02 paper/main.tex
Edited ms and SI for near final submission

@@
 {Lyndon Estes$^{\ast1, 2}$, Paul R. Elsen$^{3, 4}$, Tim Treuer$^{4}$, Labeeb Ahmed$^{5}$, \\
-Kelly Caylor$^{6, 7}$, Jason Chang$^{5}$, Jonathan J. Choi$^{5}$, and Erle Ellis$^{6}$ \\
+Kelly Caylor$^{6, 7}$, Jason Chang$^{5}$, Jonathan J. Choi$^{8}$, and Erle Ellis$^{5}$ \\
 \\
@@
 \normalsize{University of California Santa Barbara, Santa Barbara, CA 93106, USA}\\
+\normalsize{$^{8}$Nicholas School of the Environment, Duke University, Durham, NC 27708, USA}\\
 \normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  LEstes@clarku.edu.}}
@@
 \textbf{To understand ecological phenomena, it is necessary to observe their behavior across multiple spatial and temporal scales.
-This need was first highlighted in the 1980s, and since then technology has opened previously inaccessible scales to observation.
-Have there been corresponding changes in the scales observed by modern ecologists?
-To help answer this question, we analyzed the resolution, extent, interval, and duration of observations (excluding experiments) in 348 studies published between 2004-2014.
+Since this need was first highlighted in the 1980s, technology has opened previously inaccessible scales to observation.
+To help determine whether there have been corresponding changes in the scales observed by modern ecologists, we analyzed the resolution, extent, interval, and duration of observations (excluding experiments) in 348 studies published between 2004-2014.
 We found that observational scales were generally narrow because ecologists still primarily use conventional field techniques.
 In the spatial domain, most observations had resolutions $\leq$1 m$^2$ and extents $\leq$10,000 ha.
-In the temporal domain, most observations were either unreplicated or infrequently repeated ($\geq$1 month interval), and $\leq$1 year in duration.
+In the temporal domain, most observations were either unreplicated or infrequently repeated ($>$1 month interval), and $\leq$1 year in duration.
 Compared to studies conducted prior to 2004, observational durations and resolutions appear largely unchanged, but intervals have become finer and extents larger.
@@
 Furthermore, most studies did not clearly report scale, suggesting that it remains a minor concern.
-We suggest that ecologists can obtain better understanding of the scales represented by observations by incorporating autocorrelation measures, while journals can promote attentiveness to scale by implementing scale reporting standards.}
-
-%To fully understand ecological phenomena, we must observe them across multiple spatial and temporal scales.
-Beginning in the 1980s, several prominent ecological studies to this need beginning in the 1980s, and since then the ability to collect multi-scale observations has grown rapidly.
-To assess modern ecology's progress in addressing scale, we analyzed the resolution, extent, interval, and duration of observations in 348 observational studies published between 2004-2014.
-We found that most observations are collected with conventional field techniques, resulting in generally narrow observational scale domains.
-In the spatial domain, most observations have resolutions of $\leq$1 m$^2$ and extents of $\leq$10,000 ha.
-In the temporal domain, most observations were either unreplicated or of low frequency ($\geq$1 month interval), and had relatively short durations ($\leq$1 year).
-Compared to studies conducted prior to the early 2000s, observational durations and resolutions remain largely unchanged, but intervals have become finer and extents larger.
-Despite such gains, a large gulf exists between the scales at which phenomena are actually observed, and the scales those observations ostensibly represent, revealing portions of space and time not truly captured by replicates, which raises concerns about observational comprehensiveness.
-Adding to these concerns, scales were not clearly reported in most studies, suggesting that scale is a minor consideration.
-Journals can help mitigate this problem by implementing scale reporting standards, which can spur ecologists to more rapidly adopt new observational technologies, and thereby close key gaps in current observational domains.}
+Ecologists can better understand the scales represented by observations by incorporating autocorrelation measures, while journals can promote attentiveness to scale by implementing scale reporting standards.}
 
@@
 
-Our study focused on two dimensions of spatial scale, resolution (grain) and extent, and two of temporal scale, interval and duration (Table 1, and see Methods for full definitions).
-We also assessed observational scales within two additional dimensions, \emph{actual} extent (the summed area of spatial replicates) and \emph{actual} duration (the summed observational time of temporal replicates).
-We evaluated these additional dimensions to gain insight into how much the actual scales of observation (i.e.
-how much space and time is covered by the measurement) differ from the scales that the observations ostensibly represent.
-This difference may contain important information about how effectively ecological observations characterize ecological phenomena.
-First, an increasing gap between actual and ostensible observational scale implies greater interpolation or extrapolation of observed measurements, raising the odds of over-leveraging data.
-Second (and relatedly), since natural systems are frequently complex, non-linear, and non-random \cite{levin_ecosystems_1998,pringle_spatial_2017,rietkerk_regular_2008}, a larger gap may increase the likelihood of encountering unanticipated data challenges such as censoring (\emph{sensu} \cite{efron_efficiency_1977}), as phenomena may resolve themselves in the space or time between observations.
+Our study focused on two dimensions of spatial scale, resolution (grain) and extent, and two of temporal scale, interval and duration (Table 1).
+We analyzed the observational domains within each of these four dimensions and between pairs of these dimensions.
+We also assessed two additional dimensions, \emph{actual} extent (the summed area of spatial replicates) and \emph{actual} duration (the summed observational time of temporal replicates), which we used to evaluate how much the actual scales of observation (i.e.
+how much space and time is covered by the measurement) differ from the scales that they ostensibly represent.
+These differences may impact how effectively observations characterize ecological phenomena.
+For one, an increasing gap between actual and ostensible observational scales implies greater interpolation or extrapolation of observed measurements, raising the odds of over-leveraging data.
+Furthermore, since natural systems are frequently complex, non-linear, and non-random \cite{levin_ecosystems_1998,pringle_spatial_2017,rietkerk_regular_2008}, a larger gap increases the likelihood of data challenges such as censoring (\emph{sensu} \cite{efron_efficiency_1977}), as phenomena may resolve themselves in the space or time between replicates.
 
@@
 
-We undertook three analyses to characterize the scale domains of modern ecological observations.
-We first evaluated the distribution of observations (domain) within each of the four primary dimensions.
-Next, we examined observational domains within four different two-dimensional juxtapostions (interval versus resolution; interval versus duration; extent versus resolution; extent versus duration).
-Lastly, we quantified the differences between actual and ostensible scales of observation (i.e.
-between actual extent and extent, and actual duration and duration), and examined how these differences varied within the domain of the corresponding ostensible scale.
-
 \vspace{10pt}
@@
 \\
-We reviewed 348 papers randomly selected from 42,918 titles published between 2004-2014 in the top 30 (based on 2012 impact factor) ecology-themed journals.
-We extracted and analyzed scale data from 378 observations of ``natural'' (i.e.
-non-experimentally manipulated) ecological features that were reported within 133 of the reviewed papers (plus an additional 62 that these cited as the source of observations).
-The majority of sampled observations were collected using manual field methods (80\%), followed by automated \emph{in situ} sensing techniques (12.4\%), remote sensing (6.9\%), and paleo-reconstruction ($<$0.8\%).
+We reviewed 348 papers randomly selected from 42,918 titles published between 2004-2014 in the top 30 ecology-themed journals.
+We extracted scale data from 378 observations of ``natural'' (i.e.
+non-experimentally manipulated) ecological features reported within 133 of the reviewed papers (plus an additional 62 cited as the source of observations).
+The majority of sampled observations were collected using conventional field methods (80\%), followed by automated \emph{in situ} sensing techniques (12.4\%), remote sensing (6.9\%), and paleo-reconstruction ($<$0.8\%).
 
 \vspace{5pt}
-\noindent \textbf{\emph{Distributions within the four primary dimensions}}\\
-In terms of resolution, spatial replicates for the majority (67\%) of observations (across all methods) were $<$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $\geq$1 ha (Fig.
+\noindent \textbf{\emph{Observational domains within individual dimensions}}\\
+In terms of resolution, spatial replicates for the majority (67\%) of observations (across all methods) were $\leq$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $>$1 ha (Fig.
 1A).
 These distributions primarily reflect those of field observations, the dominant observational methodology.
-Examining the distributions for each observational method (Supplementary Figure 1) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $<$0.1 m$^2$) than field observations (47\% $<$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $\geq$100 m$^2$).
+Examining the distributions for each observational method (Supplementary Figure 1) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $\leq$0.1 m$^2$) than field observations (47\% $\leq$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $>$100 m$^2$).
 
-The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 11\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig.
+The extent of 19\% of observations was $\leq$10 ha, 23\% covered 10-1,000 ha, 11\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig.
 1B).
-As with resolution, the extent covered by automated sensing methods tended to be smaller (52\% $<$100 ha) than those of field observations (31\% $<$100 ha), while all but 4\% of remote observations covered areas $\geq$10,000 ha (as did the small number of paleo-reconstructions).
+As with resolution, the extent covered by automated sensing methods tended to be smaller (52\% $\leq$100 ha) than those of field observations (31\% $<$100 ha), while all but 4\% of remote observations covered areas $>$10,000 ha (as did the small number of paleo-reconstructions).
 
@@
 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals.
-With respect to temporally replicated observations (Supplementary Figure 1), automated sensing techniques had the finest intervals (61\% $\leq$1 day; 100\% $\leq$1 year), followed by remote observation (37\% $\leq$1 day; 78\% $\leq$1 year), field observations (17\% $\leq$1 day; 86\% $\leq$1 year), and paleo-reconstructions (21\% $\leq$1 decade).
+Among temporally replicated observations (Supplementary Figure 1), automated sensing had the finest intervals (61\% $\leq$1 day; 100\% $\leq$1 year), followed by remote observation (37\% $\leq$1 day; 78\% $\leq$1 year), field methods (17\% $\leq$1 day; 86\% $\leq$1 year), and paleo-reconstructions (21\% $\leq$1 decade).
 
-Duration was one day or less for 31\% of sampled observations (due to lack of temporal replication), while 10\% covered one day to one month, 23\% lasted one month to one year, 27\% covered 1-10 years, and 9\% spanned a decade or more (including several paleoecological studies covering centuries to millennia; Fig.
+Duration was $\leq$1 day for 31\% of sampled observations (due to lack of temporal replication), while 10\% covered one day to one month, 23\% one month to one year, 27\% 1-10 years, and 9\% a decade or more (several paleoecological studies covered centuries; Fig.
 1D).
-Paleo-reconstructions naturally had the longest duration (67\% $>$ 1 decade), while just $\sim$40\% of field, automated, and remote observations had durations of 1 year or longer.
+Paleo-reconstructions naturally had the longest duration (67\% $>$ 1 decade), while just $\sim$40\% of field, automated, and remote observations had durations exceeding one year.
 
 \vspace{5pt}
-\noindent \textbf{\emph{Spatial and temporal domains}}\\
-Juxtaposing these observational dimensions provides further insight into the spatial and temporal domains of observations (Fig.
-2).
-Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig.
+\noindent \textbf{\emph{Observational domains within two-dimensions}}\\
+Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack intervals) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig.
 2A).
 A less dense, oblong concentration of observations bounded on the upper left by monthly to yearly observations at 100 m$^2$ resolution and on the lower right by near-daily to monthly observations with 1-10 ha resolution is also evident.
-The four observational methods occupied substantially different portions of the domain space, as indicated by the locations of their median values (and illustrated further in Supplementary Figure 2): the median domain of field observations was between 0.1-1 m$^2$ of resolution with a monthly interval, whereas remote observations had a coarser median resolution (1,000 m$^2$) but finer median interval ($\sim$1 day).
+The four observational methods occupied substantially different domains, as indicated by the locations of their median values (and see Supplementary Figure 2): the median domain of field observations had 0.1-1 m$^2$ resolution and a monthly interval, whereas remote observations had a coarser median resolution (1,000 m$^2$) but finer median interval ($\sim$1 day).
 Paleo-reconstructions and automated sensing techniques were both finely resolved (median between 10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.
 
-Comparing the interval and duration of temporally replicated observations showed most observations had daily to decadal intervals and durations of $\geq$1 month up to 1 decade (Fig.
+Comparing the interval and duration of temporally replicated observations showed most observations had daily to decadal intervals and durations of 1 month to 1 decade (Fig.
 2B).
 The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting 1 year to 1 decade tend to have yearly to decadal intervals.
-This tendency is reflected in the median domain locations of the primary observational methods: automated sensing had the finest median interval (hour-day) and shortest duration (month-year), followed by remote sensing (median interval slightly greater than one day and median duration of 1 year), field observations (median monthly interval and duration just over 1 year), and finally paleo-reconstructions (median interval 1 decade and millennial duration).
+This tendency is reflected in the domain medians of the primary observational methods: automated sensing had the finest median interval (hour-day) and shortest duration (month-year), followed by remote sensing (median interval slightly greater than one day and median duration of 1 year), field observations (median monthly interval and duration just over 1 year), and finally paleo-reconstructions (median interval 1 decade and millennial duration).
 
-Contrasting the two spatial dimensions against one another (for all observations) shows a primary concentration of observations of 10 cm$^2$ to nearly 100 m$^2$ resolution with extents ranging from $\sim$1,000 to nearly 1,000,000 ha (Fig.
+Contrasting the two spatial dimensions against one another shows a primary concentration of observations of 10 cm$^2$ to nearly 100 m$^2$ resolution with extents ranging between $\sim$1,000-1,000,000 ha (Fig.
 2C).
-The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $<$10 ha extent.
-These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Supplementary Figure 2) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.
-Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a result that likely reflects the very low sample size for this observational type.
+The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $\leq$10 ha extent.
+These three concentrations suggest that observational extent increases with resolution, which is further evident in the median domain values (and kernel densities; Supplementary Figure 2) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.
+Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a possible artifact of small sample size.
 
-Two primary domains of observational concentration are revealed by juxtaposing duration and extent (across all observations).
-The first consists of observations spanning just over one month to one decade in time and 10-1,000 ha in space, while the second is defined by observations of one year to several decades that cover 10,000 to 1,000,000 ha (Fig.
+There are two primary observational domains within the contrast between duration and extent.
+The first consists of observations lasting one month to one decade with extents of 10-1,000 ha, while the second is defined by observations of one year to several decades that cover 10,000-1,000,000 ha (Fig.
 2D).
-Three other notable, but lesser areas of concentration are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($<$1 day) of either 1-10 ha or 10,000-1,000,000 ha.
-The median observation from automated sensing (1 year duration, 100 ha extent) lies near the center of the first major concentration, while the median extents of field (1-10,000 ha) and remote (1-10 million ha) observations bound the second major concentration at its upper and lower extents, with the median duration of both observational types falling between 1 month to 1 year.
+Three other notable, but lesser concentrations are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($\leq$1 day) of either 1-100 ha or 10,000-1,000,000 ha.
+The median observation from automated sensing (1 year duration, 100 ha extent) lies near the center of the first major concentration, while the median extents of field (1,000-10,000 ha) and remote (1-10 million ha) observations bound the second major concentration at its upper and lower extents, with the median duration of both observational types falling between 1 month to 1 year.
 
@@
 \noindent \textbf{\emph{Differences between actual and ostensible scales}}\\
-For our final analysis of observational domains, we evaluated the differences between the scales represented by the extent and duration dimensions and the scales that ecological observations actually cover.
-To make this assessment, we log$_{10}$ transformed and then subtracted the values of i) actual extent from extent and ii) actual duration from duration, yielding the magnitude of difference between each pair of dimensions for each observation.
-We plotted these values (summarized in box plots) against their corresponding extent/duration values to evaluate whether these differences varied with scale (Fig.
-3).
-Extent was on average 5.6 orders of magnitude larger than actual extent, and this difference increased with extent, reaching a maximum of 8.3 orders of magnitude between 100 million and 1 billion ha of extent (Fig.
+Observational extent was on average 5.6 orders of magnitude larger than actual extent (Fig.
 3A).
-The difference fell to 3 orders of magnitude between 1-10 billion ha, a domain that was covered by the $<$2\% of observations that were primarily collected with remote sensing.
-Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (see Supplementary Figure 3 for method-specific plots).
+This difference increased with extent, reaching a maximum of 8.3 between 100 million and 1 billion ha of extent, then falling to 3 orders of magnitude between 1-10 billion ha (these extents were covered by $<$2\% of observations, which were primarily collected with remote sensing).
+Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (Supplementary Figure 3).
 
-The magnitude of difference between the duration and actual duration of observations was somewhat smaller, averaging 3.4 across all observations, ranging from just under 2 for the shortest durations (hour-day) to over 4 for observations lasting 1 decade to 1 century (Fig.
+The magnitudes of difference between observational duration and actual duration was somewhat smaller, averaging 3.4 and ranging from just under 2 for the shortest durations (hour-day) to over 4 for observations lasting 1 decade to 1 century (Fig.
 3B).
-As with extent, the difference fell substantially for the longest durations (century to 10,000 years), as these domains were covered by paleo-reconstructions (Supplementary Figure 3 in SI), which have effectively no difference between duration and actual duration because coring techniques capture continuous temporal records.
-The mean difference magnitudes for the other three observing methods ranged from just over 3 (field observations and automated sensing) to nearly 6 (remote observations).
+As with extent, the difference fell substantially for the longest durations (century to millenia), as these domains were covered by paleo-reconstructions (Supplementary Figure 3), which have little difference between actual and ostensible duration because coring techniques capture continuous temporal records.
+The mean difference magnitudes for the other three observing methods ranged from just over 3 (field and automated sensing) to nearly 6 (remote observations).
 
@@
 \noindent \textbf{\emph{Potential biases and uncertainties in quantifying scales}}\\
-There were several potential methodological aspects that could have influenced our assessment of ecology's spatial and temporal domains.
-The first stems from our finding that many studies did not precisely report observational scales, which meant that we had to estimate, rather than simply record, these values for most observations (specifically, in 63\%, 60\%, and 69\% of cases for resolution, extent, and actual extent, and 36\%, 64\%, and 83\% of cases for interval, duration, and effective duration, respectively).
-The inevitable estimation errors may have biased our findings.
-However, we attempted to quantify and account for this error by assessing between-observer variability and incorporating this uncertainty into our resampling methodology (Supplementary Results).
+Our results were potentially influenced by several methodological issues.
+First, most studies did not precisely report observational scales, thus we had to estimate, rather than simply record, scale values for the majority of observations (we estimated 63, 60, 69, 36, 64, and 83\% of resolution, extent, actual extent, interval, duration, and actual duration values, respectively).
+Estimation errors may therefore have biased our findings.
+We attempted to quantify and account for this error by assessing between-observer variability and incorporating this uncertainty into our resampling methodology (Supplementary Results).
 The resulting confidence intervals (Fig.
-1) suggest that it was unlikely that estimation errors unduly influenced our findings.
+1) suggest that that estimation errors did not unduly influence our findings.
 
-Another potential source of bias lies within our scale-estimation protocols, chiefly with respect to our rule for estimating resolution (the smallest areal unit of \emph{complete} measurement).
+Our scale-estimation protocols may also have introduced bias, particularly our rule for estimating resolution (the smallest areal unit of \emph{complete} measurement).
 We selected this definition for the sake of consistency, but some papers reported resolution as a larger area in which sub-samples were taken.
-For these, our estimates are finer than what the studies' authors apparently considered to be the resolution of spatial replicates.
-Our domain estimates would presumably be somewhat different if we had included experimentally manipulated observations.
+For these, our estimates were finer than what the studies' authors considered to be the resolution.
+Our results would also  be somewhat different if we had included observations from experiments.
 For example, average resolution and duration would likely be finer \cite{tilman_ecological_1989,kareiva_spatial_1988}.
-Additionally, the token 1 second (Supplemental Results) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) mean that we underestimated actual duration-duration differences for this observational method (Supplementary Figure 3).
-However, the relatively low proportion of remote observations suggests that the impact on our overall findings was negligible (Fig.
-3).
+Additionally, the token 1 second (Supplemental Methods) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) caused us to underestimate the differences between their durations and actual durations (Supplementary Figure 3).
+However, the relatively small number of remote observations suggests that the impact of this bias on our overall findings was negligible.
 
-It is also possible that our findings misrepresent the characteristic scales of ecological observations, due to sampling error.
-Although we used a randomized title draw to obtain a representative sample, we reviewed just 0.8\% of papers published during the 10 year period.
-Our sample may therefore under- or over-represent observational coverage in certain scale domains, particularly for analyses broken down by observational method.
-The case where this is of most concern is for paleo-reconstructions, the small sample size of which resulted in a likely over-estimate of the typical extent of such observations (e.g.
+It is also possible that our findings misrepresent observational domains because of sampling error.
+Although we randomized our sample to ensure representativeness, we reviewed just 0.8\% of the papers published during our study period.
+Our sample may therefore under- or over-represent observational coverage in certain domains, particularly for specific methods.
+This possibility is greatest for paleo-reconstructions, where the small sample size likely resulted in an overestimate of typical observational extent (e.g.
 Fig.
@@
 
-Finally, because our review did not include papers beyond 2014, the omission of studies from the most recent years could have introduced bias into our domain estimates.
-To evaluate this possibility, we used linear regressions to examine whether the relative frequency of observing methods changed during the 10 year study period, and whether our findings regarding scale domains might be different if we included more recent studies.
-Although our sample size was too small to assign statistical significance, we found a possible positive trend in the use of remote observations and a corresponding decline in field observations over the study period.
-If these trends were not spurious, they suggest that including studies from 2015-2017 would result in a somewhat larger relative sample of remote observations, which would in turn slightly increase the mean extent of ecological observations (see SI for details of trend analyses).
+Finally, our omission of papers published after 2014 could also have biased our findings.
+Although our sample size was too small to assign statistical significance, we found a possible positive trend in the use of remote observations and a corresponding decline in field observations over the course of our study period.
+If these trends were not spurious, they suggest that including studies from 2015-2017 would result in a somewhat larger relative sample of remote observations, which would in turn slightly increase the mean extent of ecological observations (see Supplementary Results for further details).
 
@@
 \noindent \textbf{\emph{Insights into the scale domains of modern ecology}}\\
-Our results suggest that the scale domains of modern ecological observations are fairly narrow, and are collected primarily with conventional field techniques.
-Spatially, the majority of observations have grains of $\leq$1 m$^2$ and extents of $\leq$10,000 ha (Fig 1A;B).
-In the temporal domains, most observations are either un-replicated snapshots, or of low frequency ($\geq$1 month interval; Fig.
-1C) and relatively short duration ($\leq$1 year; Fig.
+Our results suggest that modern ecology's observational domains are fairly narrow, and that ecologists still primarily rely on conventional field-based observational techniques.
+In the spatial dimensions, the majority of observations have grains of $\leq$1 m$^2$ and extents of $\leq$10,000 ha (Fig 1A;B).
+In the temporal dimensions, most observations are either un-replicated or relatively infrequent ($>$1 month interval; Fig.
+1C) and have relatively short duration ($\leq$1 year; Fig.
 1D).
@@
 Contrasting observational dimensions reveals that larger extents are associated with larger spatial replicates (Fig.
-2C), while longer durations tend to be associated with longer intervals (Fig.
+2C), while longer durations are associated with longer intervals (Fig.
 2B).
-The latter association primarily reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of traditional field-based observation, the dominant observational mode, but is also reflected by the relative domain locations of the four main observational methods, suggesting that high observational frequency poses a durational cost across observational methods.
-A similar cost tradeoff is also evident in the inverse relationship between resolution and interval that dominates that domain space (Fig.
-2A), which again primarily relates to characteristics of field-based observations, in which larger spatial replicates demand greater effort that in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
-Less obvious is the opposite tradeoff that occurs with remote observation (Supplementary Figure 2), where finer resolution is desired to increase detail, but typically comes at the cost of longer intervals \cite{estes_platform_2016}.
+The latter association reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of field observations, but also appears to affect the other three methods, as evidenced by their relative domain locations.
+A similar tradeoff can be seen in the inverse relationship between resolution and interval (Fig.
+2A), which primarily relates to field observations, where larger spatial replicates demand greater effort, which in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
+Less obvious is the opposite tradeoff that affects remote observation (Supplementary Figure 2), where finer resolution (which is necessary for detail) typically necessitates longer intervals \cite{estes_platform_2016}.
 
-As a result of these tradeoffs, there are several notable gaps in the spatio-temporal domains of ecological observations, which lie in the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($\geq$1 m$^2$ up to 100 ha; Fig.
+As a result of these tradeoffs, there are several notable observational gaps, specifically within the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($>$1 m$^2$ up to 100 ha; Fig.
 2A) and 2) decadal or longer durations (Fig.
@@
 
-Have these domains changed since the seminal papers on scale first began to appear in the late 1980s \cite{wiens_spatial_1989, levin_problem_1992, tilman_ecological_1989}?
-A comprehensive answer to this question would require a similar study focused on earlier literature, but an analysis of results presented in three prior studies provides partial insight.
-The first and most comprehensive dataset consists of duration values extracted by Tilman \cite{tilman_ecological_1989} from 623 studies published between 1977-1987 in the journal \emph{Ecology}.
-The mean duration of the most comparable subset of those values (n=419; see SI) was 3.6 years, versus 3.3 years in our sample (or 5.1 if temporally un-replicated observations are excluded).
-The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who extracted resolutions from 97 community ecology experiments published in \emph{Ecology} between 1980-1986.
-The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,479,465 m$^2$), but comparing the 80th percentile value (197 m$^2$) of Kareiva and Anderson's \cite{kareiva_spatial_1988} to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
-The third dataset is provided by Porter et al \cite{porter_wireless_2005}, who compared the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}).
+Have these domains changed since the seminal papers on scale first appeared in the late 1980s \cite{wiens_spatial_1989, levin_problem_1992, tilman_ecological_1989}?
+A comprehensive answer would require a repeat similar analysis focused on earlier literature, but the data provided by three prior studies provides partial insight.
+The first dataset consists of duration values that Tilman \cite{tilman_ecological_1989} extracted from 623 studies published between 1977-1987 in the journal \emph{Ecology}.
+The mean duration of the most comparable subset of those values (n=419) was 3.6 years, versus 3.3 years in our sample (or 5.1 when excluding temporally un-replicated observations).
+The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who assessed the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986.
+The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,479,465 m$^2$), but comparing Kareiva and Anderson's \cite{kareiva_spatial_1988} 80th percentile value (197 m$^2$) to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
+Porter et al \cite{porter_wireless_2005} provide the third dataset, which compares the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}).
 The mean interval was 178 days, compared to 684 days in our sample, but the 80th percentile value in our study was 169 days compared to 329 days in theirs.
@@
 Although limited due to methodological differences (e.g.
-a focus on experiments versus un-manipulated systems), these comparisons suggest that the duration and, less clearly, resolution of ecological observations have changed little in the past 30 years, but observational frequency and particularly extent have increased.
-A weak positive trend in our own data also suggests that average extent of ecological observations is steadily increasing (Fig.
-S5 in SI), and likely corresponds to the increasing use of remote observation technologies (Fig.
-S4 in SI).
+a focus on experiments versus un-manipulated systems), these comparisons suggest that the duration and resolution of ecological observations have changed little in the past 30 years, but observational frequency and extent have both increased.
+A weak positive trend in our own data also suggests that the mean extent of ecological observations is steadily increasing (Supplementary Figure 5), which likely corresponds to increasing use of remote sensing (Supplementary Figure 4).
 
-However, even though observational extent appears to be increasing, there remains for most observations a large gulf between the area that is actually sampled and that which the spatial replicates ostensibly represent (Fig.
+Despite this apparent increase in observational extent, there remains a large gulf between the areas that ecologists actually observe and the areas their observations are intended to represent (Fig.
 3A).
-A similarly large discrepancy is also evident between the time spans covered by repeated observations and the time that is spent actually observing a phenomenon (Fig.
+A substantial discrepancy also exists between the amount of time spent observing phenomena and the time spans those observations theoretically represent (Fig.
 3B).
 These differences between the actual and ostensible scales of observation have implications for ecological understanding, as the unobserved portions of space and time may contain important patterns and processes that are not captured by replicates, due to phenomenon-dependent factors such as autocorrelation and representativeness of the sampling scheme \cite{underwood_experiments_1997,palmer_scale_1994, cao_comparison_2002, legendre_spatial_1993,collins_method_2000}.
-Brief, infrequent snapshots, or fine-grained, spatially sparse replicates, may be sufficient to characterize many phenomena (as an example with respect to temporal replication, annual changes in tree cover are well-represented by low frequency satellite imaging \cite{hansen_high-resolution_2013}), but may be inadequate for more dynamic phenomena.
+Brief, infrequent snapshots, or fine-grained, spatially sparse replicates, may be sufficient to characterize many phenomena (e.g.
+annual changes in tree cover are well-represented by low frequency satellite imaging \cite{hansen_high-resolution_2013}), but may be inadequate for more dynamic phenomena.
 For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe fire behavior \cite{clements_observing_2007}.
@@
 It is therefore important to examine whether the scales of the phenomena being observed are adequately captured by the design of replicates.
-Our methods suggest one possible procedure for assessing the \emph{scale representativeness} of replicates, which is to i) calculate the autocorrelation (spatial or temporal) within the observations (using a semi-variogram, for example), ii) find the threshold distance (or time) below which a suitably strong correlation (e.g.
-r=0.7) will exist between neighboring sampled values, iii) add that distance (or time) to the sample resolution (or duration), and iv) recalculate actual extent (or duration) using the adjusted resolution (or sampling duration).
+Our methods suggest one possible procedure for assessing the \emph{scale representativeness} of observations, which is to 1) calculate the autocorrelation (spatial or temporal) within the observations (using a semi-variogram, for example), 2) find the threshold distance (or time) below which a suitably strong correlation (e.g.
+r=0.7) will exist between neighboring sampled values, 3) add that distance (or time) to the sample resolution (or duration), and 4) recalculate actual extent (or duration) using the adjusted resolution (or sampling duration).
 The difference between this autocorrelation-adjusted actual extent (or actual duration) and extent (or duration) may provide a useful additional measure of how well the replicates represent the intended scale of observation.
-Though increasing spatial or temporal coverage may not always be the goal of a study (e.g., when spatial or temporal autocorrelation is a measure of interest), if the gap between actual and ostensible values remains large, then alternative sampling methods may be used to close it.
+Though increasing spatial or temporal coverage may not always be the goal of a study, if the gap between actual and ostensible values remains large, then alternative sampling methods may be used to close it.
 For example, remote sensing provides wall-to-wall spatial coverage of a study area, erasing the difference between actual extent and extent.
@@
 However, most of this uncertainty is attributable to unclear reporting of scale values in the majority of papers we reviewed (a problem also noted in geography studies \cite{margulies_ambiguous_2016}).
-This tendency towards vague documentation offers one final insight, which is that, despite decades of accumulated knowledge regarding the importance of scale in ecology \cite{levin_problem_1992, wiens_spatial_1989, chave_problem_2013, wheatley_factors_2009}, scale appears to remain a low priority throughout much of the discipline.
+This tendency towards vague documentation offers one final insight: despite decades of accumulated knowledge regarding its importance \cite{levin_problem_1992, wiens_spatial_1989, chave_problem_2013, wheatley_factors_2009}, scale appears to remain a low priority throughout much of the ecological discipline.
 Beyond contributing to the broader problem of scientific reproducibility \cite{baker_1500_2016}, inattentiveness to scale increases the risk that observations inadequately represent the phenomenon of interest, thereby limiting the generalizability of any derived ecological knowledge \cite{margulies_ambiguous_2016, wheatley_factors_2009, wiens_spatial_1989}.
 To mitigate this problem, we recommend that ecological journals require authors to quantify and clearly report the values of resolution, extent, interval, and duration.
-Fortunately, some journals already appear to be implementing such policies.
-For example, \emph{Global Ecology and Biogeography} now requires information on the spatial, temporal and taxonomic scale of studies to be in the abstract (a policy adopted in early 2016).
+Fortunately, some journals already appear to be implementing such policies; for example, \emph{Global Ecology and Biogeography} now requires information on the spatial, temporal and taxonomic scale of studies to be in the abstract (a policy adopted in early 2016).
 
@@
 However, the increasing extent of ecological observations, enabled by remote sensing and presumably motivated by many ecologists' appreciation of scale-related issues, suggests that ecology's scale domains are gradually changing.
-In the coming years, the accelerating gains in technology and analytical methods will allow researchers new and unprecedented capabilities to peer into, and thus close, the prominent holes in observational scale domains.
+In the coming years, the accelerating gains in technology and analytical methods will allow researchers new and unprecedented capabilities to peer into, and thus close, the prominent holes in observational domains.
 A renewed, discipline-wide focus on scale's importance, including the adoption of stricter scale-reporting standards by journals, will help spur ecologists to address these gaps, while fostering the improved transferability of knowledge within the discipline.
@@
 Bandwidths of varying resolutions were tested on kernel density estimates of sampling interval versus plot resolution to test how sensitive our results were to the bandwidth value (see Supplementary Results section).
+For comparisons involving the interval dimension, we removed temporally un-replicated observations because these lacked interval values.
 
@@
 Decades were calculated for each pair for all bootstrap replicates.
+We plotted the box plots against their corresponding bin means to evaluate how these differences varied with scale (Fig.
+3).
 

### 2018-02-03 paper/main.tex
Final edits for re-submission

@@
 Ecological patterns emerge from temporal and spatial domains that may be coarser or finer than the processes that shape them, which means that investigation across multiple scales is essential for understanding ecological phenomena \cite{levin_problem_1992,sandel_scale_2009}.
-This awareness has grown rapidly since the 1980s, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad and increasing range of scales \cite{schneider_rise_2001, chave_problem_2013}.
+This awareness has grown rapidly since the 1980s\cite{schneider_rise_2001}, accelerated by the need to understand how changes in the global climate, ocean, and land systems are affecting everything from individual populations \cite{tingley_push_2012} to entire biomes \cite{xiao_photosynthetic_2004}, while technological advances in areas such as remote sensing and genetics are making it ever-easier to quantify ecological features across a broad and increasing range of scales \cite{schneider_rise_2001, chave_problem_2013}.
 
-Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) meta-analyses \cite{tilman_ecological_1989,kareiva_spatial_1988} of ecological research scales were published nearly 30 years ago (but see \cite{porter_crop_2005,sandel_scale_2009} for more recent reviews), it is both timely and important to assess the scales of contemporary ecological investigation.
+Given the growing awareness of scale, expanding data gathering capabilities, and the fact that the most comprehensive (and arguably best-known) meta-analyses \cite{tilman_ecological_1989,kareiva_spatial_1988} of ecological research scales were published nearly 30 years ago (but see \cite{porter_wireless_2005, sandel_scale_2009} for more recent reviews), it is both timely and important to assess the scales of contemporary ecological investigation.
 To address this need, we quantified the spatial and temporal domains of empirical observations that were reported within recently (2004-2014) published ecological studies.
@@
 non-experimentally manipulated) ecological features reported within 133 of the reviewed papers (plus an additional 62 cited as the source of observations).
-The majority of sampled observations were collected using conventional field methods (80\%), followed by automated \emph{in situ} sensing techniques (12.4\%), remote sensing (6.9\%), and paleo-reconstruction ($<$0.8\%).
+Most sampled observations were collected using conventional field methods (80\%), followed by automated \emph{in situ} sensing techniques (12.4\%), remote sensing (6.9\%), and paleo-reconstruction ($<$0.8\%).
 
@@
 \noindent \textbf{\emph{Observational domains within individual dimensions}}\\
-In terms of resolution, spatial replicates for the majority (67\%) of observations (across all methods) were $\leq$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $>$1 ha (Fig.
+In terms of resolution, spatial replicates for most (67\%) observations (across all methods) were $\leq$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $>$1 ha (Fig.
 1A).
 These distributions primarily reflect those of field observations, the dominant observational methodology.
-Examining the distributions for each observational method (Supplementary Figure 1) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $\leq$0.1 m$^2$) than field observations (47\% $\leq$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $>$100 m$^2$).
+Automated sensing and paleo-reconstruction observations had resolutions that were generally finer (85\% or more $\leq$0.1 m$^2$) than field observations (47\% $\leq$0.1 m$^2$), while most remote observations were much coarser (70\% $>$100 m$^2$; Supplementary Figure 1).
 
@@
 1B).
-As with resolution, the extent covered by automated sensing methods tended to be smaller (52\% $\leq$100 ha) than those of field observations (31\% $<$100 ha), while all but 4\% of remote observations covered areas $>$10,000 ha (as did the small number of paleo-reconstructions).
-
-%16 + 29 + 26 + 67 + 14 + 37 + 10
+As with resolution, the extent covered by automated sensing methods tended to be smaller (52\% $\leq$100 ha) than those of field observations (31\% $<$100 ha), while 96\% of remote and all paleo-reconstruction observations covered areas $>$10,000 ha.
 
@@
 
-Duration was $\leq$1 day for 31\% of sampled observations (due to lack of temporal replication), while 10\% covered one day to one month, 23\% one month to one year, 27\% 1-10 years, and 9\% a decade or more (several paleoecological studies covered centuries; Fig.
+Duration was $\leq$1 day for 31\% of sampled observations (due to lack of temporal replication), while 10\% covered one day to one month, 23\% one month to one year, 27\% 1-10 years, and 9\% $>$1 decade (Fig.
 1D).
@@
 \noindent \textbf{\emph{Observational domains within two-dimensions}}\\
-Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack intervals) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig.
+Contrasting resolution with interval reveals that most temporally replicated observations had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig.
 2A).
@@
 The four observational methods occupied substantially different domains, as indicated by the locations of their median values (and see Supplementary Figure 2): the median domain of field observations had 0.1-1 m$^2$ resolution and a monthly interval, whereas remote observations had a coarser median resolution (1,000 m$^2$) but finer median interval ($\sim$1 day).
-Paleo-reconstructions and automated sensing techniques were both finely resolved (median between 10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.
+Paleo-reconstructions and automated sensing were both finely resolved (median between 10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.
 
-Comparing the interval and duration of temporally replicated observations showed most observations had daily to decadal intervals and durations of 1 month to 1 decade (Fig.
+Comparing the interval and duration of temporally replicated observations showed most observations had daily to decadal intervals and durations of one month to one decade (Fig.
 2B).
-The orientation of this concentration shows that interval increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting 1 year to 1 decade tend to have yearly to decadal intervals.
-This tendency is reflected in the domain medians of the primary observational methods: automated sensing had the finest median interval (hour-day) and shortest duration (month-year), followed by remote sensing (median interval slightly greater than one day and median duration of 1 year), field observations (median monthly interval and duration just over 1 year), and finally paleo-reconstructions (median interval 1 decade and millennial duration).
+Interval appears to increases with duration; observations lasting one month to one year tend to have daily to monthly intervals, while those lasting one year to one decade tend to have yearly to decadal intervals.
+This tendency is reflected in the domain medians of the primary observational methods: automated sensing had the finest median interval (hour-day) and shortest duration (month-year), followed by remote sensing (median interval $\sim$1 day and duration 1 year), field observations (median interval 1 month and duration $\sim$1 year), and finally paleo-reconstructions (median interval 1 decade and millennial duration).
 
@@
 2C).
-The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $\leq$10 ha extent.
+Another prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third, fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $\leq$10 ha extent.
 These three concentrations suggest that observational extent increases with resolution, which is further evident in the median domain values (and kernel densities; Supplementary Figure 2) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.
-Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a possible artifact of small sample size.
+Paleo-reconstructions were the exception, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a possible artifact of small sample size.
 
@@
 Three other notable, but lesser concentrations are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($\leq$1 day) of either 1-100 ha or 10,000-1,000,000 ha.
-The median observation from automated sensing (1 year duration, 100 ha extent) lies near the center of the first major concentration, while the median extents of field (1,000-10,000 ha) and remote (1-10 million ha) observations bound the second major concentration at its upper and lower extents, with the median duration of both observational types falling between 1 month to 1 year.
+The median observation from automated sensing (1 year duration, 100 ha extent) lies near the center of the first major concentration, while the median extents of field (1,000-10,000 ha) and remote (1-10 million ha) observations bound the second major concentration at its upper and lower extents, with the median duration of both observational types falling between one month to one year.
 
@@
 3A).
-This difference increased with extent, reaching a maximum of 8.3 between 100 million and 1 billion ha of extent, then falling to 3 orders of magnitude between 1-10 billion ha (these extents were covered by $<$2\% of observations, which were primarily collected with remote sensing).
-Remote observations had the smallest mean difference magnitude (1.9), compared to 5.7 or larger for the other three methods (Supplementary Figure 3).
+This difference increased with extent, reaching a maximum of 8.3 between 100 million and 1 billion ha of extent, then falling to 3 orders of magnitude between 1-10 billion ha (these extents comrpised $<$2\% of observations, which were primarily collected with remote sensing).
+Remote observations had the smallest mean difference magnitude (1.9), compared to $\geq$5.7 for the other three methods (Supplementary Figure 3).
 
-The magnitudes of difference between observational duration and actual duration was somewhat smaller, averaging 3.4 and ranging from just under 2 for the shortest durations (hour-day) to over 4 for observations lasting 1 decade to 1 century (Fig.
+The difference magnitudes between observational duration and actual duration were somewhat smaller, averaging 3.4 and ranging from $\sim$2 for the shortest durations (hour-day) to >4 for observations lasting one decade to one century (Fig.
 3B).
 As with extent, the difference fell substantially for the longest durations (century to millenia), as these domains were covered by paleo-reconstructions (Supplementary Figure 3), which have little difference between actual and ostensible duration because coring techniques capture continuous temporal records.
-The mean difference magnitudes for the other three observing methods ranged from just over 3 (field and automated sensing) to nearly 6 (remote observations).
+The mean difference magnitudes for the other three observational methods ranged from just over 3 (field and automated sensing) to nearly 6 (remote observations).
 
@@
 Our results were potentially influenced by several methodological issues.
-First, most studies did not precisely report observational scales, thus we had to estimate, rather than simply record, scale values for the majority of observations (we estimated 63, 60, 69, 36, 64, and 83\% of resolution, extent, actual extent, interval, duration, and actual duration values, respectively).
+First, most studies did not precisely report observational scales, thus we had to estimate, rather than simply record, scale values for most observations (we estimated 63, 60, 69, 36, 64, and 83\% of resolution, extent, actual extent, interval, duration, and actual duration values, respectively).
 Estimation errors may therefore have biased our findings.
@@
 The resulting confidence intervals (Fig.
-1) suggest that that estimation errors did not unduly influence our findings.
+1) suggest that estimation errors did not unduly influence our findings.
 
-Our scale-estimation protocols may also have introduced bias, particularly our rule for estimating resolution (the smallest areal unit of \emph{complete} measurement).
+Our scale-estimation protocols may also have introduced bias, particularly our protocol for estimating resolution (the smallest areal unit of \emph{complete} measurement).
 We selected this definition for the sake of consistency, but some papers reported resolution as a larger area in which sub-samples were taken.
@@
 For example, average resolution and duration would likely be finer \cite{tilman_ecological_1989,kareiva_spatial_1988}.
-Additionally, the token 1 second (Supplemental Methods) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) caused us to underestimate the differences between their durations and actual durations (Supplementary Figure 3).
+Additionally, the token one second (Supplemental Methods) we used to represent the duration of remotely sensed temporal replicates (which are effectively instantaneous) caused us to underestimate the differences between their durations and actual durations (Supplementary Figure 3).
 However, the relatively small number of remote observations suggests that the impact of this bias on our overall findings was negligible.
@@
 Although our sample size was too small to assign statistical significance, we found a possible positive trend in the use of remote observations and a corresponding decline in field observations over the course of our study period.
-If these trends were not spurious, they suggest that including studies from 2015-2017 would result in a somewhat larger relative sample of remote observations, which would in turn slightly increase the mean extent of ecological observations (see Supplementary Results for further details).
+If these trends were not spurious, they suggest that including studies from 2015-2017 would result in a somewhat larger sample of remote observations, which could slightly increase the mean observational extent (see Supplementary Results).
 
@@
 Our results suggest that modern ecology's observational domains are fairly narrow, and that ecologists still primarily rely on conventional field-based observational techniques.
-In the spatial dimensions, the majority of observations have grains of $\leq$1 m$^2$ and extents of $\leq$10,000 ha (Fig 1A;B).
+In the spatial dimensions, most observations have resolutions $\leq$1 m$^2$ and extents $\leq$10,000 ha (Fig 1A;B).
 In the temporal dimensions, most observations are either un-replicated or relatively infrequent ($>$1 month interval; Fig.
@@
 The latter association reflects a cost-imposed tradeoff between sampling frequency and temporal duration that is characteristic of field observations, but also appears to affect the other three methods, as evidenced by their relative domain locations.
-A similar tradeoff can be seen in the inverse relationship between resolution and interval (Fig.
-2A), which primarily relates to field observations, where larger spatial replicates demand greater effort, which in turn reduces sampling frequency \cite{kareiva_spatial_1988}.
-Less obvious is the opposite tradeoff that affects remote observation (Supplementary Figure 2), where finer resolution (which is necessary for detail) typically necessitates longer intervals \cite{estes_platform_2016}.
+A similar tradeoff is illustrated by the inverse relationship between resolution and interval (Fig.
+2A), which primarily relates to field observations, where larger spatial replicates demand greater effort, reducing sampling frequency \cite{kareiva_spatial_1988}.
+Less obvious is the opposite tradeoff that affects remote observation (Supplementary Figure 2), where finer resolution (necessary for detail) typically necessitates longer intervals \cite{estes_platform_2016}.
 
-As a result of these tradeoffs, there are several notable observational gaps, specifically within the domains defined by high frequency (daily to sub-daily intervals) observations having 1) high to moderate resolutions ($>$1 m$^2$ up to 100 ha; Fig.
-2A) and 2) decadal or longer durations (Fig.
+As a result of these tradeoffs, there are several notable observational gaps, specifically within the domains defined by high frequency (daily to sub-daily intervals) observations having high to moderate resolutions ($>$1 m$^2$ up to 100 ha; Fig.
+2A) and decadal or longer durations (Fig.
 2B).
-Another hole is evident in the high to moderate resolution, large extent (1 million-10 billion ha) domain (Fig.
+Another gap is evident in the high to moderate resolution, large extent (1 million-10 billion ha) domain (Fig.
 2C).
@@
 Have these domains changed since the seminal papers on scale first appeared in the late 1980s \cite{wiens_spatial_1989, levin_problem_1992, tilman_ecological_1989}?
-A comprehensive answer would require a repeat similar analysis focused on earlier literature, but the data provided by three prior studies provides partial insight.
-The first dataset consists of duration values that Tilman \cite{tilman_ecological_1989} extracted from 623 studies published between 1977-1987 in the journal \emph{Ecology}.
+A comprehensive answer would require a similar analysis focused on earlier literature, but the data provided by three prior studies provide partial insight.
+The first dataset consists of duration values that Tilman \cite{tilman_ecological_1989} extracted from 623 studies published between 1977-1987 in \emph{Ecology}.
 The mean duration of the most comparable subset of those values (n=419) was 3.6 years, versus 3.3 years in our sample (or 5.1 when excluding temporally un-replicated observations).
 The second dataset is found in Kareiva and Anderson \cite{kareiva_spatial_1988}, who assessed the resolutions of 97 community ecology experiments published in \emph{Ecology} between 1980-1986.
-The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,479,465 m$^2$), but comparing Kareiva and Anderson's \cite{kareiva_spatial_1988} 80th percentile value (197 m$^2$) to ours (115 m$^2$) shows that the majority of contemporary observations are finer-grained than most 1980s-era experiments.
-Porter et al \cite{porter_wireless_2005} provide the third dataset, which compares the extent and interval of 25 studies published in 2003 and 2004 (also in \emph{Ecology}).
+The average of those (12,657 m$^2$) was substantially smaller than the mean of our sample (1,479,465 m$^2$), but comparing Kareiva and Anderson's \cite{kareiva_spatial_1988} 80th percentile value (197 m$^2$) to ours (115 m$^2$) shows that most contemporary observations are finer-grained than most 1980s-era experiments.
+Porter et al \cite{porter_wireless_2005} provide the third dataset, which compares the extent and interval of 25 studies published in 2003-2004 in \emph{Ecology}.
 The mean interval was 178 days, compared to 684 days in our sample, but the 80th percentile value in our study was 169 days compared to 329 days in theirs.
@@
 a focus on experiments versus un-manipulated systems), these comparisons suggest that the duration and resolution of ecological observations have changed little in the past 30 years, but observational frequency and extent have both increased.
-A weak positive trend in our own data also suggests that the mean extent of ecological observations is steadily increasing (Supplementary Figure 5), which likely corresponds to increasing use of remote sensing (Supplementary Figure 4).
+A weak positive trend in our data also suggests that the mean extent of ecological observations is steadily increasing (Supplementary Figure 5), which likely corresponds to increasing use of remote sensing (Supplementary Figure 4).
 
@@
 annual changes in tree cover are well-represented by low frequency satellite imaging \cite{hansen_high-resolution_2013}), but may be inadequate for more dynamic phenomena.
-For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that it cannot be used to observe fire behavior \cite{clements_observing_2007}.
-To capture such behavior, long periods of continuous observation may be more important for understanding dynamics than frequent repeats.
+For example, wildfire extent and duration can be mapped by daily return satellites \cite{roy_prototyping_2005,jones_fire_2009}, but the instantaneous nature of the imaging means that they cannot be used to observe fire behavior \cite{clements_observing_2007}.
+To capture such behavior, long periods of continuous observation may be more important  than frequent repeats for understanding dynamics.
 
@@
 Furthermore, the interval of high-resolution imaging (higher resolution is preferred in images as it allows individual features to be better discerned \cite{dark_modifiable_2007, hay_comparison_2003}) is now approaching daily to sub-daily scales \cite{drusch_sentinel-2:_2012, hand_startup_2015}, allowing improved representation of spatial and temporal dynamics.
-For phenomena that can't be measured from space, either because they are not visible or because they require continuous observation, new approaches for collecting \emph{in situ} or near-surface observations (e.g.
+For phenomena that cannot be measured from space, either because they are not visible or because they require continuous observation, new approaches for collecting \emph{in situ} or near-surface observations (e.g.
 low-cost wireless sensors \cite{wolf_gsm-based_2012, collins_new_2006, porter_wireless_2005}, citizen observers \cite{dickinson_current_2012}, and autonomous vehicles \cite{anderson_lightweight_2013}) can be used to increase the spatial and temporal coverage of observations.
@@
 To mitigate this problem, we recommend that ecological journals require authors to quantify and clearly report the values of resolution, extent, interval, and duration.
-Fortunately, some journals already appear to be implementing such policies; for example, \emph{Global Ecology and Biogeography} now requires information on the spatial, temporal and taxonomic scale of studies to be in the abstract (a policy adopted in early 2016).
+Fortunately, some journals already appear to be implementing such policies.
+For example, \emph{Global Ecology and Biogeography} now requires information on the spatial, temporal and taxonomic scale of studies to be in the abstract (a policy adopted in early 2016).
 
@@
 \noindent \textbf{\emph{Paper selection and review}}\\
-We used the 2012 Web of Science impact factors to select the highest ranked ecology-themed journals that published studies with an observational component, excluding journals devoted to reviews, meta-analyses, laboratory, cellular, or experimental studies.
-To select a representative sample of recent ecology studies, we downloaded the metadata for all papers published in the 30 selected journals (Supplementary Table 1) between 2004 and 2014.
+We used the 2012 Web of Science impact factors to select the thirty highest ranked ecology-themed journals that published studies with an observational component, excluding journals devoted to reviews, meta-analyses, laboratory, cellular, or experimental studies.
+To select a representative sample of recent ecology studies, we downloaded the metadata for all papers published in the selected journals (Supplementary Table 1) between 2004 and 2014.
 Our study involved six different observers (those reviewing the papers to extract the observational scales), each of whom was given a randomly selected batch of 500 titles.
@@
 The perturbation occasionally resulted in physically impossible values (e.g.
-interval or actual duration longer than duration; actual extent larger than extent); in these cases we adjusted the perturbed values within the dimension that should be smallest to equal the value in the dimension that should be largest (e.g.
-we would set interval to equal duration, and not vice versa).
+interval or actual duration longer than duration; actual extent larger than extent).
+In these cases we capped the perturbed value in the smaller of the two dimensions (i.e.
+resolution or interval) so that it equalled the corresponding value in the largest (i.e.
+extent or duration).
 We used the resulting set of perturbed observations to quantify uncertainty within our scale estimates.
@@
 
-\noindent Where x is either extent or duration and y is its actual counterpart.
+\noindent Where x is either extent or duration and y is actual extent or actual duration, respectively.
 We then evaluated how the magnitudes of difference varied with increasing values of extent/duration, using box plots to summarize decades within the same bins used to summarize the frequency distributions of the extent and duration of observations (Fig.
@@
 1 in Kareiva and Anderson \cite{kareiva_spatial_1988}, and Fig.
-2 in Porter et al \cite{porter_crop_2005}.
-To maintain as much comparability as possible with our inclusion criteria, we excluded experiment studies in Tilman's data, as well as the values of any studies having duration values greater than 100 years (no upper time bound was provided for these), leaving duration values for 419 (out of 623) studies.
+2 in Porter et al \cite{porter_wireless_2005}.
+To maintain as much comparability as possible with our inclusion criteria, we excluded experiment studies in Tilman's data, as well as the values of any studies exceeding 100 years' duration (no upper time bound was provided for these), leaving duration values for 419 (out of 623) studies.
 Since Tilman presented duration values as a histogram, we calculated the mean duration across all studies as the weighted (by number of observations per bin) mean of bin center-point values (i.e.
@@
 Porter et al also used irregular scales for both their X (frequency) and Y (extent) axis, therefore we had to visually estimate the scale values for each data point after graphical extraction, and converted their extent values (in km) to hectares and their frequency values to intervals.
-Kareiva and Anderson presented resolution as plot diameters (m), which we squared to make comparable to our resolution metric.
+Kareiva and Anderson \cite{kareiva_spatial_1988} presented resolution as plot diameters (m), which we squared to make comparable to our resolution metric.
 
@@
 conceived the study.
-L.E., T.E., P.E., J.C., and E.E.
+L.E., T.T., P.E., J.J.C., and E.E.
 helped design the sampling protocol.
-L.E., T.E., P.E., J.J.C., L.A, and J.C.
+L.E., T.T., P.E., J.J.C., L.A, and J.C.
 extracted and analyzed scale data.
@@
 \vspace{-0.15 cm}
-\caption{Kernel density estimates of observational densities within the domains defined by A) interval and resolution (of temporally replicated observations only), B) duration and extent, C) resolution and extent, and D) interval and duration (of temporally replicated observations).
-Density estimates were applied to the log-transformed values of each observational dimension, and density estimates are rescaled to represent percentages.
+\caption{Kernel density estimates of observational densities within the domains defined by A) resolution and interval (of temporally replicated observations), B) duration and interval (of temporally replicated observations), C) resolution and extent, and D) duration and extent.
+Density estimates were applied to the log-transformed values of each observational dimension, and density estimates were rescaled to represent percentages.
 Letters in the plots denote the median values of different observational methods (f=field observations; a = automated sensing; r = remote sensing; p = paleo-observations).
@@
 \caption{The difference between extent and \emph{actual} extent (the summed area of spatial replicates) (A) and duration and \emph{actual} duration (the summed sampling duration across temporal replicates) (B).
-Difference values are expressed in terms of how many orders of magnitude larger (or longer) extent (duration) is than actual extent (actual duration), and are summarized (as box plots, with circle in box representing the mean and line the median) in bins representing increasing scales of actual extent/duration.
+Difference values are expressed in terms of how many orders of magnitude larger (or longer) extent (duration) is than actual extent (actual duration), and are summarized (as box plots, with circle in box representing the mean and line the median) in bins representing increasing scales of extent/duration.
 The percentages of observations falling within each bin are indicated by the color of the inter-quartile and the numeric value above the upper whisker.

### 2018-03-31 README.Rmd
Added README

---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This package contains the data and code for the following paper:

Estes, L., Elsen, P., Treuer, T., Ahmed, L., Caylor, K., Chang, J., Choi, J.J., Ellis, E. (2018). The spatial and temporal domains of modern ecology. __Nature Ecology and Evolution__, DOI:10.1038/s41559-018-0524-4

The package can be installed and contents explored as follows:

```{r, eval = FALSE}
devtools::install_github("agroimpacts/ecoscales", build_vignettes = TRUE)
browseVignettes("ecoscales")
```
